\paragraph{Grappling: Shoot Takedown \large} \label{man:grappling-shoot-takedown}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Attack
\item [Requirements:] Be at least Medium Range or longer
\item [Maneuver:] Grapple Attack at HA Reach at Takedown TN 8 with X dice,
  targeting Opponent’s leg.

\item [Success:] Immediately enter a Clinch Trapping the Opponent’s leg.
  Opponent must roll a Stability check at [BS] RS or be rendered \nameref{par:Prone}. You also
  immediately become \nameref{par:Prone} if the target fails their Stability check.

\item [Failure:] If Parried or Blocked, you either suffer a Swing, Thrust,
  Shield Bash, or Pommel Strike (depending on the Parrying/Blocking device, at
  the GM's discretion) to your Upper Head with 0 BS, or you are rendered \nameref{par:Prone},
  the defender chooses which.

\item [Superior:] On a successful Shoot Takedown, you may Trap both of the
  Opponent’s legs. 
\end{description}

\paragraph{ Grappling: Trap \large } \label{man:grappling-trap}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] None.
\item [Maneuver:] Grapple Attack at HA Reach at Trap TN 7 with X dice, targeting
  head, right arm, left arm, right leg, left leg, or torso.

\item [Success:] Target body part becomes Trapped. All TNs increase by 1 for the
  victim until it is released. No actions can be taken involving the Trapped
  body part (except for Force to try and free the body part) until it is
  released. If a weapon is being held in a Trapped body part, the weapon itself
  is also considered to be Trapped. If the body part that is Trapped was being
  used in a Maneuver, then Trap also inflicts [BS] Stun.
\begin{itemize}
\item You may voluntarily release a Trap or Pin at any time.
\end{itemize}
\item [Special:] If you fail any Stability test, you release all Trapped body
  parts before falling \nameref{par:Prone}. You may only have two body parts Trapped or Pinned
  at any time. Only one Trap or Pin can increase an Opponent’s TN’s at any time

\item [ Superior: ] If you have Initiative, you resolve Trap at -1 TN.
\end{description}

\paragraph{Grappling: Pin \large } \label{man:grappling-pin}
---\quad {\large [X+2]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] At least one of Opponent's body parts is Trapped.
\item [Maneuver:] Grapple Attack at HA Reach on Trapped body part at Pin TN 7 with X dice.
\item [Success:] Target body part becomes Pinned. This is identical in effect to
  being Trapped, except that TNs increase by 2, instead of 1. Some Maneuvers
  have a special bonus if used against a Pinned body part.  
\item [Special:] If you fail any Stability test, you release all Pinned body
  parts before falling \nameref{par:Prone}. You may only have two body parts
  Trapped or Pinned at any time. Only one Trap or Pin can increase an opponent’s
  TN’s at any time.  
\end{description}

\paragraph{Grappling: Subdue \large} \label{man:grappling-subdue}
---\quad {\large [X+$\rfrac{1}{2}$ Opponent’s STR]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative. Be Trapping or Pinning a body part
\item [Maneuver:] Roll Subdue TN 8 with X dice.
\item [Success:] Opponent immediately suffers [1+Half BS] Stun and must pay an
additional [1] activation cost to all maneuvers. You may choose to maintain this
Maneuver each action instead of performing another maneuver, rolling at TN 7
instead of 8. While maintained, the Stun is increased to [1+BS] Stun. You do not
need to pay the additional activation cost to maintain this maneuver. 

\item [Special:] If the Grapple is a Ground Fight and you have initiative,
  resolve this maneuver at -1 TN. You may choose to pay [$\rfrac{1}{2}$ the
  Opponent’s STR] CP each action you maintain this maneuver to inflict Pain
  instead of Stun while maintaining this Maneuver. 
  \begin{itemize}
  \item If you have Trapped or Pinned more than one body part, increase the
    Stun/Pain inflicted by 1 and the opponent’s activation cost penalty by 1.  
  \item If you have Pinned one of the Opponent’s body parts, reduce all STR
    -based activation costs by 1 for this maneuver. This effect is doubled if you
  are Pinning two body parts.  
  \item You may pay 2 CP to change this Subdue into a Submission Hold after dice
    have been declared. You immediately inflict 1 strain on the Subdued body part
  for each action Subdue has been maintained.  
  \end{itemize}
\item [Superior:] You may perform a
  \hyperref[man:grappling-limb-break]{Limb-Break} on the subdued body part at -1
  TN, Limb-Break ends the subdue.    
\end{description}

\paragraph{Grappling: Submission Hold \large} \label{man:grappling-submission-hold}
---\quad {\large [X+1]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative. Be Trapping or Pinning a body part.
\item [Maneuver:] Roll Submission hold at TN 7 with [X] dice targeting one
  Trapped or Pinned body part. 

\item [Success:] Target Trapped body parts Strain increases by your BS. If
Strain is higher than target STR, inflict [Strain-target STR] unarmed damage
to one of the following hit locations, selecting from the appropriate body
part’s list. This damage is not resisted by TOU or armor. 
  \begin{description}
  \item [Head: ] Neck, Upper Back, Shoulder
  \item [Arm: ] Elbow, Shoulder, Lower Arm
  \item [Torso: ] Lower Back, Upper Back, Belly
  \item [Leg: ] Hip, Knee, Shin, Foot
  \end{description}
\item [Special:] If the Grapple is a Ground Fight and you have initiative,
  resolve this maneuver at -1 TN. You may pay [$\rfrac{1}{2}$ the Opponent’s
  STR] CP each action you maintain this maneuver to inflict [BS] Stun   

If you have Trapped or Pinned more than one body part, increase the Stun
inflicted by 1 

If you have Pinned the target body part, treat the opponent’s STR as 1 lower for
all purposes associated with this maneuver. If you have two body parts pinned,
including the target body part, treat their STR as half it’s normal score. 

\item [ Superior: ]
 You may now inflict Pain rather than Stun using the Special section of
 Submission Hold. This Pain is cumulative, and is removed once the Submission
 Hold is released.  
\end{description}

\paragraph{Grappling: Limb-Break \large} \label{man:grappling-limb-break}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative. At least one of Opponent's body parts is Trapped.
\item [Maneuver:] Grapple Attack at HA Reach at Hit Location on Trapped body
  part at Break TN with X dice. 
\item [Success:] Inflict [STR-1+Half BS] bludgeoning damage (not
  unarmed--bludgeoning!) to Hit Location chosen from list. This attack ignores
  all AV on the chosen Hit Location, but see below. 
  \begin{description}
  \item [Arms TN 6:] Shoulder, Upper Arm, Elbow, Forearm
  \item [Head TN 7:] Neck, Lower Head
  \item [Legs TN 7:] Hip, Thigh, Knee, Shin, Foot
  \item [Torso TN 9:] Shoulder, Chest, Sides, Belly, Hip, Upper Back, Lower Back
  \end{description}
\item [Special:] 
 If the Hit Location is covered by non-weakpoint($\phi$) Plate, then apply half
 of that armor’s AV.  

If the body part being targeted is Pinned or Subdued, Limb-Break inflicts
[STR-1+BS] instead of its original amount. Additionally, if the Opponent is
Subdued, resolve this maneuver at -1 TN. 
\item [ Superior: ] Limb-Break damage is increased by 2.
\end{description}

\paragraph{Grappling: Throttle \large} \label{man:grappling-throttle}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] Opponent's head is Trapped or Pinned.
\item [Maneuver:] Grapple Attack at HA Reach at Throttle TN 7 with X dice.
\item [Success:] Opponent suffers [BS] Stun, and begins to suffocate.

Each Action after the first, Throttle can be ‘maintained’ in place of performing
another Maneuver. Maintaining a Throttle is the same as performing the Throttle
Maneuver, but at TN 6. The Opponent takes [1+$\rfrac{1}{2}$ BS] Pain every Action that
Throttle is maintained (not the Action it is initiated) along with [BS] Stun.
When a character's Pain from Throttle equals or exceeds their MOB score, they
must make a WIL test at the end of each Action (at 1 RS per point of Pain over
the character's MOB score) or lose consciousness for 1d10 minutes. 

If a character is Throttled into unconsciousness, they must make a HLT test at
RS 2 or die from a crushed windpipe. Each Throttle action after unconsciousness
forces an additional test at +1 RS.   

If a character breaks the Throttle and manages to get free, they lose 1 Pain for
every 2 Rounds they have to breathe without exertion (but only to a maximum of
the Pain inflicted by Throttle). 
\item [Special:] If a garrote, like a wire or cord, is used instead of bare
  hands or arms, the TN for Throttle is 6, maintaining it is done at TN 5, and
  the Pain inflicted every Action that it is maintained is [1+BS].

  If a character is garroted into unconsciousness, they must make a HLT test at
  RS 4 or die after going unconscious, from a crushed glottis/windpipe. Each
  Throttle action after unconsciousness forces an additional test at +1 RS.

  If the neck is protected by plate armor like a gorget or bevor, the TN of
  Throttle increases to TN 9, and maintaining it increases to TN 8. Garrotes
  cannot be used over characters with such protection.
\end{description} 

\paragraph{Grappling: Coup De Grace \large} \label{man:grappling-coup-de-grace}
---\quad {\large [X+2]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Thrusting, Initiative
\item [Requirements:] Have initiative. Be wielding a weapon with an arm that is
  not Trapped, Pinned or Subdued. This weapon must inflict piercing damage on
  the Thrust. Must have only one of Opponent’s body parts Trapped or Pinned.  
\item [Maneuver:] Thrusting Weapon Attack at Weapon Reach with X dice, rolled at
  Thrust TN+1, aimed at a Hit Location in Trapped or Pinned Body Part’s area. 
\item [Success:] Inflicts [Weapon Thrust Damage+STR+BS] piercing damage to
  chosen Hit Location. Do not roll on table. The damage from this attack is not
  reduced by AV from any armor which has a Weak Spot for the Hit Location that
  was targeted by this attack. All other AV protecting that area applies. 
\item [Special:] If the weapon being used is longer than S, increase Thrust TN
  by 1 per step longer the weapon is. Longer weapons shortened through the use
  of Half-Swording or the Grappling at the Sword Talent are exempt from these
  penalties if this would bring them to S Reach or lower. If the weapon being
  used has Thin Blade, make this Attack at -1 TN.  
\item [ Superior: ] If the Grapple is currently a Ground Fight, resolve this
  maneuver at an additional -1 TN 
\end{description}

\paragraph{Grappling: Reap Takedown \large} \label{man:grappling-reap-takedown}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative, both you and Opponent are Standing. Be Trapping
an arm, head, or torso, and have at least one of your legs not Trapped or
Pinned. 
\item [Maneuver:] Grapple Attack at HA Reach at Takedown TN 7 with X dice.
\item [Success:] Opponent rendered \nameref{par:Prone} and suffers [BS] Stun. If you are not
  being Trapped, you may remain Standing. 
\item [Failure:] Make a Stability check at [Opponent BS] RS or be rendered \nameref{par:Prone}.  
\end{description}

\paragraph{Grappling: Suplex \large} \label{man:grappling-suplex}
---\quad {\large [X+3]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative, both you and Opponent are Standing. Be
  Trapping 
the torso and one other body part. 
\item [Maneuver:] Grapple Attack at HA Reach at Suplex TN 7 with X dice.
\item [Success:] Both characters are rendered \nameref{par:Prone}. Opponent suffers [BS*2]
  Stun and 
takes falling damage to the Upper Back as if they fell [STR + BS] feet. 
\item [Failure:] Opponent’s torso is no longer trapped.
\item [Special:] If you have Trapped one of the Opponent's legs and successfully
  Suplex them, your opponent uses only half their TOU to reduce the distance
  fallen, and the distance fallen is multiplied by 2. The activation cost of
  this maneuver is reduced by 1 if you have Pinned the opponent’s Torso. 
\item [ Superior: ] You may choose to remain standing after performing a Suplex.
\end{description}

\paragraph{Grappling: Toss \large} \label{man:grappling-toss}
---\quad {\large [X+2]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Initiative
\item [Requirements:] Have initiative, be Trapping a body part, both you and
  target are Standing. 
\item [Maneuver:] Grapple Attack at HA Reach at Arm Toss TN 7 with X dice.
\item [Success:] Target is rendered \nameref{par:Prone}, and takes falling damage to the Upper
  Back as if they fell [(STR+BS)*Y] feet.  
 Y is 1 by default.
\item [Special -- Slam:]
 When declaring this maneuver, you may spend 2 CP to increase Y by 1. Y can only
 be increased this way up to a maximum of $\rfrac{1}{2}$ STR.
 \item [Superior:] You may render yourself \nameref{par:Prone} to increase the distance fallen
   for your opponent’s Falling Damage by your TOU + half your Weight not
   counting what's in your hands (Sheathed Polearms only count half their
   weight).    
\end{description}

\paragraph{Grappling: Slip \large} \label{man:grappling-slip}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] Grappling.
\item [Maneuver:] Grapple Defense at Slip TN 6 with X dice.
\item [Success:] Opponent Maneuver negated, you gain Initiative.
\item [Special:] If the Grapple is currently a Ground Fight when performing a
  successful Slip, you may spend 2 CP after success to attempt to stand. Roll a
  MOB test at terrain difficulty. The Opponent can oppose this test with their
  own MOB by spending 2 CP. Success on this test moves the Clinch to standing
  position and ends Ground Fight. If the opponent paid the 2 CP to oppose, they
  stand up as well. If you succeed with 2 or more BS, you may pay 2 CP to break
  free of a Trap or Pin.  
\end{description}

\paragraph{Grappling: Twist \large} \label{man:grappling-twist}
---\quad {\large [X]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] Grappling.
\item [Maneuver:] Grapple Defense at Twist TN 7 with X dice.
\item [Success:] Opponent's Maneuver negated, you gain Initiative, and you gain
  [BS] CP for your next Maneuver if you continue the Grapple. 
\item [Special:] If fight is a Ground Fight when performing a successful Twist,
  you may automatically move yourself to standing position. 

If successful, you may end Clinch, and return both characters to regular combat
at HA Reach. If the fight is a Ground Fight, both characters are \nameref{par:Prone}. 
\end{description}

\paragraph{Grappling: Bite \large} \label{man:grappling-bite}
 ---\quad {\large [X+1]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple, Unarmed
\item [Requirements:] Must have teeth. Cannot have a visor or other armor
  covering the face. No hands required. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Bite TN 8 at Target
  Zone on the Thrusting Attack Table (must aim for face, neck, lower arm, upper
  arm, or any Trapped body part). 
\item [Success:] Inflicts [Half STR+Half BS] cutting damage to Hit Location.
\item [Special:] Wounds inflicted by Bite have +3 Infection Chance.
If Hit Location is protected by Hard armor, you suffer [Half Total Damage]
Shock, inflict Unarmed Damage instead of Cutting, do not inflict the increased
Infection Chance, and cannot inflict higher than a Level 1 Wound. 

If you successfully performed a Bite Maneuver in the previous Action, you may
choose to have it automatically hit the same location, this time rolling at TN
6, to represent the biter latching on and savaging with his teeth. 
\end{description}

\paragraph{Grappling: Force \large} \label{man:grappling-force}
---\quad {\large [X+2]}
\vspace{-10pt} \begin{description}
\item [Type:] Universal, Grapple
\item [Requirements:] Grappling, have one body part currently Trapped or Pinned.
\item [Maneuver:] Select one of your body parts that is currently Trapped or
  Pinned.

  Grappling Attack at Force TN 7 (if Trapped) or Force TN 8 (if Pinned)
  with X dice.  
\item [Success:] Body part is no longer Trapped or Pinned.
\item [Special:] This Maneuver does not increase its TN for Traps or Pins. Use
  the TN listed. In the event of a tie, neither Maneuver resolves. Instead, the
  opponent must roll an [END+1/2 STR] check at RS equal to 1/2 your STR or no
  longer be trapping that body part.  
\end{description}

\paragraph{Grappling: Pick Up \large} \label{man:grappling-pick-up}
---\quad {\large [X+ $\rfrac{1}{2}$ Opponent’s TOU]}
\vspace{-10pt} \begin{description}
\item [Type:]
 Universal, Grapple
\item [Requirements:]
 Be Trapping or Pinning two body parts, your Opponent is \nameref{par:Prone}.
\item [Maneuver:]
 Roll X at TN 7
\item [Success:]
 Both you and your Opponent are now standing.
\item [Failure:]
 If you were standing, you are now \nameref{par:Prone}.
\end{description}
