%BEGIN_FOLD%including all packages, settings etc.
%LETTRINE and \ornamento are to be learned
\RequirePackage[l2tabu,orthodox]{nag}
\documentclass[oneside,11pt,english]{book}
\usepackage{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{xcolor}
\usepackage{titletoc,titlesec} %Used for the mini-tables of content in each chapter, and reducing the header
	\titlespacing*{\chapter}{0pt}{*-5}{*1.5}
	\titleformat{\chapter}[display]{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}
	\dottedcontents{<section>}[<left>]{<above-code>}{<label width>}{<leader width>}
\definecolor{marron}{RGB}{60,30,10}
\definecolor{darkblue}{RGB}{0,0,80}
\definecolor{lightblue}{RGB}{80,80,80}
\definecolor{darkgreen}{RGB}{0,80,0}
\definecolor{darkgray}{RGB}{0,80,0}
\definecolor{darkred}{RGB}{80,0,0}
\definecolor{shadecolor}{rgb}{0.97,0.97,0.97}
\usepackage[demo]{graphicx}
\usepackage{wallpaper}
\usepackage{wrapfig,booktabs}
\usepackage{tabu,longtable,ctable,colortbl}
	\tabulinesep = 1.2mm
\usepackage{fancyhdr}
\usepackage{lettrine}
\usepackage[font=normalsize]{caption}     %% make caption in normal size
	\DeclareCaptionLabelFormat{blank}{} %For hidden captions
\usepackage{multicol}
\usepackage{sidecap}
%\input Acorn.fd
%\newcommand*\initfamily{\usefont{U}{Acorn}{xl}{n}}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{rotating}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{
	tmargin=3cm, 
	bmargin=3cm, 
	lmargin=1.5cm,
	rmargin=1.5cm,
	headheight=1.5cm,
	headsep=0.8cm,
	footskip=0.5cm}
\usepackage[full]{textcomp}
\renewcommand*{\familydefault}{pplj}
%	\usepackage
%	[
%	final,
%	stretch=10,
%	protrusion=true,
%	tracking=true,
%	spacing=on,
%	kerning=on,
%	expansion=true]
%	{microtype}
\setlength{\parskip}{1.3ex plus 0.2ex minus 0.2ex}

\usepackage{fourier-orns}

\newcommand{\ornamento}{\vspace{2em}\noindent \textcolor{darkgray}{\hrulefill~ \raisebox{-2.5pt}[10pt][10pt]{\leafright \decofourleft \decothreeleft  \aldineright \decotwo \floweroneleft \decoone   \floweroneright \decotwo \aldineleft\decothreeright \decofourright \leafleft} ~  \hrulefill \linebreak \vspace{2em}}}
\newcommand{\ornpar}{\noindent \textcolor{darkgray}{ \raisebox{-1.9pt}[10pt][10pt]{\leafright} \hrulefill \raisebox{-1.9pt}[10pt][10pt]{\leafright \decofourleft \decothreeleft  \aldineright \decotwo \floweroneleft \decoone}}}
\newcommand{\ornimpar}{\textcolor{darkgray}{\raisebox{-1.9pt}[10pt][10pt]{\decoone \floweroneright \decotwo \aldineleft \decothreeright \decofourright \leafleft} \hrulefill \raisebox{-1.9pt}[10pt][10pt]{\leafleft}}}

\makeatletter
\def\headrule{{\color{darkgray}\raisebox{-2.1pt}[10pt][10pt]{\leafright} \hrulefill \raisebox{-2.1pt}[10pt][10pt]{~~~\decofourleft \decotwo\decofourright~~~} \hrulefill \raisebox{-2.1pt}[10pt][10pt]{ \leafleft}}}
\makeatother

\makeatletter
	\renewcommand{\l@section}{\@dottedtocline{1}{1.5em}{2.6em}}
	\renewcommand{\l@subsection}{\@dottedtocline{2}{4.0em}{3.6em}}
	\renewcommand{\l@subsubsection}{\@dottedtocline{3}{7.4em}{4.5em}}
\makeatother

\fancyhf{}%ADD a reset button in the footer, maybe linked with the leaf

\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\renewcommand{\sectionmark}[1]{\markright{#1}}

\newcommand{\estcab}[1]{\itshape\textcolor{marron}{\nouppercase #1}}

%\fancyhead[LE]{\estcab{Fran Oldstyle}}
%\fancyhead[RE]{\estcab{History of taxonomy}}
% \fancyhead[CO]{\estcab{\decoone}}
\fancyhead[LO]{\estcab{\rightmark}} % malo cuando no hay section ~~~ \thesection
\fancyhead[RO]{\estcab{\leftmark}}

 \fancyhead[RO]{\bfseries\nouppercase{ \leftmark}}
% \fancyfoot[LE]{\bf \thepage ~~ \leafNE}
%\fancyfoot[RO]{ \leafNE  ~~ \bfseries \thepage}

\fancyfoot[RO]{
	\ornimpar \linebreak \large \hfill \sffamily\bfseries \textcolor{darkgray}{\leafNE ~~~ \thepage}
}
%\fancyfoot[RE]{\ornpar   \newline \large  \sffamily\bf \textcolor{darkgray}{\thepage ~~~ \reflectbox{\leafNE}}  \hfill}

\newenvironment{Section}[1]
{\section{\vspace{0ex}#1}}
{\vspace{12pt}\centering ------- \decofourleft\decofourright ------- \par}

\setlength{\parindent}{1em} % Sangría española
\pagestyle{fancy}

\renewcommand{\footnoterule}{\vspace{-0.5em}\noindent\textcolor{marron}{\decosix \raisebox{2.9pt}{\line(1,0){100}} \lefthand} \vspace{.5em} }
\usepackage[hang,splitrule]{footmisc}
\addtolength{\footskip}{0.5cm}
\setlength{\footnotemargin}{0.3cm}
\setlength{\footnotesep}{0.4cm}

\usepackage{multirow}
\usepackage{chngcntr}
%\counterwithout{figure}{chapter}
%\counterwithout{table}{chapter}
\usepackage[colorlinks=true,linkcolor=darkblue, hyperfootnotes=false]{hyperref}
	\usepackage{etoolbox}
	\AtBeginEnvironment{longtabu}{\footnotesize}{}{}   %% change all longtabu content to foot note size
	\AtBeginEnvironment{tabu}{\footnotesize}{}{} %%Do the same for tabu content, not sure if these two can be combined
	\AtBeginEnvironment{table}{\renewcommand*{\familydefault}{ppl}}{}{}
	\AtBeginEnvironment{longtabu}{\renewcommand*{\familydefault}{ppl}}{}{}
\usepackage{microtype}
%END_FOLD
\begin{document}
\frontmatter
	%\TileWallPaper{300pt}{300pt}{Descargas/fondopapelviejo.jpg}

\begin{titlepage}
\pagestyle{empty}
	\centering
	\ornamento\\
	{\scshape\huge Song of Swords Beta 1.9.9\par}
	\vspace{0.8cm}
	{\scshape\Large{(MICRO UPDATE)}\par \scshape\Large{from} \par}
	\vspace{0.1 cm}
	{\huge Opaque Studios \par}
	\vfill
	{\Large\LaTeX ified Revision - Alpha version 0.1.2\\\today }\par
	\vspace{0.2 cm}
	 \textsc{\copyright~Zachary Irwin, James Lacombe}, 2015. All Rights Reserved
	
	\ornamento
\end{titlepage}
\newpage
\begin{center}
\vfill
\textbf{Beta Rules for Song of Swords Roleplaying Game}\par
Lead Designer: James “Rome” Lacombe\par
Executive Designer, Producer: Zachary T. Irwin\par
Supplementary Design Team: Daniel Leiendecker, Forrest Phanton, Taylor Davis\par
Art Director: Taylor Davis\par
Cover Art: Kenneth Solis\par
Book Art: Darren Tan, Frankie Perez, Duc Pham, Taylor Davis, Kassandra Swager\par


Additional Development: Justin Hazen, Travis Mitchell, Alex Gann, Nicholas Henry, Alan Baird and The Five Cells, Dusan Nadimakovic\par
CEO: Zachary T. Irwin\par


All contents copyright © 2012-2015 by Zachary T. Irwin, James T. Lacombe. All rights reserved.


This book is dedicated to Daniel Paris.
\vfill

Special thanks to our families, friends, teachers, Bret Sweet, Sherri Paris, Kenneth Solis, Charlie Krank, Nicholas Nacario, and our fans met on /tg/ and other places, everyone else, everyone else’s grandmother, the moon, God, any other deities you might care to name \textit{except Moloch because he’s a tool}, Thomas Sowell, Zeljko Raznatovic’s pet tiger, Lu Bu, Mr. Rogers, Margaret Thatcher, G.K. Chesterton, Thomas Aquinas, Hidetaka Suehiro and Andrew Jackson

\vspace{1cm}
If you can think of anyone else we didn’t name, write it in. We even left a space. 

\vfill
opaquegames.com

facebook.com/OpaqueIndustries

opaque.freeforums.net
\end{center}
\clearpage
\phantomsection
\tableofcontents %BOOKMARK needs to be added here? 
\mainmatter
\chapter{Introduction}\label{ch:intro}
%%Save original \intextsep
	\newlength{\oldintextsep}
	\setlength{\oldintextsep}{\intextsep}
\newpage
\section*{Foreword}
It’s been a long time, now, since that first appearance back in 2012. We’ve gone through three names, a dozen updates, two hundred threads, and hundreds of thousands of posts. Oceans of blood have been spilled in the countless fechts fought in this game’s testing. Our little reservation on the internet should be marked forever by a mountain of human skulls.

To put it simply, it’s been a long road and we’ve come a long way.

\noindent And now we are here. This, the release of a new version. This will mark our first (nearly) complete 
version of Song of Swords to be released in many months. There have been many great and significant 
additions to the game, and numerous systems have been added to the core rules, which hitherto have been 
separate documents. This is, at least for now, the definitive Song of Swords. 

So here it is -- enjoy it. God knows you’ve been waiting long enough. This page is looking a bit empty, so 
I’m going to stick some Kipling in here. 
\vspace{0.8 cm}
{\large
\begin{verse}
\emph{As it will be in the future, it was at the birth of Man\\
There are only four things certain since Social Progress began. \\
That the Dog returns to his Vomit and the Sow returns to her Mire,\\
And the burnt Fool's bandaged finger goes wabbling back to the Fire; \\}
\vspace{0.4 cm}
\emph{And that after this is accomplished, and the brave new world begins \\ 
When all men are paid for existing and no man must pay for his sins, \\
As surely as Water will wet us, as surely as Fire will burn, \\ 
The Gods of the Copybook Headings with terror and slaughter return! \\}
\vspace{0.3 cm}
\hfill-\textbf{Rudyard Kipling}, \textit{The Gods of the Copybook Headings} 
\end{verse}}
	\newpage
\section{What is Song of Swords?}
	\begin{quotation}
{\small \emph{Mago laughed despite himself as he vaulted the barricade. The routing nomads were scrambling to reform in their makeshift wagon fort. Mago landed before a man struggling with the match of his arquebus. The man looked up. With an exultant whoop, Mago drew his saber and cut the man’s right arm off at the elbow in the same motion. The limb fell, its fingers still holding the powder-horn, and Mago walked past, pushing the man out of his way just as he began to scream at the stump of his own arm.}

\emph{A lancer approached, and Mago wove out of the way of the polearm, two-stepped forward, and snapped out with a cut that caught fingers between blade and haft, cleaving all of the digits from the spearman’s forward hand. He barked aloud in wordless triumph, and swept in to hammer aside a high blow from another swordsman. Grasping his foe by the collar, Mago saw his face. Little more than a child, brown eyes wide, mouth moving in wordless fear. Mago blinked, and then caved the boy’s head in with three terrible blows from the spiked pommel of his sword. He threw the corpse away, and flung himself into the melee, as his brothers closed the net. The slaughter was absolute, men who had raided and pillaged Dacian land for years died on their knees, begging for Dacian mercy.}\par
\emph{Mago brought his sword down, cleaving a pleading nomad’s head down through to the teeth. He wrenched his sword free and flung his arms wide. There were no more left to kill. The Fighting Volkodavs howled in triumph on a mountain of corpses. A gunshot rang out. Mago heard someone scream. The voice sounded familiar.}\par
\emph{Then he hit the ground. White-hot pain erupted from his upper thigh, where the bullet had struck him in the back. He realized that he was screaming, and all at once it hit him. The cold mud in his hair, on his face. The clawing agony in his leg. He knew that he would live, somehow. Limping for life, but alive. He forced open his eyes, through the tears of pain, and scanned the perimeter for the source. \textbf{Who?} He tried to breathe through gritted teeth, pain and fury foaming in his mouth. \textbf{Who shot me? I must know.} He felt a tooth crack as his eyes settled on the source of his ruin.}\par
\emph{The one-armed man, pale from loss of blood and fading fast, grinned over the smoking barrel of his arquebus.}}\par
	\end{quotation}
Song of Swords is a historical and fantasy tabletop role-playing game that puts its emphasis on realism. 
It's a game about intense life and death conflicts, adventure with peril around every corner, deep stories, 
involved characters, and no holds barred combat. 

Song of Swords isn't a game with hit points or ablative health. Here, injuries are devastating, debilitating, 
and dead-serious affairs. Even blows that don't kill immediately can kill with blood loss or infection from 
sloppy treatment. Survivors may suffer limps, lameness and even crippling disability. You will need sharp 
wits, thick armor, quick feet, great martial skill, or a combination of several of them, to survive. 

You decide your characters' goals, and through the Arc Points system, your character grows by 
progressing toward these goals, in addition to those chosen by the Gamemaster to give structure to the 
game's narrative. Violence may be a means to an end, but its danger means that you must pick your 
battles wisely, instead of charging headfirst into every combat situation. 

These elements combine to create a game in which the heroism of the characters is genuinely heroic. 
When even one slip-up can mean dismemberment or death, it takes real courage to face danger, and 
courage is what heroes are made of. This isn't a game for the faint of heart. 

Whether you fancy a campaign as 15th century Englishmen fighting in France under Henry V, or a 
desperate last stand against the Mongols in Hungary, Song of Swords can provide your historical role-
playing experience. Perhaps you’re more interested in fighting Grues in the hideous depths of the 
Kaselreich, or fighting Goblins in the brutal tunnels of Cross Way beneath the iron hills of Dace. Again, 
we’ve got you covered.
Welcome to Song of Swords.

\section{What is a Roleplaying Game?}
A roleplaying game like Song of Swords is a game in which the players assume the ‘roles’ of their 
characters and go on adventures with them, using their combined imagination. One player, the 
‘Gamemaster’ or GM, is in charge of running the game, playing the minor characters (Non-Player 
Characters, or NPCs) and guiding the players in their adventures.

To add tension and fun to the game, dice are rolled to decide the outcome of certain events like combat, 
dangerous acrobatics, stealth, and so on. As characters survive their adventures and overcome obstacles, 
they grow, becoming more capable and skilled. 

The rules in this book are here to help make sure everyone has a fun role-playing experience, the Song of Swords way. Remember, the goal here is to have fun! 
\section{Basic Mechanics Explained}
Song of Swords works on a simple dice Pool system, using only 10-sided dice, or d10s. Remember, 
whenever rolling is required, it will always be d10s that you will roll. Whenever there is a situation in 
which a character must accomplish a task that has some chance of failure, or that is reliant upon his own 
abilities for degree of success or failure, this is done through rolling dice. 
\subsection{Basics of the Dice Pool}
Rolls are simple affairs. You have a number of dice (a ‘Pool’) that you roll against a ‘Target Number’ or 
‘TN.’ Each die in the pool that surpasses this TN is a Success. 
Most rolls will have ‘Required Successes,’ or ‘RS,’ meaning the number of Successes necessary for the 
rolling character to win or succeed the test. And that’s it--it’s as simple as that. This kind of roll is called a ‘Check.’ 
\subsubsection{Basic Check}
A basic Check is where a character rolls a Pool, decided by one of their Attributes or Skills or some other 
factor, against an RS. The TN for these checks is always 7. The RS is decided by the GM, who analyzes 
the situation and the rules, and decides the appropriate RS according to the rules or his own judgement.\par
\textbf{Attribute Checks} usually use a single Attribute as the Pool. So, a Strength Check at RS 3 would involve a 
character rolling a number of dice equal to their Strength at TN 7. If they got 3 or more Successes they 
would succeed the Check.\par
\textbf{Skill Checks} use a character’s Attribute plus the appropriate Skill’s Skill Level to determine the Pool. 
Skills are detailed in their own Chapter, but the principle for rolling them is exactly the same as with 
Attribute Checks. Determine the RS, roll at TN 7, count the successes. Skill Checks and Attribute Checks 
are two kinds of basic Checks that follow these rules.  
	\begin{quote}
			\emph{Soot Macross, one of the infamous Macross Sisters, is trying to climb a 30-foot tall fortress wall escape capture. The GM tells her player to roll a \nameref{skill:Climbing} check. Soot’s Pool for the check is her Mobility (which is 10) plus her Climbing Skill (which is 4). She rolls 14 dice, of which 5 roll equal to or over 7, meaning she got 5 Successes.}

			\emph{The GM decides that since the inside of the wall is rough and covered in possible handholds, the RS is only 1 per 10 feet. Since the walls are 30 feet high, Soot’s roll of 5 was more than sufficient to get her to the top.} 

			\emph{But climbing down the other side is another question entirely… }
	\end{quote}
\subsubsection{Opposed Rolls}
	An Opposed Roll is when two characters roll against each other. In Opposed Rolls, there is usually not an 
RS. Instead, the character who gets more successes is victorious in whatever sort of contest they are 
engaging in.

	So if two characters are arm-wrestling, and the GM decides this is best represented as an Opposed 
Strength Roll, they would both roll their Strength at TN 7, and whichever of them got more successes 
would win.

	In the event of a tie in successes in an Opposed Roll, either the two characters are evenly matched (if such 
a thing is plausible in the situation) or they must roll again to break the tie. The GM must decide which is 
more appropriate given the situation. 
	\begin{quote}
		\emph{Richard and Dutch are having a drinking contest. The GM decides that this should be an Opposed Health 
Roll. The two characters roll Health at TN 7. Richard gets more successes, so the GM rules that he outdrinks his friend 
and wins the contest.}
	\end{quote}
\paragraph{Non-Standard TNs:}
For most of the game, TNs are always 7. However, in a few instances, particularly in combat, there are 
different TNs for things like weapons and certain combat maneuvers. 


As a result, this book will always list what the TN for a test is. However, if in any event you see a Check 
without a listed TN, assume that that TN is 7. 
 \paragraph{Rounding:}
In any situation in which a number would be divided, such as determining Compound Attributes or the 
weight of a character’s equipment, always round down unless specifically told not to, such as when using 
the Abnormal Class/Wealth Combinations special rule, or when calculating asymmetrical armor weights. 
Otherwise, always round down!
\chapter{Character Creation}\label{ch:charcreation}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\newpage
	\begin{quote}
		\emph{“It is thus the beauty of Genosism that its means are its ends. Under such a system there can be no confusion as to the right course of action, no contrivance of the ends justifying the means, because you are not pursuing an objective the route to which must be interpreted, but exercising a lifelong discipline the very practice of which is its own purpose.”}\par

		\hfill -\textbf{Pontifatrix Lea I}, \textit{Annotations on The Solar Scriptures}, Xth Edition 
	\end{quote} 
Song of Swords is a game about characters; those controlled by the players are called Player Characters 
(PCs), while the rest are controlled by the Gamemaster (GM), and are called Non-Player Characters 
(NPCs). In order to ensure a fun and balanced game for everyone, all characters are created using a 
system called Character Creation. That is, all PCs and NPCs are created using the steps in this chapter. 

In Character Creation, the GM establishes whether the campaign will be a fantasy or historical campaign, 
and sets the power level of the campaign (Campaign Power), which allows players to balance their 
characters relative to one another. 

Characters are built using Player Creation Points (PCP) which are spent on different aspects of characters, 
called Categories, which includes their Attributes, Skills, Proficiencies, and so on. The amount of PCP a 
character has to spend at Character Creation is based on the Campaign Power. 

\section{How to Create a Character}
Here’s a step-by-step guide to making a character in Song of Swords. It might be a bit tricky at first, but 
once you understand the steps it’s actually very simple.

\subparagraph{Step 1: Set the Campaign Power}
The GM will set the Campaign Power and explain the theme of the campaign. The Campaign Power will 
determine how many PCP you have available to make your character, as well as the maximum PCP you 
can invest in any one Category. PCP are the building blocks of your character during this process, so the 
more of them you have, the more powerful your character will be. Refer to~\autoref{tab:campaign_power}, below, for a guide on Campaign Power.

\subparagraph{Step 2: Develop Your Character Concept}
Decide what kind of character you are making. Your character is your avatar for interacting with the game 
world, and the nature of the character you create will dramatically change your roleplaying experience. 
When deciding on a character concept, it’s a good idea to factor in what the GM has said the campaign 
might be like and what the other players are building. Players should communicate with one another to 
coordinate Character Creation, either to create a group with fun dynamics, prior relationships, or 
complementary and conflicting personalities. This makes for a more engaging group dynamic when play 
begins! 

\subparagraph{Step 3: Spend PCP on Categories}

Now, divide your PCP between the Categories outlined in \autoref{tab:PCPInvestment}~below. There are six mandatory 
Categories, plus one optional one, Magic. You do not have to put any points into Magic, but \emph{all other 
Categories must have at least 1 PCP put into them.} For more detail on these Categories, and what your 
PCP investment will mean, refer to their respective chapters. 

\begin{description}
\item [Race:] Your character’s birth race. Different races have dramatically different abilities. Dealt with in \autoref{ch:races}.
\item [Attributes:] Your character’s base physical and mental ability scores. Having strong Attributes makes a character more capable in general. See: \autoref{ch:attributes}.
\item [Skills:] Your character’s level of aptitude in various disciplines. Skills are very important for being able to do things both inside and outside of combat. See: \autoref{ch:skills}.
\item [Proficiencies:] Your character’s proficiency in fighting. See \autoref{ch:proficiencies}.
\item [Social Class and Wealth:] Your character’s standing in the world, and also the amount of money your character starts with. For good reference, a knight’s full kit of plate armor, a good helmet, a middle of the road warhorse, a sword, a dagger and a lance would cost over 14 gp. By contrast, a regular footman’s kit, including a padded jack, jack chains, a skullcap, and a spear, shield and dagger would cost less than 1 gp. See: \autoref{ch:wealth}.
\item [Magic:] Not included in Beta.
\item [Boons and Banes:] Your character’s various traits, quirks and flaws. See: \autoref{ch:boons}.
\end{description}

\subparagraph{Step 4: Buy Equipment and Supplies}
The resources your character has at their disposal will have a great impact on your campaign. Be sure to 
equip your character with the items that they will need. The quality and number of items available to your 
character is largely affected by your character’s Wealth. Refer to \autoref{ch:equipment} for more information about buying equipment and supplies.

\subparagraph{Step 5: Review Your Character with the GM}
At this stage it is important to review your character with the GM. Your GM will check whether you have 
spent your PCP appropriately in each of the required Categories, and supplied your character with 
necessary equipment. 

\subparagraph{Step 6: Determine Arcs}
The next step of Character Creation is to determine your character’s Arcs. Arcs represent your character’s 
goals, drives and beliefs that guide them through life, especially during the campaign itself. These provide 
Arc Points, which are used to advance your character’s Abilities and Skills. For more information on Arcs 
and Arc Points, go to \autoref{ch:arc}

\subparagraph{Step 7: Name and Describe the Character}
Finally, give your character a name and describe them. Be mindful of the setting of the campaign, as well 
as your character’s race, gender and social class when choosing a name. Perception is everything; a great 
name can evoke strong imagery, and how characters look and conduct themselves may influence how 
other characters and NPCs interact with them. 

Once you have completed Step 7, you are ready to play! 

\section{Campaign Power}
As explained above, the Campaign Power determines the type of campaign you will play, as well as the 
general potency of characters. In a campaign set at ‘Gritty Power’, players are limited in the amount of 
PCP they have available to spend. This means that characters tend to be more specialized and realistic, as 
they have less PCP to play with than, say, a ‘Legendary Power’ campaign. The following table explains 
the amount of PCP available for different levels of Campaign Power. 

\begin{table}[hb]
	\centering
\caption{Campaign Power}
\noindent\begin{tabu} to \textwidth {X[l,2] c X[c]}
\rowfont[c]{}Campaign Power & PCP & Max PCP per Category\\\toprule
Gritty Power (realistic, low-powered historical campaigns) & 18 & 6 \\
Low Power (realistic, medium-powered historical and fantasy campaigns) & 22 & 7 \\
Medium Power (heroic historical and fantasy campaigns) & 26 & 8 \\
High Power (high fantasy campaigns) & 30 & 10 \\
Legendary Power (epic fantasy) & 34 & 10 \\
\end{tabu} 
\label{tab:campaign_power}
\end{table}

\subsection{Player Creation Points}
PCP are spent on Categories, which are outlined in their own chapters. The amount of PCP available to 
spend at Character Creation is determined by the Campaign Power, as is the maximum PCP per Category 
allowed. 

\autoref{tab:PCPInvestment} below explains what your PCP investment will mean for your character. Each of the vertical columns represent a Category in Character Creation. The horizontal rows represent PCP investment. So, 
if you put 4 PCP into Race, you get Tier 3 Race options. If you put 3 PCP into Social Class and Wealth, 
your character is a Poor Freeman and starts with 3 gp. If you put 10 PCP into Attributes, you get 48 
Attribute Points to spend on various Attributes. 

\begin{table}[hb]
	\centering
\caption{PCP Investment}
\label{tab:PCPInvestment}
\begin{tabu} spread 0pt {c l c c c X c c}
PCP & Race & Attributes & Skills & Profs & Social Class/Wealth & Magic\footnote[1]{Not in Beta. Except Ascended} & Boons and Banes\\
\midrule
1 & Tier 1 & 88 & 0 & 0 & Slave/Nothing & 5 & -15 \\
2&Tier 2&92&3&3&Peasant/(1 gp)&10 & -10 \\
3&-&96&6&6&Poor Freeman/(3 gp)&15 & -5 \\
4&Tier 3&108&9&9&Freeman/(5 gp, [1W])&20 & 0 \\
5&-&124&12&12&High Freeman/(10 gp, [2W])&25&5\\
6&Tier 4\footnote[1]&140&15&15&Minor Noble/(100 gp, [3W])&30&10\\
7&-&160&18&18&Landed Noble/(250 gp, [6W])&35&15\\
8&Tier 5\footnote[1]&180&21&21&High Noble/(500 gp, [10W])&40&20\\
9&-&200&24&24&Royalty/(1000 gp, [15W])&45&25\\
10&-&224&27&27&High Royalty/(1500 gp, [20W])&50&30\
\end{tabu}
	\vspace{5pt}
		\caption*{$ ^1 $Not in Beta. Except Ascended}
\end{table}

\subsubsection{General Rules for PCP Investment}
You must invest at least 1 PCP in each Category (with the exception of Magic). Even historical 
campaigns must spend 1 PCP in the Race Category, even if the only option is a Human character. 

As \autoref{tab:campaign_power}: \nameref{tab:campaign_power} explains, no Category can have more PCP devoted to it than the maximum allowed for the 
set Campaign Power. For example, if your Campaign Power is set at 'Gritty Power`, you may not spend 
more than 6 PCP in any Category. 
\part[Vital Stats]{De Humani Corporis Fabrica}
\chapter{Arc and Advancement}\label{ch:arc}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Characters grow over the course of every story. This chapter explains how your character advances their 
own story, as well as improve their Attributes, Skills, Proficiencies and other qualities during gameplay. 

The advancement system in Song of Swords is called ‘Arc’, which represents your character’s goals, 
drives and beliefs. By working towards these goals, following their drives, and adhering to their beliefs, 
your character will improve their Attributes, Skills, Proficiencies and other qualities, as well as progress their own storyline 

Advancement is made through the expenditure of Arc Points, which is explained below. 

 
\subsection*{What is Arc?}
The Arc system is a mechanic within the game that tracks your character’s goals, beliefs and motivations, 
and allows the GM to reward the character with Arc Points for acting according to those motivations. Arc 
Points can be spent to improve a character’s Attributes, Skills, and Proficiencies and other qualities. 
Characters do not become stronger merely by killing enemies or winning fights, they become stronger by 
pursuing their goals, fighting for what they believe in, and for moving along their part in the story through 
action. 

 
Generally, whenever a character accomplishes or indulges in one of their Arcs, they gain one or more Arc 
Points. For example, if a character’s ‘Glory Arc’ is ‘glory in battle,’ and he acts gloriously in battle, then 
he should gain Arc Points. The number of Arc Points gained is at the GM’s discretion, however Table 
XX, below, provides suggestions for the amount that should be rewarded. 

 
For a several-hour session, the GM should aim to award about 4-6 Arc Points per character, so long as the 
players have been properly engaged. 
 

There are five Arc Categories, each detailed below, along with how Arc Points are gained in them. Arc 
Points are recorded on the Character Sheet. 
\section{Arc Categories}

\subsection{Saga Arc}
The Saga Arc is one shared by the entire group of PCs. It is either decided by the GM, or by the group as 
a whole, at the GM’s discretion. It can also be shared by any group or organization, such as an entire 
Crusade sharing the ‘Take Damascus’ Saga Arc, however it is best used with a group of PCs unless that 
organization is integral to the plot or premise of the campaign. 
 

The reward for pursuing a Saga is gradual, with a large payoff. Taking a significant step towards 
completing a Saga Arc, such as winning a major battle, defeating a major enemy, or overcoming a 
significant obstacle, should grant each character several Arc Points. Fulfilling the Saga Arc should gain 
the characters a large number of Arc Points, up to as many as 18. 


Examples:
\begin{itemize}
\item ‘Slay the dragon that destroyed our ancestors’ kingdom and reclaim our legacy’ 
\item ‘Win Swiss independence from Austrian rule’ 
\item ‘Fight Prince John’s tyranny until the return of the True King’ 
\end{itemize} 
 
\subsubsection{Suspending a Saga Arc}
In cases where players are not working as a team, not a part of the same storyline, or acting in opposition 
to each other, the Saga Arc can be suspended or completely removed from the campaign. 

 

\subsection{Epic Arc} 
An Epic Arc is much like a Saga Arc in that it covers a wide-ranging, profoundly important goal, save for 
the fact that it applies only to the character taking it. Epic Arcs should be deeply personal to the character 
and be one of their driving forces in life. 

 
Rewards for the Epic should be more frequently staged than Saga rewards, but the number of rewarded 
Arc Points should be fewer for their frequency. 

 

Examples: 
\begin{itemize}
\item ‘Save my wife from slavers’ 
\item ‘Avenge my father’s murder’ 
\item ‘Obtain enlightenment’ 
\end{itemize}
 

\subsection{Belief Arc}
The Belief Arc does not cover a specific storyline, but instead focuses on a core belief, concept, maxim, 
faith, oath, value, virtue or code that a character believes in, even if they don’t actively live it. When 
acting in accordance with the Belief Arc, especially in the face of adversity, or when it causes a story or 
encounter to happen, the character is rewarded. 

 
Many cultures have embraced very rigid codes of honor, not just on the field of battle, but in all areas of 
life. All characters, in both historical and fantasy campaigns, can make good use of this Arc by picking 
their culture’s code of honor, or one of their own devising. 

 
Belief Arcs differ from other Arcs in that they can cover a broad variety of concepts, principles and 
motivations. In essence, any concept that a character would be willing to fight for, not for direct gain, but 
simply out of principle, may be a Belief Arc. 
 

Examples: 
\begin{itemize}
\item ‘Islam’ 
\item ‘Christianity’ 
\item ‘The gods and glory of Rome’ 
\item ‘The honor of my family’ 
\item ‘The pursuit of knowledge’ 
\end{itemize}
 

\subsection{Glory Arc} 
The Glory Arc has no strict storyline or goal save for the acquisition of glory, defined here as praise, 
renown, honor or fame by common consent for accomplishing daring and dangerous deeds. Arc Points 
are typically rewarded for doing anything wildly impressive, such as surviving perilous situations like 
deadly combat or outrunning a massive boulder — any feat of success when the odds are stacked against 
the character. 

 
In considering Glory, Arc Points may be awarded in a number of situations, such as winning a fight 
against someone objectively ‘better,’ winning against multiple people, accomplishing a difficult skill test, 
or rolling particularly well on any pool roll or test. The character is also encouraged to define what their 
own version of Glory is, such as behaving honorably, earning adoration and fame, or triumphing over 
their peers, for which the GM should offer more Arc Points. 

 

Examples: 
\begin{itemize}
\item ‘Gaining glory in battle’ 
\item ‘Protecting the weak’ 
\item ‘Helping the poor and the sick’ 
\item ‘Performing surgical miracles’ 
\item ‘Making great discoveries’ 
\item ‘Earning fame’ 
\end{itemize}
 

\subsection{Flaw Arc} 
Only Humans have access to the Flaw Arc, because of their racial characteristic, The Human Condition. 

A Flaw Arc represents a character’s impulses that lead them contrary to their goals. Heroic characters 
often have great flaws or weaknesses of character, and these weaknesses provide part of the character’s 
narrative that is indispensable to their growth. Flaw Arcs are chosen at Character Creation, and may be 
altered in the event of the Flaw being resolved, or being rendered somehow superfluous to the character. 

A Flaw is indulged when it actively complicates a character’s larger endeavors. 

Examples:
\begin{itemize}
\item ‘Debilitating vices (alcoholism, etc)’ 
\item ‘Prejudices that lead characters to trouble (racism, religious radicalism, nationalism, etc)’ 
\item ‘Conflicting character motivations (divided loyalties between two kingdoms, conflicting oaths, 
religious loyalties contrary to goals, etc)’ 
\end{itemize}
So, a character’s list of Arc Categories and their descriptions might read something like this (\autoref{tab:ArcPointFormat}), if the 
character were a brave Tigurnian freedom fighter, trying to free his homeland from the clutches of the 
Kaselreich.
\begin{table}[!h]
	\centering
	\caption{Arc Point Format}
	\label{tab:ArcPointFormat}
	\begin{tabu} spread 0pt{X[-1,c]X}
Arc Category & Description\\ 
Saga&Defeat the Kasels and save Tigurnia \\
Epic &Defeat the Kasel army led by The Red Duke \\
Belief&Ruvian Genosism (Religion) \\
Glory &Feats of daring for the Homeland \\
Flaw&Quick to trust \\
	\end{tabu}
\end{table}

\section{Awarding Arc Points} 
Arc Points are awarded by the GM when a player has acted in accordance with any of their Arcs, and can 
be awarded either at the moment of successful accordance or at the end of the gaming session, at the 
GM’s discretion. There is no limit to the amount of Arc Points a player may earn in a session, or that a 
GM may distribute, though some moderation is advised. A very well-played session should run at about 
4-6 Arc Points to each player. 

Awarding the correct amount of Arc Points can be tricky. \autoref{tab:AwardingArcPoints} provides guidelines to a fair amountof Arc Points to distribute per character, per session, however there may be reasons to award more or less 
than this suggestion. Some characters will focus on Arcs that they enjoy, or otherwise want to see through 
to completion, warranting extra Arc Points. GMs may want to give more Arc Points to players who go the 
extra mile and through active participation, help progress and enrich the experience of the session. On the 
other hand, GMs may wish to refrain from rewarding players who mess around the entire session and do 
not play the game at all. These are simply guidelines, however; each group is truly different. 

It can be helpful to compare Arc Points to an action movie. Each different Arc represents a motivation for 
a character, whether they know it or not at the beginning. Each Arc can be an important scene in the 
session, with Belief reflecting the characters’ convictions, Glory representing the action scenes that drive 
the narrative forward, and Saga being progress for the over-arching plot, with Epic being what drove them 
into the plot in the first place. 

\begin{table}[b]
	\centering
	\caption{Awarding Arc Points}
	\label{tab:AwardingArcPoints}
		\begin{tabu} spread 0pt {X[-1,c]X[l,m] } %TABLE needs to be fixed, as the vertical lines are garbage, maybe wraptab?
\rowfont[c]{}Arc Category&Arc Points Awarded Per Session \\\cmidrule{1-2}
Saga
	&0-6\newline
		18 (upon completion of Arc)\\
	\\
Epic
	&0-6\newline
		18 (upon completion of Arc)\\
	\\
Belief&0-4\\
		\\
Glory&0-4\\
		\\
Flaw&0-2\\
		\end{tabu} 
\end{table}
The amount of Arc characters have spent is a direct correlation to how powerful they are, and players 
should keep track of the amount of Arc they spend on their Character Sheet. It is important for the GM to 
consider how much Arc characters have spent when planning future encounters with increasingly more 
powerful opponents. 

Saga and Epic Arcs have higher caps as they are representative of greater plot points and should be 
rewarded more substantially, albeit obtaining the goals of these Arcs should be more difficult as well. 
Generally, any great advancement towards the goals of the Saga and Epic Arcs should be worthy of being 
rewarded, though the amount should be tempered with the effort put in, level of roleplaying, success in 
rolls, and finally just how much was done in obtaining the goals. Final completion of a Saga or Epic Arc 
should award the character with as many as 18 Arc Points. 

 
\subsection{Arcs Ending and Changing}
It is fully expected that Arcs will be finished or changed throughout the course of play. Should a character 
be successful enough to complete their Saga or their Epic Arc, they are allowed to find new ones. If a 
character has adequate reason to change their Saga or Epic mid-session, they are fully welcome to. 
However, if there is no compelling reason for a player to change either Arc, and it is simply because they 
want to, then at the GM’s discretion the player should not be allowed to gain Arc Points for the new Arc 
until the next session. 


\subsection{Character Advancement}
At any time, including during combat, characters can spend Arc Points to upgrade elements of their 
character. Arc Points spent this way are removed from the character’s total, and the upgrades (unless 
otherwise specified by the GM or the rules) take effect immediately. 


Refer to the chapters on Attributes~(\autoref{ch:attributes}), Skills~(\autoref{ch:skills}), Proficiencies~(\autoref{ch:proficiencies}), and Boons and Banes~(\autoref{ch:boons}) for more information on how to advance these by spending Arc Points. 
\chapter{Attributes}\label{ch:attributes}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Attributes refer to your character’s raw physical and mental abilities, which affect all learned Skills, 
Proficiencies, as well as movement and general adventuring. These are the base components of a 
character’s makeup: their strength, their perceptiveness, their agility of body and mind. 


There are eight core Attributes, being Strength, Agility, Health, Endurance, Willpower, Wit, Intelligence 
and Perception. Characters also have six Compound Attributes, which are made using combinations of 
core Attributes, as they are used for tasks that would logically require the use of more than one core 
Attribute. The Compound Attributes are Adroitness, Mobility, Carry, Toughness, Charisma and Grit. 
 

Attributes can be affected by Race, Skills, Proficiencies, as well as Boons and Banes. It is a good idea to 
understand what Attributes will be useful to your character, and factor in how your Attributes can be 
modified during play. 


That sounds like a lot of attributes to worry about! Well don't worry, follow along and it'll be simpler than 
you know. Not all characters will need to worry about all of these Attributes. Some may focus on 
Strength, where others may focus on Agility. Attributes are the core tools of your character’s body and 
mind, so don't be afraid to start sculpting! 
\section{Character Creation}
When creating a character, all Attributes start at 1. You receive a number of Temporary Arc Points 
indicated on \autoref{tab:PCPInvestment}~regarding~\nameref*{tab:PCPInvestment}~in~\autoref{ch:charcreation}, and may spend them to upgrade your 
Attributes as-per Advancing with Arc. The higher an Attribute is at the start, the more Arc it costs. To 
simplify this, \hyperref[tab:AttributesatCharCreation]{here}~is a table with the total cost of upgrading an Attribute from 1 at Character Creation.
\begin{table}[hb]
	\centering
	\caption{Attributes at Character Creation}
	\label{tab:AttributesatCharCreation}
	\begin{tabu} spread 0pt {XX[-2,c]}
Attribute Level&Arc Cost\\\toprule
1&-\\ 
2&4\\ 
3&8\\ 
4 (Human Average)& 12\\
5&16\\ 
6&20\\ 
7&24\\ 
8&28\\ 
9&34\\ 
10 (Human Maximum)&40 \\
11&46\\
12&52\\
13 (Mortal Maximum)&58\\ 
	\end{tabu}
\end{table}
When determining starting Attributes, remember that no character can have an Attribute lower than 1 at 
Character Creation. This includes racial modifiers; you may not build a character whose Attributes would 
be lowered below 1 by racial modifiers or any other factors. 

\textbf{The normal limit for any Attribute at Character Creation is 8}. Races that have a bonus or penalty to an 
Attribute also change their limit for that Attribute by an amount equal to their bonus or penalty, positive 
or negative. 

Zells have a +2 bonus to PER. Therefore, the PER limit for Zells is 10 at Character Creation. 

Goblins have a -2 penalty to STR. Therefore, a Goblin’s maximum STR is 6 at Character Creation.
\section{Advancing with Arc}
During play, your character can advance their Attributes by spending Arc Points (refer to \autoref{ch:arc} for 
more information on Arc). 
 

Each Attribute must be at least 1 at Character Creation. Every level gained thereafter requires a certain 
amount of Arc Points. 

 
Each level must be gained sequentially. To have your character’s PER increase from level 7 to level 10, 
you must pay a total of 16 Arc Points (4 for level 8 + 6 for level 9 + 6 for level 10). 
 

Note that only core Attributes can be advanced using Arc, as Compound Attributes will increase naturally 
as core Attributes increase. The Compound Attribute of Grit advances separately, however, and is further 
detailed under its description. 
\begin{table}
	\centering
	\caption{Arc Cost of Attributes}
	\label{tab:AttribsArcCost}
\begin{tabu} spread 0pt {Xl}
Attribute Level & Arc Cost\\\
1& -                        \\
2& 4                        \\
3& 4                        \\
4 (Human Average)& 4        \\
5& 4                        \\
6& 4                        \\
7& 4                        \\
8& 4                        \\
9& 6                        \\
10 (Human Maximum)& 6       \\
11& 6                       \\
12& 6                       \\
13 (Mortal Maximum)& 6      \\
\end{tabu}
\end{table}
\subsubsection{Non-Human Races and Attribute Maximums}
Humans max out their Attribute stats at 10, but what about races who have bonuses to their Attributes? A 
character with a racial bonus to an Attribute can raise that Attribute to 10 plus the bonus amount, at the 
costs detailed in \autoref{tab:AttributesatCharCreation}
\section{Core Attributes}
\subsection*{Strength (STR)}
Strength is a character's ability to apply force, whether that be lifting a cart off a trapped child or 
smashing someone's head in with a warhammer. This covers your character’s full body strength, not just 
how many phone books he can rip in half. Strength influences many physical activities that are based on 
force and strength, but also determines the base damage inflicted by melee weapons in combat. 


Strength contributes to the Toughness and Mobility Compound Attributes.
\subsection*{Agility (AGI)} 
Agility refers to your character’s physical and manual dexterity. Agility’s domain includes acrobatics, 
balancing, and other acts of flexibility. It is an important Attribute in combat, as well as in various 
physical activities and skills. 

Agility contributes to the Adroitness and Mobility Compound Attributes. 

\subsection*{Endurance (END)}
Endurance determines your character’s stamina and ability to cope with physical strain and exertion. 
Endurance is a combination of qualities, including cardiovascular development, which contributes to your 
character being able to continue strenuous activity for extended time frames. Endurance also factors into 
how well your character can handle \nameref{sec:Blood Loss} and governs certain Skills. 

Endurance contributes to the Toughness, Carry and Mobility Compound Attributes. 
\subsection*{Health (HLT)}
Health represents your character's raw constitution, physical robustness, resistance to disease and 
infection, and general state of well-being. When injuries are sustained, having higher Health makes 
wounds easier to recover from. 


Health contributes to the Toughness Compound Attribute. 
\subsection*{Willpower (WIL)}
Willpower governs your character’s force of will, determination, and strength of personality. 
It allows your character the ability to focus through distraction, control their emotions, and resist pain, 
intoxication, and other (sometimes magical) influences. It is also a core component in a strong personality 
and thus factors heavily into human interactions. Plus, Willpower reduces the sum total of Pain from all 
Wounds. 

 
Willpower contributes to the Charisma Compound Attribute, and is also used in the Magic system. 
\subsection*{Wit (WIT)}
Wit is the speed and flexibility of the mind. The Attribute determines your character’s skill in 
improvisation, and ability to cope with rapid bursts of information and activity without being 
overwhelmed. From telling jokes to dueling with swords, Wit is important for many activities, especially 
certain Skills, and is one of the Attributes that no hero should be without. 

 
Wit contributes to the Adroitness and Charisma Compound Attributes. 
\subsection*{Intelligence (INT)}
Intelligence references your character's ability to collate, process, recall and connect information in a 
logical manner. It doesn’t speak to how 'smart' your character sounds or acts, but rather how good he is at 
truly understanding concepts and analyzing facts. Intelligence is required primarily in engineering, 
architecture, linguistics, and sciences, where thinking fast isn't as important as thinking clearly and 
meticulously. Intelligence also aids in analyzing ideas for contradictions or inconsistencies. The Attribute 
determines your character’s ability to examine and retain information, and also to apply logical processes 
to facts. 

Intelligence is used extensively in the Skill and Magic systems. 
\subsection*{Perception (PER)}
Perception determines your character's awareness of their surroundings, which includes visual, audible, 
and even olfactory (smell) awareness. Perception is important, as it allows your character to spot 
ambushes and gauge distances. It is the basis of many Skills, and also plays a major role when using 
ranged weapons. In addition, Perception allows characters to read the expressions of others, spotting 
miniscule movements and cues, which allows greater control over social interactions with other people. 

 
Perception contributes to the Charisma Compound Attribute. 
\section{Compound Attributes}
Compound Attributes are made from an average or aggregate of multiple core Attributes. These 
Attributes are determined after Character Creation is completed, and factor in your character’s core 
Attributes as well as any penalties or bonuses related to Race or Boons and Banes. 
\subsection*{Adroitness (ADR): $\frac{\mathrm{AGI+WIT}}{2}$}
Adroitness measures your character’s physical articulation, speed, and mental alacrity. It represents their 
reflexes and their coordination between mind and body. Adroitness is obviously very important in 
fighting, as it contributes directly to your character’s CP, but it also helps them avoid being tripped, 
knocked over, thrown from rocking horses, flung over the sides of ships, and other things that can be 
avoided with a combination of quick thinking and action. 

 
\subsection*{Mobility (MOB): ($\frac{\mathrm{STR+AGI+END}}{2}$)}
Running, jumping, climbing: these are all determined by Mobility. Your character can move a number of 
yards equal to his Mobility each Round during combat, or run twice that number, or faster with certain 
Skills and armor enhancements. 
\subsection*{Carry (CAR): $(\mathrm{END} \times 2)$}
Carry determines how much weight in armor, equipment, and other inventory your character can lug 
around before being encumbered. See \autoref{sec:exhaustion} for more details.


\subsection*{Toughness (TOU): ($\frac{\mathrm{STR+END+HLT}}{3} $)}
Toughness refers to how resilient your character is to harm. It represents thickness of skin, hardness of 
bone, and layers of callous and scar tissue. Toughness reduces the amount of damage your character takes 
when they are attacked. 

\subsection*{Charisma (CHA): ($\frac{\mathrm{WIL+PER+WIT}} {2}$)}
Charisma determines how well your character interacts with other characters. It’s their animal magnetism, 
their ability to read people and appeal to other’s emotions, or ‘get’ them. Charisma also influences your 
character’s ability to hide their own emotions, or to express themselves in an impassioned manner; it is 
used in social interactions of all sorts. Charisma is an important Attribute in the Magic system. 

 

\subsection*{Grit: ($\frac{\mathrm{WIL}}{2}$)} 
Grit is a character’s accumulated resistance to pain, fear, and the shock of injury. Grit reduces the total Pain a character has accumulated. (see \autoref{ch:injuryrecovery}). Characters with high Grit can continue fighting 
through terrible injuries, and keep calm in the face of unspeakable horror. Many people acquire high Grit 
by surviving combat, slaying foes, and enduring injuries. 


Your character’s starting Grit is determined by their WIL score, but unlike the other Compound 
Attributes, it is not tied to WIL afterwards. Increasing WIL after Character Creation has no effect on Grit, 
nor can it be purchased with Arc Points. Instead, as detailed below, Grit increases through experience. As 
your character is confronted with violence, injury, bloodshed and terror, they have the chance to increase 
their Grit scores. 
 

\section{How to Increase Grit}
When your character experiences, witnesses or perpetrates an action above their Grit level, the GM may 
decide that they should make a Grit test. This is a \textbf{WIL} check with RS equal to the difference between the character’s current Grit and the level of severity of the experience (as determined on the chart below) to a maximum RS of 3. 

\begin{longtabu} spread 0pt {lXX}
	\caption{Grit Advancement}
	\label{tab:GritAdvancement}\\
Grit&Description&Example Advancement Threshold\\
0&Totally innocent and inexperienced. Average child.&Hurting someone’s feelings, or having one’s feelings hurt. \\
1&Sensitive, sheltered, still learning about the basics of social interaction. Average young person. & Physically hurting someone else (even accidentally), being hurt oneself. \\
2& Mature person with an understanding of the way the world works. Average adult. &Being confronted with death. \\
3&Worldly person with experience in the darker elements of society.& Witnessing violent death.\\
4&Tough person, fully aware of the grim nature of life, no longer fazed by the unrelenting wickedness of the world they live in. &Killing somebody. \\
5& Hardened person, now capable of facing the harshness of life head-on.&Killing several people. \\
6& Jaded individual, increasingly callous to both enduring and witnessing pain. &Witnessing violence and death in a large scale. \\
7& Callous and unfeeling, this person’s empathy is buried deep down to shield them from the pain of enduring hardship and witnessing it in others. &Killing or facilitating the killing of large numbers of people. \\
8& Empty inside, this person has locked away their feelings so long they’ve almost completely lost touch with them. &Perpetrating genocide, mass torture, or surviving the same.\\
9&At this point this person is no longer technically sane. This level of indifference to pain and suffering, both of the self and others, is a mental illness. &Witnessing everything you have ever cared for be destroyed. \\
10& Almost completely inured to fear and pain. Totally fearless. &Destroying everything you have ever cared for with your own hands. \\
		\\
		\caption*{Success: Character gains 1 Grit. \qquad Failure: Character does not gain a point of Grit.} %NEEDS multicol or something
\end{longtabu}
\chapter{Races}\label{ch:races}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quotation}
\emph{"In the beginning, there was nothing. Then came He Who Came First, and The Twelve Great Old Ones. They created the cosmos, 
and then departed. As Steward of this world they left Genosus, God of Light, Life, and Justice, to illuminate all, and transform it 
according to His will.” }

\hfill--Genosian Creation Myth, \textit{The Solar Scriptures}
\end{quotation}
\begin{quotation}
\emph{"In the beginning, there was no light. Then came He Who Came First, and The Twelve Great Old Ones. They illuminated the 
cosmos, and saw it empty. They departed, but to bring beauty to this barren world, they left Bocanadessia, Goddess of Love, Life, 
and Beauty, to preside over all in her majesty."} 

\hfill--Dessian Creation Myth, \textit{The Moonsong} 
\end{quotation}
\begin{quotation}
\emph{/Remembering Now/Time of Starting/Predating/Memory/Story/Holocaust/Comes To Pass/\\
/Bubbles Together/Bubbles Sorted/Bubbles Purposeful/Attributed To/The Seven/Family/Together}


\emph{/To This Place/Many Hands/Arriving Loudly/Holocaust/Perfect World/Burned Away/Accident\\
/The Seven/Burned Shadows/Ropes and Sails/Nightmare/Prison/Eternity/Construct }

\hfill--Zellish Creation Myth, Burdinadin Translation Attempt 
\end{quotation}
\section*{Character Creation}
The fantasy setting of Song of Swords has a number of playable races, however historical campaigns may use only Humans. Not all races are created equal, and some possess powers, abilities, weaknesses or qualities that set them apart from (and thus worth more or less than) Humans.  
The amount of PCP devoted to a character's Race score determines which races are available. As abilities and traits made available by choosing a Race are Boons and Banes of a kind, they can be recorded in the Boons and Banes section of the Character Sheet, however racial modifiers do not affect the purchasing of 
Boons and Banes at Character Creation. 
The number of PCP placed into the Race section determines what ‘tier’ of races are available for the character, as shown in \autoref{tab:racetiers} below. 
\begin{table}[!h]
	\caption{Race Tiers}
	\label{tab:racetiers} 
	\begin{tabu}{ccX}
PCP&Tier&Races\\
1&Tier 1&Human, Goblin\\
2&Tier 2&Dwarf, Zell\\
4&Tier 3&Burdinadin, Ohanedin\\
6&Tier 4&Orredin\\
8&Tier 5&Star Vampire, Sarturi Chosen, Genosian Paladin, Dessian Silver Guard, Paleolithic Elf\\
	\end{tabu}
\end{table}

\subsubsection{Historical Campaigns and Race}
All characters in a historical campaign will, naturally, be Human (at least until the aliens figure out how 
to use wi-fi), so assume that all historical characters automatically spend 1 PCP in Race. 
\section{Humans, Dwarves and Goblins}
\subsection{Humans}
\begin{quotation}
	\emph{"Humans pursue death with zeal because they know their mortality. 
	Deep down, they do not want to live long enough to die."}\par
	\hfill--\textbf{Nikephoros}, Helian Philosopher 
\end{quotation}
\begin{quotation} 
	\emph{"My arms are tired, my feet are cold; 
I wish I were at home eating boar."}\par
	\hfill--\textbf{Sir Wilhelm the Strong}, Beleaguered Knight
\end{quotation}
\subsubsection*{Overview}
Human beings, also called the Race of Men, the Hostoadin, and a host of other names and monikers, are the most populous and widespread race in the Tattered Realms, rivaled only by the Zells, and potentially the Goblins. 


Humanity is a flexible race, and one that builds communities naturally. Humans form nomadic tribes and 
pastoral communities as easily as they form city-states and great empires. There does not seem to be an 
environment which they favor over any other; from burning deserts to frozen tundras, lush valleys to 
desolate wastelands -- anywhere men can live, men do live.


The Human tendency to build communities brings with it a sense of tribalism that often turns them 
against one another, and thus war is the one constant of Human civilization. Yet, so too have they 
contributed universally to art, philosophy, science and religious wisdom. The Human race is one of 
infinite potential, both for civilization and debauchery. 
 

\subsubsection*{Appearance} 
Human height varies depending on climate, diet and social conditions, but at this time the average man is about 5'6" and 160 lbs, while the average woman is 5'4", weighing about 120 lbs. Human beings come in many pigments and colorations, though all are descended from common ancestors and are ultimately 
similar.


Human beings are young adults by 16 years of age, middle-aged by 40 and elderly by 60, give or take a 
few years depending on constitution and lifestyle. Infant mortality is high, but in most societies, if a child can survive his first six years, he is likely to live until a ripe age.


\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} None
\paragraph{\label{par:Willing to Learn}Willing to Learn}---\quad
Humans are flexible creatures by nature, and learning is something most can do very quickly, 
particularly when instructed. Many of the other races, having the mindset of immortality, or 
simply lacking this flexibility of mind, take to structured teaching less well, and require more 
practice to reach the same level of competence. This particularly shines in the arts of battle, which 
the Human mind adopts easy. When purchasing Schools, Human characters reduce the cost of 
each Proficiency Core added to the School by 1. 


\paragraph{\label{par:The Human Condition}The Human Condition}---\quad
Human beings live short, brutal lives by the standards of the elder races. Moreso than any other 
race, Humans are characterized by their ability to overcome their own weaknesses. Only Humans 
may benefit from the Flaw Arc. 
\subsubsection*{Culture and Society} 
There is no single unified Human culture. The diaspora of humanity across Mundus has created an 
enormous field of differing cultures and civilizations, which vary dramatically from each other. No blanket statement can be said about Human cultures, except that they predominantly consist of Humans.
\subsubsection*{Ecology} 
XXX 
\subsection{Goblins}
\begin{quotation}
	\emph{"Do you see those wretched creatures, slinking in the dark? Once upon a time, they too had hopes and dreams, and built great empires to challenge the gods. From mud and grime they came, to mud and grime we returned them. Take this lesson to heart. Human."}\par
	\hfill--\textbf{Petraclius Volkanius}, \textit{Helian Dux}
\end{quotation}
\begin{quotation} 
	\emph{"The best seasoning is hunger."}

	\hfill--\textbf{King Rudger the Reclaimer}, Goblin ruler to his young prince 
\end{quotation}
\subsubsection*{Overview} 
The Goblins are a crafty, slight, subterranean race, driven into the deep places of the world in eras long 
past by other races invading their homelands and defeating them in battle. Now the Goblins live brutish 
and violent lives in the deep places, struggling for survival against the terrors of the dark, and 
occasionally slipping away to the surface, to seek better lives among their historic enemies. 
\subsubsection*{Appearance} 
Goblins are physically slight, and have thin, downy hair on their heads. They can have alarmingly large 
eyes due to their environment, and can see in the dark with great precision. The average Goblin stands at 
3'5", and weighs 50 lbs. Their eyes are large and lack whites and irises, being entirely pupillary. Their 
hair is usually white or pale yellow, but is sometimes red or a ruddy brown. Their skin tends to range 
from a fishbelly-white to a darker green tinge. Exposure to the sun darkens them as it tans Humans, 
turning them a darker green. It is rumored that if Goblins are well-fed and long-lived, some may begin to 
grow horns. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} -2 STR, +1 AGI, +1 PER 
\paragraph{Regeneration}
Goblins can recover from any injury that is not fatal. Lost limbs, eyes, and so on, regrow after 
twice the Healing Time of the wound has elapsed. As such, Goblins cannot take Banes that they 
would heal from, like Severed Limb, at Character Creation. 
\paragraph{Small}
-2 Reach, -1 MOB 
\paragraph{Sneaky}
Goblins are naturally cautious and swift creatures. They gain +1 bonus dice to the Stealth Skill, 
and are always considered trained in that skill. 
\paragraph{Scarce Environment}
Goblins only have to eat half as much as other races in order to survive. 
\paragraph{Tight Spot}
Goblins are very short, but also highly flexible and double-jointed. They can fit through any 
space large enough for their small heads to pass through. 
\paragraph{See in Darkness}
Goblins can see in Pitch Black, Poorly Lit and Dimly Lit as though it were Evenly Lit. 
\paragraph{Photophobia}
When in Brightly Lit, Goblins count as being in Poorly Lit. When in Evenly Lit, Goblins count as 
being in Dimly Lit. Blinding is still Blinding. 
\subsubsection*{Culture and Society} 
The Goblins live in the deep places, in the vast volcanic tunnels beneath the surface. In these scarce, dark 
places the Goblins form tight-knit tribes and communities where survival is a daily struggle, and there is 
no room for waste or rivalry. 

 
The males live in the outer cordon of war-camps and outposts, protecting the inner villages from the 
predation of hostile subterranean beasts, while females and the young dwell in the inner lands, scraping a 
living from the stones, and crafting weapons, armor, food, and supplies for their beleaguered defenders. 
 

Goblin leadership tends to be split between the inner (female) and the outer (male) halves of each tribe, 
and traditionally only the greatest warriors of each generation are allowed to breed, producing new 
generations to waste their lives in defense of the nest. 
 

Goblins are suicidally courageous, but fear sunlight more than death. Only the bravest and most 
determined of raiders will venture above ground to seek supplies, loot, or even allies on the surface world. 
 

Goblin religion is simple and practical. They have no concept of gods, an afterlife (the thought is 
horrifying to them, as life is grotesque enough without the added dread of being unable to die), or 
divinity. They do understand sacredness, however, and hold sacred the idea of continuation through 
transformation of the body into other forms. Thus, even the beasts which consume the Goblins are, in a 
way, sacred, because they are composed of Goblins. Even dirt can be sacred as it may someday be a 
Goblin. 

 
However, taking the place of evil, hell, and dark gods in the Goblin religion are very real demons called 
'the Horned Ones,' depicted as terrible lanky figures with sharp protrusions jutting from their heads in 
artwork and myth. They are known to hunt and kill Goblins, however not as part of the natural cycle of 
death and rebirth, but instead as part of something profane, terminal, and unnatural. Some of the Horned 
Ones are infamous, and have elaborate mythos attached to them, with fanciful names like Sharphorror, 
Paindeath, Killsfast, Toothcutter, and Circlebreaker. 
\subsubsection*{Ecology} 
The Goblins have spent millennia struggling to survive in one of the most hostile and barren 
environments conceivable. Their diet largely consists of algae, roots, insects, and the flesh of beasts 
hideous beyond imagining. They have no particular aversion to cannibalism, and often ritually eat their 
own dead, both as celebration of their fallen comrades, and to take their nutrients back into the 
community. 

 
Due to this harsh diet, Goblins can digest truly foul substances and still glean nutrition from them. 
Famously, on the surface, they can actually sustain themselves largely on the nutrient-rich earth of 
farmland, simply shoveling it into their mouths as a Human would pottage. 
 

Goblins reach maturity within three years, but rarely live longer than 10, due to their extreme 
environment and consequent malnutrition. Legends state that a Goblin with ample food and free of the 
threat of predation can live as long, or longer, than Humans. 
\subsection{Dwarves}
\begin{quotation}
	\emph{"The stone has called us for an eternity,	It rings unceasingly behind the ears, 	It calls in a voice that can be ignored, 	But only just, and not forever."}

	\hfill--\textbf{Kordi Sticcor}, Dwarvish Engineer-Poet 
\end{quotation}
\begin{quotation}
	\emph{"Dwarvish social behavior is more consistent with that of carpenter ants than of men or Din. Their appearance is a trick. The smiling face on a cobra's hood."}

	\hfill--\textbf{Igone Zorezai}, Burdinadin Writer \& Researcher, \textit{‘Encyclopedia Subterrania, Dwarves \& Star-Nosed Moles’}, Library of Glade Lagu 
\end{quotation}
\subsubsection*{Overview} 
The Dwarves, or Fedarshin (as they are called by the Din), are a race of short, sturdily-built folk who live 
primarily underground in isolated strongholds typically called forts, or holds. The Dwarves do not fit in 
on Mundus. They have a compulsive urge to dig, to scrape away dirt and to strike the earth, and seek out 
something in the deep. They combat these urges with obligation. Dwarves value anything that can keep 
them focused on life, even if the call of the depths are never far from their mind. 

 
Dwarves also know, instinctively, when the moment of their death by age will be, down to the second. 


They do not know how, or where, but they know when. Many Dwarves, upon reaching an old age and 
knowing their time has come, will finally indulge in one final dig, heading down to die in the deep, as so 
many have done before. 
\subsubsection*{Appearance} 
Dwarves stand about 4'10" tall, with black or brown hair, often braided with metals and stone insignias 
woven in. Their compact, stocky frames weigh about as much as Humans who stand a foot or more taller. 
Females are almost exactly the same height, however lack beards and have softer (if still very sturdy) 
facial features. Dwarves can see in the dark. They are hardy, and very resistant to injury, toxin, and 
hardship, and are famous for their capacity for drink. They are also long-lived, although not immortal, and 
after reaching adulthood at about 20 years, many live to the very old age of 200. Interestingly, Dwarves 
never become decrepit, they simply persist in a sort of middle age until dropping dead. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} +2 END, +1 HLT 
\paragraph{Short and Stout:}
-1 Reach, -2 MOB 
\paragraph{Robust Immunities}---\quad
-1 to Infection Chance against all wounds, +2 to \nameref{sec:Blood Loss} rolls. 
\paragraph{See in Shadow}---\quad
When in low-light conditions, Dwarves treat each stage of Lighting as one step less severe on the 
table, because of their excellent eyesight in the dark (i.e. Pitch Black becomes Poorly Lit, Poorly 
Lit becomes Dimly Lit, and Dimly Lit becomes Evenly Lit). However, they cannot see colors 
when in the dark. 
\paragraph{Prodigious Livers}---\quad
Dwarves gain a +6 to their effective HLT when rolling to resist toxins (poisons, alcohol, general toxins) that are ingested, and a +2 to their effective HLT against poisoned arrows, weapons or other toxins that enter the body through injury.
\paragraph{The Call of the Deep}---\quad
Dwarves get +4 to any roll involving digging downwards. A Dwarf must make a WIL roll at RS 5 
to stop himself from digging once he has begun. A Dwarf who is currently under the influence of 
alcohol reduces the RS to 1. 
\paragraph{Hammer of Industry}---\quad
Dwarves always get one extra die in each step of crafting the roll in. (Not in Beta) 
\subsubsection*{Culture and Society}
Dwarvish culture varies greatly depending on the Hold in question. However, most Dwarvish 
communities are close-knit, with individual clans within the Hold each having a hereditary occupation, 
jealously guarding their secrets from others while endlessly perfecting their techniques. Contact with 
outsiders is greatly restricted, and trade with the general public is often divided into two groups: the 
‘outer’ market, where strangers and poor traders deal at inflated prices, and the ‘inner’ market, where 
trusted merchants are allowed to see the real stock without price-gouging. 


It is not uncommon for three generations of merchants to deal with a clan of Dwarves before being 
granted trust (usually in the form of a tattoo by the Hold's inkweaver that grants access to the inner 
bazaar). Dwarven goods are of obscenely high quality, but increasingly, their exquisite craftsmanship has 
begun to meet competition, both from the Burdinadin's precision engineering, and the raw mass 
production power of Human production houses. 

 
The Dwarvish system of writing, which is written as dots in three columns going downwards, was born of 
their primary means of identification in the dark tunnels, where facial features are hard to discern. Thick 
beards (or necklaces, for women) with elaborate braids studded with beads, clever knots and medals, 
allowed individuals to recognize each other by ‘reading’ them. Many Dwarves of status take great pride 
in their elaborate, dignified beards, while those who leave the Holds typically cease the practice, after 
realizing that nobody can read their beards except Dwarves, or else increase the practice and flatter 
themselves by braiding their beards ostentatiously, perhaps for that very same reason. 

 
Internally, each Hold is ruled by an anonymous King, who sits upon a stone throne in the deepest heart of 
the Hold, and wears the Hold's ancestral mask. The King chooses his own successor through merit, and 
the death of the King is never announced — officially, most Dwarf Holds have only ever had one King. 
Instead, each new King is called a ‘Mood,’ such that King Ulric may have had the Mood of Economic 
Reform, the Mood of Innovation, the Mood of the Rat War, and the Mood of Decadent Feasting over the 
course of several centuries. 

 
Dwarves have a deep, overriding urge to dig. Those who tear themselves away from the underground 
eventually lose this urge, but for most Dwarves it is a constant presence in their minds. If they begin to 
dig, Dwarves find that it is difficult to stop, and if left to their own devices, most would simply dig until 
they died of thirst in the deep, hammering ever downwards in an ecstasy of picking. 


The Dwarves know this as ‘The Call,’ and they consider it a natural thing. Most of them combat it with 
alcohol, as a drunk Dwarf becomes temporarily deaf to The Call. A common strategy for ambitious 
miners is to carry two canteens, one with water, and one with strong turnip-wine or dark ale. After the 
water runs out, the Dwarf will quickly become intoxicated enough to realize that he's out of supplies, and 
return to the surface. 

 
Of course, this is a dangerous technique, given Dwarven alcohol resistance, and also given the 
carelessness of drunken miners, so the foreman and overseers of Dwarvish dig operations usually drink 
enough to keep the edge off at all times, so that they can spot anyone who has spilled his drink, or 
watered down their ale too much to pull themselves away from the pick. 
\subsubsection*{Ecology} 
The Dwarvish diet is very similar to that of Humans, with a few exceptions. Dwarves can ingest 
extremely toxic substances without fear, and their diet includes many things that would be outright 
poisonous to Humans or Din. Dwarves are also highly resistant to the effects of alcohol, and have had to 
develop singularly strong mixes to achieve the same effect as Humans get from softer drinks. Because the 
Dwarvish eye is much more optimized for tracking the outline of shapes than it is for discerning fine 
details within those shapes, Dwarves have trouble identifying people by their facial features. 

 
Dwarves mate for life, and rarely seek a second mate if separated from the first by death or misfortune. 
\section{Din}%NEEDS additional work
The Din, or Elves, are a category of humanoids characterized by their possession of ‘The Focus.’ The 
Focus is a sort of practice, aversion, taboo or belief that grants the Din immortality so long as it is held to, 
as well as certain other powers or abilities, based on the sub-race of Din. 


The Din themselves often quarrel over what precisely constitutes a Din. The Ohanedin, for example, maintain that Humans are also Din (they call them the Hostoadin) because of their ability to Ascend, which they consider a form of Focus. The Burdinadin, by contrast, do not believe that the Zells or the Ohanedin are Din at all, because both receive their immortality from higher powers, rather than from within. The Orredin believe that the Goblins are a degenerate and fallen form of Din that has forgotten its Focus (and thus, justify by precedent that the Orredin themselves are an ascended superior race, while the rest are merely average). 


Whatever the truth of the matter, for the purposes of Races, any race with a Focus can be considered Din, and remain ageless so long as they practice their Focus. They are also susceptible to any effects, magic, and so on that specifies an effect against Din.


In terms of appearance, the Din as a whole are generally similar to Humans, but are usually marked by slightly tapered ears (though this varies between the modest and delicate ears of the Orredin, and the grand, some would say obnoxiously large ears of Zells). The features of Humans (for those Din who hold that the Humans can be counted among them) are simply a logical continuation down from the Orredin.


\subsubsection{Focus}
Din are ageless beings and hold a kind of immortality as long as their Focus is maintained. Each Din has a different Focus, which provides them with certain attributes and skills.


\paragraph*{Zells}---\quad
The Zellish Focus is called ‘The Dream,’ which grants them immortality so long as they are part of The Dream. A seafaring race, the Zells need to be members of a crew to be part of The Dream. Zells have a strong bond with one another, and have telepathic links with their crew members.

\paragraph*{Burdinadin}---\quad
The Focus for Burdinadin is rather unclear. The ‘Iron Elves,’ as they are also called, only know that while they are within the sanctuary of their Iron Glade, or otherwise shielded when venturing outside the Glade, they are protected from spiritual contamination, and maintain their Focus.

\paragraph*{Ohanedin}---\quad
The Ohanedin are a feared, yet honorable race. Their Focus has three main principles, which ensures their immortality and agelessness. First, they must never clothe themselves in metal, they must never betray an oath freely given, and never eat meat from animals that cannot speak.

\paragraph*{Orredin}---\quad
[Not Released] %BLATANT LIES, can we fix this somehow?

\paragraph*{Paleolithic Elf}---\quad
[Not Released]

\subsection{Zells}
\begin{quote}
	\centering
		\emph{“Quatschte blöde an die Zelle, kriegste schnell ne Maulschelle”}\\
		\emph{<<\emph{If you talk tough to a Zell, you should expect your face to swell}>>}\par
\hfill --Sessonian Folk Saying (Rhyming Translation)
\end{quote}

\begin{quote}
	\centering
		\emph{“My race sailed these waters before yours crawled out of them.”}\par
		\hfill --\textbf{Rade Sark}, Admiral-Prince of the Sea of Eyes
\end{quote}
 
\subsubsection*{Overview} 
The Zells, or Zellish Elves, or Sugaardin, are a race of humanoids who favor a maritime existence, and are telepathically bonded to their crews, as well as to the living ships on which they sail, called Zellislava. They are immortal, so long as they are on the sea and part of a crew, and so most Zells avoid land when possible. Zells also have a strange attunement to the supernatural, drawing the attention of powerful spirits and even deities more easily than other races. 

The Zells have ruled the seas of Mundus for countless eons, and can be found in every maritime role, from mercenary to merchant. Though at heart they are not as warlike a people as Humans are, the Zells have proven that on the high seas they have no equals, and certain groups of them are highly militant.

\subsubsection*{Appearance} 
Zells resemble Humans, but it would be difficult to mistake the two. Their ears are long and tapered, and grow longer with age; truly ancient Zells having ears in excess of a foot in length. 
The average Zell of both genders stands at about 5'10". Males tend to weigh about 150 lbs, and females slightly less.


The ‘classic Zell,’ the stock from which most of them derive, is fair skinned, with straight black hair, black eyes, and an aquiline nose. However, most Zells have some Human ancestry, and so show diverse traits much as Humans do. Zellish men can grow full beards, something that most Din cannot do, and some of the Zellish ethnic groups place great importance in the growing of mustaches and beards.


\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} +1 AGI, +2 PER 

\paragraph{Elvish Agelessness}---\quad
Elves reach physical maturity by the age of 20, much like Humans do. However, after that, their 
bodies' aging process halts, so long as they retain their Focus. Each breed of Din has a different 
Focus. 


\paragraph{Focus}---\quad
Like all members of the Din, the Zells have a focus that allows them to remain Ageless. The focus for Zells is ‘The Dream.’ As long as a Zell is a member of a Dream, the Zell does not age significantly past physical maturity. This effect does not stop the Zell's ears from growing longer over time. Zells who leave their Dream begin to age normally. 
Zells lose Focus when they set foot on land, but they can stave off the effects of aging by drinking salt walter from the sea. As long as they have a steady supply of sea water and are still part of a Dream, they can retain their immortality indefinitely.

\subparagraph{Zellish Dream}---\quad
The Dream is a sort of psychic connection that occurs between Zells. It is a form of semi-voluntary communal bonding, the trigger for which is a kind of vessel on open water. When it takes root, the Zells begin to share thoughts, become aware of each others' well-being and feelings, and perhaps more importantly, begin to hear the thoughts of the ship itself, which is naturally drawn into the bond. The Dream allows Zells to communicate nonverbally over a distance of about a mile, and to communicate instructions to their ship (assuming it is a Zellislava). Being a member of a Dream is the requirement for their Focus.


Becoming part of a Dream requires that two or more Zells spend about a week as the sole occupants of a vessel or ship. Even a rowboat will do. Importantly, these Zells must at least agree on which of The Seven Gods of Chaos has primacy. They do not need to be of the same tribe or hold the same beliefs. Even Zells who were raised on land with no contact with other Zells can form crews by fulfilling these requirements. Additionally, Zells that are not part of a Dream may be incorporated into a Dream by spending a similar amount of time on the ship with its crew.


\paragraph{Zellislava}---\quad
Any ship that serves as the ‘founding’ vessel for a Zellish Dream (i.e. the rowboat that Zells spend a week on to merge their consciousnesses) becomes a Zellislava immediately.
Alternatively, if a pre-established Dream of Zells make up the primary crew of a vessel of any size for about a year, that ship will gradually become a Zellislava over that period. The Zellislava itself will be aligned to the member of The Seven Gods of Chaos that the Zellish crew worship. A Zellislava is a living ship that gradually grows and expands from the framework of its original body, increasing evermore in sophistication and complexity, and eventually gaining moving components.

\paragraph{Friends in Deep Places}---\quad
Zells in open water, either at sea, on inland oceans or large lakes, have the peculiar ability to 
vanish into the depths, and reappear at the nearest shore. This involves them emptying their lungs, 
losing consciousness, and sinking into the deep. Through a barely understood force (believed by 
the Zells to be the machinations of their gods) the Zell will emerge anywhere from a few hours to 
a month later on the nearest coast, feeling reasonably well-rested, but hungry and sore. 
Everything on their person, including objects held in all but the most airtight containers, will be thoroughly soaked. 

Additionally, anything that a Zell personally drops into the ocean, intentionally or 
unintentionally, can be recovered during the Zell's time in the deep, but this requires a WIL roll 
with an RS of at least 3, and possibly more if the item is very heavy. Objects too large or heavy to 
be carried by one person cannot be recovered. 

If a Zell is somehow prevented from sinking while attempting to use this ability, they do not 
vanish and reappear, but they also do not reawaken until either released into the deep, at which 
point the ability takes effect as normal, or until they are pulled up to the surface, at which time 
they awaken, confused and disoriented. Regardless of the status of their Focus, a Zell does not 
age while in this state, and fish and other animals will not eat or even touch a Zell who is in this 
sort of stasis. Theoretically a Zell could spend centuries submerged in this manner… 

\paragraph{The Zellish Touch}---\quad
Zells always add their PER to Missile rolls involving any weapon with circular components. If 
they use the Aim Maneuver, they may add their PER again. Weapons that ‘count’ for this rule 
include Zellish wheelbows, wheellock firearms, throwing disks like frisbees, circular sawblades, 
chakrams, and so on. 

\paragraph{Echolocation}---\quad
A Zell's ears are much more sensitive than that of a Human, and because they do not speak often 
in their native environment, they are less likely to filter out echoes that would otherwise prove 
highly confusing in regular speech. The result is that they have the ability to echolocate, that is, 
by making small noises that reverberate off of far surfaces, they have a sense of where objects 
are, even in complete darkness. 

So long as conditions are relatively quiet, a Zell can treat even Pitch Dark Lighting as Poorly Lit 
for the purposes of movement, and even shooting. If the target of such a missile attack has made a 
loud or audible noise (even a bow or crossbow shooting is sufficient) then this improves to Dimly 
Lit. Poorly Lit reduces MP by 75\%, Dimply Lit reduces MP by 50\%. 

The range of this ability is generally 10 yards per point of PER the Zell has. 

\paragraph{Racial Glossophobia}---\quad
Zells have very keen senses of hearing, and this actually proves a disadvantage when 
communicating verbally, because the echoes of both their own and other voices makes it very 
difficult to pick out small details like words and inflections. As such, Zells often seem awkward, 
are prone to pausing in conversation to ‘sift’ through the words they have just heard, and also 
tend to prefer talking either very softly, or very loudly, with little room in between. Zells suffer a 
-4 pool penalty to CHA related tests involving communicating with others verbally. This does not 
apply for Intimidation checks. 

\paragraph{Race of Sailors}---\quad
Zells gain a +2 racial pool bonus to Sailing tests. This bonus increases to +4 if the boat being 
operated is a Zellislava. This bonus decreases to -2 if the boat being operated is a Zellislava of a 
Chaos God hostile to the Zell in question. 

\paragraph{Resistance to Disease}---\quad
Zells cannot contract some diseases, including any transmitted by insects or rats, and some 
vitamin deficiencies, like scurvy. They are also curiously immune to botulism, though it does 
upset their stomachs. 

\paragraph{Strange Tastes}---\quad
Zells suffer a -2 penalty to PER to detect poison (or anything else) in food or drink, because they 
have a dulled sense of taste. However, they may survive drinking seawater, and cannot suffer 
nausea from smells or tastes, no matter how vile. 

\subsubsection*{Culture and Society} 
The Zells as a race do not have a well-organized society. They live primarily on the sea in independent 
ships, each with their own small community and goals. These ships belong to tribes, of which there are 
42, but these tribes do not have any sort of permanent leadership above the individual ship level. Instead, 
tribes have a ‘creed,’ which is decided upon every hundred years in a great meeting, in which all ships of 
the tribe gather together (if able), and hold talks about any possible changes to their creed. If a captain or 
his crew breaks the creed, and it is discovered, punishment is usually meted out quickly, according to the 
creed, by the rest of the tribe. 

The actual culture and values of the Zells vary dramatically along tribal lines, but a constant is the 
worship of The Seven, who are believed to be the source of The Dream, and who are venerated almost 
universally by seabound Zells. Each tribe has a ‘primarch,’ or a member of The Seven whom they believe 
is first among equals, and whom they believe to share a special relationship. Another constant in Zellish 
culture are wheels. Particularly wheels that spin or turn seem to hold great significance to Zells, even if 
they cannot articulate why exactly they do. Their religious services always involve wheels or rings of 
some sort, and The Seven are even portrayed as being spokes on a great wheel, in some tribes. 

Even weaponry is often made involving rings or wheels: throwing disks with razor edges, bows with the strings 
threaded through multiple wheels, and the occasional sword with rings attached to the spine. Thus far, the 
Zells haven't figured out how to make a sword that looks like a ring, preferring to use their Zellish Sabers, although it's only a matter of time. 

Many Zells have simply been assimilated into Human culture (usually because their ancestors couldn't 
afford a ship). These people are not generally much different from regular Humans, though they tend to 
be taciturn, as the natural aversion to words stays with them. The fixation with circles also seems to be 
somewhat deeper than merely cultural, as Zells, regardless of their background or upbringing, tend to 
favor rings, wheels, and other circular designs in art and craft. It is unclear if this fixation with rings is 
due to the Seven's association with the ‘Brass Wheel,’ or if The Seven's association with rings is due to 
the Zells' worship of them. 

Zells are stereotyped by humans as being highly promiscuous and uninhibited. This is mostly because the 
ones humans meet tend to be sailors getting drunk before heading back out to sea, but it is generally true 
that Zells think less of romance and relationships than humans do. Children are usually kept by (or left 
with) the mother. This is considered the normal state of affairs among Zells--it is very unusual for a Zell 
to actually know his father’s name, those who do will often adopt it and the suffix “-vic” as their surname. 

So a Zell named Boris whose father’s name was Mark would be called Boris Markovic. Having such a 
name is as close to “nobility” as it gets in Zellish society. 

\subsubsection*{Ecology} 
Zells are social animals. In their natural state, they exist in tight-knit groups that live, travel, fight, eat, and 
generally act together as one. The Zellish diet does not differ much from that of Humankind, though their 
sense of taste is far less acute, and Zells are poor at discerning tastes unless they are very strong. This has 
led to most of their food being extremely pungent. They also drink (mostly) salt water, which would be 
fatal for most other races. Zells treasure the taste of fresh water, and believe it helps with the digestion. 

Most Elves are resistant to disease, but Zells in particular have adapted immunities to a few conditions 
that are commonly the bane of sailors. They cannot contract scurvy, and their skin, blood, and bodily oils 
are highly toxic to insects of all sorts, making infections from mosquitoes, lice or fleas unlikely. 

Zells do not mate for life, and in fact have great difficulty feeling attraction for those that inhabit a Dream 
with them. They are prone to short relationships in passing, either with their own kind, other Din, or 
Humans, and the offspring are almost always Zells, regardless of the mate. 

\subsubsection{How to Roleplay a Zell}
When playing a Zell, remember that while you’re probably a sailor (unless you’re a land-born Zell) 
you’re not a human sailor. Zells are taciturn, they don’t like talking too much, they dislike loud noises 
like shouting, and they prefer to show rather than tell. Aside from that, they can be very human, they 
enjoy the same things more or less, but they tend to take a longer view on things. Zells love the sea, they 
love contemplation, solitude, and the company of close friends. For a Zell, the only truly close friends are 
those in the Dream. Friends outside of this psychic link--especially humans--may be liked, even loved, 
but they will never truly be “in” like a Zell’s crewmates are.

As a Zell, you’re a hard worker, a brave seaman, and a true comrade. You don’t weigh in your opinion unless you’re sure of what you’re saying, 
and generally you’d rather be sure of what’s going on before any decision is made. You might feel that 
you are superior to humans because of your immortality, (and also because their ears remind you of 
children) but Zells and humans have been living together for a long time, and there’s also a certain 
brotherhood there--you’re probably part human yourself, after all. 

\subparagraph{Male Names:} Vlad, Rade, Boris, Akula, Olaf, Taviye, Dusan, Slavomir, Bozemir, Radovan, Arkan
\subparagraph{Female Names:} Kasatka, Dahaka, Shahaka, Moloko, Rahoko, Draguna, Natalya, Tamara, Dusica

\subsection{Burdinadin} %BURDS NEED WORK
\begin{quote}
\emph{``I've been inside of an Iron Glade. The construction is admirable, though lacking in aesthetics. They bathe in rooms full of steam.\\
Their halls are lit by panels of glowing glass, and the great towers stretch as far underground as above, but there is no personal element.\\
It was not a home, it was an emergency shelter. A bolthole. A temporary solution become terribly permanent.''}

\vspace*{-5pt}\hfill--\textbf{Zoph Escher}, Lord of Kar-Ischil \textit{‘On Elves’}, Report to The Hurtsickle Society
\end{quote}
\begin{quote}
\emph{``If we thought death was a way out, we would have built gibbets, not Glades.''}

\hfill --\textbf{Skillithingtorix Amljatijakove}, Burdinadin Radical, \textit{‘No Way Out’}, Master Thesis, Lagu University of Natural Philosophy
\end{quote}

\subsubsection*{Overview}
The Burdinadin, or ‘Iron Elves,’ are a race of intellectuals and shut-ins who have sealed themselves away 
from the outer world within the great Iron Glades, huge metallic fortresses that shield them from the 
impurities of the outside. They are gifted scientists, in large part because the same quasi-magical senses 
that made the outside world so intimidating to them can also be turned inwards, to analyze the subtle 
mechanics of natural law around them. 

\subsubsection*{Appearance} 
The Burdinadin range from chestnut-skinned to fair, and tend to have pale hair and green or red eyes. 
They rarely stand over 5'6", and tend to hover between 100 and 140 lbs. Females are not noticeably 
shorter than males, but tend to be slighter of figure. 

\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} +1 INT, +1 AGI 

\paragraph{Elvish Agelessness}---\quad
Elves reach physical maturity by the age of 20, much like Humans do. However, after that, their 
bodies' aging process halts, so long as they retain their Focus. Each breed of Din has a different Focus. 

\paragraph{Focus}---\quad
The Burdinadin do not currently know what their Focus is, but so long as they are within the Iron 
Glades or properly shielded while outside, they retain their Elvish Agelessness. 

\paragraph{Natural Philosophy}---\quad
So long as the Burdinadin is not suffering Pain from spirit contamination (see Sterile Soul below) 
a Burdinadin may make a special Education test to discern the function, purpose, composition, 
and/or mechanics of any object, device or technology. For this test, the Burdinadin adds PER to 
INT. This kind of check is omni-sensory, utilizing sight, smell, touch, hearing, and even taste.

If the Burdinadin is allowed to disassemble and play with the subject manually, a +4 bonus is gained to the Education test. %EDUCATION TEST?
When used in this way, Natural Philosophy requires time to study the object determined by the RS of the test. See \autoref{tab:Natural Philosophy} below.

After a successful use of this ability to analyze something, any \nameref{skill:Crafting}~rolls made to improve, redesign, copy, forge, disable or otherwise take advantage of this knowledge gain a bonus equal 
to the Bonus Successes from the Natural Philosophy roll. The same object can be analyzed 
multiple times, but the Crafting bonus does not stack, only the highest number of BS is used. 

There is a cost to using this power. Every hour that is spent using Natural Philosophy, the 
Burdinadin must make a WIL test at RS 1, +1 per hour they have already been using the ability. If 
they fail, they attract the attention of That Which Stares Back, and begin suffering spirit 
contamination, as when the Burdinadin is outside and unshielded. Once this begins, the 
Burdinadin cannot break free of the trance, and thus the spirit contamination, until they make a 
WIL test at RS 1, +1 per hour since That Which Stares Back has fixed its gaze upon them. 

\begin{table}[hb]
	\centering
	\caption{Natural Philosophy}
	\label{tab:Natural Philosophy}
	\begin{tabu} spread 0pt {X[-2] X X}
\rowfont[c]{}Test RS &Time Required &Example Subject\\\toprule
1 &Less than a second &Mace \\
2 &10 Seconds &Lever \\
3 &1 minute &Mousetrap \\
4 &10 minutes &Crossbow \\
5 &1 hour &Matchlock \\
6 &12 hours &Pocket Watch \\
7 &1 day &Steam Engine \\
8 &1 week &1960s Supercomputer \\
9 &6 weeks &Nuclear Submarine \\
10+ &x10 per RS in months &Alcubierre Drive \\
	\end{tabu}
\end{table}

\paragraph{The Stains of Time}---\quad
The Burdinadin can see trails of causality through the air before them when in tense situations, 
much as Humans see strange patterns on the insides of their eyelids after looking at bright lights. 
Unlike the Human visions, which are thoroughly ignorable traces of dying cells on the cornea, the 
Burdinadin's visions are glimpses of the future, brought forth by their ability to gaze into the 
interior mechanics of the universe's functions. 

So long as a Burdinadin is not suffering from spirit contamination (see Sterile Soul below) and is not suffering more than Light Encumbrance, the TN of all Void and Parry maneuvers is reduced by 1. 

Additionally, The Stains of Time allows Burdinadin to Void or Parry missile attacks from firearms. 

\paragraph{Sterile Soul}---\quad
The Burdinadin cannot handle contact with the outside world, and all of its spirit contamination, 
without proper shielding. If a Burdinadin is ever forced to leave the Iron Glade without a 
protective suit or other means of protection, he will suffer 1 Pain per hour from spirit 
contamination. If the Pain suffered exceeds the Burdinadin's combined HLT and WIL score, he 
must pass an RS 3 WIL check each hour or die, as his nervous system is systematically 
dismantled by aggressive spirits. 

If the Burdinadin finds shelter in a sterile location, or is moved to a safe place after suffering Pain 
in this way, the Pain will fade at a rate of 1 Pain per 1 week spent in the sterile location.

For faster contamination cleansing methods, and for protective suits and items, see \autoref{ch:Burdinadin Armory}

\paragraph{Burdinadin Armory}---\quad
Special access to Burdinadin equipment. More information can be found in \autoref{ch:Burdinadin Armory}.


\subsubsection*{Culture and Society} 
The Burdinadin live almost exclusively in the Iron Glades, which were built to shield them from the 
spiritual contaminations of the outside world. Having learned to retain their sanity in such claustrophobic 
environments, the Burdinadin combine gregariousness and privacy, and greatly value their personal 
space, often retreating into seclusion for weeks or months at a time to ponder over problems, or come to 
terms with personal dilemmas. Their society recognizes and respects this withdrawal, but expects 
conformity and reservation in public. The result is a highly dignified society, with extreme standards of 
decency, conduct, and morality in public, but provides that anything is permissible in private. 

The Burdinadin cannot perform meaningful agriculture inside the Iron Glades, and so they trade for food 
and materials with the outside world, often trading back cheap steel, mass-produced goods, and wondrous 
inventions for bored kings. 

Burdinadin are known for their introversion and being slow to befriend and slower still to trust. Romantic 
relationships are rare amongst them (too much time, too much money, more interesting things to do--like 
researching beetle phenotypes!) to the point that many Glades institute a severe bachelor’s tax to extort 
their citizens into producing offspring. Despite that, once they’ve got them, Burdinadin are very dedicated 
parents. 

\subsubsection*{Ecology} 
The Burdinadin have relatively tame dietary quirks compared to Humans. They do have difficulties 
digesting milk (having almost no evolutionary history of domesticated cattle), and are especially 
susceptible to food poisoning from meat near spoiling. Most Burdinadin eat a largely vegetarian diet, with 
meat being eaten fresh, or not at all. 

Burdinadin don’t usually mate for life, though many have reoccuring relationships with the same individual over a long period of time. Pregnancies last about 12 months. They can breed with Humans and other Din, though the offspring are generally of the mother's race, except for Zells, with whom they almost invariably produce Zells. 
\paragraph{How to Roleplay a Burdinadin}
When playing a Burdinadin, remember that you grew up in a highly ordered and enclosed environment, where your primary mission in life was maintaining that environment. You’ve been conditioned since you 
were a child to have a keen attention to detail, to live an orderly and methodical life, and to pursue your work--whatever it may be, from soldiery to research--with unparalleled dedication and ambition. If your job is maintaining the air-circulation units of your Glade, then your every waking moment will be spent furiously struggling to maintain, improve and perfect that unit. If your job is killing people, then you will spend every waking moment of your life straining to be as good of a killer as you can be. 

As a Burdinadin, you’re smart, fast-talking, fact-obsessed, and a bit naive. Metaphors and analogies tend to go over your head, you take things literally and you present things literally. You’re not opposed to ideas about magic, gods, souls or whatever--after all, what is the Void if not a realm beyond this one? But you’d rather explain things with hard data than speculation and philosophy. As a Burdinadin, you might be a bit awkward around people in a non-work setting. You have hobbies that you know inside and out, but what’re the odds anyone you’re going to meet is as into whittling as you are? Not very. 

You might not see humans, or even other Din, as being equal to you. They’re irrational, illogical, they don’t care about facts the way Burdinadin do, they get too wrapped up in emotionalism and nonsense. Maybe you think of them like animals just running off of instinct, or maybe you admire them for their ability to let go of details and “go with the flow.” 
\subsection{Ohanedin}%OHANEDIN
	\begin{quote}
		\emph{"The Ohanedin hate bows, did you know that? We all think of those feisty wood-elves with their bows, but they hate 'em. It's Zells what like bows. The stick-eaters don't like bows, 'cause you've got to be far away to use 'em. Can't enjoy it that way."}

		\hfill--\textbf{Sir Roger Westridge}, Knight of the Green Brotherhood
	\end{quote}
\begin{quote}
\emph{``This isn't your land, infidel."} 

\hfill--\textbf{Zoga Ganix}, Ohanedin Ranger, \textit{‘The Slaughter at Occi’} (Attributed) 
\end{quote}
\subsubsection*{Overview} 
The Ohanedin are an Elvish race of fiercely individualistic and independent tribesmen from the northern 
hills, cliffs and forests of Iber. They treasure their freedom and their way of life, and fight ruthlessly and 
without restraint to defend their ancestral homes. They are rightly feared by all who neighbor them, and 
are often (rightly) accused of the cannibalism, Human sacrifice, and depraved tortures that befall those 
who trespass into their lands without license. 

Conversely, they are also known for their nobility, iron-bound senses of honor, and undying loyalty to those who win their friendship, and many leave their hilly homelands to seek employment in the retinues of noble families. Some Ohanedin have acted as retainers, bodyguards and teachers for ten generations of the same noble Human family. Their word is their bond — more literally than any Human could hope to understand. 
\subsubsection*{Appearance}
Ohanedin are usually fair-skinned, with eyes ranging from brown to bright blue, and black, brown, or 
greenish-blonde hair. They are taller than their Burdinadin cousins, standing between 5'8" and 6'0" on 
average, and weighing about 140 lbs. Females are not much smaller than males, but the difference is 
noticeable. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} +1 STR, +2 AGI, +1 PER 
\paragraph{Elvish Agelessness}---\quad
Elves reach physical maturity by the age of 20, much like Humans do. However, after that, their bodies' aging process halts, so long as they retain their Focus. Each breed of Din has a different Focus. 
\paragraph{Focus}---\quad
To preserve their immortality, Ohanedin must obey the mandates of their patron spirits. Never clothing themselves in metal (metal weapons are acceptable, as long as the Ohanedin doesn't actually have to touch the metal part), never betraying an oath freely given, and never eating meat from animals that cannot speak are the three basic prerequisites, however many have significantly more. Due to the complexity of the Ohanedin Focus, it is broken down into its individual qualities below. 

An Ohanedin that breaks the laws of consumption (eating the flesh of an animal that cannot speak) will violently regurgitate its meal, and will lose access to Natural Awareness and Rules of Nature for the next 24 hours.
\paragraph{Metallophobia}
Ohanedin have a particularly rigid Focus, particularly when it comes to metal. Wearing metal 
covering the torso inflicts 8 Pain on an Ohanedin for as long as it is worn. Wearing it on the head 
inflicts 10, whereas the limbs only inflict 4 each. Wielding a weapon made of metal inflicts no 
pain, unless that metal is part of the grip that must be touched. A wood grip can insulate an 
Ohanedin from this effect, but cloth, leather or similar materials do not. 

Of course, the Ohanedin also age so long as metal is in contact with their flesh, as its Focus is 
being denied, and the Ohanedin cannot use his Natural Awareness or Rules of Nature powers 
either.

Injuries inflicted by metal weapons do not inflict additional pain, but any missile made of metal that has a Stuck chance inflicts 2 additional Pain on a successful Wound if the missile becomes Stuck. 
\paragraph{Focus Oaths}---\quad
An Ohanedin who gives his word freely is compelled to keep it, until such a time as he fulfils his 
promise, or the compact is broken by the other party. An Ohanedin who breaks his oath loses 
Focus for a year and a day, and is no longer immortal, and does not benefit from Natural 
Awareness or Rules of Nature. The Ohanedin does retain access to his Murderous Speed ability, 
and is also still vulnerable to Metallophobia. This period can be cut short if the individual to 
whom he made the broken oath names a penance for him that is deemed acceptable by the spirits 
(this is a fine art, and the rules are not clear to anyone, even the highest wisemen of the 
Ohanedin) which the guilty party then fulfils. A common penance is for the Ohanedin to cut off 
one of their own fingers, another is to serve the offended party for ten years. 
\paragraph{Natural Awareness}
An Ohanedin is automatically aware of any wounded or frightened life forms within 100 yards of 
it in any natural (non-urban) environment. This awareness is accurate to an area of about 10 
yards, at which point it cannot pinpoint it any further. This ability does not work on other 
Ohanedin, or creatures with the Soulless descriptor. Ohanedin lose this ability if their Focus is lost. 
\paragraph{Murderous Speed}
Ohanedin have nearly perfect control of their bodies' movements. There are no nervous tics, no 
‘tells,’ and no involuntary flinches. They move like spiders, from perfect stillness to rapid 
deliberate motion with no transition in between. 

In the Orientation Declaration stage of Combat, Ohanedin do not have to declare which Orientation they are not throwing.

Additionally, Ohanedin gain a +2 racial bonus to any Stealth related roll. 
\paragraph{The Feeling Disappears}
The Ohanedin's spirit senses, much like their Burdinadin cousins, offer them an unusual insight 
into the world around them and an ability to sense events before they occur. Unlike the Burdinadin, however, the Ohanedin are not fettered by their need to avoid spirit contamination. 
To the contrary, it is through the spirits that this power is achieved. 

As long as an Ohanedin currently has Focus, the TN of all Void and Parry maneuvers is reduced by 1.

Additionally, The Feeling Disappears allows an Ohanedin to Void or Parry missile attacks from firearms. 
\paragraph{Rules of Nature}
The Ohanedin have a delicate relationship with their patron spirits, and it is possible to offer up 
their own life-force for consumption in return for victory. This often results in blood forcing its 
way out of the eyes, ears, mouth and other orifices, but the sheer power granted in return is worth 
any sacrifice. An Ohanedin may choose, when declaring an attack that inflicts damage, to add 
their WIL score to the amount of damage dealt. If they do so, they suffer 1d10x2 Bleed as the 
attack resolves, whether or not the attack is successful at hitting and inflicting damage. 
\subsubsection*{Culture and Society} 
The Ohanedin have one rule as a society. That rule is that no Din can be ruled except by consent. Their 
analogue to feudalism is a system of voluntary slavery, but it is not hereditary. Instead, individuals who 
excel are offered land and status in return for years of their life in service to powerful lords, during which 
time they gain experience, wealth, and connections. 


As these arrangements are sealed by oaths to the spirits, which an Ohanedin cannot break without 
sacrificing his Focus, the arrangement is highly functional, as the power structures are simply incapable 
of betraying and plotting against themselves. This results in extremely bloody battles, where the Zaldunak 
(knights, or cavaliers) can be ordered to attack by their lords, and are unable to retreat until ordered to, or 
until their masters are killed. Ohanedin culture holds a certain reverence for these sort of oaths, and it is 
commonly believed that one can become a living god by holding to an oath long enough, and 
unconditionally enough. The Ohanedin religion venerates several of these entities, some of whom are 
very much alive and active, still acting to fulfil their oaths. 


The Ohanedin have a very sophisticated and ironclad family unit that is the basis of their society. Men are 
less numerous than women in Ohanedin society (mostly because of violence) and as such, polygamy is 
considered normal among the upper classes, though the average Ohanedin will live a monogamous life. 
The Ohanedin take this sort of thing very seriously, and are utterly dedicated to their spouses. The oaths 
of marriage are the second strongest out of all of the many oaths, surpassed only by the 


\subsubsection*{Ecology} 
The Ohanedin cannot consume the flesh of beasts that cannot speak; this is not just a religious conviction, 
but a biological one imposed upon them by the spirits, and they become violently ill if they consume animal flesh. 


As such, they have complex dietary requirements, which they must supplement (when possible) with meat 
that they are allowed to eat… generally Human, or that of slain rivals or fallen friends of their own race. 
Certain mimicking birds are considered fair game as well, though these are generally scrawny and make 
for poor meals. 


Blood, however, is considered perfectly fair game wherever it comes from. Thus, the most important component of the Ohanedin diet is a mixture of milk and cow blood, which they carefully extract from 
their cows without killing them. 


Ohanedin pregnancies are similar to those of the Burdinadin, lasting 12 months and being relatively 
painless. Most interracial pairings follow the race of the mother, with the exception of Zells, who always 
produce Zells. 
\subsubsection{How to Roleplay an Ohanedin}
When roleplaying an Ohanedin, remember that they are a very martial, very honest people. The Ohanedin 
may dissemble, they may even outright lie, but they try never to break agreements or promises, no matter 
how minor. For them, everything is about honor, obligation, and duty. Ohanedin don’t commit to things 
easily, since once committed to, breaking off an engagement is difficult indeed. As an Ohanedin, be 
straightforward, but don’t commit to things unless you mean it. Be skeptical of everything, don’t be lured 
into agreeing to something until you have all the facts. 

Be just to your comrades, and absolutely merciless to your enemies. Remember to praise the spirits--they’re your friends after all--and be respectful to those 
who show you the same courtesy. You probably distrust humans. Humans are bad at keeping promises, 
they make six of them every minute and they break them just as quickly. That doesn’t mean you hate 
them, but they’re compulsive oathbreakers, and they cannot be trusted in the same way an Ohanedin can. 
Even other Din are like this, though less so. Be cautious around outsiders, but true to those who prove 
worthy of your trust. 
\subsection{Orredin}
\subsubsection*{Overview} 
The Orredin are a powerful and ancient race of Din, often called Helians, or High Elves. From their 
floating city above the Broken Sea, they have ruled empires forgotten to man for eons. When humankind 
was still knapping spears from flint, the Orredin had already developed public sanitation, double-entry 
accounting, and several varieties of espresso. Their traditions of intellectualism, philosophy, engineering, 
medicine and sorcery are all ages-long, and there is surely no race of men or din more educated and sophisticated than they. 

Of course, these great powers make them very arrogant, and given that few Orredin have ever stepped 
down from their ivory towers, most of them are in fact rather inexperienced in temporal affairs, and their 
extensive knowledge is not (as many of them believe) an actual substitute for experience and practice. 

Though beings of terrible power and intellect, Orredin who venture forth on adventures and expeditions 
often find themselves rudely awakened to the hardships of everyday life--and those who survive are 
greatly strengthened by the revelation. 
\subsubsection*{Appearance} 
The Orredin are otherworldly and dignified creatures. Humanoid, with short but delicately pointed ears 
and refined features. They are not very tall compared to the other Dinnic races. The men stand at about 
5’6, the women several inches shorter. Most have fair skin and hair. Their eyes are distinct for their 
uniformly purple hue--a product of their connection to the Warrens, the primal source of Sorcerous 
Energy. Orredin cannot grow full beards, though some can muster goatees or thin mustaches. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} +2 Intelligence, +2 Willpower, -2 Health, -1 Strength 
Always have \nameref{boon:Beautiful} Minor (can buy \nameref{boon:Beautiful} Major at difference in cost)

\paragraph{Focus:} Like all Din, the Orredin have a Focus which allows them to remain ageless and immortal. The 
Orredin focus is to simply be in a Warren with 50 or more Sorcerous Flow. The Flow can be utilized by 
devices or spells, so long as it is there, any Orredin in that Warren has Focus. Alternatively, consuming 
(usually drinking a solution of) about a teaspoon of powdered magestone (typically costing 10 gp) will 
satisfy an Orredin's Focus for a year. 
\subsubsection{Apotheosis:} Orredin have a special pool of points called Gnosis. Each point of Gnosis is a point of extra 
Flow that an Orredin has available to him regardless of what Warren they are in. Gnosis moves with the 
Orredin, and can be used, occupied etc. just like normal Flow, except that it can only be used by the 
Orredin himself.

Gnosis can also be spent to activate Bloodline Abilities. Doing this temporarily “spends” the Gnosis 
points, which cannot be used for Flow or other purposes until they return. Gnosis regenerates (up to the 
Orredin’s Gnosis Cap) at a rate of 1 Gnosis per [30-Willpower] minutes. 
An Orredin starts with 1 Gnosis, and can increase their Gnosis cap in the following ways:
\begin{itemize}
	\item An Orredin may spend 15 Arc Points to increase his Gnosis Cap by 1. Only 5 Gnosis can be gained in this manner. 
	\item Consuming (usually drinking a solution of) two and a half tablespoons (about 100 gp worth) of Magestone will increase an Orredin's Maximum Gnosis by 1. Only 5 Gnosis can be gained in this manner.
	\item Absorbing energy from certain ancient relics, powerful artifacts and rare devices will increase an Orredin's Maximum Gnosis by 1. Only 10 Gnosis can be gained in this manner. 
	\item The last way to increase Gnosis is to keep calm and Sieg Dion.
\end{itemize}
\paragraph{The Godhead is a Spring:} Every Orredin is at least a Minor Sorcerer, and is considered to have X points in the Magic Category devoted to Sorcery. The level can be increased with additional expenditure in the Magic Category at Character Creation. 
\paragraph{From it flows Everything:} Each Orredin in a Sorcerous Warren increases the Flow value of that Warren by 2, and of all adjacent Warrens by 1.
\paragraph{And then returns once more, and so is like a Spring:} All Orredin have a Bloodline Power that can be selected from the list below. Additional powers can be purchased in the Magic category at Character Creation. These are just examples. Confer with your GM about personalized Bloodline Powers that may fit your character better.
\subsubsection{Bloodline Powers}
\paragraph{\large\label{man:Killing Glare}Killing Glare}---\quad{\large[X+1]}\par
The Orredin can emit thin, visible beams of energy from his eyes that can shear through almost any material like a hot knife through butter. Activate this ability with Gnosis [X+1].

When Activated, the Orredin may shoot lasers out of their eyes. These lasers are treated as a Shooting 
Missile Weapon capable of Rapid Fire (with no limit except MP) with the following Profile:
\begin{table}[!ht]
	\caption{Killing Glare}
	\label{tab:Killing Glare}
	\begin{tabu} to \textwidth {XXX[-1,c]X[c]XX}
Weapon Name&Weapon Type&Range& Missile TN&Damage&Special\\\toprule
Killing Glare&Elf Laser&20&4&(WIP+X)c&Flaming 3\\
	\end{tabu}
\end{table}

There is no “Eye Lasers” Proficiency. For the purposes of Missile Pool, Killing Glare uses Perception instead of Proficiency. Aiming is possible, and adds Perception a second time. Bracing is not possible with Killing Glare. Other Orredin may simply deflect the ability harmlessly by spending Gnosis equal to X. Sometimes, Killing Glare will manifest itself as a beam emitted from the mouth or a hand.

\paragraph{\large\label{man:Kneel!}Kneel!}---\quad{\large[X]}\par
When activating this ability, the Orredin may issue a command, generally no longer than two or three words, directed at a living being. That being must make a Willpower test at RS equal to 2+(X/2) or do its best to obey the command. If the command is dangerous or potentially suicidal, the target gains a bonus of +3 to its Willpower test. If the command is overtly suicidal (“Die” is a valid command, causing the 
target’s heart to stop) the target doubles its Willpower for the purposes of the test. Kneel! is only binding 
for about ten seconds (5 Rounds in combat) after which the effects wear off. Multiple targets may be selected for Kneel! At a cost of +2 Gnosis per target. This ability does not work on other Orredin. 


\paragraph{\large\label{man:Majesty}Majesty}---\quad{\large[X]}\par
The Orredin may produce light, an aura of majesty and power, and create unearthly music, noises or 
whispers around them at will. It become very difficult to argue with, contradict, or even disagree with an 
Orredin in this state, much less to attack one. When activating this ability, the Orredin's Charisma doubles for a number of minutes equal to X. Any attempt to attack the Orredin suffers a reduction in CP/MP for 
the Maneuver equal to the Orredin's doubled Charisma score. 
If the Orredin makes any attack himself, or runs away (quickly, slowly backing away is fine) panics, is hurt, etc., the effect is immediately shattered. 
This ability does not work on other Orredin. 


\paragraph{\large\label{man:Purple Vortex}Purple Vortex}---\quad{\large[X]}\par
The Orredin can stare into his opponent’s eyes, and enter a state of mutual hypnosis during which he can exert subtle control over his opponent. This is useful in things like dancing or debate or darts, but it is perhaps most useful in combat. 
In Combat, after all Actions have been declared but before they are resolved, if the Orredin can look into 
an opponent’s eyes clearly (a closed helmet would prevent this, or being blind, shutting one’s eyes, etc) 
an Orredin may activate this ability by spending X Gnosis. Roll a Willpower Contest between the Orredin and the target.
\vspace*{-15pt}\subparagraph{Success:} Treat X amount of dice as if they had rolled successes on an unresolved maneuver this action.\par
This ability can be negated by other Orredin by spending Gnosis equal to X.
\subsubsection*{Culture and Society} 
The Orredin primarily live in the flying city of Helion, where their race has endured since the cataclysm that destroyed their empire millennia ago. Their numbers were sorely reduced in the disaster, and most of them were related to each other. Keenly aware of the dangers of inbreeding, the Orredin enacted a centuries-long Eugenics program, aimed at gradually assimilating fresh blood into the genepool, and 
restoring their numbers.

To find proper candidates, human and Dinnic mercenaries from across the land 
are hired and sent on suicidal missions, tested for courage, physical perfection and other attributes, and 
engaged in gladiatorial combat. Only the finest are adopted into the Orredin’s fine society, the rest are 
maimed or perish. Despite these dangers, the supply of mercenaries hungry for glory, wealth and the 
Orredins’ favor never runs dry. 

The Orredin themselves are a fastidious, tasteful and clean people. Their cities are magnificent triumphs 
of exquisite city-planning, peerless architectural genius, and almost utopian excess. What sorcery and 
good administration cannot accomplish, the endless legions of slaves who toil beneath the city can. 
Helion, the mighty capital, is the largest and most glorious city in the world, and it is itself only a 
hundredth of what it used to be, the rest having fallen into the sea below during the cataclysm that 
destroyed the old Empire. 

Orredin society is complex, revolving around public offices, familial politics, favors and obligations, and 
an ironclad yet unspoken set of social conventions that dominate every interaction. For the Orredin, status 
and advantage is all there is. The lower ranking ones, those who either could not or would not play “The 
Great Game,” often spend their days plying artisinal trades or other noble activities. Those who climb the 
ladder are often involved in labyrinthine plots spanning hundreds of years and dozens of kingdoms, 
manipulating events across the mortal realm to bring them some small advantage at home. Others are 
involved in the Purple Guard’s machinations, taking a personal interest in the potential future of the 
Orredin race. 

A few--very few--reject all of this, and leave Helion. Exiles either by choice or by law are very rare, and 
such individuals are often assumed to be involved in some scheme so important that it necessitated 
leaving Helion, something most Orredin have difficulty imagining in the first place. Sometimes this 
stereotype is true. Other times the exile’s mystique is the entire reason they enter exile. Still other times it 
is neither. Orredin too can feel the pull of curiosity, and the lust for adventure. The world as it looks from 
Helion is small and trivial, like a game board with kingdoms and nations for pieces, but every so often an 
Orredin starts to wonder if it is not worth investigating the realm of ants and pawns for themselves. 
\subsubsection*{Ecology} 
Orredin have very delicate constitutions, and their stomachs cannot tolerate any sort of turbulence without 
making them unpleasantly ill. The Orredin diet favors honey, olives, fruits, nuts, wine, pure white bread 
made of good wheat, vinegar, and sometimes fish. When meat is eaten, birds and fowl are preferred, with 
red meat only being served in small portions, often carefully prepared and seasoned for digestive 
purposes. Like most Din, the Orredin have a very high tolerance for alcohol, and they prefer drinking 
wine instead of water, for the purposes of sanitation. They are lactose intolerant, and cannot drink milk or 
eat cheese. However, they are very fond of coffee, which they prepare with almond-milk and drink 
constantly. 

Many believe that the Orredin do not require sleep, like Zells--this is not true, most of them 
are simply too caffeinated to close their eyes more than once or twice a week. The withdrawal symptoms 
of centuries of caffeine addiction do not bear thinking about, and as such the Orredin are very particular 
about maintaining a steady supply of coffee beans from Kedoua and Karthack. Wars that have killed 
millions of people have been fought over the coffee supply. 

Orredin women can birth a child once every twelve years, with a gestation period of thirty six months. As 
a result, replenishment of their population is extremely slow, and every child is an enormous investment 
of time and resources, made more complicated by the fact that most Orredin alive today are related to one 
another, and fear of inbreeding has made most of them bachelors. 
Interracial pairings always produce a non-Orredin, but it has been found that after four or five generations 
of controlled breeding, the Orredin characteristics can be made to present themselves once again. A 
program exists to select humans of sufficiently excellent stock to be inducted and married into Orredin 
society. Though their lifespans are short, their offspring make up the vanguard of a new generation of 
Orredin, each the product of hundreds of years of controlled breeding. 

\subsubsection{How to Roleplay an Orredin}
When roleplaying an Orredin, remember that they are an enormously privileged and educated people, 
used to civilization and sophisticated society. Even the poorest and most despondent Orredin who grew 
up in Helion will be used to a life of plenty and leisure. For many, adventuring, or even venturing down 
from Helion, is an awful and draining experience that they would much rather avoid. Some are 
adventurous and take to this “life in the rough” with gusto, others tolerate it, and settle for occasionally 
escaping into a cup of coffee or a fine meal.

They are also obsessively cleanly. Orredin have delicate constitutions and are frightened of illness. They 
would bathe multiple times per day if they could. 

As an Orredin, you are smart, you know better (and if you don’t, it’s only because things are so 
uncivilized down here) and frankly, you are better than just about everyone else. You won’t just say it 
casually, but humans, dwarves, even other din, they’re all inferior to you. Deep down, that’s been bored 
into you from childhood, and compared to the glory of Helion, most things you see on the surface just 
confirm it. You might not be a jerk about it, but when push comes to shove, it’s one of the cornerstones of 
your cultural worldview. 

\subparagraph{Male Names:} Diomedes, Benedictus, Alexander, Volkanius, Kelkotattius, Heracles, Damocles 
\subparagraph{Female Names:} Camilla, Phyrne, Gorgo, Hydna, Anyte, Hipparchia 

\subsubsection*{Culture and Society} 
XXX 
\subsubsection*{Ecology} 
XXX 
\subsection{Paleolithic Elf}
\subsubsection*{Overview} 
XXX 
\subsubsection*{Appearance} 
XXX 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} XXX 
XXX 
XXX 
\subsubsection*{Culture and Society} 
XXX 
\subsubsection*{Ecology} 
XXX 
\section{Ascended Humans}
Ascended Humans, or ‘Ascendants’ are Human beings who have returned from death changed. Their faith, dedication and willpower in life was such that they were able to achieve a sort of reincarnation, 
returned to the mortal realm through the patronage of a deity. Ascendants are the immortal servants of 
their patrons, returned to see the will of their god done. No other race besides Humans can Ascend, and 
this has been the subject of many serious inquiries as to the nature of Humankind relative to the other 
races. The Orredin, for example, resolutely believe that the Ascendants are proof that Humans are actually 
Din, and that their Focus is dying. Whatever the case, the Ascendants exist. They are proof that there is 
something in the Human spirit more powerful than death. 

There are several varieties of Ascendant, each will be described in detail below, but they also all share some basic traits. 

\paragraph{Bonuses}
All Attribute caps are fixed at 10.
\subsubsection{Touched by the Source}
Ascended Humans can never truly die, so long as their faith is maintained. An Ascendant's body can be 
destroyed or killed just like that of a regular Human, but after a certain amount of time it will inevitably 
be reborn. The exact time frame and method of this varies depending on the type of Ascendant, and 
whether certain conditions have been met. Ascendants never return into a situation in which they would 
immediately be killed or captured, unless they (or their divine patron) wills it. The Ascendant also has no 
say whatsoever in whether or not they are reincarnated; once the contract is signed, it's for eternity. 
 

\subsection*{Sarturi Chosen}
After a number of years equal to 1,000 minus one year for every Human that the Sarturi Chosen killed in 
his most recent ‘life,’ the Sarturi Chosen will return. If a great battle or massacre occurs that sheds great 
quantities of Human blood, the Chosen will emerge from one such pool, as though it were of much 
greater depth, and strike once more into the world of the living. 
\subsection*{Genosian Paladin}
After a number of years equal to 1,000 minus one year for every point of Divinity that the Genosian 
Paladin either had in reserve, or had spent within 24 hours of his death, the Paladin will return through the 
Solar Gate on Mt. Genosus. 
\subsection*{Dessian Silver Guard}
After a number of years equal to 1,000 minus one year for each Human Adherent the Silver Guard had at 
the time of death, the Silver Guard will be reborn to Dessia in the Black Citadel of Kar-Karthack. 

Additionally, Ascended Humans do not age. 
\subsubsection{Divine Power}
Ascendants power their abilities using Divinity, a sort of mystical energy they gain by performing their 
duties. Each type of Ascendant has different means to acquire Divinity. 

If an Ascendant's Divinity is exhausted (i.e. they run out of Divinity, depleting to 0 or below) the 
Ascendant ‘Falls.’ A Fallen Ascendant immediately disintegrates from the mortal realm, as their bodies 
are sustained by Divinity. If an Ascendant's Fall occurs for reasons that their patron deity would approve 
of, they will return as normal. However, if the Fall was incurred through misbehavior, sheer stupidity, or 
incompetence, they may neglect to reincarnate the Ascendant. 

All Ascended Humans have Divinity equal to their WIL score at Character Creation, or when returning from death. 
\subsection{Genosian Paladin} 
\begin{quote}
\emph{"God is Justice."}

\hfill--\textbf{Mirza Kovac}, Genosian Paladin 
\end{quote}
\subsubsection*{Overview} 
The Paladins are the agents of the Indefatigable Sun, the deity called Genosus. Genosus is a god of labor, 
industry, dignity, and order, and his churches are well-integrated into the societies of countries where they 
have large followings. The Genosian Paladins are the direct servants of Genosus. It is said that the 
afterlife given to Genosus' followers is an eternity on the Solar Ziggurat, a mighty spiritual construction 
that is always being expanded so as to house the souls of more and more followers. The Paladins are those 
whose willpower, industry, and commitment to good works are so great that they are able to scale the 
Ziggurat and reach the Solar Gate — a stable portal between Genosus' sacred realm and the mortal world — and then cross over. There they are given new flesh, and act as Genosus' living agents on Mundus.

Paladins have tremendous powers both to destroy and to heal; they are terrible foes to face in war, and 
mighty allies to have in peace. They are most commonly used to hunt witches, heretics, abominations and 
infidels who threaten the Genosian way of life, but occasionally they offer counsel to rulers, heal the sick 
or wounded, and even provide direction for great labor projects. 
\subsubsection*{Appearance} 
Paladins generally look more or less the way they did in life, although their time on the Solar Ziggurat 
tends to bronze those who had lighter skin, and they retain that tan indefinitely. Even albinos who become 
Paladins find that they can no longer suffer sunburns. In direct sunlight, a Paladin's head always seems to 
have a glow to it, like a corona of light, which marks them clearly as agents of Genosus. Aside from this, 
Paladins can look more or less like any other person, though it has been noticed that no Paladin has ever 
been returned to life with any sort of physical defect that would threaten their effectiveness. In Genosus' 
mercy he repairs such flaws by hand before delivering them back to life. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} None 
\paragraph{Genosian Ideal}
Paladins cannot have begin play with any sort of Bane that would physically cripple them such as 
scars, missing limbs, old wounds, lasting pain, lost eyes, brain damage, and so on (though being 
tall or short, fat or skinny is acceptable). These Banes can be gained during play, but cannot 
normally occur at Character Creation. 

At the GM's discretion, a Paladin may have acquired an 
injury during his career after emerging from the gate, and so may have one or more of these
Banes, but the cost of the Bane is halved for the purposes of B\&B points, but for the purposes of
Healing (Paladin ability, explained below) or other requirements, the cost of the Bane counts as normal.
Paladins cannot be Complete Monsters, because they must have a Faith Arc in order for them to become Ascendants. 
\paragraph{Glory, Amen}
Paladins gain Divinity in the following ways:
\begin{itemize}
\item Successfully using Judgement to destroy something evil gains the Paladin between 1 and 
10 Divinity depending on seriousness of a threat, with 1 being a petty arsonist, and 10 
being an Elder Dragon that was bent on consuming the world. 
\item Spending a full day engaging in honest labor gains the Paladin 1 Divinity. 
\item Organizing the honest labor of good Genosians (or potential converts) for a full day gains 
the Paladin between 1 and 3 Divinity. 
\item If an infidel genuinely converts to Genosism because of the Paladin's words or actions, he 
gains 1 Divinity. 
\item Killing any sort of infidel, heretic or apostate whose has deliberately set themselves 
against Mighty Genosus, his people, or his church(es) gains the Paladin between 1 and 5 
\end{itemize}
Divinity depending on the severity of their crimes (1 being a peasant who abandoned the faith out of frustration, 5 being a prominent priest who defected to Chaos Worship or worse).
\paragraph{Judgement}
A Paladin can channel the power of Genosus through any weapon—even his own hands—to strike the enemies of his God with terrible vengeance. When a Paladin performs a melee maneuver that inflicts damage (such as Swing and Thrust, though Shoot and Melee Shoot do not count) he may also declare that he is performing Judgement.\par

If the attack hits, in addition to any damage from the attack itself, Judgement inflicts a Fire Damage state not just on the Hit Location, but on all Hit Locations in the Target Zone. So, if a Paladin uses Judgement and strikes an opponent, aiming for the Lower Arm (Target Zone), and hits the victim on the Hand (Hit Location), the victim's Elbow, Forearm and Hand all erupt into flames.

Judgement's Fire Damage is X/TN 5, 3 rounds. X is the Paladin's WIL score.

However, Judgement will not work on just any target. The target must be one who is an enemy of 
Genosus and Genosism, who has somehow threatened, wronged, or attacked Genosians... or 
someone who has mocked Genosus, Genosian teachings, or the divine works of Genosus. In 
short, it must be someone Genosus would want dead. How they feel about or view themselves, or 
how they justify their actions, has no bearing on the effect of Judgement. 

If Judgement is performed on an innocent person, not only does the attack not inflict Fire Damage, but the Paladin loses 3 Divinity immediately. 

If a target struck by Judgement perishes as a result of the attack (either the fire or the raw 
damage) the Paladin gains an amount of Divinity based on how great of a threat the victim was 
(see Glory, Amen for more details). This is generally between 1 and 10, with 1 being a common 
criminal, and 10 being a threat to the entirety of Genosism. 
\paragraph{Shield of Dawn}
Paladins are protected by the will of Genosus, the Indefatigable Sun. This extends to the point 
that missiles flung at them seem to be slowed by the very light itself.

Paladins gain AV versus Missiles and Explosions (this counts as Cover AV) of all varieties so 
long as they are in contact with light. The amount of AV they gain depends on the intensity of the 
light they are exposed to.\par
\begin{table}[!ht]
	\caption{Shield of Dawn AV}
	\label{tab:Shield of Dawn AV}
	\begin{tabu} spread 0pt {Xc}
		Light Intensity& AV vs Missiles\\
		Faint luminescence (certain mushrooms, fireflies, the barest light of dying coals)& 1\\
		Moonlight or Firelight &2\\
		Overcast day, intense artificial light, Paladin's own radiance& 3\\
		Direct sunlight& 4\\
		Direct sunlight within a minute of having used the Judgement ability& 5\\
	\end{tabu}
\end{table}
If there is no light to be had, the Paladin can spend 1 Divinity to produce radiance of their own 
for a span of about ten seconds. If this is used in conjunction with existing light worth 10 or more 
AV, it improves the AV granted by Shield of Dawn by 2. 
\paragraph{Healing}
A Paladin can use Divinity to work the miracle of healing on the wounded or the sick. They can 
even, in extreme cases, restore destroyed limbs, cure blindness, and even madness.

Spending 1 Divinity per level of the Wound will heal, over the course of a few seconds (in 
combat, one Round per one level of the Wound) and halt the bleeding until it is healed. The 
Paladin must remain in contact with the wounded party for this entire time. The Wound is not 
reduced over time, but is entirely removed when the healing process is complete. 

A Paladin can heal himself at double this speed (one Round per two levels of the Wound) and can 
continue taking other actions as he does so. A Paladin's Healing Touch is doubly effective against 
Burn Wounds, with each point of Divinity curing two Wound levels. 

A Paladin can cure most diseases with 1 to 3 Divinity (1 being a cold or flu, and 3 being The Red 
Death) by praying for several moments, and then slapping the subject on the side of the head. If it 
doesn't hurt, it doesn't count. 

Injuries that involve a Bane (a lost eye, a severed limb, a crippled limb, brain damage, and so on) 
are more difficult to repair. The Paladin must spend several moments collecting himself (as long 
as a minute in some cases) and then strike the subject with his hand, and shout. At this point, both 
the Paladin AND the subject need to ‘buy off’ the injury Bane. The Paladin must spend an 
amount of Divinity equal to the cost of the Bane, whereas the subject must spend Arc Points (hey, 
miracles don't work themselves—you need faith!). If either party doesn't have enough points to 
do it, the Paladin will be able to tell before actually performing the ritual. 


If successful, limbs regrow instantly (the new flesh tends to have a golden hue to it, and is numb 
and difficult to use for about an hour) eyes regrow, bones reset themselves, scarred lungs become 
pink and new, madness fades, and brains restore themselves from damage.
\paragraph{Wrath of God}
Judgement isn't a power fueled by Divinity. It is a fountain of Divinity. Performing their duty to 
judge the wicked and the depraved is one of the chief sources of a Paladin's power. But it is one 
thing to merely punish the wicked, and quite another to bring their sinfulness to Genosus’ 
attention, and to let him deal with it in his own way. 


Wrath of God creates an explosion centered on the Paladin. This explosion will not harm the 
Paladin himself, or individuals that he explicitly doesn't want to injure, but it will harm 
everything else around him, animate or inanimate. Wrath of God automatically consumes all of a 
Paladin's Divinity except for one point. 


The explosion caused by Wrath of God has a Radius of 1 yard and a Power of 1 per point of 
Divinity spent in this way. This is Bludgeoning Damage, and also inflicts Fire Damage of 3/TN 5 
that burns for 3 rounds to any Hit Location that suffers damage from the explosion. The Fire 
Damage can be avoided if a character benefits from enough cover that a Hit Location's damage 
from the explosion itself is reduced to 0, in which case that Hit Location suffers no Fire Damage 
either. 


Individuals killed or enemies destroyed by Wrath of God can never give a Paladin Divinity. 
Essentially, it doesn't count as the Paladin killing them. It's Genosus doing the heavy lifting for 
once. 


A Paladin with 12 Divinity uses Wrath of God. He creates an explosion 22 yards in diameter, centered on him, with 11 
Power. One poor man is within 11 feet of him, and so suffers 11 damage, minus his TOU and AV, for a total of 7, 
scattered across him on the Random Missile Hit Table. He suffers several Level 1 and 2 Wounds, and one serious Level 
3 Wound on the chest. However, each of these locations is also lit on fire, suffering additional Fire damage for 3 
Rounds. 
 
\paragraph{I See You}
Paladins do not have any special ability to tell if someone is an infidel or a heretic just by looking 
at them. However, a Paladin can ‘Mark’ any living thing that he would be able to use Judgement 
on as an enemy of Genosus. This involves touching the target in any way—even just a light brush in passing will do—and willing it to be so.


When this is done, the Paladin must spend 1 Divinity, and assign a ‘name’ to the target. It doesn't have to be the thing's real name, but as long as the Paladin associates that name with it, it is enough. From then on, whenever ANY Paladin thinks of the name assigned to this marked target while concentrating, they will feel a pulling sensation in the direction that the target is, varying in strength depending on distance.


Two targets cannot share a name. If a Paladin attempts to use a name that is already taken, it doesn't stick. Of course, once a target has been destroyed, its name is up for grabs again.


The marked subject may not be initially aware of this effect, but just as the Paladin feels a tugging sensation towards his target, the target will itself become aware of the Paladin.
Sometimes this manifests in the same way it does for the Paladin—a tugging sensation on the ear or tongue—other times it manifests as phantasmal sounds.
Generally this effect is disturbing, and even if they don't realize that it signifies a Paladin's coming, the marked individual will generally be aware that something bad is on its way. 


Examples include:
\begin{itemize}
\item The creaking of chains that grows louder and more frantic as the Paladin approaches. 
\item The beating of drums and chanting, growing more furious as the Paladin approaches. 
\item A tugging sensation on the tongue, and whispers that intensify as the Paladin approaches. 
\item A pleasant tune in the distance that grows more distorted and ominous as the Paladin 
approaches. 
\end{itemize}
\subsubsection*{Culture and Society} 
XXX 
\subsubsection*{Ecology} 
XXX 
\subsection{Dessian Silver Guard}
\begin{verse}
	\emph{Bocanadessia's hand is long,\\
It stretches like mist from the sea,\\
Around it flesh chills and puckers,\\
But souls are saved by ivory touch.}\\

--\textit{‘Karthack is Eternal’}, National Epic of Karthack
\end{verse}
\subsubsection*{Overview}
The Silver Guard are Ascendants born of the living god Bocanadessia, the reincarnated souls of her finest 
servants, changed by the unearthly moonlit path they traveled in death to rejoin the living. The Silver 
Guard have very personal relationships with their mother goddess, and take orders from her directly 
through the Lunar Song. They are the living will of Bocanadessia, and when they leave the great fortress 
that is their home, it is often to travel for years enacting the will of the divine. To this end they have 
powerful abilities both to protect themselves from harm, and also to benefit those who commit themselves 
to the worship of Bocanadessia. 

Each Silver Guard is a sort of miniature deity in Dessian culture, demigods that deserve worship in their 
own right. Most Dessians are ‘Adherents’ of a Silver Guard, and pray to Bocanadessia through the 
conduit that is the Silver Guard. This is, in fact, the source of the Silver Guards' power—each follower 
they mark as their own allots them a greater share of the Divinity supplied to them from above each 
month. Some of the Silver Guard act as warriors, generals, and even assassins. Others act as spymasters, 
administrators, and coordinators. All are feared and respected, even by those who pay no homage to Bocanadessia. 
\subsubsection*{Appearance} 
Silver Guards look much like the people they were in life. Their flesh and eyes are usually the same color 
(occasionally they are born albinoid, with an ivory white complexion), but their hair is always straight and 
bleached of all pigment, with rare exceptions. They are often beautiful and fine to look upon. The Silver 
Guard share a kind of ‘family resemblance’, that is undoubtedly the result of their divine heritage; each 
have minor differences from their original looks, and seem to share physical traits. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} XXX 
\paragraph{Perfect}
Silver Guards cannot begin play with any sort of Bane that would physically cripple them, such 
as scars, missing limbs, old wounds, lasting pain, lost eyes, brain damage, and so on. These Banes 
may be gained during play, but cannot normally occur at Character Creation. At the GM's 
discretion, a Silver Guard may have acquired an injury during his career after emerging into this 
world, and so may have one or more of these Banes, but the cost of the Bane is halved for the 
purposes of B\&B Points, although for the purpose of buying off Banes, they are counted at the 
standard cost. 


Silver Guards, like Paladins, cannot be Complete Monsters, because they must have a Faith Arc
in order for them to become Ascendants. 


Additionally, unlike Paladins, Silver Guards cannot take any sort of Bane that would render them unattractive. They are chosen specifically for their looks. They cannot take the Skinny, Fat, or Ugly Banes. 
\paragraph{The Moon Also Rises}
Silver Guards gain Divinity in the following ways: 
\begin{itemize}
\item Completing a task set for the Silver Guard by Bocanadessia or her Song grants between 
1-5 Divinity, depending on the importance or difficulty of the task (1 being a minor task 
or a daily function, 5 being a great service to Dessianism as a whole). 
\item Slaying or defeating any enemy of Sacred Karthack or Bocanadessia herself grants a 
Silver Guard between 1-5 Divinity, depending on the severity of the threat (1 being a 
minor foe, 5 being a mortal enemy of Bocanadessia). 
\item Each Adherent a Silver Guard has grants him 1 Divinity per month, on the night of the 
Eclipse. 
\item Marking an Adherent grants a Silver Guard 1 Divinity immediately. 
\end{itemize}
\paragraph{Moonstrike}
When a Silver Guard performs a melee maneuver that inflicts damage (such as Swing and Thrust, 
though Shoot and Melee Shoot do not count) he can declare that attack to be a Moonstrike. 


If the Moonstrike inflicts a Wound of any sort, the attack also inflicts Cold Damage (equal to the 
Silver Guard's WIL score) to the attack’s Hit Location. 


In addition, the Silver Guard may spend any amount of Divinity. Each point of Divinity spent has 
the following effects: 
\begin{itemize}
\item The victim suffers 1 point of Cold Damage randomly allocated on the body. 
\item The victim loses the ability to benefit from one of his Arcs for one month, or until any 
Silver Guard grants him Succor. This can stack. 
\item The Silver Guard regains 1 CP. 
\end{itemize}
If the target of this ability is another Ascendant, they can negate (or reduce the effects of) Moonstrike by spending Divinity of their own, on a 1-for-1 basis. 
\paragraph{The White Fog} %REALLY don't like that little wraptable break on the second-to-last row
	\setlength{\intextsep}{0pt}
\begin{wraptable}{r}{9 cm}
	\caption{Conditions for the White Fog}
	\label{tab:Dessian White Fog}
	\begin{tabu} spread 0pt {Xr}
		Condition&Modifier\\
		Daytime&-10 yards\\
		Night&+10 yards\\
		Extremely Hot&-20 yards \\
		Hot&-10 yards \\
		Cold&+30 yards \\
		Extremely Cold&+50 yards \\
		No Moisture&-10 yards \\
		Some Moisture&+10 yards \\
		Tons of Moisture (snow, rain, ocean)&+20 yards \\
		Another Silver Guard is present&+10 yards each \\
	\end{tabu}
\end{wraptable}
	\setlength{\intextsep}{\oldintextsep}
A Silver Guard may choose to passively drain the heat out of an area around himself, and to cause 
a mist to envelop the area, reducing visibility by one light stage. If there are multiple Silver 
Guards and their fields of White Fog overlap, visibility is reduced by one light stage per field 
overlapping. Silver Guards and their Adherents can see through this fog clearly, though they 
cannot normally see in the dark. 

Additionally, within this area, the ambient temperature drops by 
one stage. As before, if there are multiple Silver Guards creating fields of White Fog, the effects 
combine. Light sources still function in the White Fog, but the range that the light carries is  halved, and weak flames, like candles, may be extinguished. 

Additionally, any Fire Damage inflicted in the White Fog has its duration reduced by 1 per 
overlapping field of White Fog, to a minimum of 1 round of burning. 

Any missile attack or explosion that would effect a Silver Guard in the White Fog treats the White Fog as cover with an AV value equal to the Silver Guard's WIL, +1 per overlapping White Fog from other Silver Guards. 


The size of the White Fog field depends on the current climate and time of day. By default, the area has a radius of 50 yards.
\paragraph{Moonlit Twostep}
The Silver Guard is not constrained by physical laws in the same way men are. They move with 
an unearthly grace, sometimes seeming to glide rather than walk, and in battle, they can briefly 
gain altitude and launch themselves like birds of prey down upon their enemies. 


Silver Guards gain a +4 bonus to their MOB for the purposes of movement and for mobility 
maneuvers, as well as a +2 bonus to Skill and Attribute tests made to jump, climb, maneuver, or 
otherwise move. A Silver Guard may spend one Divinity to perform the Soaring Charge 
Maneuver in the first Action of combat if his Orientation is Red. 


\paragraph{\label{man:Soaring Charge}Soaring Charge}---\quad{[2+Encumbrance Penalty]}
\begin{description}
\item [Special:] Silver Guard, Instant. 
\item [Requirements:] Be a Silver Guard. Use this Maneuver only if you threw Red. 
\item [Maneuver:] Declare an Attack as normal. However, you gain a +2 bonus to Initiative rolls made during this Action (this includes Red/Red tests) and a +2 bonus to any damage you inflict with Attack Maneuvers (except Melee Shoot) during this Action.
\item [Special:] For the rest of the Round, your Defenses gain the following rule: "\textbf{Failure:} Enemy attack is not reduced by Successes from this maneuver, and hits with all Successes as BS."
\end{description}

\paragraph{Mark of the Moon}
A Silver Guard can ‘Mark’ an individual as an ‘Adherent’ if they are a true believer in 
Bocanadessia. This has no cost, and can be done to any willing individual who has a Faith Arc 
that venerates Bocanadessia, or who is willing to convert (the act of Marking instills a new and 
virtually unshakeable faith in the individual), and isn't a Complete Monster. Being a Complete 
Monster prevents this ability from working. Marking an Adherent involves putting a hand on 
their head or face, and willing it to be so. 

Marked Adherents gain a spectral imprint of a handprint and a moon (in the phase that it was in at the time the Adherent was Marked) that can only be seen by Ascended Humans. Paladins and 
Sarturi Chosen can see it as well. 

Marked Adherents gain the following benefits so long as they remain true to their faith: 
\begin{itemize}
\item They have a +2 bonus to their HLT for the purpose of resisting disease, poison, or blood 
loss. 
\item They can hear the Song of the Moon more clearly. 
\item They can be granted Succor by any Silver Guard. 
\item They gain double the normal amount of Arc Points for following their Faith. 
\end{itemize}
The downsides, though, are noteworthy: 
\begin{itemize}
\item If their Faith ever changes from the pure veneration of Bocanadessia, the Moon-mark on 
their face will explode. This will inflict Concussive Damage equal to the Silver Guard's 
WIL, to the part of the Adherent's body that the mark is on. This damage is not reduced 
by TOU or Armor, but if the Adherent's WIL is higher than the Silver Guard's, than the 
difference between the two scores is subtracted from the damage received. (i.e. if the 
Adherent has 8 WIL and the Silver Guard has 5 WIL, then the damage is reduced by 3, 
bringing it down from 5 to 2). If they somehow survive this, they are no longer an 
Adherent. 
\item If the Silver Guard who Marked the Adherent gives the Adherent an order, the Adherent 
must pass a WIL test at an RS equal to the Silver Guard's WIL in order to not obey it. 
This includes an order to, say, ‘drop dead’ or ‘swallow that carving knife.’ 
\item If the Silver Guard is killed violently, all of his Adherents must make a HLT test at RS 5 
or die instantly. One Adherent at random may be spared from this fate per point of 
Divinity the Silver Guard had in his repository when he died (so, if a Silver Guard has 30 
Adherents and 25 Divinity, and is crushed by a boulder, only 5 of his Adherents will have 
to make HLT tests). 
\end{itemize}
Marking can be removed by other Ascendants without triggering the death of the Adherent. The 
Adherent doesn't need to be willing, but the Ascendant must spend 15 Divinity, and put their 
hand over the Mark. Then, a \textbf{WIL} contest is made between the other Ascendant and the Silver 
Guard who marked the Adherent. If the Ascendant is successful, the Adherent's Mark is removed. 
If their Faith was changed by the Marking, it is now changed back to what it originally was. 
\paragraph{Forcible Marking}
It is actually possible to forcibly turn someone into an Adherent, but this is a secret that most 
Silver Guards would rather not be revealed, and not many of them are capable of it to begin with; 
only the most powerful can do it. 


This ability is not available to regular Silver Guards, but can be purchased for 50 Arc Points.


Performing it requires the expenditure of Divinity equal to the WIL of the victim, and an 
additional expenditure of X. The Silver Guard must successfully grasp the victim's face (trapping 
the head in a grapple would work) before doing this.
The Victim must make a \textbf{WIL} test at RS equal to X, or become an Adherent against his will, 
complete with faith in Bocanadessia. While generally the new Adherent will still remember, and 
be confused by the fact that they weren't an Adherent until a few moments ago, the use of Succor 
to drain away their negativity can keep them under control. 


This ability cannot be used on characters with the Complete Monster Bane. 
\paragraph{Succor}
Silver Guards have the ability to grant healing to their Adherents (or the Adherents of other Silver 
Guards) through the use of Divinity. This is not as powerful as Paladin healing--the Silver Guard 
cannot instantly close wounds, nor can they regrow lost limbs, and it cannot be used on the Silver 
Guard himself, but it is much easier for the Silver Guard to do. 

A Silver Guard can cure most diseases by praying over the kneeling or prostrate Adherent, and 
then laying a hand on the spectral brand they are marked with. This costs 0 Divinity if the disease 
is something non-fatal but merely unpleasant, and 1 Divinity if the disease is potentially or actively lethal. 

Mental illnesses of all kinds (including hereditary ones that have no actual cure or treatment) can 
be removed with a touch. The Silver Guard only needs to spend 1 Divinity, but the Adherent must 
spend Arc Points equal to the cost of the Mental Illness Bane, ‘buying it off.’ 


A Silver Guard may cause a physical injury to heal at double the normal speed by spending 1 
Divinity per 2 levels of the Wound. Frostbite and other cold-related injuries can be cured 
immediately at no Divinity cost. The Infection Chance of such a Wound is reduced by 4. 


Finally, for no Divinity cost, a Silver Guard can take away all negative thoughts, emotions, or 
painful memories away from an Adherent, by holding their head and willing it to be so. An 
ethereal white mist emerges from the Adherent's mouth and nose, and enters the Silver Guard. 
The Adherent will forget all of his negativity, while the Silver Guard will have to bear it himself. 
Silver Guards live with some of the most soul-crushingly horrific knowledge to begin with, the 
petty problems of a Human being, even a thousand Human beings, are so trivial that most Silver 
Guards wouldn't notice. 

After about a month, the memories taken away will start to return, but can be relieved again 
through another Succor. 

\paragraph{A Thousand Stars, One Moon}
A Silver Guard can cause their Adherents to become stronger by spending Divinity. This allows 


Silver Guards who are leaders in war to empower their armies--it also has domestic uses. Silver 
Guards can empower groups of farmers or laborers in a similar manner. 


Each point of Divinity spent has the following effects. The effect lasts for 1 hour, at which time 
the Silver Guard must pay the cost again to sustain the effect. 
\begin{itemize}
\item Increase one attribute by 1 point per point of Divinity spent. The Silver Guard may 
choose how these points are distributed. 
\item Increase Pain resistance by 1 per point of Divinity spent. 
\item Ignore 1 point of total \nameref{sec:Blood Loss} per point of Divinity spent. 
\item Complete immunity to cold or cold-related damage for the full duration. 
\end{itemize}
\subsection{Sarturi Chosen}
\begin{quotation}
	{\emph{“Raiders, pirates, reavers, thieves. These things are not natural. The are the result of inequality. An unnatural state, where the humors of the land are out of balance. The poor take from the wealthy because they have no choice. They have no choice because the world is broken. We can fix it. All we need to do is adjust the balance. Balance the scales. Return to Sartur that which is His, and He will redistribute it fairly. And what is Sartur’s, my brothers?”}

\hfill--\textbf{Coavva the Black}, Sarturi Chosen, \textit{‘The Battle of the Red River’}}
\end{quotation}

\subsubsection*{Overview} 
The Sarturi Chosen are the chosen servants of the Blood God, Sartur. Sartur is a little-understood deity in 
Vosca, and his followers are seen by most as little more than a depraved cult of demon-worshippers. The 
Chosen exemplify both this stereotype, and also the truth of their religion. Sartur is the God of Blood, and 
his religious tenets are essentially simple: if things are bad, it is because there is either too little, or too 
much blood in the land. Too much blood can be solved by shedding it into the ground, which is Sartur's 
domain. Too little blood can be solved by shedding it elsewhere, thus allowing Sartur to redistribute the 
blood to those more needy. The Sarturi believe that Vosca is suffering from a terrible imbalance of blood, 
hence the horrible conditions of the frozen north, and the decadence and wealth of the south. They believe 
that by killing enough people and animals in the south, they will be able to cure both the decadence and 
pettiness of the southern peoples, and the starvation and poverty of the northerners. 


From their perspective, they are perfect altruists. They just happen to go about their philanthropic duty by 
murdering hundreds or thousands of people with axes. 
\subsubsection*{Appearance} 
The Chosen are not changed in any way from their original forms before death. They look exactly like 
normal Humans. Of course, the huge majority of Sarturi Chosen were great warriors in life, and so tend to 
be robust and powerful stock. 
\subsubsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} None 
\subsubsection{Give unto Sartur\dots} 
The Chosen gains Divinity in the following ways: 
\begin{itemize}
\item Killing people. Divinity equal to the slain character's HLT score per Human killed, 
unless they were: pregnant women, children, or those sworn to peace or poverty. Killing 
such a person (or even being party to their death) instead costs the Chosen 10 Divinity. 
\item Killing domesticated animals. 1 Divinity per 100 lbs that the animal weighs (animals 
lighter than 100 lbs provide no Divinity.) 
\item Making people bleed in combat. Every time anyone within 50 feet of a Sarturi fails a 
\nameref{sec:Blood Loss} test from wounds inflicted by the Sarturi, the Sarturi gains Divinity equal to 
the number of HLT points lost. 
\item Being injured. A Sarturi who suffers any cutting or piercing wound in legitimate battle 
(practice or self-inflicted wounds don't count) gains Divinity equal to the Level of the 
Wound (this Divinity cannot be used to negate the Wound that provided it with the Not 
Today! ability). 
\item Suffering \nameref{sec:Blood Loss}. If a Sarturi fails a \nameref{sec:Blood Loss} test, he gains Divinity equal to 2x 
the HLT points lost. 
\end{itemize}
\paragraph{From the Depths to the Soil}
\begin{wraptable}{r}{7 cm}
	\caption{Area Affected by ``From the Depths to the Soil''}
	\label{tab:sarturflowerpower}
	\begin{tabu} spread 0pt {cX}
		Divinity&Area\\
		1&1 square yard\\
		2&10 square yards\\
		3&100 square yards\\
		4&1,000 square yards\\
		5&10,000 square yards\\
	\end{tabu}
\end{wraptable}
The Chosen have the ability to spend Divinity to enrich the soil and life of an area. This can cause 
the accelerated growth of plant life, the restoration of health to flora and fauna alike, and the 
elimination of pollution. At base, wherever a Chosen steps, dead plant matter quickly and visibly 
rejuvenates (if it's in season) and no animal kept in close proximity to a Chosen will ever grow 
sick. 


By spending Divinity, the Sarturi can cause a larger area to burst into health and life. Any crops 
planted in this land will grow to their potential no matter what else happens, no animals raised 
and grazed on this land will grow sick or die except through external influence. Even the weather 
will be milder and better for crops and pastoralism here, even if it is horrible all around the area. 
The effect of this ability lasts for 1 month per Divinity spent--so a 1 Divinity expenditure 
improves 1 square yard for 1 month. A 5 Divinity expenditure improves 10,000 square yards for 5 
months. Past 5 Divinity, the size of the area no longer increases, but the duration does. So 300 
Divinity will improve a 10,000 square yard area for 300 months, or about 25 years. 


Every time a member of a Sarturi Chosen's Band (see below) is killed, the Sarturi may activate this ability, centered on his dead companion, rather than himself, as though it had been activated 
with 1 Divinity. This has no cost.
\paragraph{A River of Blood}
A Chosen may spend 4 Divinity to increase his STR and HLT by 1 each for a span of 10 Turns 
(or about 1 minute) at will. Repeated uses of this ability can increase the Chosen's STR and HLT 
by an amount up to his WIL score. 


If inside an area under the effects of From the Depths to the Soil, the amount of STR and HLT 
gained through the use of this ability, and the maximum amount that can be gained, are both 
doubled.
\paragraph{Not Today!}
Sarturi Chosen are held together by more than muscle and sinew. When they sustain injuries, 
their flesh can be made to close immediately upon being sustained, and blood will visibly contract 
back into the body before the wound closes. The process is not even truly painful to them. 


Any time a Chosen would sustain a Wound, they may instead spend 2 Divinity per level of the 
Wound to negate it immediately. They still suffer half the Stun, but none of the Pain of the 
Wound, and of course no \nameref{sec:Blood Loss}. 


If inside an area under the effects of From the Depths to the Soil, this ability costs 1 Divinity per 
level of the Wound instead of 2. 

 

\paragraph{Band of Blood:}
Before a battle, any companion of the Chosen may don a bandana stained with the Chosen's blood. If they do so, they are considered part of that Chosen's ‘Band’ until the next time they remove the bandana. 


A member of the Band gains the following bonuses: 
\begin{itemize}
\item +1 bonus to all damage inflicted by weapons that inflict Cutting damage. 
\item Ignores the first 2 points of Pain suffered. 
\item Any time \nameref{sec:Blood Loss} is inflicted or suffered, increase the amount of \nameref{sec:Blood Loss} gained 
by 3. 
\end{itemize}
Whenever a member of a Sarturi Chosen's Band is killed, all remaining members of the Band 
gain 1 CP, up to a maximum of their WIL for 24 hours. The Chosen gains 2 CP up to a maximum 
of 2x his WIL for 24 hours, and 1 Divinity. 


If the Chosen is killed, every member of the Band will suffer 5 \nameref{sec:Blood Loss} per Turn until they 
take off the bandanas. 
\paragraph{Sartur Hungers!:}
Wounds just seem to want to bleed more when a Chosen is on the battlefield. To increase the 
Blood Loss of every Wound by 1, which every living thing within 10 yards of a Chosen suffers, 
the Chosen may spend 1 Divinity. This effect does not increase the \nameref{sec:Blood Loss} of the Sarturi 
himself, and it cannot affect a specific Wound more than once. 
\paragraph{No Hard Feelings:}
If a Sarturi Chosen is killed in battle, whoever struck the finishing blow gains a number of Faith 
Arc Points equal to the total Divinity the Sarturi has, plus the amount spent in the last 24 hours. 
\subsubsection*{Culture and Society} 
XXX 
\subsubsection*{Ecology} 
XXX 
\section{Star Vampires}
\subsection*{Overview} 
XXX
\subsubsection*{Appearance}
In mortal form, the Vampire looks most like a human, or perhaps a Din (it depends) with pale, 
bloodless skin. Contrary to popular belief, vampires are not albino -- they have melanin in their 
skin like any human, but the absence of blood causes those of them with less of it to appear 
ghastly pale. Those who know what they are looking for, however, can spot a darker skinned 
vampire by the lack of blood visible on certain more translucent parts of the skin, particularly the 
ears and cheeks.

Vampires seem to be able to craft their appearances before adopting them for the first time, so 
most are very handsome specimens, with few physical imperfections. There are some 
interesting exceptions to this -- several famous vampires have been horribly misshapen by 
design.
\subsection*{Racial Abilities and Characteristics} 
\paragraph{Attribute Modifiers:} A Star Vampire’s End and Health are automatically 6, at no cost to the player at Character Creation.

A Star Vampire’s TOU is $ \frac{\text{STR}+\text{WIP}}{2} $

Star Vampires may resemble humans outwardly, but on the inside they lack human organs. They instead have large bladders for storing blood, bones, and additional muscle. Vampires are far stronger than humans, and they cannot be injured in the same way, since they are animated by Dark Fire.

\subsubsection{Blood Requirements}
A Star Vampire's Attributes, Skills, and pools for supernatural abilities are normally halved. 
Spending a Blood Point will restore these to full dice for 24 hours. A Vampire may spend a point 
of Blood at any time, but it takes a full minute for their abilities to come into effect. This state is 
referred to as “Full Form.”

\paragraph{Red Feast}---\quad
Star Vampires can consume blood by biting the necks of their victims, or by drinking it out of 
containers, etc. This inflicts Bloodloss 20 on the victim, which remains so long as the Vampire 
drinks. For each point of HET lost (equivalent to 2 pints of blood) the Vampire gains one “Point” 
worth of Blood.

A Vampire can store 12 Points of blood (about 24 Pints) in its body for later 
usage. Once a character reaches 0 HET, they die, and the Vampire can glean no further blood 
from them.

\paragraph{Red Fountain}---\quad
A Star Vampire in Full Form may spend 2 Blood Points to attempt to force the blood of a victim 
to erupt from their body, and flow towards the Vampire, who absorbs it through the skin. The 
Vampire makes an opposed WIP test against the victim’s HET. If successful, the victim instantly 
suffers 5 Bloodloss per point the Vampire succeeded by, and the Vampire gains Blood Points 
for lost HET as normal. This may be maintained for multiple rounds so long as the Vampire is 
not interrupted at no cost. The blood loss only lasts until the Vampire stops feeding, or 
(obviously) until the victim perishes.

The victim of this ability suffers massive hallucinations during this process, and must make a 
WIP test at RS 4 to take any action while suffering Bloodloss from this ability. The hallucinations often take the form of terrible darkness, massive eyes staring out of the cosmos, and shrill,
all-consuming screaming (or is it laughter?) from a thousand inhuman throats. 

\paragraph{Blood-Charged Form}---\quad
A Star Vampire in Full Form may spend 2 additional points of Blood to gain a +2 bonus to all 
physical Attributes for one hour. During this time they take on a monstrous appearance as their 
skin draws back, exposing their fangs, elongating their fingers into talons, and giving them a 
hideous skeletal look. 

\paragraph{Dark Swarm}---\quad
A Star Vampire can summon a swarm of shadowy, immaterial forms out of the ether to swarm 
an opponent, blinding and confusing them. The shapes range from dark grasping hands to 
skulls, formless tendrils, and even alien, winged shapes when used offensively, they tend to 
take the forms of their target’s greatest fears and insecurities. This counts as a Blind Toss 
Maneuver, but with 1 TN. Performing this Maneuver  is only possible if the Vampire is in Full 
Form, but has no additional cost. 

\paragraph{Wisp}---\quad
A Star Vampire in Full Form may vanish and reappear an instant later into a cloud of dark 
smoke and spectral stars (or are they eyes?). The Vampire will be “gone” for about half of a 
second, and cannot see what is happening around itself until it reappears. The Vampire can 
reappear up to 5 yards from its current position, or may spend a Blood Point to appear up to 50 
yards away.

This can be done in Combat. It counts as an Outmaneuver at 2 TN.
A Vampire cannot perform this ability if it is in within eyesight of a mirror. (Not any reflective 
surface will suffice, it must be a silver and glass mirror.)

\paragraph{Spiderclimb}---\quad
A Vampire may walk, crawl or climb on any surface as though it were a spider or gecko, 
adhering to surfaces through unnatural means. 

\paragraph{Hypnotism}---\quad
Simply by speaking to an individual, a Vampire can cause them to act more favorably towards 
them, as though they were a friend. Vampires always count as having +4 Charisma. 
When in Full Form, a Vampire may cause an individual to take a Willpower Test at RS 2 or obey 
any simple or reasonable request under the effects of hypnosis. 

When in Full Form, a Vampire may spend a Point of blood to make an opposed Willpower Test 
against a target. If successful, the Target must obey the Vampire for so long as they retain eye 
contact. The orders can be foolish, even suicidal, and the Target will obey. 

If a character resists any form of a Vampire's hypnotism, they are immune to any repeat 
attempts by the Vampire for a year and a day. A Vampire has no way of knowing whether or not 
its Hypnotism was actually effective on the target except for how it reacts to orders. 

\paragraph{Inhuman Anatomy}---\quad
Vampires cannot suffer Pain\footnote{*Weapons made of Silver (this includes Silversteel) do inflict Pain on Vampires as normal, as certain Vampire Weaknesses.} from wounds. Stun is always halved for Vampires. They cannot 
suffer from Infection or Instant Death with four exceptions: Decapitation, destruction of the brain, 
piercing of the central blood sack of their body (located in the chest or the abdomen, choose at 
character creation) with a Wooden or Silver weapon, or through burning of the body until death. 
Vampires cannot suffer from sickness, poison, radiation, etc. 

Vampires can suffer Bloodloss, but Vampires do not lose HET as normal if they fail Bloodloss 
Rolls, they instead lose 1 Blood Point per HET that would be lost. Vampires cannot perish from 
Bloodloss, they can only run out of Blood Points. 

Vampires that are killed immediately crumble into piles of powdered carbon, resembling ash. 

\paragraph{Regeneration}---\quad
Vampires heal at a rate of 1 Wound Level per day, so long as they are in Full Form. Otherwise, 
they do not heal at all. 
If in Full Form, a Vampire may spend 2 Blood Points to reduce all wounds by 1 level. This 
process takes 1 Round, and the Vampire cannot move while healing in this way. 

\paragraph{Dark Heart}---\quad
A Vampire has 4 bonus Fire for use in Pyromancy so long as it is in Full Form. 
A Vampire in Full Form can transmute 1 Blood Point into 4 extra Fire at any time. This extra Fire 
lasts for one Round. 

All Vampires have +4 Lustre. 

All Vampires can perform Dark Pyromancy. 

\paragraph{Spurn the Earth, Consume the Stars}---\quad
A Vampire in desperate straits may return to its true form. The true form of a Star Vampire is a 
hideous,cancerous-looking mass of polyps, tubes, and coral-like bones. In its true form, every 
living thing within 150 meters of the Vampire suffers Bloodloss 15. Every 3 points of blood lost in 
this manner restore 1 Blood Point to the Star Vampire. The Star Vampire has a mobility of 12 in 
this form, and can move in any direction, but it cannot attack, communicate or perform any other 
action except for moving and drinking blood. 

Upon returning to its true form, a Vampire's only conscious goal can be to return to the Sky, 
smashing through the roof or simply flying straight upwards as quickly as possible. Once there, 
the Vampire will re-assimilate with its Dark Master, and either return to the world again, in mortal 
form once more, or will cease to exist as its purpose is served. 

\paragraph{Dominate Undead}---\quad
A Vampire can force other undead (and Minor Vampires) to obey its commands through sheer 
psychic force of will. When faced with an undead or minor vampire, a Vampire in Full Form may 
make a WIP test to dominate them, unless they are already under the control of another 
Vampire or a necromancer.
This WIP test is at base RS 2, but can increase if there are multiple undead, as detailed in the 
table below. 

Dominated Undead and Minor Vampires cannot perform any action without the order of their 
master (though they can defend themselves, and can be given instructions to react in certain 
ways or to improvise within limits by their master) and will obey their master utterly and without 
fear or hesitation. They are under the psychic control of the Vampire and can relay what they 
see and hear instantly to the Vampire at any distance, and the Vampire can order them to take 
actions at any distance with a thought. 

A Vampire can attempt to wrest control of undead from a rival Vampire or Necromancer with an 
opposed WIP test, with the currently controlling figure gaining a +2 bonus to its roll. Once 
control has been torn away, the rival cannot attempt to take it back until the next moonrise. 

\subsubsection{Vampiric Weaknesses}
Vampires have a number of weaknesses.
\begin{enumerate}
	\item They find the sight and proximity of religious symbols to be physically painful, akin to the 
		sensation of burning alive. A Vampire not in Full Form suffers 20 Pain if within eyesight of a 
		religious symbol. In Full Form, this is reduced to 10 Pain. A Vampire may spend 6 Blood Points 
		to cause any normal (man portable, nonmagical) religious symbol in eyesight of them to erupt 
		into flames and burn, melt, or disintegrate. This weakness only applies if the Vampire is aware 
		that the object is a religious symbol if the Vampire doesn't recognize it, it does not count. 
	\item They detest plants with very strong smells, as their bodies refuse to process vegetable 
		matter. Garlic is the stereotypical example, but onions, cabbage, and even mustard all work as 
		well. A vampire cannot digest Blood if the smell of vegetable matter is strong in the 
		surroundings presently. 
	\item A Vampire can be instantly killed if it suffers a level 5 wound to the chest with either a wooden 
		or silver spike. This includes arrows, crossbow bolts made of wood, swords or other weapons 
		made of Silversteel, etc. Silver bullets also work. 
	\item Holy water, oil, or similar chemicals burn Vampires like boiling oil or burning sand burns 
		humans (see burning rules for more details.) These wounds do not heal. Ever.
	\item Vampires always count as Enemies of Genosus for the purposes of Judgement. Unlike 
		Ghouls (see below) the fires are agonizingly painful for them. 
	\item Vampires exposed to direct sunlight when not in Full Form suffer Fire Damage equivalent to a 
		Bonfire to exposed areas. 
\end{enumerate}

\subsubsection{Infecting Mortals}
Star Vampires are born beyond the limits of the sky, and are not created out of human stock, as 
is commonly believed by mortals. However, they are capable of creating lesser versions of 
themselves out of mortal creatures, including humans, Din, and even (it is speculated) goblins. 

These creatures are often \emph{called} vampires by those who witness them feeding, but more strictly 
speaking would be Ghouls, as they lack many of the characteristic powers of the Vampire, and 
retain only its thirst for blood. Ghouls are created when a Vampire kills a mortal through 
consumption of their blood, and then wills them to live. The resulting creature has the same 
Immortal Anatomy, Vampiric Weaknesses and Blood-Charged Form of a Vampire (there is a 
period of vomiting up organs and otherwise hideous feverish transformation that lasts for two to 
three days) but lacks the powers of Hypnotism, Dark Swarm!, Wisp, Spiderwalk, and of course 
Spurn the Earth, Consume the Sky. 

Ghouls are more or less incapable of disobeying their vampiric masters, and most of them, 
twisted by their infection, wouldn’t want to to begin with. A Ghoul must make a Willpower Test at 
RS equal to their master’s WIP to even bend the wording of their master’s orders. Outright 
disobedience is all but impossible (RS equal to WIP+2) but has happened on rare occasions. If 
the Vampire a Ghoul serves is killed, or returns to the sky, they will lose 1 point of Strength, 
Endurance or Health (in that order) per day until all of those Attributes are reduced to 1 or less, 
at which point they will perish. If a Ghoul can find a new master before dying, they can establish 
a new relationship, and regain their lost attributes at a rate of 1 point per day until back to 
normal.

A Ghoul without a master may delay the loss of Attributes by 1 week by consuming 3-4 pounds 
of human (or Dinnic) flesh. This does not have to be fresh -- it can even be rotting. Ghouls are 
immune to the negative effects of consuming such ghastly fare, and it will stave off their decay, 
though it does not restore lost attributes. Many masterless ghouls form packs, roving between 
graveyards and small villages, searching for a new master while consuming the dead (or the 
living!) to sustain themselves. 

Ghouls always count as Enemies of Genosus for the purposes of a Paladin’s Judgement ability, 
but while the flames do consume their flesh, the sensation is not one of agony, but of ecstasy 
and comfort. This has the same effects as Pain for the purposes of disrupting the Ghoul’s 
combat abilities, but it is commonly believed by Paladins that Judging a Ghoul is a mercy that 
can save it from damnation, hence the pleasurable salvation. Some Ghouls and Vampires 
suspect that this is actually a clever trap devised by Genosus to make hunting Ghouls easier for 
his servants.

\chapter{Boons and Banes}\label{ch:boons}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Boons and Banes are qualities and quirks that affect characters in small but enduring ways. Boons are 
positive characteristics, such as ambidexterity, a knack for languages, or a particular aptitude for 
climbing. Banes, however, are negative characteristics, and include traits such as haemophilia, a prominent speech impediment, or a total inaptitude for swimming. 

Boons and Banes are predominantly chosen at Character Creation, but during play characters can buy new 
Boons, buy off Banes or encounter situations that cause a character to gain one! 
\section{Character Creation}
Every character must invest at least 1 PCP in Boons and Banes. There are advantages to spending more 
than this amount, however. Generally speaking, the more PCP you invest in Boons and Banes, the more 
Boons you are able to take. 

\autoref{tabwrap:Boons and Banes CharCreation} details the number of B\&B Points your PCP investment will provide you. 

\begin{wraptable}{r}{5cm}
	\caption{Boons and Banes Costs at Character Creation}
	\label{tabwrap:Boons and Banes CharCreation}
	\begin{tabu}{X[c]X[c]}
		PCP&B\&B Points\\
		1&-15\\
		2&-10\\
		3&-5\\
		4&0\\
		5&5\\
		6&10\\
		7&15\\
		8&20\\
		9&25\\
		10&30\\
	\end{tabu}
\end{wraptable}
Characters must start play at or above a zero (0) B\&B Point score, and cannot start play with a negative 
B\&B Point score. As an investment of 1 PCP means you have -15 B\&B Points, this clearly poses a 
problem. 


Taking Banes positively impacts your character's B\&B Point score, with limited effect. After investing 
PCP into Boons and Banes, characters at a deficit of B\&B Points must take Banes in order to buy off the 
debt. Once you have reached an even zero, the deficit is gone and there is no need to take more Banes. 


Taking Boons negatively impacts your character's B\&B Point score, and so it is necessary to have a 
surplus of B\&B Points to take Boons. If you have invested enough PCP into Boons and Banes such that 
you have a surplus of B\&B Points, then you may take as many Boons as you wish, so long as your B\&B 
Points remain at or above zero. 


Regardless of your PCP investment, however, \emph{you may take only up to 15 additional B\&B Points worth 
of Banes, after at least breaking even with zero B\&B Points}. With these additional points, you are allowed 
to purchase a proportionate amount of Boons, if desired. You may continue to take Banes after these 
additional 15 B\&B Points, however these Banes do not continue to positively impact your Character’s 
B\&B Point score. 


\subparagraph{1-3 PCP Investment}
Investing 2 PCP will grant you -10 B\&B Points, and taking the moderate Enemies Bane (at a value of 10 B\&B Points) will 
ensure an even zero B\&B Point score. At this point, you can begin play (with only one Bane), or opt to use up to 15 additional 
B\&B Points' worth of Banes, allowing you to also take a number of Boons that leaves your B\&B Point score at or above zero.

\subparagraph{4-10 PCP Investment}
Investing 8 PCP will grant you 20 B\&B Points, and taking the Prodigious Wit and Prodigious Strength Boons (at a value of 10 
B\&B Points each) will ensure an even zero B\&B Point score. At this point, you can begin play (with two Boons), or opt to use up 
to 15 additional B\&B Points’ worth of Banes, allowing you to also take a number of Boons that leaves your B\&B Point score at or above zero. 
\section{Advancing with Arc}
During play it is possible to add Boons by spending Arc Points (except those marked as ‘Character 
Creation only’). Any Boons bought must have an in-game reason, for example, the Allies Boon cannot be 
bought unless the character does acquire powerful allies in game—allies cannot just spawn out of thin air. 


It is typically difficult to lose Boons, however at the GM’s discretion, certain events and situations may 
render Boons ineffective. An example may be when a character has the Boon ‘Beautiful,’ but during play, 
becomes severely disfigured and scarred. There is no cost involved in losing a Boon. 


Most Banes are able to be removed by spending Arc Points (except those marked as ‘Cannot be 
removed,’ which are permanent). 


Banes may be acquired during the course of play, through battle and other game developments. There is 
no cost involved in gaining a Bane, except those suffered through Wounds and unfortunate circumstance. 
An example may be gaining the ‘Wanted’ Bane by committing deeds that offend those in power. 


A single Boon or Bane may typically be taken once, and many have multiple levels of purchase. Those 
marked with a * can be taken multiple times. For example, you can take Languages (1) multiple times to 
represent knowing multiple native languages, or you could take Old Wound multiple times to have Old 
Wounds on different parts of your body. Boons and Banes cannot stack on themselves: you cannot take 
Old Wound multiple times on the exact same part of your body, or take Prodigious Strength twice to raise 
your character creation attribute caps by 4. 


Unless specified otherwise, a higher level of a Boon or Bane replaces the effects of its lower levels.
\clearpage
\section{Boons List}
	\begin{longtabu} spread 0pt {XXX[-1,r]}
			\caption{List of Boons}
			\label{tab:boons list}\\
		\rowfont[c]{}Boon& Description&B\&B Point / Arc Point Cost\\\toprule
		\nameref{boon:Allies}*&&1/5/10\\
		\nameref{boon:Ambidextrous}& Character Creation only& 3\\
		\nameref{boon:Animal Affinity}&&2/4/6\\
		\nameref{boon:Beautiful}&&3/6\\
		\nameref{boon:Berserker}&& 6/12\\
		\nameref{boon:Bloodthirsty}& &4\\
		\nameref{boon:Brave}&&3\\
		\nameref{boon:Contacts}*& &1/4/6/8\\
		\nameref{boon:Direction Sense} &&3\\
		\nameref{boon:Estate}* &Character Creation only &10\\
		\nameref{boon:Favor}*&&1-3\\
		\nameref{boon:Famous}&&2-4\\
		\nameref{boon:Folks Back Home}&&3/6/8\\
		\nameref{boon:Follower}*& &15\\
		\nameref{boon:Good Ears}&& 3\\
		\nameref{boon:Good Eyes}& &3\\
		\nameref{boon:Good Nose}&& 3\\
		\nameref{boon:Hale and Hearty}&& 2/4\\
		\nameref{boon:Impressive Voice}* &&2\\
		\nameref{boon:Known for Virtue}& &5\\
		\nameref{boon:Languages}* &&1/2/3\\
		\nameref{boon:Literate}*&& 1\\
		\nameref{boon:Natural Born Killer}& &6/12/18\\
		\nameref{boon:Natural Leader}&&3\\
		\nameref{boon:Retinue}&&5/10/15\\
		\nameref{boon:Tireless Lobster}&Fatigue Rules Only& 6 \\
		\nameref{boon:Tall} &Character Creation only &8/12\\
		\nameref{boon:True Grit} & Character Creation only &2/4/6\\
		\nameref{boon:Rich}& Character Creation only &1/3/5\\
\end{longtabu}
\section{Banes List}
	\begin{longtabu} spread 0pt {X[1.5] X X[-2,r]}
		\caption{List of Banes}
		\label{tab:Banes List}\\
\rowfont[c]{}	Bane&Description&B\&B Point / Arc Point Cost\\\toprule
\nameref{bane:Arrow Magnet}	&&3\\
\nameref{bane:Bad Ears} 		&Cannot be removed &2/4\\
\nameref{bane:Bad Eyes} 		&Cannot be removed &4/6\\
\nameref{bane:Bad Reputation} &&3/6/9\\
\nameref{bane:Barren/Sterility} &Cannot be removed &1/3\\
\nameref{bane:Blind} &Cannot be removed&20\\
\nameref{bane:Braggart} &&3\\
\nameref{bane:Brain Damage}* &Cannot be removed &4/8\\
\nameref{bane:Broken Limb/Appendage}* &Cannot be removed& -\\
\nameref{bane:Complete Monster} &&10\\
\nameref{bane:Craven} &&7/15\\
\nameref{bane:Crippled Limb/Appendage}* &Cannot be removed &8\\
\nameref{bane:Dead} &Cannot be removed &100\\
\nameref{bane:Debt} &&2/4/8\\
\nameref{bane:Dependent} &Cannot be removed &6/12\\
\nameref{bane:Dire Past} &Cannot be removed &0\\
\nameref{bane:Enemies} &&3/10/15\\
\nameref{bane:Facial Deformity} &Cannot be removed &2/4/8\\
\nameref{bane:Fat} &&5\\
\nameref{bane:Haemophilia} &Cannot be removed &7\\
\nameref{bane:Hothead} &&3\\
\nameref{bane:Honorable} &&15\\
\nameref{bane:Lasting Pain} &Cannot be removed &4/8\\
\nameref{bane:Mute} &Cannot be removed &5/8\\
\nameref{bane:Oath} &&2 to 10\\
\nameref{bane:Old Wound} &Cannot be removed &1\\
\nameref{bane:One-Eyed} &Cannot be removed &10\\
\nameref{bane:Poor} &&4/6/8\\
\nameref{bane:Severed Limb/Appendage}* &Cannot be removed &15\\
\nameref{bane:Sheltered} &&2/4/6\\
\nameref{bane:Short} &&8/15\\
\nameref{bane:Skinny} &&3\\
\nameref{bane:Technologically Impaired} &&5\\
\nameref{bane:Unhappily Married} &&1/2/3\\
\nameref{bane:Virtuous} &&10\\
\nameref{bane:Wanted} &&5/10/15\\
\end{longtabu}
\clearpage
\section{Boons}
\paragraph{\label{boon:Allies}Allies}---\quad\textbf{(1/5/10)}\par
The Allies Boon represents a character's positive relationship with a powerful figure, organization, or 
group. At its lowest level, it could simply represent someone important or competent around town. At its 
highest, it should represent the attention and favor of kings. This Boon can be taken multiple times to gain 
multiple allies, and can be gained (or lost) in the course of play through player actions. 


A character can appeal to one of their Allies for aid, but that is no guarantee that help will come, or that it 
will be sufficient for the task at hand. More often, Allies will be able to tie up loose ends, or swoop in to 
save a character's bacon when things get too hairy. 
\begin{itemize}
\item 1: Minor local power. Crime boss, mayor, town sheriff, and so on. 
\item 5: Moderate regional power. Baron, trade tycoon, colonel in the army, and so on. 
\item 10: Major international power. King, powerful cardinal, the Pope, and so on. 
\end{itemize}
\paragraph{\label{boon:Ambidextrous}Ambidextrous}---\quad \textbf{(3)}\par
\textit{Character Creation only}\par
Ambidextrous characters are equally capable of using either hand well, and are not affected negatively by 
using a weapon or tool in their ‘off’ hand. Ambidextrous characters gain certain benefits when using two 
weapons, as detailed in the relevant Maneuvers and Proficiencies. In addition, if a character's arm is 
injured and they are forced to use their weapon in their ‘off’ hand, an Ambidextrous character suffers no 
penalties. 
\paragraph{\label{boon:Animal Affinity}Animal Affinity}\!---\quad \textbf{(2/4/6)}\par
A character with this ability has an unusual affinity for animals. 
\begin{itemize}
	\item 2: Gain a +2 bonus to CHA when interacting with animals, and to Ride checks. 
	\item 4: As above, and may select one type of domestic animal (i.e. dogs, cats, horses). These animals will always cooperate with you, within reason. 
	\item 6: As above, and you may select one type of wild animal (i.e. wolves, cougars, bears). These animals will never attack you unprovoked, and may even come to your aid in certain situations. 
\end{itemize}

\paragraph{\label{boon:Beautiful}Beautiful}---\quad \textbf{(3/6)}\par
You're pretty. Really pretty. A character with this Boon is easy on the eyes. This can take many forms, 
and while different cultures have greatly different standards of beauty, for the most part, this is a universal 
Boon. Beautiful people are generally treated better, as they make a good first impression. 

\begin{itemize}
\item [2:] You gain a +2 bonus to CHA when dealing with everyone, if your beauty is apparent. You gain a +4 bonus to CHA when dealing with characters who would find you sexually attractive.
\item [4:] You cause carriage crashes. You gain a +3 bonus to CHA when dealing with everyone, if your beauty is apparent. You gain a +6 bonus to CHA when dealing with characters who would find you sexually attractive. Beauty of this caliber often comes with unwanted attention.
\end{itemize}
\paragraph{\label{boon:Berserker}Berserker}---\quad \textbf{(6/12)}\par
You are prone to flying into uncontrollable spells of bloodthirsty, shield-gnawing lunacy, during which 
you cannot be prevented from cleaving every head within arm's reach from its shoulders. 


Berserk rages last for 3d10 Rounds. To come down from a berserk rage early, you must initiate a WIL 
test at RS 5. Success ends the Berserk Rage. This can be attempted once per Round. 


How you enter a berserk rage varies depending on the level of this Boon, but the effects of going berserk are generally the same. 
\begin{itemize}
\item You reduce all Pain by half. 
\item You ignore ALL Stun inflicted by Wounds, but not those inflicted by weapon effects. 
\item You ignore all effects of Exhaustion. 
\item You reduce the RS of Bleed tests by 2. 
\item You may not take any sort of Parrying or Blocking Action (you may still Void). 
\item You must attack the nearest active enemy, or, barring the presence of enemies, the nearest living 
creature. 
\end{itemize}
After a character comes down from a berserk rage, all effects end, and they are immediately Exhausted, 
regardless of their END or current Encumbrance. If a Berserker's \nameref{sec:Blood Loss} total is above 25 at this 
point, they must make a HLT check at RS 5 or drop dead instantly. 


Characters that go into a berserk rage often may develop mental derangements over time, including 
depression, alcoholism, and schizophrenia. 
\begin{itemize}
\item 6: Any frustration or irritation, physical pain, insult, or vexation requires that the character make a WIL check at RS 4 or fly into a berserk rage. He can voluntarily fly into such a rage at will, though only when not Exhausted, by taking a few seconds (one Round) to roar, gnaw upon their shield, and generally work 
themselves into a fury.
\item 12: As above, however involuntarily entering a berserk rage requires a WIL check at RS 2. The character can voluntarily enter a berserk rage at will, though only when not Exhausted. In the case that a character enters a berserk rage as a result of an attack, any effects the Wound causes happen before the rage is entered.
\end{itemize}
\paragraph{\label{boon:Bloodthirsty}Bloodthirsty}---\quad \textbf{ (4)}\par
Your eagerness to join battle has stripped from you any form of hesitation or restraint when moving in for 
the kill. Whenever you take an Initiative test brought on by going Red/Red (and only Red/Red – tests forced by maneuvers do NOT count) you gain +2 dice for that test. There is no limit to the number of 
times this can occur in combat.


Any character who faces you in combat will immediately become aware that you have this Boon unless you pass a Subterfuge vs PER contest. 
\paragraph{\label{boon:Brave}Brave}---\quad\textbf{(3)}\par
You're courageous and firm of heart. Whenever you would need to make a WIL test vs Fear, you may 
choose instead to automatically pass the test. This does not apply to any Fear tests involving magic, however you gain +2 dice to your roll.
\paragraph{\label{boon:Contacts}Contacts}---\quad\textbf{(1/4/6/8)}\par
You have access to a network of informants, friends of friends, cousins, and entire flocks of talkative little birds that you can call upon for information.


You can use Contacts to gain insight and intelligence on any subject you please. Usually this requires the expenditure of a few coins (rarely more than 2-3 sp per RS of the roll, see below), and may take anywhere from ten minutes to a week or so. 


Roll your CHA at the TN decided by your level in this Boon, and if you meet the RS, you gain the 
information you need, with more BS supplying more information. The GM can set the RS from 1 
(effortless) to 8 (world-shatteringly obscure). If there is absolutely no way that information can be gained 
through Contacts, it should be evident after the first roll. 

\begin{itemize}
\item 1: TN 7: You know some guys. 
\item 4: TN 6: You know some guys who know some guys. 
\item 6: TN 5: You know some guys in every city and service in the country.
\item 8: TN 4: You know everyone.
\end{itemize}
\paragraph{\label{boon:Direction Sense}Direction Sense}---\quad\textbf{(3)}\par
You have an infallible sense of direction. Wherever you are, you can figure out which way is north. Your skill in this regard borders on the supernatural, and grants you a +4 to Navigation tests.
\paragraph{\label{boon:Estate}Estate}---\quad\textbf{(10)}\\
\textit{Character Creation only}\par
You own lands, businesses, or other static wealth that provides you with a steady stream of income. 
Taking this Boon grants you an Asset worth 2 Wealth (see the Assets section of Chapter 5 for more 
details). This Boon can be taken multiple times, either to gain multiple Assets (each valued at 2 Wealth), 
or to combine them into a bigger, badder Asset. 


You cannot acquire Assets with a Wealth value greater than the amount your Social Class granted you. 
That is, if you are a Lesser Noble, and start the game with 6 Wealth, you could take Estate a maximum of 
three times for 6 additional Wealth, for a total of 12, but no more. 
\paragraph{\label{boon:Favor}Favor}---\quad\textbf{(1/3)}\par
Someone owes you a favor. You may take this Boon multiple times, and for each time you do, you gain 
an additional favor that you can ‘call in’ to get something done. Confer with the GM to see who could 
potentially owe your character one. Once a Favor is called in, the debt is settled. However, if things are 
still amicable, there's no reason why another exchange of favors couldn't be arranged. 
\begin{itemize}
\item 1: Minor Favor. A serious favor from a regular man, or a trifling favor from someone of power (‘let me hide in your house from the Baron's men,’ or ‘get me off of this larceny charge’).
\item 3: Major Favor. A serious favor from a powerful individual (‘vouch for my innocence, your majesty’).
\end{itemize}

\paragraph{\label{boon:Famous}Famous}---\quad\textbf{(2/4)}\par
People know you. Maybe you killed the Manticore of Marienburg, or saved a princess or something. 
Whatever the reason, you gain a bonus to CHA when dealing with people overawed by your star-power. 
This Boon has its drawbacks though—being recognizable makes it hard to hide, and harder still to get 
away with mischief.
\begin{itemize}
\item 2: +2 bonus to CHA, -2 to Disguise or Bluff rolls to avoid recognition. 
\item 4: +4 bonus to CHA, -4 to Disguise or Bluff rolls to avoid recognition. 
\end{itemize}
\paragraph{\label{boon:Folks Back Home}Folks Back Home}---\quad\textbf{(3/6/8)}\par
You've got a family, a clan, a tribe or some other social group that you're part of that cares for you. You're 
probably away from them right now, but they're there, waiting for your return. The level of this Boon 
determines just how influential and wealthy these folks are, and how much they might be able to help in 
the event that you go to them for aid. Not having this Boon doesn't necessarily mean your character is an 
orphan or without family, just that the family is scattered, disunited, or otherwise unlikely to be able to 
provide cohesive aid. 
\begin{itemize}
\item 3: Nobody important, but they’re there. 
\item 6: Well to do, and somewhat influential. 
\item 8: Powerful, important people.
\end{itemize}
\paragraph{\label{boon:Follower}Follower}---\quad\textbf{(10)}\par
You've got a sidekick, a buddy, a comrade, a henchman, or a right-hand-man. Followers can be anything 
from unusually loyal mercenaries, to childhood friends, and spouses who aren't about to stand by while 
your character fights heretics, bandits and shoggoths without them. 

 

A Follower is distinguished from Allies or a Retinue member in that the Follower is an actual character. 
Either you, or your GM, can write up the stats for the Follower (at the GM's discretion), usually of a 
power level the same as or lower than their PC leader. In some situations, an existing NPC can be drafted 
as a Follower, at the GM's discretion. 

 

Followers aren't directly under the player's control, in that they are still technically NPCs under the GM's 
purview, but there's nothing wrong with a player controlling a Follower while his character is injured or 
tied to an interrogator's chair somewhere. 

 

If the current PC of the Follower dies, the Follower ceases to be a Follower. At the GM's discretion, the 
Follower may stay of their own volition, leave, or even become a PC under the control of the player 
whose character died. 

 

Some guidelines for this Boon: it is not advised to let Followers take Follower, Retinue, or Contacts. It 
just makes things confusing. As well as this, don't let this Boon get out of control, and swell the group to 
double its original number. It is acceptable to limit characters taking this Boon, or of obtaining Followers 
altogether, see Optional Rules for more details. See the Followers and Retinue section in Optional 
Rules for more details. 

 

\paragraph{\label{boon:Good Ears}Good Ears}---\quad\textbf{ (3) }\par
You've got the ears of a fox. When making a PER check involving your auditory senses, you gain a bonus 
of +2 to the roll. 


A character cannot have Good Ears at the same time as the Bad Ears Bane. 


\paragraph{\label{boon:Good Eyes}Good Eyes}---\quad\textbf{(3) }\par
You've got the eyes of a hawk. When making a PER check involving sight, you gain a bonus of +2 to the roll. 


A character cannot have Good Eyes at the same time as the Bad Eyes Bane. 


\paragraph{\label{boon:Good Nose}Good Nose}---\quad\textbf{ (3) }\par
The nose knows. You've got an exceptionally good sense of smell. Whenever making a PER check of any 
sort, you gain a bonus of +1 to the roll. This does not work underwater, or in situations where an 
overpowering smell would mask everything around you. If you lose your nose, you lose this Boon. 

 

\paragraph{\label{boon:Hale and Hearty}Hale and Hearty}---\quad\textbf{ (2/4) }\par
You're really healthy, and you rarely become sick or suffer from poorly-healed wounds. 

\begin{itemize}
\item 2: You reduce the Infection Chance of injuries you suffer by 1. Your HLT for the purposes of resisting 
poison, disease and other ill effects (not Infection) is increased by 2. 
\item 4: You reduce the Infection Chance of injuries you suffer by 2. Your HLT for the purposes of resisting 
poison, disease and other ill effects (not Infection) is increased by 4. 
\end{itemize}
\paragraph{\label{boon:Impressive Voice}Impressive Voice}---\quad\textbf{(2)}\par
You've got a characteristic tone to your voice that lets you influence people in certain ways. This comes 
in three varieties: Powerful, Grating, and Soothing. You can take this Boon up to three times to gain other 
voice qualities. 

\begin{itemize}
\item Powerful: Your voice is strong, commanding, and intimidating. You gain a +2 bonus to Leadership and 
Intimidation. 
\item Grating: You have a sneering honk of a voice that can peel the skin straight from someone's bones with a 
single insult. You gain a +2 bonus to Ridicule rolls, and a +2 bonus to resisting them as well. 
\item Soothing: You have a pleasant voice, that can calm the heart and soothe the soul. You gain a +2 bonus to 
Diplomacy rolls with the intent to sooth, calm, or convince. You also gain a +4 bonus to any Perform 
tests that involve the voice (song, drama, and so on). 
\end{itemize}
\paragraph{\label{boon:Known for Virtue}Known for Virtue}---\quad\textbf{(5)}\par
You are known as a paragon of virtue, a protector of the weak, a champion of justice, a practitioner of 
mercy, or a truly honorable man. Whether this is true is not important. Being Known for Virtue gives you 
a degree of moral authority over the virtuous. People will assume that you are working for the best, and 
when dealing with principled individuals, their favorable view of you grants you a +2 bonus to CHA for 
all social rolls, and these individuals will also be much more inclined to give you the benefit of the doubt 
if you're caught in a sticky situation. 


If you ever gravely dishonor yourself or otherwise compromise your perceived moral high ground, you 
may lose this Boon, at the GM's discretion. For particularly egregious acts, you may even gain the Bad 
Reputation Bane. 



\paragraph{\label{boon:Languages}Languages}---\quad\textbf{(1/2/3) }\par
You know how to speak languages other than your native tongue. 
\begin{itemize}
\item 1: You know another native language. You may purchase this Boon multiple times during Character 
Creation, each time learning a new language. 
\item 2: You have a knack for languages. You know another native language, and you may purchase additional 
languages after Character Creation for 1 Arc Point each, assuming you have a month or so to practice. 
\item 3: You're a proper polyglot. You know a number of new languages equal to your INT+2. In addition, you 
automatically learn any language you are exposed to for more than a week by spending 1 Arc Point. 
\end{itemize}
\paragraph{\label{boon:Literate}Literate}---\quad\textbf{(1) }\par
You know how to write using your own language. If you can speak another language, and it has the same 
alphabet or writing system, then you are likely able to understand and write in that language as well. This 
Boon can be taken multiple times to learn unfamiliar writing systems. 


\paragraph{\label{boon:Natural Born Killer}Natural Born Killer}---\quad\textbf{(6/12/18) }\par
You were born to kill. Maybe you just have an exceptional, intuitive understanding of violence, or maybe 
you just lack inhibitions that normal people have against striking hard, fast, and first. Either way, your 
killer instincts give you an edge over other fighters. 

 
\begin{itemize}
\item 6: You gain +1 to your CP. 
\item 12: You gain +2 to your CP. 
\item 18: You gain +3 to your CP. 
\end{itemize}
\paragraph{\label{boon:Natural Leader}Natural Leader}---\quad\textbf{(3) }\par
Some people just have a knack for leadership. You're confident when dealing with other people, and your 
force of personality makes it easy for you to organize and lead groups of people. When making CHA-
based rolls to lead others, either in a military setting or for some other purpose, you gain +2 to your score. 
\paragraph{\label{boon:Retinue}Retinue}---\quad\textbf{(5/10/15) }\par
(Note: You may only take this boon if your campaign is using the optional Followers and Retinue rules) 
You're a leader of men, and have a group of loyal followers who aid you in your endeavors. When you 
take this Boon, you gain a group of NPCs who act as your personal bodyguards, aids, or crew. The size of 
this group varies on the version of this Boon you buy. See the Followers and Retinue section in 
Optional Rules for more details. 
\begin{itemize}
\item 5: You have a group of 10 Minions. 
\item 10: You have a group of 20 Minions and 2 Henchmen. 
\item 15: You have a group of 30 Minions and 3 Henchmen. 
\end{itemize}
\paragraph{\label{boon:Tall}Tall}---\quad\textbf{(8/12) }\\
\textit{Character Creation only}\par
You're a pretty big guy. For you, this means that you have an advantage in reach and stature over most 
people, which is especially useful in combat. 


A character cannot have Tall at the same time as the Short Bane. 
\begin{itemize}
\item 8: You’re noticeably taller than most other people, around a head or so. 
You gain +1 to Initiative as you have a slightly longer reach and better vantage over your peers. You gain a +1 bonus to Charisma for the purposes of persuasion, intimidation and leadership.
\item 12: You dwarf your peers, to the point you have to squeeze yourself through normal doorways. You gain +1 to the effective Reach of all weapons and unarmed Maneuvers you perform, and you gain a +2 bonus to Charisma for the purposes of persuasion, intimidation and leadership.
\end{itemize}
\paragraph{\label{boon:True Grit}True Grit}---\quad\textbf{(2/4/6) }\par
Character Creation only. 
Whether by hard experience, early exposure to a grim, violent world, or some natural strength of 
character, you have an amazing resistance to fear, pain, and shock. Each level of True Grit awards you an 
additional starting point of Grit. 


A character cannot have True Grit at the same time as the Sheltered Bane. 
\begin{itemize}
\item 2: Will of Iron: +1 Grit 
\item 4: Nerves of Steel: +2 Grit 
\item 6: Heart of Stone: +3 Grit 
\end{itemize}
\paragraph{\label{boon:Rich}Rich}---\quad\textbf{(1/3/5) }\\
\textit{Character Creation only}\par
Money, dinero, shekels, cash, bank, the sinews of war, the wealth of nations! You've got an unusual 
amount of money for someone of your Social Class. This is cold cash, not Assets, and once you've spent 
it, it's gone. 
\begin{itemize}
\item 1: Extra 10\% of your base money 
\item 3: Extra 50\% of your base money. 
\item 5: Extra 100\% of your base money. 
\end{itemize}
\paragraph{\label{boon:Tireless Lobster}Tireless Lobster}---\quad\textbf{(6)}\\
\textit{Only for use with Optional Fatigue Rules}
You’ve gotten so used to wearing armor constantly that it’s basically a second skin for you, and doesn’t tire you out meaningfully anymore. Wearing armor for more than 4 hours no longer causes its Weight to increase by half, and sleeping in armor only increases its effective Weight by half, instead of doubling it.

\section{Banes}
\paragraph{\label{bane:Arrow Magnet}Arrow Magnet}---\quad\textbf{ (3)}\par
You've got a way with arrows. They just can't stay away from you. In any situation where someone would 
be targeted at random by missiles, you will be targeted first, and anyone making missile attacks gain a +1 dice bonus against you. 
\paragraph{\label{bane:Bad Ears}Bad Ears}---\quad\textbf{ (2/4)}\\
\textit{Cannot be removed}\par
You just don't have good hearing. Maybe you worked on a howitzer range, or maybe it's congenital. 
\begin{itemize}
	\item [2:] You suffer a -2 to all PER checks involving hearing. 
	\item [4:] You're stone deaf. You suffer a -2 to all PER checks involving hearing, and if a check requires hearing, 
you can't roll at all. You can only understand people if you can see their mouths, or through sign 
language. 
\end{itemize}
A character cannot have the Bad Ears Bane at the same time as the Good Ears Boon. 
\paragraph{\label{bane:Bad Eyes}Bad Eyes}---\quad\textbf{(4/6)}\\
\textit{Cannot be removed}\par
You don't see too well. Corrective eyeglasses may be available in your time period, but if not... well, 
maybe you shouldn't be the one carrying the blunderbuss. 
\begin{itemize}
\item [4:] You suffer a -2 to PER if not wearing glasses. 
\item [6:] You suffer a -4 to PER if not wearing glasses, and a -2 to PER even if you are. 
\end{itemize}
A character cannot have the Bad Eyes Bane at the same time as the Good Eyes Boon. 
\paragraph{\label{bane:Bad Reputation}Bad Reputation}---\quad\textbf{(3/6/9) }\par
You aren't a popular fellow. Perhaps you're a member of an unsavory group that people generally 
mistrust, or maybe your personal deeds (real or attributed) have made you a pariah. Either way, folks are less likely to cooperate with you and may even try to hurt you if they take particular offense to you! Of 
course there will often be people who don't care about your reputation, and some may even approve! 


In certain situations, the GM may choose to waive your penalty, or even turn it into a bonus. While 
intimidating some yokels, for example. 

\begin{itemize}
	\item [3:] You're unpopular. You suffer a -2 to all social rolls made with people who disapprove of your 
	reputation. 
	\item [6:] You're generally disliked. You suffer a -4 to all social rolls made with people who disapprove of your 
	reputation. 
	\item [9:] You're nearly universally despised. You suffer a -6 to all social rolls made with people who disapprove 
of your reputation. 
\end{itemize}
\paragraph{\label{bane:Barren/Sterility}Barren/Sterility}---\quad\textbf{(1/3)}\\
{Cannot be removed}\par
You are incapable of producing offspring. Perhaps it is genetic, or perhaps you are a eunuch… or perhaps 
you had an unfortunate encounter with a low-flying halberd. This Bane is not unique to either gender. 
\begin{itemize}
\item [1:] Barren/Sterile. You’re simply sterile. No children for you (or no more children, if you already have them). You may still feel sexual urges, depending on the nature of your condition.
\item [3:] Eunuch. You were made a eunuch before puberty (male only). This is a different situation from the above, as this has a serious effect on the development of young men. In addition to all of the obvious effects of the process and the sterility associated with it, you must also pay 2 additional Arc to level up STR or HLT. However, as a side effect, you gain a +2 bonus against all manner of Social rolls made with the intent to deceive, and a +2 bonus against Intimidation and Taunt attempts.
\end{itemize}
\paragraph{\label{bane:Blind}Blind}---\quad\textbf{(20)}\\
\emph{Cannot be removed}\par
You're stone fucking blind. This brings with it certain problems. While blind men have been known to fight, it is by no means easy, and few would recommend it. 

Blind Characters cannot make sight-based PER checks, nor can they do anything that strictly requires the 
use of their eyes. However, they do gain a +2 bonus to PER checks involving sound, touch, and other 
senses to which they are naturally better attuned, due to their lack of sight. In combat, a Blind Character 
must make a PER check as part of declaring a Maneuver. The number of Successes on that check is the 
maximum number of CP or MP he can dedicate to that Maneuver. (This does not include Activation 
Costs, which are paid normally.) 
This check uses the other four senses, and benefits from the +2 bonus to PER checks mentioned above. 
Blind Characters treat all light levels as if they were Evenly Lit. Pages 202-203 
A character cannot have Blind at the same time as the Good Eyes Boon or One-Eyed Bane. 
\paragraph{\label{bane:Braggart}Braggart}---\quad\textbf{(3)}\par
You've got a big mouth. You can't help but boast of your strength, intelligence, or achievements, and you 
also can't stand the notion of being shown up. Maybe your claims are baseless... maybe they aren't. Either 
way, you get yourself in trouble a lot, and you're incredibly easy to rile up. 


You suffer a -2 WIL penalty to resist Ridicule rolls, and you must regularly relate your greatness to 
anyone who will listen. You can suppress your boastful urges for a few minutes when it's important by 
making a WIL test at RS 2. 

 

\paragraph{\label{bane:Brain Damage}Brain Damage}\quad\\
\textit{Cannot be removed}\par
You've suffered an injury that has significantly damaged your brain. You suffer from decreased mental 
faculties, and possibly some neurological failures (tics, partial paralysis of limbs, and so on). As stated in \autoref{ch:injuryrecovery}, whenever you are knocked out or pass out from \nameref{sec:Blood Loss} or excessive Pain, you must pass a TOU vs. 2 RS test or you gain this Bane. Certain wounds may also give you this Bane. 


When you gain Brain Damage, roll 1d10 on the \nameref{tab:Brain Damage} and immediately suffer its effects.\\
Old Wounds typically develop on the head, neck or around the spine but it’s possible that they may affect hands or joints (GM’s choice). If you would gain a Bane you already have, instead increase the severity of that Bane, if you can. Otherwise ignore that particular effect.

For each instance of Brain Damage you currently suffer from, add +1 RS to all INT-based rolls. You no think so good after getting struck on the head too much. 
\begin{table}[!h]
	\centering
	\caption{Brain Damage Table}
	\label{tab:Brain Damage}
	\begin{tabu} to \linewidth {cX}
		Roll&Effect\\
		1 &Mild Concussion: This Bane fades in 1d10 days \\
		2 &Concussion: This Bane fades in 2d10 days \\
		3 &Severe Concussion: This Bane fades in 3d10 days \\
		4 &Gain Bad Eyes (or lose Good Eyes if you have it) \\
		5 &Gain Bad Ears (or lose Good Ears if you have it) \\
		6 &Gain Old Wound and Bad Ears (or lose Good Ears if you have it) \\
		7 &Gain Old Wound and One-Eyed (determine which eye randomly) \\
		8 &Gain Old Wound and Mute \\
		9 &Gain Old Wound and Lasting Pain \\
		10&Gain Old Wound, Lasting Pain, and Blind \\
	\end{tabu}
\end{table}
\paragraph{\label{bane:Broken Limb/Appendage}Broken Limb/Appendage}\quad\\
\textit{Cannot be removed}\par
A broken limb is a serious injury, but with some time and rest, it’ll be as good as new. Right? 
Well, in theory anyway. If a limb is used while broken, it may not heal properly, and repeated injuries to a wounded limb can permanently cripple it. A broken limb cannot be used for anything (wielding a sword, walking, and so on), until the Wound that caused it heals (a treated broken leg can be walked on with the aid of crutches, but at 1/4th of normal MOB, and no intense movement is possible). 

You may not choose this Bane at Character Creation, and it cannot be bought off. The Wound must heal normally. 
\paragraph{\label{bane:Complete Monster}Complete Monster}---\quad\textbf{(10) }\par
Some people are cruel, petty or spiteful due to their poor upbringing, the rough environment they were 
exposed to, or an unpleasant childhood. Their insecurities manifest themselves in antisocial traits 
developed as coping mechanisms to deal with the pain these situations caused them. 


Not you. You may have excuses, but they're not legitimate. You are wolf to man. Maybe you're crazy, or maybe you're just too sane. You view human social concessions as an idiotic charade, proof that the 
people around you are just machines, automatons with no real agency in their own behavior. You're the 
only real person. They're just puppets made out of meat. You can play the game, but your definition of 
winning and losing is much more practical. You win if you get what you want, you lose if you don't, and 
everyone else is just an tool to be used or an obstacle to be overcome in the pursuit of your desires. 


You also have an uncontrollable urge to kick puppies and steal pies. 


You gain a +2 bonus to Persuasion rolls, because of your superficial charm, and this remains as long as 
the true depth of your soulless evil remains unknown. It is possible to retain a good reputation with this 
Bane, if you're very clever, but anyone who realizes your true nature sees the depths of your monstrosity, 
and you gain a permanent -4 penalty to Social rolls against them, as well as losing your normal bonus to 
Persuasion against them. Characters may glimpse your true nature whenever you fail a Persuasion roll 
against them (context is everything, GM's discretion), but are much more likely to realize it if they 
witness you actually behave like a Complete Monster. 


You also do not have a Belief Arc, and you cannot gain one unless this Bane is bought off. This Bane can be bought off, but only at double its purchasing cost ,A total of 20 Arc Points, and requires some sort of serious soul-searching epiphany. 


A character cannot have Complete Monster at the same time as the Honorable and/or Virtuous Banes. 
\paragraph{\label{bane:Craven}Craven}---\quad\textbf{(7/15) }\par
You're a coward. There's a difference between feeling fear, and being incapable of overcoming it. A coward cannot bring himself to confront any sort of danger head-on, and will often try to hide, avoid, or 
simply flee from any possible injury or harm to himself. 

\begin{itemize}
	\item [7:] You suffer a -2 penalty to CP in any combat situation in which you do not have a decisive and obvious advantage. You must make a WIL test at RS 2 to bring yourself to confront any sort of danger. If you are injured by violence, you must make a WIL test at RS 4 or panic and try to escape. 
	\item [15:] You suffer a -4 penalty to CP in any combat situation in which you do not have a decisive and obvious advantage. You must make a WIL test at RS 4 to bring yourself to confront any sort of danger. If you are injured by violence, you must make a WIL test at RS 6 or panic and try to escape.
\end{itemize}
Craven cannot be taken with the \nameref{bane:Honorable} Bane.
\paragraph{\label{bane:Crippled Limb/Appendage}Crippled Limb/Appendage}---\quad\textbf{(8)}\\
\textit{Cannot be removed}\par
A crippled limb is one that has suffered significant damage and has not been allowed to heal properly. As 
a result, it has lost most or all of its functionality, and is essentially dead weight, or close to it. Someone with a crippled leg is still better off than someone with no leg at all, but not by much.


The Crippled Limb can be used, but at severe penalties. Any Skill tests made that require the limb suffer a +4 to their RS. Attack or Defense Maneuvers made with the limb suffer a +3 to their TN. If the crippled limb is being used for locomotion (like a leg when walking, or an arm while climbing) MOB is reduced by half.
\paragraph{\label{bane:Dead}Dead}---\quad\textbf{(100) }\\
\textit{Cannot be removed}\par
You are dead. You may take no Actions except the Decompose Maneuver.
\paragraph{\label{bane:Debt}Debt}---\quad\textbf{(2/4/8) }\par
You owe people money. The amount is significant, but you have some time to pay it off before bad things 
start happening. 


In many cases, nobody can actually force you to repay your debts (particularly if you're an armed man 
with few, if any, solid assets and a horse), but moneylenders have long arms and longer memories, and a 
great incentive to either make you pay in gold, or in pounds of flesh, as an example to others. Characters 
that start with no money cannot take this Bane. 
\begin{itemize}
	\item [2:] Minor. You owe an amount of money equal to the Wealth you have at the time of Character Creation, not including Assets. This is determined with Boons and Banes that modify Wealth and Assets. 
	\item [4:] Moderate. You owe an amount of money equal to half again your starting Wealth. This is determinedwith Boons and Banes that modify Wealth and Assets. 
	\item [8:] Major. You owe an amount of money equal to twice your starting Wealth, and starting Assets. This is determined with Boons and Banes that modify Wealth and Assets.
\end{itemize}
\paragraph{\label{bane:Dependent}Dependent}---\quad\textbf{(6/12) }\par
Someone relies on you. Someone you can’t leave behind. And if anything were to happen to them…


Your Dependant is an NPC created by you and the GM (confer with your GM during creation) who tags 
along with your character. The Dependant can’t just be sent off--maybe you just can’t bear to part with 
them, or they wouldn’t be safe anywhere else, or you’re protecting them from something nobody else can 
even be allowed to know about. Whatever the reason, the Dependant is a chink in your armor that must be 
defended at all costs. 
\subparagraph{6:} Create a Follower with 4 less PCP than the PC’s. They may not spend more than 4 PCP on Boons \& Banes, may not be a Tier 5 Race, and should be of a PCP 4 Social Class or lower (Same or lower than the Player). The Dependant may gain Arc Points through their own Arcs, or however else the GM sees fit. Should the Dependant be in danger, you must roll a simple WIL test with RS of 2, or immediately rush to their defence. If the Dependant is injured, you have a -1 to all actions that does not involve saving, rescuing or healing your dependant. If your Dependant dies, and it is not your fault, you could have done nothing to intervene, or the Dependant themselves tried to save someone, then you must pay back your next 6 Arc Points as you slip into a depressed state. If you COULD have saved them, but failed due to being overwhelmed by enemies, failing a skill roll, or whatever other reason, then the next 12 Arc Points must be paid as a cost, and you could be in danger of receiving an Insanity.\par

\subparagraph{12:} Create a Follower with 8 less PCP than the PCs. They may not spend more than 2 PCP on Social Class, and may only put 1 PCP into Boons \& Banes. They are restricted to half of the Bane holder’s Proficiency PCP. The Dependant gains Arc Points at half the rate they normally would. Should the Dependant be in danger, you must make a WIL test with an RS of 4, or immediately rush to their aid. You cannot leave your Dependant behind anywhere, unless you make arrangements for their safety (A castle, with owner and guards friendly to the players counts, likewise does a tavern with lockable doors), even if this means you bring them into battle with you. If you must leave them behind anywhere that you cannot be sure is safe, or if you know they are in danger, you take a -2 to all actions that do not lead towards their rescue. If the Dependant should die for any reason, you must pay the next 10 Arc Points you gain as your world shatters around you. You have lost a very important part of your world, and you may never recover. If, however, they die due to you a failure on your part, in a combat you are also taking part in, or because you failed a Skill roll that would have saved them, the next 20 Arc Points must be paid in cost, the GM is HIGHLY encouraged to provide you with an Insanity, or two, and you lose the ability to gain Arc Points from one of your Arcs for a period determined by your GM.\par

\subparagraph{Special:} If the GM allows it, you may, instead of taking the Arc Point hit, take a number of Banes that would represent your character’s inability to process or cope with what has happened. The GM may decide these for you, or you may put forward a suggestion. This is a serious event in your character’s life, and thus the Banes you take should represent this.
\paragraph{\label{bane:Dire Past}Dire Past}---\quad\textbf{(0) }\\
\emph{Cannot be removed}\par
You've been through some serious stuff. Maybe you're a survivor of a terrible battle, or maybe you're a veteran of the Crusades. Maybe you spent your childhood as a cabin boy on a witch-hunter's ship, fighting infidels and blasphemous cults on the islands along the Barbary Coast. Maybe you were born in the dark, where others merely adopted it.


Write or explain a brief backstory for your character to the GM. He will choose (or design) several Banes 
for you, to represent the scars and looming shadows of your dark past (you are not awarded B\&B Points 
for these Banes directly).


As you have survived your Dire Past, however, you gain 10 additional B\&B Points to spend on Boons. 
These B\&B Points do NOT count towards your maximum points from Banes. 


GMs should not be lenient when choosing Banes to suit a character’s Dire Past. Be vicious. Old Wounds, 
One-Eyed, Enemies, Bad Reputation... even things like Hothead, Honorable and Virtuous are appropriate 
to apply for this Bane. The character isn’t just awarded 10 B\&B Points, that’s 10 Points in addition to what they can earn by taking Banes, make them work for it! 
\paragraph{\label{bane:Enemies}Enemies}---\quad\textbf{(3/10/15) }\par
You have some powerful enemies who mean to do you serious harm. They may just want to ruin you and 
crush your name, or perhaps they want to cut your head off and put it on a pike. Either way, they're 
willing to go out of their way—potentially FAR out of their way—to do it. The level of this Bane 
indicates just how serious an enemy you've made. Enemies can either be individuals of significant power, or entire organizations or countries. 
\begin{itemize}
	\item [3:] Single, dangerous individual, small group or minor organization (rival merchant, local guild). 
	\item [10:] Single, powerful individual, large group or organization, regional authorities (Sheriff of Nottingham, the City Guard).
	\item [15:] Incredibly powerful enemies with long arms, continent-spanning organizations (the Holy Roman Emperor, the Church, the Teutonic Order).
\end{itemize}
\paragraph{\label{bane:Facial Deformity}Facial Deformity}---\quad\textbf{(2/4/8) }\\
\textit{Cannot be removed}\par
People remember your face, and not in a good way. An injury, disease, or birth defect has marred your 
looks, and you tend to draw attention from your peers. You may not buy off this Bane, barring some 
miraculous treatment or magic (at the GM's discretion). 
\begin{itemize}
	\item [2:] You have a distinctive (though not particularly disfiguring) mark on your face. You suffer a -2 to checks to disguise yourself or lie about your identity, and people will have an easy time describing you. The mark is not significant enough to disturb people. Examples include Otto Skorzeny and Ernst Blofeld. 
	\item [4:] Your face has suffered serious injury or disfigurement. You suffer a -4 to checks to disguise yourself or lie about your identity, and people will have a very easy time describing you. In addition, you also suffer a -2 to the first Social checks you make with any person, as your first impressions are poor. Intimidation rolls are exempt from this penalty, and may (situationally) be enhanced. Examples include Sandor Clegane and Tycho Brahe (without prosthetic).
	\item [8:] Your face is a horrific ruin, and people may have difficulty realizing that you are human, if they see you without warning. You suffer -6 to checks to disguise yourself or lie about your identity, and people will have an incredibly easy time describing you. In addition, you suffer a -4 to the first Social checks you make with any person, as your first impressions are poor, and continue to suffer -2 afterwards, as you simply make people uncomfortable (long-time friends and the particularly understanding might be exempt from these penalties). Intimidation rolls are exempt from this penalty, and may (situationally) be enhanced. Examples include Baldwin IV of Jerusalem and Joe Bonham.
\end{itemize}
\paragraph{\label{bane:Fat}Fat}---\quad\textbf{(5)}\par
You are rotund! A lot of folks have a bit of extra meat on them, but you look like you just caused a 
famine in England. Put the fork down, fatty! 


You suffer a -2 to your HLT for the purposes of determining Fatigue, and a -2 to MOB. 
On the upside, you gain a +1 to Stability rolls. 


Fat cannot be taken at the same time as the Skinny Bane.
\paragraph{\label{bane:Haemophilia}Haemophilia}---\quad\textbf{(8) }\\
\textit{Cannot be removed}\par
Haemophilia is a genetic disorder that impairs the body’s ability to control blood clotting or coagulation, 
and the character’s body has trouble stopping bleeding when any vessel is broken. This is an absolutely 
terrible condition to have. 


Whenever you suffer any sort of \nameref{sec:Blood Loss}, increase the amount of Bleed points you gain by 3. The RS 
for medical tests made to stop the bleeding are always increased by 2. 


\paragraph{\label{bane:Hothead}Hothead}---\quad\textbf{(3) }\par
‘Someone get this hothead outta here!’ You get angry easily, and you're prone to overreacting to 
perceived threats or insults. Whenever you feel threatened, insulted, or aggravated by somebody, or 
generally frustrated with a situation, you must make a WIL check at RS 2 or begin either a serious verbal 
or physical confrontation immediately. 


If you are in a situation in which starting such a confrontation would clearly not solve anything (not even 
by venting your anger by smashing something), or would very obviously result in your own death, the 
WIL check is reduced to RS 1. 


\paragraph{\label{bane:Honorable}Honorable}---\quad\textbf{(15) }\par
You're genuinely honorable. You might not be a nice guy, you might not be a philanthropist, but you have 
a real sense of honor. Honor can't be bought, nor can it be awarded. Honor is fought for, acquired, and 
maintained with diligence. To never break one's word, freely given, to never violate certain rules, to treat 
others, and oneself, with respect. You've got this, and it is representative of real strength of character. 
However, those bound by honor can also be dragged down by it. Honorable behavior is not always smart 
behavior, and those who break their own codes of honor are lessened by it, as an essential part of 
themselves dies in the act. 


Honorable characters gain a +2 bonus to resist subterfuge, intimidation, torture, and other means of 
prying information from them. Additionally, he gains a +1 bonus to all social checks made to persuade, 
reason with, or debate those who know the character to be honorable (this could be everyone, if the 
character is famous, or just those who know him well or have been impressed with his conduct). 

However, a character with this Bane must always conduct himself in an honorable manner (discuss with 
your GM what ‘honor’ means for your character, establish the principles, and stick to them) or else suffer serious consequences. 

A character who goes against his own principles loses this Bane, and must pay the next 10 Arc Points 
earned as a penalty, as his character works through the crisis of conscience brought on by the collapse of 
his worldview. At his discretion, the GM can waive this penalty if the character's collapse was not due to 
moral weakness so much as extreme circumstance. Honorable can also be purchased again (at the GM's 
discretion) after being lost, to represent the character regaining his principles for 2 Arc Points. 
 
A character cannot have the Honorable Bane at the same time as the Complete Monster Bane. 
\paragraph{\label{bane:Lasting Pain}Lasting Pain}---\quad\textbf{(4/8) }\\
\emph{Cannot be removed}\par
You've got an injury that not only hasn't healed properly, but that causes you chronic or constant pain in the limb. You may learn to live with the pain, but it is never far from your mind.


Choose a Hit Location, such as hand, thigh, face, belly, and so on, which will be the location of the injury or defect that causes the Lasting Pain. If you acquire this Bane through an injury as detailed in \autoref{ch:injuryrecovery}:~\nameref{ch:injuryrecovery}, you do not get to choose the area. 
\begin{itemize}
	\item [4:] Minor. It hurts, but not too much. Each day, roll 1d10. On a 1-8, you suffer 2 Pain throughout the day. Any injury to the area awakens this pain until the injury is fully healed. 
	\item [8:] Major. The pain can be crippling. Each day, roll 1d10. On a 1-9, you suffer 4 Pain throughout the day. Any injury to the area awakens this pain until the injury is fully healed.
\end{itemize}
\paragraph{\label{bane:Mute}Mute}---\quad\textbf{(5/8) }\\
\emph{Cannot be removed}\par
You can't talk. Perhaps you've suffered a throat injury or had your tongue cut out. Perhaps you've been 
that way since birth, or maybe a disease ravaged your vocal chords. Either way, you cannot communicate verbally in a meaningful fashion.
\begin{itemize}
	\item [5:] You cannot form words, but you can cry out, shout, or otherwise make vocal sounds to, say, alert your sleeping friends that you are surrounded by giant spiders.
	\item [8:] You can neither speak nor make any other vocal sounds, and must communicate entirely by writing or using sign language.
\end{itemize}
\paragraph{\label{bane:Oath}Oath}---\quad\textbf{(2 to 10) }\par
You've sworn a solemn oath, in good conscience and of your own accord. An Oath is a vow or obligation 
that a character has made, to himself, to others, or to God. This Bane assumes that the Oath in question 
was made honestly—simply taking an Oath and then breaking it casually when out of sight doesn't count. 


If you ever break the Oath (setting aside mitigating circumstances, for example, breaking a Vow of 
Silence to warn a child of danger, or to inform the Pope of the assassin drawing up behind him) you must 
pay Arc Points as you earn them equal to twice the value of the Bane, as you struggle emotionally with 
your failure. You may choose to retain your Oath after these Arc Points have been paid, or to abandon it 
and lose this Bane at no further cost. 


Oaths can be worth between 2 and 10 depending on the severity of the Oath being taken. An Oath of 
Fealty for a knight might be worth 2, since it's not something that would be hard for him to keep, whereas 
a Vow of Silence could be worth more, and a Vow of Pacifism could be worth as much as 10, depending 
on the character's background and premise. Consult with your GM to decide what an Oath should be 
worth. Generally, the harder it would be to keep, the more it should be worth. 
\paragraph{\label{bane:Old Wound}Old Wound}---\quad\textbf{(1) }\\
\emph{Cannot be removed}\par
You've suffered a severe injury that has never quite healed properly, and still pains you from time to time, and is particularly sensitive to further injury. 


Choose a Hit Location (such as hand, thigh, face, belly, and so on), which will be the location of the Old Wound. If you acquire this Bane through an injury as detailed in \autoref{ch:injuryrecovery}:~\nameref{ch:injuryrecovery}, you do not get to choose the area.


Any attack that hits this location automatically inflicts Stun equal to a level 1 Wound to that area, ignoring all reductions, even if the attack inflicts no Wound.


If this Bane is bought off, its cost to remove it is 5, instead of the 1 suggested by its cost. 
\paragraph{\label{bane:One-Eyed}One-Eyed}---\quad\textbf{(10) }\\
\emph{Cannot be removed}\par
You've lost an eye! Perhaps it was an accident, or through injury in combat, or maybe you were born with only one functional eye. Stuff happens. Either way, having one eye can be a disadvantage in combat and in daily life. One-Eyed people can learn to compensate for their lack of depth perception with simple tricks, however it is still difficult to gauge distances.


PER tests to spot things, gauge distances, and so on, suffer a +1 RS. You also suffer a -1 to your CP in combat, and a -2 to your MP when making ranged attacks, (though not for using missile weapons in melee, those use the CP penalty listed earlier).


If you buy off this Bane, you don't grow a new eye, but become so accustomed to only having one that you no longer suffer the penalties. You cannot remove the Bane, however the penalties don't apply anymore.


If somehow you acquire two One-Eyed Banes, then you regrettably lose both One-Eyed Banes, and gain the Blind Bane, unless you have more than two eyes (i.e. if you are a giant goliath spider possessed by the ghost of Black Hand Lee).


At Character Creation, you may take this Bane as a -2 Facial Deformity to represent the One-Eyed Bane you have ‘bought off’ before Character Creation. Do not apply penalties from One-Eyed, but other rules still take effect.


A character cannot have One-Eyed at the same time as the Blind Bane.
\paragraph{\label{bane:Poor}Poor}---\quad\textbf{(4/6/8) }\par
You're unusually poor for your class in society. You lack funds, perhaps because you or your ancestors made poor investments, without falling too far in social status. Either way, it's likely that you're still trying to claw your way back into relative affluence. Maybe it's even what motivates you.

\begin{itemize}
	\item [4:] You start with half (50\%) the Wealth of a normal character of your Social Class and Wealth level. You may only take this level of the Bane if your character is of Wealth level 2 or higher, or with GM's permission.
	\item [6:] You start with a quarter (25\%) the Wealth of a normal character of your Social Class and Wealth level.You may only take this level of the Bane if your character is of Wealth level 3 or higher, or with GM's permission.
	\item [8:] You start with no money.You may only take this level of the Bane if your character is of Wealth level 4 or higher, or with GM's permission.
\end{itemize}
A character cannot have the Poor Bane at the same time as the Rich Boon.
\paragraph{\label{bane:Severed Limb/Appendage}Severed Limb/Appendage:}---\quad\textbf{(10/15/18) }\\
\emph{Cannot be removed}\par
You're missing an arm or a leg. This could be the whole arm or leg, or just the hand or foot, but the loss of the manipulator at the end is what's important.

If you choose this Bane at Character Creation, choose a limb affected by this Bane, and how much of the limb is missing. If you suffer this Bane because of an injury, then of course you have no say in where the limb has been lost.
\begin{itemize}
	\item [10:] The hand or foot has been lost.
	\item [15:] The lower half of the limb has been lost (from the elbow or knee).
	\item [18:] The full limb has been lost (from the shoulder or hip).
\end{itemize}
The penalty for having lost a limb is serious.
\vspace*{-10pt}\subparagraph*{Hand:}
You cannot perform any task that requires both hands, or use any two-handed weapon. Gripped 
shields cannot be used (strapped shields can, at the GM’s discretion, with minor modification). 
When Grappling, making a Slip attempt to prevent this arm from being Trapped or Pinned has a -
1 TN, unless you have a prosthetic, and Force attempts to untrap the limb are made at a -1 TN as 
well. Your Punches inflict -2 damage without a prosthetic. 
\vspace*{-10pt}\subparagraph*{Lower Arm (from the Elbow):}
You cannot perform any task that requires both hands, or use any two-handed weapon. No shields 
of any kind can be used. All Weapon Maneuvers suffer a +1 Activation Cost because of your loss 
of balance. You may pay 5 Arc Points to remove the Activation Cost (representing your character 
getting used to the missing limb). You can no longer perform the Punch or Elbow Maneuvers 
without a prosthetic. All Grappling Maneuver TNs are increased by 1. 
\vspace*{-10pt}\subparagraph*{Full Arm (from the Upper Arm to Shoulder):}
You cannot perform any task that requires both hands, or use any two-handed weapon. No shields 
of any kind can be used. All Weapon Maneuvers suffer a +2 Activation Cost because of your loss 
of balance. You may pay 8 Arc to remove the Activation Cost (representing your character 
getting used to the missing limb). All Grappling Maneuver TNs are increased by 2. 

Missing both hands or arms prevents you from using any held weapons. It is still possible to 
Grapple unless you are missing both arms to the elbow or more, but the penalties are cumulative. 

\vspace*{-10pt}\subparagraph*{Foot (from the Foot or lower Shin):}
You suffer a -2 penalty to MOB, and must make a Stability roll at RS 2 whenever moving faster 
than a slow limp (a quarter of normal speed) or fall. All Stability rolls forced upon you by other 
sources have their RS increased by 1. You suffer a CP penalty of 2 in combat in addition to these 
other penalties, and all Dodge Maneuvers have their TNs increased by 2. 
\vspace*{-10pt}\subparagraph*{Lower Leg (from the upper Shin to Knee) and Full Leg (from the Thigh to Hip):}
You cannot walk except to possibly hop at a quarter of normal MOB. You must make a Stability 
test every turn you move like this at RS 3, or fall and be prone. All Stability rolls forced upon you 
by other sources have their RS increased by 2. You suffer a CP penalty of 4 in combat in addition 
to these other penalties, and all Dodge Maneuvers have a +2 Activation Cost, and their TNs 
increased by 2. A Crutch can allow you to move at half the normal MOB without a chance of 
falling, and reduce the CP penalty to 2, but you cannot run, and a crutch precludes using a 
weapon on the lost-leg-side hand, and the use of 2H weapons. You cannot perform Knee or Kick 
maneuvers with your lost leg, and may only perform kicks with your remaining leg (if you have 
one) while prone, or with a crutch at +1 TN and +1 Activation Cost. 
\vspace*{-10pt}\subparagraph*{Full Leg (from the Thigh to Hip):}
XXXXXXXX

If both legs are lost, movement is extremely difficult. MOB is reduced to an effective 1, running 
is not possible. Combat is unthinkable. You are permanently Prone, and can perform no Dodge 
Maneuvers. 


If a severed limb is used without giving it time to heal, the user immediately gains 4 Shock and 1 Pain.


Prosthetics are available in \autoref{ch:Prosthetics} and will mitigate some of the issues with missing limbs.
\paragraph{\label{bane:Sheltered}Sheltered}---\quad\textbf{(2/4/6) }\par
For some reason, you aren’t quite as used to the world as you should be. You’re less hardened against the 
harshness of the world, and things tend to affect you more than your fellows. You start with less Grit than 
you normally would. You cannot reduce your Grit beyond 0. 
\begin{itemize}
	\item [2:] Softy. -1 Grit.
	\item [4:] Seriously Sheltered. -2 Grit.
	\item [6:]You don’t even know what color blood is. -3 Grit.
\end{itemize}
Cannot have the Sheltered Bane at the same time as the True Grit Boon. 
\paragraph{\label{bane:Short}Short}---\quad\textbf{(8/15) }\par
You're much shorter than normal for your race. This negatively affects your Reach and your MOB, but you also tire less easily, and you're a harder target for archers and other attackers with missiles.
\begin{itemize}
	\item [8:] You're about 20\% shorter than the average person. You suffer -1 to both Reach and MOB, but you reduce your Exhaustion Rate by 1x (to a minimum of 1x) for the purposes of acquiring Fatigue. Missile Attacks reduce their MP against you by 1 (or their CP by 1, if they are made as Melee Maneuvers in close combat). 
	\item [15:] You're about 30\% shorter than the average person, and quite possibly have some form of dwarfism. You suffer a -2 to both Reach and MOB, but you reduce your Exhaustion gain by 1x (to a minimum of 1x) for the purposes of acquiring Fatigue. Missile Attacks reduce their MP against you by 1 (or their CP by 1, if they are made as Melee Maneuvers in close combat). This is not cumulative with the minor version of Short.
\end{itemize}
A character can not have the Short Bane at the same time as the Tall Boon. Duh. 

\paragraph{\label{bane:Skinny}Skinny}---\quad\textbf{(3) }\par
You’re thin as a rail, a regular beanpole, and unwholesomely thin! Italian grandmothers everywhere 
actively seek you out and attempt to feed you delicious pasta.


You suffer a -1 to your Stability rolls, and your effective CAR for determining Encumbrance is reduced by 1. However, your Exhaustion Rate is reduced by 1x (to a minimum of 1x) for the purposes of acquiring Fatigue.


A character cannot have the Skinny Bane at the same time as the Fat Bane.
\paragraph{\label{bane:Technologically Impaired}Technologically Impaired}---\quad\textbf{(5) }\par
You lack a working knowledge of modern technology, and have great difficulty understanding how all 
these newfangled devices work. You may not have any Proficiency in a weapon considered modern, nor 
knowledge of or Skills pertaining to any sciences that are on the cutting edge.You may have outdated or 
obsolete expertise, though—confer with your GM. 


If you want to acquire a new Proficiency involving a modern weapon, or to acquire Skills pertaining to 
modern technology, the first point or rank you buy in any of these Proficiencies or Skills costs double the 
normal amount of Arc Points, if you are learning it as part of a School. 


If you’re trying to acquire a new Proficiency without instruction, you must first spend twice the normal 
amount of Arc Points to begin a Proficiency, and then make an INT check at an RS determined by the 
complexity of the device (a hand-spanned crossbow being RS 2, a cranequin-spanned crossbow being RS 
4, a wheellock musket being RS 5, and a Puckle gun being RS 6. The GM has the final say on what the 
RS for figuring out a weapon is. Some may be impossible without instruction). Success means that you manage to figure out the weapon sufficiently to take Proficiency in it. Failure means you still spend the Arc Points, but do not gain Proficiency in it. However, you may try again, this time with a cumulative -1 to the RS.
\paragraph{\label{bane:Unhappily Married}Unhappily Married}---\quad\textbf{(1/2/3) }\par
Your significant other doesn’t like you very much, whether or not the feeling is mutual. They will go to lengths to make your life more difficult, and will generally be a pain in the neck. They can be eluded for a short time, but they always catch up to you in the end.
\begin{itemize}
	\item [1:] The spouse only makes a minor fuss, whether treating guests badly and bringing disgrace to your name or spreading rumors about you while you’re out adventuring. 
	\item [2:] The spouse actively tries to disrupt your life in some way. The gravity of this torment depends on the Social Class and Wealth of the spouse—a disgruntled peasant wife might intentionally undercook your lunch, but your treacherous husband the Duke might humiliate you in court. 
	\item [3:] The spouse really has it out for you, and they have your in-laws to back them up. A poorer family might just regularly harangue you or work to make your life miserable. In noble courts, this sort of bitterness results in assassins with ropes hiding in your bedroom, and having to drink out of a hip flask at every meal.
\end{itemize}
\paragraph{\label{bane:Virtuous}Virtuous}---\quad\textbf{(10) }\par
You possess that greatest of weaknesses, the one flaw that has been the bane of more otherwise perfectly 
competent and ambitious heroes than any other: you are a genuinely moral, honest, and righteous person. Perhaps it's just in your nature, or perhaps you were just raised well. You've got a conscience, a desire to help people, and while you may not be a pacifist, you're no murderer, and you despise senseless violence, drawing the sword only when necessary. This isn't an easy way to live. Historically, paragons of morality were few and far between, and it has often been said that ‘the good die young.’


If you ever act in an immoral, unnecessarily cruel or ruthless fashion, you must pay the next 10 Arc Points you earn, as you are wracked by your conscience. However, if you immediately act to try and make up for your moral transgression, by making amends with the wronged or through atonement to society or God, this loss is reduced to 5 Arc Points.


A character cannot have the Virtuous Bane at the same time as the Complete Monster Bane
 

\paragraph{\label{bane:Wanted}Wanted}---\quad\textbf{(5/10/15) }\par
You're wanted by the law, either in your own country, or others. Believe it or not, even in the medieval 
era, people often acknowledged the criminals of other kingdoms. Often, but not always. Whomever you've committed a crime against, they're willing to go to lengths to retrieve you. 
\begin{itemize}
	\item [5:] Alive.
	\item [10:] Dead.
	\item [15:] Alive, and to be kept alive. For a long, long time. 
\end{itemize}
\chapter{Skills}\label{ch:skills}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage 
\section{Skills}\label{sec:skills}
Skills represent a character’s acquired knowledge and expertise in a wide variety of activities, from 
translating ancient scripture, to carving stone, and rigging the sails on a ship. The full list of Skills can be found in \autoref{ch:skills}:~\nameref{ch:skills}.


At Character Creation, the number of Skill Points a character receives is determined by the PCP placed in 
the section, as well as a number of points equal to the character’s INT. Buying Skills with these points 
works identically to buying them with Arc. These points can be spent freely on any Skill within the 
section, barring those the GM excludes from the campaign, and up to the maximum starting skill level. 
The maximum skill level attainable at Character Creation is equal to your Intelligence (INT) score. 


Skills represent a character's acquired knowledge and abilities, including things like historical knowledge, 
surgical skill, and tracking. There are dozens of such skills, each detailed in the Skills section. Skills are leveled up by spending Skill Points on them. 


The Skills chapter is one that represent most of a character’s knowledge, expertise and abilities, especially in non-combat areas.


Skills are a system within the game that cover things like how well and far a character can jump or a character’s knowledge of the history of a neighboring kingdom. Each skill comes into play in the form of a skill test, which uses a character’s attributes and skill level to generate a pool against a static difficulty provided further in the book. Skills are used in “skill tests” where a character uses their skill pool to roll against the situation’s difficulty (measured in required successes or RS) at a static Target Number (TN).


Skills are different from Proficiencies and other systems in that they typically do not often directly affect combat, but they are still very useful both in and out of combat. A knight has to know how to ride his horse, an assassin has to know how to blend into a crowd, and a hunter had better know how to live off the land.


Skills List:
\begin{itemize}
	[noitemsep]
	\item \nameref{skill:Athletics}
	\item Chymistry --- (Not in Beta)
%\nameref{skill:Chymistry}
	\item \nameref{skill:Climbing}
	\item \nameref{skill:Cooking}
	\item \nameref{skill:Crafting}
	\item Drill --- (Not in Beta)
%	\item \nameref{skill:Drill} 
	\item \nameref{skill:Engineering}
%	\item \nameref{skill:Gather Information}
	\item \nameref{skill:History}
	\item \nameref{skill:Hunting}
	\item \nameref{skill:Intimidate}
	\item \nameref{skill:Navigation}
	\item \nameref{skill:Observation}
	\item \nameref{skill:Orate}
	\item \nameref{skill:Perform}
	\item \nameref{skill:Persuasion}
	\item \nameref{skill:Politics}
	\item \nameref{skill:Research}
	\item \nameref{skill:Riding}
	\item \nameref{skill:Sailing}
	\item \nameref{skill:Stealth}
	\item Strategy --- (Not in Beta)
%\nameref{skill:Strategy}
	\item \nameref{skill:Subterfuge}
	\item \nameref{skill:Surgery}
	\item \nameref{skill:Swimming}
	\item Tactics --- (Not in Beta)
%\nameref{skill:Tactics}
\end{itemize}

\paragraph{\label{par:Trained}Trained} No longer count as ‘Untrained’ when using the skill.

\vspace{-10pt}\paragraph{\label{par:Teacher}Teacher} Grants the character the ability to give another character one free level of the skill the teacher is proficient in. This ability may only be used once on a person for a single skill, and only a number of skills equal to the receiving character’s INT may be trained this way total. Using this takes about a month of in-game time, give or take at the GM’s discretion, and requires the location and materials needed to perform the trained skill. Multiple characters may be trained at the same time as long as their training starts at the same time and there are adequate resources and facilities.

\setlength{\intextsep}{0pt}
\begin{wraptable}{r}{8 cm}
	\centering
	\begin{tabu} spread 0pt{X[c]X[c]XX[c]}
		Skill Level&Skill Pool&	Skill Bonus&Arc Price\\\toprule
		1&			+1&			\nameref{par:Trained}&	1\\
		2&			+2&			-&			1\\
		3&			+3&			-&			1\\
		4&			+4&			\nameref{par:Teacher}&	1\\
		5&			+5&			-&			2\\
		6&			+6&			-&			2\\
		7&			+7&			-&			2\\
		8&			+8&			-&			3\\
		9&			+9&			-&			3\\
		10&			+10&		\nameref{par:Mastery}&3\\
	\end{tabu}
	\caption{Skills Overview}\vspace{-15pt}
	\label{tab:Skill Levels Overview}
\end{wraptable}
\setlength{\intextsep}{\oldintextsep}

\paragraph{\label{par:Mastery}Mastery} Once per session, you may automatically succeed with 1/2 dice rolled for one Skill roll.
\paragraph{\label{par:Skill Level}Skill Level} Each skill has a level (ranged from 1 to 10) decided by the character’s investment of Skill Points (at character creation) or Arc (later game). The skill level gives the character dice equal to their skill level for skill tests, as well as other bonuses at various skill levels. Skills cost Arc to increase. Upon obtaining a level with a listed Skill Bonus, the character receives the stated bonus.
\paragraph{\label{par:Skill Pool}Skill Pool (SP)} A skill pool is a pool of dice used during a skill test. The pool is generated from the relevant Attribute at one die per Attribute value, added with dice equal to the character’s skill level in the relevant skill. Only one attribute can be referenced per skill test; if the situation ever calls upon a different attribute, it becomes a second test. For instance, when using the Climb skill, the character might be called on to make both STR and END based tests.
\paragraph{\label{par:TN}TN} The Target Number (TN) for skill tests is 7. It is always 7. No force can change it from 7 to anything else. Anything that suggests that it is possible to change the Target Number for skill tests to a number other than 7 is a lie. Ignore those anythings.
\paragraph{\label{par:Skill Tests}Skill Tests} There are many situations in the game in which a character’s skills will be tested, such as trying to negotiate a hostile deal or simply leaping from one roof to another during a chase. Every test of a skill will state a required number of successes (RS) for the skill test to be successful. The character must roll their pool and get equal to or above the skill’s target number (TN), and get equal to or more than the number of required successes (RS) for the test to be passed. Skills’ tests have modifiers to the situation that affect the skill check’s RS, such as trying to climb a wall slanted against the character or knowing the history of a foreign kingdom. If the number of RS is ever raised above the character’s dice pool plus automatic successes the character might get from other factors, then the skill attempt immediately becomes impossible. Situational modifiers can lower RS as well, though it can never be lowered beneath 1. If there is a situation that a skill doesn’t seem to cover, the GM should use their best judgement to create a test based on similar difficulties to the situation.
\paragraph{\label{par:Untrained}Untrained} If a character has no levels in a skill they are attempting to use, then they are considered unstrained in it. Some skills cannot be used if the character is untrained in them:
\begin{itemize}
	[noitemsep]
	\item Chymistry (Not in Beta)
	\item \nameref{skill:Crafting}
	\item \nameref{skill:Engineering}
	\item \nameref{skill:History}
	\item \nameref{skill:Navigation}
\end{itemize}
\paragraph{Leveling up Skills} Skills have a leveling system based on the amount of points spent in any given skill to increase it to the next level. Skill levels are bought simply by paying the Arc price.
\paragraph{Retrying Skill Attempts} For physical skills, they can be retried any amount of times, unless injuries or situation stops the character from doing so. For mental ones, such as History and Politics, a single check is allowed, any retries have to be done through the Research skill. For social skills, it’s best to wait before trying again, as if tried again too soon, the target may not react positively, at the GM’s discretion.

\setlength{\intextsep}{0pt}
\vspace{-5pt}\begin{wraptable}{r}{8 cm}
	\centering
	\caption{Encumbrance Level}
	\label{tab:Encumbrance Penalties}
	\begin{tabu} spread 0pt{XX[c]X[r]}
		Encumbrance		&Mobility Effect&Skill Penalty\\\toprule
		Unencumbered	&0				&0\\
		Light			&-2				&+1 RS\\
		Medium			&-4				&+2 RS\\
		Heavy			&-6				&+3 RS\\
		Overloaded		&-8				&+4 RS\\
	\end{tabu}
\end{wraptable}
\setlength{\intextsep}{\oldintextsep}

\paragraph{Encumbrance} The amount of gear worn can affect certain skills. There are five levels of Encumbrance: Unencumbered, Light, Medium, Heavy, and Overloaded. See \autoref{sec:Encumbrance}

\subsection{Character Creation}
At Character Creation, Skills can be purchased by spending Skill Points, given at an amount depending on the invested PCP plus an amount equal to the character’s INT, by purchasing individual skills.


SIDENOTE: Skill Descriptions %MAKE this a sidebar?
A Note to GMs: Though we endeavor to make the skill system as accurate as possible, the reality may be 
a situation makes more sense for a character to use their STR when doing something rather than their 
END. We provide the best examples we can, but it is your job to ensure what your players are doing is 
represented as accurately as possible. 

\section{Skill Descriptions}
\paragraph{\label{skill:Athletics}Athletics}---\quad\textbf{(Various)}\par%THIS TABLE IS REALLY UGLY
Athletics is the skill for moving, jumping, and performing other feats of physical prowess. Gymnastics, sprinting, jumping, and other classic physical activities fall within its purview. Different attributes may be used with Athletics depending on what the character is doing. Someone lifting weights, for example, will probably use STR, while a sprinter would use MOB. A long-distance or marathon runner would use END. Terrain will apply for running. Walking generally doesn’t require an Athletics test, no matter the rigor.


Note: The fastest a mortal person can run is their MOB x4 in yards.
\begin{longtabu} to 1\linewidth {X[1.5]X[r]}
	\caption{Athletics}
	\label{tab:Athletics}\\
\rowfont[c]{}Action & Base Attribute vs. RS or other effects\\\toprule\endhead
	\\
		\multicolumn{2}{c}{Lifting}\\\tabucline[1pt on 2pt] {1-2}
Lifting 10$ \times $STR in lbs. above your head & STR vs. 1 RS \\
Lifting more than 10$ \times $STR in lbs. above your head & +1 RS per 10$ \times $STR (to maximum of 50$ \times $) \\
\\
Holding 10$ \times $STR in lbs. above your head & STR vs. 1 RS \\
Holding that same weight for more than 1 second & +1 RS per second held.\\
	\\
	\\
		\multicolumn{2}{c}{Sprinting}\\\tabucline[1pt on 2pt] {1-2}
Sprinting 2$\times$ your MOB (for up to 5 seconds) &MOB vs. 2 RS\\
Sprinting more than 2$\times$ your MOB (for up to 5 seconds) &+1 RS per 1$\times$ more (to maximum of 4)\\
Sprinting for longer than 5 seconds &+1 RS per 10 seconds of sprinting.\\
	\\
Encumbrance (+1 at Light, +4 at Overloaded)&+1 to +4 RS\\
	\\
	\\
		\multicolumn{2}{c}{Running}\\\tabucline[1pt on 2pt] -
Running $ \frac{1}{4} $ mile without stopping. &No Test Required\\
Running a mile without stopping. &END vs. 2 RS\\
Running 5 miles without stopping. &END vs. 5 RS\\
Running an entire marathon without stopping. (~26 mi.) &END vs. 8 RS\\
	\\
	\\
		\multicolumn{2}{c}{Jumping}\\\tabucline[1pt on 2pt] {1-2}
Vault a Waist Height Obstacle &AGI vs. 2 RS \\
Vault a Shoulder Height Obstacle &AGI vs. 3 RS \\
Vault a Reach Height Obstacle &AGI vs. 5 RS \\
Vault a Jumping Reach Height Obstacle &AGI vs. 6 RS \\
\\
Jump Horizontally 1 yard &MOB vs. 1 RS \\
Jump Horizontally further than 1 yard &+2 RS per yard\\
No Running Start (for horizontal) &+2 RS\\
\\
Jump Vertically 1 foot &MOB vs. 1 RS \\
Jump Vertically more than 1 foot &+2 RS per 1 foot \\
\\
Encumbrance (+1 at Light, +4 at Overloaded) &+1 to +4 RS\\
\end{longtabu}

\paragraph{\label{skill:Climbing}Climbing}---\quad\textbf{(END)}\par
Climbing is the skill that lets characters scale mountains, cliffs, walls, and ropes. In some cases, climbing is easy: climbing a stable ladder that’s at 90 degrees or less, or any dry, sticking, or bogged surface that’s less than 40 degrees up is a simple climbing test. However, even a 10 degree surface will need at least be tough if it’s covered in ice. Terrain can also be a huge factor when rolling Climbing tests. Failing by more than your character’s Climbing skill requires a Stability test against an RS equal to the initial check, with failing resulting in a fall.


To properly roll a climb test, \emph{set the RS at the total distance that is to be climbed}, (1 RS per 10 feet) then apply the modifiers. Failure without falling can still get the person up the amount of successes, at the GM’s discretion.

\begin{longtabu} to \linewidth{X[1.65]X[r]}
	\caption{Climbing}
	\label{tab:Climbing Skill}\\
		\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Climbing with many hand-holds (rockwall, rigging, stable ladder, knotted rope) & END vs. 1 RS per 10 feet\\
Climbing with some hand-holds and support (swaying rigging, an unsecured ladder, or a knotless rope against a wall) &END vs. 2 RS per 10 feet\\
Climbing rigging flailing in the wind, a shaky, unstable ladder, or a free-hanging, knotless rope&END vs. 3 RS per 10 feet\\
Climbing with no handholds (requires tools) &END vs. 5 RS\\
	\\
		\rowfont[c]{}Modifiers &Effects\\\cmidrule{1-2}
Decent climbing tools (climbing picks, spiked boots) &-1 RS (when appropriate)\\
Fear (-1 Climbing away from something scary, +1 towards) &-1 to 1 RS\\
Wall Slanting (-4 helpful slant, -2 favorable slant, +2 difficult slant, +4 malicious slant)& -4 to +4 RS\\
Weather (Wet/Windy at +1, Soaked/Storming at +3)&+1 to +3 RS\\
Encumberance (+1 at Light, +4 at Overloaded)&+1 to +4 RS\\
Time (Rushed at +1 RS, Hastened at +3 RS)&+1 to +3 RS\\
\end{longtabu}
Sidebar: Only apply the modifiers for situations that make sense. Climbing picks don’t help that much with rope!

\paragraph{\label{skill:Cooking}Cooking}---\quad\textbf{(INT) }\par
Cooking is an important skill in a world where most vegetables are grown in unsanitary conditions, and most water is unsafe to drink without boiling. The rise of urban civilization made this doubly important, as the water and ground pollution that comes along with large masses of people make the preparation of food essential. Some foods cannot be eaten at all without preparation--rice consumed without boiling, for example, cannot be properly digested. So in a way, cooking allows for inedible substances to be made edible, making it even more important in an environment where food may be scarce. Cooks require ingredients to make meals. For abstraction purposes, each meal takes about 1 lb of ingredients total per person fed (so 4 lbs for a basic 4 person meal) and needs at least 2 ingredients per level (a fine meal, for example, requires 6 different ingredients). For every ingredient needed that is missing, the RS increases by 1. If the chef is missing more than 3 ingredients, they cannot make a meal of that level and must make a lower-quality meal. The chef may opt to have up to 2 spices act as ingredients, but loses the bonus they would normally give. The meal requires at least one non-spice ingredient. Meals also require 30 minutes of prep time per level, with a lack of time or extra time giving penalties or bonuses accordingly. If the prep time is too short for the maximum RS, the meal cannot be made. A “Meal” is here assumed to be enough foor for 3-4 people at base.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Cooking}
	\label{tab:Cooking}\\
		\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Preserving food &INT vs. 1 per lb\\
Make a poor meal &No Test Required\\
Make a good meal &INT vs. 2\\
Make a fine meal &INT vs. 3\\
Make a lavish meal &INT vs. 4\\
Hide poison in a meal &INT vs. 3+1 per extra dose\\
Make a poisonous food safe to eat &INT vs. 5 to 10 per lb, depending the food\\
	\\
		\rowfont[c]{}Modifiers &Effects\\\cmidrule(lr){1-2}
Large Meal (10 people) &+1 RS \\
Banquet (30 people) &+2 RS \\
Massive Meal (100+ people) &+3 RS \\
Spices (-1 at Cheap, -3 at Expensive &-1 to -3 RS )\\
Haste (Rushed at +1 RS, Hastened at +3 RS) &+1 to +3 RS \\
Leisure (Per extra hour, minimum RS 1) &-1 RS\\
	\\
\multicolumn{2}{c}{Assistants take test at $ \frac{1}{2} $ RS. They give +$ \frac{1}{2} $ Successes, to a maximum of the main Cook’s Skill Level.}\\
\end{longtabu}


Sidebar:
Both making the meal and adding the poison to it are two separate checks.\\
Number of assistants that can be used is at the GM’s discretion. Too many chefs spoils the soup!


\paragraph{\label{skill:Crafting}Crafting}---\quad\textbf{(INT)}\par
Crafting is a general skill umbrella that covers hundreds of potential artisan disciplines. Upon taking the Crafting skill, a character should choose what exact sort of crafting the skill covers. A Shoemaker’s skill is definitely different from a Clockmaker’s. A character can have multiple Crafting skills, each in different disciplines, which are leveled separately. The examples in this list are mostly a general guide. GMs are encouraged to decide just how hard a particular test should be and what attribute to use, especially if it doesn’t quite fit the examples given.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Crafting}
	\label{tab:Crafting}\\
		\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Make a simple item (dagger, mace, staff or pole, lumber, chain, vest, horseshoes, wooden doll, sketch)&INT vs. 2 per item\\
Make a basic item (spear, shortbow, hand axe, gun barrel, maille vest, munitions cuirass, breeches, basic landscape, bottle) &INT vs. 3 per item\\
Make a complex item (longsword, greataxe, maille hauberk, light crossbow, cuirass, layered dress, jointed doll, decent portrait, cheap glass lens) &INT vs. 6 per item\\
Make an advanced item (greatsword, fitted suit, gauntlets, high-quality painting, anatomical drawings, bulletproof cuirass, large glass pane) &INT vs. 10 per item\\
Make a great item (advanced weapon or armor with modifications, papal vestments, famous artwork, fine glass lens, precision watch gears) &INT vs. 15 per item\\
	\\
		\rowfont[c]{}Modifiers & Effects\\\cmidrule(lr){1-2}
Tools (Good Quality at -2, Poor Quality at +2) &-2 RS to +2 RS \\
Time (Rushed at +1 RS, Hastened at +3 RS) &+1 to +3 RS\\
\caption*{Assistants take test at $ \frac{1}{2} $ RS, they give +$ \frac{1}{2} $ Successes, to a total of the assistant’s skill level in the chosen skill, with Unskilled still equally 1.}
\end{longtabu}

\paragraph{\label{skill:Engineering}Engineering}---\quad\textbf{(INT)}\par
Engineering is similar to Crafting, but with a more specific focus: Engineers plan and implement large works, like bridges, siege weapons, buildings, ships, etc. Essentially any construction task that would be physically impossible to do alone will fall under an Engineering Skill rather than a Crafting one. Even mediocre laborers can accomplish great things with a skilled engineer at their head, so the Engineering skill of the extra hands employed is not absolutely important (though it can’t hurt). Many of the parts an Engineer will want to work with are on the market, but many more will need to be custom-made by the appropriate crafter, especially for more complex projects. Engineers would be smart to make friends with crafters or learn to craft for themselves lest they end up unable to make what they want or need. Like with crafting, the examples in this list are only a guide. The GM is free to use their own interpretation. The time it would take to construct these devices is also negotiable--in some cases, it could take years even with sufficient manpower.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Engineering}
	\label{tab:Engineering}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Make a simple item (hut, beam bridge, raft, dugout, roasting jack) &INT vs. 2 per item \\
Make a basic item (log cabin, sailboat, arch bridge, palisade) &INT vs. 3 per item \\
Make a complex item (two-story house, galley, rope bridge) &INT vs. 6 per item \\
Make an advanced item (mansion, castle wall, ship of the line) &INT vs. 8 per item \\
Make a great item (ironclad steamship, suspension bridge, skyscraper) &INT vs. 12 per item \\
	\\
\rowfont[c]{} Modifiers & Effects\\\cmidrule(lr){1-2}
Tools (Good Quality at -2, Poor Quality at +2)&-2 RS to +2 RS\\
Laborers (Extra Laborers at -2, No Laborers at +2) &-2 RS to +2 RS\\
Time (Rushed at +1 RS, Hastened at +3 RS)&+1 to +3 RS\\
\end{longtabu}

\paragraph{\label{skill:History}History}---\quad\textbf{(INT)}\par
History is the record of events both human and natural as chronicled throughout the ages. History is a patchwork quilt of stories, accounts, perspectives and guesswork that makes up the understanding that the present has of the past. How closely history coincides with reality is always uncertain, but it provides the foundation, the context for understanding current events and is thus indispensable, even with its flaws. The History skill allows a character to have or gain knowledge of a past event, people, place, culture, etc. When a character encounters something which they might’ve heard about at some point in the past, or upon which an understanding of history might shed some light or special understanding, the GM may prompt them to make a History roll at an RS determined by the obscurity of the event and the resources available to the character.
\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{History}
	\label{tab:History}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
\rowfont[c]{} Modifiers & Effects\\\cmidrule(lr){1-2} 
Scale of Event (World-changing at -4, Major war/disaster at -2, Minor War/battle/treaty at 0, Small war/disaster at +2, Obscure Skirmish/story/treaty at +4, Forgotten war/legend/treaty at +6) &-4 RS to +6 RS\\
Time (Very Recent -2, Recent 0, Ancient History +4) &-2 RS to +4 RS\\
Location (Home at -3, Local at -2, Nation at -1, Foreign at +2, Far Away Land at +4, Literally a different world at +6)&-3 RS to +6 RS\\
Personal (Happened to me at -5 RS, Happened to Family/Friends at -2 RS &-5 RS to -2 RS)\\
\end{longtabu}

\paragraph{\label{skill:Hunting}Hunting}---\quad\textbf{(Varies) }\par
In Song of Swords, hunting has less to do with killing game and more to do with getting near it in the first place. Obviously, this includes tracking, but it also includes knowing how to hide your scent, how to control and entrap animals, and how to make and disguise simple traps to capture or kill game. Obviously, these skills aren’t useful just for hunting animals. People have been referred to as ‘the most dangerous game’ for a reason, after all.


In most cases, a Hunting test will take only a few moments, with the exception of making Traps. Traps require ten minutes of uninterrupted work per RS on the skill test to make them.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Hunting}
	\label{tab:Hunting}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
	\\
		\multicolumn{2}{c}{Tracking}\\
		\multicolumn{2}{c}{PER vs. 1 or Target’s AGI/Stealth}\\\cmidrule(lr){1-2}
Animal size (1 being an elephant, Humans being a 2, 3 being a housecat, 5 being a single ant)&+1 to 5 RS \\
Flock size (0 being singular, 5 being a pack of wolves, 10 being army ants) &0 RS to -10 RS\\
Age of tracks (0 being a few minutes old, 1 being a few hours, 5 being a day or more) &0 RS to +5 RS\\
Target is bleeding &-1 RS per 5 Blood Loss on the target\\
	\\
	\\
		\multicolumn{2}{c}{Trapping}\\\cmidrule(lr){1-2}
Constructing a simple trap (Pitfall, Snare) &INT vs. 2 RS\\
Constructing an intermediate trap (Punji Pit, tripwire and cage, Spiked branch traps) &INT vs. 4 RS \\
Constructing a complex trap (Tripwire triggered crossbow, falling cage, collapsing floor) &INT vs. 6 RS \\
Especially complex design (Needing a time delay at +1, an especially deadly Rube Goldberg machine +5)&+1 RS to +5 RS\\
\end{longtabu}

\paragraph{\label{skill:Intimidate}Intimidate}---\quad\textbf{(CHA) }\par
The key skill of some bodyguards, a lot of bouncers, and pretty much every good gang enforcer or commissar, Intimidate is used to control people through fear. Naturally, any actions taken with Intimidate are going to be contested actions, usually against WIL.


The effects of successfully intimidation can vary greatly upon the situation, but the basic result is the target is much more likely to do what the intimidator wants.


Modifiers can affect either the Intimidator or the Intimidated, depending on who has the benefit. It’s important to note that the table below is an example of how the particular situation can modify the Intimidate roll. Ultimately, it is up to the GM to confer bonuses and penalties based on the nature of the characters involved.


For Combat purposes, successfully Intimidating a character who doesn’t have the Brave Boon removes their ability to declare Offensive either as what they are “not” going to Declare, or as what they are actually going to Declare, in the first round of Combat with the party that Intimidated them. In combat using the Intimidate skill counts as an action.


Example: Gizka intimidates Goofus, overcoming her short height with the double-barreled dragon pistol she has on the table. She beats Goofus’s score, and immediately attacks. Goofus cannot declare that he isn’t going Offensive, but he also can’t go Offensive. So, he must declare that he is not going Cautious, leaving him only with Defensive. Gizka declares Offensive, and gives him both barrels.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Intimidate}
	\label{tab:Intimidate}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Intimidate & CHA vs. RS at Target’s WIL\\
\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Opponent is in a situation where they can’t defend themselves &-3 RS (does not stack with Injury)\\
Reputation (For being Tough at +1, Violent +2, Dangerous at +3) &+1 RS to +3 RS \\
Injury &-1 RS per wound level (of highest level wound, multiple wounds do not stack)\\
Loyalty (Pledged at +1, Fanatic at +3)& +1 RS to +3 RS \\
Will be hurt for divulging &+3 RS \\
In a position to harm a friend of the target &-1 RS\\
Given example of power (Show of Strength at -1, Harmed an ally/target at -2, killed an ally at -3)&-1 RS to -3 RS \\
Exploiting Fear (Minor Fear at -1, Major Phobia at -2)&-1 RS to -2 RS\\
\end{longtabu}

\paragraph{\label{skill:Navigation}Navigation}---\quad\textbf{(PER) }\par
Knowing how to get around is really important. It’s easy enough to find your way around your own home, but unfamiliar streets can be a nightmare, especially if you can’t read the street signs. Trekking through the wilderness is even worse, but at least the wilderness often has landmarks. The open seas rarely give you obvious clues as to where you are, and often, only the sun and the stars guide you. Clouds and fog can take everything away from you.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Navigation}
	\label{tab:Navigation}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Navigating a familiar area&No test required\\
Finding a notable location (a visible mountain, town hall, a major business on the main street or a port, a city connected by marked roads, a large river, etc.)&No test required\\
Finding a certain building on marked streets &PER vs. 2\\
Finding a clear path through a forest &PER vs. 4 \\
Travelling through a new area, using the stars and sun to guide you &PER vs. 6 \\
Sailing in the right direction on the open sea under an overcast night sky with a new moon &PER vs. 20 \\
\end{longtabu}

\paragraph{\label{skill:Observation}Observation}---\quad\textbf{(PER, Opposed) }\par
While anyone can look at something and see the gist of what’s going on, a good Observer can notice little details that others might miss. They can spot the errors in a forgery, notice the lone person in a sea of people who’s not moving the same way, or can see the glimmer of moonlight on the eye staring at them from a nearby bush. Cooks with particularly good Observation can tell which mulling spices were used in a cider or which breed of cattle their steak came from. A master smith can tell roughly the quality of a metal object by its heft, its color, and how it rings when hit against something.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Observation}
	\label{tab:Observation}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Notice (Sight/Hearing) &PER vs. 1\\
Search Area &PER vs. 3 \\
Finding a Needle in a Haystack &PER vs. 15 \\
Notice Concealed/Hidden Weapon &PER vs. Stealth \\
Wake up from minor disturbance (slightly creaky floorboard, whistling wind, distant thunder) &PER vs. 3 \\
Wake up from moderate disturbance (loudly squeaky door, nearby battle, far off shouting) &PER vs. 5 \\
Listening in on Whispering &1 RS per foot of distance\\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2}
Lighting (Brightly at -1, Dimly at +1, Poorly at +3) &-1 RS to +3 RS \\
Notice/Search (Trying to be Seen at -1, Shortly there at +2, Poorly Hidden at +4, Small Movement/Object at +6, Well Hidden +9) &-1 RS to +9 RS \\
Hear through Barrier (Minor Disturbance at +1, Thick Wall at +2, Battle at +4) &+1 RS to +4 RS \\
\end{longtabu}

\paragraph{\label{skill:Orate}Orate}---\quad\textbf{(CHA)}\par
Talking to a sea of people tends to be pretty intimidating. More-so if that sea of people are really angry. Orators, however, have a few tricks up their sleeve to make their message not just heard, but received well. A good Orator knows how to make a crowd sympathetic, whether by using propaganda, starting a chant to keep people from thinking, or in some cases just being sincere.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Orate}
	\label{tab:Orate}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Convince hungry desperate angry peasants that “that guy!” is responsible for their woes... &CHA vs. 1\\
And that they should hang him! &CHA vs. 3\\
And his soldiers! &CHA vs. 5 \\
And the senate! &CHA vs. 8 \\
And the king! &CHA vs. 10 \\
And make you King instead! &CHA vs. 20 \\
Convince previously convinced hungry desperate angry peasants that you AREN’T responsible for their woes.&CHA vs. 6 \\
Inspire people to act &CHA vs. 1 \\
Stir angered people to violence &CHA vs. Average WIL \\
Diffuse Violence &CHA vs. Average WIL \\
Spread Word/Lies &CHA vs Average PER \\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Disposition (Supportive at -2, Agreeable at -1, Disillusioned at +1, Hostile at +4)&-2 RS to +4 RS\\
Foreigner/Outsider&+2 RS \\
Reputation (Beloved at -2, Good at -1, Bad at +1, Horrible at +2) &-2 to +2 RS \\
Authority Figure (Lord/Ruler at -3, Respected Official at -2, Lowly Official at -1) &-3 RS to -1 RS\\
Outsider, Lower Class, Inferior, etc. &+1 to 10 RS\\
Disposition of Crowd (Supportive at -2, Agreeable at -1, Disillusioned at +1, Hostile at +4) &-2 RS to +4 RS \\
\end{longtabu}

\paragraph{\label{skill:Perform}Perform}---\quad\textbf{(Varies) }\par
Singing, dancing, acting, storytelling, and playing instruments are all ways Performers have entertained people, likely since before recorded history. The Perform skill is used for many reasons: for some, personal accomplishment is all that matters, while others use it to entertain friends and family or to keep spirits up in rough situations. Still more seek fame, glory, money, or the eyes of powerful people. Perform is separated into different specialties, like Crafting is. Singing, dancing, acting, and storytelling are all individual specialties, while most musicians specialize in a particular instrument. They may, of course, try to play an instrument they are unfamiliar with if it is similar enough to their specialty (at the GM’s discretion), but they will suffer a 1 or 2 die penalty as it’s not quite what they’re familiar with. Perform uses CHA unless a particular performance would do better with another attribute: for example, an extended solo for a brass or woodwind instrument might use END as the character cannot run out of breath, or a fast-paced or complicated dance might use AGI. Ad-libbing something is usually WIT. Perform’s RS is generally related to the audience’s expectations. A family watching a child’s recital might only expect a bare success, while an actor performing for a theater full of nobles might be expected to get 6 or even more successes.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Perform}
	\label{tab:Perform}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Can-Can with friends &CHA vs. 1 \\
Belly dancing in a decent club &AGI vs. 3 \\
Telling a riveting tale &CHA vs. 3 \\
Ad-libbing Antony and Cleopatra from a quick description &WIT vs. 6 \\
Moving Musical Performance worthy of Crowd &CHA vs. 8 \\
Flyting &Opposed WIT test \\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Disposition of Crowd (Supportive at -2, Agreeable at -1, Disillusioned at +1, Hostile at +4) &-2 RS to +4 RS\\
\end{longtabu}

\paragraph{\label{skill:Persuasion}Persuasion}---\quad\textbf{(Varies)}\par
Where intimidation uses fear to control people, Persuasion uses trust. A good Persuader either uses their own inherent trustworthiness to convince people of something, or they find someone or something else that the person is willing to trust. Obviously, most of the time the Persuader will be using CHA, although sometimes INT will come into play, or even WIT if the Persuader truly needs to think fast. They will be opposed by the Target’s own INT, WIT, or even an Observation or Research roll if the character is inclined to trust their own work over someone else’s.


\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Persuasion}
	\label{tab:Persuasion}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Persuade &CHA vs. RS at Target’s WIL \\
	\\
Gather Information &CHA vs. 1\\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Disposition (Supportive at -2, Agreeable at -1, Disillusioned at +1, Hostile at +4)&-2 RS to +4 RS\\
Scale of Knowledge (Common Knowledge at -4, Talked About at -2, Known by Few at +2, Guarded Secrets at +4) &-4 RS to +4 RS\\
Time (Very Recent -2, Recent 0, Ancient History +4) &-2 RS to +4 RS\\ 
\end{longtabu}
\paragraph{\label{skill:Politics}Politics}---\quad\textbf{(INT) }\par
Politics is a character’s knowledge of current events, from the local level to the national to all of the known world. It covers things from what nations are at war with each other, what nobles are courting others, and other, similar topics. If it makes sense for the character’s background, it can also apply to knowledge in the criminal realm.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Politics}
	\label{tab:Politics}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Know Current Events &INT vs. 1\\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Time (Very Recent -2, Recent 0, Ancient History +4) &-2 RS to +4 RS\\
Location (Home at -3, Local at -2, Nation at -1, Foreign at +2, Far Away Land at +4, Literally a different world at +6) &-3 RS to +6 RS\\
Culture (Gossipy at -2, Distrustful at +1, Secretive at +2) &-2 RS to +2 RS\\
Scale of Knowledge (Common Knowledge at -4, Talked About at -2, Known by Few at +2, Guarded Secrets at +4) &-4 RS to +4 RS\\
\end{longtabu}
\paragraph{\label{skill:Research}Research}---\quad\textbf{INT}\par
Research is all about knowing how to find something out. Someone with low-level Research skill might know how to find an alphabetically filed document or how to use a table of contents. A talented Researcher can scan through a few dozen peer-reviewed journals in a few minutes to find that one obscure article someone wrote about a really rare bone condition.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Research}
	\label{tab:Research}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Commonly Known/Recent Information&INT vs. 1
	\\
		\rowfont[c]{}Modifiers&RS Change\\\cmidrule(lr){1-2}
Materials (Well-Written at -1, Scattered Notes at +2, Poorly Understood at +4)&-1 RS to +4 RS\\
Scale of Knowledge (Common Knowledge at -4, Talked About at -2, Known by Few at +2, Guarded Secrets at +4)&-4 RS to +4 RS\\
Time (Very Recent -2, Recent 0, Ancient History +4)&-2 RS to +4 RS\\
Familiar Subject&-1 RS\\
\end{longtabu}

\paragraph{\label{skill:Riding}Riding}---\quad\textbf{Varies}\par
Sitting on the back of another creature can be surprisingly hard. Riding is how good a character is at staying on top of his mount and leading or driving it accurately. It is a very important skill for cavalrymen and knights. Riding’s uses are, for the most part, detailed in the Maneuvers section, or alongside information on different mounts.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Riding}
	\label{tab:Riding}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Simple Trick (Jump, Kneel, Call, Sit) &ADR vs. 1\\
Difficult Trick (Sharp Turn, Dangerous Jump, Dangerous Charge) &ADR vs. 3 \\
Calm &CHA vs. 3\\
	\\
		\rowfont[c]{}Modifiers &RS Change\\
Untrained&+4 RS\\
Encumberance (+1 at Light, +4 at Overloaded)&+1 to +4 RS \\
\end{longtabu}

\paragraph{\label{skill:Sailing}Sailing}---\quad\textbf{WIT}\par
Ships are more than just weighing the anchor and hoisting the sails. Sailors are expected to know a variety of knots, which sails are used for what kinds of wind, how to sail upwind, and how to prepare a ship to weather the weather. Sailing is the skill used to operate all kinds of watercraft, from a rowboat to a ship of the line. This handles the physical acts of handling heavy rigging, moving cargo, steering from the helm, and rowing. This is a sailor’s primary skill. The officers and captain of a ship usually have this skill to use in emergencies, so that they understand their own vessel, and because most of them were probably sailors before they were captains or officers, but most of their work is done with Navigation, Tactics, and Engineering, while their crew takes care of actually sailing the ship.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Sailing}
	\label{tab:Sailing}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Avoid Simple Obstacle &WIT vs. 1 RS \\
Prep to Sail &WIT vs. 1 RS \\
Traverse Hazard &WIT vs. 3 RS \\
	\\
		\rowfont[c]{}Modifiers& RS Change \\
Crew (Skilled -2, Well-staffed, Understaffed at +1, Untrained at +3) &+1 RS to +3 RS\\
Weather (Windy at +1, Storming at +3, Massive Storm at +6)&+1 to +6 RS\\
\end{longtabu}

\paragraph{\label{skill:Stealth}Stealth}---\quad\textbf{(AGI)}
Sneaking around is a good way to deal with one’s problems, as long as one’s problems have eyes and ears. Camouflage, a soft step, and an eye for shadowed areas help characters avoid detection. It is also used for hiding things. Stealth is made using AGI as an opposed check versus Observation or Hunting.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Stealth}
	\label{tab:Stealth}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Slinking without sound &AGI vs. Observation \\
Falling/Jumping without sound (max 10 feet) &AGI vs. Observation \\
Trailing without notice &AGI vs. Observation \\
Hiding with Concealment &AGI vs. Observation \\
Applying camouflage &2 RS \\
Concealing a weapon &vs. Observation \\
	\\
		\rowfont[c]{}Modifiers& RS Change\\\cmidrule(lr){1-2}
Running&+4 RS \\
Distraction&-2 RS \\
Ground Material (soft at -1, loud at +1)&+1 RS \\
Encumberance (+1 at Light, +4 at Overloaded)&+1 to +4 RS\\
\end{longtabu}

\paragraph{\label{skill:Subterfuge}Subterfuge}---\quad\textbf{(CHA)}\par
Sometimes, it’s not yourself or your possessions that need to be hidden. When you need to lie, bluff, or sometimes just to keep your mouth shut in the face of threats, torture or intoxication, Subterfuge is your skill. Subterfuge is made using CHA as an opposed check. The opposer uses any skill that would give them insight into the information being given, as determined by the GM. The player may default to $ -\frac{1}{2} $ their Observation skill if they prefer.

\begin{longtabu} to \linewidth{X[1.5]X[r]}
	\caption{Subterfuge}
	\label{tab:Subterfuge}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Lie &CHA vs. Observation\\
Present Fake Emotions &CHA vs. Observation \\
\\
	\rowfont[c]{}Modifier&Effect\\\cmidrule(lr){1-2}
Rehearsed &-1 RS \\
Believability (Very Believable at -2, Unbelieveable at +2, Near Impossible at +8) &-1 RS to +8 RS \\
\end{longtabu}

\paragraph{\label{skill:Surgery}Surgery}---\quad\textbf{(Varies)}\par
Doctors are well-respected members of society for good reason. Cleaning and bandaging a small cut is one thing, but precisely breaking the skull so that the brain doesn't get crushed by the blood filling it from a broken aneurysm without causing further damage is another.


\subparagraph{Perform Surgery:} A lengthy process generally requiring tools, the character can attempt to at the least patch up a wound. In order to stop a wound from Bleeding, and to reduce the Pain it causes, the Surgery skill must come into play. A Surgery check made at appropriate difficulty for wound level Treats the wound. Treated wounds have their Bloodloss reduced to 0 (unless reopened somehow), and their Pain reduced by half. Treated wounds begin healing immediately. Healing time depends on the character’s maximum normal HLT, and the level of the wound. Once a Wound has fully healed, the Wound, and any remaining pain, are removed from the character.\par


\begin{longtabu} to 1\textwidth {X[1.5]X[r]}
	\caption{Surgery}
	\label{tab:Surgery}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Perform Surgery&	AGI vs. Wound Level

					Stops (Removes) Bloodloss and half of Pain.

					Test done per wound.\\
Diagnosing a Wound&INT vs. 2 \\
Stopping \nameref{par:Internal Bleeding}&INT vs. 6\\
Autopsy&INT vs. 4 \\
Amputation&AGI vs. 3 \\
Remove Projectile&AGI vs. Wound Level \\
Staunching Blood Loss&INT vs. 1, -1 Bleed per success \\
	\\
		\rowfont[c]{}Modifier&Effect\\\cmidrule(lr){1-2}
Projectile Size (Javelin at -2 RS, Bullet at +2 RS)&-2 RS to +2 RS\\
Tools (Good Quality at -2, Poor Quality at +2) &-2 RS to +2 RS\\
Unsuitable environment &+2 RS\\
\caption*{Assistants take test at $ \frac{1}{2} $ RS, they give +$ -\frac{1}{2} $ Successes, to a total of the assistant’s skill level in the chosen skill, or 1 if Unskilled.}
\end{longtabu}

\paragraph{\label{skill:Swimming}Swimming}---\quad\textbf{(Varies)}\par
Most characters will be from land-bound races. Swimming is not an instinct: a character must learn to move through the water as easily as they move on land. Zells are naturally pretty good at this, but for most other people it can be difficult, especially with harsh waters or heavy equipment.

\begin{longtabu} spread 0pt {X[2]X[r]}
	\caption{Swimming}
	\label{tab:Swimming}\\
	\rowfont[c]{}Action &Base Attribute vs. RS or other effects\\\toprule
Swimming in calm water &END vs. 1\\
Level 1 rapids&END vs. 2 \\
Level 2 rapids&END vs. 3 \\
Level 3 rapids&END vs. 4 \\
Level 4 rapids&END vs. 5 \\
Level 5 rapids&END vs. 6 \\
Level 6 rapids&END vs. 10 \\
	\\
		\rowfont[c]{} Modifier & Effect\\\cmidrule(lr){1-2} 
Encumberance (+1 at Light, +4 at Overloaded)&+1 to +4 RS\\
Arm/Leg unusable&+2 RS per limb\\
Weather (Windy at +1, Storming at +3, Massive Storm at +6)&+1 to +6 RS\\
\end{longtabu}

\subparagraph{When failed: Drowning}---\quad
If the character is in danger of drowning, they must immediately make a swim check using their STR at an increasing difficulty. For each round they fail the strength check, their HLT is decreased by one until zero, where the character has drowned. The HLT lost from drowning is temporary and returns to the character once they are revived. All modifiers for swim tests apply. If a character makes a successful strength check, they can then make another normal swim check to make it the rest of the way. If they fail again, repeat the process. For Drowning, losing all of one’s HLT represents the character passing out and not their immediate death; if the character is pulled from the water within 5 minutes, they can be revived by a Medical test at RS 3; revival attempts can be made up to the drowned character’s total HLT score. If a character cannot be revived, they die. If the character is pulled out within 5 minutes plus half their normal HLT score, they can still be revived, but they must make a HLT check at RS 3 or gain the bane \nameref{bane:Brain Damage}~(major)
\setlength{\intextsep}{0pt}
\begin{wraptable}{r}{8 cm}
	\centering
	\caption{Drowning}
	\label{tab:Drowning}
		\begin{tabu}spread 0pt {X[l]X[r]}
\rowfont[c]{}Check&Modifier\\\toprule
First STR Check&No modifier.\\
2nd STR Check &+1 RS, -1 HLT\\
3rd STR Check &+2 RS, -1 HLT\\
4th STR Check &+3 RS, -1 HLT\\
\rowfont[c]{}+\ldots*&+\ldots*\\
		\end{tabu}
	\caption*{*~Continue on in like fashion until the character makes a successful strength check, or until the character reaches 0 HLT.}
\end{wraptable}
\setlength{\intextsep}{\oldintextsep}
\chapter{Social Class and Wealth}\label{ch:wealth}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Every soldier who marches to war does so with the labors of a hundred other men and women on his back. His pistol was made in Vienna, the frog for his sword was crafted in York. The fur for his cap was hunted in Russia, the cowhide for his pack was herded in Scotland, his boots were cobbled in Normandy, his sword was made in Solingen, his halberd is Swiss, his breastplate is from Milan, his codpiece is from Flanders, and his booze is from Toulouse. Your characters need to be outfitted if they’re going to survive an adventure. Outfitting costs money.

This chapter explains how the currency system of Song of Swords works, how your characters acquire 
Wealth and Assets to purchase equipment, and what their Social Class means for them.
\section{Character Creation}
A character’s Social Class represents their standing in society, which determines the amount of Wealth available to them at base (this can be modified by the GM, as well as Boons and Banes), and what their privileges and obligations are within society.


At Character Creation, you need to invest at least 1 PCP in Social Class and Wealth. Your investment is very important to your character, as it will impact the amount of equipment and influence they may have at the start of play.


The Social Classes listed in \autoref{tab:Social Classes and Wealth} are generic; every culture will have different names for these 
rankings, often multiple for each one. Confer with your GM to decide what your character’s Social Class means.
	\setlength{\intextsep}{0pt}
\begin{wraptable}{r}{7.5 cm} %THIS WRAPTABLE NEEDS THAT SPACE REMOVAL MACRO
	\centering
	\begin{tabu} spread 0pt{cXX}
PCP		&Social Class	&Wealth\\\cmidrule{1-3}
1		&Slave			&Nothing\\
2		&Peasant		&1 gp\\
3		&Poor Freeman	&3 gp\\
4		&Freeman		&5 gp, [1W]\\
5		&High Freeman	&10 gp, [2W]\\
6		&Minor Noble	&100 gp, [3W]\\
7		&Landed Noble	&250 gp, [6W]\\
8		&High Noble		&500 gp, [10W]\\
9		&Royalty		&1000 gp, [15W]\\
10		&High Royalty	&1500 gp, [20W]\\
	\end{tabu}
	\caption{Social Classes and Wealth}
	\label{tab:Social Classes and Wealth}
\end{wraptable}
	\setlength{\intextsep}{\oldintextsep}
\subsection{Social Class Benefits} Each Social Class has access to certain benefits, and you may choose several of them from the list associated with your character’s Social Class, at no further cost. As Social Class benefits are Boons of a kind, they can be recorded in the Boons and Banes section of the Character Sheet, however Social Class benefits do not affect the purchasing of Boons and Banes at Character Creation.
The benefits presented here are examples, and your GM may decide that some of these are inappropriate for your campaign, or that others should be added to the list. Confer with your GM about Social Class and its effects.

\begin{table}[hb]
	\centering
	\caption{Social Class Benefits}
	\label{Social Class Benefits}
	\begin{tabu} spread 0pt{X[-1]X}
Social Class	&Boons\\\toprule
Slave			&\textbf{Pick 1:} \nameref{boon:Hale and Hearty} (2), \nameref{boon:Beautiful} (2), \nameref{boon:Languages} (1)\\
Peasant			&\textbf{Pick 1:} \nameref{boon:Hale and Hearty} (2), \nameref{boon:Folks Back Home} (3)\\
Poor Freeman	&\textbf{Pick 1:} \nameref{boon:Hale and Hearty} (2), \nameref{boon:Folks Back Home} (3), \nameref{boon:Literate} (1)\\
Freeman			&\textbf{Pick 2:} \nameref{boon:Hale and Hearty} (2), \nameref{boon:Folks Back Home} (3), \nameref{boon:Literate} (1),\\
High Freeman	&\textbf{Pick 2:} \nameref{boon:Folks Back Home} (6), \nameref{boon:Literate} (1), \nameref{boon:Languages} (1), \nameref{boon:Contacts} (1)\\
Minor Noble		&\textbf{Pick 2:} \nameref{boon:Allies} (5), \nameref{boon:Famous} (2), \nameref{boon:Hale and Hearty} (2), \nameref{boon:Literate} (1), \nameref{boon:Languages} (1), \nameref{boon:Contacts} (1), \nameref{boon:Folks Back Home} (6)\\
Landed Noble	&\textbf{Pick 2:} \nameref{boon:Allies} (5), \nameref{boon:Famous} (3), \nameref{boon:Hale and Hearty} (2), \nameref{boon:Literate} (2), \nameref{boon:Languages} (2), \nameref{boon:Contacts} (1), \nameref{boon:Folks Back Home} (6)\\
High Noble		&\textbf{Pick 2:} \nameref{boon:Allies} (5), \nameref{boon:Famous} (4), \nameref{boon:Hale and Hearty} (2), \nameref{boon:Literate} (2), \nameref{boon:Languages} (2), \nameref{boon:Contacts} (4), \nameref{boon:Folks Back Home} (6)\\
Royalty			&\textbf{Pick 2:} \nameref{boon:Allies} (10), \nameref{boon:Famous} (4), \nameref{boon:Hale and Hearty} (2), \nameref{boon:Literate} (2), \nameref{boon:Languages} (2), \nameref{boon:Contacts} (4), \nameref{boon:Folks Back Home} (6)\\
High Royalty	&\textbf{Pick 3:} \nameref{boon:Allies} (10), \nameref{boon:Famous} (4), \nameref{boon:Hale and Hearty} (2), \nameref{boon:Literate} (3), \nameref{boon:Languages} (3), \nameref{boon:Contacts} (6), \nameref{boon:Folks Back Home} (6)\\
	\end{tabu}
\end{table}
\subsection{Abnormal Social Class and Wealth Combinations}
If your character concept involves being an unusually wealthy or poor member of your social class (say, a penniless errant prince,or an escaped prisoner who found huge wealth on a small island where he was told to search by his dying cellmate) then you can do so, using the following system:\par
First, decide on the Social Class and Wealth you would like your character to have, and their individual PCP cost (for example, a Landed Noble Social Class [7] and the Wealth of a Peasant [2]). Combine the two costs, and divide the result by 2, rounding up. 

\emph{This is the only time in the game in which you round up.}

This is the total PCP cost for your character’s Social Class and Wealth. You gain all of the benefits of your Social Class, and all of the benefits of your Wealth. Examples are provided in the following table. For their potentially unbalancing nature, all abnormal Social Class and Wealth combinations are subject to GM approval.

\begin{table}[hb]
	\centering
	\caption{Abnormal Social Class and Wealth Combinations}
	\label{tab;Abnormal Class/Wealth Combinations}
	\begin{tabu} spread 0pt {X[l]X[R]X[R]X[-1,c]}
Concept						&Social Class		&Wealth				&PCP \\\toprule
Pauper Prince				&Royalty -- (8)		&0 gp -- (1)		&5\\
Poor Knight					&Minor Noble -- (6)	&5 gp [1W] -- (4)	&5\\
Incredibly Lucky Peasant	&Peasant -- (2)		&100 gp [3W] -- (6)	&4\\
Janissary					&Slave -- (1)		&10 gp [2W] -- (5)	&3\\
	\end{tabu}
\end{table}
\section{Wealth}
In reality, currency is a tricky business, particularly when you're talking about multiple eras, kingdoms, 
languages and trade routes.

To keep things from becoming too labyrinthine, Song of Swords uses a generic wealth system, with each item assigned a value in a currency system based on the Carolingian Pound, a system developed by Charlemagne that was used in Europe for nearly 1,000 years in one form or another. This system is divided into three denominations of currency: Copper, Silver and Gold, representing (in English terms) the Denarius, the Shilling, and the Pound, the lattermost of which was worth (theoretically) a ‘pound’ of silver.

This system may be familiar to players of other RPGs, which have drawn upon similar inspiration. We adopt it because of that familiarity, and because the bulk of materials priced in historical documents from Medieval Europe are either already in, or can be easily translated to the Carolingian system.

If you prefer a local currency for your campaign, fear not! Conversion tables for older, newer, or foreign currencies will be provided in upcoming Song of Swords Culture Expansions.

\begin{table}[hb]
	\centering
	\caption{Currency Value}
	\label{tab:Currency Value}
	\begin{tabu} spread 0pt {XXX}
Currency &Worth &Historical value\\\toprule
1 gold piece (gp) &20 sp &1 Pound\\
1 silver piece (sp) &12 cp &1 Shilling\\
1 copper piece (cp) &A loaf of black bread &1 Denarius\\
	\end{tabu}
\end{table}

\section{Assets}
Assets represent less tangible sources of wealth than simple piles of money. Most people, particularly the 
wealthy, keep most of their wealth in the form of lands, cattle, business and debt, rather than cold hard cash. 

Assets come in three levels: Minor, Moderate and Major. You can buy Assets with money for 150\% of 
their Liquidated Value, at Character Creation, or through the events of a campaign. 

\begin{table}[hb]
	\centering
	\caption{Assets}
	\label{tab:Assets}
	\begin{tabu} spread 0pt {X[-1,l]X[-2,c]X[c]X[4]}
\rowfont[c]{}Asset Level &Withdraw Amount&Liquidated Value&Example\\\toprule
Minor &20 sp &6 gp (120 sp) &Small Farm, Small Business, Small Cattle Herd \\
Moderate &40 sp &12 gp (240 sp) &Large Farm, Respectable Business, Large Cattle Herd \\
Major &60 sp &18 gp (360 sp) &Huge Farm, Powerful Guild, Large Coaching House, Massive Cattle Herd, Bank\\
	\end{tabu}
\end{table}

Once per session (or month, whichever comes first), a character can ‘withdraw’ Wealth from his assets, 
essentially picking up the profits for his personal use. This has no consequences, and assumes that enough 
of the funds have been cycled back into the Asset to keep it running. 

A character can also liquify an asset in order to convert it directly into Wealth, permanently exhausting it, 
but extracting a large sum of cash from it. This removes the Asset from the character’s control—or at 
least renders it null until it is revitalized through re-investment (the same as purchasing a new asset of the 
same value).

The act of withdrawing Wealth from an Asset generally takes from several hours to a month, depending 
on the distance from the place and the ease of having the money (or letters of credit) sent over. 
Liquefaction can take a longer or shorter period of time, at the GM’s discretion. 

In the above \autoref{tab:Social Classes and Wealth}, the number in brackets is a character’s Wealth, which can be used either to acquire additional money or assets, as detailed in \autoref{tab:Wealth/Asset/Gold Chart}, below. These can be purchased multiple times, and Wealth must be spent at Character Creation.

\begin{table}[hb] %THIS table seems superfluous, merge with Assets table? Either that or make it a wraptab
	\centering
	\caption{Wealth/Asset/Gold Chart}
	\label{tab:Wealth/Asset/Gold Chart}
	\begin{tabu} spread 0pt {X[-2,c]X[2]X[-1]}
\rowfont[c]{}Wealth Spent &Asset Gained &Gold Gained\\\toprule
1 &Minor Asset &6 gp\\
2 &Moderate Asset &12 gp\\
3 &Major Asset &18 gp\\
	\end{tabu}
\end{table}
\part{Equipment}
\chapter{Equipment}\label{ch:equipment}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quotation}
	\emph{“Imagine walking five hundred miles over the course of two weeks, carrying an arquebus, a bardiche, a stone of grain, another stone of water, ten pounds of shot, your own armor, your tent, whatever amenities you want for yourself, and your lord’s favorite dog. In the rain. In winter. With dysentery. Alright, are you imagining that?}

	\emph{Now imagine that as soon as you’re done with that, you need to actually fight the enemy. You have a horse, but a senator’s nephew is riding it. You’re knee-deep in mud, and you’ve just been assigned a new recruit to train. He speaks four languages, none of which are yours, and has something to prove. Now he’s drunk and arguing with your superiors, you haven’t slept in thirty hours, you’ve just discovered that the fop nephew has broken your horse’s leg in a gopher hole, and your gun’s wheellock is broken.}

	\emph{At that moment, the beating of war-drums echoes from the darkness. Someone screams, and a cannonball lands in your cooking fire, where you were drying your boots.}
		
	\emph{Welcome to war. Enjoy your stay.”}

		\hfill--\textbf{Mago Straddock}, Dacian Volkodav
\end{quotation}
Your character may be the most feared swordsman in the lands, but without his weapon, he is just and 
only a man—and probably a soon-to-be-dead man, at that. 

Equipping your character is very important. Not only will they need a weapon (or five!), but they may 
also need armor, food and drink, transport, medical supplies, tools for survival, and even slaves. This 
chapter outlines all that you need to know about weaponry, armor, goods and services.

If you need to learn more about gold and Wealth, refer to \autoref{ch:wealth}. Information on carrying your equipment and what that means for your character’s Fatigue is explained in \autoref{sec:exhaustion}:~\nameref{sec:exhaustion}
\chapter{Melee Weapons}\label{ch:Melee Weapons}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quote}
	\emph{“Invincible Khan. We have exhausted our supply of arrows. We are out of water and the men are dying of thirst. We are running out of horses. The rats are without number. They give us no sleep and they may soon overtake us. We have killed many thousands but we require aid. Your loyal servant begs you for aid."}
	
	\hfill -- Missive from \textbf{Noyan Tsubokai} (Deceased) to \textbf{Gunsukh Khan}\par
	\hfill Intercepted by Bangle Guard riders near Pishka 
\end{quote}
Melee weapons are those intended for close-range and in hand-to-hand combat. Any character who expects to enter 
combat should carry a melee weapon. 

Below you will find a full list of these weapons, as well as information on each of their qualities.

\subsection{Which Weapon for Me?}
Let’s be serious. You probably flipped straight to the weapon section as soon as you found out there was one; 
everyone loves weapons! But you probably weren’t expecting there to be this many. Song of Swords has over 150 
weapons, many of which can be used in combination with shields, parrying daggers, or other weapons. 

You may be thinking: ‘So how am I supposed to pick a weapon?’ Well, don’t worry. That’s what this little section is 
for.

First, remember that you’re not limited to just one. A well-equipped warrior may carry a spear, a shield, a sword and a dagger, and expect to use any or all of them in combat. It’s good to have options, and it’s handy to have weapons of varying Reach, so that you never end up trying to fight at knife-range with a pike.

Here are some general tips for choosing a weapon:
\subsubsection{Attack TN}
Weapons with low Attack TNs (displayed in this chapter as ‘Swing TN’ and ‘Thrust TN’) are easier to hit with, and 
weapons with higher Attack TNs are harder to hit with. Remember, a TN of 8 means that statistically, you’ll succeed 
with 3 out of every 10 dice you roll (8, 9, and 10) whereas a TN of 7 means you’ll have an average of 4 Successes, a 
TN of 6 means 5 average Successes, and so on. Plan accordingly! BS carry directly into damage with most attacks.
\subsubsection{Attack Damage}
One of the counterpoints to TN is attack damage. Most weapons with good TNs do not inflict huge amounts of 
damage; typically, large and clunky weapons are harder to land hits with. Weapons with high damage modifiers can be
more useful in some situations, particularly on the battlefield, so you can crack open enemy armor.
\subsubsection{Defense TN / Guard}
If you have a shield, your weapon’s Defense TN isn’t a huge priority, but if you’re fighting with just the weapon, 
consider picking a weapon with a good Defense TN in order to parry attacks. Also remember that shields can be 
beaten away or even destroyed by the enemy; you don’t want to be left out in the cold. The Guard value of the weapon
is also important for protecting your weapon-hand, which is a popular target when left undefended.
\subsubsection{Reach}
Reach dictates the terms of engagement more than most other factors. Generally speaking, if two fighters of similar 
skill have mismatched Reach, the one with the longer Reach will win. This is not universally true, but it’s a fair 
enough guideline that you should account for it. However, some weapons are so long that they become very clumsy 
when used in close combat, like pikes. Additionally, the longer your weapon, the more penalties you suffer if the 
enemy closes within your Reach. Consider keeping a sidearm of shorter Range alongside a weapon with longer 
Reach, or be prepared to close past an enemy’s longer weapon with a short one of your own.
\subsubsection{Special}
The special rules of a weapon can completely change its uses. Always pay close attention to the ‘special’ section of the
weapon, as Draw, Multi-Hit, Hook, and other special considerations can give you an edge over your enemies.

\subsubsection{How Many Weapons?}
You can carry multiple weapons! Don’t think you have to limit yourself to just one. Also, be very careful to read the 
Proficiencies that each weapon uses, as some Proficiencies give access to good Maneuvers earlier than others!

\subsection{Melee Weapon Characteristics}
\subsubsection{Weapon Name}
The name of the weapon.
\subsubsection{Weapon Type}
The type of the weapon, and thus the Proficiency that it can be used with.
\subsubsection{Hands}
This specifies the number of hands required to use the weapon, either one (1H) or two (2H). 
\subsubsection{Reach}
The weapon’s Reach, as \autoref{tab:Reach} describes. Reach has important effects in combat, which are further explained in \autoref{ch:combat}, \autoref{sec:Reach}
\begin{table}[!ht]
	\centering
	\caption{Reach}
	\label{tab:Reach}
	\begin{tabu} spread 0pt {cX[c]X[c]X}
Number	&\multicolumn{2}{>{\centering\setlength\hsize{2\hsize}}c}{Stages of Reach}&Example\\\toprule
1 		&Half-Arm 			&HA				&Knee, Elbow, Grapple\\
2		&Hand				&H				&Dagger, Punch\\
3		&Short 				&S				&Short Sword, Kick\\
4		&Medium				&M				&Arming Sword\\
5		&Long				&L				&Poleaxe, Longsword\\
6		&Very Long 			&VL				&Arming Spear\\
7		&Extremely Long 	&EL				&Halberd\\
8		&Ludicrously Long	&LL				&Pike\\
	\tabuphantomline
	\end{tabu}
\end{table}
\subsubsection{Swing}
Swing represents the weapon’s use for swinging attacks. The first number is the Swing TN, the number in parenthesis 
is the weapon’s base damage modifier, and the type of damage (either cutting, piercing or bludgeoning) is represented 
by a letter (c, p or b, respectively). 
\subsubsection{Thrust}
Thrust represents the weapon’s use for thrusting attacks. The first number is the Thrust TN, the number in parenthesis 
is the weapon’s base damage modifier, and the type of damage (either cutting, piercing or bludgeoning) is represented 
by a letter (c, p or b, respectively). 
\subsubsection{Defense / Guard}
The first number represents the weapon’s Defense TN, against which defenses that use the Parry TN are rolled. The 
number in parenthesis is the weapon’s Guard value. The Guard value is applied as Hard AV against any attacks that hit
the weapon-hand of the wielder while the weapon is being held normally. This AV layers with any actual armor worn 
on the weapon-hand. The kind of Guard it represents depends on the Guard value
\begin{itemize}
	\item [0:] No Guard (Shaska, khyber knife)
	\item [1:] Small Guard (Katana, Smallsword)
	\item [2:] Crossguard (Arming Sword, Messer, Falchion, Longsword)
	\item [3:] Complex Guard (Rapier, Sidesword)
	\item [4:] Basket Hilt (Schiavona, Heavy Backsword)
\end{itemize}
Note that only swords with Guard Value 2 (neither lower nor higher) can use the \nameref{man:Murder-Strike} Maneuver.
\subsubsection{Special}
The weapon’s special qualities, which are further described in \autoref{sec:weapons-special-qualities}
\subsubsection{Weight}
The weight of the weapon added to Encumbrance when sheathed, holstered or stored. When being held, a weapon has 
no Weight value.
\subsubsection{Cost}
The weapon’s cost. May vary regionally.

\section{Daggers}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{Daggers}\\
	\label{tab:Daggers List}\\\endfirsthead
Weapon Name						&Weapon Type	&Hands&Reach&Swing&Thrust&Defense (Guard)&Special						&Weight	&Cost\\\toprule\endhead
Ballock Dagger					&Dagger			&1H	&H	&7(-1c)	&6(+1p)	&8(1)	&Fluid Thrusts, Light Blade				&0	&5 cp\\
Baselard						&Dagger			&1H	&S	&6(+0c)	&6(+0p)	&7(1)	&Fluid Thrusts							&0	&2 sp\\
Bowie Knife/Pishkan Toothpick	&Dagger			&1H	&S	&6(+0c)	&6(+1p)	&8(1)	&-										&0	&10 cp\\
Cinquedea						&Dagger, Sword	&1H	&S	&6(+0c)	&7(+2p)	&8(0)	&-										&0	&3 sp\\
Degan							&Dagger, Sword	&1H	&S	&6(-1c)	&6(+1p)	&7(0)	&Fluid Thrusts, Light Blade				&0	&1 sp\\
Dirk							&Dagger			&1H	&H	&7(-1c)	&6(+0p)	&7(0)	&Fluid Thrusts, Light Blade				&0	&4 cp\\
Khyber Knife					&Dagger, Sword	&1H	&S	&7(+0c)	&7(+1p)	&7(0)	&Draw 2, Thin Blade						&0	&10 cp\\
Knife (Large)					&Dagger			&1H	&S	&6(-1c)	&6(+0p)	&9(0)	&Draw 1, Light Blade					&0	&7 cp\\
Knife (Small)					&Dagger			&1H	&H	&6(-2c)	&6(-1p)	&9(0)	&Draw 1, Light Blade, Thin Blade		&0	&6 cp\\
Kukri							&Dagger			&1H	&S	&6(+1c)	&7(+0p)	&9(0)	&Draw 2, Light Blade, Forward Swept		&0	&1 sp\\
Main Gauche						&Dagger			&1H	&H	&7(-1c)	&7(+0p)	&6(3)	&Companion Weapon, Light Blade			&0	&3 sp\\
Misericorde						&Dagger			&1H	&H	&9(-2c)	&5(-2p)	&10(0)	&Fluid Thrusts, Light Blade, Thin Blade	&0	&3 cp\\
Navaja							&Dagger			&1H	&H	&6(+0c)	&7(-1p)	&8(0)	&Draw 2, Light Blade					&0	&6 cp\\
Poignard						&Dagger			&1H	&H	&6(-1c)	&6(+0p)	&8(1)	&Fluid Thrusts, Light Blade, Thin Blade	&0	&6 cp\\
Pugio							&Dagger			&1H	&H	&7(-1c)	&6(+1p)	&7(0)	&Light Blade, Spatulate Tip 2			&0	&6 cp\\
Puukko							&Dagger			&1H	&H	&6(-1c)	&6(-0p)	&9(0)	&Fluid Thrusts, Light Blade				&0	&4 cp\\
Rondel Dagger					&Dagger			&1H	&S	&7(-1c)	&6(+1p)	&7(1)	&Fluid Thrusts, Light Blade, Thin Blade	&0	&6 cp\\
Short Katzbalger				&Dagger, Sword	&1H	&S	&7(+0c)	&7(+0p)	&7(1)	&Spatulate Tip 2						&0	&6 sp\\
Stiletto						&Dagger			&1H	&H	&7(-1c)	&6(+0p)	&8(1)	&Fluid Thrusts, Light Blade, Thin Blade	&0	&8 cp\\
Swordbreaker					&Dagger			&1H	&S	&6(-1c)	&6(+0p)	&7(1)	&Companion Weapon, Parrying Teeth		&0	&1 sp\\
Tanto							&Dagger			&1H	&H	&7(+0c)	&6(+0p)	&8(1)	&Draw 1, Light Blade, Thin Blade		&0	&3 sp\\
\end{longtabu}

\section{Swords}
\subsection{1H Swords}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{One-Handed Swords}
	\label{tab:1H Swords}\\\endfirsthead
Weapon Name					&Weapon Type    &Hands   &Reach  &Swing  &Thrust &Defense (Guard)  &Special                                 &Wt  &Cost\\\toprule\endhead
Arming Sword (Early)        &Sword          &1H      &M      &7(+1c) &7(+0p) &7(1)           &Draw 2, Spatulate Tip 2                 &0.5 &1 gp\\
Arming Sword (Chivalric)    &Sword          &1H      &M      &7(+2c) &7(+0p) &7(2)           &-                                       &0.5 &1 gp\\
Arming Sword (Late)         &Sword          &1H      &M      &7(+0c) &7(+2p) &7(2)           &-                                       &0.5 &1 gp\\
Backsword (Standard)        &Sword          &1H      &M      &7(+1c) &7(+1p) &7(4)           &-                                       &0.5 &1 gp\\
Backsword (Heavy)           &Sword          &1H      &L      &7(+2c) &8(+1p) &8(3)           &-                                       &0.5 &25 sp\\
Bastard Sword*              &Sword          &1H      &L      &7(+0c) &7(+1p) &8(2)           &Heavy Weapon                            &1   &50 sp\\
Cinquedea                   &Dagger, Sword  &1H      &S      &6(+0c) &7(+2p) &8(0)           &-                                       &0   &3 sp\\
Claymore (Basket Hilted)	&Sword		&1H      &M      &7(+1c) &7(+0p) &7(4)           &Tight Grip 2                           &0.5 &1 gp\\
Cutlass						&Sword          &1H      &S      &7(+1c) &8(+0p) &7(3)           &-                                       &0.5 &1 sp\\
Degan&Dagger, Sword  &1H      &S      &6(-1c) &6(+1p) &7(0)           &Fluid Thrusts, Light Blade              &0   &1 sp\\
Dussack                     &Sword          &1H      &S      &7(+1c) &7(-1p) &7(3)           &Draw 2, Tight Grip 1                    &0.5 &1 sp\\
Espada Ropera               &Sword          &1H      &L      &7(+0c) &7(+1p) &7(3)           &-                                       &0.5 &25 sp\\
Falcata                     &Sword          &1H      &S      &6(+1c) &8(-1p) &8(3)           &Forward Swept                           &0.5 &1 gp\\
Falchion                    &Sword          &1H      &M      &7(+2c) &8(+0p) &8(2)           &-                                       &0.5 &18 cp\\
Gladius                     &Sword          &1H      &S      &7(+1c) &7(+1p) &7(0)           &-                                       &0   &1 sp\\
Karthacki Dueling Sword     &Sword          &1H      &M      &6(-1b) &6(+0p) &7(0)           &Fluid Thrusts, Light Blade, Thin Blade  &0.5 &10 sp\\
Katana*                     &Sword          &1H      &M      &7(+1c) &8(+1p) &8(1)           &Draw 3, Heavy Weapon                    &1   &50 sp\\
Katzbalger                  &Sword          &1H      &S      &7(+1c) &8(+0p) &7(2)           &Spatulate Tip 2                         &0.5 &4 sp\\
Katzbalger (Short)          &Dagger, Sword  &1H      &S      &7(+0c) &7(+0p) &7(1)           &Spatulate Tip 2                         &0   &6 sp\\
Khyber Knife                &Dagger, Sword  &1H      &S      &7(+0c) &7(+1p) &7(0)           &Draw 2, Thin Blade                      &0   &10 cp\\
Kilij                       &Sword          &1H      &M      &7(+1c) &8(+0p) &7(2)           &Draw 2, Cavalry Sword                   &0.5 &30 sp\\
Koncerz                     &Sword          &1H      &VL     &7(-2c) &8(+2p) &8(3)           &AP Thrust 2, Couched Charge             &0.5 &25 sp\\
Khopesh                     &Sword, Blunt   &1H      &S      &6(+1c) &9(-1p) &8(0)           &Forward Swept, Hook                     &0.5 &1 sp\\
Kriegsmesser*               &Sword          &1H      &M      &7(+1c) &8(+0p) &8(2)           &Draw 2, Heavy Weapon                    &1   &10 sp\\
Langes Messer               &Sword          &1H      &M      &7(+1c) &7(+0p) &8(2)           &Draw 2                                  &0.5 &1 sp\\
Longsword*                  &Sword          &1H      &L      &7(+1c) &7(+0p) &8(2)           &Heavy Weapon                            &1   &50 sp\\
Pallasch                    &Sword          &1H      &L      &7(+1c) &8(+1p) &7(3)           &Cavalry Sword                           &0.5 &35 sp\\
Rapier                      &Sword          &1H      &L      &6(-1c) &6(+0p) &7(3)           &Fluid Thrusts, Thin Blade               &0.5 &30 sp\\
Saber (Early)               &Sword          &1H      &M      &7(+0c) &7(+0p) &8(1)           &Draw 2, Cavalry Sword                   &0.5 &10 sp\\
Saber (Hussar)              &Sword          &1H      &M      &7(+1c) &7(-1p) &7(3)           &Draw 3, Cavalry Sword                   &0.5 &25 sp\\
Saber (Karabela)            &Sword          &1H      &M      &7(+0c) &8(-1p) &7(3)           &Draw 4, Cavalry Sword                   &0.5 &30 sp\\
Saber (Zellish)             &Sword          &1H      &M      &7(+1c) &8(+0p) &8(3)           &Draw 2, Forward Swept                   &0.5 &15 sp\\
Schiavona                   &Sword          &1H      &L      &7(+0c) &7(+0p) &7(3)           &Tight Grip 2                            &0.5 &1 gp\\
Scimitar                    &Sword          &1H      &M      &7(+0c) &7(-1p) &7(2)           &Draw 2, Cavalry Sword                   &0.5 &15 sp\\
Scythe-Sword                &Sword          &1H      &S      &7(+1c) &9(-1p) &8(1)           &Draw 2, Forward Swept                   &0.5 &16 cp\\
Shaska                      &Sword          &1H      &M      &7(+0c) &7(+0p) &7(0)           &Draw 2, Cavalry Sword                   &0.5 &15 sp\\
Short Sword                 &Sword          &1H      &S      &7(+0c) &6(+1p) &7(1)           &-                                       &0   &1 sp\\
Sidesword                   &Sword          &1H      &L      &7(+1c) &7(+1p) &7(3)           &-                                       &0.5 &15 sp\\
Smallsword                  &Sword          &1H      &M      &6(-2c) &6(+0p) &7(3)           &Fluid Thrusts, Light Blade, Thin Blade  &0.5 &30 sp\\
Spadroon                    &Sword          &1H      &M      &6(-1c) &6(+0p) &7(3)           &Fluid Thrusts, Light Blade              &0.5 &30 sp\\
Tesshake                    &Sword          &1H      &S      &7(+1c) &7(+0p) &8(3)           &-                                       &0.5 &1 sp\\
Tulwar                      &Sword          &1H      &S      &7(+0c) &7(-2p) &8(2)           &Draw 4, Tight Grip 2, Cavalry Sword     &0.5 &25 sp\\
Urumi                       &Sword          &1H      &VL     &7(+0c) &NA     &9(1)           &Chain 3, Draw 2                         &0   &30 sp\\
Wakizashi                   &Sword          &1H      &S      &7(+0c) &8(+0p) &7(1)           &Draw 1                                  &0   &10 sp\\
War Cleaver                 &Sword          &1H      &M      &7(+2c) &8(-1p) &8(0)           &Heavy Weapon                            &0.5 &6 cp\\
Yataghan                    &Sword          &1H      &S      &7(+1c) &8(+0p) &7(0)           &Forward Swept, Draw 1                   &0.5 &10 sp\\
\end{longtabu}

\subsection{2H Swords}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{Two-handed Swords}\\
	\label{tab:2H Swords}\\\endfirsthead
Weapon Name					&Weapon Type	&Hands	&Reach	&Swing	&Thrust	&Defense (Guard)	&Special						&Wt		&Cost\\\toprule\endhead
Bastard Sword				&Sword			&2H		&L		&7(+1c)	&7(+2p)	&7(2)				&Hand-Off, Heavy Weapon 			&1		&50 sp\\
Claymore					&Sword			&2H		&L		&7(+2c)	&7(+0p)	&7(2)	&Heavy Weapon &2 &50 sp\\
Estoc						&Sword 			&2H		&L		&8(+0b)	&6(+2p)	&7(2)	&AP Thrust 1, Thin Blade, Fluid Thrusts			&1		&45 sp\\
Flammenschwert				&Sword			&2H		&VL		&7(+2c)	&8(+3p)	&7(3)	&Draw 1, Crushing 1, Heavy Weapon				&2		&4 gp\\
Godenak &Sword &2H &M &7(+3c) &8(-1p) &8(0) &AP Swing 1, Heavy Weapon &1 &2 sp\\
Grosses Messer &Sword &2H &L &7(+2c) &8(+0p) &8(1) &Draw 3, Heavy Weapon &2 &1 gp\\
Katana 					&Sword &2H &M 	&7(+2c) &7(+1p) &7(1) &Draw 4, Hand-Off, Heavy Weapon 			&1 	&50 sp\\
Katzbalger (Large) 		&Sword &2H &L 	&7(+2c) &8(+0p) &7(2) &Spatulate Tip 2, Heavy Weapon 				&1 	&2 gp \\
Kriegsmesser 			&Sword &2H &M 	&7(+2c) &7(+0p) &7(2) &Draw 3, Hand-Off, Heavy Weapon 			&1 	&10 sp\\
Longsword 				&Sword &2H &L 	&7(+2c) &7(+1p) &7(2) &Hand-Off, Heavy Weapon 					&1 	&50 sp\\
Massive Godenak 		&Sword &2H &VL	&7(+4c) &8(+0p) &8(0) &AP Swing 2, Heavy Weapon, Freakishly Large	&	&\\    
Montante 				&Sword &2H &VL &7(+2c) &7(+1p) &7(2) &Heavy Weapon 								&2 	&4 gp\\
No-Dachi 				&Sword &2H &L 	&7(+2c) &8(+0p) &8(1) &Draw 3, Heavy Weapon 						&1 	&3 gp\\
Swiss Saber (Schnepf) 	&Sword &2H &L 	&7(+1c) &7(+0p) &7(3) &Draw 3, Heavy Weapon 						&1 	&45 sp\\
Zweihander 				&Sword &2H &VL &7(+3c) &8(+2p) &7(2) &Heavy Weapon 								&2 	&75 sp\\
Zweihander (Grutte) 	&Sword &2H &EL &7(+4c) &8(+3p) &7(2) &Heavy Weapon, Freakishly Large 			&3 	&6 gp\\
\end{longtabu}

\section{Blunt}
\subsection{1H Blunt}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{One-Handed Blunt Weapons}\\
	\label{tab:1H Blunt}\\\endfirsthead
Weapon Name						&Weapon Type	&Hands&Reach&Swing&Thrust&Defense (Guard)&Special						&Weight	&Cost\\\toprule\endhead
Battleaxe* &Blunt &1H &M &7(+2c) &7(-2b) &9(0)& AP Swing 1, Hand-Off, Hook, Thrusting Slot, Shock 2, Swinging Slot &1 &6 sp\\
Bar Mace &Blunt &1H &S &7(+1b) &7(-2b) &7(0)& AP Swing 1, Shock 1, Thrusting Slot &0 &1 sp\\
Club (Short)/Truncheon &Blunt &1H &S &6(+0b) &7(-2b) &6(0)& Swinging Slot, Thrusting Slot &0 &1 cp\\
Club (Medium)/Rod &Blunt &1H &M &7(+0b) &7(-1b) &6(0)& Swinging Slot, Thrusting Slot &0 &1 cp\\
Double-Headed Flail &Blunt &1H &M &8(+1b) &8(-2b) &10(0)& Chain 1, Multi-Hit Swing 2, Shock 2 &1 &9 cp\\
Hand Axe/ Hatchet &Blunt &1H &S &7(+2c) &7(-2b) &8(0)& AP Swing 1, Hook, Shock 1, Swing Slot, Thrusting Slot &0 &6 cp\\
Khopesh Sword, &Blunt &1H &S &6(+1c) &9(-1p) &8(0)& Forward Swept, Hook 0.&5 &1 sp\\
Light Flail &Blunt &1H &M &7(+2b) &8(-2b) &9(0)& Chain 2, Shock 2 &1 &6 sp\\
Light Mace (Flanged) &Blunt &1H &S &7(+2b) &7(-2b) &7(0)& AP Swing 2, Crushing 1, Shock 2, Thrusting Slot &0 &3 sp\\
Light Mace (Knobbed) &Blunt &1H &S &7(+2b) &7(-2b) &7(0)& AP Swing 1, Shock 2, Thrusting Slot &0 &4 cp\\
Light Mace (Spiked) &Blunt &1H &S &7(+2b) &7(-2b) &7(0)& Bleed 2, Shock 2, Thrusting Slot &0 &8 cp\\
Tabarzin* &Blunt &1H &L &7(+2c) &8(+0p) &9(0)& AP Swing 1, Hand-Off, Hook, Shock 1, Swinging Slot, Thrusting Slot &2 &10 sp\\
Warhammer (Standard) &Blunt &1H &S &7(+2b) &7(-2b) &8(0)& AP Swing 1, Crushing 1, Reverse Slot, Shock 1, Thrusting Slot &0 &6 cp\\
War Pick &Blunt &1H &S &7(+2p) &7(-2b) &8(0)& AP Swing 2, Hook, Reverse Slot, Thrusting Slot, Thin Blade &0 &2 sp\\
	\\
	\\
		\multicolumn{10}{c}{1H Blunt Weapon Attachments}\\\tabucline{1-10}
Arming Spike	&Blunt	&1H	&-	&-		&(+0p)	&-		&Weapon Attachment (Thrust) &0 &1 cp\\
Reverse Blade	&Blunt	&1H	&-	&7(+1c)&-	&-		&Weapon Attachment (Swing) &0 &6 cp\\
Reverse Hammer	&Blunt	&1H	&-	&7(+0b)&-	&-		&AP Swing 1, Crushing 1, Shock 1, Weapon Attachment (Swing) &0 &6 cp\\
Reverse Spike	&Blunt	&1H	&-	&7(+0p)&-	&-		&AP Swing 2, Hook, Weapon Attachment (Swing) &0 &6 cp\\
Weighted Chain	&Blunt	&1H	&VL	&8(+1b)&-	&8(0)	&Chain 3, Hook,Shock 2, Weapon Attachment (Swing) &1 &6 cp\\
\end{longtabu}
%BEGIN_FOLD%For the 1H Blunt attachment tabu outside the longtabu environment
%\subsubsection{1H Blunt Weapon Attachments}%MAKE THIS A SUBTABLE OF THE ABOVE TABLE
%\begin{tabu} to \linewidth {X[3] X *{5}{X[-3,c]} X[5] *{2}{X[-2]}}
%Weapon Name						&Weapon Type	&Hands	&Reach&Swing&Thrust&Defense (Guard)&Special						&Weight	&Cost\\
%Arming Spike	&Blunt	&1H	&-	&-		&(+0p)	&-		&Weapon Attachment (Thrust) &0 &1 cp\\
%Reverse Blade	&Blunt	&1H	&-	&7(+1c)&-	&-		&Weapon Attachment (Swing) &0 &6 cp\\
%Reverse Hammer	&Blunt	&1H	&-	&7(+0b)&-	&-		&AP Swing 1, Crushing 1, Shock 1, Weapon Attachment (Swing) &0 &6 cp\\
%Reverse Spike	&Blunt	&1H	&-	&7(+0p)&-	&-		&AP Swing 2, Hook, Weapon Attachment (Swing) &0 &6 cp\\
%Weighted Chain	&Blunt	&1H	&VL	&8(+1b)&-	&8(0)	&Chain 3, Hook,Shock 2, Weapon Attachment (Swing) &1 &6 cp\\
%\end{tabu}
%END_FOLD
\subsection{2H Blunt}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{Two-Handed Blunt}
	\label{tab:2H Blunt}\\\endfirsthead
Weapon Name				&Weapon Type	&Hands	&Reach	&Swing&Thrust&Defense (Guard)&Special																	&Weight	&Cost\\\toprule\endhead
Battleaxe				&Blunt			&2H		&M		&7(+3c)&7(-2b)&8(0)&AP Swing 1, Hand-Off, Heavy Weapon, Hook, Thrusting Slot, Shock 2, Swinging Slot	&1 &6 sp\\
Club (Large) 			&Blunt			&2H		&M		&6(+1b)&7(-1b)&7(0)&Heavy Weapon, Shock 1, Swinging Slot, Thrusting Slot								&1 &3 cp\\
Dane Axe 				&Blunt			&2H		&L		&7(+4c)&8(+0p)&8(0)&AP Swing 2, Crushing 2, Heavy Weapon, Hook, Shock 2, Swinging Slot, Thrusting Slot	&2 &7 sp\\
Goedendag 				&Blunt, Polearm	&2H		&L		&7(+3b)&8(+2p)&8(0)&AP Swing 2, AP Thrust 2, Crushing 1, Heavy Weapon, Shock 2							&2 &1 sp\\
Gunstock Club 			&Blunt			&2H		&M		&7(+2b)&7(-1b)&7(0)&AP Swing 1, Heavy Weapon, Shock 1, Swinging Slot									&1 &7 cp\\
Heavy Mace (Flanged)	&Blunt			&2H		&M		&7(+3b)&7(-2b)&7(0)&AP Swing 2, Crushing 2, Heavy Weapon, Shock 1										&1 &6 sp\\
Heavy Mace (Knobbed)	&Blunt			&2H		&M		&7(+3b)&7(-2b)&7(0)&AP Swing 1, Heavy Weapon, Shock 4													&1 &1 sp\\
Heavy Mace (Spiked) 	&Blunt			&2H		&M		&7(+3b)&7(-2b)&7(0)&AP Swing 1, Bleed 3, Crushing 1, Heavy Weapon										&1 &2 sp\\
Morningstar 			&Polearm, Blunt	&2H		&L		&7(+3b)&8(+1p)&8(0)&AP Swing 2, Bleed 3, Heavy Weapon													&5 &1 sp\\
Poleaxe 				&Polearm, Spear, Blunt &2H&L	&7(+2c)&7(+2p)&7(0)&AP Swing 1, AP Thrust 1, Heavy Weapon, Shock 2, Swinging Slot,Thin Blade			&5	&8 sp\\
Pole-Flail				&Blunt, Polearm	&2H		&L		&7(+2b)&7(-2b)&8(0)&AP Swing 2, Chain 2, Crushing 2, Heavy Weapon, Shock 2								&2 &2 sp\\
Tabarzin				&Blunt			&2H		&L		&7(+3c)&8(+0p)&8(0)&AP Swing 2, Hand-Off, Heavy Weapon, Hook, Shock 2, Swinging Slot, Thrusting Slot 	&2 &10 sp\\
Warhammer (Heavy)		&Blunt			&2H		&M		&7(+2b)&7(-2b)&8(0)&AP Swing 1, Crushing 2, Heavy Weapon, Shock 2, Swinging Slot, Thrusting Slot		&1	&1 sp\\
War Maul				&Blunt			&2H		&M		&7(+3b)&8(-2b)&8(0)&AP Swing 1, Crushing 2, Heavy Weapon, Shock 4 &2 &2 sp\\
Wood Axe				&Blunt			&2H		&M		&7(+2c)&8(-2b)&8(0)&AP Swing 1, Heavy Weapon, Hook, Shock 2 &1 &2 sp\\
\\
\\
	\multicolumn{10}{c}{2H Blunt Weapon Attachments}\\\tabucline{1-10}
Arming Spike &Blunt &2H &- &-(+0p) &- &- &Weapon Attachment (Thrust) &0 &1 cp\\
Reverse Blade &Blunt &2H &- &7(+3c) &- &- &Heavy Weapon, Weapon Attachment (Swing) &0 &6 cp\\
Reverse Hammer &Blunt &2H &- &7(+2b) &- &- &AP Swing 1, Crushing 1, Heavy Weapon, Shock 1, Weapon Attachment (Swing) &0 &6 cp\\
Reverse Spike &Blunt &2H &- &7(+2p) &- &- &AP Swing 2, Heavy Weapon, Hook, Weapon Attachment (Swing) &0 &6 cp\\
\end{longtabu}

\section{Spears}
\begin{longtabu} to 1\linewidth {X[2,l] X X[-1,c] X[-1,c] XX X[-1,c] X[2,l] X[-3,c] X[-3,r]}
	\caption{List of Spears}\\
	\label{tab:Spears}\\\endfirsthead
Weapon Name						&Weapon Type			&Hands	&Reach	&Swing		&Thrust	&Defense (Guard)	&Special																	&Weight	&Cost\\\toprule\endhead
Ahlspiess 						&Polearm, Spear			&2H		&L		&8(-1b)		&7(+1p) &7(0)				&AP Thrust 3,Thin Blade 													&5		&4 sp\\
Fauchard 						&Polearm, Spear			&2H		&VL 	&7(+2c)		&8(+1p) &7(0)				&Forward Swept, Heavy Weapon, Swinging Slot,								&4		&1 sp\\
Ge (Chinese Dagger-Axe Halberd)	&Polearm, Spear			&2H		&VL 	&7(+3p)		&7(+2p) &7(0)				&AP Swing 1, Heavy Weapon, Hook,											&5		&1 sp\\
Half-Pike* 						&Spear					&1H		&L		&8(-2c)		&7(+1p) &8(0)				&Fluid Thrusts																&-		&-\\
Half-Pike 						&Spear 					&2H		&L		&7(-1c)		&6(+2p) &7(0)				&Fluid Thrusts, Hand-Off 													&4		&1 sp\\
Lance (Light) 					&Spear 					&1H		&VL		&10(-1c)	&8(+2p) &10(0)				&Couched Charge, Heavy Weapon 												&4		&1 cp\\
Lance (Heavy) 					&Spear 					&1H		&EL		&10(-1c)	&9(+3p) &10(0)				&Couched Charge, Heavy Weapon 												&6		&1 sp\\
Lance (Hollow) 					&Spear 					&1H		&LL		&10(-1c)	&9(+3p) &10(0)				&Couched Charge, Heavy Weapon 												&8		&4 sp\\
Longstaff 						&Polearm, Spear			&2H		&EL		&7(+0b)		&7(-1b) &7(0)				&Heavy Weapon																&6		&4 cp\\
Military Fork 					&Polearm, Spear			&2H		&VL		&10(-1c)	&7(+1p) &7(0)				&Multi-Hit Thrust 2, Swinging Slot, 										&5		&3 cp\\
Musket (Butt) 					&Spear					&2H		&H		&7(+1b)		&7(-1b) &8(0)				&Heavy Weapon, Shock 2														&0		&-\\
Musket (Bayonet) 				&Spear					&2H		&L		&9(-1c)		&7(+2p) &8(0)				&Heavy Weapon, Thin Blade													&0		&3 cp\\
Partisan 						&Polearm, Spear			&2H		&EL		&8(+2c)		&7(+3p) &7(0)				&Fluid Thrusts,																&5		&10 sp\\
Pike 							&Spear 					&2H		&LL		&10(-2c)	&8(+3p) &9(0)				&Brace																		&6		&1 sp\\
Poleaxe 						&Polearm, Spear, Blunt	&2H		&L		&7(+2c)		&7(+2p) &7(0)				&AP Swing 1, AP Thrust 1, Heavy Weapon, Shock 2, Swinging Slot, Thin Blade	&5		&8 sp\\
Quarterstaff 					&Polearm, Spear			&2H		&VL		&6(+0b)		&6(-1b) &6(0)				&Fluid Thrusts, Heavy Weapon												&4		&3 cp\\
Ranseur 						&Polearm, Spear			&2H		&EL		&8(+0c)		&7(+2p) &7(0)				&																			&5		&1 sp\\
Spear (Arming)* 				&Spear					&1H		&VL		&9(-1c)		&7(+1p) &9(0)				&																			&4		&6 cp\\
Spear (Arming) 					&Spear					&2H		&VL		&7(+0c)		&7(+2p) &7(0)				&Fluid Thrusts, Hand-Off,													&4		&6 cp\\
Spear (Short) 					&Spear					&1H		&M		&8(+0c)		&7(+2p) &8(0)				&Fluid Thrusts 																&0		&3 cp\\
Spear (Standard)* 				&Spear					&1H		&EL		&9(+0c)		&7(+2p) &9(0)				&																			&4		&8 cp\\
Spear (Standard) 				&Spear					&2H		&EL		&8(+0c)		&7(+3p) &8(0)				&Fluid Thrusts, Hand-Off, 													&4		&8 cp\\
Spetum 							&Polearm, Spear			&2H		&VL		&8(+1c)		&7(+2p) &7(0)				&Fluid Thrusts, 															&4		&2 sp\\
Swordstaff 						&Polearm, Spear			&2H		&EL		&8(+3c)		&7(+3p) &8(0)				&																			&8		&3 sp\\
Trident* 						&Polearm, Spear			&1H		&L		&10(-1c)	&8(-1p) &9(0)				&Multi-Hit Thrust 3, 														&4		&5 cp\\
Trident 						&Polearm, Spear			&2H		&L		&9(+0c)		&7(+0p) &7(0)				&Hand-Off, Multi-Hit Thrust 3, 												&4		&5 cp\\
Yari (Jumonji) 					&Polearm, Spear			&2H		&VL		&7(+2c)		&7(+2p) &7(0)				&Heavy Weapon, 																&4		&2 sp\\
Yari (Bishamon) 				&Polearm, Spear			&2H		&VL		&7(+3c)		&7(+2p) &7(0)				&AP Swing 1, 																&4		&3 sp\\
Yari (Kata Kama) 				&Polearm, Spear			&2H		&VL		&7(+2p/+2c)	&7(+2p) &7(0)				&AP Swing 1, Heavy Weapon, Hook, 											&4		&3 sp\\
	\\
	\\
	\multicolumn{10}{c}{Spear Weapon Attachments}\\\tabucline{1-10}
		Reverse Blade &Spear&* &- &7(+2c) &- &- &Heavy Weapon, Weapon Attachment (Swing) &0 &6 cp\\
		Reverse Hammer &Spear&* &- &7(+1b) &- &- &AP Swing 1, Crushing 1, Heavy Weapon, Shock 1, Weapon Attachment (Swing) &0 &6 cp\\
		Reverse Spike &Spear&* &- &7(+1p) &- &- &AP Swing 2, Heavy Weapon, Hook, Weapon Attachment (Swing) &0 &6 cp\\
		Weapon Queue &Spear&* &- &7(-2c) &7(+0p) &8(0) &AP Swing 1,Thin Blade, Weapon Attachment (Thrust) &0 &4 cp\\
	&\multicolumn{8}{c}{*Use attached weapon's Proficiency or Hands}&\\
\end{longtabu}

\section{Polearms}
\begin{longtabu} to 1\linewidth {X[2,l]XX[-1,c]X[-1,c]XXX[-1,c]X[2,l]X[-3,c]X[-3,r]}
	\caption{List of Polearms}\\
	\label{tab:Polearms}\\\endfirsthead
Weapon Name						&Weapon Type			&Hands	&Reach	&Swing		&Thrust		&Defense (Guard)	&Special																	&Weight	&Cost\\\toprule\endhead
Ahlspiess						&Polearm, Spear 		&2H		&L		&8(-1b)		&7(+1p)		&7(0)				&AP Thrust 3,Thin Blade														&5		&4 sp\\
Bardiche						&Polearm				&2H		&VL		&7(+4c)		&9(-1p)		&8(0)				&AP Swing 2, Crushing 2, Heavy Weapon, Shock 2, Swinging Slot				&5		&2 sp\\
Bec de Corbin					&Polearm				&2H		&VL		&7(+2p)		&7(+2p)		&7(0)				&AP Swing 2, Heavy Weapon, Hook, Swinging Slot, Thin Blade					&5		&1 sp\\
Bill (Black)					&Polearm				&2H		&VL		&8(+3c)		&7(+2p)		&7(0)				&AP Swing 1, Crushing 1, Heavy Weapon, Hook, Shock 2, Swinging Slot,		&5		&1 sp\\
Bill (Forest)					&Polearm				&2H		&EL		&8(+2c)		&7(+2p)		&8(0)				&AP Swing 2, Crushing 1, Heavy Weapon, Hook, Shock 2, Swinging Slot,		&5		&1 sp\\
Falx							&Polearm				&2H		&L		&8(+3c)		&10(-1p)	&8(0)				&AP Swing 2, Forward Swept, Heavy Weapon									&4		&1 sp\\
Fauchard 						&Polearm, Spear			&2H		&VL		&7(+2c)		&8(+1p) 	&7(0)				&Forward Swept, Heavy Weapon, Swinging Slot,								&4		&1 sp\\
Ge (Chinese Dagger-Axe Halberd)	&Polearm, Spear			&2H		&VL		&7(+3p)		&7(+2p) 	&7(0)				&AP Swing 1, Heavy Weapon, Hook,											&5		&1 sp\\
Glaive 							&Polearm				&2H		&VL		&7(+2c)		&8(+1p) 	&7(0)				&Draw 2, Heavy Weapon, Swinging Slot,										&5		&3 sp\\
Guandao (Chinese Glaive)		&Polearm				&2H		&VL		&7(+2c)		&8(+2p) 	&7(0)				&AP Swing 1, Draw 2, Heavy Weapon,											&5		&5 sp\\
Halberd							&Polearm				&2H		&VL		&7(+3c)		&8(+2p) 	&7(0)				&AP Swing 2, Crushing 2, Heavy Weapon, Swinging Slot,						&5		&3 sp\\
Lochaber Axe					&Polearm				&2H		&VL		&7(+2c)		&8(+1p) 	&8(0)				&AP Swing 2, Heavy Weapon, Hook, Shock 2									&5		&2 sp\\
Longstaff						&Polearm, Spear			&2H		&EL		&7(+0b)		&7(-1b) 	&7(0)				&Heavy Weapon																&6		&4 cp\\
Lucerne Hammer					&Polearm				&2H		&L		&7(+2b)		&7(+2p) 	&7(0)				&Crushing 2, Shock 2, AP Thrust 1, Heavy Weapon, Swinging Slot, Thin Blade	&5		&8 sp\\
Military Fork					&Polearm, Spear			&2H		&VL		&10(-1c)	&7(+1p) 	&7(0)				&Multi-Hit Thrust 2, Swinging Slot,											&5		&3 cp\\
Morningstar						&Polearm, Blunt			&2H		&L		&7(+3b)		&8(+1p) 	&8(0)				&AP Swing 2, Bleed 3, Heavy Weapon											&5		&1 sp\\
Naginata						&Polearm, Spear			&2H		&VL 	&7(+2c)		&7(+1p) 	&7(0)				&Draw 2, Fluid Thrusts														&4		&10 sp\\
Partisan						&Polearm, Spear			&2H		&EL 	&8(+2c)		&7(+3p) 	&7(0)				&Fluid Thrusts,																&5		&10 sp\\
Pike							&Spear					&2H		&LL		&10(-2c)	&8(+3p) 	&9(0)				&Brace																		&6		&1 sp\\
Poleaxe							&Polearm, Spear, Blunt	&2H		&L		&7(+2c)		&7(+2p) 	&7(0)				&AP Swing 1, AP Thrust 1, Heavy Weapon, Shock 2, Swinging Slot, Thin Blade	&5		&8 sp\\
Quarterstaff					&Polearm, Spear			&2H		&VL		&6(+0b)		&6(-1b) 	&6(0)				&Fluid Thrusts, Heavy Weapon												&4		&3 cp\\
Saber-Halberd					&Polearm				&2H		&VL		&7(+3c)		&8(+0p) 	&7(0)				&AP Swing 1, Crushing 1, Draw 2, Heavy Weapon, Swinging Slot				&5		&15 sp\\
Spetum							&Polearm, Spear			&2H		&VL		&8(+1c)		&7(+2p) 	&7(0)				&Fluid Thrusts																&4		&2 sp\\
Swordstaff						&Polearm, Spear			&2H		&EL		&8(+3c)		&7(+3p) 	&8(0)				&																			&8		&3 sp\\
Trident*						&Polearm, Spear			&1H		&L		&10(-1c)	&8(-1p) 	&9(0)				&Multi-Hit Thrust 3															&		&\\
Trident							&Polearm, Spear			&2H		&L		&9(+0c)		&7(+0p) 	&7(0)				&Hand-Off, Multi-Hit Thrust 3												&4		&5 cp\\
War-Scythe						&Polearm				&2H		&VL		&7(+3c)		&9(+0p) 	&8(0)				&Forward Swept, Heavy Weapon,												&4		&4 cp\\
Yari (Jumonji)					&Polearm, Spear			&2H		&VL		&7(+2c)		&7(+2p) 	&7(0)				&Heavy Weapon,																&4		&2 sp\\
Yari (Bishamon) 				&Polearm, Spear			&2H		&VL		&7(+3c)		&7(+2p) 	&7(0)				&AP Swing 1,																&4		&3 sp\\
Yari (Kata Kama)				&Polearm, Spear			&2H		&VL		&7(+2p/+2c)	&7(+2p) 	&7(0)				&AP Swing 1, Heavy Weapon, Hook,											&4		&3 sp\\
		\\\\
		\multicolumn{10}{c}{Polearm Weapon Attachments}\\\tabucline{1-10}
			Reverse Blade &Polearm &- &- &7(+2c) &- &- &Heavy Weapon, Weapon Attachment (Swing) &0 &6 cp\\
			Reverse Hammer &Polearm &- &- &7(+1b) &- &- &AP Swing 1, Crushing 1, Heavy Weapon, Shock 1, Weapon Attachment (Swing) &0 &6 cp\\
			Reverse Spike &Polearm &- &- &7(+1p) &- &- &AP Swing 2, Heavy Weapon, Hook, Weapon Attachment (Swing) &0 &6 cp\\
			Weapon Queue &Polearm &- &- &7(-2c) &7(+0p) &8(0) &AP Swing 1,Thin Blade, Weapon Attachment (Thrust) &0 &4 cp
\end{longtabu}

\section{Special Qualities}\label{sec:weapons-special-qualities}
\vspace{-5pt}\paragraph{\label{wepqual:AP Swing}AP Swing [X]}---\quad Armor Piercing Swing [X].

When making a Swinging attack against a Hit Location with Hard armor protection, this weapon Inflicts X additional damage against any Armor, up to the AV of the Armor.

\vspace{-5pt}\paragraph{\label{wepqual:AP Thrust}AP Thrust [X]}---\quad Armor Piercing Thrust [X].

When making a Thrusting attack against a Hit Location with Hard armor protection, this weapon Inflicts X additional damage against any Armor, up to the AV of the Armor.

\vspace{-5pt}\paragraph[Bleed]{\label{wepqual:Bleed}Bleed [X]}---\quad
Any Swinging attack from this weapon that either inflicts a Wound, or fails to inflict a Wound by 2 or fewer damage, inflicts X \nameref{sec:Blood Loss}.

\vspace{-5pt}\paragraph{\label{wepqual:Brace}Brace}---\quad
When using this weapon, you gain 2 automatic additional successes on Initiative Tests made as a result of \nameref{par:Aggressive/Aggressive} Orientation.

\vspace{-5pt}\paragraph{\label{wepqual:Cavalry Sword}Cavalry Sword}---\quad
When making a Swing with \nameref{man:Ride-By} attack, this weapon’s \nameref{wepqual:Draw} value is increased by 1.

\vspace{-5pt}\paragraph[Chain]{\label{wepqual:Chain}Chain [X]}---\quad
Any attack made at Swing TN with this weapon ignores the first X successes of an enemy’s Parry or Block defense, including Parrying and Blocking Maneuvers like Riposte or Arm Parry. Additionally, this weapon can neither engage in, nor be engaged in a Hilt Push of any sort.

\vspace{-5pt}\paragraph{\label{wepqual:Companion Weapon}Companion Weapon}---\quad
When this weapon is used to make any sort of defensive Maneuver, one of the dice devoted automatically succeeds before rolling.

\vspace{-5pt}\paragraph{\label{wepqual:Couched Charge}Couched Charge}---\quad
When making a \nameref{man:Lance Charge} attack with this weapon, reduce its Thrusting TN by 2, and use the Horse’s STR as the base for damage instead of your own.. However, your Reach is not increased by the Tall Boon. In addition, if a Lance Charge attack made by this weapon is Blocked, it carries through on the intended Target Zone anyway, as though it hit with 0 BS, and is reduced by the Shield’s AV regardless of where it hit. If the total damage inflicted by an attack that takes advantage of Lance Charge exceeds the Durability of the weapon by 5 or more, then the weapon breaks immediately after the attack is made.

\vspace{-5pt}\paragraph[Chain]{\label{wepqual:Crushing}Crushing [X]}---\quad
When this weapon successfully hits with a Swing attack on a Hit Location with any sort of Hard armor protection, reduce all AVs of the armor on that Hit Location by X. This is done after damage is resolved for this attack, and only affects future attacks. Only the highest AV armor on that location is reduced by Crushing. If that armor is reduced to AV 0 in all areas, then Crushing begins to affect lesser Hard armors beneath it. 

\vspace{-5pt}\paragraph[Draw]{\label{wepqual:Draw}Draw [X]}---\quad
When making a Swinging attack that would inflict cutting damage with this weapon, inflict X additional cutting damage if you roll 3 or more BS. This special rule does not come into play if the target benefits from any sort of Hard armor on the Hit Location being hit.

\vspace{-5pt}\paragraph{\label{wepqual:Fluid Thrusts}Fluid Thrusts}---\quad
This weapon can Feint from a Thrust.

\vspace{-5pt}\paragraph{\label{wepqual:Forward Swept}Forward Swept}---\quad
When this weapon is used in a Swing attack against which the target would receive an AV bonus, use the lower value between the armor’s AVC and AVP. The Swing still inflicts cutting damage regardless of which AV is used.

\vspace{-5pt}\paragraph{\label{wepqual:Freakishly Large}Freakishly Large}---\quad
This weapon is so massive that unless the character using it has either 9 STR or the greater \nameref{boon:Tall} Boon, the TNs for the weapon are all increased by 1.

\vspace{-5pt}\paragraph{\label{wepqual:Hand-Off}Hand-Off}---\quad
This weapon can be wielded with one hand in addition to wielding with both hands. If used in this fashion, you may continue using it with its normal Proficiency, or with a 1H version of either Blunt or Sword proficiency (depending on the base type of the weapon). However, instead of using the normal profile of the weapon, use the 1H version of the weapon with an * next to it. Switching from 1H to 2H versions of a weapon can be done at the beginning of any Action (or in response to an injury) and requires no CP expenditure.

\vspace{-5pt}\paragraph{\label{wepqual:Heavy Weapon}Heavy Weapon}---\quad
When weapons with the Light Blade quality Parry Swinging attacks from this weapon, the Defense TN of the Light Blade is increased by 2, instead of 1.

\vspace{-5pt}\paragraph{\label{wepqual:Hook}Hook}---\quad
This weapon can be used with the \nameref{man:Hook} Maneuver.

\vspace{-5pt}\paragraph{\label{wepqual:Light Blade}Light Blade}---\quad
When Parrying Swinging attacks from any weapon that does not have this quality, this weapon’s Defense TN is increased by 1. Against Heavy Weapons, this penalty increases to 2.

\vspace{-5pt}\paragraph[Multhit]{\label{wepqual:Multi-Hit}Multi-Hit [Y] [X]}---\quad
When this weapon is used in a Y type (Swing or Thrust) attack that targets Hit Locations that successfully hits the target, instead of resolving one attack, X identical attacks of equal strength hit instead. Roll on the Target Zone as normal for the first hit. The second hit lands as though the roll had been one lower, the third hit as though the roll had been one higher, etc. Roll on the the same Target Zone to see where these additional attacks land.

\vspace{-5pt}\paragraph{\label{wepqual:Parrying Teeth}Parrying Teeth}---\quad
-1 Parry TN vs attacks made by Swords and Daggers.

\vspace{-5pt}\paragraph[Shock]{\label{wepqual:Shock}Shock [X]}---\quad
Any hit from this weapon inflicts additional Stun equal to the listed number.

\vspace{-5pt}\paragraph[Spatulate Tip]{\label{wepqual:Spatulate Tip}Spatulate Tip [X]}---\quad
When making a Thrusting attack that would inflict Piercing damage with this weapon, inflict X additional cutting damage if you roll 3 or more BS. This special rule does not come into play if the target benefits from any sort of Hard armor on the Hit Location being hit.

\vspace{-5pt}\paragraph{\label{wepqual:Swinging Slot}Swinging Slot}---\quad
This weapon can have a Swinging Weapon Attachment added to its Swing profile. This does not replace the current Swing profile, but can be used instead of it when declaring a Maneuver.

\vspace{-5pt}\paragraph{\label{wepqual:Thin Blade}Thin Blade}---\quad
When making a Coup de Grace or Joint Thrust attack, this weapon’s Thrust TN is decreased by 1. In addition, against mail armor this weapon gains AP Thrust 3. If the weapon has an AP Thrust quality, this replaces that value, it does not add to it.

\vspace{-5pt}\paragraph{\label{wepqual:Thrusting Slot}Thrusting Slot}---\quad
This weapon can have its Thrust profile replaced with a Thrusting Weapon Attachment.

\vspace{-5pt}\paragraph[Tight Grip]{\label{wepqual:Tight Grip}Tight Grip [X]}---\quad
When using the Power Attack option on a Swing attack, inflict X additional damage. When making a Stability Test to retain your weapon against a Disarm Maneuver or similar event, you may roll 2 additional dice.

\vspace{-5pt}\paragraph{\label{wepqual:Weapon Attachment}Weapon Attachment}---\quad
This weapon can be attached to any weapon with a Swinging Slot (for Reverse Weapons) or a Thrusting Slot (for Arming Spikes) and used instead of the weapon’s basic swinging/thrusting profile. When using the Weapon Attachment, use its attack TN and Damage instead of the original where listed, use the original where no new number is listed, i.e., a Battleaxe with a Reverse Spike Weapon Attachment can make a Strike using the stats of a Reverse Spike instead of the stats of the Battleaxe.

\section{Materials}
\begin{itemize}
	\item [] \textbf{Bronze}
	\subitem -2 Durability
	\subitem +25\% Cost
\end{itemize}
\begin{itemize}
\item [] \textbf{Iron}
\subitem -2 Durability
\subitem -1 Damage of all types
\subitem -25\% Cost
\end{itemize}

\begin{itemize}
\item [] \textbf{Steel}
\subitem No change
\end{itemize}
\begin{itemize}
\item [] \textbf{Orichalcum}
\subitem Cannot break
\subitem +1 Damage of all types
\subitem +5,000\% Cost + 5 gp
\subitem Limited Supply (1)
\end{itemize}
\begin{itemize}
\item [] \textbf{Silversteel}
\subitem -2 Weight
\subitem +2 Durability
\subitem +1 Damage of Cutting and Piercing types
\subitem +1,000\% Cost + 1 gp
\subitem Limited Supply (3)
\end{itemize}
\paragraph{Limited Supply}--- \quad
 This material is in very limited supply. It is so rare that most characters, even the exceedingly wealthy cannot find enough to produce much equipment for themselves. At Character Creation, you may not take more individual items with the Limited Supply quality than the number listed in parenthesis without suffering an increase in cost. If you do take more items of that type, than all of them double in cost. You must also retroactively pay for those you have already taken.

\section{Customization Options}
\paragraph{\label{par:Ridiculously Sharp}Ridiculously Sharp}---\quad \textbf{(+50\% Cost)}\par
This only applies to weapons that inflict cutting damage on hit. It can also be applied to the Reverse Blades of weapons.

Ridiculously Sharp weapons improve their \hyperref[wepqual:Draw]{Draw} value by 1 (or gain Draw 1 if they do not already have Draw) the first time they inflict damage in a fight with a Swinging attack. Afterwards, they must be resharpened to regain this benefit.

Resharpening generally costs a copper piece, but can be done in about an hour with a wheel or stone. Orichalcum weapons never require re-sharpening, and Silversteel weapons can be used twice before they must be resharpened.

\paragraph{\label{par:Bizarre Gimmick}Bizarre Gimmick}---\quad \textbf{(+100\% Cost)}\par
This customisation allows you to add hidden quirks to your weapon. Think hidden parrying hooks, a hollow blade containing an extendable point, a hollow grip filled with blinding dust, or a blade coated in a magnesium solution that can explode into blazing light for an instant.

Bizarre Gimmick worsens the weapon’s best TN by 1 (if its TNs are all the same, worsen the attack TN that inflicts the most damage. If those are the same, reduce the Defense TN), but in return, the Maneuver for which the Gimmick was designed (which must be chosen at creation, with GM approval) gets a special bonus. Once per fight when performing the chosen Maneuver, you may increase your opponent’s Defense TN against it by 2. It can also, optionally, give a weapon the special quality ‘Hook’ if that is the Maneuver chosen.

An additional downside is that professionals and experts tend to look down on this sort of tomfoolery.
\paragraph{Examples for Bizarre Gimmick:}
\begin{itemize}
	\item A spring-loaded sword whose blade juts out another six inches when a button is pressed. Affects Steal Initiative.
	\item Movable crossguard hooks that snap shut to help capture enemy weapons. Affects Disarm.
	\item Flip-out flukes that allow the weapon to hook and catch. Affects Hook. Gives the weapon the special quality Hook if it didn’t already have it.
	\item A hollow grip filled with powder. Affects Blind Toss, and counts as something to ‘throw.’
\end{itemize}
\paragraph{Custom Hilt (+50\% Cost)}\par
A more advanced hilt that provides better protection to the hand. The weapon gains 2 points of Guard value.
\paragraph{Custom Grip (+50\% Cost)}\par
A handgrip that corrects the form and facilitates a certain sort of strike. The weapon either gains Tight Grip 1 (or improves its current Tight Grip value by 1) or gains Draw 1 (or improves its current Draw value by 1). It is possible to buy this upgrade twice to give or upgrade both Tight Grip and Draw.
\paragraph{Exquisite Decoration}---\quad \textbf{(+100-2,000\% Cost)} \par
The weapon is beautifully decorated with gemstones, reliefs, inlays, and other fineries. This has no effect on the weapon’s usefulness in battle, but it may be impressive enough to provide circumstantial bonuses to social rolls involving glamor and grandeur.

\paragraph{Fine Forging (+50\% Cost)}---\quad
The weapon’s Durability increases by 3. You may take this up to three times, increasing the weapon’s Durability by as much as 9 for a total 150\% increase in cost.

\paragraph{Integrated Pistol (+100\% Cost AND 1 gp)}---\quad
The weapon gains an \hyperref[tab:Integrated Firearms]{Integrated Pistol} as listed below. This pistol can be fired as a normal pistol. Alternatively, in the event that you use a Swinging or Thrusting attack that hits an enemy, 1 CP can be spent to make an immediate Melee Shoot attack on the target with BS from the Swing or Thrust. This shot always hits in the same Hit Location as the Swing or Thrust. Because of the bulk of the pistol, the weapon’s best Attack TN is worsened by 1. If the Attack TNs are equal, you may choose which is worsened.

This customization can be taken multiple times to represent a revolver-type weapon, allowing the weapon to be fired multiple times before needing to be reloaded. Additional purchases do not worsen the weapon’s Attack TNs further.

You may purchase a Firing Mechanism for the Integrated Pistol, if you are so inclined, otherwise it is assumed to use a flintlock.

The idea of designing a sword with a pistol built into it may sound like fantasy, but if so, it’s the sort of fantasy that appealed to the medieval mind enough for them to give it a shot. Integrated pistols usually fire in-line with the blade, but a few are built into the crossguard, and fire forward from the hand like regular pistols. Either way, they’re not very accurate and they tend to make the weapon more cumbersome. However, they can be a huge advantage at close range, or when closing with longer-ranged adversaries.
\begin{table}[ht]
	\centering
	\caption{Integrated Firearms}
	\label{tab:Integrated Firearms}
	\begin{tabu} spread 0pt {*{7}{X[-1]}}
Weapon Name &Weapon Type &Range &Missile TN &Damage &Load &Ammunition\\\toprule
Integrated Pistol &Firearm &5 &7 &7p &12 &Ball, Shot\\
Integrated Dragon &Firearm &2 &6 &7p &10 &Ball, Shot\\
	\end{tabu}
\end{table}

\chapter{Missile Weapons}\label{ch:Missile Weapons}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quote}
\emph{“No person of Albish descent will bear or train with, or own the weapon called the Warbow or Longbow without special dispensation from his Lord. Any who violate this order shall tied to a post and speared.” }

--\textit{Text of Krajini Bow Prohibition}
\end{quote}

\section{Missile Weapon Characteristics}
\subsubsection{Weapon Name}
The name of the weapon.

\subsubsection{Weapon Type}
The type of the weapon, and thus the Proficiency that it can be used with. 

\subsubsection{Range}
This is the distance in yards that a missile weapon can be used, and considered to be at “Short” range. 
Every multiple of this distance the weapon is used at past the first decreases the Missile Pool for Shoot or 
Fling by 1. See Missile Weapons in Combat in the \nameref{ch:combat}~for more details.  %NEEDS ref to missile weapons in combat

\subsubsection{Missile TN}
This is the TN used when making a missile attack with this weapon.

\subsubsection{Damage}
This is the power (before modification by arrowheads or Maneuvers/Talents) of an attack made by a 
missile weapon, as well as the damage type of the attack. The damage type of missile weapons is 
invariably piercing (p) damage. 


\subsubsection{Special}
The weapon’s special qualities, which are further described in \autoref{sec:Missile Special Qualities}.

\subsubsection{Weight}
This is the Weight of the weapon when it is being stored. When in use, a weapon’s Weight is 0. 

\subsubsection{Cost}
This is the cost of the weapon.

\subsubsection{Required STR}
\vspace{-10pt}\emph{Bow}

If a character does not have the Required STR of a bow, he cannot string it. If already strung, the bow can 
be used with inferior STR, but the Missile TN increases by 1, and damage dealt decreases by 1, per point 
of the character’s STR beneath the Required STR. 

\subsubsection{Span}
Bows and throwing weapons can be drawn and fired in the same Action. However, crossbows need to be 
spanned, and firearms need to be loaded prior to firing (see Span - Crossbow, and Load - Firearm, 
respectively)

\vspace{-10pt}\emph{Crossbow}

This number determines how many Successes are required before the crossbow is spanned and ready to 
be fired. Reference this number with the Spanning Tool of the weapon, which determines the TN of these 
Successes, as well as whether they can be accumulated over Rounds, or only in one Round. The Reload 
Maneuver is used to Span crossbows. 

\subsubsection{Spanning Tool}
\vspace{-10pt}\emph{Crossbow}

This refers to the Spanning Tool that is used to span the crossbow. The Span TN is the minimum number 
required for Span Successes (see Span - Crossbow). Store Span refers to whether the Span Successes can 
be accumulated over Rounds, or only in one Round (‘yes’ for accumulating Successes, ‘no’ when all 
Successes must be gained in one Round). 

\subsubsection{Load}
\vspace{-10pt}\emph{Firearm}\par
This number determines how many Successes are required before the firearm is loaded and ready to be 
fired. Reference this number with the weapon’s Loading Mechanism, which determines the TN of these 
Successes. Firearms universally accumulate Load Successes over Rounds, whereas crossbows may or 
may not allow that. The Reload Maneuver is used to Load firearms. 

\subsubsection{Loading Mechanism}
\vspace{-10pt}\emph{Firearm}\par
This refers to the Loading Mechanism that is used to load the firearm, and details the type of cartridge 
used, which greatly affects the ease of loading. The Load TN is the minimum number required for Load 
Successes (see Load - Firearm). The Ammunition Cost Modifier determines how much this Loading 
Mechanism affects the cost of ammunition. Brass cartridges would be more expensive than paper cartridges, for example. 

\subsubsection{Firing Mechanism}
\vspace{-10pt}\emph{Firearm}\par
The Firing Mechanism determines the types of Loading Mechanisms available for firearms. Firing 
Mechanisms may have Special qualities that greatly affect how your firearm operates.

\subsubsection{Ammunition}
\vspace{-10pt}\emph{Bow, Crossbow and Firearm}\par
This outlines the type of ammunition that can be fired from the weapon. Ammunition must first be loaded 
before it can be fired. Ammunition may have Special qualities, which may make it more or less effective 
in certain circumstances.

\paragraph{Catch Chance}\label{par:Catch Chance}
---\quad
Some missile weapons, and their projectiles, have a significant chance of becoming stuck in their victims. 
If a missile with a Catch Chance inflicts a Wound on a character, roll 1d10. If the result is equal to or 
under the Catch Chance, then the head of the missile becomes stuck in the Wound. This adds +4 Pain to 
the Wound, and increases its Infection Chance by 2 until it is removed. 

If the stuck missile is not removed (Surgery Check RS 3), and the Wound heals with the missile (or its 
head, in the case of arrows) still stuck inside, then the character immediately gains either the minor or 
major \nameref{bane:Lasting Pain}~Bane in that area. The GM can decide on the level of Lasting Pain, or you can roll for it: 1-6 minor, 7-10 major.

\subsubsection{Ammunition Capacity}
\vspace{-10pt}\emph{Firearm}\par
This specifies how many shots your firearm holds before it requires reloading. Ammunition Capacities 
have Special qualities which affects whether firearms can be fired with simultaneous Shots, and how 
firearms are reloaded. 

\clearpage
\section{Missile Weapons}
\subsection{Bows}
Explain how weapons are used/what is required for it. 
\begin{table}[hb]
	\centering
	\caption{Bows}
	\label{tab:Bows}
	\begin{tabu} spread 0pt {X[3]X[-1,c] *{3}{X[-2,c]} X[-1,c] X[2] *{2}{X[-2]}}
\rowfont[c]{}Weapon Name&Weapon Type&Range&Missile TN&Damage&Required STR&Special&Wt&Cost\\\toprule
Composite Horsebow &Bow &25 &7 &5p &4 &Cavalry Bow 3 &0 &6 sp\\
Composite Warbow &Bow &30 &7 &6p &5 &Cavalry Bow 2 &0.5 &8 sp\\
Horsebow &Bow &20 &7 &4p &3 &Cavalry Bow 3 &0 &2 sp\\
Longbow &Bow &25 &7 &5p &4 &-&0.5 &3 sp\\
Warbow &Bow &30 &7 &6p &6 &- &0.5 &4 sp\\
Zellish Wheelbow &Bow &40 &7 &7p &5 &- &3 &5 gp\\
	\end{tabu}
\end{table}

\begin{table}
	\centering
	\caption{Ammunition for Bows and Crossbows}
	\label{tab:Ammunition for Bows and Crossbows}
	\begin{tabu} spread 0pt {XX[2.5] X[-2,c] X[-2,l]}
\rowfont[c]{}Ammunition Name&Special&\nameref{par:Catch Chance}&Cost\\\toprule
Barbed Broadhead &Winged 2 &10/10 &4 sp/20\\
Bodkin &+10 Range, Narrow &1/10 &1 sp/20\\
Bludgeon/Stun &Bludgeon, Shock 2, -5 Range &0/10 &10 cp/20\\
Broadhead &+1 Damage, Winged 2 &5/10 &2 sp/20\\
Fire Arrow &-1 Damage, +1 Missile TN, -5 Range, Flaming 1 &5/10 &2 sp/20\\
Heavy Broadhead &+2 Damage, Winged 2, -5 Range &5/10 &3 sp/20\\
Lozenge-Head &-5 Range, AP 2 &2/10 &2 sp/20\\
Swallowtail &-1 Required STR, Winged 1 &8/10 &4 sp/20\\
	\end{tabu}
\end{table}

\subsection{Crossbows}
Explain how weapons are used/what is required for it. 
\begin{table} [!hb] %CROSSBOWS LOOKS LIKE SHIT
	\centering
	\caption{Crossbows}
	\label{tab:Crossbows}
	\begin{tabu} to \linewidth { X X[-2,c]X[-2,c]X[-2,c]X[-2,c]  X[4] X[-2]X[-2]}
		Weapon Type & Range & Missile TN           & Damage & Span & Spanning Tool               & Wt & Cost  \\
		Crossbow    & 20    & 6                    & 10p    & 25   & Crank, Windlass             & 3  & 2 gp  \\
		Crossbow    & 5     & 5                    & 4p     & 2    & Lever, Screw                & 0  & 5 sp  \\
		Crossbow    & 15    & 6                    & 8p     & 10   & Hand, Lever, Stirrup, Crank & 2  & 1 gp  \\
		Crossbow    & 10    & 6                    & 7p     & 6    & Hand, Lever, Stirrup        & 1  & 15 sp \\
		Crossbow    & 10    & 6                    & 6p     & 4    & Lever                       & 1  & 10 sp \\
		Crossbow    & 25    & 10(7)\textasciitilde & 15p    & 40   & Winch                       & 5  & 5 gp
	\end{tabu}
\vspace{5pt}	\caption*{\textasciitilde~This weapon’s Missile TN is 10 if it is fired while moving, standing normally, from a horse, and so on. In order to use TN 7, it must be Braced against the ground, a wall, on a stand or rest, and so on.}
\end{table}
\subsubsection{Ammunition}
Bows and Crossbows use the same ammunition. Refer to \autoref{tab:Ammunition for Bows and Crossbows} under Bow - Ammunition. %COULD MAYBE USE REF TO BOW - AMMUNITION

\begin{table}
	\centering
	\caption{Spanning Tools}
	\label{tab:Spanning Tools}
	\begin{tabu} spread 0pt {X *{2}{X[-1,c]} *{2}{X[-2,c]}}
		\rowfont[c]{}Spanning Tool & Span TN & Store Span & Wt & Cost \\ \toprule
		Crank (Crannequin)         & TN 7    & Yes        & 0  & 1 sp \\
		Hand                       & TN 6    & No         & 0  & -    \\
		Lever                      & TN 5    & No         & 0  & 5 cp \\
		Screw                      & TN 8    & Yes        & 0  & 1 sp \\
		Stirrup                    & TN 4    & No         & 1  & 5 cp \\
		Windlass                   & TN 6    & Yes        & 1  & 2 sp \\
		Winch                      & TN 5    & Yes        & 2  & 3 sp
	\end{tabu}
\end{table}

\subsection{Firearms}
Explain how weapons are used/what is required for it. 
\begin{table}[!hb]
	\caption{Firearms}
	\label{tab:Firearms}
	\begin{tabu} to \linewidth {X[3] X *{4}{X[-2,c]}  X[3] *{2}{X[-2]}}
		\rowfont[c]{}Weapon Name & Weapon Type & Range & Missile TN           & Damage & Load & Ammunition                & Wt  & Cost   \\ \toprule
		Arquebus                 & Firearm     & 10    & 7                    & 6p     & 20   & Ball, Shot                & 1   & 12 sp  \\
		Blunderbuss              & Firearm     & 3     & 6                    & 6p     & 18   & Ball, Heavy Shot          & 1   & 12 sp  \\
		Deck Gun/Abus Gun        & Firearm     & 25    & 10(7)\textasciitilde & 18p    & 50   & Ball, Heavy Shot, Spike   & 10  & 100 sp \\
		Dragon                   & Firearm     & 2     & 6                    & 5p     & 10   & Ball, Shot                & 0.5 & 18 sp  \\
		Hand Bombard             & Firearm     & 8     & 8                    & 12p    & 30   & Ball, Shot, Spike         & 2   & 10 sp  \\
		Hand Gonne               & Firearm     & 5     & 8                    & 6p     & 20   & Ball                      & 1   & 8 sp   \\
		Musket                   & Firearm     & 15    & 7                    & 8p     & 20   & Ball, Shot, Buck and Ball & 2   & 20 sp  \\
		Pistol                   & Firearm     & 5     & 7                    & 5p     & 12   & Ball, Shot                & 0   & 10 sp  \\
		Rifle                    & Firearm     & 20    & 7                    & 10p    & 25   & Ball, Rifle Ball          & 1   & 40 sp
	\end{tabu}
\vspace{5pt}\caption*{\textasciitilde~This weapon’s Missile TN is 10 if it is fired while moving, standing normally, from a horse, and so on. In order to use TN 7, it must be Braced against the ground, a wall, on a stand or rest, and so on.}
\end{table}

\begin{table}[!hb]
	\centering
	\caption{Ammunition}
	\label{tab:Ammunition for Firearms}
	\begin{tabu} spread 0pt {XX[2.5] X[-2,c] X[-1,l]}
\rowfont[c]{}Ammunition Name &Special &Catch Chance &Cost\\\toprule
Ball &AP 6 &9/10 &1 cp/10\\
Buck and Ball &AP 4 (first hit only), Scatter 3/6 &9/10 &1 cp/10\\
Heavy Shot &Scatter 8/6 &9/10 &3 cp/10\\
Rifle &Ball AP 6, -1 Missile TN, -10 Load &8/10 &6 cp/10\\
Shot &Scatter 6/6, -3 Load &9/10 &1 cp/10\\
Spike &+2 Damage, -1 Missile TN, +3 Range, +10 Load &1/10 &1 cp/Each\\
	\end{tabu}
\end{table}

\subsubsection{Ammunition Capacity}
\paragraph{Double}---\quad 2 Shots - Wt +50\% Cost

Can be fired twice before reloading, or both barrels can be fired simultaneously. 
Declare a single Shot normally, resolve the second Shot with the same amount of 
dice as the first. Each attack hits and is resolved separately. Each chamber must be 
reloaded separately. 
This weapon can now use Rapid Shot.


\paragraph{Single}---\quad 1 Shot - Wt - Cost

Standard for most weapons.

\paragraph{Magazine (X)}---\quad X+1 Shots 1 Wt +2,000\% Cost base, +50\% per X.

This weapon now has an internal magazine that chambers X rounds (plus the one loaded into the chamber already), and loads them into the weapon one at a time as they are fired.
After firing a round, the next round can be fired without Reloading until all the rounds in the magazine have been fired. Each purchase of this Capacity increases X by 1, to a maximum of 15.

When performing the Reload Maneuver, it is possible to reload more than 1 round into the magazine. If the successes rolled on the Reload are higher than the Load value of the weapon, then for each time the successes multiply the Load value, an additional round is loaded.

(Example: A Dragon Pistol with a Magazine has a Load value of 10. If a highly capable shooter were to roll 20 successes on a Reload Action, then he could reload 2 rounds into the gun at once!)

This weapon can now Rapid Fire.

This weapon can only load Brass or Paper Mache cartridges.

\paragraph{Multishot (X)}---\quad X+1 Shots 0.5 Wt +50\% Cost

This weapon now has X additional rounds that fire simultaneously. Declare a single Shot normally, then resolve the other X with the same amount of dice as the first.
Each attack hits and is resolved separately. However, any attack made with this weapon suffers an MP reduction equal to X due to increased recoil. Each chamber must be reloaded separately. Each purchase of this Capacity increases X by 1, to a maximum of 10.

This weapon cannot use Rapid Shot, as the barrels do not fire sequentially. 

\paragraph{Revolver (X)}---\quad X+1 Shots 0.5 Wt +100\% Cost per X

Can be fired X additional times before reloading. Each chamber must be reloaded separately. Each purchase of this Capacity increases X by 1, to a maximum of 10. 
After the first purchase, additional purchases do not further increase the weapon’s Weight. 

This weapon can now use Rapid Shot. 

\paragraph{High Caliber (X)}---\quad 0 Wt +25\% Cost

This modification changes the size and power of the weapon’s ammunition instead of the weapon’s capacity. This can be applied to any weapon, and if combined with other Ammunition Capacity options, affects all loads in the weapon.

Weapon Damage is increased by X. Missile TN is increased by X.


\subsubsection{Firing Mechanism}
\begin{table}[hb]
	\centering
	\caption{Firing Mechanism}
	\label{tab:Firing Mechanism}
	\begin{tabu} to \linewidth {X X[5] X[2]X[-1]}
\rowfont[c]{}Firing Mechanism &Special &Loading Mechanism &Cost\\\toprule
Caplock$^{\#}$ 
	&-5 Load, -1 Missile TN, \newline
		Requires a Percussion Cap for each Shot (Percussion Caps cost 1 cp each).
	&Manual,\newline
		Paper Cartridge.
	&1 gp\\
Firelock 
	&Must be touched off by hand to fire (+10 Load), \newline
		Firearm Explodes on Critical Failure (as Light Hand Grenade)
	&Manual.
	&-\\
Flintlock
	&Flint must be changed every 20 Shots, Flint Breaks on Critical Failure, Changing flint requires Load 5 Reload Action
	&Manual, \newline
		Paper Cartridge.
	&3 sp\\
Matchlock &Fuse must be lit before being fired (+20 Load for the first Shot only), it won’t work if wet, and burns 1 foot of match per hour if kept lit. Match is extinguished on Critical Failure (Matchcord costs 1 cp per foot) 
	&Manual, \newline
	Paper Cartridge.
&6 cp\\
Needlefire$^{\#}$ &-1 Missile TN, -8 Load 
	&Papier-Mâché Cartridge,\newline
	Brass Cartridge.
&5 gp\\
Snaplock &Flint must be changed every 10 Shots, Flint Breaks on Critical Failure, Changing flint requires Load 5 Reload Action 
	&Manual, \newline
	Paper Cartridge.
&2 sp\\
Wheellock &-1 Missile TN, Wheel Breaks on Critical Failure 
	&Manual, \newline
	Paper Cartridge.
	&8 sp\\
	\end{tabu}
\vspace{5pt}\caption*{\#~These weapons are several centuries more advanced than the rest, and would not be available in any sort of medieval campaign.}
\end{table}

\begin{table}
	\centering
	\caption{Firearm Loading Mechanism}
	\label{tab:Loading Mechanism}
	\begin{tabu} spread 0pt {XXX}
\rowfont[c]{}Loading Mechanism &Load TN &Ammunition Cost Modifier\\\toprule
Brass Cartridge &TN 3 &1,000\% \\
Manual &TN 8 &100\% \\
Paper Cartridge &TN 6 &200\% \\
Papier-Mâché Cartridge &TN 4 &300\% \\
	\end{tabu}
\end{table}

\subsection{Throwing Weapons}
Explain how weapons are used/what is required for it.

\begin{table}[hb]
	\centering
	\caption{Throwing Weapons}
	\label{tab:Throwing Weapons}
	\begin{tabu} spread 0pt {*{9}{X[-1]}}
		\rowfont[c]{}Weapon Name & Weapon Type & Range & Missile TN & Damage       & Special            & \nameref{par:Catch Chance}&Wt   & Cost \\ \toprule
		Axe                      & Thrown   & 10  & 8  & +3c/+1b      &-                   & 5/10    & -  & -\\
		Chakram                  & Thrown   & 15  & 7  & +1c          &Winged 2 & 0/10               & -  & 1 sp\\
		Club                     & Thrown   & 10  & 7  & +1b          & -                  & 0/10    & -  & -    \\
		Hammer                   & Thrown   & 10  & 7  & +2b          & AP 1               & 0/10    & -  & -    \\
		Heavy Dart               & Thrown   & 20  & 8  & +2p          & AP 1               & 6/10    & 0  & 1 sp \\
		Javelin (Heavy)          & Thrown   & 10  & 8  & +3p          & AP 1               & 5/10    & 1  & 2 sp \\
		Javelin (Lead)           & Thrown   & 5   & 8  & +4p          & AP 2, Shield Stick & 10/10   & 1  & 3 sp \\
		Javelin (Light)          & Thrown   & 15  & 7  & +2p          & -                  & 5/10    & 0  & 1 sp \\
		Knife                    & Thrown   & 5   & 8  & +1p/+0c      & -                  & 4/10    & 0  & -    \\
		Knife (Throwing)         & Thrown   & 10  & 7  & +1p/+0c      & -                  & 4/10    & 0  & 8 cp \\
		Metal Weight             & Thrown   & 10  & 7  & +2b          & -                  & 0/10    & 1  & 1 cp \\
		Pole-Sling               & Thrown   & 25  & 7  & +3b          & -                  & 0/10    & 2  & 5 cp \\
		Rock                     & Thrown   & 15  & 7  & +1b          & -                  & 0/10    & 0  & -    \\
		Sling (Light)            & Thrown   & 15  & 7  & +1b          & -                  & 0/10    & 0  & 1 cp \\
		Sling (Long)             & Thrown   & 20  & 7  & +2b          & -                  & 0/10    & 0  & 2 cp \\
		Soliferrum               & Thrown   & 5   & 8  & +3p          & AP 3, Shield Stick & 8/10    & 1  & 4 sp \\
		Spear                    & Thrown   & 10  & 7  & +2p          & -                  & 5/10    & -  & -    \\
		Sword                    & Thrown   & 5   & 8  & +3p          & -                  & 8/10    & -  & -
	\end{tabu}
\end{table}

\section{Special Qualities}\label{sec:Missile Special Qualities} %MISSILE SPECIAL QUALITIES

\vspace{-5pt}\paragraph[AP]{\label{rangequal:AP [X]}AP [X]}---\quad Armor Piercing [X]. Inflicts X additional damage against any Armor, up to the AV of the Armor. Has no effect on Armor with the Bulletproof quality. 

\vspace{-5pt}\paragraph[Bleed]{\label{rangequal:Bleed [X]}Bleed [X]}---\quad Any Wounds inflicted by this missile cause X additional Bleeding damage. 

\vspace{-5pt}\paragraph{\label{rangequal:Bludgeon}Bludgeon}---\quad Inflicts Bludgeoning damage regardless of the weapon’s damage type. 

\vspace{-5pt}\paragraph[Cavalry Bow]{\label{rangequal:Cavalry Bow [X]}Cavalry Bow [X]}---\quad
This bow is easier to use while riding horseback. The penalty for shooting from horseback while 
moving (4 CP normally) is reduced by X for this weapon. 

\vspace{-5pt}\paragraph[Flaming]{\label{rangequal:Flaming [X]}Flaming [X]}---\quad
On hit inflicts X/TN 5 Burn for 3 Rounds to Hit Location. 

\vspace{-5pt}\paragraph{\label{rangequal:Narrow}Narrow}---\quad
This weapon gains AP 4 against armor with the Mail special quality. 

\vspace{-5pt}\paragraph[Scatter]{\label{rangequal:Scatter [X/Y]}Scatter [X/Y]}---\quad 
In addition to the primary Shot, which resolves normally, you automatically hit with a number of 
additional attacks equal to X (roll for Hit Location on the Random Missile Hit Table with each for a 
Shoot Maneuver, roll for Hit Location on appropriate Thrust Table for Melee Shot Maneuver) at 
Damage Y. Attacks that hit the same Hit Location combine their Damage before subtracting AV and 
TOU. 

Hard Armor without the Mail quality halves Scatter damage. 

No special rules apply to these additional attacks unless they are listed after X/Y. Scattered attacks do 
not gain BS as damage. 

Weapons with Scatter do not suffer the normal penalties for Range. Instead, each Range increment 
past the first reduces X by 1. Once X is 0, the next Range increment renders the weapon totally 
ineffective. It may be possible (GM’s discretion) to hit multiple opponents with Scatter (see 
Scatterfire in the Combat chapter) 

Voids made against attacks with Scatter (if they can be made at all, i.e. with Stains of Time) have an 
additional Activation Cost equal to X. 

Shield Stick 

If a Fling attempt made by this missile is blocked, or if it hits an area protected by a shield’s passive 
AV, it becomes ‘stuck’ in the shield. The shield’s Block TN is increased by 2, and its Weight is 
increased by 2 as well. Removing the missile is very difficult, requiring several minutes of work that 
cannot be done in combat. 

Shock [X] 
Successful hit inflicts Stun equal to X. 

Winged 
[X] 

When this weapon hits, it inflicts 1 bonus damage per X BS scored. This damage is not inflicted if the 
missile hits Hard armor. The Textile quality of armor is negated by a Winged missile. 

\chapter{Prosthetics}\label{ch:Prosthetics} %WHOLE PROSTHETICS CHAPTER
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quotation}
\emph{“You are sure this will work, elf?”}

\vspace{5pt}\emph{“Yes, Herr Muts, it will work. I will admit that accommodating the socket for your musculature was difficult, we have never had such a well developed specimen on the table before, and we had to accommodate for the great strain you are likely to be placing on the limb. Many of the muscles have been re-anchored to the socket, though they won’t really do anything except hold it in place. Additional support rods connect to the scapula and clavicle. You should notice an increased appetite as long as it is active\ldots~ That is normal. When you want to disengage the limb, perhaps to sleep, simply envision a blue triangle in your mind’s eye. To activate it, attempt to touch your thumb to little-finger on the mechanical hand, and that should re-engage it.”}

\vspace{5pt}\emph{“Are you sure this isn’t witchcraft?”}

\vspace{5pt}\emph{“Of course not. You’re still human right down to the marrow of your\ldots~remaining bones.”}

\vspace{5pt}\emph{“What about the special feature I requested?”}

\vspace{5pt}\emph{“Oh yes, let me go set up some targets\ldots”}
\end{quotation}

Prosthetics are artificial replacement limbs designed to restore some functionality to amputees. At their 
most basic, a prosthetic can be a simple stick, like the idiosyncratic ‘peg leg’ that everyone imagines 
pirates as sporting. At their most complex, they can be jointed mechanical replicas of the lost limb that 
can actually provide surprising functionality, even in combat. 

If you are missing more than one part of a limb, you must take prosthetics for the pieces closer to the 
remaining arm before you take prosthetics for the extremities. That is, if you are missing your arm below 
the elbow, and you want to take a Clamp, you must first take an Extender.

Some notes with the following tables: 
\begin{itemize}
\item Armor may not be worn over prosthetics unless otherwise specified. 
\item Damage does not stack between attacks. 
\item Location refers to the area the prosthetic is replacing. 
\item Prosthetics count as “Hard” armor for all purposes, including Draw, Armor Piercing effects, etc.
\end{itemize}

\subsubsection{Example Prosthetic}
	\vspace{-5pt}\paragraph{Prosthetic Name}---\quad Location, cost
	\vspace{-15pt}\subparagraph{Effect:} Rules effect of the prosthetic in question.
\section{Arm Prosthetics}
\paragraph{Stick}---\quad Any Location, 8 cp
	\vspace{-15pt}%\begin{description}
	\subparagraph{Effect:} The limb can punch or arm-parry as normal, but will break upon receiving 4 or more damage. May be armored.
\paragraph{Extender}---\quad Forearm, 2 cp
	\vspace{-15pt} \subparagraph{Effect:} A hand prosthetic may be attached to this. 5 damage breaks the prosthetic.
\paragraph{Hook}---\quad Hand, 5 sp
	\vspace{-15pt} \subparagraph{Effect:} Grappling suffers no penalties, but a Hook cannot hold a weapon, including a 1H or 2H weapon. Punch becomes a swing and does piercing damage. 6 damage breaks the prosthetic.
\paragraph{Clamp}---\quad Hand, 10 sp 
	\vspace{-15pt} \subparagraph{Effect:} Grappling suffers no penalties. A Clamp can hold either a 1H weapon, or be used in combination with a real hand to hold a 2H weapon. CP is reduced by 2 when using a Clamp in combat. Because of the locked design of the Clamp, Disarming or Beating a weapon held by a Clamp has a +2 Activation Cost. A successful disarm breaks the clamp, rendering it useless. 6 damage breaks the prosthetic. 
\paragraph{Weapon}---\quad Hand/Forearm, Weapon Cost $+$ 10 sp
	\vspace{-15pt} \subparagraph{Effect:} A 1H melee weapon, or a hand crossbow, pistol, or dragon, may be attached to the stump, and functions as the weapon would. If attached at the Forearm, decrease Range and Damage by 1 unless ranged. Ranged weapons must be fired using the other hand, or by the lost limb using a special mechanism costing 5 sp.
\paragraph{Articulated Hand}---\quad Hand, 25 gp
	\vspace{-15pt} \subparagraph{Effect:} The hand functions as normal except for a +1 Activation Cost to Weapon Maneuvers when using it to hold a 1H weapon by itself. Wielding a 2H weapon with an Articulated Hand and a real hand incurs no penalty. Wielding a 2H weapon with two Articulated Hands incurs a +1 Activation Cost for all maneuvers. 3 damage breaks the prosthetic
\paragraph{Armored Articulated Hand}---\quad Hand, 75 gp
	\vspace{-15pt} \subparagraph{Effect:} The hand functions as normal except for a +1 Activation Cost on all maneuvers involving it. 8 damage breaks the prosthetic.
\paragraph{Burdinadin Prosthetic Hand}---\quad Hand, 30 gp
	\vspace{-15pt} \subparagraph{Effect:} Functions like an Armored Hand until activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.
\paragraph{Burdinadin Prosthetic Forearm}---\quad Hand/Forearm, 50 gp
	\vspace{-15pt} \subparagraph{Effect:} Functions like an Armored Hand until activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.
\paragraph{Burdinadin Prosthetic Full Arm}---\quad Full Arm, 80 gp
	\vspace{-15pt} \subparagraph{Effect:} Does not function unless activated. While Active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.


%\end{description}
%%Decided to not make a table because it'd look like shit, but keeping the code in case I change my mind
	%\begin{longtabu} to \linewidth{*{3}{X[-1]}X[4]}
	%	\captionsetup{textformat=empty, labelformat=blank}
	%	\caption{Arm Prosthetics}\vspace{-25pt}
	%	\label{tab:Arm Prosthetics}\\
	%	Prosthetic &Location &Cost &Effect\\\toprule
	%	Stick &Any &8 cp &The limb can punch or arm-parry as normal, but will break upon receiving 4 or more damage. May be armored.\\
	%	Extender &Forearm &2 cp &A hand prosthetic may be attached to this. 5 damage breaks the prosthetic. \\
	%	Hook &Hand &5 sp &Grappling suffers no penalties, but a Hook cannot hold a weapon, including a 1H or 2H weapon. Punch becomes a swing and does piercing damage. 6 damage breaks the prosthetic. \\
	%	Clamp &Hand &10 sp &Grappling suffers no penalties. A Clamp can hold either a 1H weapon, or be used in combination with a real hand to hold a 2H weapon. CP is reduced by 2 when using a Clamp in combat. Because of the locked design of the Clamp, Disarming or Beating a weapon held by a Clamp has a +2 Activation Cost. A successful disarm breaks the clamp, rendering it useless. 6 damage breaks the prosthetic. \\
	%	Weapon &Hand/Forearm &Weapon + 10 sp &A 1H melee weapon, or a hand crossbow, pistol, or dragon, may be attached to the stump, and functions as the weapon would. If attached at the Forearm, decrease Range and Damage by 1 unless ranged. Ranged weapons must be fired using the other hand, or by the lost limb using a special mechanism costing 5 sp. \\
	%	Articulated Hand &Hand &25 gp &The hand functions as normal except for a +1 Activation Cost to Weapon Maneuvers when using it to hold a 1H weapon by itself. Wielding a 2H weapon with an Articulated Hand and a real hand incurs no penalty. Wielding a 2H weapon with two Articulated Hands incurs a +1 Activation Cost for all maneuvers. 3 damage breaks the prosthetic\\
	%	Armored Articulated Hand &Hand &75 gp &The hand functions as normal except for a +1 Activation Cost on all maneuvers involving it. 8 damage breaks the prosthetic.\\
	%	
	%	Burdinadin Prosthetic Hand &Hand &30 gp &Functions like an Armored Hand until activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.\\
	%	Burdinadin Prosthetic &Hand/Forearm &50 gp &Functions like an Armored Hand until activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties Forearm whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.\\
	%	Burdinadin Prosthetic Full Arm &Full Arm &80 gp &Does not function unless activated. While Active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and the hand can do anything (such as write, manipulate fine objects, and press buttons) that a normal hand can do. Burdinadin Prosthetics can sustain 15 damage before being destroyed.\\
	%\end{longtabu}
\section{Leg Prosthetics}
\paragraph{Wood Foot}---\quad Foot, 10 cp
\vspace{-15pt} \subparagraph{Effect:} Allows you to move with -2 MOB. 5 damage breaks the prosthetic. May be armored.
\paragraph{Pegleg}\label{par:Pegleg}
---\quad Shin, 1 sp
\vspace{-15pt} \subparagraph{Effect:} Allows you to move with $ -\frac{1}{4} $ MOB. 4 damage breaks the prosthetic. May be armored.
\paragraph{Long Pegleg}\label{par:Long Pegleg}
---\quad Thigh/Shin, 1 sp
\vspace{-15pt} \subparagraph{Effect:} Allows you to move with $ -\frac{1}{2} $ MOB. 4 damage breaks the prosthetic. May be armored.
\paragraph{Cane}---\quad Any, 10 cp
\vspace{-15pt} \subparagraph{Effect:} Allows you to move with $ -\frac{1}{2} $ MOB. Requires a free hand to use. May include a weapon of medium length or shorter (add the cost of the weapon). 5 damage breaks the cane.

\paragraph{Crutch}---\quad Any, 10 cp
\vspace{-15pt} \subparagraph{Effect:} Allows you to move with $ \frac{1}{2} $ MOB, or normal MOB if two are used. Each requires a free arm to use. May be used as a Large Club with +1 to Range and all TNs. 5 damage breaks the crutch.

\paragraph{Burdinadin Prosthetic Foot}---\quad Foot, 10 gp
\vspace{-15pt} \subparagraph{Effect:} Functions like a normal foot.

\paragraph{Burdinadin Prosthetic Lower Leg/Foot}---\quad Foot/Lower Leg, 25 gp
\vspace{-15pt} \subparagraph{Effect:} Functions like a \nameref{par:Pegleg} unless activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and can do anything that a real leg can do (such as wiggle toes, tap dance, and C-walk). Burdinadin Prosthetics can sustain 15 damage before being destroyed. 

\paragraph{Burdinadin Prosthetic Full Leg}---\quad Foot/Lower Leg/Upper Leg, 50 gp
\vspace{-15pt} \subparagraph{Effect:} Functions like a \nameref{par:Long Pegleg} unless activated. While active, the limb reduces your END by 1 temporarily, but suffers NO penalties whatsoever, and can do anything that a real leg can do (such as wiggle toes, tap dance, and C-walk) Burdinadin Prosthetics can sustain 15 damage before being destroyed.

\section{Burdinadin Prosthetics}
Do these sound like weird cybernetics? 

That’s because they are. The Burdinadin in the Tattered Realms have access to technologies that even we today have not quite matched because of their Natural Philosophy powers. Burdinadin limbs range from 
appearing smooth and realistic, to rugged and boxish. They are typically attached to sockets grafted to the 
flesh of the wearer, so that they can be removed and replaced in the event of catastrophic damage, or for maintenance. 

These limbs are powered by human body heat, and the sockets contain devices that essentially siphon caloric energy from the body to burn and use to power the limb. Much as a regular limb does, but far, far less efficiently. Wearers of these prosthetics are advised not to keep them active for long periods of time, 
as the siphoning of energy causes fatigue and hunger, and can eventually lead to malnutrition, and even death. 

Any Burdinadin limb can have a weapon installed in it, to jut out from some angle and be used in 
violence, at an additional cost of 5 gp plus the cost of the weapon. Obviously the weapon would have to 
fit inside the limb. Inside a half limb you may fit the Pistol, Dragon, Hand Crossbow, any Melee weapon 
of Short or lower Reach, or a Buckler. You may additionally mount a Blunderbuss, Light Crossbow, or 
Melee Weapon of Medium Reach or shorter inside a full limb. This weapon must be Drawn like any other 
weapon before it can be used, but counts as “readily available” for the purposes of Quickdraw. 

\subsubsection{Optional Fatigue Rules}
If you are using Fatigue rules to track character Exhaustion in combat, then instead of having activated 
Burdinadin Prosthetics reduce END, simply have them add 1 Fatigue per round per replaced limb. So a 
character with a prosthetic hand and a full prosthetic leg would add 2 Fatigue per round, whereas one with 
two prosthetic legs and a prosthetic forearm would add 3. 

There are also Enhancements that can be attached to limbs to increase their performance, also detailed in 
the Burdinadin Armory. 
\chapter{Shields}\label{ch:Shields}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\newpage
Shields are typically held in a fighter’s off-hand, and are used to block, bind and bash in conjunction with 
a proper weapon held in the dominant hand. 

Shields have a long history, and come in a huge variety of shapes and compositions.

\section{Shield Characteristics}
\paragraph{Shield Name}---\quad
The variety of shield. Some of these are generalizations, as dozens of different cultures may have had 
very minor permutations on the same shield design, but each called it something different. 


\vspace{-10pt}\paragraph{Shield Bash TN/Damage}---\quad
This number determines how effectively a shield can be used to bash or otherwise actively disrupt an opponent, and the amount of additional Bludgeoning damage it does on strike. Spiked shields, when used this way, inflict Piercing instead of Bludgeoning damage. The Bash TN is used in \nameref{man:Shield Bash}, and similar Maneuvers.

\vspace{-10pt}\paragraph{Block TN}---\quad
This number determines how effectively a shield can be used to block, deflect, or otherwise thwart 
opponents. The Block TN is used in Block, and similar Maneuvers. 

\vspace{-10pt}\paragraph{AV}---\quad
This determines the amount of passive AV applied to any body part covered by the shield when it is attacked. Certain effects can ignore shield AV or remove it temporarily from the body part. The AV 
stated is the same for all damage types.

Shield AV counts as \nameref{par:Hard}.

\vspace{-10pt}\paragraph{Durability}---\quad
Durability refers to the shield’s resistance to physical damage. If a shield suffers damage (from a \nameref{man:Hew}~Maneuver, or similar source) equal to its Durability, it is destroyed. Each time a shield sustains a \nameref{man:Hew}~attempt that does not destroy it, its Durability is permanently reduced by 2. Each time an attack of any 
sort that lands on a Shield’s protected areas inflicts damage higher than the shield's Durability, Durability 
is permanently reduced by 1.

\vspace{-10pt}\paragraph{Coverage} ---\quad
Coverage represents the Hit Locations that are covered by the shield. Much like armor, when a shield is 
carried on the arm, it provides AV equal to the amount listed in its AV column to the Hit Locations listed. 
All shield AV stacks with regular armor AV. 

\vspace{-10pt}\paragraph{Weight}---\quad
This is the weight added to encumbrance as long as the shield is carried on the arm. When a shield is held 
on the back or belt, it does not add to effective weight. 

\vspace{-10pt}\paragraph{Cost}---\quad
The cost of the shield.
\newpage
\section{Shield List}
\begin{longtabu} to \linewidth {X[2,l] X[c] *{3}{X[-1,c]} X[3] *{2}{X[-1,c]}}
	\caption{Shields}
	\label{tab:Shields}\\
Shield Name								&Bash TN&Block TN&AV&Dur.&Coverage& Wt.& Cost\\\toprule
Buckler									&6(+0b)	&7	&8	&15	&Hand*, Forearm*&0&5 sp\\
Oval Shield 							&7(+1b)	&7	&6	&10	&Hand*, Forearm* Elbow, Upper Arm, Shoulders, Neck, Chest, Sides, Belly, Hips, Groin, Thighs&1&2 sp\\
Wooden Targe 							&7(+0b)	&7	&6	&10	&Hand, Forearm, Chest&0&10 cp\\
Wicker Targe 							&7(-2b)	&7	&4	&6	&Hand, Forearm, Chest&0&2 cp\\
Rodela 									&8(+0b)	&7	&8	&15	&Hand, Forearm, Elbow, Upper Arm, Chest, Side &1& 10 sp\\
Bulletproof Rodela/Round Shield&8(+1b)	&7	&10	&20	&Hand, Forearm, Elbow, Upper Arm, Chest, Side &2&30 sp\\
Heater Shield 							&8(+0b)	&7	&6	&10	&Hand, Forearm, Elbow, Upper Arm, Shoulder, Chest, Side, Belly, Hip, Groin, Thigh&1&1 sp\\
Kite Shield 							&9(+0b)	&7	&6	&10	&Hand, Forearm, Elbow, Upper Arm, Shoulder, Chest, Side, Belly, Hip, Groin, Thigh, Knee&2&2 sp\\
Large Round Shield 						&8(+0b)	&7	&6	&10	&Hands, Forearm, Elbow, Upper Arm, Shoulders, Neck, Chest, Sides, Belly, Hips, Groin, Thighs&1&1 sp\\
Scutum 									&7(+1b)	&7	&6	&12	&Hand*, Forearm* Elbow, Upper Arm, Shoulders, Neck, Chest, Sides, Belly, Hips, Groin, Thighs, Knees, Shins&2&5 sp\\
Round Shield 							&6(+0b)	&7	&6	&8	&Hand*, Forearm* Elbow, Upper Arm, Shoulders, Neck, Chest, Sides, Belly, Hips, Groin, Thighs&1&2 sp\\
Hand Pavise 							&9(+0b)	&7	&6	&12	&Hands, Forearm, Elbow, Upper Arm, Shoulders, Neck, Chest, Sides, Belly, Hips, Groin, Thighs, Knees, Shins&1&5 sp\\
	\caption*{*As long as this shield's AV is in effect, it applies its AV to both hands and forearms}
\end{longtabu}

\chapter{Armor}\label{ch:Armor}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\section{What Armor Should I Wear?}
\begin{quote}
	\centering
	\emph{“If you have to choose one piece, get a helmet.”}
\end{quote}
Before you proceed, there’s something you should know. There are dozens and dozens of armor pieces and types in 
the section to follow. It’s very easy to be overwhelmed by the sheer number of choices you’re about to encounter. 

Don’t hyperventilate. Remain calm. If sifting through \textasciitilde150 armor pieces to custom-build your own outfit doesn’t
appeal to you, we’ve got your back. Before this chapter details the many armor components available to you, there are
Pre-made Armor Sets that you can choose from. These are full suits of armor (not including helmets) complete with 
references for the total weight of the armor, how much it costs, what it protects and how well it protects it, as well as a description of the armor set.

You’ll still have to pick a helmet, but that’s just one list, in \autoref{sec:Head and Neck Armor}

[SIDEBAR] ARMORING TIPS %SIDEBAR?

-Plate armor provides the highest protection, but many pieces have Weak Points that can be targeted by certain 
maneuvers. It’s always smart to wear something beneath Plate.

-Mail armor provides good coverage, but is weak against weapons with the Thin Blade quality.

-Brigandine has no weaknesses, but is less protective against cutting than Mail or Plate.

-Padded armor is very useful against missiles, particularly when combined with a heavier armor. Most armors should be worn with padding.

-Some weapons can severely reduce the effectiveness of armor, but none can completely ignore it through sheer 
damage. It’s always good to be armored.

-Any protection is better than nothing. If you have an exposed point, the enemy WILL attack it. 

-Partial armor is not terribly useful in melee combat, but since missile attacks land randomly, it is useful if you expect to be shot at.

-Don’t do the Hollywood thing and not wear a helmet. The head is one of the easiest places to attack.

-If you expect to be fighting on horseback, remember that leg armor’s weight is negated while riding!

[END SIDEBAR]

\subsection{Armor Value}
The Armor Value, or ‘AV’ of a piece of armor, has three entries: against cutting damage (AVC), against piercing 
damage (AVP) and against bludgeoning damage (AVB).

When a character is attacked with a weapon and suffers damage, the armor adds its AV to his TOU for the purpose of 
resisting that damage. The AV used is obviously of the category of damage that is being suffered (either cutting, 
piercing or bludgeoning). So if an armor component has an AVC of 8, AVP 7 and AVB 6, and a character suffers a 
piercing attack to the area protected by this armor, then he benefits from 7 AV.

\subsubsection{Cutting Through Armor (Optional Rule)} %MAKE LIST OF ALL OPTIONAL RULES?
Some people may be bothered by the idea of Humans possessing the strength to seriously cleave through metal armor.
It is certainly true that such feats are rare in history. Julius Caesar spoke of a soldier whose helmet and face were 
equally mangled by a Spanish saber, and William the Conqueror was said to be able to cleave a man’s helmet and 
head in half with one blow of his sword. Charles the Bold of Burgundy was also killed in this manner by a halberd. 
Japanese accounts also mentioned similar feats, and for a time there was a sort of sport built up around splitting 
helmets, called Kabutowari.

However, with the exception of rare events such as these, cleaving through helmets and other metal armors with 
swords and the like was simply not something that happened, and so encountering it regularly in Song of Swords may 
leave a bad taste in your mouth. If that is so, this optional rule is here to save the day! 

Whenever a cutting weapon inflicts damage to an area protected by Hard armor, change it to bludgeoning damage, but
still resolve it against the Cutting AV of the armor. At the GM’s discretion, if the damage inflicted exceeds TOU+AV 
by 4 or more, it may cleave through the armor and inflict a cutting wound.


\section{Armor Characteristics}

\subsubsection{Component Name}
The name of the armor component.

\subsubsection{AVC / AVP / AVB}
The armor component’s armor value against cutting, piercing and bludgeoning damage.

\subsection{Layering Armor}
Armor can be layered (worn over other armor) within reason, but the AVs do not normally layer, unless otherwise 
stated. However, layered armor always uses the highest AV of the layered pieces against any attack to that area. This is
after factoring in Special Qualities like AP, Textile, Hard, Draw, Layer, and so on.
Whichever Armor’s AV is used to defend against an attack, only that Armor’s Special Qualities (with the exception of 
Layer) are in play for the purposes of that Attack. In the event of a tied AV, the wearer of the armor may choose which 
Armor’s AV to use. The qualities of that Armor are then in play, all others are ignored.

Example 1: An attack with Saber is made on a body part armored with Quilt and Mail armor. The Quilt’s AVC is only 
2, whereas the Mail’s AVC is 6. The Mail’s AV is used, and so the Hard and Mail Qualities are also in play, while the 
Textile Quality is not.

Example 2: An attack is made with an Ahlspiess against a body part armored with Plate Armor and Mail. The Plate 
Armor’s AVP is 7, while the Mail’s AVP is 4. The Plate Armor’s AVP of 7 is increased to 8 by the Mail, and is used 
against the attack. The Ahlspiess’s Thin Blade Quality is not used, since the armor it is targeting is Plate, but its AP Thrust of 3 is still in effect.

\subsection{Underarmor}
Realistically, no armor is worn on its own. Plate armor was usually worn with clothes specifically designed to pad the 
wearer from blows, and sometimes had loops to attach elements of the armor to it, with the intention of better 
distributing the armor’s weight. Mail was often worn with cloth backing to cushion impacts and to snag missiles with 
thin tips that penetrate through them.
The components presented in this chapter assume the bare minimum amount of underarmor in order for the armor to 
function. Realistically, you will want to purchase quilt armor to accompany a great many armors in this section.

\subsubsection{Coverage}
The Coverage of an armor component details what Hit Locations the piece covers and provides AV to. Armor of the 
same and differing types can be layered, however unless they have the Layers Special Quality, multiple helmets 
cannot be worn simultaneously.

If an armor component’s Coverage lists ‘Full X,’ with X being a limb (Arm, Leg), the Torso, or the Head, then every 
Hit Location on that section of the body is covered by the armor. 

\begin{itemize}
\item Full Leg refers to the Thigh, Knee, Shin and Foot.
\item Full Arm refers to the Shoulder, Upper Arm, Elbow, Forearm and Hand.
\item Full Torso refers to the Chest, Side, Hip, and Belly.
\item Full Head refers to the Upper Head, Lower Head, and Face.
\item The Neck and Groin are never included in ‘Full X’ armor listings, and are not covered unless they are listed 
separately.
\end{itemize}

Further considerations to armor coverage include:
\begin{itemize}
\item All armor is considered symmetrical by default. The cost and weight of any Shoulder, Arm or Leg armor 
includes two pieces, one for each Shoulder, Arm, Leg, and so on.
\item If only one piece of a symmetrical armor is worn, as in, one gauntlet or one pauldron, halve the Weight of 
that armor. This only applies when no matching armor (or Weight 0 armor) is being worn on the opposite 
side.
\item If two different pieces of symmetrical armor are worn, as in, a gauntlet on one hand and a mail glove on 
the other, then the Weight of the armor is averaged between them (combine the Weight of the two armors, and then halve the result).
\end{itemize}

\subsubsection{Half AV Coverage}
Some Hit Locations will be listed with this symbol: *. This means that the armor provides only half AV Coverage to that Hit Location. 
	
\subsubsection{Weak Spots}\label{sec:Weak Spots}
Some Hit Locations will be listed with this symbol: $\phi$. This indicates a weak spot in the armor. Weak spots can be targeted using the \nameref{man:Joint Thrust} Maneuver.

\subsubsection{Special}
The Special Qualities of an armor component note unusual and special effects of that armor, such as it layering with other armor in a unique way, providing unique armor bonuses against certain types of attacks, and influencing Encumbrance/Exhaustion differently than normal, and so on. These Special Qualities are detailed at the end of the chapter.

\subsubsection{Weight}
Armor Weight measures the amount of subjective weight that wearing the armor imposes on the character, not its 
literal weight in pounds. Much as modern backpacks endeavor to distribute weight evenly between shoulders and hips,
armor worn on the torso is much less encumbering than armor worn on limbs and extremities.
Armor Weight across the character is added together, and used for determining their current Encumbrance. Don’t 
forget that backpacks and other non-armor loads are also used in determining Encumbrance.

\subsubsection{Perception Penalty}
Only helmets and other headgear have a penalty to PER (noted in the tables as PP). It is a penalty applied to PER so long as the helmet is being worn. This represents the helmet obscuring vision and limiting breathing. This penalty can sometimes be alleviated by raising or lowering the helmet’s visor, as detailed in the helmet’s Special Quality section, or the section of its visor if the visor is attached.

\subsubsection{Cost}
This is the cost of the piece of armor. At the GM’s discretion, this number can vary greatly depending on availability 
and the local economic climate. If a gigantic battle was fought in the fields outside of the city, and the markets are 
now flooded with the looted swords, armor and other goods of the dead, then it is quite reasonable for the cost of 
second-hand gear to have gone down significantly.

\section{Pre-made Armor Sets}
The following are pre-made sets of armor consisting of individual armor components from the options below. Each 
armor set specifies the components included, which details their individual AV values, Coverage, Special, Weight and 
Cost. A description of the set, including who would have worn it, when and where it is from, is also detailed. 

You can easily modify these sets by removing or adding a component here and there. Be sure to factor in the new component’s values, when you do. None of these pre-made armor sets come with helmets; you can choose your own from the lists below.

\subsection*{Full Plate Armor}
This is full plate armor which would commonly be worn by knights, wealthy mercenaries and men at arms, as well as 
nobility. Some versions would have Pauldrons instead of Spaulders and Besagews, or a Bevor instead of a Gorget.
Helmet not included.

\begin{longtabu} to \linewidth {X[2,l] *{3}{X[-2,c]} X[3,l] X[3,l] *{2}{X[-1,c]}}
	\caption{Full Plate Armor}
	\label{tab:Full Plate Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Besagews& 8& 7& 6& Shoulders (Thrust only), Upper Arm* (Thrust only) &Hard, When layered with Spaulders use Besagews value against Thrusts.&0.5&8 sp\\
Couters& 8& 7& 6& Elbow\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1& 5 sp\\
Cuirass& 8& 7& 6& Belly, Chest, Lower 
Back, Side, Upper Back& Hard& 2.5& 2 gp\\
Cuisses& 8& 7& 6& Thigh& Hard& 2& 8 sp\\
Fauld& 8& 7& 6& Groin\hyperref[sec:Weak Spots]{$\phi$}, Hip\hyperref[sec:Weak Spots]{$\phi$}, Thighs\hyperref[sec:Weak Spots]{$\phi$} &Hard, Layers 2 (Groin, Hip, 
Thighs)&1.5& 10 sp\\
Gauntlets& 8& 7& 6& Hand\hyperref[sec:Weak Spots]{$\phi$}& Hard& 2& 5 sp\\
Gorget& 8& 7& 6& Neck& Hard& 1& 10 sp\\
Greaves& 8& 7& 6& Shin& Hard& 1.5& 6 sp\\
Mail Coat& 6& 4& 3& Full Torso, Groin, Hip, 
Knees, Shoulders, 
Thighs &Hard, Layer 1, Mail& 2.5& 2 gp\\
Mail Long Sleeves& 6& 4& 3& Elbow, Forearm, 
Shoulder, Upper Arm &Hard, Layer 1, Mail& 1.5& 1 gp\\
Poleyns& 8& 7& 6& Knee\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1.5& 5 sp\\
Quilted Shirt& 2& 3& 4& Full Torso, Groin, Hip, 
Thighs& Textile& 0& 1 sp\\
Quilted Leggings& 2& 3& 4& Groin, Hip, Knees, 
Shins, Thighs& Textile& 0& 2 cp\\
Quilted Long 
Sleeves& 2& 3& 4& Elbow, Forearm, 
Shoulder, Upper Arm& Textile& 0& 6 cp\\
Rerebraces& 8& 7& 6& Upper Arm& Hard& 1& 5 sp\\
Sabatons& 8& 7& 6& Foot& Hard& 1.5& 5 sp\\
Spaulders& 8& 7& 6& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}* &Hard, Provides half AV against Thrust attacks to the Shoulder 
and Upper Arm& 0.5& 6 sp\\
Vambraces& 8& 7& 6& Forearm& Hard& 1& 5 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 21.5 Wt \quad 6 gp 19 sp 8 cp}\\
\end{longtabu}

\subsection*{Half-Plate Armor}
Half-Plate Armor is that which covers about half of one's body in plate. This is the sort of armor popular amongst infantry (such as the Rodeleros of Spain and pikemen of various nations), and also that of mercenaries and noblemen 
who dislike the weight of heavier armor. Helmet not included.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Half-Plate Armor}
	\label{tab:Half-Plate Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Cuirass& 8& 7& 6& Belly, Chest, Lower Back, Side, Upper Back& Hard& 2.5& 2 gp\\
Fauld& 8& 7& 6& Groin\hyperref[sec:Weak Spots]{$\phi$}, Hip\hyperref[sec:Weak Spots]{$\phi$}, Thighs\hyperref[sec:Weak Spots]{$\phi$}& Hard, Layers 2 (Groin, Hip, Thighs)& 1.5& 10 sp\\
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Leather Leggings& 3& 2& 2& Groin, Hip, Knees, Shins, Thighs& -& 0& 2 sp\\
Reinforced Leather Gloves& 4& 3& 3& Hand& -& 0& 2 sp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 0.5& 17 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\
Spaulders& 8& 7& 6& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}*& Hard, Provides half AV against Thrust attacks to the Shoulder and Upper Arm& 0.5& 6 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill  5 Wt \quad 3 gp 2 sp 11 cp}\\
\end{longtabu}
\subsection*{Plate and Mail Armor}
This is an older sort of plate armor, where plate components are worn over a coat of mail, so that the gaps where the 
plate doesn’t cover are still armored. Attacks that slip through gaps must still defeat the mail to injure the wearer. It is cheaper and lighter than Full Plate Armor, but it provides less comprehensive protection. Helmet not included.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Plate and Mail Armor}
	\label{tab:Plate and Mail Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special&Wt.&Cost\\
Couters& 8& 7& 6& Elbow \hyperref[sec:Weak Spots]{$\phi$}& Hard& 1& 5 sp\\
Cuirass& 8& 7& 6& Belly, Chest, Lower Back, Side, Upper Back& Hard& 2.5& 2 gp\\
Greaves& 8& 7& 6& Shin& Hard& 1.5& 6 sp\\
Leather Leggings& 3& 2& 2& Groin, Hip, Knees, Shins, Thighs& -& 0& 2 sp\\
Mail Boots& 6& 4& 3& Feet& Hard, Layer 1, Mail& 1& 3 sp\\
Mail Coat& 6& 4& 3& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Hard, Layer 1, Mail& 2.5& 2 gp\\
Mail Gloves/Mittens& 6& 4& 3& Hand& Hard, Layer 1, Mail& 0.5& 3 sp\\
Mail Long Sleeves& 6& 4& 3& Elbow, Forearm, Shoulder, Upper Arm& Hard, Layer 1, Mail& 1.5& 1 gp\\
Poleyns& 8& 7& 6& Knee\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1.5& 5 sp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 0.5& 17 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill  12.5 Wt \quad 6 gp 5 sp 11 cp}\\
\end{longtabu}
\subsection*{Leather Outfit}
Leather wasn't generally worn as armor. While leather clothing was common enough, and plenty of warriors wore 
gloves, jerkins and so forth, it wasn't really expected to protect the wearer from swords or spears. Rawhide, or thick 
leather like that used in buffcoats, however, could definitely thwart a sword-blow, and for a time, leather coats were 
issued to infantry and cavalrymen alike to wear as armor. This suit takes it a bit further, it's just a full suit of leather. It 
weighs virtually nothing, isn’t that expensive, and it provides comprehensive, if minimal, protection to the whole 
body. Helmet not included, but why would you bother? Just buy a hat.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Leather Outfit}
	\label{tab:Leather Outfit}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Leather Buffcoat& 3& 2& 2& Full Torso, Groin, Hip, 
Knees, Shoulders, 
Thighs& -& 0& 5 sp\\
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Leather Gloves& 3& 2& 2& Hand& -& 0& 1 sp\\
Leather Leggings& 3& 2& 2& Groin, Hip, Knees, 
Shins, Thighs& -& 0& 2 sp\\
Leather Neck-Stock& 3& 2& 2& Neck& -& 0& 8 cp\\
Leather Sleeves& 3& 2& 2& Elbow, Forearm, 
Shoulder, Upper Arm& -& 0& 2 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 0 Wt \quad - gp 11 sp 8 cp}\\
\end{longtabu}

\subsection*{Lamellar Outfit}
This armor is a full suit of lamellar, though it could pass just as easily for a full suit of scale. Lamellar is tough, and 
provides good protection against all damage types. The suit is actually very heavy, as one might expect a suit that 
uniformly covers and protects every inch of one's body beneath the neck. It provides excellent protection and does not
require particularly advanced technology to produce. This general sort of armor, with small scales or squares of metal 
attached either to each other with lacing or to a backing (which is the chief difference between lamellar and scale) 
goes way back. Bronze Age warriors wore this sort of armor, as did certain Roman and Persian cavalrymen, as well as
Medieval Chinese, Japanese and Korean warriors. The ubiquity of lamellar and scale armor even exceeds that of mail. 
Helmet not included.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Lamellar Outfit}
	\label{tab:Lamellar Outfit}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Lamellar Boots& 5& 5& 5& Feet& Hard& 1.5& 10 sp\\
Lamellar Coat& 5& 5& 5& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Hard& 3& 3 gp\\
Lamellar Gloves/Mittens& 5& 5& 5& Hand& Hard& 1& 3 sp\\
Lamellar Leggings& 5& 5& 5& Groin, Hip, Knees, Shins, Thighs& Hard& 3& 12 sp\\
Lamellar Long Sleeves& 5& 5& 5& Elbow, Forearm, Shoulder, Upper Arm& Hard& 4& 1 gp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 12.5 Wt \quad 5 gp 5 sp - cp}\\
\end{longtabu}

\subsection*{Winged Hussar Armor}
This is an example of the highly-varied armor worn by the Winged Hussars of Poland. In the Tattered Realms, a very 
similar armor might be worn by the Husaria of the Lendian Commonwealth, and also by the Pontifatrix Guard, who 
defend the Empress of Illegon. This is by no means a uniform; every Hussar was self-equipped, though there was a 
dominant fashion among them. This example armor provides good protection against missiles, swords, and even 
lances and bullets, while not impeding movement too much.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Winged Hussar Armor}
	\label{tab:Winged Hussar Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Anima Cuirass& 7& 7& 7& Belly, Chest, Side& Hard& 3& 48 sp\\
Anima Cuisses& 7& 7& 7& Thigh& Hard& 3& 1 gp\\
Anima Greaves& 7& 7& 7& Shin& Hard& 2& 1 gp\\
Anima Poleyns& 7& 7& 7& Knee& Hard& 2& 10 sp\\
Anima Rerebraces& 7& 7& 7& Upper Arm& Hard& 1& 1 gp\\
Anima Spaulders& 7& 7& 7& Shoulder, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}*& Hard, Provides half AV against Thrust attacks to the Shoulder and Upper Arm& 2& 1 gp\\
Gorget& 8& 7& 6& Neck& Hard& 1& 10 sp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 0.5& 17 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\
Vambraces& 8& 7& 6& Forearm& Hard& 1& 5 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 15.5 Wt \quad 7 gp 14 sp 11 cp}\\
\end{longtabu}

\subsection*{Krajini Footman's Uniform}
The military of Krajina are known for being professional, disciplined, and fabulously dressed. These outfits are 
produced by the Guild of Tailors, Armorers and Hatmakers, and are magnificently decorated. The armor provides 
excellent protection against all sorts of attacks, but primarily focuses on defending the core of the body with layered brigandine and quilt armor. The limbs are protected with hidden jack chains on the inside of the surcoat, and the neck 
is protected with a thick and high collar of leather that also makes the wearer look more imposing. This armor does 
not include the standard-issue kettle helm, but you can buy one for yourself for 3 sp, a bargain at any price!

The officer's version of this armor doesn't have any protective improvements, but it costs twice as much and is 
obscenely fashionable.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Krajini Footman’s Uniform}
	\label{tab:Krajini Footman's Uniform}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Brigandine Coat& 5& 6& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Hard& 3& 3 gp\\
Couters& 8& 7& 6& Elbow\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1& 5 sp\\
Jack Chains& 6& 0& 3& Elbow\hyperref[sec:Weak Spots]{$\phi$}, Forearm\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& Hard, No AV against Thrust attacks of any damage type& 0.5& 1 sp\\
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Leather Leggings& 3& 2& 2& Groin, Hip, Knees, Shins, Thighs& -& 0& 2 sp\\
Leather Neck-Stock& 3& 2& 2& Neck& -& 0& 8 cp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 0.5& 17 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\
Reinforced Leather Gloves& 4& 3& 3& Hand& -& 0& 2 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 5 Wt \quad 3 gp 13 sp 7 cp}\\
\end{longtabu}

\begin{longtabu} to \linewidth {X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{DIS Uniform}
	\label{tab:DIS Uniform}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Mirror Plate& 10& 9& 8& Chest\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1.5& 1 gp\\
Quilted Leggings& 2& 3& 4& Groin, Hip, Knees, Shins, Thighs& Textile& 0& 2 sp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 0.5& 17 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\
Reinforced Leather Gloves& 4& 3& 3& Hand& -& 0& 2 sp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 2 Wt \quad 1 gp 7 sp}\\
\end{longtabu}

\subsection*{Fighting Volkodav Outfit}
The Fighting Volkodavs are a Dacian paramilitary organization formed by the Colonial Sich Authority to range the 
Steppe. Officially, their mission is to protect the colonies. Unofficially, their mission is to exterminate every living 
thing in the Steppe that isn't a Dacian. The Marju steppelanders have long been the enemy of Dace, and the Volkodavs
were created to be the final resolution of the feud. The Volkodavs wear armor intended for long deployment. The 
heavy quilt keeps them warm on the freezing steppe, and combined with their chainmail coats, they are virtually 
immune to arrows. Just in case, though, many of them also wear a mirror plate on their chests. Volkodav armor is 
much cheaper than full suits of plate, and much lighter. In the highly mobile environment of the Infinite Steppe, this armor is invaluable. The Volkodavs wear special helmets, Zischagges, with a movable visor (made to resemble a 
“hounskull,” a hound's skull, or dog's face) visor, and intimidating wolf fur crests. You can get one too, they're only 1 gp.

\begin{longtabu} to \linewidth {X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\caption{Fighting Volkodav Outfit}
	\label{tab:Fighting Volkodav Outfit}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\
Greaves& 8& 7& 6& Shin& Hard& 1.5& 6 sp\\
Heavy Quilted Coat& 3& 4& 5& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Textile& 1& 5 sp\\
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Mail Coat& 6& 4& 3& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& Hard, Layer 1, Mail& 2.5& 2 gp\\
Mail Long Sleeves& 6& 4& 3& Elbow, Forearm, Shoulder, Upper Arm& Hard, Layer 1, Mail& 1.5& 1 gp\\
Mirror Plate& 10& 9& 8& Chest\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1.5& 1 gp\\
Poleyns& 8& 7& 6& Knee\hyperref[sec:Weak Spots]{$\phi$}& Hard& 1.5& 5 sp\\
Quilted Leggings& 2& 3& 4& Groin, Hip, Knees, Shins, Thighs& Textile& 0& 2 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& Textile& 0& 6 cp\\\cmidrule(lr){1-8}
\multicolumn{8}{c}{\textbf{TOTAL}:\hfill 9.5 Wt \quad 4 gp 17 sp 8 cp}\\
\end{longtabu}

\section*{Armor Components}
\section{Head and Neck Armor}\label{sec:Head and Neck Armor}
\begin{longtabu}to \linewidth {X *{3}{X[-1,c]} X[2] X[2] *{3}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Helmets} \vspace{-15pt}
	\label{tab:Helmets}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&PP& Cost\\\toprule
Armet& 8& 7& 6& Full Head, Face\hyperref[sec:Weak Spots]{$\phi$}, Neck &\nameref{par:Restricts Breathing} 2, Integrated Visor, \nameref{par:Hard}& 4& -2& 2 gp\\
Barbute& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}*, Lower Head, Upper Head &\nameref{par:Hard}& 1 &-1& 10 sp\\
Bascinet& 8& 7& 6& Lower Head, Upper Head &\nameref{par:Hard}, +1 AV vs Downward Swings, Can be worn with Great Helm for \nameref{par:Layer} 1 on Upper Head and Lower Head& 1.5& 0& 12 sp\\
Burgonet& 8& 7& 6& Lower Head, Upper Head &\nameref{par:Hard}, +1 AV vs Downward Swings &2 &-1& 10 sp\\
Close Helmet& 8& 7& 6& Full Head, Face\hyperref[sec:Weak Spots]{$\phi$}, Neck &\nameref{par:Restricts Breathing} 2, Integrated Visor, \nameref{par:Hard}& 3.5 &-1& 3 gp\\
Closed Sallet& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}, Lower Head*, Upper Head &\nameref{par:Hard}, No AV on Face vs Upward Swings &2 &-1& 10 sp\\
Conical Helm& 8& 7& 6& Upper Head &\nameref{par:Hard}, +2 AV vs Downward Swings 2& 0& 2& sp\\
Great Bascinet& 8& 7& 6& Full Head, Face\hyperref[sec:Weak Spots]{$\phi$}, Neck, Shoulder &\nameref{par:Restricts Breathing} 2, Integrated Visor, \nameref{par:Hard}& 3.5 &-2& 15 sp\\
Great Helm& 8& 7& 6& Full Head, Face\hyperref[sec:Weak Spots]{$\phi$}, Neck\hyperref[sec:Weak Spots]{$\phi$} &\nameref{par:Restricts Breathing} 3, \nameref{par:Hard}, Can be worn with a Skullcap or Bascinet for \nameref{par:Layer} 1 on overlapping areas& 5 &-3& 30 sp\\
Hat/Cap& 1& 1& 1& Upper Head &-& 0& 0& 1 cp\\
Kettle Helm& 8& 7& 6& Lower Head*, Upper Head &\nameref{par:Hard}, No AV vs Upward Swings &0& 0& 3 sp\\
Lamellar Pot Helm& 5& 5& 5& Lower Head, Upper Head &\nameref{par:Hard}& 2 &-1& 8 sp\\
Leather Pot Helm& 4& 3& 3& Lower Head, Upper Head &-& 1 &-1& 10 cp\\
Leather Skullcap& 4& 3& 3& Upper Head &-& 0& 0& 6 cp\\
Morion& 8& 7& 6& Lower Head*, Upper Head &\nameref{par:Hard}, +1 AV vs Downward Swings 1& 0& 1&2 sp\\
Open Sallet& 8& 7& 6& Lower Head*, Upper Head &\nameref{par:Hard}& 2& 0& 5 sp\\
Quilted Helmet& 2& 4& 5& Lower Head, Upper Head &\nameref{par:Textile}& 0& 0& 8 cp\\
Reinforced Roman Helm& 8& 7& 6& Lower Head, Neck*, Shoulders*, Upper Head &\nameref{par:Hard}, +4 AV on Upper Head& 2.5& 0& 1 gp\\
Roman Helm (Gallic Style)& 8& 7& 6& Lower Head, Neck* Shoulders*, Upper Head, &\nameref{par:Hard}& 2& 0& 15 sp\\
Skullcap& 8& 7& 6& Upper Head &\nameref{par:Hard}, Can be worn with Great Helm for \nameref{par:Layer} 2 on Upper Head& 0& 0& 1 sp\\
Stechhelm& 10& 9& 8& Full Head, Neck &\nameref{par:Restricts Breathing} 3, \nameref{par:Hard}, +4 AV on Lower Head, Face, and Neck& 8 &-3& 1 gp\\
Tailed Sallet& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}, Lower Head*, Neck*, Shoulders*, Upper Head &\nameref{par:Hard}, No AV on Face vs Upward Swings &2& -1& 12 sp\\
Zischagge/ Szyszak& 8& 7& 6& Lower Head, Neck*, Shoulders*, Upper Head &\nameref{par:Hard}, +4 AV on Upper Head& 2.5& 0& 10 sp\\
\end{longtabu}

\subsection{\label{sec:Visors}Optional Visors}
Any Helmet that does not cover the Face can be equipped with a visor, which can be chosen from the list below. The visor then becomes part of the helmet, applying AV and Coverage as normal. If a visor type covers more than the Face, it cannot be attached to any helmet that already covers any of those additional areas.

\paragraph{\label{par:Movable Visors}Movable Visors}---\quad
A visor can be opened or closed at will (in combat this requires an ADR test at RS 1, and 3 CP). When open, a visor ceases to provide any AV for the Hit Location, but also removes all PP and Weight penalties from the visor for as long as it is open. Closing a visor is much easier than opening one, and can be done with a sharp nod of the head (no ADR test, and only 1 CP while in combat). There is a darker side to this as well. Opening an enemy’s visor during a grapple is easier than trying to pull off his helmet (see the \nameref{ch:Maneuvers}~chapter for more on grappling).

\begin{longtabu} to \linewidth {X *{3}{X[-1,c]} X[2] X[2] *{3}{X[-1,c]}}
	\caption{Visors}
	\label{tab:Visors}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&PP&Cost\\\toprule
Bar-Nasel& 4& 2& 1& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, No AV vs Upward Swings &0& 0& 1 sp\\
Cage Visor& 7& 3& 5& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, Movable& 1& 0& 5 sp\\
Falling Buff 
(Hinged Gorget)& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}, Neck& \nameref{par:Restricts Breathing} 2, \nameref{par:Hard}, 
Movable& 2 &-1& 1 gp\\
Full Visor& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Restricts Breathing} 2, \nameref{par:Hard}, 
Movable& 1.5 &-1& 10 sp\\
Goggle Type& 5& 4& 3& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, No AV vs Upward Swings &0.5 &-1& 3 sp\\
Mail Ventail& 6& 4& 3& Face& \nameref{par:Layer} 1, \nameref{par:Hard}& 1 &-1& 2 sp\\
Tri-Bar Nasel& 6& 3& 2& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, No AV vs Upward Swings &0.5& 0& 3 sp\\
\end{longtabu}

\subsubsection{Optional Helmet Crests}
You may add an optional decorative crest to your helmet by paying an additional cost and adding the listed weight (if any) to the helmet.

\vspace*{-10pt}\paragraph{Identifying Crest}---\quad 0 Weight \quad 1 gp\par
Large and visible crest or plume (like that of a Roman Centurion), which aids in troop identification. You gain a +1 bonus to Tactics or Strategy rolls made in combat.

\vspace*{-10pt}\paragraph{Intimidating Crest}---\quad 0 Weight \quad 1 gp\par
Intimidating crests could be war-crests, horns, depict demonic imagery, or have a 
heightening effect, and so on. Opponents who have WIL lower than yours suffer a -1 
penalty to their CP for the first Round of combat they fight with you unless they have 
the \nameref{boon:Brave}~Boon. If they are \nameref{bane:Craven}, the -1 penalty will remain throughout combat, so long as the helmet is worn. 

\vspace*{-10pt}\paragraph{Ruthlessly Fashionable Crest}---\quad 1 Weight \quad 3 gp\par
These are magnificent paper maché or wooden crests, and possibly have mechanical 
movements. You gain a +1 bonus to Leadership rolls made in combat.

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{3}{X[-1,c]}}
	\caption{Non-Helmet Head Protection}
	\label{tab:Non-Helmet Head Protection}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&PP&Cost\\\toprule
Cloth Mask/Scarf& 1& 1& 1& Face, Lower Head &\nameref{par:Textile}&0&0&1 cp\\
Lamellar Ventail& 5& 5& 5& Face, Neck &\nameref{par:Hard} &1.5 &-1 &2 sp\\
Mail Camail& 6& 4& 3& Lower Head, Neck, Shoulders &\nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1& 0& 4 sp\\
Mail Coif& 6& 4& 3& Lower Head, Neck, Upper Head& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1& 0& 2 sp\\
Mail Ventail& 6& 4& 3& Face& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1.5 &-1& 1 sp\\
Quilted Camail& 2& 3& 4& Lower Head, Neck, Shoulders& \nameref{par:Textile}& 0& 0& 8 cp\\
Quilted Coif& 2& 3& 4& Lower Head, Neck, Upper Head& \nameref{par:Textile}& 0& 0& 4 cp\\
Quilted Ventail& 2& 3& 4& Face& \nameref{par:Textile}& 0 &-1& 1 cp\\
Steel Facemask& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Restricts Breathing} 1, \nameref{par:Hard}& 2 &-1& 8 sp\\
\end{longtabu}

\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{3}{X[-1,c]}}
	\caption{Neck Protection}
	\label{tab:Neck Protection}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&PP& Cost\\\toprule
Bevor& 8& 7& 6& Face\hyperref[sec:Weak Spots]{$\phi$}*, Lower Head, Neck& \nameref{par:Restricts Breathing} 1, \nameref{par:Hard}& 2& 0& 12 sp\\
Gorget& 8& 7& 6& Neck& \nameref{par:Hard}& 1& 0& 10 sp\\
Leather Neck-Stock& 3& 2& 2& Neck &-& 0& 0& 8 cp\\
Neck-Wrap (such as a Scarf)& 1& 1& 1& Neck& \nameref{par:Textile}& 0& 0& 1 cp\\
Reinforced Leather Neck-Stock& 4& 3& 3& Neck& \nameref{par:Hard}& 0& 0& 1 sp\\
\end{longtabu}

\section{Torso Armor}
\subsection{Plate Torso Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Plate Torso Armor} \vspace{-15pt}
	\label{tab:Plate Torso Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Breastplate& 8& 7& 6& Belly, Chest, Side& \nameref{par:Hard}& 1.5& 1 gp\\
Bulletproof Breastplate& 10& 9& 8& Belly, Chest, Side& \nameref{par:Bulletproof}, \nameref{par:Hard}& 3.5& 2 gp\\
Bulletproof Cuirass& 10& 9& 8& Belly, Chest, Lower Back, Side, Upper Back& \nameref{par:Bulletproof}, \nameref{par:Hard}& 4.5& 4 gp\\
Bulletproof Codpiece& 10& 9& 8& Groin& \nameref{par:Bulletproof}, \nameref{par:Hard}& 0.5& 1 gp\\
Cuirass& 8& 7& 6& Belly, Chest, Lower Back, Side, Upper Back& \nameref{par:Hard}& 2.5& 2 gp\\
Codpiece& 8& 7& 6& Groin& \nameref{par:Hard}& 0& 5 sp\\
Fauld& 8& 7& 6& Groin\hyperref[sec:Weak Spots]{$\phi$}, Hip\hyperref[sec:Weak Spots]{$\phi$}, Thighs\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, \nameref{par:Layer} 2 (Groin, Hip, Thighs)& 1.5& 10 sp\\
Heartguard& 8& 7& 6& Chest\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 0& 1 sp\\
Mirror Plate& 10& 9& 8& Chest\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 1.5& 1 gp\\
Munitions Breastplate& 7& 6& 5& Chest& \nameref{par:Hard}& 2.5& 10 sp\\
Munitions Cuirass& 7& 6& 5& Belly\hyperref[sec:Weak Spots]{$\phi$}, Chest\hyperref[sec:Weak Spots]{$\phi$}, Side\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3.5& 15 sp\\
Munitions Fauld& 7& 6& 5& Groin\hyperref[sec:Weak Spots]{$\phi$}, Hip\hyperref[sec:Weak Spots]{$\phi$}, Thighs\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, \nameref{par:Layer} 1 (Groin, Hip, Thighs)& 2.5& 10 sp\\
Plackart& 8& 7& 6& Belly\hyperref[sec:Weak Spots]{$\phi$}, Hip\hyperref[sec:Weak Spots]{$\phi$}, Side\hyperref[sec:Weak Spots]{$\phi$},& \nameref{par:Hard}, \nameref{par:Layer} 2 (Belly, Side)& 0.5& 1 gp\\
\end{longtabu}
\subsection{Laminar Torso Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Laminar Torso Armor} \vspace{-15pt}
	\label{tab:Laminar Torso Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Anima Cuirass& 7& 7& 7& Belly, Chest, Side& \nameref{par:Hard}& 3& 48 sp\\
Anima Fauld& 7& 7& 7& Groin\hyperref[sec:Weak Spots]{$\phi$}, Hip, Thighs\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3& 1 gp\\
Laminar Cuirass& 8& 7& 5& Belly\hyperref[sec:Weak Spots]{$\phi$}, Chest\hyperref[sec:Weak Spots]{$\phi$}, Side\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3& 1 gp\\
Segmented Tassets& 8& 7& 5& Groin, Hip& \nameref{par:Hard}& 3& 12 sp\\
\end{longtabu}
\subsection{Mail Torso Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Mail Torso Armor} \vspace{-15pt}
	\label{tab:Mail Torso Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Brigandine Coat& 5& 6& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Hard}& 3& 3 gp\\
Brigandine Shirt& 5& 6& 4& Full Torso, Groin, Hip, Thighs& \nameref{par:Hard}& 2& 30 sp\\
Brigandine Vest& 5& 6& 4& Full Torso& \nameref{par:Hard}& 1& 1 gp\\
Lamellar Coat& 5& 5& 5& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Hard}& 3& 3 gp\\
Lamellar Shirt& 5& 5& 5& Full Torso, Groin, Hip, Thighs& \nameref{par:Hard}& 2& 2 gp\\
Lamellar Vest& 5& 5& 5& Full Torso& \nameref{par:Hard}& 1& 1 gp\\
Mail Coat& 6& 4& 3& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 2.5& 2 gp\\
Mail Shirt& 6& 4& 3& Full Torso, Groin, Hip, Thighs& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1.5& 1 gp\\
Mail Vest& 6& 4& 3& Full Torso& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 0.5& 18 sp\\
\end{longtabu}
\subsection{Textile Torso Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Textile Torso Armor} \vspace{-15pt}
	\label{tab:Textile Torso Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Cloth Coat& 1& 1& 1& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Textile}& 0& 2 cp\\
Cloth Shirt/Vest& 1& 1& 1& Full Torso, Groin, Hip, Thighs& \nameref{par:Textile}& 0& 1 cp\\
Heavy Quilted Coat& 3& 4& 5& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Textile}& 1& 5 sp\\
Heavy Quilted Shirt& 3& 4& 5& Full Torso, Groin, Hip, Thighs& \nameref{par:Textile}& 0.5& 3 sp\\
Heavy Quilted Vest& 3& 4& 5& Full Torso& \nameref{par:Textile}& 0& 2 sp\\
Quilted Coat& 2& 3& 4& Full Torso, Groin, Hip, Knees, Shoulders, Thighs& \nameref{par:Textile}& 0.5& 17 cp\\
Quilted Shirt& 2& 3& 4& Full Torso, Groin, Hip, Thighs& \nameref{par:Textile}& 0& 1 sp\\
Quilted Vest& 2& 3& 4& Full Torso& \nameref{par:Textile}& 0& 5 cp\\
\end{longtabu}
\subsection{Leather Torso Armor}
\begin{longtabu}to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Leather Torso Armor} \vspace{-15pt}
	\label{tab:Leather Torso Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Leather Buffcoat& 3& 2& 2& Full Torso, Groin, Hip, Knees, Shoulders, Thighs &-& 0& 5 sp\\
Leather Jack& 3& 2& 2& Full Torso, Groin, Hip, Thighs &-& 0& 2 sp\\
Leather Lamellar Coat& 3& 4& 3& Full Torso, Groin, Hip, Knees, Shoulders, Thighs &-& 1& 9 sp\\
Leather Lamellar Shirt& 3& 4& 3& Full Torso, Groin, Hip, Thighs &-& 0.5& 6 sp\\
Leather Lamellar Vest& 3& 4& 3& Full Torso &-& 0& 3 sp\\
Leather Vest& 3& 2& 2& Full Torso &-& 0& 1 sp\\
\nameref{par:Hard}ened Leather Breastplate& 4& 3& 3& Belly, Chest, Side &-& 0& 5 sp\\
\nameref{par:Hard}ened Leather Cuirass& 4& 3& 3& Belly, Chest, Lower Back, Side, Upper Back &-& 1& 8 sp\\
Shagreen Greatcoat& 4& 3& 3& Full Torso, Full Arms, Groin, Hip, Knees, Shins, Thighs &-& 0& 1 gp\\
\end{longtabu}

\section{Arm Armor}
\subsection{Plate Arm Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Plate Arm Armor} \vspace{-15pt}
	\label{tab:Plate Arm Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Besagews& 8& 7& 6& Shoulders (Thrust only), Upper Arm* (Thrust only)& \nameref{par:Hard}, When \nameref{par:Layer}ed with Spaulders use Besagews value against Thrusts.& 0.5& 8 sp\\
Bulletproof Vambraces& 10& 9& 8& Forearm& \nameref{par:Bulletproof}, \nameref{par:Hard}& 2& 1 gp\\
Couters& 8& 7& 6& Elbow\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 1& 5 sp\\
Gauntlets& 8& 7& 6& Hand\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 2& 5 sp\\
Jack Chains& 6& 0& 3& Elbow\hyperref[sec:Weak Spots]{$\phi$}, Forearm\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}, No AV against Thrust attacks of any damage type& 0.5& 1 sp\\
Munitions Pauldrons& 8& 7& 6& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 4& 8 sp\\
Munitions Rerebraces& 7& 6& 5& Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3& 2 sp\\
Munitions Spaulders& 7& 6& 5& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}*& \nameref{par:Hard}, Provides half AV against Thrust attacks to the Shoulder and Upper Arm& 1& 4 sp\\
Munitions Vambraces& 7& 6& 5& Forearm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 1.5& 2 sp\\
Pauldrons& 10& 9& 8& Shoulder, Upper Arm& \nameref{par:Bulletproof}, \nameref{par:Hard}& 3& 1 gp\\
Rerebraces& 8& 7& 6& Upper Arm& \nameref{par:Hard}& 1& 5 sp\\
Spaulders& 8& 7& 6& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}*& \nameref{par:Hard}, Provides half AV against Thrust attacks to the Shoulder 
and Upper Arm& 0.5& 6 sp\\
Vambraces& 8& 7& 6& Forearm& \nameref{par:Hard}& 1& 5 sp\\
\end{longtabu}
\subsection{Laminar Arm Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Laminar Arm Armor} \vspace{-15pt}
	\label{tab:Laminar Arm Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Anima Pauldrons& 7& 7& 7& Shoulder, Upper Arm& \nameref{par:Hard}& 2& 2 gp\\
Anima Rerebraces& 7& 7& 7& Upper Arm& \nameref{par:Hard}& 1& 1 gp\\
Anima Spaulders& 7& 7& 7& Shoulder, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}*& \nameref{par:Hard}, Provides half AV against 
Thrust attacks to the Shoulder and Upper Arm& 2& 1 gp\\
Anima Vambrace& 7& 7& 7& Forearm& \nameref{par:Hard}& 2& 3 gp\\
Laminar Forearm  Guard& 8& 7& 5& Forearm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 2& 3 sp\\
Laminar Elbow Joint& 8& 7& 5& Elbow\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 2& 3 sp\\
Laminar Pauldrons& 8& 7& 5& Shoulder\hyperref[sec:Weak Spots]{$\phi$}, Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3& 10 sp\\
Laminar Rerebrace& 8& 7& 5& Upper Arm\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 1& 3 sp\\
\end{longtabu}
\subsection{Mail Arm Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Mail Arm Armor} \vspace{-15pt}
	\label{tab:Mail Arm Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Brigandine Gloves/Mittens& 5& 6& 4& Hand& \nameref{par:Hard}& 1& 3 sp\\
Brigandine Long Sleeves& 5& 6& 4& Elbow, Forearm, Shoulder, Upper Arm& \nameref{par:Hard}& 3& 2 gp\\
Brigandine Short Sleeves& 5& 6& 4& Shoulder, Upper Arm& \nameref{par:Hard}& 2& 14 sp\\
Lamellar Gloves/Mittens& 5& 5& 5& Hand& \nameref{par:Hard}& 1& 3 sp\\
Lamellar Long Sleeves& 5& 5& 5& Elbow, Forearm, Shoulder, Upper Arm& \nameref{par:Hard}& 4& 1 gp\\
Lamellar Short Sleeves& 5& 5& 5& Shoulder, Upper Arm& \nameref{par:Hard}& 2& 12 sp\\
Mail Gloves/Mittens& 6& 4& 3& Hand& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 0.5& 3 sp\\
Mail Long Sleeves& 6& 4& 3& Elbow, Forearm, Shoulder, Upper Arm& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1.5& 1 gp\\
Mail Short Sleeves& 6& 4& 3& Shoulder, Upper Arm& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 0.5& 12 sp\\
\end{longtabu}
\subsection{Leather Arm Armor}
\begin{longtabu}to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Leather Arm Armor} \vspace{-15pt}
	\label{tab:Leather Arm Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
\nameref{par:Hard}ened Leather Pauldrons& 4& 3& 3& Shoulder, Upper Arm\hyperref[sec:Weak Spots]{$\phi$} &-& 0& 5 sp\\
\nameref{par:Hard}ened Leather Rerebraces& 4& 3& 3& Upper Arm &-& 0& 3 sp\\
\nameref{par:Hard}ened Leather Vambraces& 4& 3& 3& Forearm &-& 0& 3 sp\\
Leather Gloves& 3& 2& 2& Hand &-& 0& 1 sp\\
Leather Sleeves& 3& 2& 2& Elbow, Forearm, Shoulder, Upper Arm &-& 0& 2 sp\\
Leather Lamellar Long Sleeves& 3& 4& 3& Elbow, Forearm, Shoulder, Upper Arm &-& 0& 3 sp\\
Leather Lamellar Short Sleeves& 3& 4& 3& Shoulder, Upper Arm &-& 0& 1 sp\\
Reinforced Leather Gloves& 4& 3& 3& Hand &-& 0& 2 sp\\
\end{longtabu}
\subsection{Textile Arm Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Textile Arm Armor} \vspace{-15pt}
	\label{tab:Textile Arm Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Cloth Gloves& 1& 1& 1& Hand& \nameref{par:Textile}& 0& 1 cp\\
Cloth Long Sleeves& 1& 1& 1& Elbow, Forearm, Shoulder, Upper Arm& \nameref{par:Textile}& 0& 2 cp\\
Cloth Short Sleeves& 1& 1& 1& Shoulder, Upper Arm& \nameref{par:Textile}& 0& 1 cp\\
Quilted Gloves& 2& 3& 4& Hand& \nameref{par:Textile}& 0& 1 cp\\
Quilted Long Sleeves& 2& 3& 4& Elbow, Forearm, Shoulder, Upper Arm& \nameref{par:Textile}& 0& 6 cp\\
Quilted Short Sleeves& 2& 3& 4& Shoulder, Upper Arm& \nameref{par:Textile}& 0& 3 cp\\
Reinforced Quilted Gloves& 4& 4& 5& Hand& \nameref{par:Textile}& 0& 1 sp\\
\end{longtabu}

\section{Leg Armor}
\subsection{Plate Leg Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Plate Leg Armor} \vspace{-15pt}
	\label{tab:Plate Leg Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Bulletproof Cuisses& 10& 9& 8& Thigh& \nameref{par:Bulletproof}, \nameref{par:Hard}& 4& 2 gp\\
Bulletproof Greaves& 10& 9& 8& Shin& \nameref{par:Bulletproof}, \nameref{par:Hard}& 2& 2 gp\\
Bulletproof Poleyns& 10& 9& 8& Knee& \nameref{par:Bulletproof}, \nameref{par:Hard}& 2& 1 gp\\
Cuisses& 8& 7& 6& Thigh& \nameref{par:Hard}& 2& 8 sp\\
Greaves& 8& 7& 6& Shin& \nameref{par:Hard}& 1.5& 6 sp\\
Munitions Cuisses& 7& 6& 5& Thigh\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 3& 4 sp\\
Munitions Greaves& 7& 6& 5& Shin\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 2& 3 sp\\
Poleyns& 8& 7& 6& Knee\hyperref[sec:Weak Spots]{$\phi$}& \nameref{par:Hard}& 1.5& 5 sp\\
Sabatons& 8& 7& 6& Foot& \nameref{par:Hard}& 1.5& 5 sp\\
\end{longtabu}
\subsection{Laminar Leg Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Laminar Leg Armor} \vspace{-15pt}
	\label{tab:Laminar Leg Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Anima Cuisses& 7& 7& 7& Thigh& \nameref{par:Hard}& 3& 1 gp\\
Anima Greaves& 7& 7& 7& Shin& \nameref{par:Hard}& 2& 1 gp\\
Anima Poleyns& 7& 7& 7& Knee& \nameref{par:Hard}& 2& 10 sp\\
Laminar Boots& 8& 7& 5& Foot& \nameref{par:Hard}& 2& 4 sp\\
Laminar Cuisses& 8& 7& 5& Thigh& \nameref{par:Hard}& 3& 4 sp\\
Laminar Greaves& 8& 7& 5& Shin& \nameref{par:Hard}& 3& 4 sp\\
\end{longtabu}
\subsection{Mail Leg Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Mail Leg Armor} \vspace{-15pt}
	\label{tab:Mail Leg Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Brigandine Boots& 5& 6& 4& Feet& \nameref{par:Hard}& 1.5& 8 sp\\
Brigandine Leggings& 5& 6& 4& Groin, Hip, Knees, Shins, Thighs& \nameref{par:Hard}& 3& 10 sp\\
Lamellar Leggings& 5& 5& 5& Groin, Hip, Knees, Shins, Thighs& \nameref{par:Hard}& 3& 12 sp\\
Lamellar Boots& 5& 5& 5& Feet& \nameref{par:Hard}& 1.5& 10 sp\\
Mail Boots& 6& 4& 3& Feet& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 1& 3 sp\\
Mail Groin Gusset& 6& 4& 3& Groin& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 0& 3 sp\\
Mail Leggings& 6& 4& 3& Groin, Hip, Knees, Shins, Thighs& \nameref{par:Hard}, \nameref{par:Layer} 1, \nameref{par:Mail}& 2.5& 8 sp\\
\end{longtabu}
\subsection{Leather Leg Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Leather Leg Armor} \vspace{-15pt}
	\label{tab:Leather Leg Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Leather Boots& 3& 2& 2& Feet& -& 0& 1 sp\\
Leather Leggings& 3& 2& 2& Groin, Hip, Knees, Shins, Thighs& -& 0& 2 sp\\
\nameref{par:Hard}ened Leather Cuisses& 4& 3& 3& Hip, Thighs& -& 0& 8 cp\\
\nameref{par:Hard}ened Leather Greaves& 4& 3& 3& Shins& -& 0& 1 sp\\
\nameref{par:Hard}ened Leather Knee-Guards& 4& 3& 3& Knee& -& 0& 6 cp\\
Leather Lamellar Cuisses& 3& 4& 3& Hip, Thighs& -& 0& 3 sp\\
Leather Lamellar Greaves& 3& 4& 3& Shins& -& 0& 3 sp\\
\end{longtabu}
\subsection{Textile Leg Armor}
\begin{longtabu}to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Textile Leg Armor} \vspace{-15pt}
	\label{tab:Textile Leg Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Cloth Leggings& 1& 1& 1& Groin, Hip, Knees, 
Shins, Thighs& \nameref{par:Textile}& 0& 1 cp\\
Quilted Leggings& 2& 3& 4& Groin, Hip, Knees, 
Shins, Thighs& \nameref{par:Textile}& 0& 2 cp\\
\end{longtabu}

\section{Horse Armor}
\begin{longtabu} to \linewidth{X *{3}{X[-1,c]} X[2] X[2] *{2}{X[-1,c]}}
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Horse Armor} \vspace{-15pt}
	\label{tab:Horse Armor}\\
Component Name&AVC&AVP&AVB& Coverage& Special& Wt&Cost\\\toprule
Brigandine 
Barding& 5& 6& 4& Horse& -& 8& 20 gp\\
Caprisoning& 2& 4& 5& Horse& \nameref{par:Textile}& 1& 5 sp\\
Hardened Leather Barding& 4& 3& 3& Horse& -& 3& 1 gp\\
Lamellar Barding& 5& 5& 5& Horse& -& 8& 15 gp\\
Mail Barding& 6& 4& 3& Horse& -& 8& 10 gp\\
Plate Barding& 8& 7& 6& Horse& -& 12& 30 gp\\
\end{longtabu}

\section{Special Qualities}
\paragraph{\label{par:Hard}Hard}---\quad
	Armor with this quality is more resistant to certain weapon Special Qualities (such as Draw), and triggers certain effects with others. Otherwise, it has no effect
\vspace*{-10pt}\paragraph{\label{par:Mail}Mail}---\quad
	Armor with this quality consists of interlocking ringlets of metal, and triggers certain effects with other Special Qualities. Otherwise, it has no effect.

\vspace*{-10pt}\paragraph[Layer]{\label{par:Layer}Layer [X]}---\quad
	If worn on the same Hit Location as another armor component with equal to or higher AVC, AVP, and/or AVB, this armor increases the appropriate AV on that area (or multiple, if multiple are higher) by X. Multiple instances of Layer do not combine if multiple armors with Layer overlap. Only the highest Layer value is used. Armor that benefits from the Textile Special Quality does benefit from Layer if its new AV is higher than the AV of the armor with Layer.
\vspace*{-10pt}\paragraph{\label{par:Textile}Textile}---\quad
	This armor has double AVP against non-firearm missile weapons. Double the AVP before adding inother bonuses like Layer. For the purposes of layering armor, double the AV of this armor before deciding whether or not to use it.
\vspace*{-10pt}\paragraph{\label{par:Bulletproof}Bulletproof}---\quad
	Halves AP from missile weapons.
\vspace*{-10pt}\paragraph[Restricts Breathing]{\label{par:Restricts Breathing}Restricts Breathing [X]}---\quad
	As long as this helmet is on, or the visor is down, add X to the multiplier for your current Fatigue Gain Modifier.

\subsubsection{Materials}
\begin{itemize}
	\item [] \textbf{Bronze}
	\subitem -2 AV from all sources
	\subitem +25\% Cost
\end{itemize}

\begin{itemize}
	\item [] \textbf{Bronze}
	\subitem -2 AV from all sources
	\subitem -25\% Cost
\end{itemize}

\begin{itemize}
\item [] \textbf{Steel}
\subitem No change
\end{itemize}

\begin{itemize}
	\item [] \textbf{Orichalcum}
		\subitem Converts all cutting/piercing damage to bludgeoning, ignores crushing damage
		\subitem Halves Shock effect from weapons
		\subitem +4 Bludgeoning AV
		\subitem Ignores AP Swing/Thrust X for Melee Weapons and AP X for Missile Weapons
			\subsubitem Does not ignore the AP Thrust bonus from weapons with Thin Blade if armor is mail
		\subitem Negates all unarmed damage
		\subitem +5,000\% Cost
\end{itemize}

\begin{itemize}
	\item [] \textbf{Silversteel}
		\subitem +2 AV from all sources
		\subitem Armor has half normal Weight 
		\subitem +1,000\% Cost
\end{itemize}

\chapter{Burdinadin Armory}\label{ch:Burdinadin Armory}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quote}
	
	\emph{“I have an idea.”}\par
		\hfill --\textbf{Nikola Garrastazu} (Deceased), Head of Glade Apa Translocation Research Bureau
\end{quote}
\begin{quote}
	\emph{“Early experiments showed a suppressive effect on enemy personnel merely by activating the device, as the high and sustained 
rate of fire is a strong psychological challenge to the advancing enemy to commit to an assault. We have found that a sweepin g 
field of fire is superior to a penetrating ‘deep’ firing pattern, as it causes the full front ranks to hesitate, thus slowing the overall 
alacrity of the unit and allowing for longer sustained fire. At an estimated 200 rounds per minute with a range of nearly 500 
yards thanks to the longer barrels and heavier platform, we speculate that two firing teams could repel a force of a thousand men 
with relative ease, and given terrain and proper defensive position, possibly as many as four or five times that number. 
Considerations include a lack of mobility and an inability to fire and advance, making the weapons most useful in a defensive 
capacity or as fire support for advancing troops. Additional research is required, pending another opportunity for field testing.”}\par
	\hfill --\textbf{Boye Lagetsu}, Head of Glade Itsu Artillery Design Bureau 
\end{quote}
\begin{quote}
\emph{“Through a complex process involving rock salt, brimstone, purified alcohol and other chemicals, one can create a yellowish 
liquid that, once released from pressure, rapidly evaporates into a gas. It is inexpensive, easy to transport, and because it 
remains in liquid form for a time, could be used more creatively than the gas dispensers previously developed by the Lagu Design 
Bureau. For instance, flooding tunnels, or with the use of bonding agents controlling the evaporation process, ensuring that the 
targets do not detect and flee from the agent before it has time to take effect. The actual effects of this gas have a short onset time.}

\emph{Here we have several silversalt imprints of various stages of the compound’s effect on an exposed subject. 
...I would remind the ethics committee that this weapon would probably never be used against actual people. It would be 
employed primarily against vermin, aggressive animals, and humans.”}\par
--\textbf{Xabi Beloki}, Zorezai Zaibatsu Senior Designer. \textit{‘Brimstone Gas’} Development Grant Request, Glade Lagu Records 
\end{quote}

The Burdinadin’s Natural Philosophy allows them the ability to produce many wondrous things; 
technology so far ahead of its time that most simply assume it to be magic. The closed nature of the Iron 
Glades guarantees that most of these devices never see the light of day, and that’s how the Burdinadin 
like it. They rightly fear what would happen if humans reverse-engineered their equipment.

Burdinadin hardware isn’t normally available to humans. They don’t usually sell it, and they are very 
wary about even letting their trusted human allies use it. Most equipment is designed to destroy itself 
quickly without regular micro-maintenance, the sort of thing that only Burdinadin could perform. This 
way, if the hardware falls into non-Burdinadin hands, it will quickly become unusable. 

\section{Burdinadin Armor}
Burdinadin armor differs from the armors of other races and cultures in that it is not intended merely for 
protection from violence, but also as a proof against spirit contamination. They wear it to insulate 
themselves from the outside world, and to protect their vulnerable flesh from the predations of monsters 
from the Otherworld. 

As such, Burdinadin armor has an ‘Insulation’ factor. As long as the Insulation is above a certain threshold, the Burdinadin is protected from contamination. The Burdinadin’s HLT also factors into this. 
Insulation is removed from armor by Crushing attacks. Each point of Crushing damage removes 1 point of Insulation. 

\begin{table}
	\centering
	\caption{Insulation~$+$~HLT Requirements to avoid Contamination}
	\label{tab:Insulation Requirements}
	\begin{tabu} spread 0pt {XX[-1]}
Environment&Insulation + HLT Required\\\toprule
Sterile (Glade, Iron Tent, Deadzone)&0\\
Average (Outdoors, Old Mine, City)&10\\
High (Forest, Underwater, Temple, Magical Hotspot)&15 \\
Extreme (Schwarzvald, The Spirit Realm)& 20\\
	\end{tabu}
\end{table}
\subsection{Burdinadin Armor}
Words around creating your own armor set by mixing and matching components.

\begin{table}[!ht]
	\caption{Burdinadin Headgear}
	\label{tab:Burdinadin Headgear}
	\begin{tabu} to \linewidth {X[2] *{3}{X[-1,c]} X[3]X[4] *{3}{X[-2,c]}}
Armor &AVC &AVP &AVB &Coverage &Special &Wt. &PP &Cost\\\toprule
Burdinadin Helmet&10&9&8&Upper Head, Lower Head&Hard, Insulation 2, Augmentations (1)&3&-1&3 gp\\
Burdinadin Respirator Mask &10 &9 &8 &Face$\phi$ &Hard, Insulation 2, Augmentations (1)&1&-1&1 gp\\
	\end{tabu}
\end{table}

\begin{longtabu} to \linewidth {X[2] *{3}{X[-2,c]} X[3]X[4] *{2}{X[-2,c]}}
\caption{Burdinadin Plate Armor}
\label{tab: Burdinadin Plate Armor}\\
\rowfont[c]{}Component Name&AVC&AVP&AVB&Coverage&Special&Wt.&Cost\\\toprule
Burdinadin Couters &8 &7 &6 &Elbow$\phi$ &Hard, Insulation 0 &1 &5 sp\\
Burdinadin Cuirass &8 &7 &6 &Belly, Chest, Lower Back, Side, Upper Back& Hard, Insulation 3, Augmentations (3) &2.5& 2 gp \\
Burdinadin Cuisses &8 &7 &6 &Thigh &Hard, Insulation 1 &2 &8 sp \\
Burdinadin Fauld &8 &7 &6 &Groin$\phi$, Hip$\phi$, Thighs$\phi$ &Hard, Insulation 0 &1.5 &10 sp\\
Burdinadin Gauntlets &8 &7 &6 &Hand$\phi$ &Hard, Insulation 0, Augmentations (1) &2 &5 sp \\
Burdinadin Gorget &8 &7 &6 &Neck &Hard, Insulation 0 &1 &10 sp \\
Burdinadin Greaves &8 &7 &6& Shin &Hard, Insulation 0, Augmentations (1) &1.5 &6 sp\\
Burdinadin Poleyns &8 &7 &6& Knee$\phi$ &Hard, Insulation 0 &1.5 &5 sp\\
Burdinadin Rerebraces &8 &7 &6& Upper Arm &Hard, Insulation 1, Augmentations (1) &1 &5 sp\\
Burdinadin Sabatons&8&7&6&Foot& Hard, Insulation 0, Augmentations (1) &1.5 &5 sp\\
Burdinadin Pauldrons&10&9&8&Shoulder, Upper Arm& \nameref{par:Bulletproof}, Hard, Insulation 1, Augmentations (1) &3 &1 gp\\
Burdinadin Vambraces&8&7&6&Forearm& Hard, Insulation 1&1&5 sp\\
\end{longtabu}

\subsection{Myrmidon Armor}
Myrmidon armor is a marvel of modern Burdinadin engineering, and far outclasses any other form of 
armor yet developed.

Myrmidons are heavy mechanized armor suits that are ridden in, rather than worn. The pilot primarily 
controls the armor with their mind, as well as levers and switches that connect directly to their spinal 
cords and skulls.

The armor suits themselves are bipedal monstrosities of steel and machine, often laden with weaponry 
and other equipment. Some are elaborately painted and decorated by their pilots with murals representing 
their deeds.

The Myrmidons are rare. Most Glades have only two or three of them, but each is a war machine capable 
of tearing dozens of armed men to pieces, and challenging all but the largest of monsters at range or in 
hand to hand combat. 

[Myrmidon armor is beyond the scope of this Release] 

\subsection{Armor Augmentations}
Burdinadin armor often have Augmentation slots listed in their special qualities section, which specifies 
the number of Augmentations that the armor can be fitted with. These Augmentations are listed below,
along with the details of their function and their additional costs.

\paragraph{Augmentation Name}---\quad [Number of Augmentation slots required]

Brief description of the Augmentation.
\vspace{-15pt}\subparagraph{Cost:} X gp/sp/cp
\vspace{-15pt}\subparagraph{Fitted To:} Torso/Arms/Legs/Helmet/Mask/Any\\ (Pauldrons count as both Torso and Arms for the purposes of Augmentations)

\paragraph{Illumination}---\quad[0]

A magitek (or sometimes just tech) lighting device mounted to some part of the armor. It illuminates a 
roughly 60 degree arc (one sixth of an arc) in the direction the light is pointed, making it Evenly Lit for 
about 5 yards, and dimly lit for another 10 yards. 
\vspace{-15pt}\subparagraph{Cost:} 10 sp
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph{Quick-Release}---\quad[0]

A system by which the entire armor (or components of it, specify which at time of installation) can be 
\textbf{completely ejected} from the wearer’s person instantly. This can be done as a generic action in combat, 
and takes less than one second. Only the wearer can activate this system. The armor can be placed back 
on without issue. This device is useful for Burdinadin who fear they might need to quickly ditch their 
armor in some situation--say, falling into the ocean, or needing to flee across a bridge with a low weight threshold.
\vspace{-15pt}\subparagraph{Cost:} 8 cp
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph[Enhanced Strength]{Enhanced Strength\footnote{Note: You can mix and match Enhanced Strength and Enhanced Endurance, treating the total number of enhancement levels cumulatively for the purposes of cost.} }---\quad[2+1 per level of STR]

Some Burdinadin suits utilize a system of strength-enhancing hydraulics, springs, and torsion devices to 
enhance the strength of the wearer. These systems are very expensive, and are usually reserved only for 
pioneer units or construction experts. 
\vspace{-15pt}\subparagraph{Cost:} 10 gp every +1 STR up to a maximum of +3 STR 
\vspace{-15pt}\subparagraph{Fitted To:} Torso, Arms

\paragraph[Enhanced Endurance]{Enhanced Endurance\footnotemark[\value{footnote}] }---\quad[2 +1 per level of END] 
Designed to absorb shock, distribute weight to self-supporting elements of the armor, and to generally 
reduce straight on the wearer’s body, this suite of subsystems effectively improves Endurance by making 
the armor much easier to wear. This also prevents a Burdinadin from aggravating injuries through 
exertion--thus making \nameref{sec:Blood Loss} Tests easier. 
\vspace{-15pt}\subparagraph{Cost:} 10 gp for every +1 END up to a maximum of +3 END 
\vspace{-15pt}\subparagraph{Fitted To:} Torso, Legs 

\paragraph{Sensory Suite}---\quad[1 for single, or 2 for full suite]

Various tools, devices, and nifty tricks to enhance a Burdinadin’s use of one or all of the senses. 
\vspace{-15pt}\subparagraph{Cost:} 10 sp for the effects of Good Ears, Good Eyes or Good Nose (effects stack for a character with the Boon), 3 gp for all 3 effects
\vspace{-15pt}\subparagraph{Fitted To:} Helmet, Mask
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph{Hemostatic Pressure System}---\quad[1]

The interior of the suit auto-inflates when damaged significantly, acting as a tourniquet to prevent blood 
loss. Activating the Hemostatic Pressure System reduces your current Bleed from any wounds in the area 
by 20 for the purposes of \nameref{sec:Blood Loss} rolls. The Bleed is still there, and has to be treated separately, but
will not contribute to your immediate \nameref{sec:Blood Loss}.
\subparagraph{Cost:} 5 gp
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph{Hardened}---\quad[1]

Additional armoring and technology to seal up potential gaps in the armor through which attacks might 
penetrate. All Weak Points on this piece of Armor are removed, preventing \nameref{man:Joint Thrust} and \nameref{man:Grappling: Coup de Grace} Maneuvers from functioning against them. %NEEDS REF TO WEAK POINTS
\vspace{-15pt}\subparagraph{Cost:} 1 gp
\vspace{-15pt}\subparagraph{Fitted to:} Any

\paragraph{Ablative Armor}---\quad[1]

Cheap plates designed to take the brunt of an attack by breaking, generally placed over the ‘proper’ armor. Adds 4 extra AV, which treats all attacks as having 2 more levels of Crushing than normal until it runs out. Damage to the Ablative Armor cannot be repaired; the ceramic plates must be replaced entirely. This augmentation can only be taken once for any piece of armor. 
\vspace{-15pt}\subparagraph{Cost:} 1 sp per Hit Location
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph{Integrated Weapon}---\quad[1 per weapon or 2 if melee weapon has greater Reach than Long]

Choose one melee or missile weapon. It is now built into the suit, and can be drawn and utilized as though 
quickly and readily available (for the purposes of \nameref{man:Quickdraw}). The weapon’s Weight is reduced by half if it is not currently being used.
\vspace{-15pt}\subparagraph{Cost:} 1 gp plus the cost of the weapon
\vspace{-15pt}\subparagraph{Fitted To:} Arms, Torso, Legs

\paragraph{Integrated Tool System}---\quad[1 per toolset]

Include surgeon’s tools, smith’s tools or tradesman’s tools in your armor. These have no weight and can 
be drawn for use at any time.
\vspace{-15pt}\subparagraph{Cost:} 1 gp per toolset
\vspace{-15pt}\subparagraph{Fitted To:} Arms

\paragraph{Independent Air Supply}---\quad[2/1]

A tank of oxygen that can be fed through the suit’s mask (or through a mouthpiece if no mask is available) to allow breathing in environments with no or dangerous air. The supply will last half an hour (additional air supplies can be acquired for 1 augmentation slot, adding another 30 minutes.)
\vspace{-15pt}\subparagraph{Cost:} 5 gp
\vspace{-15pt}\subparagraph{Fitted To:} Torso

\paragraph{Dark Vision}---\quad[2]
Using Magestone panels on the exterior of the helmet, radiation can be emitted into the dark that is visible 
through a specially treated visor. This allows the Burdinadin to treat any level of light as though it were 1 
level brighter than normal. The visor is movable and usually takes a form similar to a knightly helmet’s 
visor, or else a pair of goggles mounted on the forehead that can be temporarily removed. In direct 
sunlight, these goggles produce Blinding light, and cannot be used. The light of the sun also contains 
radiation much stronger than that emitted by Magestone. Any Crushing damage to the helmet has a 
significant chance (50\%) of destroying this system.
\vspace{-15pt}\subparagraph{Cost:} 3 gp
\vspace{-15pt}\subparagraph{Fitted To:} Helmet

\paragraph{Mobility Enhancement}---\quad[1]

The suit is adapted for a specific environment, allowing it to traverse more quickly. This could be 
anything from a system of skates, motors and compressed air jets, to skis or climbing apparatus. The 
Augmentation increases the suit’s MOB by X in certain environments, for example, a cave-adapted suit 
will not be able to use its equipment while trying to cross a field. A suit can be adapted for multiple 
environments. This Augmentation can be taken multiple times, each time increasing Mobility by 2 up to a 
maximum of 6.
\vspace{-20pt}\subparagraph{Example Environments:} Snow, Desert, Caves, Water, Plains, Urban, Swamp, Jungle/Forest, Mountains, Hills

\vspace{-15pt}\subparagraph{Cost:} 1 gp
\vspace{-15pt}\subparagraph{Fitted To:} Torso, Legs, Arms

\paragraph{Self-Sealing}---\quad[1]

This collection of additional armor plating and reinforcement systems allows the user to retain the suit’s 
properties of Insulation even if damaged by a Crushing attack. 
\vspace{-15pt}\subparagraph{Cost:} 2 sp
\vspace{-15pt}\subparagraph{Fitted To:} Any

\paragraph{Helian Fire Caster}---\quad[X]

Named not because of its inventor, but due to the intense blue and purple flame that the Fire Caster burns. 
A small, cylindrical weapon is attached to the armor’s left or right vambrace, either as a standalone unit, 
or connected to a tank somewhere on the person. It takes one Action to prepare the weapon to fire, but 
after that initial step, it can then be fired instead of a regular attack. It has a range of L, with no range 
penalties for shooting closer than L, and for each BS you have, you gain 1 Burn damage to a random Hit 
Location on your target. You may instead sweep the Fire Caster, using triple fuel, however you set the 
ground in front of you on fire at a range of M. Anyone that wishes to pass through it must pass an ADR 
test or immediately catch on fire. 
\vspace{-15pt}\subparagraph{Cost:} X gp/sp/cp
\vspace{-15pt}\subparagraph{Fitted To:} Suit/Helmet/Any

\section{Burdinadin Weapons} %BURD WEAPONS
Most Burdinadin weapons are eminently similar to human ones -- if somewhat more advanced or refined in 
design. This can be adequately represented by applying Burdinadin Craftsmanship, as listed below, to %BURDINADIN CRAFTSMANSHIP REF
regular weapons. However, there are certain weapons which are totally unique to the Burdinadin, which will be listed below. 

\subsection{Burdinadin Ukabilla}
The Ukabilla is a steel tube in which a complex system exists to ignore the fuse of a chemical rocket and shoot it out of one side towards the enemy. The Burdinadin developed these weapons to fill a gap in their weapons capabilities between stationary artillery pieces and personal firearms. The Ukabilla can fire a variety of rockets, each designed for a different purpose.
\begin{table}[hb]
	\centering
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Ukabilla Firearm}\vspace{-20pt}
	\label{tab:Ukabilla Firearm}
	\begin{tabu} spread 0pt {XXXXXXXXX}
\rowfont[c]{}Weapon Name &Weapon Type &Range &Missile TN &Damage &Load &Ammunition &Wt &Cost\\\toprule
Ukabilla &Firearm &25 &8 &See Round &50 &See Below &10 &100 sp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Ukabilla Rocket Types}
	\label{tab:Ukabilla Rocket Types}
	\begin{tabu} spread 0pt {XXXX}
\rowfont[c]{}Rocket Type &Damage &Special &Cost\\\toprule
Explosive &15p
	&\hyperref[]{text}Explodes R:10 P:15 T:p/b
	&5 gp \\
Canister Shot &12p
	&Scatter 10/8\newline
		Explodes R:10 P:8 T:p
	&5 gp \\
Armor Piercing &20p &- &5 gp\\
	\end{tabu}
\end{table}

\subsection{Burdinadin Lanzalamas}
The Lanzalamas is not actually that complex of a weapon, it has merely been refined from its earlier 
designs (used by the Helians, mostly) using Burdinadin expertise. The weapon projects a jet of flammable 
liquid at the target, and sets it ablaze after it has left the hose. The momentum of the burning fluid 
prevents it from “following” the stream back to the weapon and potentially causing an explosion, and the 
sophisticated pressure mechanism makes sure that that momentum is maintained until the weapon is out 
of fuel. 

The Lanzalamas is a unique weapon in that it does not inflict damage through force or impact, but rather 
by setting the target on fire. When using a Lanzalamas, do not inflict damage for Bonus Successes. Any 
target hit using the Lanzalamas is instead set on fire. The fire damage is Full Body (in certain situations 
the GM may decide otherwise) and is 5/TN5. The flame burns for 10 rounds, and even water will not 
extinguish it. Sand, mud, and clay can extinguish it. 

A Lanzalamas cannot be reloaded normally--the tank of fuel that accompanies must be refilled using 
special tools. It can be fired 10 separate times (Multi-Shots and similar things only count as a single 
firing) before it must be refilled. 

Taking Cover does not reduce the effect of this attack in any way, unless the cover is some sort of solid 
barrier with no permeation that would actually halt a jet of burning fuel. 

\begin{table}[!hb]
	\centering
	\captionsetup{textformat=empty, labelformat=blank}
	\caption{Lanzalamas Firearm}\vspace{-20pt}
	\label{tab:Lanzalamas Firearm}
	\begin{tabu} spread 0pt {*{9}{X[-1]}}
		\rowfont[c]{}Weapon Name &Weapon Type &Range &Missile TN &Damage &Load &Ammunition &Wt &Cost\\\toprule
Lanzalamas &Firearm &10 &5 &See Description &- &See Description &15 &300 sp\\
	\end{tabu}
\end{table}

\subsection{Burdinadin Weapon Enhancements}
\paragraph{REQUIRED: Burdinadin Craftsmanship}---\quad [+150\%]

The Burdinadin’s Natural Philosophy ability allows them to inspect devices and materials on a level far deeper than anyone could do with their regular senses. This lets them find weaknesses in steel, or 
inadequacies in construction that would never be detected in a regular weapon until far too late. 
Burdinadin craftsmen use this ability to find and remove these weaknesses during construction. 
As a result, Burdinadin hardware is as close to flawless as it is possible to be. Burdinadin weapons have 
+9 Durability and Burdinadin firearms never break on a misfire. For swords and other melee weapons, 
this can be combined with Fine Forging.
\vspace{-15pt}\subparagraph{Cost:} Weapon Cost$~\times~1.5$
\vspace{-15pt}\subparagraph{Fitted To:} Any

Burdinadin Craftsmanship is required before any other Burdinadin weapon enhancements can be added.

\paragraph{Electrified}---\quad[+500\%] \quad (Melee only)

An Electrified weapon is powered using a small copper turbine powered with a pull-string. Once 
activated, which can be done at the same as drawing or Quickdrawing a weapon, the weapon is Electrified 
for about ten seconds, or 5 Rounds. Any Maneuver that brings this weapon either in contact with the 
enemy (such as Strike or Thrust), or in contact with a metal weapon that the enemy is wielding (such as Parry), inflicts +4 Stun, regardless of whether or not the attack inflicts damage. This does not work against the Undead. 

\vspace{-15pt}\subparagraph{Cost:} Weapon Cost$~\times~5$
\vspace{-15pt}\subparagraph{Fitted To:} Melee Weapon

\paragraph{Magestone Sight}---\quad[10 gp] \quad (Ranged Weapon only)

A small lens or tube, marked with glowing crosshairs that are painted using a Magestone derivative. 
Mounted on the weapon, Magestone Sight provides a very visible and useful tool for aiming, even in low-light conditions. When making a shot that benefits from the Aim Maneuver, you gain an additional +2 bonus to the attack. Alternatively, this Enhancement is a piece of Magestone under a collimating lens and half-silvered mirror, making the Magestone appear to be, effectively, a dot on the horizon. The user may operate the gun as though they had already performed an Aim Maneuver. 
Cost: 10 gp
Fitted To: Ranged Weapon 
 
\chapter{Other Items}\label{ch:Other Items} %OTHER ITEMS NEED DOING
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage

I feel there needs to be an introduction of sorts here. Why are we buying other items?

\section{Everyday Items}
\begin{table}[hb]
	\centering
	\caption{Animals}
	\label{tab:Animals}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item& Notes &Cost\\\toprule
Chicken &Annoying &5 c\\
Courser (battle horse) &Fast &60 sp\\
Cow &Mercifully ignorant of the ways of the world &10 sp\\
Donkey &Plotting something &5 sp\\
Friendly doge &Much fur &1 cp\\
Goat &Will eat anything &8 cp\\
Destrier (heavy warhorse) &Thirsts for blood &100 sp\\
Hunting falcon &Ruthless towards rabbits &20 sp\\
Ox &Stubborn &12 sp\\
Palfrey (riding horse) &Noble &40 sp\\
Pig &Disgusting &3 sp\\
Sheep &Fluffy &6 sp\\
Tiger &Fearfully symmetrical, not at all tame &15 gp\\
Workhorse &Feathering optional &15 sp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Clothing}
	\label{tab:Clothing}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Notes &Cost\\\toprule
Work clothes &Full outfit &1 cp\\
Traveling cloak &Single item &1 cp\\
Urban attire &Full outfit &1 sp\\
Wealthy garment &Full outfit &1 gp\\
Noble dress &Full outfit &5 gp\\
Court attire &Full outfit &10 gp\\
Finery fit for a king &Full outfit &20-50 gp\\
Fine horse caparison &Single item &10 gp\\
Obscenely decadent horse ornamentation &Single item &20-50 gp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Drink}
	\label{tab:Drink}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Ale/Beer &Flagon &1 cp per 10\\
Ale/Beer &Gallon &1-2 cp\\
Brandy &Flask &2 cp\\
Cider &Gallon &1 cp\\
Mead &Gallon &2 cp\\
Rakia &Flask &1 cp\\
Waterskin &Skin (1 Gallon) &1 cp (water not included)\\
Wine &Flagon &1 cp per 8\\
Wine &Gallon &2-4 cp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Food}
	\label{tab:Food}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Traveling bread &Brick &1 cp/10\\
Black bread &Loaf &1 cp\\
Brown bread &Loaf &5 cp\\
Fine white bread &Loaf &10 cp\\
Cheap meat (beef, pork, chicken) &Pound &3 cp\\
Cheese &Pound &1 cp\\
Common fish &Pound &1 cp\\
Expensive meat (swan, dormouse, giraffe, bork) &Pound &1 sp\\
Fruit &Pound &1 cp\\
Lavish banquet &Per 10 people &5 sp\\
Meal (poor) &Per 1 person &1 cp\\
Meal (good) &Per 1 person &4 cp\\
Meal (fine) &Per 1 person &1 sp\\
Meal (exquisite) &Per 1 person &1 gp\\
Vegetables &Pound &1 cp\\
Rare fish &Pound &6 cp\\
Week's rations &Pound &4 cp\\
Cheap grain (kesh, sorghum, rye) &Pound &1 cp\\
Regular grain (oats, barley, hops) &Pound &3 cp\\
Fine grain (wheat) &Pound &5 cp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Torches and Illumination}
	\label{tab:Torches and Illumination}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Candle (low quality) &5 items &2 cp\\
Candle (high quality) &5 items &7 cp\\
Torch &10 items &3 cp\\
Lamp &Per item &5 cp\\
Candle lantern &Per item &9 cp\\
Oil lantern &Per item &1 sp\\
Oil &Pint &1 sp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Medical and Surgical Items}
	\label{tab:Medical and Surgical Items}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Inclusions &Cost\\\toprule
Surgeon’s tools &Needles, blades, splints, tongs, pliers, tweezers, mirrors, prongs, clamps &15 sp\\
Surgeon’s kit (5 uses) &Bandages, plasters, ointments, oils, select herbs. Enough to treat ~5 serious injuries &5 sp\\
Surgeon’s kit (1 use) &Bandages, plasters, ointments, oils, select herbs. Enough to treat 1 serious injury &6 cp\\
Poultice (single use) &Made from honey, bread, mercury, and the gods know what else. Cannot be stored for long! &1 gp\\
Poultice (ingredients) &All ingredients required to make a Poultice (See Skill [Insert Skill]) &10 sp\\
Prescription remedy &One month’s supply for illness &1 cp to 1 gp\\
Hired surgeon (cheap) &1 operation (Skill 3) &1 sp\\
Hired surgeon (regular) &1 operation (Skill 6) &1 gp\\
Hired surgeon (excellent) &1 operation (Skill 9) &10 gp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Miscellanea}
	\label{tab:Miscellanea}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Anvil &One, heavy &5 sp\\
Farm implement (hoe, shovel, etc) &One, handy &1 sp\\
Flints (flintlock) &10, also good for lighting fires &1 cp\\
Gunpowder &Pound, good for about 50 shots from an arquebus &5 cp\\
Light Hand Grenade &One, throwable &1 sp\\
Heavy Hand Grenade &One, flingable &5 sp\\
Incindiary Bomb &One, careful now &4 sp\\
Light Landmine &One, watch your feet &1 gp\\
Heavy Landmine &One, do not drop &1 gp, 10 sp\\
Instrument (lute, drum, flute, horn) &One, pleasant &1 sp to 10 gp\\
Leather (cowskin) &One, tanned &1 sp\\
Lanyard for Weapon &One, in color of choice &1 cp\\
Matchcord &50 feet, you’ll go through it quick &1 cp\\
Musketballs/shot &10 balls/10 loads of shot &1 cp\\
Paper &One sheet, useful &4 cp\\
Parchment &One sheet, fancy &8 cp\\
Rope &50 feet, useful &2 cp\\
Smith’s tools &Set, anvil not included &4 sp\\
Tinderbox &One, rugged &3 cp\\
Tradesman’s tools &Set, guild approved! &10 sp\\
Whetstone &One, dry &1 cp\\
Woodsman’s axe &One, trusty &1 sp\\
	\end{tabu}
\end{table}
\clearpage
\section{Services}
\begin{table}[hb]
	\centering
	\caption{Services}
	\label{tab:Services}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Duration &Cost\\\toprule
Guide (frontier) &1 day &6 cp\\
Mercenary (poor) &1 week &1 sp\\
Mercenary (mediocre) &1 week &3 sp\\
Mercenary (good) &1 week &6 sp\\
Mercenary (excellent) &1 week &1 gp\\
Servant (domestic) &1 year &1 gp (including board)\\
Ship (cog) &1 month &10 gp\\
Ship (caravel) &1 month &25 gp\\
Inn (poor) &1 night &1 cp/Person\\
Inn (decent) &1 night &4 cp/Person\\
Inn (fine) &1 night &1 sp/Person\\
Stabling (summer/temperate) &1 night &1 cp/horse\\
Stabling (winter/cold) &1 night &3 cp/horse\\
	\end{tabu}
\end{table}


\section{Trade Goods}
\begin{table}[hb]
	\centering
	\caption{Textiles}
	\label{tab:Textiles}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Canvas &Yard &2 cp\\
Linen &Yard &6 cp\\
Kersey &Yard &3 cp\\
Broadcloth &Yard &20 cp\\
Broadcloth (dyed) &Yard &30-40 cp\\
Broadcloth (scarlet) &Yard &150 cp\\
Silk (plain) &Yard &100 cp\\
Velvet (plain) &Yard &180 cp\\
Velvet (exquisite) &Yard &280 cp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Spices}
	\label{tab:Spices}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Cyrv &Pound &1 cp\\
Pepper &Pound &18 cp\\
Ginger &Pound &12 cp\\
Cinnamon &Pound &24 cp\\
Cloves &Pound &35 cp\\
Salt &Pound &12 cp\\
Saffron &Pound &182 cp\\
Sugar &Pound &16 cp\\
Karspice &Pound &150 cp\\
	\end{tabu}
\end{table}

\begin{table}[hb]
	\centering
	\caption{Metals}
	\label{tab:Metals}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Copper &Ingot: 7x3x2 inches, 11 lbs &Ingot: 33 cp, Pound: 3 cp\\
Tin &Ingot: 7x3x2 inches, 9 lbs &Ingot: 54 cp, Pound: 6 cp\\
Iron &Ingot: 7x3x2 inches, 10 lbs &Ingot: 30 cp, Pound: 3 cp\\
Silver &Ingot: 7x3x2 inches, 14 lbs &Ingot: 14 gp, Pound: 1 gp\\
Gold &Ingot: 7x3x2 inches, 27 lbs &Ingot: 135 gp, Pound: 5 gp\\
Bronze &Ingot: 7x3x2 inches, 11 lbs &Ingot: 55 cp, Pound: 5 cp\\
Steel &Ingot: 7x3x2 inches, 10 lbs &Ingot: 50 cp, Pound: 5 cp\\
Magestone &Not available in ingots, stones are of varying sizes &Stone: 50 gp\\
Orichalcum &Ingot: 7x3x2 inches, 11 lbs &Ingot: 5500 gp, Pound: 500 gp\\
	\end{tabu}
\end{table}


\begin{table}[hb]
	\centering
	\caption{Goods}
	\label{tab:Goods}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Unit &Cost\\\toprule
Cotton &Ton &30 gp\\
Coffee &Pound &1 sp\\
Coffee (bulk) &Ton &100 gp\\
Cocoa &Pound &10 cp\\
Cocoa (bulk) &Ton &80 gp\\
Fine Rug &Throw &5 gp\\
Common tea &Pound &1 sp\\
Common tea (bulk) &Ton &100 gp\\
Fine tea &Pound &1 gp\\
Fine tea (bulk) &Ton &2000 gp\\
	\end{tabu}
\end{table}
\clearpage
\section{Slaves}
\begin{table}[hb]
	\centering
	\caption{Slaves}
	\label{tab:Slaves}
	\begin{tabu} spread 0pt {*{3}{X[-1]}}
		Item &Notes &Cost\\\toprule
Labor slave &Oh, anyone with all their limbs will do. &1 gp\\
House slave &A house slave must be made of sharper, gentler stuff than a worker. &3 gp\\
Educated slave &A ‘marsupial attendant,’ always carrying around his notes in a pouch on his apron. &10 gp\\
Pit fighter/gladiator &Why is it easier to train a vicious man than a vicious dog? &20 gp\\
	\\
Exotic (feature) &Strange, but alluring features and language. &Cost $\times\ 2 $\\
Inferior race (feature) &Goblins, ex-bandits. &$ \dfrac{\text{Cost}}{2} $\\
Exotic race (feature) &Non-Human race (Burdinadin, Zell, etc). &Cost $ \times\ 3$\\
Especially loyal (feature) &Conditioned to be loyal to the death. &Cost $ \times\ 1.5 $\\
Especially beautiful (feature) &Beauty isn’t just about features, it’s about habit, too. &Cost $ \times\ 2 $\\
Especially intelligent (feature) &Any idiot can learn to scribble on a tablet. True intelligence is rare. &Cost $ \times\ 2 $\\
Especially strong (feature) &Ever move a piano up a staircase? Don’t worry, you’ll never have to. &Cost $ \times\ 2.5 $\\
	\end{tabu}
\end{table}
\part{The Riddle of Steel}
\chapter{Proficiencies}\label{ch:proficiencies} 
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quote}
	\centering
\emph{“When I was in Illegon, I had an apprentice. I forget his name. A hundred times I showed him the right way to parry, and a hundred times he did it the wrong way. On the hundred and first time, I didn’t hold back. Only in swordsmanship can a man lose an ear, but become a better listener.”}\\
\hfill -\textbf{Sarah Gizka}, \textit{On Swords}
\end{quote}

The Proficiency system allows characters to track their combat capabilities. An individual Proficiency 
determines skill in a broad classification of weapon (for example, 1H Sword, a 2H Sword, and so on), not simply how effective a character is at wielding certain instruments (such as a rapier, broadsword or 
mace). Proficiency levels directly affect a character’s CP, allowing them to perform more deadly 
maneuvers, and also providing characters with Talents (\autoref{ch:Talents}), which offer special bonuses to certain actions in combat.

Characters can take and level Proficiencies at Character Creation or during regular play.

\section{Character Creation}
When purchasing Proficiencies with PCP, you receive a number of Proficiency Points as indicated in \autoref{tab:PCPInvestment}~in \autoref{ch:charcreation}, and may then use those points to purchase and advance one or more Proficiencies.

Characters unlock Maneuver Tiers and Talents as they level their Proficiency. Maneuvers are certain 
techniques and moves that a character employs in combat (and are further detailed in \autoref{ch:Maneuvers}), and 
Talents are small bonuses which may be universal, or specific to a Proficiency (refer to Chapter XX). 

\autoref{tab:Proficiency Progression} below indicates how many Proficiency Points are required to gain different Proficiency levels. 

As an example, a character whose Proficiency level is ‘2’ will have unlocked Core Maneuvers for their 
chosen Proficiency and have access to a Novice Talent, and a character with a Proficiency level 10 can 
use Tier 2 Maneuvers and has access to up to three Journeyman Talents and two Novice Talents. A 
character has access to every Maneuver in his Tier, as well as in the Tiers below. 

\begin{longtabu} spread 0pt {X[-2,c] X[c] X[c] X[-1,c]}
	\caption{Proficiency Progression}
	\label{tab:Proficiency Progression}\\
\rowfont[c]{}Proficiency Level&Maneuver Tiers Unlocked&Talents Unlocked (Not available to Missile Proficiencies)&Proficiency Point / Arc Point Cost\\\toprule
1&	Core Maneuvers&	-&	1\\
2&	-& 	\hyperref[sec:Novice Talents]{Novice Talent}&	1\\
3&	-& 	-&	1\\
4&	-& 	\hyperref[sec:Novice Talents]{Novice Talent}&1\\
5&	Tier 1 Maneuvers&	-&	2 \\
6&	-&  	\hyperref[sec:Journeyman Talents]{Journeyman Talent}&2 \\
7&	-&  	-  	&2 \\
8&	-&  	\hyperref[sec:Journeyman Talents]{Journeyman Talent}		&2 \\
9&	-&  	-  	&3 \\
10&	Tier 2 Maneuvers&	\hyperref[sec:Journeyman Talents]{Journeyman Talent}		&3 \\
11&	-&  	-  	&3 \\
12&	-&  	\hyperref[sec:Master Talents]{Master Talent}  	&3 \\
13&	-&  	-  	&3 \\
14&	-&  	\hyperref[sec:Master Talents]{Master Talent}  	&6 \\
15&	\hyperref[sec:Master Talents]{Master Talent}&	-&	6 \\
16&	-&-&6 \\
17&	-&-&6 \\
18&	-&-&8 \\
19&	-&-&10 \\
20&	???&  	???  	&10\\
\end{longtabu}

\subsection{Advancing with Arc}

During play, your character can advance their Proficiencies and gain new Maneuvers and Talents by spending Arc Points (refer to \autoref{ch:arc} for more information on Arc).

When you take a new Proficiency that you didn’t purchase at Character Creation, simply pay the level 1 Arc Point cost (detailed in \autoref{tab:Proficiency Progression}) and then note it on your Character Sheet.

To level a Proficiency from level 1 to level 5, you must purchase levels 2, 3, 4 and 5. Proficiency levels 
must be bought in order (you cannot buy only level 5 if you are currently level 1, but need to first buy 
levels 2, 3 and 4). Once you have spent the required Arc Points to level your Proficiency, note it on your 
Character Sheet. 

\section{Using Proficiencies}
When in combat, a character determines their CP by combining their Proficiency level with their ADR. 
The question then is, which Proficiency should the character use? 

A character must use a Proficiency of the type of weapon that they are holding.  

If a character is holding a sword in their dominant hand, then they must use the 1H Sword Proficiency. If 
the sword is double-handed, then the Proficiency being used will almost always be 2H Sword. If the 
character is wielding a spear, then they must use the Spear Proficiency. There are a few exceptions to this, 
such as the Hand-Off rule for some 2H weapons that allows them to be wielded one-handed while still 
using the 2H Proficiency of their type. 
\begin{itemize}
	\item When Dual Wielding, a character must use the Proficiency for the weapon held in their dominant hand. 
	\item If Ambidextrous and Dual Wielding, the character must decide which hand is being treated as dominant at the beginning of each Round.
	\item When unarmed, a character must either use Pugilism or Wrestling as their Proficiency. 
	\item If a character is forced to use a different Proficiency, either by Grappling, Disarming, or other circumstances, they immediately gain or lose CP equal to the difference between their current Proficiency and the one being switched to. If they’re the same level, then nothing is lost. This can result in a character being forced to use a Default from their current Proficiency, often to disastrous results. A character with CP reduced to 0 in this way is not subject to Overshock.
	\item If a character is eligible to use multiple Proficiencies (i.e. while wielding a 2H Sword in one hand using Hand-Off) they may not change between them except at the beginning of the Round.
	\item If a character draws a new weapon mid-action, say with Quickdraw, the new Proficiency is immediately applied.
\end{itemize}

\subsection{Alternate Weapon Usage}
Many weapons can be used in ways that differ from their primary usage. This includes wielding two 
weapons at once, one in each hand, and wielding certain two-handed swords in one hand. These have very 
simple rules for ease of use. 

\subsubsection{Proficiency Defaults}
If a character ends up wielding a weapon they are not familiar with, and has no appropriate Proficiency to 
use it with, they can instead ‘Default’ to one of their other Proficiencies at a slight penalty. When 
Defaulting, the character effectively gains a temporary Proficiency, which is a number of levels lower 
than the Proficiency being Defaulted from as illustrated in Table XX. When Defaulting, a character does 
not enjoys the benefits of any School bonuses. They can only use the CP and Maneuvers appropriate to 
the level of the Proficiency they are Defaulting to. 

\begin{quote}
	\emph{Bale has been disarmed of his sword (1H Sword), and must defend himself with only a hatchet (1H Blunt). He was never trained in the weapon’s use, but he has an alright idea of how it works, and only suffers a -1 penalty with it (using a 1H Blunt weapon when Defaulting from 1H Sword). His 1H Sword level is 5, which allows him access to Tier 1 Maneuvers, but with the -1, his level in 1H Blunt is only 4, meaning he can only use Core Maneuvers. Bad luck Bale! }
\end{quote}

\begin{table}
	\centering
	\caption{Default Matrix}
	\label{tab:Default Matrix}
	\begin{tabu} spread 0pt {*{10}{X[-1,c]}}
&Grappling &Pugilism &Dagger &1H Sword &1H Blunt &2H Blunt &2H Sword &Spear &Polearm\\\toprule
Grappling &0 &-1 &-1 &-3 &-3 &-4 &-4 &-5 &-5\\
Pugilism &-1 &0 &-1 &-3 &-3 &-4 &-4 &-5 &-5\\
Dagger &-1 &-1 &0 &-1 &-1 &-3 &-3 &-4 &-4\\
1H Sword &-3 &-3 &-1 &0 &-1 &-3 &-2 &-4 &-4\\
1H Blunt &-3 &-3 &-1 &-1 &0 &-1 &-3 &-3 -&3\\
2H Blunt &-4 &-4 &-3 &-3 &-1 &0 &-2 &-3 &-3\\
2H Sword &-4 &-4 &-3 &-2 &-3 &-2 &0 &-3 &-3\\
Spear &-5 &-5 &-4 &-4 &-3 &-3 &-3 &0 &-1\\
Polearm &-5 &-5 &-4 &-4 &-3 &-3 &-3 &-1 &0\\
	\end{tabu}
\end{table}

\subsubsection{Dual Wielding}
Dual Wielding involves using a weapon in each hand. This has come up many times throughout history 
and legend, and is often associated with exceptional skill or prodigious talent. In reality, two-weapon 
fighting was not merely the province of legendary swordsmen like Miyamoto Musashi, or Pal Kiniszi. 
Though fighting with two swords was very unusual, fighting with a sword and a smaller companion 
weapon, like a dagger, was not at all unusual at many points in history. References to double-rapier styles 
(the infamous ‘Case of Rapiers’) abound, though it was never taken as seriously as the more formal styles 
of dueling.

The chief advantage of using two weapons instead of a single weapon, or a weapon and a shield, is that 
the secondary weapon can be used to parry, trap, or bind the opponent’s weapon, or make a counter-attack 
on its own. The disadvantages range from not having a shield to the awkwardness of using a large weapon 
in one’s off-hand, or the inability to quickly grasp something with the free hand. 
\begin{itemize}
	\item When Dual Wielding, always use the Proficiency of the weapon in your primary hand (if you are 
Ambidextrous, you may choose which Proficiency you use) for terms of CP and available 
Maneuvers. 
	\item Dual Wielding with a dagger or other weapon with Short Reach or less in the off-hand and a 
another weapon in the primary hand incurs no penalty. 
	\item Dual Wielding with two larger weapons in each hand incurs a -2 penalty to CP, unless the 
character is Ambidextrous, in which case no penalty is suffered. 
\end{itemize}

\subsubsection{Hand-Off Weapons}
Certain weapons, while designed primarily to be used with two hands, could be wielded in one when the 
situation demanded it, most often as cavalry swords while mounted, where one hand is needed to control 
the reins. 

To represent this, certain 2H weapons (generally swords) have an ‘Hand-Off’ rule, and a second profile 
that is 1H instead of 2H. Use the stats appropriate to the number of hands being used to hold the weapon. 

Switching from the 2H grip of a sword to a 1H grip has no cost, and can be done at the beginning of any 
Action. Switching back, assuming the other hand is still free (and attached to the body), can be done the 
same way. 

\section{Schools}
A School in Song of Swords is a group of Proficiencies that are all leveled up at once for a discounted 
cost. Schools cost more Arc Points for a character to take, but they’re much cheaper to level up, thus 
allowing a character to have multiple high-level Proficiencies at a lower cost. 

There are several different kinds of Schools, each with a different structure and unique benefits, as well as 
their own costs for buying in (either at Character Creation or afterwards) and leveling. 

\paragraph{Becoming part of a School}---\quad
A character can be part of a School at Character Creation, or can form a School during gameplay, as long 
as they have access to the proper facilities or training. (Self-Taught schools require no teachers or 
infrastructure). 

Buying into a School in order to gain its benefits is easy. The base cost in Arc of a School is listed in 
brackets in its Cost section. After this is paid, a number of Proficiencies can be selected and purchased at 
level 1 as part of the School. Each Proficiency added to a School must be paid for normally, though 
Humans, because of their racial characteristic Willing to Learn, purchase Proficiencies for 1 point instead 
of 2. Some Schools have an additional cost per Proficiency that must be paid to add one to the School. In 
the case of ‘Primary’ and ‘Secondary’ Proficiencies in Schools, these sometimes have different costs. 

Once Proficiencies are chosen and paid for, the School is complete. All Proficiencies within the School 
are at level 1. Some Schools have bonuses or penalties for Primary or Secondary Proficiencies, listed in 
brackets (i.e. 1 Primary [0]) that alter the CP of the Proficiencies in question. 

\begin{quote}
	\emph{Tosca is part of a Self-Taught School, and he has the School at level 1. His Primary Proficiency is 1H 
Sword. His Secondary Proficiency is Wrestling.
Self-Taught Schools have Primary at 0, so his 1H Sword Proficiency’s CP is 1 (1+0). Self-Taught Schools 
have Secondary at -1, so his Wrestling Proficiency’s CP is actually 0 (1-1), meaning that while he has the 
School and access to Core Maneuvers, he does not get any CP from the Proficiency when he fights with 
it.}
\end{quote}
\paragraph{Leveling a School}---\quad
Once you’ve bought into a School, it levels just like any other proficiency, using the same number of 
points as individual proficiencies. The ‘level’ of the School is tracked independently. The individual 
proficiencies in the School are listed as ‘-1,’ ‘0,’ or ‘+1,’ this being their level relative to the School’s 
level, and representing the different focus that the School places on different weapons. 

\begin{quote}
	\emph{Gallant is part of a Self-Taught School that has 2H Sword at 0, Dagger at -1 and 1H Sword at 1. His School Level is 8, which means that he effectively has 2H Sword at 8 (8+0), Dagger at 7 (8-1) and 1H Sword at 9 (8+1)}
\end{quote}

\section{School Types}

\subsection{Self-Taught}
You taught yourself--or learned through hard knocks--that it’s better to have a few Jacks up your sleeve 
than one Ace. Self-Taught Schools tend to contain few weapons, and practitioners tend to develop bad 
habits, but at the same time they are also uniquely well-suited for picking up dirty tricks and thinking 
outside of the box. 

\vspace{-15pt}\subparagraph{Entry Requirements:} Nothing. Who would you pay, yourself?
\vspace{-15pt}\subparagraph{Who Uses it:} Criminals, peasants, militia without formal training, and bored eccentric nobles. What Weapons: Self-Taught Schools can involve any weapons, but tend to use swords and polearms used afoot, pugilism, wrestling, and perhaps a missile proficiency.
\vspace{-15pt}\subparagraph{Proficiencies:} 1 Primary [0], 0-2 Secondary [-1]
\vspace{-15pt}\subparagraph{Bonuses:} Gain Special Move for either Weapon Throw, Blind Toss, or Quickdraw (Special Move means that one time per opponent, you may use a Maneuver at -1 TN and no Activation Cost. You may never use your Special Move against an opponent after you’ve used it once--they’re onto you). 
\vspace{-15pt}\subparagraph{Penalties:} When you reach a new Tier of Maneuvers (including Core Maneuvers) with any Proficiency within this School, you must pick two non-Universal Maneuvers that you can perform in or before that Tier. You suffer from Poor Form when using those Maneuvers with that Proficiency (Poor Form increases the Activation Cost of the Maneuver when used in combat by 1. An instance of Poor Form can be bought off for 1 Arc Point so that it no longer applies to that Maneuver).
\vspace{-15pt}\subparagraph{Cost:} [0]+1 per Proficiency at adoption.

\subsection{Noble}
Noble Schools exist to teach wealthy aristocrats and other interested, affluent parties how to conduct 
themselves in battle. Noble schools are expensive, and this means they can afford the best tutors to school 
their charges in how to properly murder their fellow men. The primary advantage of Noble Schools is the 
exquisite form and speed that comes from constant repetition, constant correction, and no shortage of 
training equipment. These assets come together to produce excellently practiced fighters with a great 
technical understanding of combat. 

\vspace{-15pt}\subparagraph{Entry Requirements:} 5 gp upon entry. Additionally, each Tier costs 5 gp to ‘unlock,’ as a member of a Noble School must pay exorbitant dues to his school in order to gain access to their secrets (this also serves to keep poor people out on the street, where they belong).
\vspace{-15pt}\subparagraph{Who Uses it:} Young aristocrats, the sons of very wealthy merchants, prodigious peasants allowed access because of natural talent.
\vspace{-15pt}\subparagraph{What Weapons:} Noble Schools tend to include civilized upper-class weapons, like long swords, sabers, lances, rapiers, and perhaps some missile weapons like bows or crossbows. Pugilism and wrestling are generally beneath the dignity of these men, but sometimes are included if dueling in armor is in vogue.
\vspace{-15pt}\subparagraph{Proficiencies:} 1-2 Primary [+2], 1-5 Secondary [+0] 
\vspace{-15pt}\subparagraph{Bonuses:} You automatically gain a Novice Talent for your Primary Proficiency at School level 1, a Journeyman Talent at level 5, and a Master Talent at level 10. You may also choose whether the Noble School you attended teaches Armored, Unarmored, or Missile fighting:
	\begin{description}
		\item [Armored:] Your END counts as being 2 higher for the purposes of Carry.
		\item [Unarmored:] For each Primary Proficiency you have in the Noble School, you may choose an individual weapon or shield that is compatible with that Proficiency. When using that weapon or shield with that Proficiency, you reduce the Parry/Block TN of the weapon/shield by 1.
		\item [Missile:] When taking this School, Choose either Rapid Shot or Power Draw. You gain +2 MP whenever performing that Maneuver.
	\end{description}
\vspace{-15pt}\subparagraph{Penalties:} None. You’re just superior. 
\vspace{-15pt}\subparagraph{Cost:}[2]+2 per Primary, +1 per Secondary

\subsection{Soldier}
Soldier Schools emphasize quantity and practicality. These schools are intended to instruct large numbers 
of men on how to use a large variety of weapons, and to confront whatever situation they may be forced 
into on the battlefield. The methods used are often rough, with less focus on sparring and more on the use 
of pells, training posts and courses to build strength and skill. 

\vspace{-15pt}\subparagraph{Entry Requirements:} 10 sp upon entry, or 5 sp if the character served in an army, town guard, or militia. 
\vspace{-15pt}\subparagraph{Who Uses it:} Soldiers, mercenaries, civilian militia with training.
\vspace{-15pt}\subparagraph{What Weapons:} Soldier Schools can include a broad array of weapons, including swords, daggers, spears, maces, axes, bows, crossbows, javelins and slings, and as well as these, almost certainly including wrestling in their regimen. 
\vspace{-15pt}\subparagraph{Proficiencies:} 3-9 Primary [+0]
\vspace{-15pt}\subparagraph{Bonuses:} Tier 1 Maneuvers and Tier 2 Maneuvers cost 1 less for this School. 
\vspace{-15pt}\subparagraph{Penalties:} Mastery Maneuvers Tier costs 1 more for this School. 
\vspace{-15pt}\subparagraph{Cost:}[1]+1 per Proficiency at adoption.

\subsection{Officer}
Officer Schools can be considered as more refined than a Soldier School, but lacking the finesse and 
prestige of a Noble School. Officers may not necessarily be of noble birth, but they have standards of 
dignity and decorum to uphold, while also expecting to get their hands dirty from time to time. Officer 
Schools teach the weapons of a gentleman as well as the weapons of war. 

\vspace{-15pt}\subparagraph{Entry Requirements:} 1 gp upon entry. Additionally, each Tier costs 1 gp to ‘unlock.’ The officer’s corps will generally try to dissuade the unmonied classes from advancing anywhere, but particularly in martial skill. 
\vspace{-15pt}\subparagraph{Who Uses it:} Officers, the sons of officers, martially-minded aristocrats, the sons of martially-minded aristocrats, janissaries. 
\vspace{-15pt}\subparagraph{What Weapons:} Officer Schools teach a mix of weapons of both martial and civilian bent, including missile and melee weapons. Many officers will expect to do their fighting mounted even if they command 
infantry units, and schools that cater to cavalry officers may focus on weapons like sabers, lances, 
horsebows, carbines, and so on. 
\vspace{-15pt}\subparagraph{Proficiencies:} 1-3 Primary [+1], 1-6 Secondary [+0] 
\vspace{-15pt}\subparagraph{Bonuses:} You automatically gain a Novice Talent for your Primary Proficiency at School level 1, a 
Journeyman Talent at level 5, and a Master Talent at level 10. 
\vspace{-15pt}\subparagraph{Penalties:} None.
\vspace{-15pt}\subparagraph{Cost:} [1]+2 per Primary at adoption, +1 per Secondary 

\subsection{Esoteric}
Esoteric Schools are those in which the chief aim is some intellectual or spiritual goal, where martial 
prowess and skill at arms is considered a contributor. Martial skill as a pathway to enlightenment, or some 
deeper understanding of the self, was not a new concept when the first words clawed their way from a 
human throat. The existential dread brought on by violence and the fear of death have been motivators for 
introspection since the first spear was knapped from flint. 

Esoteric Schools tend to be minimalistic with how they deal with actual violence, even brutish. The 
fighter himself must vanish into the technique, because it is the act of performing the technique perfectly 
that is relevant--not the fighter, not the enemy, not even the fight. A certain fearlessness, born of fatalism 
and discipline, is born of this sort of style, and it creates unnerving and skilled warriors who are difficult 
to read, and even more difficult to counter. Archery traditions in Esoteric Schools often involved 
controlled breathing and a holistic approach to aiming rather than a technical one, and are frequently 
practiced in conjunction with shouts or chants. 
\vspace{-15pt}\subparagraph{Entry Requirements:} 1/10th of starting Wealth at Character Creation. Esoteric Schools cannot be entered during regular play without the GM’s approval.
\vspace{-15pt}\subparagraph{Who Uses It:} Mystics, sword-sages, wandering duelists, Illuminati, and certain religious men. 
\vspace{-15pt}\subparagraph{What Weapons:} Esoteric Schools tend to focus on civilian weapons, simply because warfare is not 
usually their element. Swords are most common, but some may practice the quarterstaff, truncheon, spear, or even unarmed combat and wrestling if the culture is right. Bows and sometimes throwing weapons are 
common in Esoteric Schools, but it is EXTREMELY rare for to see crossbows or firearms included (+2 
cost when adding to School) 
\vspace{-15pt}\subparagraph{Proficiencies:} 1-2 Primary [+1], 0-3 Secondary [+0] 
\vspace{-15pt}\subparagraph{Bonuses:} Any Advanced offensive or defensive maneuver made against you has a +2 Activation Cost. 
Missile Attacks made while benefiting from Aim may adjust the Hit Location by 1 automatically. 
\vspace{-15pt}\subparagraph{Penalties:} All Advanced offensive or defensive maneuvers you perform have +2 Activation Cost. 
\vspace{-15pt}\subparagraph{Cost:} [2]+2 per Primary, +1 per Secondary 

\section*{Proficiency List}
These are the Proficiencies available to be taken either individually or as part of Schools. Each weapon 
may be able to be used with more than one Proficiency (e.g. certain pole weapons can be used with both 
the spear and polearm Proficiencies), so be sure to check which Proficiencies your character’s favored weapons use.

\section{Melee Proficiencies}
\subsection{Wrestling}
Grappling, wrestling, and throwing, both in unarmed and armed situations. Very useful in and out of armor, wrestling is one of the few ways to defeat a foe in armor without a heavy armor penetrating weapon.

Wrestling has access to a great number of superior grappling Maneuvers. 

\subsubsection{Core Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item \nameref{man:Clinch (Defense)} [Superior]
	\item Upgrade one Clinch Maneuver to Superior
\end{itemize}
\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Upgrade one Clinch Maneuver to Superior
	\item \nameref{man:Disarm} [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
All remaining Clinch Maneuvers gain Superior

\subsection{Pugilism}
Fighting with fists, feet, knees, elbows, and so on. Pugilism is the striking component of unarmed combat, and focuses less on wrestling and grappling, and more on striking the opponent. 

Pugilism has access to a great number of Superior Maneuvers. 

\subsubsection{Core Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Upgrade an Unarmed Maneuver to Superior (choose three: \nameref{man:Kick}, \nameref{man:Straight Punch}, \nameref{man:Hook Punch}, \nameref{man:Elbow}, \nameref{man:Knee}, \nameref{man:Shoulder Charge})
	\item \nameref{man:One-Two Punch}
\end{itemize}

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt} \begin{itemize}
	[itemsep=0.5mm]
	\item Upgrade an Unarmed Maneuver to Superior (choose one: Kick, Straight Punch, Hook Punch, Elbow, Knee, Shoulder Charge)
	\item Disarm [Superior]
	\item Arm Parry [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Upgrade an Unarmed Maneuver to Superior (choose one: Kick, Straight Punch, Hook Punch, Elbow, Knee, Shoulder Charge)
	\item Clinch [Superior]
\end{itemize}

\subsection{Dagger}
Fighting with knives, daggers, and other short cutting and stabbing weapons. Certain daggers may tread 
the line between dagger and sword. Dagger fighting is useful both in a civilian and military setting -- the 
dagger is often the last line of defense a fighter has, as well as being useful for finishing off grounded 
opponents. The weapon also has darker applications for assassins and murderers, as its size lends well to 
easy concealment. Stilettos, Rondel Daggers, Cinquedeas, and Misericordes are all useful daggers for various situations.

Using a dagger allows access to a combination of both armed and unarmed Maneuvers, making it highly 
versatile. 

\subsubsection{Core Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Deep Draw Cut
	\item Pommel Strike
\end{itemize}

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Dacian Rush
	\item Arm Parry [Superior]
	\item Pommel Strike [Superior]
	\item Disarm [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
Clinch [Superior]

\subsection{1H Sword}
The single-handed sword has always held a special place in warfare as the companion weapon of the 
professional soldier. As such, practice with the sword has always been associated with martial virtue and 
professionalism, from the ascendancy of the Roman Legion to the collapse of Napoleon’s Empire. The 
Gladius, the Arming Sword, the Rapier, Saber and Backsword, all carry a respectable legacy of violence. 

1H Sword has access to many highly technical Maneuvers, as well as early access to many advanced Maneuvers, including Shield Maneuvers.

\subsubsection{Core Maneuvers}
Pommel Strike

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Shield Bind
	\item Deep Draw Cut
	\item Riposte
	\item Murder-Strike
	\item Half-Sword
	\item Quickdraw
	\item Shield Bash [Superior]
	\item Shield Beat [Superior]
	\item Pommel Strike [Superior]
	\item Simultaneous Parry/Attack
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Beat [Superior]
	\item Simultaneous Parry/Attack [Superior]
	\item Disarm [Superior]
	\item Hilt Push [Superior]
	\item Push Cut [Superior]
	\item Riposte [Superior]
	\item Shield Bind [Superior]
\end{itemize}

\subsection{2H Sword}
Two handed swords, though less common than their single-handed cousins throughout 
history, were frequently weapons of war intended for use in the thick of battle by elite troops. The Landsknechts with their Zweihanders were the defenders of battle standards and were also the bodyguards 
of generals. Spanish Montante manuals contain detailed plays for the fighting of many foes at once. The 
more modest Longsword, Bastard Sword and other similar weapons dominated the dueling scene in 
Europe for centuries as well. 

2H Sword has access to many highly technical Maneuvers, as well as early access to many advanced Maneuvers. 

\subsubsection{Core Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Pommel Strike
	\item Simultaneous Parry/Attack
	\item Half-Sword
	\item Murder-Strike
\end{itemize}
\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Quickdraw
	\item Riposte
	\item Deep Draw Cut
	\item Hilt Push [Superior]
	\item Disarm [Superior]
	\item Pommel Strike [Superior]
	\item Simultaneous Parry/Attack [Superior]
	\item Shield Bash [Superior]
	\item Shield Beat [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Beat [Superior]
	\item Riposte [Superior]
	\item Break [Superior]
	\item Push Cut [Superior]
\end{itemize}
\subsection{1H Blunt}
Swords are all well and good, but sometimes you’ve just got to hit someone with an axe. Blunt 
weapons weren’t always blunt--axes meant for war were actually thinner than wood-splitting axes, and were lighter and sharper. The end-heavy weighting of the weapons does lend them a similar technique. 
Warhammers, Maces, Battleaxes and Truncheons are more effective against armor than slashing weapons 
like swords, and were useful tools against plate-armored adversaries. 

1H Blunt has access to fewer Maneuvers than the Sword Proficiencies, but makes up for this with 
powerful versions of Maneuvers meant to destroy enemy armor, shields, and weapons, as well as Shield Maneuvers.

\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Shield Bind
	\item Beat [Superior]
	\item Hook [Superior]
	\item Break [Superior]
	\item Hew [Superior]
	\item Shield Bash [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Riposte
	\item Disarm [Superior] 
	\item Shield Beat [Superior] 
	\item Shield Bind [Superior] 
\end{itemize}
\subsection{2H Blunt}
Heavy blunt weapons hold a particular place in history, as both the tools of elite warriors and 
peasants alike. The Dane Axe was used by the Huscarl elite of Scandinavia and England, whereas the 
Pole-Flail was reportedly used by the peasant armies of the Hussites and the Belgians during their various 
revolts. The Morningstar and Heavy Mace were often used by professional soldiers as late as the 16th 
Century. Cheaper than swords, but more effective against heavy armor, these were often the weapons of choice in the age of high chivalry.

2H Blunt has access to fewer Maneuvers than the Sword Proficiencies, but makes up for this with 
powerful versions of Maneuvers meant to destroy enemy armor, shields, and weapons. 

\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Beat [Superior]
	\item Shield Bind
	\item Hook [Superior]
	\item Break [Superior]
	\item Hew [Superior]
\end{itemize}

\subsubsection{Tier 2 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Riposte
	\item Disarm [Superior]
\end{itemize}
\subsection{Spear}
Spears have always been ubiquitous weapons. Cheap, easy to make, and easy to train with, they 
were useful for every kind of soldier from the conscript to the professional, and remained the most 
common weapon for most of history, with only various swords coming anywhere close in terms of 
deployment. The Spear came in various lengths, from the handy Arming Spear to the extremely long Pike. 
Certain specialized versions, such as the Ahlspiess, were designed for duties for which the regular spear 
was unsuited, like penetrating plate. 

Spear has access to many highly technical Maneuvers, just like Sword, but it lacks some of the destructive 
Maneuvers of the Polearm and Blunt. 

\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Riposte
	\item Simultaneous Parry/Attack
	\item Shield Bind
	\item Shield Bash [Superior]
	\item Shield Beat [Superior]
	\item Push Cut [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
\begin{itemize}
	[itemsep=0.5mm]
	\item Simultaneous Parry/Attack [Superior] 
	\item Riposte [Superior] 
	\item Shield Bind [Superior]
\end{itemize}
\subsection{Polearm}
Fighting with Polearms is the most important skill for most soldiers. Polearms were present on 
battlefields the world over, but especially in the Medieval world. Polearms can thrust, smash through 
armor, hook foes off horses, and otherwise ruin anyone’s day. The Halberd is the best known polearm, 
but the Bill, the Voulge, the Glaive and the Lucerne Hammer were equally common. 

Polearm has access to a variety of useful Maneuvers, including a mixture of destructive and technical 
Maneuvers. It lacks Shield Maneuvers since most polearms were meant to be used two-handed. 

\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item \nameref{man:Riposte}
	\item \nameref{man:Hook} [Superior] 
	\item \nameref{man:Disarm} [Superior] 
	\item \nameref{man:Hew} [Superior] 
	\item \nameref{man:Break} [Superior] 
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
Simultaneous Parry/Attack 

\section{Missile Proficiencies}
Missile Proficiencies work very similarly to Melee Proficiencies, but with a few small differences. 
Notably, Missile Proficiencies have no Defaults (see below, in Alternate Weapon Usage) or Talents 
available to them. 

\subsection{Bow}
\subsubsection{Core Maneuvers}
Universal Maneuvers only 

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Power Draw
	\item Shoot [Superior]
\end{itemize}
\subsubsection{Tier 2 Maneuvers}
Rapid Shot [Superior] OR Power Draw [Superior]

\subsection{Crossbow}
\subsubsection{Core Maneuvers}
Universal Maneuvers only 

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Aim [Superior]
	\item Shoot [Superior]
\end{itemize}
\subsection{Firearms}
\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
\vspace{-5pt}\begin{itemize}
	[itemsep=0.5mm]
	\item Aim [Superior]
	\item Shoot [Superior]
\end{itemize}
\subsection{Throwing Weapons}
\subsubsection{Core Maneuvers}
Universal Maneuvers only

\subsubsection{Tier 1 Maneuvers}
Fling [Superior]

\chapter{Maneuvers}\label{ch:Maneuvers} %NEED to make some concept photos of possible maneuver layouts. Maneuvers are the most important part of the book, gotta make sure they work! Possibly the actual list of maneuvers should have its own chapter since it's quite extensive? This would also prevent bookmark bloat
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Maneuvers are essentially ‘moves’ used in combat. They can range from parrying a blow to firing a pistol and swinging an axe. Most Maneuvers involve devoting CP dice to them, and then rolling these as an attack or a defense against an Opponent. Some have more subtle effects, like changing the way a weapon is held, distracting or blinding an Opponent, or maneuvering around them in combat.

Maneuvers are divided into simple categories based on what they do, and when they can be used. For example, Parry is a Defense Weapon Maneuver, because it involves parrying an attack with a weapon. A Lance Charge is an Attack Mounted Weapon Maneuver, since it involves a kind of attack that can only be performed when mounted with a weapon. Most Maneuvers will require the character be wielding an appropriate weapon, which is listed in the Requirements section of each Maneuver.


\subsection{Gaining Maneuvers}
Some Maneuvers are automatically gained at Character Creation, and are able to be used by all characters with any Proficiency. These Maneuvers are identified as ‘Universal.’

Other Maneuvers, identified as ‘Advanced’ must be earned through leveling Proficiencies and unlocking Maneuver Tiers (refer to \autoref{ch:proficiencies} for more information). Advanced Maneuvers can only be used when a character is currently using a Proficiency with the Advanced Maneuver unlocked. That means, if your character is using a Bow, they cannot utilize the Advanced Maneuvers available to a 2H Blunt weapon, even if that Maneuver is unlocked.

\subsection{Using Maneuvers}
The Superior versions of Maneuvers are considered Advanced for all purposes.

Some Maneuvers, such as Grappling and Hilt Push, result in a special state in which additional Maneuvers become available. These are indicated by having the parent Maneuver’s name before theirs (i.e. Hilt Push: Grapple, or Hilt Push: Lever-Down).

\section{\label{sec:Maneuvers}The Anatomy of a Maneuver}
All Maneuvers follow this format:


\paragraph{\large Maneuver Name}---\quad [\large Cost]
		\vspace{-10pt}\begin{description}
		\item [Type:] This is the type of the maneuver, for reference to how it can be used. If the proficiency is tagged as ‘Universal,’ it can be used with any Proficiency. If it is tagged as ‘Advanced’ instead, the entry will also list which Proficiencies receive the Maneuver. Some Maneuvers are tagged as being ‘Instant’ and specify their base type, such as ‘Swinging’ or ‘Thrusting.’
		\item [Requirements:] These are the requirements to perform the Maneuver.
		\item [Maneuver:] This is the actual mechanic for performing the Maneuver.
		\item [Success:] This is what happens if the Maneuver succeeds.
		\item [Failure:] This is what happens if the Maneuver fails. If a Maneuver has no Failure listed, then it is simply ‘Negated,’ a generic failure that has no special effect.
		\item [Special:] Additional mechanics, conditions and restrictions are explained here.
		\item [Superior:] Some Maneuvers have ‘Superior’ versions that increase the effectiveness of the Maneuver, reduce its cost, or otherwise improve it. Using the Superior version of a Maneuver is optional, the lesser version can always be used. Superior Maneuvers count as the same Maneuver for the purposes of Poor Form, Good Form, and so on.
\end{description}
The table below lists all Maneuvers by their Primary and Secondary roles (a Swing is an Attack using a 
Weapon). It also shows whether a Maneuver is Universal or Advanced (noted as ‘U’ and ‘A’), as well as 
if the Maneuver has a Superior version (‘S’). The cost of the Maneuver is shown, however it is always a 
good idea to read the full text of a Maneuver to understand how it is used and how much it will cost you 
to activate. Some Maneuvers are categorized by type, and certain types of Maneuvers may be affected by 
others, or Boons, Banes and Talents.

\begin{longtabu} to \linewidth {*{8}{X[-1]}}
Maneuver& U &A &S &Primary &Secondary &Type &Cost\\\toprule
\nameref{man:Beat} &U &&S &Attack&Weapon&Swinging& X \\
\nameref{man:Break} &U &&S &Attack&Weapon&Swinging& X+2 \\
\nameref{man:Butt-Strike} &U&& &Attack&Weapon&Swinging& X+2 \\
\nameref{man:Dacian Rush} &&A& &Attack&Weapon&Thrusting& 2+X+Y \\
\nameref{man:Deep Draw Cut} &&A& &Attack&Weapon&Swinging& 0 \\
\nameref{man:Disarm} (Attack) &U&& S &Attack&Weapon& (Swinging, Punching) &X+1\\
\nameref{man:Feint} &U&& &Attack&Weapon&Instant& 2+Variable\\
\nameref{man:Half-Sword} &  &A &  &Attack&Weapon&Instant& 1\\
\nameref{man:Hew}  &U&&S &Attack&Weapon&Swinging& 0 \\
\nameref{man:Hook} &U&&S &Attack&Weapon& (Swinging, Thrusting) &1\\
\nameref{man:Joint Thrust} &U&& &Attack&Weapon&Thrusting& X+2\\
\nameref{man:Melee Shoot} &U &&&Attack&Weapon&Thrusting& X+Variable\\
\nameref{man:Murder-Strike} &&A& &Attack&Weapon&Instant& 2\\
\nameref{man:Pommel Strike} &&A& S &Attack&Weapon&Thrusting& X \\
\nameref{man:Push Cut} &U&& S &Attack&Weapon&Thrusting& X+1\\
\nameref{man:Quickdraw} &U&&&Attack&Weapon&Instant& 1 \\
\nameref{man:Swing} &U&& &Attack&Weapon&Swinging& X \\
\nameref{man:Thrust} &U&& &Attack&Weapon&Thrusting& X \\
\nameref{man:Weapon Throw} &U&& &Attack&Weapon&Thrusting& X+1 \\
\nameref{man:Clinch (Attack)} &U&& S& Attack &Unarmed &Grappling &X\\
\nameref{man:Elbow} &U&& S &Attack &Unarmed &Unarmed & X \\
\nameref{man:Headbutt} &U &&&Attack &Unarmed &Unarmed & X+1 \\
\nameref{man:Hook Punch} &U&&S &Attack &Unarmed &(Unarmed, Punching) &X\\
\nameref{man:Kick} & U&&S& Attack &Unarmed &Unarmed & X \\
\nameref{man:Knee} & U&&S& Attack &Unarmed &Unarmed & X \\
\nameref{man:One-Two Punch} & &A&& Attack &Unarmed &(Unarmed, Punching) &X+2 \\
\nameref{man:Shoulder Charge} & U &&&Attack &Unarmed &Unarmed & X \\
\nameref{man:Straight Punch} & U&&S &Attack &Unarmed &(Unarmed, Punching) &X\\
\nameref{man:Shield Bash} & U&&S &Attack &Shield &Bash & X \\
\nameref{man:Shield Beat} & U&&S &Attack &Shield &Bash & X+2 \\
\nameref{man:Shield Feint} & U&&& Attack &Shield & ????????????? &1 \\
\nameref{man:Blind Toss} & U&& &Attack &Special &????????????? &X+Variable \\
\nameref{man:Strangle} & U&&& Attack &Special &Clinching &X \\
\nameref{man:Parry} & U &&&Defense &Parry &Parrying &X \\
\nameref{man:Quick Defense} & U &&&Defense &Parry &Instant &2 \\
\nameref{man:Riposte} & &A &S &Defense &Parry& (Parrying, Instant) &X+2 \\
\nameref{man:Arm Parry} & U&&S &Defense &Unarmed &Parrying &X\\
\nameref{man:Clinch (Defense)} & ? &? &? &Defense &Unarmed &Grappling &???\\
\nameref{man:Missile Catch} & U &&&Defense &Unarmed &????????????? &2+X \\
\nameref{man:Block} & U&& &Defense &Block &Blocking &X \\
\nameref{man:Shield Bind} & &A&& Defense &Block &Blocking &X+1 \\
\nameref{man:Total Block} & U &&&Defense &Block &Blocking &X \\
\nameref{man:Closing Void} & U&&& Defense &Void &Void &X \\
\nameref{man:Dart-Out} & U &&&Defense &Void &Void &X+ENC \\
\nameref{man:Flee} & U &&&Defense &Void &Void &X+ENC \\
\nameref{man:Take Cover} & U &&&Defense &Void &Void &X+ENC \\
\nameref{man:Void} & U &&&Defense &Void &Void &X+ENC \\
\nameref{man:Pass (Do Nothing)} & U &&&Defense &Special &????????????? &0 \\
\nameref{man:Cleaving Blow} & U &&&Cross Fighting &Attack &Swinging &2 \\
\nameref{man:Target Change} & U &&&Cross Fighting &Attack &????????????? &1 \\
\nameref{man:Ally Defense} & U &&&Cross Fighting &Defense &????????????? &2 \\
\nameref{man:Double Attack} & U&&&Simultaneous &General &Attack &X+Y+1 \\
\nameref{man:Double Shot} & U&&&Simultaneous &General &Attack &X+Y+2 \\
\nameref{man:Simultaneous Block/Attack} & U&&&Simultaneous &General &(Blocking, Attack) &X+Y \\
\nameref{man:Simultaneous Off-Hand Parry/Attack} & U&&&Simultaneous &General &(Parrying, Attack) &X+Y+2 \\
\nameref{man:Simultaneous Parry/Attack} & &A &S& Simultaneous &General &(Parrying, Attack) &X+Y+2 \\
\nameref{man:Simultaneous Void/Attack} & U&&&Simultaneous &General &(Void, Attack) &X+Y+2 \\
\nameref{man:Aim} & U&&S &Missile &General &????????????? &0 \\
\nameref{man:Braced Shot} & &A&& Missile &Shooting &Shooting &???\\
\nameref{man:Power Draw} & &A &S &Missile &Shooting &Shooting &??? \\
\nameref{man:Rapid Shot} & U&&S &Missile &Shooting &Shooting &??? \\
\nameref{man:Reload} & U&& &Missile &Shooting &???????????? &??? \\
\nameref{man:Shoot} & U&&S& Missile &Shooting &Shooting& ??? \\
\nameref{man:Double Fling} & U&&S &Missile &Throwing &Throwing &??? \\
\nameref{man:Fling} & U&&S& Missile &Throwing &Throwing &??? \\
\nameref{man:Grappling: Choke} & U &&&Grapple &Advantage& ???????????? &X \\
\nameref{man:Grappling: Coup de Grace} & U &&&Grapple &Advantage &Thrusting &X+2\\
\nameref{man:Grappling: Limb-Break} & U&&S &Grapple &Advantage &???????????? &X \\
\nameref{man:Grappling: Pin} & U&& &Grapple &Advantage &???????????? &X \\
\nameref{man:Grappling: Takedown} & U &&&Grapple &Advantage &???????????? &X \\
\nameref{man:Grappling: Throw} & U&&S &Grapple &Advantage &???????????? &X+2 \\
\nameref{man:Grappling: Trap} & U&&S &Grapple &Advantage &??????????? &X \\
\nameref{man:Grappling: Slip} & U &&&Grapple &Disadvantage &??????????? &X \\
\nameref{man:Grappling: Twist} & U &&&Grapple &Disadvantage &???????????? &X \\
\nameref{man:Grappling: Bite} & U &&&Grapple &General &Unarmed &X+1 \\
\nameref{man:Grappling: Force} & U&& &Grapple &General &???????????? &X+1 \\
\nameref{man:Grappling: Unarmed Attack} & ? &?& ? &Grapple &General &???????????? &??? \\
\nameref{man:Grappling: Use Weapon} & ? &? &? &Grapple &General& ???????????? &??? \\
\nameref{man:Hilt Push} & U&&S& Hilt Push &??? &Instant &2 \\
\nameref{man:Hilt Push}: Butt-Strike &U&&&&&& \\
\nameref{man:Hilt Push}: Evasion &U&&&Hilt Push &??? &???& \\
\nameref{man:Hilt Push}: Clinch &U&&&Hilt Push &Advantage& &??? \\
\nameref{man:Hilt Push}: Half-Sword &U &&&Hilt Push &Advantage& &??? \\
\nameref{man:Hilt Push}: Lever Down &U &&&Hilt Push &Advantage& &2+X \\
\nameref{man:Hilt Push}: Pommel Strike& U&&& Hilt Push &Advantage &&??? \\
\nameref{man:Hilt Push}: Slide-Cut &U &&&Hilt Push &Advantage &&??? \\
\nameref{man:Hilt Push}: Slide-Thrust &U&& &Hilt Push &Advantage &&??? \\
\nameref{man:Hilt Push}: Bind &U&& &Hilt Push &Disadvantage &&??? \\
\nameref{man:Hilt Push}: Wind &U&& &Hilt Push &Disadvantage &&??? \\
\nameref{man:Hilt Push}: Withdraw &U &&&Hilt Push &Disadvantage &&??? \\
\nameref{man:Obstruction} & U &&&Mobility &General& &??? \\
\nameref{man:Outmaneuver} & U &&&Mobility &General &Instant &X+2+ENC \\
\nameref{man:Rapid Rise} & U &&&Mobility &General &Instant &3+(ENC*2) \\
\nameref{man:Slip-About} & U &&&Mobility &General &Instant &X+ENC+Special \\
\nameref{man:Steal Initiative} & U &&&Mobility &General &Instant &X+Opponent PER \\
\nameref{man:Thread the Needle} & U &&&Mobility &General &Instant &X+Y+ENC \\
\nameref{man:Lance Charge} & U &&&Mounted &Attack& &X+2\textasciicircum \\
\nameref{man:Mount Bite} & U&& &Mounted &Attack& &2\textasciicircum \\
\nameref{man:Mount Kick} & U &&&Mounted &Attack& &2\textasciicircum \\
\nameref{man:Disengage} & ? &? &? &Mounted &Defense& &??? \\
\nameref{man:Lean} & U &&&Mounted &Defense &Void &X+ENC\textasciicircum \\
\nameref{man:Mount Dodge} & U &&Mounted &Defense &Void &&2\textasciicircum \\
\nameref{man:Ride-By} & U &&&Mounted &Passing& &2\textasciicircum \\
\nameref{man:Trample} & U &&&Mounted &Passing& &4\textasciicircum \\
\nameref{man:Dog Bite} & U&&& Dog &Attack &&X \\
\nameref{man:Savage} & U&&& Dog& Attack &&X \\
\nameref{man:Tackle} & U&&& Dog& Attack &&2+X \\
\\
\caption*{\textasciicircum Check required. Refer to Maneuver details.}\\
\caption*{U = Universal Maneuver \hfill A = Advanced Maneuver \hfill S = Superior Maneuver}\\
\caption*{ENC = Encumbrance Penalty (presumably to CP).}\\
\end{longtabu}

\section{Attack Maneuvers}
Attack Maneuvers are generally (though not always) declared with Initiative (see \autoref{ch:combat}). These all 
involve some sort of attack aimed at an Opponent. Attack Maneuvers are divided into sub-categories, 
which when referenced should be denoted with Attack (i.e. an Unarmed Maneuver in this category is an 
Unarmed Attack, a Weapon Maneuver in this category is a Weapon Attack, etc.) 

\subsection{Weapon Attack Maneuvers}
\paragraph{\large\label{man:Beat}Beat}---\quad{\large[X]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Swinging
\item [Requirements:] Have a weapon with a Swing TN. It must be the Round immediately after Orientation Declaration, and you must have declared Aggressive.
\item [Maneuver:] Swinging Weapon Attack at Swing TN of weapon, targeting Opponent weapon or shield with X dice. Halve all Reach costs for this Maneuver. For the purpose of Initiative tests involving this Maneuver, Reach costs are not factored into the Initiative roll. 
\item [Success:] Opponent cannot use targeted weapon or shield to perform any action until Refresh. Opponent suffers [BS] Stun. This attack does not change the Reach of combat. 
\item [Superior:] You may now declare this Maneuver in any Action immediately after Refresh if you have Initiative. 
\end{description}
\begin{quotation}
\emph{Vladock is having trouble getting past his opponent’s spear. He declares a Beat with 8 dice. His opponent declares a Void with 7. Vladock gets 3 BS to his opponent’s 2, and so beats aside his opponent’s weapon and costs him 1 Stun. Vladock can now make an attack in the next Action without fear of being counter-attacked by the spear, since it cannot be used for anything. He’ll still have to pay Range costs, but at least it’s a safe bet instead of a dangerous move. }
\end{quotation}

\paragraph{\large\label{man:Break}Break}---\quad{\large[X$ + $2]}
	\vspace{-10pt}
\begin{description}
	\item[Type:] Universal, Swinging
	\item[Requirements:] Have a weapon with a Swing TN. Opponent has a weapon.
	\item[Success:] Opponent’s weapon sustains damage equal to [STR+Weapon Swing Damage+BS]. If the amount of damage dealt equals or exceeds the damage threshold (listed below) for the targeted weapon, it breaks, and becomes useless except as a potential improvised weapon. At the GM’s discretion, a spear or polearm may effectively become a quarterstaff after being broken, a wood-hafted blunt weapon might become a truncheon, and so on. 
	\item[Special:] This Maneuver ignores Reach modifiers. The user does not move to their weapon’s Reach on a successful attack, but may if the Opponent’s weapon is broken. 
	\item[Superior:] Break inflicts +2 additional damage. 
\end{description}

\begin{table}[!hb]
	\centering
	\caption{Weapon Damage Threshold}
	\label{tab:Weapon Damage Threshold}
	\begin{tabu} spread 0pt {XX[-1,r]}
Weapon&Damage Threshold \\\toprule
Dagger&12\\
Light Blade Sword &15 \\
Regular Sword &20 \\
Wood-Hafted Blunt &10 \\
Metal-Hafted Blunt&15 \\
Spear/Polearm&10\\
Langetted/Reinforced Spear/Polearm&12 \\
	\end{tabu}
\end{table}

\begin{quotation}
\emph{Hector is charging a pikeman with his zweihander. Rather than try to close the distance immediately to attack the man, Hector declares a Break against the man’s weapon with 10 dice, paying 2 for activation. His opponent attempts a Void with 8 dice. Hector succeeds with 3 dice to 2, and inflicts 7+3+1 (strength+weapon+bonus successes) damage for a total of 11, breaking the pike, whose durability was 10. The Pikeman now has only a broken stick, and Hector can deal with him on more even terms.}
\end{quotation}


\paragraph{\large\label{man:Butt-Strike}Butt-Strike}---\quad{\large[X]}
\begin{description}
\item [Type:] Universal, Attack, Swinging 
\item [Requirements:] Have a Polearm or Spear weapon, or a Sword being held Half-Swording. 
\item [Maneuver:] Swinging Weapon Attack at Short Reach at Weapon Swing TN, aimed on Swing Table. 
\item [Success:] Inflicts Bludgeoning damage equal to [STR+0+BS] to Hit Location. 
\item [Special:] If this weapon has a metal cap of some sort on the weapon’s reverse end (or if it is a sword being used) this attack inflicts +1 Bludgeoning damage.  If Half-Swording the weapon being used for Butt-Strike, the weapon does not suffer the penalty to Swinging damage normally associated with Half-Swording. 
\end{description}

\paragraph{\large\label{man:Dacian Rush}Dacian Rush}---\quad{\large[2+X+Y]}
\vspace{-10pt}\begin{description}
\item [Type:] Advanced, Attack, Thrusting 
\item [Requirements:] Have a Dagger. If opponent declares a Quickdraw, you may treat this Maneuver as an Instant that occurs before their Quickdraw.
\item [Maneuver:] Unarmed Attack at Hand Reach with X at Hand Blind TN 7 vs Opponent PER at TN 7. If Opponent has declared a Quickdraw, it is delayed until after this Action is completed. Any defense declared against this Maneuver is not resolved against the Hand Blind, but against the Thrust to follow. Success [Hand Blind]: Opponent suffers [BS] Stun. If a Quickdraw was declared, it is cancelled. You may immediately roll a Thrust Attack with Y as normal. Any non-Weapon Defense against your Thrust Attack is resolved at +2 TN.
\item [Failure - Hand Blind:] You may immediately roll a Thrust Attack with Y as normal. 
\item [Success Thrust:] Resolve Thrust as normal. 
\item [Special Rapid Thrust:] Weapons of Short Reach or Shorter only. If Thrust is successful, you may pay 1+X additional CP to make another Thrust with X dice as part of the same Action. Resolve Thrust as normal. The Opponent may use Quick Defense against this Maneuver. 
\end{description}
\begin{quotation}
\emph{Vlatko rushes out of concealment to assassinate a Marju Khan, who has just left the bathhouse in a robe. He has a Pishkan Toothpick as his weapon, and declares a Dacian Rush. Vlatko has 15 CP. The Khan has 20 normally, but this Action he is surprised, and has only 10. Vlatko declares a Dacian Rush with 2+6+7 dice. The Khan tries to Quickdraw his saber by spending 1 CP, so that he can Parry the Dacian's knife with 8 dice.}

\emph{First the Khan must roll his PER (5) versus Vlatko's 6 dice, both at TN 7. Vlatko succeeds by 1 BS, and in doing so inflicts 1 Stun, cancels the Khan's Quickdraw, and also his Parry, returning half of the dice spent (0 from Quickdraw, 4 from the Parry) to his pool. Stun reduces this to 3, meaning the Khan only has 4 CP left. Vlatko now resolves his Thrust with 7 dice, the Khan tries to use Quick Defense with an Arm Parry by spending 2 and then defending with his remaining 2 dice. However, this Arm Parry is made at +2 TN (TN 9 for Arm Parry) meaning the Khan is less than likely to succeed. Vlatko gets 4 successes, the Khan gets 0. Vlatko inflicts STR+Weapon Damage+BS for 5+1+4 piercing damage to the Khan’s TOU of 5 and the AV of his clothing (ruled to be 1) inflicting a level 4 piercing Wound to the Belly through the Khan's robe. The Khan takes more damage than he can handle, and fails his Overshock test, becoming prone.}

\emph{Both parties have 0 CP left. If Vlatko had a few more dice, he could declare a Rapid Thrust to stab the man again, but instead he'll just wait for the Refresh to finish the injured warlord off.}
\end{quotation}

\paragraph{\large\label{man:Deep Draw Cut}Deep Draw Cut}---\quad{\large[0]}
\vspace{-10pt}\begin{description}
\item [Type:] Advanced, Attack, Swinging 
\item [Requirements:] Wielding a weapon which inflicts Cutting damage on Swing. 
\item [Maneuver:] Declare a Swing Attack as normal, but decrease the Reach of the weapon by one step (Medium to Short, Short to Hand, and so on) paying Reach costs for the new Reach if necessary. Reduce the number of BS needed to trigger Draw Cut with this Swing by 1.
\end{description}

\begin{quotation}
\emph{Jerzy, who is fighting with brigands, and has been stabbed in the side, has just stolen Initiative, but he only has 4 remaining dice. His Hussar Saber has Draw 3, and he means to take advantage of it. He declares a Deep Draw Cut, moving his weapon to S Reach (convenient, as the fight is now at S Reach because of his stabbing) putting 4 dice on a Strike to the neck. He gets 2 successes, and thanks to Deep Draw Cut, that’s enough to activate the Draw damage of his Hussar Saber. His Strength of 4, plus 1 for the weapon, plus 3 for the Draw, plus 2 for the BS equals 10 total damage. Jerzy scores a Level 5 Wound, and decapitates his foe. If he hadn’t used Deep Draw Cut, he would’ve only inflicted 7 damage, which would’ve only been a level 2 Wound, and his opponent’s attack might’ve gone through\ldots}
\end{quotation}

\paragraph{\large\label{man:Disarm}Disarm}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Attack, Swinging or Punching 
\item [Requirements:] None 
\item [Maneuver:] Swinging Weapon Attack at Weapon Reach at Parry TN, OR Punching Unarmed Attack at Hand Reach at Punch TN 6, targeting Opponent weapon. 
\item [Success:] Opponent must make an AGI test with RS equal to BS. If this test is failed, the targeted weapon is removed from the Opponent's grasp, and either retained by you, dropped at your feet, or thrown several yards away at your discretion.
\item [Special:] If an Opponent’s weapon is a 2H weapon, the Opponent gains a +1 bonus to his AGI test.  If the weapon is connected to your character by some means (a lanyard or a chain, etc) the weapon is dropped, but can be retrieved as though Drawing the weapon, or Quickdraw. 
\item [Superior:] When disarming an Opponent with a 2H weapon, they do not gain the normal +1 bonus to the 
stability test.
\end{description}
\begin{quotation}
\emph{Dusan and Mirza are practicing their swordsmanship. Dusan has declared a Disarm with 6 dice, and Mirza has declared a Parry with 5. Dusan scores 3 successes to Mirza’s 2. Mirza must now make an AGI test at RS 3 (2 base +1 for BS). He has 5 AGI, but only rolls 2 successes. He is disarmed, and his weapon is tossed some distance away. Mirza grins bitterly as he goes to retrieve his sword.}
\end{quotation}

\paragraph{\large\label{man:Feint}Feint}---\quad{\large[2+Variable]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Instant 
\item [Requirements:] Activate after having declared a Swinging or Thrusting Maneuver, and after any other Maneuvers have been declared, but before any are resolved. You may only activate this Maneuver from a Thrusting Maneuver if the weapon being used has the Fluid Thrusts quality. 
\item [Maneuver:] When activated, change either the type of the attack (to either a Swinging or Thrusting Maneuver) or (if the Maneuver targets an area) the Hit Location of the attack (Face to Chest, Chest to Upper Arm, etc.). You may elect to change both if applicable, but you must choose at least one. You may pay 2 dice to add 1 dice to this attack, as many times as you can afford. Additionally, your opponent suffers a +1 TN to any Defense made against this Attack unless he plays 2 CP.
\end{description}

\paragraph{\large\label{man:Half-Sword}Half-Sword}---\quad{\large[1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Instant 
\item [Requirements:] Have any Sword without the Light Blade special rule. 
\item [Maneuver:] This weapon is now Half-Swording. Reduce the Reach of your weapon by 1 stage, and reduce Swing damage by 2. If the weapon being used was 1H with Hand-Off, it is now 2H. If it was 2H, it remains 2H. When performing a Joint Thrust Maneuver while Half-Swording, your weapon counts as having the Thin Blade rule, if it did not previously, and the damage of the Joint Thrust is made at [Weapon Damage+STR+BS] instead of [Weapon Damage+Half STR+BS]. Additionally, Pommel Strike Maneuvers performed while Half-Swording inflict +1 damage. You may deactivate this Maneuver instantly and at no cost before an Action. The weapon immediately regains its original characteristics and loses all bonuses. This Maneuver can be activated instantly before declaring another Maneuver. This Maneuver cannot be activated and deactivated in the same Action.
\item [Special:] If the weapon being Half-Sworded has the Hand-Off quality, determine its Half-Sword attributes 
based on its 2H version, regardless of whether the Maneuver was activated from 1H or 2H mode. 
\end{description}
\paragraph{\large\label{man:Hew}Hew}---\quad{\large[0]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Swinging 
\item [Requirements:] Have a weapon with a Swing TN. 
\item [Maneuver:] Swinging Weapon Attack at Weapon Reach, at Weapon Swing TN targeting Opponent shield.
\item [Success:] Opponent takes [half BS] Stun. Opponent’s shield then sustains damage equal to [STR+Weapon Damage+BS] against its Durability. If the damage equals or exceeds its Durability, the shield is destroyed. If it does not, the shield remains intact but permanently loses 2 Durability, weakening it to further Hews attempts.
\item [Special:] Blocking Maneuvers made to defend against this Maneuver are made at +1 TN. 
\item [Superior:] If this Hew fails to destroy a shield it hits, it reduces that shield’s Durability by an additional 1. 
\end{description}

\paragraph{\large\label{man:Hook}Hook}---\quad{\large[1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Swinging OR Thrusting 
\item [Requirements:] Have a weapon with the Hook quality. 
\item [Maneuver:] Swinging Weapon Attack at Weapon Reach at Weapon Swing TN, OR Thrusting Weapon 
Attack at Weapon Reach at Thrust TN, targeting Opponent shield or the Opponent. 
\item [Success:] If targeting Opponent, Opponent takes [BS] Stun, and Opponent must make a stability roll at 
[2+BS] RS or be rendered Prone. 
If targeting Shield, Opponent cannot use his shield in the following Action, and must make an AGI roll at 
[2+BS] RS or lose shield altogether as it is dragged away. 
\item [Special:] You cannot use Hook against a mounted Opponent unless the weapon being used is of L Reach 
or longer. 
\item [Superior:] If targeting an Opponent on a mount, add BS to the number of feet that the character 
effectively falls after being pulled from saddle. 
\end{description}

\paragraph{\large\label{man:Joint Thrust}Joint Thrust}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Thrusting 
\item [Requirements:] Wielding a weapon with a Thrust value that inflicts piercing damage. 
\item [Maneuver:] Thrusting Weapon Attack at Weapon Reach with X dice, rolled at Thrust TN+1, aimed at Hit 
Location with the Weak Spot quality on Opponent armor. 
\item [Success:] Inflicts piercing damage equal to [Weapon Thrust Damage+Half STR+BS] to chosen Weak Spot 
Hit Location. Do not roll on table. The damage from this attack is not reduced by AV from any armor 
which has a Weak Spot for the Hit Location that was targeted by this attack. All other AV protecting that 
area applies. 
\item [Special:] When rolling an Initiative test to determine attack order while making a Joint Thrust, you may 
roll 1 additional die in the test. 
As listed in the Special Qualities section for weapons, those with Thin Blade reduce their Thrust TN by 1 
when they perform this Maneuver (since Joint Thrust already worsens the Thrust TN of the weapon by 1, 
this simply means that the attack is made at Thrust TN.) 
This attack does not benefit from AP Thrust [X] unless there is Hard Armor on the Hit Location aside 
from that which is being ignored by this Maneuver, such as mail worn beneath the plate, in which case it 
does benefit from AP Thrust [X], and also from rules like Thin Blade’s AP Thrust 3 vs mail. 
\end{description}

\paragraph{\large\label{man:Melee Shoot}Melee Shoot}---\quad{\large[X+Variable]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Attack, Thrusting 
\item [Requirements:] Have a loaded/spanned Bow, Crossbow or Firearm ready in the hand, or declare this 
Maneuver with Quickdraw on a holstered/slung loaded Bow, Crossbow or Firearm. 
\item [Activation Cost:] If weapon is already in hand and readied, 0. If weapon has just been drawn using 
Quickdraw this Action, 2. 
\item [Maneuver:] Thrusting Weapon Attack with X not exceeding MP at current Combat Range OR Current 
Range with Missile TN of weapon, aimed on the Thrusting Attacks Table. 
\item [Success:] Inflicts damage appropriate for a ranged attack with the weapon used, using BS from the roll as 
normal, but rolling on the Thrusting Attacks Table. 
\item [Special:] If this attack is defended against with a Weapon Defense Maneuver, there is an additional 
Activation Cost to the defense equal to the Range difference between the character and the Opponent. If 
the Weapon Defense is successful, move to Defender’s Weapon Reach. 
When rolling an Initiative test to determine attack order while making a Melee Shoot Maneuver, you may 
roll 1 additional die in the test. 
\end{description}
\paragraph{\large\label{man:Murder-Strike}Murder-Strike}---\quad{\large[2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Instant 
\item [Requirements:] Have a 1H or 2H Sword with base Guard Value 2 (custom modification bonuses don’t 
count). Have two hands available to hold the weapon. This Maneuver can only be activated if you have 
used no other Maneuvers first. 
\item [Maneuver:] When activated, roll ADR at 3 RS. This is instant, and does not resolve in Initiative. If 
successful, this Maneuver takes effect immediately, and you may declare another Maneuver this Action. 
If unsuccessful, this Maneuver takes effect at the end of the Action, and the weapon cannot be used for 
any other Maneuvers until then. 
\item [Success:] This weapon is now Murder-Striking. Change the Swing damage to +1b if the weapon is 1H and 
+2b if the weapon is 2H. Change the Thrust damage to -2b. The weapon loses all special qualities and 
gains Shock 2 and Heavy Weapon instead. If the weapon was 1H, it is now 2H. If the weapon was Half-
Swording, or if it was a 2H Sword held 1H through Hand-Off, revert it to its regular form before applying 
these changes. 
\item [Special:] You may deactivate this Maneuver in an identical manner to activating it, returning the weapon 
to its normal stats. 
\end{description}
\paragraph{\large\label{man:Pommel Strike}Pommel Strike}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Attack, Thrusting 
\item [Requirements:] Have a 1H or 2H Sword or any other weapon with a pommel. 
\item [Maneuver:] Thrusting Weapon Attack at H Reach at Thrust TN 6, aimed on Thrusting Attack Table. 
\item [Success:] Inflicts bludgeoning damage equal to [STR+0+BS] to Hit Location. 
\item [Superior:] Pommel Strike also inflicts [2+BS] Stun upon hitting the Face or Upper Head. 
\item [Special:] When rolling an Initiative test to determine attack order while making a Pommel Strike 
Maneuver, you may roll 1 additional die in the test. If you are Half-Swording when you resolve your 
Pommel Strike, add +1 to its damage, as explained in the Half-Sword maneuver. 
\end{description}

\paragraph{\large\label{man:Push Cut}Push Cut}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Thrusting 
\item [Requirements:] You have a weapon with both a Thrust TN, and a Swing TN that inflicts cutting damage. 
\item [Maneuver:] Thrusting Weapon Attack at Weapon Reach, rolled at Thrust TN, aimed on the Thrusting 
Attack Table.
\item [Success:] Inflicts cutting damage equal to [STR+Weapon Swing Damage-1+BS] to Hit Location. 
\item [Special:] When rolling an Initiative test to determine attack order while making a Push Cut, you may roll 1 
additional die in the test. This is identical to the bonus gained for the Thrust Maneuver. 
This attack does not apply AP Swing/Thrust, Crushing, or Shock effects. 
This attack inflicts Draw damage. 
\item [Superior:] When performing a Push Cut, reduce the number of BS needed to trigger Draw by 1.
\end{description}
\paragraph{\large\label{man:Quickdraw}Quickdraw}---\quad{\large[1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Instant 
\item [Requirements:] Have a weapon that you want drawn, and have the capacity to draw it. This Maneuver can only be activated if no other Maneuvers have been used first.
\item [Maneuver:] When activated, roll ADR at 1 RS per Reach level of the weapon past H Reach. This is instant, and does not resolve in Initiative. H Reach weapons automatically succeed. Weapon with greater than L Reach cannot be Quickdrawn.
The RS for Quickdrawing missile weapons of various sorts is listed in the table below. (If the missile weapon cannot be)
\item [Success:] You immediately draw the weapon, and may declare a Swing, Thrust or Parry Maneuver in this same Action. These Maneuvers must be made with the Quickdrawn weapon. 
\item [Special:] If the weapon being Quickdrawn has the Draw special quality, reduces the RS to successfully Quickdraw it by 2. 
\end{description}
\begin{table}[!ht]
	\centering
	\caption{Missile Weapon Quickdraw RS}
	\label{tab:Quickdraw RS Missile Weapons}
	\begin{tabu} spread 0pt{XX}
Missile Weapon &Quickdraw RS \\\toprule
Pistol &1 \\
Dragon Pistol& 2 \\
Blunderbuss &3 \\
Musket &4 \\
Light Crossbow &3 \\
Heavy Crossbow &4 \\
Small Bow &4 \\
Large Bow &6 \\
\end{tabu}
\end{table}

\paragraph{\large\label{man:Swing}Swing}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Swinging 
\item [Requirements:] Wielding a weapon with a Swing TN. 
\item [Maneuver:] Weapon Attack at Weapon Reach with X dice rolled at Swing TN, aimed on the Swinging 
Attack Table. 
\item [Success:] Inflicts damage equal to [STR+Weapon Swing Damage+BS] to Hit Location. Damage is of 
weapon’s Swing damage type. 
\item [Power Attack (1-2):] When declaring a Swing, you may spend up to 2 CP to increase Swing damage by 
an amount equal to the number of CP spent. 
\end{description}
\paragraph{\large\label{man:Thrust}Thrust}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Thrusting 
\item [Requirements:] Wielding a weapon with a Thrust value. 
\item [Maneuver:] Weapon Attack at Weapon Reach with X dice, rolled at Thrust TN, aimed on the Thrusting 
Attack Table. 
\item [Success:] Inflicts damage equal to [STR+Weapon Thrust Damage+BS] to Hit Location. Damage is of 
weapon’s Thrust damage type. 
\item [Special:] When rolling an Initiative test to determine attack order while making a Thrust, you may roll 1 
additional die in the test. 
\item [Rapid Thrust:] This may be used by weapons of S Reach or shorter only. If Thrust is successful, you may 
pay X+2 additional CP to make another Thrust with X dice as part of the same Action. Resolve Thrust as 
normal. The Opponent may only use Quick Defense against this Maneuver. 
\end{description}
\paragraph{\large\label{man:Weapon Throw}Weapon Throw}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Thrusting 
\item [Requirements:] Have a Throwing Weapon. 
\item [Maneuver:] Thrusting Weapon Attack at current Combat Range with X dice at the Missile TN of the 
weapon aimed on Thrusting Attack Table. Weapon is thrown. 
\item [Success:] Inflicts [STR+Weapon Missile Damage+BS] to Hit Location. Weapon may be embedded in 
target, or fallen nearby (see Catch Chance rules) This attack loses 1 damage per step of Range within the 
throwing weapon’s Range value that the attack was made at. 
\item [Failure:] In addition to having no effect, weapon has likely flown past target or is stuck in a shield. 
\item [Special:] These attacks can be Parried at +2 Activation Cost, or Blocked/Voided as normal. When rolling 
an Initiative test to determine attack order while making a Weapon Throw, you may roll 1 additional die 
in the test.
\end{description}
\begin{quotation}
\emph{Duke rushes into battle with a warhammer in either hand. He is faced by a foe with a spear who outreaches him by two yards. 
Duke declares Offensive, as do his opponent, meaning it is now do-or-die time. Rather than try to defend against his opponent’s 
attack and strike at the same time, or win an Initiative test using only his S Reach, Duke declares a Weapon Throw with 8 dice. 
This lets him make the attack at the EL Reach of his Opponent’s weapon so that he doesn’t suffer any penalties, and he gains the 
+1 bonus to Initiative for the Maneuver’s special quality.}

\emph{He rolls more successes than his Opponent, so he makes his attack with 8 dice, using the Missile TN of Hammers (7) and scores 3 successes. The hammer slams into the spearman’s face, and he collapses into a heap. }
\end{quotation}


\subsection{Unarmed Attack Maneuvers}
\paragraph{\large\label{man:Clinch (Attack)}Clinch (Attack)}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Defense, Grappling 
\item [Requirements:] Must have limbs. 
\item [Maneuver:] Grappling Attack at HA Reach with X dice at Clinch TN 8. 
\item [Success:] You and your Opponent are now Grappling. You regain CP equal to BS. As part of this action 
you may immediately declare and resolve a Grapple Advantage Maneuver. Your Opponent may only 
defend against this with a Quick Defense. If your Opponent has an unresolved Action that the Grapple 
precedes, first resolve this Grapple and its secondary Maneuver, then resolve the Opponent’s Action, if 
possible. 
Immediately upon entering the Grapple, but before declaring your Grapple Advantage Maneuver (this 
applies for both you and your Opponent) you must switch to (or Default to) your Wrestling Proficiency. If 
that would change your total CP from the Proficiency you had been using, add or subtract the difference 
from your current total. After this, gain CP equal to your STR+Miscellaneous bonuses. Your maximum 
CP is always increased by your STR+Miscellaneous bonuses while Grappling. The bonus is lost when the 
Clinch ends. 
Do not proceed with declaring and resolving any other Maneuvers until the Proficiency and 
STR+Miscellaneous changes to CP are made. 
As long as you are Grappling, the Reach between you and your opponent is locked at HA Reach. Normal 
Maneuvers can still be performed in the Grapple, but can be hindered by Trappings, or Pinnings. 
While Grappling, no Maneuver that would change the Reach of the combat does so, no Maneuver that 
would render one of those in the Grapple unable to be targeted by the other does so, and no Maneuver that 
would remove the character from combat does so. 
If, at the end of an Action, you are neither Trapping nor Pinning an Opponent's body part, the Grapple 
ends. By default, when the Grapple ends, the two characters remain at HA Reach. 
\item [Failure:] If Parried or Blocked, you suffer a Swing, Thrust, or Shield Bash (depending on the Parrying/Blocking device, at the GM's discretion) to your arms (left 1-3, right 4-6), chest (7-8) or head (9-10) without BS.
\item [Superior:] Clinch resolves at Clinch TN 7. 
\end{description}


\paragraph{\large\label{man:Elbow}Elbow}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed 
\item [Requirements:] An arm with which to strike. This can be done while holding a weapon, at the GM's 
discretion. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Elbow TN 7 to Target Zone on the Swinging or 
Thrusting Attack Table. 
\item [Success:] Inflicts Unarmed Damage equal to [STR+0+BS] to Hit Location. 
\item [Failure:] If Parried, suffer a Swing or Thrust (Opponent’s choice, GM adjudicates) to Elbowing arm with 
dice equal to half of Defense successes from Parrying weapon. 
\item [Special:] If elbow is armored in metal, add +1 to damage. 
If aiming an Elbow below the waist, Activation Cost increases by 1. 
\item [Superior:] Your elbows now inflict an additional +2 damage. 
\end{description}
\begin{quotation}
\emph{Gallant has Pinned his opponent’s arm, torso, and neck using the Grappling rules, and delivers a savage Elbow to his victim’s 
head with 12 dice. He rolls 8 successes to the defenseless fool, and inflicts STR-2(4)+8 damage, for a total of 12! }
\end{quotation}

\paragraph{\large\label{man:Headbutt}Headbutt}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed 
\item [Requirements:] None. No hands required. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Headbutt TN 6 at either Face or Chest Target 
Zone on the Thrusting Attack Table. 
\item [Success:] Inflicts Unarmed Damage equal to [STR-1+BS] to Hit Location. Target suffers [BS] Stun. Half 
of damage done minus BS is inflicted back on your upper head. You suffer [half BS] Stun. 
\item [Failure:] If Parried, suffer a Swing or Thrust (opponent’s choice, GM adjudicates) to Upper Head with 
dice equal to half of Defense successes from Parrying weapon. If Blocked, suffer a Shield Bash to Upper 
Head with dice equal to half of Defense successes from shield used. 
\item [Special:] If head is protected with a helmet (chain coifs do not count) the Head receives no damage or stun 
back upon it from this attack, and the Headbutt instead inflicts [STR+0+BS] Damage. 
\end{description}
\paragraph{\large\label{man:Hook Punch}Hook Punch}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed, Punching 
\item [Requirements:] Have a hand with which to strike. This can be done while holding a weapon, at GM's 
discretion. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Punch TN 6 at Target Zone on the Swinging 
Attack Table. 
\item [Success:] Inflicts Unarmed damage equal to [STR-1+BS] to Hit Location. Half of damage done minus BS 
is inflicted back on hand. If you strike a hard surface (metal armor, the upper head, a shield) you receive 
full damage back upon your hand. 
\item [Failure:] If Parried, suffer a Swing or Thrust, (opponent’s choice, GM adjudicates) to punching arm with 
dice equal to half of Defense successes from Parrying weapon or shield. 
\item [Special:] If hand is armored in metal, or has brass knuckles/knuckledusters, add +2 to damage, and suffer 
no damage back upon hand. If you have Superior Hook Punch, you inflict +1 additional Shock as well. 
If aiming a Hook Punch below the waist, Activation Cost increases by 1. 
\item [Superior:] Your punches now inflict an additional +1 damage, and inflict 2 Shock. 
\end{description}
\paragraph{\large\label{man:Kick}Kick}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed 
\item [Requirements:] Have an available leg to Kick. 
\item [Maneuver:] Unarmed Attack at S Reach with X dice at Kick TN 8 at Target Zone on the Thrusting or 
Swinging Attack Table. 
\item [Success:] Inflicts Unarmed damage equal to [STR+0+BS] to Hit Location. Opponent must make a stability 
test at BS or be rendered Prone. 
\item [Failure:] If Parried, suffer a Swing or Thrust (opponent’s choice, GM adjudicates) to kicking leg with 
dice equal to half of Defense successes from Parrying weapon or shield. 
\item [Special:] If aiming a Kick above the waist, Activation Cost increases by 1. If wearing sabatons or steel 
boots, add +1 to damage. 
When rolling an Initiative test to determine attack order while making a Thrusting Kick, you may roll 1 
additional die in the test. 
\item [Whip Kick (1-2):] When declaring a Swinging Kick,, you may spend up to 2 CP to increase Kick damage 
by an amount equal to the number of CP spent. 
If the Target is currently prone, a Thrusting kick inflicts an additional +2 damage. 
\item [Superior:] Your Kicks now inflict an additional +2 damage. 
\end{description}
\begin{quotation}
\emph{Gaston has just discharged his blunderbuss into a dubious brigand, but he has no melee weapon with which to defend himself from yon brigand’s comrade! He levels a Kick at the blaggard with 14 dice, and his attack succeeds with 3 BS. Gaston inflicts negligible damage to the brigandine armored brigand, but the Opponent must make an RS 3 stability test or be rendered prone. The brigand fails, and is flung backwards onto the ground, giving Gaston enough time to pour a fistfull of silverware down the barrel of his blunderbuss, and\ldots}
\end{quotation}

\paragraph{\large\label{man:Knee}Knee}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed 
\item [Requirements:] Have an available leg to Knee. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Knee TN 7 at Target Zone on the Thrusting 
Attack Table. 
\item [Success:] Inflicts Unarmed damage equal to [STR+1+BS] to Hit Location. 
\item [Failure:] If Parried, suffer a Swing or Thrust (opponent’s choice, GM adjudicates) to kneeing leg with 
dice equal to half of Defense successes from Parrying weapon or shield. 
\item [Special:] If aiming a Knee above the waist, Activation Cost increases by 2 unless Knee is targeting Target 
Zone currently Trapped in a Clinch. If wearing metal knee-armor, add +1 to damage. 
\item [Superior:] Your knees now inflict an additional +2 damage. 
\end{description}

\paragraph{\large\label{man:One-Two Punch}One-Two Punch}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description}
\item [Type:] Advanced, Attack, Unarmed, Punching 
\item [Requirements:] Have two hands with which to strike. This can be done while holding a weapon or 
weapons, at GM’s discretion. 
\item [Maneuver:] Activate this Maneuver after resolving a Hook Punch or a Straight Punch, Pay 2 CP, and 
immediately declare a second Maneuver of the same type as the previous to the same Target Zone as the 
first with X. X cannot be more than the dice devoted to the original Maneuver. This new Punch resolves 
in the same Action, at the same Initiative, as the first Punch. The Opponent’s defense to the original 
Punch (if any) is used against this second Punch without rolling, at the same Successes as the original. 
The target may declare a Quick Defense to add dice to the original defense, but if this is done, all dice 
must be re-rolled at the new TN for Quick Defense. 
\item [Special:] You may declare another One-Two Punch after resolving a One-Two Punch, but you must pay 
the costs again. 
\end{description}
 

\paragraph{\large\label{man:Shoulder Charge}Shoulder Charge}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed 
\item [Requirements:] Be at Medium range or further to your target. 
\item [Maneuver:] Unarmed Attack at HA Reach, at Shoulder Charge TN 6 at Target Zone on the Thrusting 
Attack Table. 
\item [Success:] Inflicts Unarmed damage equal to [STR-1+BS], and [BS] Stun. Opponent must make a stability 
test at BS or be rendered prone. 
\item [Failure:] If Parried, suffer a Swing or Thrust, (opponent’s choice, GM adjudicates) to neck with dice 
equal to half of Defense successes from Parrying weapon or shield. 
\item [Special:] If wearing pauldrons, spaulders, or other plate shoulder protection, add +1 damage. 
\end{description}
 

\paragraph{\large\label{man:Straight Punch}Straight Punch}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Unarmed, Punching 
\item [Requirements:] Have an available hand to Punch. This can be done while holding a weapon, at the GM's 
discretion. 
\item [Maneuver:] Unarmed Attack at H Reach with X dice at Punch TN 5 at Target Zone on Thrusting Attack 
Table. 
\item [Success:] Inflicts Unarmed damage equal to [STR-2+BS] to Hit Location. Half of damage done minus BS 
is inflicted back on hand. If you strike a hard surface (metal armor, the upper head, a shield) you receive 
full damage back upon your hand. 
\item [Failure:] If Parried, suffer a Swing or Thrust, (opponent’s choice, GM adjudicates) to punching arm with 
dice equal to half of Defense successes from Parrying weapon. 
\item [Special:] If hand is armored in metal, or has brass knuckles/knuckledusters, add +2 to damage, and suffer 
no damage back upon hand. If you have Superior Straight Punch, you inflict +1 additional damage, and 1 
additional Shock as well. 
When rolling an Initiative test to determine attack order while making a Straight Punch, you may roll 1 
additional die in the test. 
If aiming a Straight Punch below the waist, Activation Cost increases by 1. 
\item [Superior:] Your punches now inflict an additional +1 damage, and inflict 1 Shock. 
\end{description}
\subsection{Shield Attack Maneuvers}
\paragraph{\large\label{man:Shield Bash}Shield Bash}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Bash 
\item [Requirements:] Have a shield. 
\item [Maneuver:] Bashing Shield Attack with X dice at H Reach at Shield Bash TN at Target Zone on 
Thrusting Attack Table. 
\item [Success:] Inflicts bludgeoning damage equal to [STR+Shield Bash Damage+BS] to Hit Location. 
\item [Special:] Weapon Defense Maneuvers made against a Shield Bash are made at +1 TN. Shield Bash cannot 
be performed as part of any Simultaneous Maneuver, Off-Hand Maneuver or Double Attack. 
\item [Superior:] Successful Shield Bash additionally forces a stability test at [BS] RS or be knocked prone. 
\end{description}
\begin{quotation}
\emph{Dietrich, having just spit a Bohemian rebel on his sidesword, finds an officer in full armor approaching him from an unseen 
quarter. His sword won’t be able to penetrate that armor, so he declares a Shield Bash to the face on the Hussite rascal with his 
rodela with 8 dice. The Hussite is in full armor, so his Void costs would be high. He attempts to Parry with 10 dice. Dietrich 
rolls 5 successes, while the Hussite rolls 5, but because Parry Successes are halved against Shield Bash, Dietrich succeeds with 5 
BS, and inflicts 11 damage, inflicting a level 1 wound to the Hussite through his helmet. That might give him the leg-up he needs 
to take this guy down.}
\end{quotation}

\paragraph{\large\label{man:Shield Beat}Shield Beat}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Bash 
\item [Requirements:] Have a shield.
\item [Maneuver:] Bashing Shield Attack at Shield Bash TN targeting Opponent shield or weapon. 
Success (Shield): Target shield does not grant AV for this or the next Action, and cannot be used to Block 
for the next Action. You do not gain AV from your shield in the next Action. 
Success (Weapon): Target weapon cannot be used as part of any Maneuvers in the next Action. You 
cannot Block with your shield in the next Action. This does not change the Reach of combat. 
\item [Superior:] By paying 2 additional CP, you may target both an Opponent’s weapon and shield with this 
Maneuver with the same Shield Beat, and both effects occur if the Maneuver is successful. 
\end{description}
\paragraph{\large\label{man:Shield Feint}Shield Feint}---\quad{\large[1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack 
\item [Requirements:] Have a shield, make an attack with another weapon. 
\item [Maneuver:] Spend Activation Cost to change declared attack from current one to a Shield Bash 
Maneuver, aimed at the same or a new location, using the same dice. Pay CP for Reach costs, if 
necessary. You may activate this Maneuver after an Opponent has declared a defense against your other 
attack. 
\end{description}
\begin{quotation}
\emph{Gallant declares a Swing with his arming sword to his Opponent’s neck with 6 dice. His Opponent, who is well-armored, 
declares a Parry with 8 dice. Gallant activates Shield Feint, switching his Swing to a Shield Bash with 6 dice to the face. He has 
to pay an additional 2 dice for Reach, but his Opponent’s defensive successes are halved, because of Parry’s penalties against 
Shield Bash.}
\end{quotation}

\subsection{Special Attack Maneuvers}
\paragraph{\large\label{man:Blind Toss}Blind Toss}---\quad{\large[x+Variable]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack 
\item [Requirements:] Have something to throw in-hand or within quick and easy reach (hat, scarf, cloak, sand, 
dirt, etc) must be at EL Reach or shorter. 
\item [Maneuver:] Attack at TN 5, ignores all Reach modifiers. 
\item [Success:] Opponent's ATN and DTN is increased by 1 per BS to a maximum of 10 until the end of the 
next Action phase. 
\item [Special:] This Maneuver’s Activation Cost is equal to X+half of the target's PER, +1 for each time you 
have used this Maneuver on that character before. This is not limited to one fight—every single time you 
use this Maneuver on a character, your Activation Cost for it increases by 1 against that character. 
This Maneuver cannot be used as part of any sort of Simultaneous Maneuver. 
If this man %this was missing in the main document
This Maneuver cannot be Parried, but may be Blocked or Voided. 
When rolling an Initiative test to determine attack order while making a Blind Toss, you may roll 3 
additional dice in the test.
\end{description}
\begin{quotation}
\emph{Gallant’s saber has just been knocked out of his hand by his intrepid foe, Goofus. Desperate, and having just won the Initiative 
back with a Duck, Gallant flings a fistful of sand at his foe with 5 dice. Goofus ignores it, and tries to launch a Swing past it. 
Gallant scores 3 successes, moving Goofus’s Swing TN to 10. Goofus scores no successes, having been caught in the face with a 
fistful of sand. He wipes the stuff from his eyes, and opens them, enraged, to the sight of Gallant pulling back the hammer of his 
raised pistol with a thumb. Alas, poor Goofus\ldots}
\end{quotation}


\paragraph{\large\label{man:Strangle}Strangle}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Clinching 
\item [Requirements:] Are behind an Opponent, either through stealth, Outmaneuver, and so on. Have two arms 
(hands optional), a wire/cord, some sort of bar, or any other potential strangling device. 
\item [Maneuver:] Clinching Attack with X dice at Strangle TN 8. 
\item [Success:] Immediately enter the Clinch, with your Opponent’s head trapped, and resolve a Strangle 
Maneuver (using whatever tool you initiated the Strangle with) with BS as automatic successes. You may 
maintain the Strangle as normal in the coming rounds. 
\item [Special:] If using a wire or a bar, Strangle TN is 7.
\end{description}

\section{Defense Maneuvers}
\subsection{Parry Defense Maneuvers}
\paragraph{\large\label{man:Disarm (Defense)}Disarm (Defense)}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Parrying
\item [Requirements:] Have a weapon, device or anything with a Defense TN. 
\item [Maneuver:] Defense at X with Defense TN. 
\item [Success:] Opponent attack negated. Opponent must make an AGI test with [2+BS] RS. If this test is failed, 
the targeted weapon is removed from the Opponent's grasp, and either retained by you, dropped at your 
feet, or thrown several yards away at your discretion. 
\item [Failure:] Opponent attack is not reduced by Successes from this Maneuver, and hits with all Successes as BS.
\end{description}
\paragraph{\large\label{man:Parry}Parry}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Parrying 
\item [Requirements:] Have a weapon, device or anything with a Defense TN. 
\item [Maneuver:] Defense with X at Defense TN. If you are using this Maneuver without Initiative, declare it 
against an Opponent attack. 
\item [Success:] Opponent attack negated. Gain Initiative if any BS are scored. 
\item [Special:] You may only declare this Maneuver once per Action with a weapon. If you have two weapons, 
you may declare a Parry and an Off-Hand Parry simultaneously, against two separate attacks. 
Optional [Handguards]: If you successfully defend against a Swinging or Thrusting Weapon Attack 
Maneuver made by an Opponent, but do not have any BS, your opponent resolves a Swing or Thrust at 0 
BS against the hand holding the weapon you used to Parry (if the weapon was two-handed, or if you are 
Half-Swording, roll odd/even to decide which hand is hit). The Guard value of the weapon reduces this 
damage as normal, unless you are Half-Swording, in which case the forward hand is not protected. The 
Parrying party still gains Initiative, and Reach does not change despite the attack being resolved. It counts 
in all ways as the Parrying party being successful, except that the extra hit to the hand resolves. Multi-Hit 
does not apply here. 
\end{description}

\paragraph{\large\label{man:Quick Defense}Quick Defense}---\quad{\large[2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Instant 
\item [Requirements:] You may declare this Maneuver in response to having any attack declared at you, 
regardless of whether or not you have already declared Maneuvers for this Action. You must have some 
means of mounting a Defense available. 
\item [Maneuver:] Declare any Defense Maneuver in response to the attack, paying for that Maneuver as 
normal, OR alter the X value of a Defense Maneuver you have already declared which has not yet 
resolved by adding CP to it on a 1 for 1 basis. 
\item [Special:] You may declare any number of Quick Defenses in an Action, against any number of attacks. 
However, the Defensive Maneuvers chosen cannot be used more times than normally allowed. 
\end{description}

\paragraph{\large\label{man:Riposte}Riposte}---\quad{\large[2+X]}
\vspace{-10pt}\begin{description}
\item [Advanced:] 1H Sword, 2H Sword, Polearm, Spear, 1H Blunt, 2H Blunt %this is not following the standard format
\item [Requirement:] Have a weapon with a Parry TN and a Strike or Cut TN. 
\item [Maneuver:] Parry-Like Defense with X at Weapon Parry TN. 
If you are using this maneuver with Initiative, wait until characters without Initiative have declared their 
Defenses to declare the target of this defense. If nobody declares attacks against you, you can return the 
Defensive Dice to your Combat Pool. Then, resolve maneuvers at Initiative. 
If you are using this maneuver without Initiative, declare it against an enemy attack. 
\item [Success:] Negates enemy attack and gains initiative. You may follow through Riposte next Action by 
declaring a Strike-Like or Thrust-Like Maneuver on enemy whose attack was negated by this maneuver. 
If you do so, your attack gains bonus dice equal to enemy successes on Countered attack. 
Superior Riposte: You may, in addition to a Strike or a Thrust, follow through Riposte with a Grapple, 
Half-Sword, or Hook (paid for normally) gaining bonus dice equal to enemy successes on the Countered 
attack as normal. 
\item [Special:] You may only declare this maneuver once per Action with a weapon. 
\end{description}

\subsection{Unarmed Defense Maneuvers}
\paragraph{\large\label{man:Arm Parry}Arm Parry}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Parrying 
Requirement: Have a free arm that extends (at least) to the forearm (prosthetics can be used for this) with 
which to Parry. This arm can not be holding or wearing a shield. 
\item [Maneuver:] Unarmed Defense with X at Arm Parry TN 7. If you are using this Maneuver without 
Initiative, declare it against an Opponent attack. 
\item [Success:] Opponent Attack negated. If Opponent attack was a Swing or Thrust, or Swinging/Thrusting, it 
inflicts a Swing or Thrust to Parrying limb with 0 BS. Total damage is reduced by Arm Parry BS. 
\item [Failure:] Opponent attack resolves with damage reduced by Successes. If Opponent attack was Swing or 
Thrust, or Swinging/Thrusting, it inflicts a Swing or Thrust to the lower arm Target Zone of the Parrying 
limb with 0 BS in addition to the Opponent’s attack. 
\item [Special:] If Arm Parry is used against an attack targeting the limb used for Arm Parry, the Arm Parry TN 
increases to 8. 
\item [Superior:] Your Arm Parry TN is improved by 1, to a base TN of 6. If you Parry an Unarmed Attack that 
allows a return Strike/Thrust on Parry, use an Elbow instead. This Elbow does not benefit from the effects 
of Superior. 
\end{description}

\paragraph{\large\label{man:Clinch (Defense)}Clinch (Defense)}---\quad{[X]}
Refer to \nameref{man:Clinch (Attack)}, as the Maneuver is mechanically the same.

\paragraph{\large\label{man:Missile Catch}Missile Catch}---\quad{\large[2+X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense 
\item [Requirements:] Either you or someone adjacent to you is being attacked with a thrown missile weapon, 
and you have at least one unused hand. 
\item [Maneuver:] Defense at Catch TN 9 with X. If you are using this Maneuver without Initiative, declare it 
against an Opponent attack. 
\item [Success:] Opponent missile attack is negated, AND you grab the missile out of the air, and can (if the 
weapon does not require some sort of throwing device) throw it back next Action, or simply discard it 
immediately. 
\item [Failure:] If the missile was aimed at you, it hits, and catching limb also suffers a hit from the missile at 
half BS. If the missile was aimed at someone else, the attack simply resolves. 
\end{description}
\subsection{Block Defense Maneuvers}
\paragraph{\large\label{man:Block}Block}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Blocking 
\item [Requirements:] A shield or other item with a Block TN. 
\item [Maneuver:] Block with X dice at Block TN. If you are using this Maneuver without Initiative, declare it 
against an Opponent attack. 
\item [Success:] Opponent attack negated. 
\item [Special:] You may declare this Maneuver any number of times per Action, if there are multiple incoming 
attacks to defend against. 
\end{description}

\paragraph{\large\label{man:Shield Bind}Shield Bind}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Defense, Blocking 
\item [Requirements:] Have a shield, and Opponent is using weapon of M Reach or longer. 
\item [Maneuver:] Blocking Defense at Block TN. 
\item [Success:] Opponent attack negated. Attacking weapon cannot be used as part of any Maneuvers in the next 
Action. You cannot Block with your shield in the next Action. 
\item [Superior:] When you successfully Shield Bind, you may advance a number of Reach steps towards your 
opponent up to BS. 
\end{description}
\paragraph{\large\label{man:Total Block}Total Block}---\quad{\large[Half Maximum CP]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Blocking 
\item [Requirements:] Have a Shield. Only usable in first Action of a Round. If it is First Action, only usable if 
you declared Defensive. 
\item [Maneuver:] Enemy Attack resolves as normal. If attack causes damage (to body only, not weapon or 
shield) add Shield AV to defense of the area struck. 
Lower Shield Durability by opponent’s successes. 
\item [Special:] This Block cannot be used in conjunction with any sort of Simultaneous Attack/Defense. If you 
declare Total Block you cannot declare any attack during this Action for any reason. 
\end{description}
\subsection{Void Defense Maneuvers}
\paragraph{\large\label{man:Closing Void}Closing Void}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Void 
\item [Requirements:] None. 
\item [Maneuver:] Void Defense at TN 8 with X. 
\item [Success:] Opponent attack negated. You close a number of Reach steps equal to BS. You may choose to 
close fewer steps than your BS. 
\item [Failure:] Opponent attack is not reduced by Successes from this Maneuver, and hits with all Successes as 
BS. 
\item [Special:] You may declare this Maneuver any number of times per Action, if there are multiple incoming 
attacks to defend against. 
Fatigue Rules Only: You may choose to gain [2+ENC Penalty] Fatigue when performing this Maneuver 
(multiply this as you would normal Fatigue!). If you do, the TN of the Maneuver is reduced by 1, to a 
base of 7. Additional factors (Talents, Schools, and so on) may reduce this further. 
\end{description}
\paragraph{\large\label{man:Dart-Out}Dart-Out}---\quad{\large[X+ENC Penalty]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Void 
\item [Requirements:] None. 
\item [Maneuver:] Void Defense at TN 8 with X. 
\item [Success:] Opponent attack negated. You retreat a number of Reach steps equal to BS, until you are at 
chosen Reach. 
\item [Failure:] You must make a stability test at [1+Opponent BS] RS or become prone. 
\item [Special:] If you increase the Reach between you and an Opponent to LL Reach, and still have BS left, and 
no additional enemies are targeting you who are closer than LL Reach or Outflanking, you may choose to 
leave the Bout. 
Fatigue Rules only: You may choose to gain [2+ENC Penalty] Fatigue when performing this Maneuver 
(multiply this as you would normal Fatigue!). If you do, the TN of the Maneuver is reduced by 1, to a 
base of 7. Additional factors (Talents, Schools, and so on) may reduce this further. 
\end{description}
\paragraph{\large\label{man:Flee}Flee}---\quad{\large[X+ENC Penalty]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Void 
\item [Requirements:] Have not declared an Attack in either this or the previous Action. If it is the first Round 
of the Bout, must have chosen Defensive Orientation. 
\item [Maneuver:] Void Defense with X dice at TN 5. You may not declare more dice on this Maneuver than 
your current MOB. 
\item [Success:] Opponent attack negated. You leave the Bout. 
\item [Special:] If it is the first Round, and you have just declared Orientation, you cannot use this Maneuver 
unless you declared Defensive. 
You may declare this Maneuver any number of times per Action, if there are multiple incoming attacks to 
defend against. 
Fatigue Rules Only: Performing this Action causes you to immediately gain [3+ENC Penalty] points of 
Fatigue (multiply this as you would normal Fatigue Gain!) that are applied immediately after the roll. 
\end{description}

\paragraph{\large\label{man:Take Cover}Take Cover}---\quad{\large[X+ENC Penalty]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Void 
\item [Requirements:] The ability to move. This Maneuver can be used defensively against an Opponent attack, or passively to assume a defensive position against future attacks. 
\item [Maneuver:] Void Defense with X at TN 5. 
\item [Success (vs. Maneuver):] You gain Cover AV+BS AV against Opponent missile attack (and all sub-
attacks, for example, from multiple barrels, scatter, and so on) and retain it until you Break Cover. If 
Opponent attack does not defeat Cover AV, it is not considered to have hit you at all. 
\item [Success (Passive):] You gain a full Cover AV (see combat rules for missiles in Chapter XX for details) equal to Cover AV+BS, and retain it until you Break Cover.
\subitem [Failure:] You do not gain Cover AV, but your Successes still reduce Opponent Successes.
\item [Special:] Take Cover immediately ends your actions. If there is nothing at all that you can use to Take Cover, you can instead hit the dirt, and gain the lowest Cover AV (0) on the table below. You become prone as a result of ‘hitting the dirt,’ and may become prone when Taking Cover behind other terrain, depending on context.\par
If you are being targeted by multiple missile attacks, a Take Cover Maneuver applies against all of them, there is no additional cost, and the Maneuver must be declared only once. Apply the Take Cover roll individually to each separate attack.\par
If Riding, you cannot Take Cover normally, but you can hide behind your horse. This is a Ride Test at RS 4, and if successful, the attack hits the horse instead of you (resolve an attack against the horse. If the attack is successful, resolve it against yourself as normal using the Cover AV of a horse as listed below.) Breaking Cover: When you cease to be prone or move out from behind the Cover, you lose the Cover AV bonus. At the GM’s discretion, performing certain Maneuvers may also Break Cover (returning fire with a weapon that requires one to stand up, attacking an Opponent, and so on).
\end{description}


\begin{table}[!ht]
	\centering
	\caption{AV of Cover}
	\label{tab:Cover AV}
	\begin{tabu} spread 0pt{XX}
Cover Type& Cover AV\\
Tall grass/hit the dirt& 0\\
Hiding behind Horse& Mount TOU+Mount AV\\
Fence& 6\\
Wooden wall/tree& 12\\
Piled dirt& 18\\
Stone wall& 24\\
	\end{tabu}
\end{table}

\paragraph{\large\label{man:Void}Void}---\quad{\large[X+ENC Penalty]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense, Void 
\item [Requirements:] None. 
\item [Maneuver:] Void Defense with X dice at TN 8. 
\item [Success:] Opponent attack negated. You do not take Initiative with this Maneuver without 2 or more BS. 
\item [Special:] You may declare this Maneuver any number of times per Action, if there are multiple incoming 
attacks to defend against. 
\item [Fatigue Rules only:] You may choose to gain [2+ENC Penalty] Fatigue when performing this Maneuver 
(multiply this as you would normal Fatigue!). If you do, the TN of the Maneuver is reduced by 1, to a 
base of 7. Additional factors (Talents, Schools, and so on) may reduce this further. 
\end{description}
\begin{quotation}
\emph{Gallant is being assaulted by a psychotic carpenter with a sledgehammer. The carpenter is making a Swing with 8 dice. Seeing 
an attack incoming, and having no weapon with which to Parry, he declares a Void with 7 dice. Not liking his odds, but noticing 
that he is unencumbered, Gallant elects to take 2 Fatigue and reduce the TN of his Void by 1. This gives him the edge he needs to 
avoid having his head smashed in.}
\end{quotation}


\subsection{Special Defense Maneuvers}
\paragraph{\large\label{man:Pass (Do Nothing)}Pass (Do Nothing)}---\quad{\large[None]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Defense
\item [Requirements:] None. You may ALWAYS choose to Pass. 
\item [Maneuver:] You do nothing this Action. As you are undergoing no activity, you count as idle for purposes of Fatigue, meaning you may actually regain your breath in this time. If you are attacked and choose to perform a Quick Defense, you do not gain the Fatigue bonus from this Maneuver.
\end{description}
\section{Cross-Fighting Maneuvers}
\subsection{Attack Cross-Fighting Maneuvers}
\paragraph{\large\label{man:Cleaving Blow}Cleaving Blow}---\quad{\large[2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Swinging 
\item [Requirements:] Wielding a weapon with a Swing TN, facing multiple opponents that are valid Opponents 
for you. 
\item [Maneuver:] At any time that a Swing is successful and the requirements are met, you may pay the 
Cleaving Blow Activation Cost to immediately (as same Action) declare and resolve a Swing Maneuver 
against new Opponent with BS. If the new Opponent is at the same Reach as the previous Opponent, you 
do not need to pay Reach costs. If the Opponent is at a different Reach, you must pay the difference 
between the previous target and the new target’s Reach. 
You may add additional dice to this Maneuver at a 2:1 ratio (2 CP spent for 1 CP added). You may 
continue doing this until you are out of CP, until you are out of potential targets, or until you have 
activated Cleaving Blow a number of times equal to the Reach value of your weapon (4 times for a M 
Reach weapon, 6 times for a VL Reach weapon, and so on). 
\end{description}

\paragraph{\large\label{man:Target Change}Target Change}---\quad{\large[1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack 
\item [Requirements:] Facing multiple Opponents that are valid targets for you. 
\item [Maneuver:] Spend 1 to change target to a new valid Opponent immediately upon declaring this 
Maneuver. You may then declare Maneuvers normally against this new Opponent. 
\end{description}
\subsection{Defense Cross-Fighting Maneuvers}
\paragraph{\large\label{man:Ally Defense}Ally Defense}---\quad{\large[2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Defense 
\item [Requirements:] Facing multiple Opponents that are valid targets for you, at least one of which is 
launching an attack that is directed at a character other than you. 
\item [Maneuver:] Pay Activation Cost, and then declare a Defense Maneuver against Opponent’s attack aimed 
at a character other than you. Resolve this defense normally. If you succeed in this defense, the Opponent 
may choose to target you. 
\end{description}

\section{Simultaneous Maneuvers}
\subsection{General Simultaneous Maneuvers}
\paragraph{\large\label{man:Double Attack}Double Attack}---\quad{\large[X+Y+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack
\item [Requirements:] Have two weapons.
\item [Maneuver:] Declare a Weapon Attack Maneuver with your primary hand weapon (X), and an off-hand 
Weapon Attack Maneuver with your off-hand weapon (Y). These Maneuvers resolve Simultaneously, and must be defended against separately. 
\item [Special:] If a character has the Ambidextrous Boon, this Maneuver’s Activation Cost is reduced to [X+Y.] 
\end{description}

\paragraph{\large\label{man:Double Shot}Double Shot}---\quad{\large[X+Y+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack 
\item [Requirements:] Have two single-handed shooting missile weapons (Firearm or Crossbow) that are loaded 
and ready to shoot. 
\item [Maneuver:] Declare a Melee Shoot Maneuver with your primary-hand weapon (X) and your off-hand 
weapon (Y). These Maneuvers resolve simultaneously, and must be defended against separately. 
\item [Special:] If a character has the Ambidextrous Boon, this Maneuver’s Activation Cost is reduced to 
[X+Y+1] 
\end{description}
\paragraph{\large\label{man:Simultaneous Block/Attack}Simultaneous Block/Attack}---\quad{\large[X+Y]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Blocking 
\item [Requirements:] Have both a shield and a weapon. 
\item [Maneuver:] Declare a Weapon Attack Maneuver (X) and a Blocking Maneuver with your weapon and 
shield, dividing dice between the two. 
If you are using this Maneuver with Initiative, declare an Attack first, but wait until characters without 
Initiative have declared their Defenses to declare your Block. If nobody declares attacks against you, you 
can return the defense dice to your CP. 
If you are using this Maneuver without Initiative, declare your Block and your Attack at the same time, 
and resolve them at Initiative. 
\end{description}
\paragraph{\large\label{man:Simultaneous Off-Hand Parry/Attack}Simultaneous Off-Hand Parry/Attack}---\quad{\large[X+Y+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Parrying 
\item [Requirements:] Have two arms with which to attack, weapons optional. 
\item [Maneuver:] Declare a Weapon OR Unarmed Attack Maneuver (X) with one of your arms/weapons, and a 
Weapon OR Unarmed Parry Maneuver (Y) with the other arm/weapon. You may declare these in any 
combination. 
If you are using this Maneuver with Initiative, declare an Attack first, but wait until characters without 
Initiative have declared their Defenses to declare your Parry. If nobody declares attacks against you, you 
can return the defense dice to your CP, then, resolve Maneuvers at Initiative. 
If you are using this Maneuver without Initiative, declare your Parry and your Attack at the same time, 
and resolve them at Initiative. 
\item [Special:] If a character has the Ambidextrous Boon, this Maneuver’s Activation Cost is reduced to [X+Y]. 
\end{description}
\paragraph{\large\label{man:Simultaneous Parry/Attack}Simultaneous Parry/Attack}---\quad{\large[X+Y+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Attack, Parrying 
\item [Requirements:] Have a weapon. 
\item [Maneuver:] Declare a Weapon Attack Maneuver (X) and a Parry Maneuver (Y) with the same weapon 
simultaneously, dividing CP between the two. 
If you are using this Maneuver with Initiative, declare an Attack first, but wait until characters without 
Initiative have declared their Defenses to declare your Parry. If nobody declares attacks against you, you 
can return the defense dice to your CP, then, resolve Maneuvers at Initiative. 
If you are using this Maneuver without Initiative, declare your Parry and your Attack at the same time, 
and resolve them at Initiative. 
\item [Superior “Masterstrike”:] If used without Initiative, you may add any BS from your Parry, if successful, 
to your Attack roll. If used with Initiative, and there is no attack incoming after Actions are declared, you 
may add up to half of the Parry dice you set aside to your Attack instead of returning them to your CP. 
\end{description}
\paragraph{\large\label{man:Simultaneous Void/Attack}Simultaneous Void/Attack}---\quad{\large[X+Y+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Void 
\item [Requirements:] None. 
\item [Maneuver:] Declare an Attack Maneuver (X) and a Void Manuever (Y), dividing dice between the two. 

If you are using this Maneuver with Initiative, declare an Attack first, but wait until characters without Initiative have declared their Defenses to declare your Parry. If nobody declares attacks against you, you can return the defense dice to your CP. Then, resolve Maneuvers at Initiative.

If you are using this Maneuver without Initiative, declare your Parry and your Attack at the same time, and resolve them at Initiative. 
\end{description}
\section{Missile Maneuvers}
Missile Maneuvers refer specifically to Maneuvers that can only be used in the Bystander Phase with 
missile weapons. Using missile weapons in close quarters or melee (that is, in a Bout) is covered by 
Maneuvers like Melee Shoot and Fling. 

Only missile Proficiencies can have these Maneuvers, even if they are listed as Universal. 

Remember, a character’s MP is normally equal only to their Proficiency with the missile weapon, but can 
be increased with Talents, and with the Aim Maneuver below. 
\subsection{General Missile Maneuvers}
\paragraph{\large\label{man:Aim}Aim}---\quad{\large[0]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile 
\item [Requirements:] Have missile weapon that is ready to shoot or fling. 
\item [Maneuver:] Spend Action aiming at visible target. No roll is required, failure is only possible if you are 
interrupted by force, injury, or death. 
\item [Success:] Next Action, if you take a Missile Attack with the weapon that is aimed at the target, you gain 
an Aim bonus equal to your PER to your MP for that attack. 
\item [Superior:] Upon taking an Aim action, add 10 yards to the Range increments of your weapon. 
\end{description}
\subsection{Shooting Missile Maneuvers}
\paragraph{\large\label{man:Braced Shot}Braced Shot}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Missile, Shooting 
\item [Requirements:] Have a shooting weapon that can be fired from a rest, and a rest ready to be used. A 
portable gun rest can be used in conjunction with this action, but a fence, table, crenellation, mantlet, 
pavise, and so on, can be used if available. You cannot be running or riding a horse while using this 
Maneuver (there may be very circumstantial exceptions, at the GM’s discretion). 
\item [Maneuver:] Make a Shooting Missile Attack as detailed for that Maneuver, except that instead of making 
the attack with full MP, make the attack with full MP+PER, as though you had performed an Aim 
Maneuver in the previous Action. If you performed an Aim Maneuver in the previous Action, the effects 
are cumulative. 
\item [Special:] This Maneuver cannot be made if you have moved in this phase, and it cannot be made at all on 
horseback or while mounted, barring unusual circumstances (you are riding an elephant, which has a 
carriage with a gun rest, or you are riding a camel with a harness-mounted musket on a pivot). 
\end{description}
\paragraph{\large\label{man:Power Draw}Power Draw}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Advanced, Missile, Shooting 
\item [Requirements:] Bows only. This Maneuver can only be used if your STR is at least 1 point higher than 
the Required STR of the bow being used. 
\item [Maneuver:] Shooting Missile Attack with MP-2-1 per range increment at Missile TN. 
\item [Success:] Inflicts Weapon Damage+Required STR+BS to Target Zone rolled on Missile Hit Table. 
\item [Special:] You cannot perform this Maneuver while moving. 
This attack cannot be Parried. Only Block, Void, and Cover Maneuvers can be made against it. 
Before rolling this attack, you may spend up to 4 MP. If you do so, and the attack hits, you may modify 
your roll on the Missile Hit Table by a number up to the amount of MP spent in this manner. This is in 
addition to other sources of Hit Table modification. 
Special [Din]: The Burdinadin racial characteristic ‘The Stains of Time’ and the Ohanedin characteristic 
‘The Feeling Disappears’ allows them to Void and Parry attacks made by firearms. A Parry made against 
a firearm attack counts as a Break attempt against the parrying weapon with 0 BS. If the base damage is 
sufficient to destroy the weapon, it does so, and the attack hits the defender as normal. 
\item [Superior:] If your target is within the first Range increment of your weapon, you may spend up to 2 MP 
before making the Power Draw to increase damage by an amount equal to MP spent. 
\end{description}
\paragraph{\large\label{man:Rapid Shot}Rapid Shot}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile, Shooting 
\item [Requirements:] Have a bow, or a missile weapon with multiple ‘shots’ (any regular bow, double-barreled 
pistol, revolver, multi-prod crossbow, repeating crossbow). 
\item [Maneuver:] When activating this Maneuver, you may declare multiple Shoot Maneuvers at once. You 
may aim these Maneuvers at multiple targets, or at a single target. Each shot after the first reduces your 
‘full MP’ for all of them by 2. Resolve each shot separately, after all have been declared. You may not 
declare more Shoot Maneuvers than the number allowed by your weapon, or 5 in the case of bows. 
\item [Superior:] Each shot after the first now reduces your total MP by 1 instead of 2. 
\end{description}
\paragraph{\large\label{man:Reload}Reload}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile 
\item [Requirements:] Have a shooting weapon that requires reloading (firearm or crossbow) that is unloaded. 
Have ammunition with which to reload it. 
\item [Maneuver:] Roll [Weapon Proficiency+AGI] at TN determined by Loading Method. 
\item [Success:] The weapon gains a number of Load Points or Span Points equal to Successes. If the weapon 
reaches its Load or Span requirement, it is now loaded and ready to be used. 
\item [Special:] If moving while Reloading, the TN for this Maneuver is increased by 2. Weapons with the 
Cavalry Bow special quality do not reduce this penalty. 
\end{description}
\paragraph{\large\label{man:Shoot}Shoot}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile, Shooting 
\item [Requirements:] Have a Bow, Crossbow or Firearm that is ready to shoot or be loosed. 
\item [Maneuver:] Shooting Missile Attack with MP-1 per Range increment at Missile TN. 
\item [Success:] Inflicts Weapon Damage+BS to Hit Location rolled on Missile Hit Table. 
\item [Special:] If you are Shooting while moving, reduce your MP by 4. If you are on horseback and your 
weapon has the Cavalry Bow special quality, this penalty can be alleviated as detailed. 
This attack cannot be Parried. Only Block, Void, and Cover Maneuvers can be made against it. 
Voids always gain a +2 penalty to Void TN versus Shoot.

If the weapon being used is a firearm, then in addition to Parry, Block and Void Maneuvers cannot 
normally be made against it either (see Special [Din]). 

Before rolling this attack, you may spend up to 4 MP. If you do so, and the attack hits, you may modify 
your roll on the Missile Hit Table by a number up to the amount of MP spent in this manner. This is in 
addition to other sources of Hit Table modification. 
\item [Special (Din):] The Burdinadin racial characteristic ‘The Stains of Time’ and the Ohanedin characteristic 
‘The Feeling Disappears’ allows them to Void and Parry attacks made by firearms. A Parry made against 
a firearm attack counts as a Break attempt against the parrying weapon with 0 BS. If the base damage is 
sufficient to destroy the weapon, it does so, and the attack hits the defender as normal. 
\item [Superior:] When spending MP to adjust the Missile Hit Location of this Maneuver, you may modify your 
roll on the Missile Hit Table by an additional 1 space per 2 MP spent (i.e. if you spend 2 MP to modify 
your roll, you may change it by 3, instead of by 2). 
\end{description}
\subsubsection{Missile Attack Penalties}
\begin{table}[!ht]
	\caption{Missile Attack Penalties}
	\label{tab:Missile Attack Penalties}
	\begin{tabu} spread 0pt {X[5,L]X[-1,r]}
Slightly Obscured Visibility (gunsmoke, scattered people, light fog):& -2 MP \\
Mildly Obscured Visibility (thick smoke, crowd, light forest, fog):& -4 MP \\
Heavily Obscured Visibility (smokescreen, dense crowd, forest, dense fog):& -6 MP \\
	\\
Target is Moving Slowly (walking):& -2 MP \\
Target is Moving Quickly (running):& -4 MP \\
Target is Moving Very Quickly (sprinting):& -6 MP \\
Target is Moving Super Quickly (Horse Sprinting, Falling):& -8 MP \\
	\\
Target is Moving Towards/Away relative to Shooter:& -1 MP \\
Target is Moving Laterally relative to Shooter:& -2 MP \\
	\\
Target is Fighting someone in Melee:& -4 MP \\
Target is Figthing an Ally in Melee:& -6 MP \\
	\end{tabu}
\end{table}

\subsection{Throwing Missile Maneuvers}
\paragraph{\large\label{man:Double Fling}Double Fling}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile, Throwing 
\item [Requirements:] Have Throwing Weapons in either hand that are ready to be thrown. Slings cannot be 
used for this Maneuver. The weapons must be of the same general type (javelins, knives, hammers/axes, 
various rocks or lead weights, and so on). 
\item [Maneuver:] Perform two Fling attacks, each with half MP-1-1 per Range increment at Missile TN. These 
Flings resolve as normal, but with the half-1 MP. 
\item [Special:] If using this Maneuver when moving, reduce MP by 2. 
\item [Superior:] If your target is within the first Range increment of your weapon, you may spend up to 2 MP to 
increase damage by 1 damage per MP spent. 
\end{description}
\paragraph{\large\label{man:Fling}Fling}---\quad{\large[???????]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Missile, Throwing 
\item [Requirements:] Have a Throwing Weapon that is ready to be thrown. 
\item [Maneuver:] Missile Attack with full MP-1 per Range increment at Missile TN. 
\item [Success:] Inflicts [STR+Weapon Damage+BS] to Hit Location rolled on Missile Hit Table. 
This attack cannot be Parried. Only Block, Void, and Cover Maneuvers can be made against it. 
As this attack uses a Throwing Weapon (javelin, thrown sword, axe, hammer, spear, and so on) it can be 
countered with the Missile Catch Maneuver. 
\item [Special:] If using this Maneuver when moving, reduce MP by 2. 
\item [Superior:] If your target is within the first range increment of your weapon, you may spend up to 2 MP to 
increase damage by XX damage per MP spent. 
\end{description}
\section{Grapple Maneuvers}
Any Maneuver Grapple with the Advantage tag may only be defended against using a Maneuver with the 
Disadvantage tag.

\subsection{Advantage Grapple Maneuvers}
\paragraph{\large\label{man:Grappling: Choke}Grappling: Choke}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage, Opponent's head is Trapped or Pinned. 
\item [Maneuver:] Grapple Attack at HA Range at Choke TN 7 with X dice. 
\item [Success:] Opponent suffers [BS] Stun, and begins to suffocate. 
Each Action after the first, Choke can be ‘maintained’ in place of performing another action. Maintaining 
a Choke is the same as performing the Choke action, but at TN 6, against Opponent defense (if any). The 
Opponent takes 1 Pain every Action that Choke is maintained (not the Action it is initiated) along with 
the Stun. 
When a character's Pain from Choke becomes equal to his MOB score, he must make a WIL test each 
Action (at RS equal to 1 per point of Pain over the character's MOB score) or lose consciousness for 1d10 
minutes. If a character is Choked into unconsciousness, they must make a HLT test at RS 2 or die from a 
crushed windpipe. Each Choke action after unconsciousness forces an additional test at +1 RS. 
If a character breaks the Choke and manages to get free, they lose 1 Pain for every 10 seconds they have 
to breathe without exertion but only to a maximum of the Pain inflicted by Choke. 
\item [Special:] If a garrote, like a wire or cord, is used instead of bare hands or arms, the TN for Choke is 6, 
maintaining it is done at TN 5, and the Pain inflicted every Action that it is maintained is 2 instead of 1. If 
a character is garroted into unconsciousness, they must make a HLT test at RS 4 or die after going 
unconscious, from a crushed windpipe. Each Choke action after unconsciousness forces an additional test 
at +1 RS.
If the neck is protected by plate armor like a gorget or bevor, the TN of Choke increases to TN 9, and 
maintaining it increases to TN 8. Garrotes cannot be used over characters with such protection. 
\end{description}

\paragraph{\large\label{man:Grappling: Coup de Grace}Grappling: Coup de Grace}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage, Thrusting 
\item [Requirements:] Grappling, have Advantage, wielding a weapon of S Reach or shorter with an arm that is 
not Trapped, Pinned or Locked. This weapon must inflict piercing damage on the Thrust. 
\item [Maneuver:] Thrusting Weapon Attack at Weapon Reach with X dice, at Thrust TN+1, aimed at Hit 
Location with the Weak Spot quality on Opponent armor. 
\item [Success:] Inflicts piercing damage equal to [Weapon Thrust Damage+STR+BS] to CHOSEN Hit Location 
with the Weak Spot quality. Do not roll on table. The damage from this attack is not reduced by AV from 
any armor which has a Weak Spot for the Hit Location targeted by this attack. All other AV protecting 
that area apply. 
This attack does NOT suffer the increased Activation Cost for Ground Fighting. 
\item [Special:] If the weapon being used is of S Reach or less, make this Attack at -1 TN. If the weapon being 
used has Thin Blade, make this Attack at -1 TN. It is possible to benefit from both S Reach and Thin 
Blade simultaneously when making this Attack. 
\end{description}
\paragraph{\large\label{man:Grappling: Limb-Break}Grappling: Limb-Break}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage, at least one of Opponent's body parts are Trapped. 
\item [Maneuver:] Grapple Attack at HA Reach on Trapped body part at Break TN of body part with X dice 
(Arms: TN 6, Leg: TN 7, Neck: TN 7, Torso: TN 9). 
\item [Success:] Inflict bludgeoning damage (not unarmed--bludgeoning!) equal to [STR-2+Half BS] to area 
chosen from list. This attack ignores all armor (Arms: Elbow, Forearm, Upper Arm, Shoulder; Neck: 
Neck, Lower Head; Leg: Knee, Foot, Shin, Thigh, Hip; Torso: Chest, Shoulder, Side, Belly, Hip, Lower 
Back, Upper Back). 
\item [Special:] If the limb being targeted is Pinned, Break inflicts [STR-2+BS] instead of its original amount, 
making it more effective. 
\item [Superior:] Break damage is increased by 2. 
\end{description}
\paragraph{\large\label{man:Grappling: Pin}Grappling: Pin}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage, at least one of Opponent's body parts is Trapped. 
\item [Maneuver:] Grapple Attack at HA Reach on Trapped body part at Pin TN 8 with X dice. 
\item [Success:] Target body part becomes Pinned. This is identical in effect to being Trapped, except that TNs 
increase by 2, instead of 1. Some Maneuvers have a special bonus if used against a Pinned body part. If 
the limb that is Pinned was being used in a Maneuver, then Pin also inflicts Shock equal to Double BS. 
\item [Special:] If you suffer Overshock while you are Pinning body parts, and you fail your stability test, you 
release them. 
\end{description}
\paragraph{\large\label{man:Grappling: Takedown}Grappling: Takedown}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage. 
\item [Maneuver:] Grapple Attack at HA Reach at Takedown TN 8 with X dice. 
\item [Success:] Grapple goes to ground. While on the ground, all Maneuvers have a +1 Activation Cost, and the 
fighter with Advantage gains a -1 bonus to all Maneuver TNs while retaining it. 
\end{description}
\paragraph{\large\label{man:Grappling: Throw}Grappling: Throw}---\quad{\large[X+2]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage. 
\item [Maneuver:] Grapple Attack at HA Reach at Throw TN 7 with X dice. 
\item [Success:] Opponent is rendered prone, and suffers falling damage as though falling a number of feet equal 
to (5+STR+BS)x2 onto whatever surface they are being thrown upon. (See falling damage in Chapter XX 
for more details).
Grapple ends unless the Takedown Maneuver is used. If the Grapple ends, the Reach of combat is set to 
your choice of Reach from HA to L. 
\item [Special:] If you have Pinned one of the target's body parts and successfully Throw the target, you may 
move the result on the Falling Table by 1 per body part Pinned. 
The +2 Activation Cost for throwing on the ground includes the penalty for being on the ground. 
\item [Superior:] You may re-roll on the Falling Table to see where damage is inflicted. 
\end{description}
\paragraph{\large\label{man:Grappling: Trap}Grappling: Trap}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Advantage 
\item [Requirements:] Grappling, have Advantage. 
\item [Maneuver:] Grapple Attack at HA Reach at Trap TN 7 with X dice, targeting head, right arm, left arm, 
right leg, left leg, or torso. 
\item [Success:] Body part targeted becomes Trapped. All TNs increase by 1 for the victim until it is released. No 
actions can be taken involving the Trapped body-part (except for Force, to try and free the limb, or 
defenses as appropriate) until it is released. If a weapon is being held in a Trapped limb, the weapon itself 
is also considered to be Trapped. If the Limb that is Trapped was being used in a Maneuver, then Trap 
also inflicts Shock equal to BS. 
\item [Special:] If you suffer Overshock while you are Trapping body parts, and you fail your stability test, you 
release them. 
\item [Superior:] Trap TN is reduced to 6. 
\end{description}
\subsection{Disadvantage Grapple Maneuvers}
\paragraph{\large\label{man:Grappling: Slip}Grappling: Slip}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Disadvantage 
\item [Requirements:] Grappling. 
\item [Maneuver:] Grapple Defense at Slip TN 6 with X dice. 
\item [Success:] Opponent Maneuver negated, you gain Advantage. 
\item [Special:] If fight is on the ground when performing a successful Slip, you may spend 2 CP after success to 
attempt to stand (MOB test at terrain difficulty, opponent can oppose by spending 2 CP). Success moves 
the Clinch to standing position. 
If BS exceed 2, you may pay 2 CP to end Clinch, and return both characters to regular combat. If the fight 
is on the ground, both characters are Prone. 
\end{description}

\paragraph{\large\label{man:Grappling: Twist}Grappling: Twist}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Disadvantage 
\item [Requirements:] Grappling. 
\item [Maneuver:] Grapple Defense at Twist TN 7 with X dice. 
\item [Success:] Opponent Maneuver negated, you gain Advantage, you gain CP equal to BS for the next 
Maneuver. 
\item [Special:] If fight is on the ground when performing a successful Twist, you may automatically move to 
standing position. 
If successful, you may end Clinch, and return both characters to regular combat at HA Reach. If the fight 
is on the ground, both characters are Prone. 
\end{description}
\subsection{General Grapple Maneuvers}
\paragraph{\large\label{man:Grappling: Bite}Grappling: Bite}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple, Unarmed 
\item [Requirements:] Grappling. Must have teeth. Cannot have a visor or other armor covering the face. No 
hands required. You do not need to have Advantage to perform this Maneuver, but it resolves after the 
character with Advantage's Maneuver. 
\item [Maneuver:] Unarmed Attack at HA Reach with X dice at Bite TN 8 at Target Zone on the Thrusting 
Attack Table (must aim for head, neck, arm, or any Trapped body part). 
\item [Success:] Inflicts cutting damage equal to [Half STR+Half BS] to Hit Location. 
\item [Special:] Wounds inflicted by Bite have +3 Infection Chance. If Hit Location is protected by Hard armor, 
you suffer Shock equal to Half of the total damage inflicted on the target. 
If Bite is used consecutively between two Actions, you may choose to have it automatically hit the same 
location, this time rolling at TN 6, to represent the biter latching on and savaging with his teeth. 
This attack does not suffer the increased Activation Cost for ground fighting. 
\end{description}

\paragraph{\large\label{man:Grappling: Force}Grappling: Force}---\quad{\large[X+1]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Grapple 
\item [Requirements:] Grappling, have one body part currently Trapped or Pinned. You do not need to have 
Advantage to perform this Maneuver, but it resolves after the character with Advantage's Maneuver, and 
may not undo a Trap or Pin implemented by that Maneuver. 
\item [Maneuver:] Select one of your body parts that is currently Trapped or Pinned. Grappling Attack at HA 
Reach at Force TN 7 (if Trapped) or Force TN 8 (if Pinned) with X dice. 
\item [Success:] Body part is no longer Trapped or Pinned. 
\item [Special:] This Maneuver does not increase this TN any further for Traps or Pins than is detailed in this 
Maneuver, and is immune to the normal cumulative +1 TNs. 
\end{description}

\paragraph{\large\label{man:Grappling: Unarmed Attack}Grappling: Unarmed Attack}---\quad{\large[??????]}
\vspace{-10pt}\begin{description} 
\item [Type:] XXXXXXXX 
\item [Requirements:] XXXXXXXX 
\item [Maneuver:] XXXXXXXX 
\item [Success:] XXXXXXXX 
\item [Special:] XXXXXXXX 
\end{description}
\paragraph{\large\label{man:Grappling: Use Weapon}Grappling: Use Weapon}---\quad{\large[??????]}
\vspace{-10pt}\begin{description} 
\item [Type:] XXXXXXXX 
\item [Requirements:] XXXXXXXX 
\item [Maneuver:] XXXXXXXX 
\item [Success:] XXXXXXXX 
\item [Special:] XXXXXXXX 
\end{description}
\section{Hilt Push Maneuvers}
\subsection{Advantage Hilt Push Maneuvers}
\paragraph{\large\label{man:Hilt Push}Hilt Push}---\quad{\large[2]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Instant 
\item [Requirements:] Activate Hilt Push under any of the following circumstances, in descending order of 
priority. If your opponent does not declare a Hilt Push when they have a higher priority “trigger” for this 
maneuver, you may then do so.
	\begin{itemize}
		\item You have just successfully defended against a Weapon Attack with a Parry Maneuver and both the 
weapon being used and the weapon used to parry are within 2 Reach Stages or fewer of each other. 
		\item Your Weapon Attack has just been defeated by a Parry Maneuver and both the weapon being used and 
the weapon used to parry are within 2 Reach Stages or fewer of each other.
	\end{itemize}
\item [Maneuver:] Both of the weapons involved immediately enter the Hilt Push. 
While in the Hilt Push, a weapon cannot be used for any normal maneuvers, it can only perform 
Maneuvers with the Push tag. However, you may use other Maneuvers that do not involve the weapon in 
the Hilt Push.
	\begin{itemize}
		\item Hilt Push Maneuvers always use the Parry TN of their weapons instead of Thrust TN or Strike TN. 
		\item Hilt Push Maneuvers cannot be defended against by non-Hilt Push Maneuvers, and vice-versa. 
		\item While in the Hilt Push, if the Reach of Combat would ever become longer than the Reach of the longest 
weapon in the Push, it does not. Instead, it stays at the that Reach. 
		\item If either party in a Hilt Push becomes Prone, the Push ends. 
		\item If either party would become Untargetable or Outflanking, the Push ends. 
		\item If either party is Disarmed of or drops their weapon, the Push ends, and the Disarmed party immediately 
suffers Stun equal to half of their maximum CP. 
		\item Hilt Push Maneuvers always resolve simultaneously, but are declared normally. Initiative still passes 
between characters as normal. 
		\item If the end of the Clash Count is reached, the character with Initiative can choose to either end the Push, or continue it. If they choose the former, they may also exit the Bout as normal, or continue it without the Push. If they choose the Latter, obviously they cannot end the Bout.
		\item When the Push ends, the fighters are placed at the Reach of the last Attack that hit (if an Attack ended the Push) or at the Reach of the character who successfully Evaded.
		\item It is possible to Half-Sword while in the Hilt Push, or to enter a Hilt Push while Half-Swording. 
		\item For the purposes of Simultaneous Maneuvers involving two weapons, it is possible to declare a Push Maneuver as the primary or Off-Hand attack, and a different Maneuver with another weapon that is not a Push Maneuver.
	\end{itemize}
\item [Superior:] The Activation Cost for this Maneuver is reduced to 0. 
\item [Hilt Push Maneuvers:] Push Maneuvers count as both attacks and defenses. When declaring a Hilt Push: 
Cut against a Hilt Push: Thrust, they Defend against each other, and the victorious Maneuver only 
resolves at BS over the other, which is canceled. 
In the event of a tie in a Hilt Push, neither Maneuver resolves. 

The following Maneuvers act exactly as their normal versions, but count as separate Maneuvers for the purposes of Talents, School Bonuses, etc. Additionally, they can only defend against/be defended against by other Hilt Push Maneuvers. Hilt Push Maneuvers are all Universal regardless of whether or not the original Maneuver is.

Hilt Push Maneuvers marked with a * end the Push if they are successful. 
	\begin{itemize}
		[noitemsep]
		\item Hilt Push: Cut 
		\item Hilt Push: Draw Cut 
		\item Hilt Push: Thrust 
		\item Hilt Push: Push Cut 
		\item Hilt Push: Joint Thrust 
		\item Hilt Push: Pommel Strike* 
		\item Hilt Push: Disarm* 
		\item Hilt Push: Grapple 
		\item Hilt Push: Butt-Strike* 
		\item Hilt Push: Shoulder Charge 
	\end{itemize}
\end{description}
\paragraph{\large\label{man:Hilt Push: Lever Down}Hilt Push: Lever Down}---\quad{\large[2+X]}
\vspace{-10pt}\begin{description}
\item [Type:]
\item [Requirements:] Be in the Push 
\item [Maneuver:] Push Attack with X at Weapon Parry TN. 
\item [Success:] Opponent suffers Shock equal to BS. You may CHOOSE to immediately end the Push. If you 
do not, and you instead perform a Push Maneuver in the next Action, that Maneuver, if successful, is not 
reduced by Opponent BS, and resolves at full BS. The Maneuver still reduces Opponent BS as normal. 
\end{description}

\paragraph{\large\label{man:Hilt Push: Evasion*}Hilt Push: Evasion*}---\quad{\large[2+Encumbrance+X]}
\vspace{-10pt}\begin{description} 
\item [Type:] 
\item [Requirements:] Be in the Push. 
\item [Maneuver:] Void with X at TN 8. 
\item [Success:] You may CHOOSE to immediately end the Push. 
\item [Failure:] Opponent Attack is not reduced by your Successes with this Maneuver. 
\end{description}
\section{Mobility Maneuvers}
\subsection{General Mobility Maneuvers}
\paragraph{\large\label{man:Obstruction}Obstruction}---\quad{\large[????????]}

XXXXXXXXXXXXXXXXXXXXXX 

\paragraph{\large\label{man:Outmaneuver}Outmaneuver}---\quad{\large[X+2+Encumbrance Penalty]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mobility, Instant 
\item [Requirements:] You and at least one other Character are Targeting an Opponent. Activate this Maneuver 
before declaring an Action. You may not perform this Maneuver while Prone. 
\item [Maneuver:] Roll X+Mobility versus Battlefield Terrain TN. Targeting Opponent rolls Mobility at 
Battlefield TN, and may spend X+Encumbrance Penalty CP to add X dice to his Mobility roll. 
\item [Success:] You are now Outflanking the Target, and the Target Un-Targets you. 
\item [Failure:] You do not Outflank your Opponent. If your Opponent succeeds by 2 or more BS, he may elect 
to Untarget himself from you, while also Untargeting you. 
\item [Special:] Performing this Action increases your Fatigue Gain for this Round by 1. 
When you activate this Maneuver, the other Character (or Characters) targeting your Opponent may elect 
to Activate it as well by spending X+Encumbrance Penalty CP, and resolving identically. This can result 
in multiple successes, or you failing to succeed but another Character succeeding. 
At GM’s discretion, the Battlefield TN can be lower for the defender than the attacker, or vice-versa. I.E., 
if the defending character is in a defensible position like a doorway or a breach in a wall where getting 
past him to outflank him would be more difficult. 
\end{description}
\paragraph{\large\label{man:Rapid Rise}Rapid Rise}---\quad{\large[3$+ (\mathrm{Encumbrance~Penalty}\times2) $]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mobility, Instant 
\item [Requirements:] Must be Prone. Declare this Maneuver before declaring Maneuvers for this Action. You 
may not Activate this Maneuver if you are currently suffering from any Wound that has the Auto 
Knockdown tag. 
\item [Maneuver:] Roll Mobility vs TN 5+Encumbrance Penalty, at RS 1+Terrain Penalty. 
\item [Success:] You are no longer Prone and may declare any other Maneuver for this action. 
\item [Failure:] You take too long to stand up and can only Quick Defend for the action. After the action has 
ended, and if you have not been affected by a Stability test or Auto-prone, you are no longer Prone.
\end{description}
\paragraph{\large\label{man:Slip-About}Slip-About}---\quad{\large[X+Encumbrance Penalty+Special]}
\vspace{-10pt}\begin{description} \item [Type:] Universal, Mobility, Instant 
\item [Requirements:] None. Declare this Maneuver before declaring Maneuvers for this Action. 
\item [Maneuver:] Pay CP equal to Opponent Mobility. Roll X$ + $Mobility vs Battlefield Terrain TN. Opponent rolls Mobility at Battlefield TN. 
\item [Success:] You are now Outflanking the Target, and the Target Un-Targets you. This condition does not 
require an ally, and ends at the end of the Action. 
\item [Special:] Performing this Action increases your Fatigue Gain for this Round by 1. 
At GM’s discretion, the Battlefield TN can be lower for the defender than the attacker, or vice-versa. i.e. 
if the defending character is in a defensible position like a doorway or a breach in a wall where getting 
past him to outflank him would be more difficult. 
\end{description}
\paragraph{\large\label{man:Steal Initiative}Steal Initiative}---\quad{\large[X+Opponent Per]}
\vspace{-10pt}\begin{description} 
\item [Type:] XXXXXX 
\item [Requirements:] Don't have Initiative. 
\item [Maneuver:] Roll an Initiative Contest against target, with all normal modifiers. You may spend X extra 
CP, and add 1 die to your Initiative Roll for every 3 extra CP spent. 
\item [Success:] Gain Initiative. 
\item [Tie:] You and Opponent resolve maneuvers simultaneously as-per Offense/Offense tie. 
The Activation Cost of Steal Initiative is equal to the target's modified Perception score, plus optional X. 
\end{description}
\paragraph{\large\label{man:Thread the Needle}Thread the Needle}---\quad{\large[X+Y+Encumbrance Penalty]}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Mobility, Instant 
\item [Requirements:] Multiple Adversaries. Activate this Maneuver before Actions are declared in a Round. 
You may not perform this Maneuver while Prone. 
\item [Maneuver:] Pay 2 CP per Opponent who is currently Targeting you (Y) and then roll X+Current Mobility 
versus Battlefield Terrain TN. All Targeting Opponents roll Mobility at TN+1, but may spend 2+X CP to 
roll Mobility+X at this difficulty. Compare Successes individually to your Opponents. 
\item [Success:] You may choose to Un-Target yourself from Targeting Opponent. If you do, you cannot 
continue Targeting this Opponent. 
\item [Failure:] Targeting Opponent retains you as their Target. 
\item [Special:] Performing this Action increases your Fatigue Gain for this Round by 1. 
\end{description}
\section{Mounted Maneuvers}
\subsection{Attack Mounted Maneuvers}
\paragraph{\large\label{man:Lance Charge}Lance Charge}---\quad{\large[X+2] (RS 3 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Mounted 
\item [Requirements:] Charging into a Bout while mounted, have a weapon with the Couched Charge quality. 
\item [Maneuver:] Attack at Weapon Reach with X, rolled at Weapon Thrust TN, aimed on the Thrust Table. 
\item [Success:] Inflicts damage equal to [Mount Strength+Weapon Thrust Damage+BS] of Weapon Thrust 
Damage Type to Thrust Location. 
\item [Failure:] If this maneuver is Blocked, it is not negated and inflicts damage as-per Success, but 
automatically adds Shield AV to the attack, regardless of its ultimate hit location. 
\item [Special:] When rolling an Initiative Test to determine attack order while making a Thrust, you may roll 2 
additional dice in the Test for the Thrust. 
Lance Charge ignores the first 4 successes from any Weapon Defense made against it. 
\item [Ride Fail:] Your mount rears in shock at the last minute and refuses to charge home. You enter the Bout, 
but you lose all CP devoted to Lance Charge.
\end{description}

\paragraph{\large\label{man:Mount Bite}Mount Bite}---\quad{\large[2] (RS 4 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Attack, Mounted 
\item [Requirements:] In a Bout, Mounted, Mount has a Bite attack. 
\item [Maneuver:] Mount makes Unarmed Attack at Medium Reach with Horse’s Full CP at Bite TN 6 aimed on 
the Thrusting Table to either Head, Neck, or Arm. You do not pay Reach Costs for this Attack, the Horse 
does. 
\item [Success:] Inflict damage equal to [Half Mount Strength+Mount Bite Damage+BS] in Cutting Damage to 
Thrusting Hit Location. 
[Ride Fail:] You cannot convince your mount to bite anything this Action.
\end{description}

\paragraph{\large\label{man:Mount Kick}Mount Kick}---\quad{\large[2] (RS 3 Ride Check)}
\vspace{-10pt}\begin{description}
\item [Type:] Universal, Attack, Mounted 
\item [Requirements:] In a Bout, Mounted, Mount has a Kick attack. 
\item [Maneuver:] Mount makes Unarmed Attack at Long Reach with Full CP at Kick TN 7 aimed on the Thrust 
Table. 
\item [Success:] Inflict damage equal to [Mount Strength+Mount Kick Damage+BS] in Bludgeoning Damage to 
Thrust Location. 
\item [Ride Fail:] You cannot convince your mount to kick anything this Action.
\end{description}
\subsection{Defense Mounted Maneuvers}
\paragraph{\large\label{man:Disengage}Disengage}---\quad{\large[??????]}
\vspace{-10pt}\begin{description} 
\item [Type:] XXXXXX 
\item [Requirements:] XXXXXX 
\item [Maneuver:] XXXXXX 
\item [Success:] XXXXXX 
\end{description}

\paragraph{\large\label{man:Lean}Lean}---\quad{\large[X+Encumbrance Level] (RS 2 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mounted, Void 
\item [Requirements:] Mounted, you are being attacked above the Waist. 
\item [Maneuver:] Void at Lean TN 8 with X. Make a Ride Check. 
\item [Success:] Opponent attack negated. You may pay 1 CP to gain Initiative. 
\item [Special:] Remember, Armor Weight is halved while Mounted for purposes of Encumbrance. 
\item [Ride Fail:] Lean still resolves, but you potentially fall from the saddle! Make a Knockdown test at RS 3 
or suffer Falling Damage equal to 6 feet onto ground, and be rendered Prone and Dismounted.
Fatigue Rules Only: You may choose to gain 2+Encumbrance Penalty Fatigue when performing this 
maneuver (Multiply this as you would normal Fatigue!) If you do, the TN of the maneuver is reduced by 
1, to a base of 7. Additional factors (Talents, Schools, etc) may reduce this further.
\end{description}
\paragraph{\large\label{man:Mount Dodge}Mount Dodge}---\quad{\large[2] (RS 2 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mounted, Void 
\item [Requirements:] Mounted, either your Horse or your own body beneath the Waist is being attacked. 
\item [Maneuver:] If Successful, Horse rolls full CP to Void at Horse Dodge TN 7 against Opponent Attack. 
\item [Success:] Opponent Attack negated. You gain Initiative. 
\item [Ride Fail:] You fail to convince your horse to swerve away! Opponent attack resolves as normal.
\end{description}
\subsection{Passing Mounted Maneuvers}
\paragraph{\large\label{man:Ride-By}Ride-By}---\quad{\large[2] (RS 3 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mounted 
\item [Requirements:] Charging into a Bout while mounted. Declare this Maneuver in addition to any other 
Maneuver. It is always resolved after all other Maneuvers, and can be used in combination with any 
except for other Passing Mounted Maneuvers. 
\item [Maneuver:] Make a Ride Check after resolving Maneuvers for this Action Phase. 
\item [Success:] You leave the Bout immediately, and re-enter the Bystander Phase. You may not act in (or be 
targeted in) the next Action. 
\item [Special:] If you successfully make a Weapon Attack in conjunction with a Ride-By, you must make a 
Stability test at RS 2, or lose your weapon. If the weapon has the Cavalry Weapon quality, no Stability 
test is necessary and you automatically retain your weapon. If the weapon is connected to your character 
by some means (a lanyard or a chain, etc) the weapon is dropped, but can be retrieved as though Drawing 
the weapon, or Quickdraw. 
\item [Ride Fail:] Your horse panics, and accidentally rams the thing you were trying to ride past. Resolve a 
Horse Crash as detailed in the Mounted Combat section.
\end{description}
\paragraph{\large\label{man:Trample}Trample}---\quad{\large[4] (RS 4 Ride Check)}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Mounted 
\item [Requirements:] Charging into a Bout while mounted, declare this maneuver in addition to any other 
Maneuver. It is always resolved after all other Maneuvers, and can be used in combination with any 
except for other Passing Mounted Maneuvers. 
\item [Maneuver:] Make a Ride Check after resolving Maneuvers for this Action Phase. 
Success (Ride): Mount makes Unarmed Attack at Half-Arm Reach with Full CP at Trample TN 6 on 
target. 
Success (Trample): Inflicts Trampling Damage equal to [Mount Strength+Mount Weight+BS] must make 
Stability test equal to 3+BS or be rendered prone. 
\item [Ride Fail:] Your Horse does not successfully overrun the target, and misses it. If Critical Failure, Resolve 
a Horse Crash as detailed in the Mounted Combat section!
\end{description}

\section{Dog Maneuvers}
\subsection{Attack Dog Maneuvers}
\paragraph{\large\label{man:Dog Bite}Dog Bite}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Dog 
\item [Requirements:] Be Dog (or Wolf, Hyena, or similar animal using Dog rules) and have teeth not constrained by a muzzle. 
\item [Maneuver:] Unarmed Attack at Half-Hand Reach with X at Dog Bite TN 6 aimed on Thrusting Table. 
\item [Success:] Inflict damage equal to [Str-1+BS] in Cutting Damage to Thrusting Hit Location. If there is 
Hard Armor on the area, this attack only inflicts [Str-2+Half BS] damage, and inflicts Bludgeoning 
Damage instead.

If you hit a limb (Arm, Leg) or the Throat, Head or Groin, you may spend 2 CP (if you inflicted either a 
Cutting or Bludgeoning Wound with this Attack, there is no CP cost) to Grab the Hit Location Bitten. If 
you do so, Target must make a Stability test at RS 2+BS or be rendered Prone. 

While Grabbed by a Dog, performing any Maneuver that involves a Grabbed Limb has an Activation 
Cost of 4, and suffers a TN increase of 2. This includes Void and Mobility Maneuvers if the Leg/Legs are 
grabbed, and Attacks involving a Grabbed Arm. A Dog can use special Maneuvers on a Grabbed Victim. 
A Dog remains Grabbing the affected character until any of the following circumstances: 
\begin{itemize}
\item The Dog is rendered Prone by Overshock. 
\item The Dog makes another Dog Bite attack. 
\item The Dog makes any sort of Defensive Maneuver. 
\item The Dog voluntarily releases the victim. 
\item The Dog (or someone else) rips the Victim’s limb off. 
\end{itemize}
\end{description}
\paragraph{\large\label{man:Savage}Savage}---\quad{\large[X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Dog 
\item [Requirements:] Be Dog (or Wolf, Hyena, or similar animal using Dog rules) and have teeth not 
constrained by a muzzle. Be currently Grabbing a Target. 
\item [Maneuver:] Unarmed Attack at Half-Hand Reach with X at Savage TN 4, targeting Grabbed Hit 
Location. This Maneuver cannot be defended against. 
\item [Success:] Inflict damage equal to [Str-1+BS] in Cutting Damage to Grabbed Hit Location. If there is Hard 
Armor on the area, this attack inflicts [Str-4+BS] in Bludgeoning Damage instead. Target must make a 
Stability Test at RS 2+BS or be rendered Prone. If already Prone, Target takes [BS] Stun. 
This Attack does not end a Grab. 
\end{description}
\paragraph{\large\label{man:Tackle}Tackle}---\quad{\large[2+X]}
\vspace{-10pt}\begin{description} 
\item [Type:] Universal, Dog 
\item [Requirements:] Be Dog (or Wolf, Hyena, or similar animal using Dog rules). 
\item [Maneuver:] Unarmed Attack at Half-Hand Reach with X at Furious Dog Tackle TN 5. 
\item [Success:] Opponent suffers [BS] Stun, and must make a Stability Test at RS 2+BS or be rendered Prone 
and suffer Falling Damage as though having fallen Feet equal to [5+Dog Strength+BS.] 
\end{description}

\chapter{Talents}\label{ch:Talents} %TALENTS NEED DOING
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
Talents are small improvements to a character’s combat skills that are acquired as the character’s Proficiency score increases. At set intervals in a Proficiency’s progression, a character earns Talents of Novice, Journeyman, or Master rank. These Talents may be chosen from the Talent list at the end of this chapter

\section{Gaining Talents}
A character gets a Talent whenever a Proficiency or a School reaches a level in which a Talent is 
rewarded (see \autoref{tab:Proficiency Progression}~for details.)

Proficiencies that are a part of a School do not gain Talents. Instead, the School as a whole gains only one 
set of Talents, but these Talents apply to all Proficiencies within the School, and can use any Proficiency 
in the School to satisfy their requirements. 

\subsection{Choosing Talents}
When your character receives a Talent from a Proficiency level that they’ve just acquired, you may choose any Talent under the Talent rank they have gained (Novice, Journeyman or Master) or lower, provided the character meets the requirements of the Talent. You may take a Novice Talent in a Journeyman Talent slot, but you may not take a Journeyman level Talent in a Novice slot.

Some Talents have Attribute requirements, which your character must match or exceed in order to take the Talent. Others have Proficiency requirements, meaning that they can only be taken as Talents for that Proficiency or a School containing that Proficiency.

Talents cannot be taken multiple times.

\section{Using Talents}
Any time a Talent would be applicable for activation, simply announce that it is in effect or not in effect.
If it is simply an automatic improvement to a Maneuver or similar, assume that it is in effect unless you say otherwise. 

Unless the Talent is chosen for a School, Talents can not be used for Proficiencies they are not linked to, or when Defaulting from a Proficiency. 

\section*{Talent List}
\section{Novice Talents}\label{sec:Novice Talents}
\paragraph{\label{talent:Accuracy}Accuracy}
	\begin{description}
		\item [Requirements:] Universal, AGI 4. 
		\item [Description:] You have a natural knack for hitting in the right spot. It's not always conscious, but your 
attacks just seem to find their way to your intended target. 
		\item [Effect:] Whenever you hit with a melee attack that rolls for Hit Location on a Target Zone, you may modify the Hit Location roll by 1. If your opponent has Rapid Reaction, the effects of both Talents cancel 
out, and the attack lands normally. 
Additionally, whenever you perform a Joint Thrust Maneuver, add +1 CP to that attack. 
	\end{description}
\paragraph{\label{talent:Boar Piercer}Boar Piercer}
	\begin{description}
		\item [Requirements:] STR 5, Throwing 
		\item [Description:] Either you’ve got a thing for pork, or you just really like throwing spears at things. You can 
throw a spear like nobody’s business, especially in combat. 
		\item [Effect:] When making a Melee Throw or Fling with a Spear or any variety of Javelin, you inflict +1 
damage. 
	\end{description}
\paragraph{\label{talent:Bonk!}Bonk!}
	\begin{description}
		\item [Requirements:] Str 6, 1H or 2H Blunt 
		\item [Description:] See Title 
		\item [Effect:] Whenever you Power Attack with a Blunt weapon, you inflict additional Stun with a successful attack equal to the CP spent to increase damage. 
	\end{description}

\paragraph{\label{talent:Flourishing Drills}Flourishing Drills}
	\begin{description}
		\item [Requirements:] Dagger, Sword or Blunt Proficiency (1H and 2H varieties are both eligible). 
		\item [Description:] You have practiced a series of tight, circular cutting or striking drills to the point that you can perform them in your sleep. In battle, you are a whirlwind of rapid strikes, each snapping into another faster than the eye can see. 
		\item [Effect:] Whenever you perform a Swinging Maneuver, if in the previous Action you also performed a Swinging Maneuver, the Maneuver rolls 2 additional dice. 
	\end{description}
	
\paragraph{\label{talent:Good Form}Good Form}
	\begin{description}
		\item [Requirements:] Noble School, or high-quality tutelage. 
		\item [Description:] You have been drilled in the performance of a specific maneuver, and can now execute it 
flawlessly, every time. This is the hallmark of high-quality training in swordsmanship, and it requires 
more than just practice, but the careful guidance of more experienced masters who can correct 
inadequacies as they form. 
		\item [Effect:] Choose one Maneuver. Reduce that Maneuver’s Non-[X] and Non-Range Activation Cost by 1. 

	\end{description}
\paragraph{\label{talent:Helm-Splitter}Helm-Splitter}
	\begin{description}
		\item [Requirements:] Universal. 
		\item [Description:] Maybe you've spent a lot of time practicing for helmet-cutting competitions, or maybe 
you've just chopped a lot of wood. You can hit things REALLY hard, as long as you’re swinging straight 
down. 
		\item [Effect:] If you use the Power Attack function of Strike, you may spend 1 more CP to increase damage 
than normal. 

	\end{description}
\paragraph{\label{talent:Shield-Breaker}Shield-Breaker}
	\begin{description}
		\item [Requirements:] 1H or 2H Blunt. 
		\item [Description:] You have a knack for smashing shields, and have practiced at sundering them to deprive 
your enemies of their defenses. 
		\item [Effect:] You inflict +2 damage when making Hew attacks. When passively inflicting damage to a Shield, 
increase that damage by 1. 

	\end{description}
\paragraph{\label{talent:Shield Trick}Shield Trick}
	\begin{description}
		\item [Requirements:] Shield, STR 4. 
		\item [Description:] When using a shield, you know how to hide your weapon behind it so that your movements 
are concealed until the moment you strike. 
		\item [Effect:] After making a successful Block, Shield Bash, or other shield-related Maneuver, if you perform a 
Swinging or Thrusting Maneuver with the weapon in your other hand in the next Action, you gain +1 dice 
to the attack. This includes Melee Shoot.

	\end{description}
\paragraph{\label{talent:Slowpoke}Slowpoke}
	\begin{description}
		\item [Requirements:] Spear or Polearm 
		\item [Description:] Getting past the range of a spear is hard. You know how to make it even harder.
		\item [Effect:] When using a Spear or Polearm, when your opponent makes an attack during which he would 
have to pay a Range Cost, increase the total Range Cost by 1. 

	\end{description}
\paragraph{\label{talent:Impaler}Impaler}
	\begin{description}
		\item [Requirements:] Spear 
		\item [Description:] Like ol’ Uncle Vlad, you’ve got a thing for impaling people. 
		\item [Effect:] If you inflict a Level 5 Wound with a any sort of Thrusting attack, you regain 2 CP. 

	\end{description}
\paragraph{\label{talent:Special Move}Special Move}
	\begin{description}
		\item [Requirements:] Universal. 
		\item [Description:] You have a very unconventional way of performing a certain move. Those who haven't seen 
it before are put off by the speed and sureness with which you conduct this strange Maneuver, and, at 
least the first time, it's much harder to defend against. 
		\item [Effect:] Choose a Maneuver from the following list: Weapon Throw, Blind Toss, Quickdraw, or Melee Shoot. When you use this Maneuver against an individual for the first time, you perform a Special Move 
version of that Maneuver. A Special Move has no Activation Cost (this includes Range) and its TN is 
reduced by 1. If a character witnesses you using a Special Move against someone else, on a PER test at 
RS 3, he can count as having had it used against him before, and as such you can never gain the bonus 
against him. 

	\end{description}
\paragraph{\label{talent:Rapid Reaction}Rapid Reaction}
	\begin{description}
		\item [Requirements:] Universal, AGI 4. 
		\item [Description:] While you might not always be able to move out of the way, you sure know how to take a hit better than most. 
		\item [Effect:] When you are hit by a melee attack in combat, you may alter the roll to see which location is hit by 1 in either direction. If your opponent has Accuracy, the effects of both Talents cancel out, and the 
attack lands normally. 


	\end{description}
\paragraph{Specialist}\label{talent:Specialist}
	\begin{description}
		\item [Requirements:] Universal.
		\item [Description:] One Maneuver is just more important to you than the rest. You’ve drilled it for countless hours, and can perform it in your sleep. 
		\item [Effect:] Choose one Advanced Maneuver. Reduce that Maneuver’s Non-[X] and Non-Range Activation Cost by 1.
	\end{description}
\paragraph{Smash Punch}\label{talent:Smash Punch}
	\begin{description}
		\item [Requirements:] Universal 
		\item [Description:] You can deliver a monstrous uppercut. This technique favors hook-heavy infighters. 
		\item [Effect:] When making a Hook Punch, you may elect to perform a Smash. If you do, you inflict +2 
damage, and your Hit Location is automatically the Lower Head. However, this attack suffers a -1 penalty 
to Initiative Rolls. 
	\end{description}
\paragraph{Head Guard}\label{talent:Head Guard}
	\begin{description}
		\item [Requirements:] Universal. 
		\item [Description:] You’ve got a good handle on protecting your head. The head is the most commonly 
armored part of the human body, and for good reason: it’s the obvious place to attack with any sort of 
weapon, and it’s not far from arm-level, making it convenient as well. 
		\item [Effect:] Whenever Parrying or making a Parry Maneuver to defend an attack targeting your Head, Face, or 
Neck, you may add 2 dice to the Parry (this does not let you Parry if you have no CP remaining). 

	\end{description}
\paragraph{Infighter}\label{talent:Infighter}
	\begin{description}
		\item [Requirements:] Universal. 
		\item [Description:] You’re not afraid to get up close and personal. 
		\item [Effect:] When fighting an enemy inside your weapon's reach, your penalty to attacks is 1 less than it would be normally. When attacking an enemy within their weapon’s reach, their penalty to attacks and 
defenses is 1 higher than it would be normally. Multiple people with Infighter cannot stack, and if both a 
combatant and his current target have Infighter, it cancels out. 

	\end{description}
\paragraph{Counter Attack}\label{talent:Counter Attack}
	\begin{description}
		\item [Requirements:] Universal, AGI 4. 
		\item [Description:] You know what they say about the best defense... 
		\item [Effect:] If you attempt to Steal Initiative to launch an attack against an opponent who is already attacking you, you gain a +1 dice bonus to the Initiative test.
	\end{description}

\section{Journeyman Talents}\label{sec:Journeyman Talents}
\paragraph{\label{talent:Art of Eight Limbs}Art of Eight Limbs}
	\begin{description}
		\item [Requirements:] Pugilism, Wit 6 
		\item [Description:] How is it even possible to attack from a position like that!? 
		\item [Effect:] You can Feint from any Unarmed Maneuver into any other Unarmed Maneuver. 
	\end{description}
	
\paragraph{\label{talent:Bear Piercer}Bear Piercer}
	\begin{description}
		\item [Requirements:] STR 5, Spear, Boar Piercer 
		\item [Description:] You’re really good at throwing spears. REALLY good. 
		\item [Effect:] When making a Melee Throw or Fling with a Spear or any variety of Javelin, any target you hit must make a Stability Test at RS 2+BS or be rendered prone. 
	\end{description}
	
\paragraph{\label{talent:Grappling at the Sword}Grappling at the Sword}
	\begin{description}
		\item [Requirements:] Sword or Blunt proficiency (1H or 2H varieties are both eligible). 
		\item [Description:] You know how to use a sword in the grapple, using the weapon to gain leverage and lock joints, and also how to avoid the weapon getting in your way during a struggle on the ground. 
		\item [Effect:] You may use an M Reach weapon in a Grapple without dropping it. This includes longer weapons that are being shortened with the Halfsword Maneuver. You may also ignore the Reach Costs for making 
attacks or defenses using weapons in the Grapple.

	\end{description}
\paragraph{\label{talent:Swift Sword}Swift Sword}
	\begin{description}
		\item [Requirements:] Sword proficiency (1H or 2H varieties are both eligible), AGI 5. 
		\item [Description:] You’ve got a swift hand when it comes to swords. 
		\item [Effect:] You gain a +1 bonus to Initiative contests when using a 1H or 2H Sword. 

	\end{description}
\paragraph{\label{talent:Roll With It}Roll With It}
	\begin{description}
		\item [Requirements:] Universal, AGI 5. 
		\item [Description:] It's natural to cringe from potential injury, but your reactions are much more practical, and you have a natural understanding of how force transfers from one body to another. You instinctively roll 
away from blows and other sources of harm, and this tends to lessen their effects on you. 
		\item [Effect:] Once per Round, if you are struck by an attack (missile attacks, including those made in melee, 
are not eligible) that would inflict damage, and you are not currently Prone, you may immediately become 
Prone to reduce the damage inflicted by your AGI. This has all the normal effects of being made Prone. 

	\end{description}
\paragraph{\label{talent:Ground Wrestler}Ground Wrestler}
	\begin{description}
		\item [Requirements:] Wrestling. 
		\item [Description:] You’re more practiced in grappling on the ground than standing. Many fighters fear this 
kind of fight for how it can go disastrously wrong, but you revel in it. 
		\item [Effect:] When in a Clinch, and the fight has gone to the ground, you do not suffer the associated 
Activation Cost penalty for Clinch Maneuvers, and you add 1.5 your Strength instead of 1.0 to your CP 
(i.e., when you go to Ground, if you have 4 Strength normally, you gain a bonus of 6 to your CP as long 
as you are on the Ground.) 

	\end{description}
\paragraph{\label{talent:Stepping Parry}Stepping Parry}
	\begin{description}
		\item [Requirements:] MOB 7. 
		\item [Description:] You have learned to step close, to take the initiative, and most importantly, how to do so 
through clever manipulation of the opponent’s weapon, and its position relative to you. 
		\item [Effect:] Add the following section to your Parry or Parrying Maneuver with one Proficiency. 
		\item [Special:] On success, you may change the range between yourself and your opponent by 1 for every 2 BS.

	\end{description}
\paragraph{\label{talent:Kebab Master}Kebab Master}
	\begin{description}
		\item [Requirements:] Impaler Talent, Spear 
		\item [Description:] You know a thing or two about skewering things. 
		\item [Effect:] If you inflict a Piercing Wound with a Thrusting attack, increase the Level of the Wound by 1. 

	\end{description}
\paragraph{\label{talent:Rapid Strike}Rapid Strike}
	\begin{description}
		\item [Requirements:] Universal, AGI 5. 
		\item [Description:] You're very skilled at attacking with weapons in both hands, striking with one while the other returns to an ideal position to strike from, and repeating this over and over to strike fast, hard, and often.
		\item [Effect:] Whenever you make a Double Attack, you gain a +1 CP bonus to the Off-Hand attack. Each consecutive Double Attack increases this bonus by +1, to a maximum of 4. 

	\end{description}
\paragraph{\label{talent:Spartan Sole}Spartan Sole}
	\begin{description}
		\item [Requirements:] Pugilism, Str 5 
		\item [Description:] Talk to the Sandal, ‘cause the shield don’t care. 
		\item [Effect:] Whenever you force a Stability Test with a Kick to knock your opponent down, increase the RS of that Test by 2. 

	\end{description}
\paragraph{\label{talent:Slippery}Slippery}
	\begin{description}
		\item [Requirements:] Universal, AGI 6. 
		\item [Description:] You know how to move. In combat you're fast, fluid, and can change direction on a dime. 
Your enemies often find themselves looking at nothing—and nothing is the last thing they see. 
		\item [Effect:] As long as you are Unencumbered, you gain a -1 TN when performing the Outmaneuver and Thread the Needle Maneuvers.

	\end{description}
\paragraph{\label{talent:Quick Reload}Quick Reload}
	\begin{description}
		\item [Requirements:] Pistol or Arquebus/Musket, AGI 5. 
		\item [Description:] A great deal of practice and some finely-tuned muscle memory have made reloading a breech-loading firearm into one quick, fluid motion. 
		\item [Effect:] When performing the Reload Maneuver, you gain an additional automatic success for every two successes you roll. This may cause your Reload’s total successes exceed the actual number of dice you rolled. 

	\end{description}
\paragraph{\label{talent:Spear-Sunderer}Spear-Sunderer}
	\begin{description}
		\item [Requirements:] Universal, STR 5. 
		\item [Description:] Cleaving through the shafts of spears, axes, polearms and the like isn’t easy, but it is certainly possible.
		\item [Effect:] You inflict +2 damage when making Break attacks. 

	\end{description}
\paragraph{\label{talent:Straight Blast}Straight Blast}
	\begin{description}
		\item [Requirements:] Pugilism, AGI 6. 
		\item [Description:] Commonly associated with Wing Chun and certain styles of boxing, you have mastered a rapid, piston-like punching style that can wear down any defense. Combined with a strong center and 
good footwork, this technique can be devastating. 
		\item [Effect:] When making a One-Two Punch with a Straight Punch into another Straight Punch, you may reduce the Activation Cost of the Maneuver chained from the first by 1. 

	\end{description}
\paragraph{\label{talent:Rolling Hooks}Rolling Hooks}
	\begin{description}
		\item [Requirements:] Pugilism, STR 6. 
		\item [Description:] To the untrained eye, a feral and apelike series of haymakers. But to the trained boxer, this technique is a monstrous culmination of brute strength, skill, and pugilist rhythm. 
		\item [Effect:] When making a One-Two Punch with a Hook Punch into another Hook Punch, you may reduce the Activation Cost of the Maneuver chained from the first by 1. 

	\end{description}
\paragraph{\label{talent:Long Jab}Long Jab}
	\begin{description}
		\item [Requirements:] Pugilism, INT 5. 
		\item [Description:] You are a more measured fighter, preferring to fight at longer ranges, peppering opponents with strikes intended to wear them down, and only striking them when they present an opening. 
		\item [Effect:] Your Straight Punch Maneuvers (not the Hook Punch or One-Two Punch) are made at +1 Reach.

	\end{description}
\paragraph{\label{talent:Ironwall}Ironwall}
	\begin{description}
		\item [Requirements:] Universal, END 5. 
		\item [Description:] You’re hard to dislodge from your feet. It takes a singular force to put you on the ground.
		\item [Effect:] You gain +2 bonus dice to Stability tests made to avoid falling over or being shoved, pushed, or dislodged from your position. 

	\end{description}
\paragraph{\label{talent:Shield Charge}Shield Charge}
	\begin{description}
		\item [Requirements:] Shield, STR 5, MOB 8, Superior Shield Bash. 
		\item [Description:] When slamming a shield properly into someone’s face, you know how to throw your shoulder into it, potentially flinging them from their feet. 
		\item [Effect:] When you force a Stability test through a Superior Shield Bash, add 1 to the Stability test RS. 

	\end{description}
\paragraph{\label{talent:Rim Bash}Rim Bash}
	\begin{description}
		\item [Requirements:] Shield, STR 5, MOB 8, Superior Shield Bash, Shield Charge 
		\item [Description:] You know just the right spot to get that extra bit of Oomph out of a shield bash. 
		\item [Effect:] When you force a Stability test through a Superior Shield Bash, add 1 to the Stability test RS. 
This is cumulative with Shield Charge. 
	\end{description}
\section{Master Talents}\label{sec:Master Talents}
\paragraph{\label{talent:It Was On Purpose!}It Was On Purpose!}
	\begin{description}
		\item [Requirements:] Missile Weapon, WIT 6.
		\item [Description:] No really, it was totally intentional! You seem to have a skill for putting small missiles 
through narrow slits and chinks in armor. You’re every knight’s worst nightmare. 
		\item [Effect:] Whenever a Shoot, Melee Shoot or Fling attack you make hits an opponent on a Hit Location 
where the target’s armor has a Weak Point, the attack inflicts +2 damage. This does not apply to Scatter 
damage, except for the primary shot from the weapon. 

	\end{description}
\paragraph{\label{talent:Whirlwind Defense}Whirlwind Defense}
	\begin{description}
		\item [Requirements:] Spear, Polearm, 2H Sword or 2H Blunt, AGI 6 
		\item [Description:] You can move your weapon in a whirlwind of defensive patterns, intercepting enemy attacks as you do. 
		\item [Effect:] There is no limit to the number of Parries you can make with a weapon in one Action. After any successful Parry, you regain 2 CP. 

	\end{description}
\paragraph{\label{talent:Bad War}Bad War}
	\begin{description}
		\item [Requirements:] 2H Sword or Polearm, STR 6, AGI 6. 
		\item [Description:] It’s a rare man who can survive in the Bad War. When two pike formations meet, and the world turns into an infinity of spear points, halberds, katzbalgers and screaming, dying men, they’ll be glad they brought you along. 
		\item [Effect:] You may use Cleaving Blow with Breaks in addition to Swings. Each time you activate Cleaving Blow, you regain 2 CP. 

	\end{description}
\paragraph{\label{talent:Stern}Stern}
	\begin{description}
		\item [Requirements:] Universal, HLT 6. 
		\item [Description:] Few people can match your physical resolve. 
		\item [Effect:] Once per Combat, at any Refresh, you may choose to ignore all Pain, and regain your CP as normal. Next Refresh, the Pain returns in full. 

	\end{description}
\paragraph{\label{talent:Superior Void}Superior Void}
	\begin{description}
		\item [Requirements:] Universal, AGI 6. 
		\item [Description:] You don’t even remember what it feels like to fear being hit. 
		\item [Effect:] You perform all Void maneuvers at -1 TN. 

	\end{description}
\paragraph{\label{talent:Juggernaut}Juggernaut}
	\begin{description}
		\item [Requirements:] Universal, END 6. 
		\item [Description:] ‘Even if one’s head were to be cut off, he should be able to perform one more action with certainty.’ 
		\item [Effect:] Once per Session, you may elect to delay the full effect of any Wound until the end of the Action. 
However, if the Wound inflicted has a “Limb Severed” or “Death” result, halve the CP of any Maneuvers 
you currently have declared, but then resolve them as normal. 

	\end{description}
\paragraph{\label{talent:Jarring Blows}Jarring Blows}
	\begin{description}
		\item [Requirements:] Universal, STR 6. 
		\item [Description:] You really know how to hit people, and can throw enemies off their feet with powerful blows. 
		\item [Effect:] Whenever any Maneuver you perform would cause an opponent to make a Stability test or become Prone (including Overshock) you increase the RS of that test by 2. 

	\end{description}
\paragraph{\label{talent:Bruiser}Bruiser}
	\begin{description}
		\item [Requirements:] Universal. 
		\item [Description:] A veteran of many battles, you know where to strike to offset a foe. 
		\item [Effect:] Any Wounds you cause with Unarmed or Melee Weapon Attacks inflict +1 Stun per level of the Wound. 

	\end{description}
\paragraph{\label{talent:Flicker Jab}Flicker Jab}
	\begin{description}
		\item [Requirements:] Pugilism. AGI 8 
		\item [Description:] Your jabs are so rapid, and come from such unorthodox angles, that most opponents cannot even perceive them. 
		\item [Effect:] If you declare a Straight Punch with fewer than 5 CP, your opponent must make a Perception Test at RS 3 in order to perform any sort of Defense against it. You may not in any way increase the CP of the Straight Punch once this Perception Test is made.
	\end{description}
\paragraph{\label{talent:The Spear that Pierced the Sky}The Spear that Pierced the Sky}
	\begin{description}
		\item [Requirements:] Spear, Kebab Master Talent 
		\item [Description:] Your capabilities with the spear are compelling proof that God is a lancer. 
		\item [Effect:] Range Costs paid by opponents making attacks against you are doubled. 
	\end{description}

\chapter{Adventuring}\label{ch:adventuring} %ADVENTURING CHAPTER
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
%Environment, Terrain, Temperature, Lighting missing from the original document

\section{Encumbrance}\label{sec:Encumbrance}
A character’s Encumbrance is the degree to which they are hindered by the equipment, baggage and 
armor that they’re wearing and carrying. This is as much tied to the awkwardness of the things being carried as it is to the weight.

A character’s Encumbrance Level is determined by their Carry Score, and the total Weight they are 
currently carrying. Weight is determined by the following:
	\begin{itemize}
		\item Stowed/Sheathed/Holstered Weapons (Weapons currently being wielded do not add to Encumbrance, because they are generally very light and handy when wielded. It is the awkwardness of carrying them on your person that contributes Weight. Imagine trying to do anything with a 15 foot pike strapped to your back!)
		\item Worn or carried Armor
		\item Miscellaneous Gear/Equipment/Items carried in pockets, satchels, and backpacks. Generally, 20 lbs in a backpack or 10 lbs in pockets/over the shoulder should equal 1 Weight. Superior carrying equipment, like a large framed backpack or a harness may further decrease these numbers, situationally.
	\end{itemize}
There are five Levels of Encumbrance. Unencumbered, Light, Medium, Heavy, and Overloaded. 
Each Encumbrance Level has a different effect on a character’s Combat Pool, Mobility, certain Skill Checks, and also on Exhaustion.

Skills that suffer additional difficulty from Encumbrance have a special tag, as shown in the Skills section.

\begin{table}[hb]
	\centering
	\caption{Encumbrance Level}
	\label{tab:Encumbrance}
	\begin{tabu} spread 0pt {X[1.5]X*{5}{X[c]}}
\rowfont[c]{}Weight& Encumbrance Level &CP Effect &Mobility Effect &Skill Penalty &Exhaustion Speed &Recover Speed\\\toprule
	Less than CARRY &Unencumbered &0 &0 &0 &$ \times1 $ &$ 1\times $\\
	Equal to CARRY &Light &-1 &-2 &+1 RS &$ \times1.5 $ &$ 1\times $\\
	2$ \times $ CARRY& Medium &-2 &-4 &+2 RS &$ \times2 $ &$ \tfrac{1}{2}\times $\\
	3$ \times $ CARRY& Heavy &-3 &-6 &+3 RS &$ \times3 $ &$\tfrac{1}{4}\times $\\
	4$ \times $ CARRY& Overloaded &-4 &-8 &+4 RS &$ \times3.5 $ &None\\
	\\
	\end{tabu}
	\caption*{At 5x Carry and beyond, the character is too heavily encumbered to move, and cannot engage in combat or make any sort of skill or attribute tests that involves moving. Reducing the load is the only way to continue moving.}
\end{table}

\section{\label{sec:exhaustion}Exhaustion}
\subsection{Exhaustion \& Fatigue}
As characters exerts themselves physically, either in combat or other strenuous activities like running, jumping or severe lifting, fatigue will begin to set in.

-side bar-
[Fatigue rules add another layer of realism to combat, but also another layer of complexity. These should be considered Optional Rules, and can be safely ignored in campaigns with new players, those more interested in heroic action, and those who wish to keep bookkeeping to a minimum.]

\subsubsection{Fatigue Points}
At the end of each Round, while in combat, all characters in the combat who acted in the previous Round gain one Fatigue point. Encumbrance and Environmental effects can increase this amount (being in Medium Encumbrance, for example, causes a character to gain double the normal number of Fatigue).

Additionally, if a character has a visor, helmet etc with the Constricts Breathing special rule, add an additional 1x to Fatigue Gain per point of the quality (i.e. Constricts Breathing 3 adds 3x to the Fatigue Gain.). If the visor/whatever is raised, then it has no effect. Once a character reaches the amount of Fatigue to move to the next level of Exhaustion, they suffer penalties until they’ve had a chance to rest.

Additionally, any Wound inflicts Fatigue equal to the level of the Wound, regardless of what sort of 
Wound it is. Some Wounds will inflict additional Fatigue, and \nameref{sec:Blood Loss}, with its ability to reduce 
Health, can also effect Fatigue. 

\begin{table}[hb]
	\centering
	\caption{Fatigue Level}
	\label{tab:Fatigue Level}
	\begin{tabu} spread 0pt {X[-1,r]@{\ to\ }X[-1,l]X*{3}{X[-1,c]}}
\rowfont[c]{}\multicolumn{2}{l}{Fatigue Level} &Exhaustion Level &CP Penalty &Mobility Penalty &Skill Penalty\\\toprule
(0&5)\hfill~$+$~HLT&Fresh &0 &0 &0\\
(6&10)\hfill~$+$~HLT&Winded&-1 &-1 &-1\\
(11&15)\hfill~$+$~HLT&Tired &-2 &-2 &-2\\
(16&20)\hfill~$+$~HLT&Very Tired &-4 &-4 &-4\\
(21&25)\hfill~$+$~HLT&Exhausted &-6 &-6 &-6\\
	\end{tabu}
\end{table}

\paragraph{Resting}---\quad
A character who takes no Actions during a Bout Cycle (or who only takes Actions that do not cause 
Fatigue) loses Fatigue equal to his Endurance score, modified by his Encumbrance level (see Recovery 
Speed). Full-Face Helmets and Visors halve a character’s Endurance for this purpose, unless lifted or
removed (see \nameref{par:Movable Visors}~in \autoref{par:Movable Visors}~for more details.)
Sitting down and having a drink adds 2 to a character’s effective Endurance (after Encumbrance) for the 
purposes of Resting. 


\subsubsection{Wearing Armor Long-Term}
Fighting in armor is one thing, but living in armor is quite another. Wearing armor for long periods of 
time, such as while traveling, performing everyday tasks, or especially while sleeping, istiresome and 
uncomfortable. 
If an individual has already been wearing armor for more than 4 hours, the armor’s effective Weight 
increases by half until he has had a chance to take it off and rest. 
If an individual sleeps in armor, the armor’s effective Weight doubles for the next day. 
Generally people who expected to be wearing armor all the time wore lighter and more comfortable stuff, 
so that as the day dragged on it would not reduce their performance as much. 


[Sidebar: Comfortable Armoring] 
Wealthier aristocrats and warriors who had heavier armor usually wouldn’t just walk around in it, they 
would be armored by servants and squires once it was clear that battle was inevitable. Some sorts of 
armor, like mail and lamellar, had the advantage that they were light and could be worn all the time with 
little problem. 
[End Sidebar] 


\chapter{Combat}\label{ch:combat}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{quote}
	\centering
		\emph{“If you’re an easy fright,\\
			You shouldn’t learn to fight.”}\par
		\hfill--\textbf{Johannes Lichtenhauer}
\end{quote}
Let’s face it. The game’s called Song of Swords; you’re going to end up in combat sooner or later. The Song of Swords’ combat system is detailed, but straight-forward; combat is intense, lethal, and fast. A
single good blow from a weapon can put anyone down, even the mightiest hero. You won’t be racing to
deplete an enemy’s hit points until they reach zero and die -- you’re inflicting Wounds. Injuries and sheer
exhaustion seriously deteriorate a character’s ability to fight.
\section{Combat Encounters}
\subsection{Rounds}
A Round represents about six seconds of fighting. During the Round, characters will act once or more
depending on their Initiative. At the end of the Round, if anyone’s still alive, a new Round starts and the
fighting continues. Combat ends when the fighting stops. Simple enough, right?
\subsubsection{3-Phase Combat Round}
Each Round has three Phases: 1, 2, and 3. Every character can act in Phase 3, but only characters with
higher ADR (as shown in the table below) can act in Phases 2 and 1. This represents the speed of the
fighters involved. A fast man can get a lot more done in a short time than a slow one. Characters who
can’t act in a Phase simply do not take any Actions. If a character who cannot act in a Phase is engaged in
Melee Combat by another character, however, they may participate as normal in the Melee.
Characters declare Actions in reverse ADR order (from lowest to highest), but the Actions are resolved in ADR Order (from highest to lowest.)
\begin{table}[!hb]
	\centering
	\caption{Round Structure}
	\label{tab:Round Structure}
	\begin{tabu} spread 0pt {X[2,l] *{3}{X}}
		Initiative&ADR 1-3&ADR 4-6&ADR 7+\\\toprule
		Phase 1&  	N  &	N & 	Y\\
		Phase 2&  	N  &	Y & 	Y\\
		Phase 3&  	Y  &	Y & 	Y\\
	\end{tabu}
\end{table}
\subsubsection{Stealing the Initiative}
In the general Round, you can temporarily raise your ADR to act in lower Phases and to improve your
place in the Initiative Order, but doing so incurs a penalty to all Actions this Round equal to the amount
by which you raise your ADR. (This penalty also applies to your Combat Pool if you end up in Melee this
Round). You cannot raise your ADR to more than double its normal amount this way. Opponents and
other characters may raise their ADR in response to any other character doing so, in the same way.
Beware overusing this option, as the penalties can quickly become so high as to make you useless, even if
you can act!

\subsection{Actions}
You may perform one Action per phase. Sometimes Actions can span multiple phases if they take longer
than a single beat in combat.

\paragraph{Base Movement}---\quad 
You can move a number of yards equal to your Mobility per Phase normally, or more
when Sprinting. If moving into Cover, you do not gain the Cover Bonus to your Defense until the
beginning of the next Phase.
Moving normally does not count as an Action, but any Action taken while moving suffers a -4 Penalty.
Conversely, moving raises your Base Defense by +1. Sprinting and Active Defense Actions are exempt
from these effects, since they are forms of Movement, and have their own penalties/bonuses.
\subsubsection{Action Types}
\paragraph{Aim}---\quad For Missile Attacks only. You focus in on one target, and increase your odds of hitting it. Your next
Missile Attack against that target gains a +2 bonus. You may Aim for multiple Phases in a row, each time
gaining a cumulative +2 bonus to your next attack, to a maximum of +6 for 3 Phases of aiming. If you
successfully Aim three times at a target in succession, you gain an additional bonus equal to your
Perception to the attack that follows. If you shift targets, lose sight of the target, or are otherwise
disrupted, you lose the Aim Bonus.

\paragraph{Missile Attack}---\quad You may perform a Missile Attack Action. There are a few varieties of these. In the event
that the target you were attempting to shoot at is incapacitated or becomes unavailable before you get to
resolve your Shoot, you may pick a new target against which to make the attack, but only at half your
base dice. Other modifiers apply as normal
\subparagraph{Shoot} lets you make a single missile attack against a target in view. These all have the same basic mechanic: Your Missile Proficiency+Aim Bonuses (if you used the Aim Action) make up your Missile Pool. For every multiple of your weapon’s Range value the enemy is further than the first,your Missile Pool is reduced by 1.
Roll Missile Pool at your weapon’s Missile TN. The target value is the enemy’s Missile Defense.
If you equal or exceed that value, you hit, and may roll on the Missile Hit Table to see where
exactly you hit.
\subparagraph{Rapid Shot} lets you make multiple missile attacks against a target in view. These function the
same as regular Shoot missile attacks, but declare beforehand how many attacks you intend to
make. Each attack after the first reduces your Missile Pool for all of the attacks made in this way by 1. So, a character making 5 Rapid Shots suffers a -4 MP penalty to all five of his attacks. Not
all weapons are capable of being used with Rapid Shot. Weapons that can only make a certain
number of attacks “normally” may be able to exceed this limit with the appropriate Talents or
weapon special rules.
\begin{description}
	\item [Bows:] Up to 5 shots can be made with Rapid Shot normally.
	\item	[Crossbows:] Cannot use Rapid Shot.
	\item	[Firearms:] Cannot use Rapid Shot normally. Some weapons with multiple-shot capacity can make up to 5.
	\item 	[Throwing Weapons:] Up to 2 shots can be made with Rapid Shot normally.
\end{description}

\paragraph{\label{par:Covering Fire} Covering Fire}---\quad lets you wait for an enemy to show himself, and then shoot him if he does.
Covering Fire suffers an additional -4 penalty if have Moved in this Phase. If a character steps
into your field of view or otherwise exposes himself (you can choose another trigger, like waiting
for the guy to draw his weapon before shooting him, etc) you may immediately take a Shoot
Action. Covering Fire lasts until the next Action you take, but you can choose to simply sustain
Covering Fire every Action until something happens. This can allow a character who cannot
normally act in Phase 1 to act in Phase 1 if he Covered in Phase 3.

\paragraph{Throw}---\quad You spend the Phase throwing a grenade or other object. You may throw the object up to its
Throwing Distance. Roll ADR at TN 7. The RS for landing the grenade within a meter of the target-area
is 4. If you fail, the throw will deviate in a random direction (1-3 too close, 4-6 too far, 7-8 to the left,
9-10 to the right) by 2 yards for every success you fail by. If you critically fail, roll deviation. The object
will land 10 yards closer to you in a straight line from its deviation point.
Throwing a weapon with the intent of actually hitting an enemy directly to injure them is resolved
differently, as a Shoot action.

\paragraph{\label{par:Active Defense:} Active Defense}---\quad You spend the Phase ducking and dodging enemy attacks. Roll an ADR Test and add BS
to your Defense RS. If you declare Active Defense, roll it before resolving any other Actions--it applies
against attacks that occur before your Initiative in the Phase, as well as those after. This can be performed
either standing still behind one piece of cover, or while running along or between close (2-3 yards apart)
pieces of cover.

\paragraph{\label{par:Take Cover} Take Cover}---\quad Similar to Active Defense, but only usable behind at least Half Cover, you simply duck down and hide. You now have Full Cover, and cannot be hit by any attacks that cannot penetrate that
Cover or otherwise circumvent it (you are still vulnerable to flanking attacks, though you count as Prone
against them.) including Suppression. You cannot move faster than 2 yards per Phase while Taking
Cover.

\paragraph{\label{par:Reload}Reload}---\quad Useful only for weapons like Firearms or Crossbows which must be loaded prior to being used, the Reload Action allows you to immediately utilize the Reload Maneuver. Moving incurs a -4 penalty to this Reload Roll.

\paragraph{\label{par:Sprint} Sprint}---\quad You move up to double your Mobility score in yards this Action in lieu of taking any non-Mobility Actions. While sprinting, you gain a +2 bonus to Defense.

\paragraph{\label{par:Melee Attack} Melee Attack}---\quad Move up to 10y into contact (2m away from) with a target, and Enter Melee Combat with a target. This Combat lasts for one Melee Round, after which the Phase ends. If you attack from a flank or
behind, this counts as a Surprise Attack. If you cannot come within 2m of the target, you cannot initiate
Melee Combat. However, for every Reach category of your melee weapon past L, you may engage an
opponent from 2 additional yards away. I.E., an EL reach weapon could engage an opponent from 8 yards
away. (This includes movement of the fighters and is not strictly indicative of the weapon’s length)
For details on Melee Combat, see below.

\paragraph{\label{par:Charge} Charge}---\quad Move up to your regular Mobility score in yards, and Enter Melee Combat with a target. This
Combat lasts for one Melee Round, after which the Phase ends. You gain a +4 bonus to your CP in this
Round, but you cannot declare any Orientation except Aggressive. If you attack from a flank or
behind, this counts as a Surprise Attack. If you cannot come within 2m of the target, you cannot initiate
Melee Combat. However, for every Reach category of your melee weapon past L, you may engage an
opponent from 2 additional yards away. I.E., an EL reach weapon could engage an opponent from 8 yards
away. (This includes movement of the fighters and is not strictly indicative of the weapon’s length)
For details on Melee Combat, see below.

\subsection{Missile Defense}
A character’s Missile Defense is, by default, 4. This means that a missile attack aimed at a character who
has no other modifiers to their Missile Defense needs 4 successes to hit. Things like movement, cover,
and other factors can alter a character’s Missile Defense. A short list is included below.
\subsubsection{Missile Hits \& Damage}
When a character is hit by a Missile Attack, roll on the \hyperref[tab:Missile Target Zones]{Random Hit Table} to determine what location the 
attack lands on. If the attack is coming from the character’s front arc, and the character has a shield, it is possible that the attack will resolve in a Hit Location covered by the shield, in which case its damage is
reduced by the shield’s AV. Attacks from the sides and rear will not hit a shield being carried normally,
but may hit a shield being carried on the character’s back.

Damage is inflicted in the following manner: $ Base~Weapon~Damage+Bonus~Successes+Misc $~versus the target’s TOU+AV. Excess damage over the target’s total inflicts Wounds as-per melee attacks. The \emph{one} exception is attacks that hit with 0 BS (exactly matching the target’s Defense) which, \emph{unlike melee attacks}, still inflict damage, but can never inflict a Wound higher than Level 1.

\section{Melee Combat}
Melee Combat occurs when characters are forced to fight up-close and personal. When this happens, the characters enter a Melee Round. This Round occurs very quickly compared to the regular combat, and so is resolved instantly.

During a Melee Round, the characters use Combat Pool, or CP, to perform Melee Maneuvers against each other. A CP normally represents one die to be rolled at a Target Number to try and score successes, to try and overpower an opponent’s defense, inflict more damage, or repel an opponent’s attack. A character only has a certain amount of CP to spend in the Melee Round, and so must conserve it and carefully assign it between the two Actions. At the start of a new Round, the characters’ CP are refreshed to their limits, though these can be reduced by Pain, being rendered Prone, or other inflictions.

Each Combat begins with the two characters rolling “Orientation,” essentially declaring whether they will be defensive, offensive or cautious in the first Action of combat. If a Combat drags on for multiple 
Rounds, only the first requires an Orientation roll, since the initiative of the characters will carry over from the previous Round.
\subparagraph{An Example of a Melee Combat}
\begin{quotation}
	\emph{Marcus and Franz have just entered combat. Marcus has 10 CP, and Franz has 12.
	The two roll for Orientation. Marcus declares Cautious, Franz declares Offensive. This means that Franz
	will be going first.
	In the first Action, Franz declares a Strike to Marcus’s Head with 8 CP, using his saber. Marcus Blocks
	with 8 CP using his shield. The two roll off, and Marcus successfully defends, which allows him to take
	the initiative.
	In the second Action, Marcus, who now has the initiative, declares a Strike to Franz’s neck with his
	battleaxe using his 2 remaining CP. Franz attempts a Parry with his saber with 4 CP. Franz succeeds in
	his Parry, and retakes the Initiative. This ends the Round. Since neither party was killed or disabled, and
	neither party retreated, the combat will continue in the next Phase with another Melee Round.}
\end{quotation}

\subsection{Round Preparation}
\subsubsection{Orientation}
Before the Round begins, each character involved in the Bout must declare their Orientation.
Orientations set the mood of the Bout by revealing whether characters intend to attack, defend or remain
cautious in the first Round. There are three types of Orientation to choose from: Aggressive, Cautious or
Defensive. It also helps to determine Initiative in the first Round.

All Orientations are declared simultaneously by both characters. Orientations are declared only once,
during the Round Preparation, except in the case of Defensive/Defensive declarations. Each Round after
the initial Round Preparation will determine Initiative and Targets from the previous Round.

Characters declare their Orientations at the same time by revealing a colored die (red for Aggressive,
yellow for Cautious, blue for Defensive). Alternatively, the GM may opt to have characters reveal one
Orientation they are NOT using first, to narrow the field, and then reveal the Orientation that they are
declaring.

\paragraph{\label{par:Aggressive}Aggressive}---\quad
Aggressive means that the character is actively engaging an enemy. When a character declares
Aggressive they commit to offensive action in the first Round, they are able to choose a Target
before Defensive and Cautious characters do, and they gain Initiative against any Defensive or
Cautious character they Target. Characters who declare Aggressive have access to the Beat and
Lance Charge Maneuvers in the first Round of combat.
\paragraph{\label{par:Cautious}Cautious}---\quad
Cautious means that the character is approaching cautiously, ready to take advantage of openings,
but not willing to leave themselves open. When a character declares Cautious they commit to
nothing in the first Round, they choose a Target after Aggressive characters do, and gain Initiative
against any Defensive character they Target. If Targeted by an Aggressive character, Cautious
characters automatically Target them back, but do not gain Initiative. Characters who declare
Cautious have access to the Dart Maneuvers in the first Round of combat.
\paragraph{\label{par:Defensive}Defensive}---\quad
Defensive means that the character is approaching Defensively, and intends to repel enemy
attacks rather than looking for openings to attack. When a character declares Defensive they
cannot declare any Attack Maneuvers in the first Action, and choose a Target after both
Aggressive and Cautious characters. Characters who declare Defensive have access to the Flee
and Total Shielding Maneuvers in the first Round of Combat.
\subsubsection*{GMs and multiple NPCs declaring Orientation}
So you’re the GM, and you’ve got three baddies going up against one PC. How the heck do you declare
Orientations?

Our advice: don’t bother holding the wrong type of dice in your hand, just decide what each mook is
going to throw, and then put those dice under a cup and lift it all at once, or load ‘em all into your hand
and drop them, or, if you’re the boring sort, just write them down and reveal them when the player
dramatically tosses his die. If the mooks are totally outclassed, there’s nothing wrong with simply telling
the player what they’re going to do.
\subsubsection*{What Orientations Mean}
Once the Orientations have been decided, all that remains is to determine who has Initiative, and who
declares their attacks first as a result.

\paragraph{\label{par:Aggressive/Aggressive}Aggressive/Aggressive}---\quad
Both characters act as though they have the Initiative over the other, declaring attacks. The
character with the lower ADR must declare his attack first, after which the other character
declares his. In the event of a tie, the PC declares second. If the characters are both PCs, they write down their maneuvers and reveal them simultaneously.

After this, both characters roll an Initiative Test (ADR+Length Advantage+Misc) to determine
whose attack resolves first. In the event of a tie, then the two characters resolve their maneuvers
with %%this part is missing in the actual document

If two Aggressive characters Target each other, they both act as though they have the Initiative
over the other. Roll an ADR test. The loser must declare his Action first, after which the winner
declares. Both still proceed as though they had Initiative (i.e. they must both declare Attack
Maneuvers), right up until the time comes to resolve those Maneuvers. At that time, the two
characters must roll an Initiative test (ADR+Length Advantage+Misc). In the case of a tie, the
two characters resolve their Maneuvers against each other simultaneously, with all Stun and Pain
being deferred until after both Maneuvers go through (this has the great potential to kill both
fighters!). 

If one fighter wins, than he resolves his Maneuver first, applies all effects of that Maneuver against his opponent, and only afterwards (if able), does the other character resolve their Maneuver.
\paragraph{\label{par:Aggressive/Anyone Else}Aggressive/Anyone Else}---\quad
If an Aggressive character and any other character Target each other, the Aggressive character
has Initiative, declares his Maneuver first, and then resolves it first.

\paragraph{Cautious/Cautious}---\quad
The two characters roll an Initiative Test (ADR+Length Advantage+Misc), prior to declaring
Maneuvers. The winner gains Initiative, the loser does not, and both may then declare Actions
and resolve them in proper order. In the event of a tie, treat it as a Defensive/Defensive situation.
\paragraph{\label{par:Cautious/Defensive}Cautious/Defensive}---\quad
The Cautious character has Initiative.
\paragraph{\label{par:Defensive/Defensive}Defensive/Defensive}---\quad
If two Defensive characters Target each other, neither of them can declare Attack Maneuvers
against each other this Round. After the first Round, if neither character has Initiative and they
are still Targeting each other, they re-declare their Orientations, in exactly the same manner as
they did during the Round Preparation.

Declaring Orientations only occurs in the first round of a combat. If a combat endures over multiple
phases, initiative is maintained from the previous round. If another character enters an existing combat,
\emph{they do not roll for Orientation} -- rather they gain initiative automatically over their opponent in the
combat.

\subsection{Maneuver Resolution}
Maneuvers are resolved as detailed in their descriptions in \autoref{ch:Maneuvers}, however the basic concepts will
be repeated here.
As an example, the Thrust Maneuver’s description:
\paragraph{\large THRUST}---\quad{\large[X]}
	\begin{description}
		\item [Type:] Universal, Thrusting
		\item [Requirements:] Wielding a weapon with a Thrust value.
		\item [Maneuver:] Weapon Attack at Weapon Reach with X dice, rolled at Thrust TN, aimed on the Thrusting Attack Table.
		\item [Success:] Inflicts damage equal to [STR+Weapon Thrust Damage+BS] to Hit Location. Damage is of weapon’s Thrust damage type.
		\item [Special:] When rolling an Initiative test to determine attack order while making a Thrust, you may roll 1 additional die in the test.
		\item [Rapid Thrust:] This may be used by weapons of S Reach or shorter only. If Thrust is successful, you may pay 2+X additional CP to make another Thrust with X dice as part of the same Action. Resolve Thrust as normal. The Target may only use Quick Defense against this Maneuver.
	\end{description}

What this means is that when a character resolves a Thrust on their Target, they do so by rolling the CP invested into the Maneuver (X) at the Thrust TN of the weapon they are using to make the attack, to a location on the Thrusting Attack Table, as shown on \autoref{tab:Hit Locations} 

If successful, this attack deals the amount of damage described in the Success section (STR+the weapon’s thrust damage rating+any Successes above the enemy Defense).
Defense works similarly, let’s look at the Block Maneuver’s description:

\paragraph{\large BLOCK}---\quad{\large[X]}
	\begin{description}
		\item [Type:] Universal, Defense, Blocking
		\item [Requirements:] A shield or other item with a Block TN.
		\item [Maneuver:] Block with X dice at Block TN. If you are using this Maneuver without Initiative, declare it against a Target attack.
		\item [Success:] Target attack negated.
		\item [Special:] You may declare this Maneuver any number of times per Action, if there are multiple incoming attacks to defend against.
	\end{description}
What this means is that when a character declares a Block against an Attack Maneuver aimed at them, he rolls his CP investment in that Maneuver at the Block TN of his shield. If he scores as many as, or more Successes, than the attacker does, than the Attack Maneuver is negated, and the character who Blocked gains the Initiative over his opponent.

\subsubsection{Target Changes}
After Maneuvers are resolved, Targets may shift if one or more characters were killed. These are thecircumstances in which Targets may change:

If a character was attacked by a character other than his current Target, and successfully defended
against that attack, the character may switch his Target to the character that attacked him, and
gain Initiative over that character if his Defense Maneuver granted it.

If a character’s current Target has been rendered un-Targetable, either by being incapacitated,
dying, or through the use of a Maneuver, the character may choose a new Target, prioritizing
those who have attacked him this Action.

If a Character’s Maneuver allows them to switch Targets, they may change to a new Target in the
manner prescribed by the Maneuver.

\subsection{Overwatch Phase}
The Overwatch Phase is identical in purpose as the Melee Phase, however it is for characters engaged in
Overwatch combat in the Bout (that is, characters with ranged weapons who have opted to Overwatch in
the General Action Phase). The Phase allows Overwatching combatants to take two Actions, including
Maneuvers.

Overwatching characters may only attack other characters engaged in their Bout. They are not able to fire
upon unengaged characters, or characters engaged in other Bouts. To do this, they must End the Watch
and either use the General Action Phase, or engage in another Bout by using the Overwatch Action.

\subsection{Round Conclusion}
At the end of the Round, if Targets are alive and present, whichever character currently has Initiative may
choose to either Disengage or Lock.
	\begin{description}
		\item [Lock:] The characters engaged in the Bout are still `in combat,' and continue fighting in the next Round. As they are still engaged, they may not participate in the General Action Phase between Rounds.
		\item [Disengage:] The characters are no longer in combat, and the Encounter ends.
If Targets are unavailable because they are dead, incapacitated, or escaped, then the Encounter ends.
Characters who were engaged in a Bout that ended due to it being Disengaged or Targets being
unavailable are able to Engage in other existing Bouts, if they are within move distance. They will be able
to use the General Action Phase so long as they are unengaged.
	\end{description}
	
\paragraph{Encounter Ends}
When all Bouts have resolved, through surrender, death or flight, the Encounter ends and characters may
interact normally.


\section{Combat Concepts}
\subsection{Combat Pool (CP)}
\[\text{Melee CP} = (\text{ADR} + \text{Proficiency Pool} + \text{Misc}) - (\text{Encumbrance}+\text{Fatigue}+\text{Pain}+\text{Misc}) \]
\[\text{Missile CP} = (\text{Proficiency Pool} + \text{Misc})-(\text{Encumbrance} + \text{Fatigue} + \text{Pain} + \text{PER Penalty}+ \text{Misc}) \]

CP represents a character's skill with the weapon they’re using, and is used to activate and power
Maneuvers of all sorts. It is determined by a character's ADR score (melee only) combined with their Proficiency Pool; thus a character who is quicker and better trained will have a higher score than one who is slower and less experienced.

CP can be reduced by a character's Encumbrance, their injuries or exhaustion, PER penalties from %HOW DO PERCEPTION PENALTIES APPLY?
headgear, and other miscellaneous factors. Because CP refreshes at the beginning of every Round, not
between Actions, it is important for a character to keep careful track of how much CP they use in the first Action, so as to have some left over for the second Action.

Maximum CP is a character's ADR (melee only), Proficiency Pool, School Bonus, Boon Bonuses, and
any miscellaneous bonuses that apply when Maximum CP is called for, minus the character's
Encumbrance Penalty, Exhaustion Penalty, Pain, and any miscellaneous penalties that apply when
Maximum CP is called for.
%BEGIN_FOLD %EXTRA target zone shit
%\section{Target Zones and Hit Locations}
%When a character is hit by an attack, there are two key aspects of the hit that need to be noted: the Target Zone (which is the general area of the body that the attack hits) and the Hit Location (which is the precisepoint of the body that is hit).
%
%Each Target Zone contains three to four Hit Locations, which are listed below. Swinging, thrusting and missile attacks have different Target Zones available to them.
%
%If an attack hits, roll 1d10 to decide the Hit Location that the attack actually lands on, and then calculate damage based on the following formula:
%
%[Total Damage (Maneuver Damage + Miscellaneous)] versus [Total Defense (TOU+AV of appropriate damage type + Miscellaneous)]
%
%If the Total Damage exceeds the Total Defense, the result is a Wound of a level equal to the excess damage. The amount of damage done is then applied on the Wound charts located in \autoref{ch:Damage Tables}
%Attacks and Missing Limbs
%If the Target is missing a body part or limb on the Target Zone that is being attacked, and the roll dictates
%that that Hit Location would be hit, the attack automatically misses and has no effect.
%Which Thigh, Though?
%When declaring an attack against an arm or a leg, or even when attacking the neck, be sure to declare
%which limb, or from which side the cut is made from, in the case of neck or torso cuts. Obviously if you
%aim for the right arm, all of the hit results should be on that arm. The same applies when making a cut to
%the neck and hitting the shoulder; specify which side you’re attacking.
%However, when making downwards or upwards swings, you may hit a shoulder or the thigh. But which
%one?
%Here’s an easy way to decide, if you don’t feel like flipping a coin: fighters tend to lead with the dominant
%side of their bodies when fighting, so if an opponent is right-handed, and you land a hit on his thigh with
%an upwards swing, you probably hit him on his right thigh.
%Alternatively, flip a coin or roll for it. GMs, be sure to be consistent about this sort of thing!
%
%	\subsection{Swinging Target Zones}
%	\begin{longtabu}spread 0pt{XX}
%		\caption{Swinging Attacks}
%		\label{tab:Swinging Hit Locations}\\
%		\rowfont[c]{}Target Zone&Hit Locations\\\toprule
%		\multirow{4}{*}{Downwards Head Swing}
%		&	1-3: Shoulder\\
%		&	4-5: Lower Head\\
%		&	6-7: Face\\
%		&	8-10 Upper Head\\
%		\\
%		\multirow{4}{*}{Upwards Head Swing}
%		&	1-3: Chest\\
%		&	4-5: Neck\\
%		&	6-8: Lower Head\\
%		&	9-10: Face\\
%		\\
%		\multirow{4}{*}{Neck Swing}
%		&	1-3: Shoulder\\
%		&	4-6: Neck\\
%		&	7-8: Lower Head\\
%		&	9-10: Face\\
%		\\
%		\multirow{3}{*}{Torso Swing (Horizontal)}
%		&	1-3: Belly\\
%		&	4-7: Side\\
%		&	8-10: Chest\\
%		\\
%		\multirow{3}{*}{Upper Arm Swing}
%		&	1-3: Elbow\\
%		&	4-7: Upper Arm\\
%		&	8-10: Shoulder\\
%		\\
%		\multirow{3}{*}{Lower Arm Swing}
%		&	1-3: Hand\\
%		&	4-7: Forearm\\
%		&	8-10: Elbow\\
%		\\
%		\multirow{3}{*}{Upper Leg Swing}
%		&	1-3: Knee\\
%		&	4-7: Thigh\\
%		&	8-10: Hip\\
%		\\
%		\multirow{3}{*}{Lower Leg Swing}
%		&	1-3: Foot\\
%		&	4-7: Shin\\
%		&	8-10: Knee\\
%		\\
%		\multirow{3}{*}{Upwards Groin Swing}
%		&	1-3: Thigh\\
%		&	4-7: Groin\\
%		&	8-10: Belly\\
%	\end{longtabu}
%	
%	\subsection{Thrusting Target Zones}
%	\begin{longtabu}spread 0pt{XX}
%		\caption{Thrusting Attacks}
%		\label{tab:Thrusting Hit Locations}\\
%		\rowfont[c]{}Target Zone&Hit Locations\\\toprule
%		\multirow{3}{*}{Head}
%		&	1-3: Neck\\
%		&	4-7: Face\\
%		&	8-10: Upper Head\\
%		\\
%		\multirow{3}{*}{Neck}
%		&	1-3: Chest\\
%		&	4-7: Neck\\
%		&	8-10: Face\\
%		\\
%		\multirow{4}{*}{Chest}
%		&	1-3: Belly\\
%		&	4-7: Chest\\
%		&	8-9: Shoulder\\
%		&	10: Neck\\
%		\\
%		\multirow{4}{*}{Belly}
%		&	1: Groin\\
%		&	2: Hip\\
%		&	3-7: Belly\\
%		&	8-10: Chest\\
%		\\
%		\multirow{3}{*}{Upper Arm}
%		&	1-3: Elbow\\
%		&	4-7: Upper Arm\\
%		&	8-10: Shoulder\\
%		\\
%		\multirow{3}{*}{Lower Arm}
%&	1-3: Hand\\
%&	4-7: Forearm\\
%&	8-10: Elbow\\
%		\\
%\multirow{3}{*}{Groin}
%&	1-3: Thigh\\
%&	4-7: Groin\\
%&	8-10: Belly\\
%		\\
%\multirow{3}{*}{Upper Leg}
%&	1-3: Knee\\
%&	4-7: Thigh\\
%&	8-10: Hip\\
%		\\
%\multirow{3}{*}{Lower Leg}
%&	1-3: Foot\\
%&	4-7: Shin\\
%&	8-10: Knee\\
%\end{longtabu}
%END_FOLD

\subsection{Missile Target Zones}
When rolling for a Hit Location with a missile weapon, first roll 1d10 on the \nameref{tab:Missile Target Zones}~table, and then roll again on the indicated Thrusting Target Zone on the \nameref{tab:Hit Locations} table that your result indicates to determine the final Hit Location of the missile attack.
\begin{table}[!ht]
	\centering
	\caption{Missile Target Zone}
	\label{tab:Missile Target Zones}
	\begin{tabu} spread 0pt {X[-1,c]X[2]}
		Roll&Target Zone\\\toprule
		1&Head\\ %SHOULD BE FACE?
		2&Chest\\
		3&Chest\\
		4&Belly\\
		5&Belly\\
		6&Groin\\
		7&Upper Arm\newline
			(1-5 left, 6-10 right)\\
		8&Lower Arm \newline
			(1-5 left, 6-10 right)\\
		9&Upper Leg \newline
			(1-5 left, 6-10 right)\\
		10&Lower Leg \newline
			(1-5 left, 6-10 right)\\
	\end{tabu}
\end{table}

\subsection{Initiative}
Unless otherwise stated, if a Defense Maneuver successfully defends against any Attack
Maneuver, the Defender gains Initiative. If a Defense Maneuver resolves but there is no Attack
Maneuver for it to defend against, it does not grant Initiative unless otherwise stated.

Action Initiative is the order in which Maneuvers declared in an Action resolve. Action Initiative, or just Initiative, is held by one character against another character that they are
``Targeting.'' If two characters are Targeting each other, only one of them can hold
Initiative.

When Maneuvers in a Bout resolve, all Maneuvers made with Initiative resolve in the reverse
order that they were declared, (unless a Steal Initiative attempt is made), and then all Maneuvers
made without initiative resolve in reverse order of declaration.

The exception to this is that Defensive Maneuvers resolve simultaneously with the Attack
Maneuver they are defending against. What this means is that if two characters are attacking each
other, the one with Initiative resolves their Maneuver first, and only afterwards does the
Initiative-less character (if they survive) resolve their Maneuver.

\subsection{Reach}\label{sec:Reach} %SECOND REACH TABLE
Reach is an important factor in combat. The fighter with longer reach tends to have an advantage, so long as the fight stays at their desired range. Entire philosophies of warfare are built around the simple advantage of having reach over one’s enemies. Of course, a long weapon reach also has its weaknesses.

When the distance is closed, longer weapons lose much of their strength and can become a liability for a fighter without a sidearm or backup weapon!

In combat, there are eight (8) stages of Reach, from Half-Arm to Ludicrously Long.

Table XX Stages of Reach
Number
Stages of Reach
Example
1
Half-Arm
HA
Knee, Elbow, Grapple
2
Hand
H
Dagger
3
Short
S
Short Sword
4
Medium
M
Arming Sword
5
Long
L
Poleaxe
6
Very Long
VL
Arming Spear
7
Extremely Long
EL
Halberd
8
Ludicrously Long
LL
Pike

Here are some important notes on Reach and how it affects combat:

The basic distance between any two characters is the Reach of the longer weapon between the two. If one of the characters has a missile weapon, then the Reach of that weapon is considered LL.

If a character successfully hits their Target with an attack that has a Reach value, the combat moves to the Reach of the attack, which will usually be the favored Reach of the attacker.

When declaring an Attack Maneuver against a Target at a longer Reach than the character’s own, an Activation Cost must be paid in addition to any others equal to the difference between the disadvantaged attacker and the defender with longer Reach.

When declaring an Attack or Defense Maneuver against a Target at a shorter Reach than the character’s own, an Activation Cost must be paid in addition to any others equal to the difference between the disadvantaged party and the opposition at the shorter Reach. This does not affect

Defense Maneuvers that do not involve defending with a weapon. Shield-based and Void Maneuvers suffer no penalty.

When making Initiative tests for Aggressive/Aggressive contests or Steal Initiative, a character whose Reach is currently favored (either the enemy is outranged, or the fight is occurring within the enemy’s range to their disadvantage) gains a +1 bonus to the Initiative test per step of Reach advantage they enjoy over their enemy.

If a character leaves an Encounter or Bout, and returns later or is re-engaged, the Reach is refreshed to the longest between the two fighters.
	\begin{quotation}
		\emph{Goofus has a short sword, whereas Gallant has a pike. Goofus’ Short Sword has S Reach, whereas Gallant’s pike has LL Reach.}

		\emph{If Goofus wants to attack Gallant, that’s five (5) stages of Reach he would need to pay for to try and make the attack (5 CP). If the attack succeeds, then he would move the fight to S Reach, since he got close enough to hit.
		At S Reach, Gallant is now the one at the disadvantage, and Attack AND Defense Maneuvers using his weapon each suffer a 5 CP Activation Cost.}

		\emph{He could still Void incoming attacks without penalty, since these do not use his now cumbersome weapon.}
	\end{quotation}
\subsubsection{Range}
When making missile attacks, the range of the weapon is measured in Yards (about three feet or one meter each) from the attacker to the target. 

Each missile weapon has a Range value, which is the number of yards at which the weapon may be used before it suffers a penalty to accuracy. For each multiple of that range at which the attack is made, the attacker suffers a -1 Penalty to his Missile attack.

So, if a weapon has 40 yard range, an attack made at up to 40 yards would be made with no penalty, one at 41-80 yards would be made at a -1 penalty, an attack at 81-12 yards would be made at a -2 penalty, etc.

\section{Additional Combat Rules}
\subsection{Declaring Defense Maneuvers without Initiative}
If you are using a Defense maneuver with Initiative, wait until characters without Initiative have declared
their Defenses to declare the target of that Defense. If nobody declares attacks against you, you may
return the Defense Dice to your Combat Pool. Then, resolve maneuvers at Initiative.
\subsection{Battlefield Terrain}
Sometimes combat will occur on uneven or otherwise irregular terrain. When fighting on rough terrain that might cause a character to fall, trip or stumble, a Mobility Test is required at the start of each Round
at an RS determined by the nature of the terrain. 

Failure inflicts Stun equal to the amount of successes by which the roll failed. This can cause Overshock and tripping!
\subsubsection{Terrain} %TERRAIN
Loose Gravel, Light Mud, Slippery Marble
Deep Mud, Shallow Water, Potholes
Collapsing Floor, Intense Mud, Deliberate Pitfalls
Difficult Terrain RS
4
6
8
Terrain which might provide one character with a defensive bonus (a wall or rampart being defended, an
overturned table being fought from behind, etc.) the attacker must pay a CP cost to make an attack,
determined by the size and nature of the defensive work.
Defense
Overturned Table, Fence
CP Cost for Attacker
2
Low Wall, Makeshift Barricade
High Wall, Rampart, Crenelations
4
6

Some terrain might produce an active danger. In this case, treat it as uneven or irregular terrain, but failing by more than X dice (represented as the second number beyond the slash in the table) results in falling, slipping into, or otherwise falling victim to the danger.

\subsection{Surprise Attacks}
If characters are caught unawares in an ambush or other surprise attack, they may not act in the first Phase, and if caught in melee their CP is halved. The Defense value of characters who have been caught unawares is reduced by 2, and they cannot benefit from Talents or other bonuses to Defense that are not from external causes like Cover or the like. 

Characters who detect the ambush beforehand may still act normally; only those surprised suffer these penalties.

\subsection{Outflanking}
Certain situations in combat can result in a character Outflanking another character. When a character Outflanks another character, they gain certain benefits:
	\begin{itemize}
		\item The Outflanked character cannot Target or perform Attack Maneuvers against the Outflanking character.
		\item The Outflanked character cannot use any Defense Maneuvers against Attack Maneuvers made by the Outflanking character without spending CP equal to the Outflanking character’s PER. Even then, only the Void or Flee Maneuver (if available) can be used.
		\item The Outflanking character makes Back Attacks when performing Attack Maneuvers against the Outflanked character.
		\item The Outflanking character ignores the Shield AV of the Outflanked character.
	\end{itemize}
Outflanking ends if the Outflanked character Targets the Outflanking character. However, if in combat with multiple opponents, Targeting one who is Outflanking, and thus ending their Outflanking status, may cause another enemy to become Outflanking, at the GM’s discretion.
	\begin{quote}
		\emph{The twins Bruce and Hazel are attacking a norse warrior named Snorri. Hazel declares an Outflank Maneuver, and succeeds, so that he is now Outflanking Snorri. Now, Snorri must pay CP equal to Hazel’s PER to defend against any of his attacks, and Hazel can target his less-defended back. If Snorri wants to get out of this, he’s going to have to defeat Bruce quick, and then turn to face his outflanking foe before he gets a few new holes carved in his back.}
	\end{quote}

\subsection{Back Attacks} %BACK ATTACKS
Any attack made from the rear, either because a character has snuck up on his Target or is Outflanking (see above) is a Back Attack.

Back Attacks replace the Chest Target Zone with the Upper Back Target Zone, and the Belly Target Zone with the Lower Back Target Zone, but otherwise roll exactly the same as normal attacks. Both Swinging and Thrusting attacks work this way.
	\begin{quote}
		\emph{Bobby is Outflanking his enemy, Bill, and declares a Torso Swing on Bill, and hits. The Hit Location roll determines that theattack hits the Chest. However, because Bobby is making a Back Attack, it actually hits Bill’s Upper Back.}
	\end{quote}

\subsection{Scatterfire} %SCATTERFIRE
Weapons that use ammunition with the Scatter special rule (Buck and Ball, Heavy Shot, and Shot) can be fired in such a way that multiple enemies are hit at once. Only enemies who are very close to each other (within a yard of one another) can be hit in this way. Targeting a character with two others within a yard’s distance away from them could result in hitting all three characters.

To start, resolve the primary attack against the center Target as normal. If it misses, all of the Scatter shots will miss as well. The primary shot resolves as normal, however the Scatter damage should be randomly divided between the potential Targets (assign numbers on a D10 to each potential Target, favoring the primary one, and roll each scatter die to determine who it hits, then determine Hit Locations as normal) and then resolved separately.

The individuals targeted by these multiple Scatters can benefit from Take Cover and other defenses as normal.

At the GM’s discretion, this rule may be applied against friendly targets as well. Firing a shotgun into a crowd of mixed friends and enemies is dangerous!

\subsection{Mounted Combat}
Characters in combat will often fight on horseback, or else mounted on exotic beasts like elephants or dinosaurs. Fighting in such a manner confers special bonuses in combat, as well as opening up new opportunities for characters to use Mounted Maneuvers.

At the start of combat, most characters who fight mounted will already be mounted. However, in the event that they are not, or if an opportunity presents itself for them to climb onto a horse, it can be done as a Move Action. If the mount is stationary, this requires an easy [RS 2] ride check. If the mount is moving, it requires a routine [RS 3] check.

Dismounting requires an effortless [RS 1] check if the mount is stationary, but a routine [RS 3] check if the mount is moving. Dismounting and mounting a mount are considered one Action.

\subsection{Explosions}\label{sec:Explosions}
Explosives such as bombs and grenades have devastating effects when used well. This section will explain the impact of explosives and how they are resolved in combat.
	\begin{description}
		\item Radius: The range within which an explosion is at its most optimal power.
		\item Power: The amount of damage that an explosion inflicts.
		\item Type: The type of damage inflicted by the explosion. If two types are listed, then the explosion inflicts half of one type, and half of the other.
		\item Special: Any special effects of the explosion. It might set all areas that suffer damage from the base explosion on fire, it may ignore some armor, or inflict extra Stun.
	\end{description}

\subsubsection{Resolving an Explosion}
All figures within the Radius of the explosion in yards suffer damage equal to the Power of the explosive, minus half of their TOU (at the GM's discretion, a fully-armored man might subtract half of their armor's Bludgeoning AV as well). The damage after reduction is randomly distributed across the body using \autoref{tab:Missile Target Zones}:~\nameref{tab:Missile Target Zones} 

Damage that lands on the same Target Zone combines before determining the Hit Location.
For each multiple of the Radius past the first, the Power of the explosion is halved, until it can no longer inflict damage. 

If an Explosion inflicts more than a Level 5 Wound to any area, add Limb Severed to the Wound's effects, regardless of its damage type.
A Take Cover Maneuver, or any sort of Cover AV, reduces damage from Explosions.

	\begin{quotation}
		\emph{Michael triggers a booby-trap, and a large bomb explodes next to him. The bomb has a power of 16, and a Range of 10 that Michael is well within. He has TOU 5. He suffers 14 (16-2) damage randomly distributed on the Missile Target Zone table. He ends up suffering a Level 4 Wound to the torso (chest), a Level 5 Wound to the right arm (elbow), a Level 4 Wound to the head (face), and a Level 1 Wound to the left leg (thigh)}

		\emph{Because the bomb relied on shrapnel to do its damage, Michael is inflicted with the Piercing damage type. Michael is left in a state of anguish, rapidly bleeding to death, and unable to move.If Michael had been 15 feet away from the bomb, he would've suffered 6 (8-2) damage, because the Power of the bomb would have been halved. If he had been 33 feet away, he would've suffered 2 (4-2) damage, as the Power would be halved yet again. Ten more feet, and he would not have suffered any damage at all, as his TOU would have cancelled out the explosion's Power entirely}
	\end{quotation}
\subsubsection{Explosives}
	\paragraph{Light Hand Grenade}
		\begin{description}
			[noitemsep]
			\item [Radius:] 5
			\item [Power:] 12
			\item [Type:] Piercing / Bludgeoning
			\item [Special:] N/A
		\end{description}
	\paragraph{Heavy Hand Grenade}
		\begin{description}
			[noitemsep]
			\item [Radius:] 10
			\item [Power:] 15
			\item [Type:] Piercing / Bludgeoning
			\item [Special:] N/A
		\end{description}
	\paragraph{Incendiary Bomb}
		\begin{description}
			[noitemsep]
			\item [Radius:] 3
			\item [Power:] 5
			\item [Type:] N/A
			\item [Special:] Does not inflict regular damage, but any area hit suffers Burn 3/TN 5 for 5 Rounds. If multiple damage lands on the same area, increase the Burn Duration by 1 per point.
		\end{description}

	\paragraph{Light Landmine}
		\begin{description}
			[noitemsep]
			\item [Radius:] 5
			\item [Power:] 15
			\item [Type:] Piercing / Bludgeoning
			\item [Special:] Can be set as a trap to explode upon being stepped on. First half of damage is always inflicted to the leg that triggers the mine.
		\end{description}

	\paragraph{Heavy Landmine}
		\begin{description}
			[noitemsep]
			\item [Radius:] 5
			\item [Power:] 30
			\item [Type:] Piercing / Bludgeoning
			\item [Special:] Can be set as a trap to explode upon being stepped on. First half of damage is always inflicted to the leg that triggers the mine.
		\end{description}

	\paragraph{Explosive Cannonball}
		\begin{description}
			[noitemsep]
			\item [Radius:] 15
			\item [Power:] 20
			\item [Type:] Piercing / Bludgeoning
			\item [Special:] Within 5 yards, add 10 Power
		\end{description}

	\paragraph{Barrel of Gunpowder}
		\begin{description}
			[noitemsep]
			\item [Radius:] 20
			\item [Power:] 30
			\item [Type:] Bludgeoning
		\end{description}
	\paragraph{Barrel of Gunpowder Filled with Shrapnel (nails, ball bearings, broken glass, etc)}
		\begin{description}
			[noitemsep]
			\item [Radius:] 30
			\item [Power:] 45
			\item [Type:] Piercing / Bludgeoning
		\end{description}
\subsection{GM Arbitration}
In a game like Song of Swords, you must expect the unexpected. With so many possible situations,
character options and potential playstyles, at some point or another you are bound to find yourself in a situation where something is technically legal by the rules, but just doesn’t make sense.

We have designed this game to the best of our ability to minimize these incidents, but if you ever encounter one, GMs, it is your responsibility to ensure suspension of disbelief (to say nothing of the enjoyment of your players!) by ruling in the direction of common sense, regardless of what the rules say.
If some bizarre twist of system terminology makes it ‘legal’ to do something patently absurd, like Shield Bash an enemy with a Pike instead of a Shield, or Grapple an opponent despite you not having arms or legs, you need not, and should not allow it.

If you feel like it, take note of such failures in the rules, and let us know so that we can fix them!
\chapter{Injury and Recovery}\label{ch:injuryrecovery}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\section{Wounds --- The Basics of Injury}
When characters sustain injuries, they sustain them in the form of Wounds. These come in five levels, from Level 1 (the least serious) to Level 5 (crippling or fatal). In minor cases, Wound Levels are temporary inconveniences, though the Pain and Stun they cause can cause a character great trouble in battle. In major cases however, high level wounds can inflict permanent Banes, like disfiguring injuries or lost limbs. Wound Levels can quite possibly cause death by blood loss, infection, or sheer damage. A character can suffer multiple Wounds of the same Damage Type to the same Area of differing levels -- Each Wound is a separate effect from any others, unless otherwise stated. (With Fire Damage, etc.)

\subsection{Anatomy of a Wound}
Wounds have a few components, illustrated here. When a Wound is suffered by a character, the level, location and type of the wound should be recorded, in addition to the \nameref{sec:Blood Loss}, Stun and Pain. 
Injuries are not removed from a character’s sheet until they are healed (see Recovery and Healing below).

\paragraph{Hit Location}---\quad
The Hit Location of the injury. This is identical to the Hit Location determined on the Hit Location Tables below.

\paragraph{Type}---\quad
This is the kind of damage that has been inflicted. Most weapons inflict Cutting, Piercing, or Bludgeoning damage. Other types of damage include \hyperref[wound:Burn Damage]{Fire} and \hyperref[wound:Cold Damage]{Cold}.
\paragraph{Bleed\label{par:Bleed}}---\quad
This is the amount of blood being lost from the wound. Different kinds of wounds inflict different amounts of Blood Loss, as detailed in their Wound Description. \nameref{sec:Blood Loss} is used for Bleeding Rolls, detailed below. 

\paragraph{Stun\label{par:Stun}}---\quad
Stun is a one-time deduction from a character's Combat Pool. It does not decrease a character's maximum CP, but rather their current reserves. Stun cannot reduce a character's CP below 0. If Stun would bring a character's CP below 0, the CP is only reduced to 0, and the character suffers Overshock.

\label{par:Overshock}Overshock forces a Stability Test at RS equal to the excess Stun after 0. Failure results in the character becoming \nameref{par:Prone}. %REFERENCES TO OVERSHOCK SHOULD BE ADDED

\begin{itemize}
\item If Stun is sustained while a Character has an unresolved Maneuver declared, the Stun is first 
subtracted from any CP devoted to the Maneuver, and after that is depleted, the rest of the Stun is 
applied to their remaining CP. 
\item If there are multiple Maneuvers declared, the effects of Stun cannot be divided up between them, 
but must be applied to one at a time. First subtract CP from one Maneuver until its CP is fully 
depleted, then do the same to the other, and finally the Character’s remaining CP. The Character 
sustaining the Stun chooses which maneuvers to subtract from. 
\end{itemize}

\paragraph{\label{par:Pain}Pain}---\quad
Pain is similar to Stun, but rather than the immediate jarring effect of the injury, it is a lasting effect. Pain inflicted by any source reduces the character’s Refresh by 1 per point of Pain, reducing the amount of Combat Pool the character gains at the start of each Round to a minimum of 0. 

If Pain reduces a character’s Combat Pool at Refresh to 0, the character is rendered helpless and incapable of doing anything meaningful (talking may or may not be possible, GM’s choice).

The character must also make a Willpower test at RS equal to any excess Pain over that required to reduce the character’s CP to 0 or lose consciousness.

A character’s total Pain is reduced by Grit before being subtracted from the character’s Combat Pool.
So, a character who has 5 Grit and who is currently suffering 12 Pain suffers a -7 penalty to their Maximum Combat Pool until the pain is alleviated. 

\section[Blood Loss]{Blood Loss and Bleeding Rolls}\label{sec:Blood Loss}
Whenever a character has an amount of Blood Loss equal to 5 or more, they must make an END test at 
the end of each Round. For every five points of Blood Loss that a character has, the test requires one more success. If the Blood Loss test is ever failed, the character suffers HLT loss equal to the amount they failed the test by.

\begin{table}[!ht]
	\centering
	\caption{Blood Loss}
	\label{tab:Blood Loss}
	\begin{tabu} to 0.4\linewidth {X[-2,r]@{\ to\ }X[-2,l]X[r]}
\multicolumn{2}{c}{Total Blood Loss}&Required Successes to Stave Off\\\toprule
0&4 &0 successes - No need to roll \\
5&9 &1 success \\
10&14 &2 successes \\
15&19 &3 successes \\
20&24 &4 successes \\
\multicolumn{2}{c}{+\ldots} &+\ldots\\
	\end{tabu}
\end{table}

\subsubsection{Losing Health:}
Any time that a Blood Loss test is failed, the Character loses an amount of HLT equal to the number of 
successes by which the test failed.

When a character's HLT reaches 1 or lower from Blood Loss, all of their other Attributes are temporarily halved. When a character’s HLT reaches 0 or or lower, they immediately pass out and continue to bleed, making END tests and losing HLT into negative amounts. When their HLT reaches a negative amount equal to their unmodified total HLT, they die. Note that when the character passes out, the TOU vs. 2 RS to avoid \nameref{bane:Brain Damage}~is done with their newly-modified TOU. (See \nameref{par:Knockout}, below.)

\begin{quote}
	\emph{Duke has been shot several times with a repeating crossbow. He is suffering 22 Blood Loss. He must therefore make an Endurance test with his END of 5 against 4 successes.}
\end{quote}
\subsubsection{Regaining Health:}
Once a wound has been treated via \nameref{skill:Surgery}~so that no more \nameref{sec:Blood Loss} can occur, HLT lost to \nameref{sec:Blood Loss} will begin replenishing at a rate of 1 HLT per week.

HLT lowered by \nameref{sec:Infection} is also recovered in this manner, but only after all of a character’s Infections have been defeated.

\subsubsection{Staunching Blood Loss}
Characters can attempt to staunch the loss of blood with the \nameref{skill:Surgery} skill and whatever materials are at hand. A wound must be targeted with this skill which, if successful, terminates all \nameref{sec:Blood Loss} from that wound.

If a Character goes for over an hour without losing a point of HLT, any remaining \nameref{sec:Blood Loss} is automatically staunched, including \nameref{par:Internal Bleeding}.

\subsubsection{Treating Wounds}
In order to stop a wound from bleeding, and to reduce the Pain it causes, the \nameref{skill:Surgery} skill must come into 
play. A Surgery check made at appropriate difficulty for wound level Treats the wound. Treated wounds 
have their \nameref{sec:Blood Loss} reduced to 0 (unless reopened somehow), and their Pain reduced by half. Treated 
wounds begin healing immediately. Healing time depends on the character’s maximum normal HLT, and 
the level of the wound. 

Once a Wound has fully healed, the Wound, and any remaining pain, are removed from the character.

Unarmed Wounds heal in Days instead of Weeks. 

\begin{table}[!ht] %THIS is the ugliest fucking table of all time holy shit
	\centering
	\caption{Wound Healing Times (in Weeks)}
	\label{tab:Wound Healing}
	\begin{tabu} spread 0pt{XX[-0.5,c] *{8}{X[-2,c]}}
 	& &\multicolumn{8}{c}{Health}\\
 	& &	1&2&3&4&5&6&7&8+\\\toprule
\parbox[t]{2 mm}{\multirow{5}{*}{\rotatebox[origin=c]{90}{Wound Level}}}
	&\multicolumn{1}{c|}{1}&	2&2&2&1&1&1&1&0\\
	&\multicolumn{1}{c|}{2}&	3&3&2&2&2&1&1&1\\
	&\multicolumn{1}{c|}{3}&	4&4&4&3&3&3&2&2\\
	&\multicolumn{1}{c|}{4}&	8&8&8&6&6&6&5&5\\
	&\multicolumn{1}{c|}{5}&	60&60&60&48&48&48&36&36\\
	\end{tabu}
\end{table}

\section{Damage Conditions and Tests}
\paragraph{\label{par:Knockout}Knockout}---\quad
A Knockout (KO) test, is an END test. You need a number of successes specified by the event. Failure causes the character to be knocked unconscious, \nameref{par:Prone}, and unable to act for 1d10 minutes. A character that is knocked out cannot defend themselves and is entirely helpless. Whenever a character is knocked out, they must pass a TOU vs. 2 RS test, or gain the \nameref{bane:Brain Damage}~ (Minor) Bane.

\subparagraph{Optional:} If a character is knocked out while standing up or fighting, the GM may rule that he suffers \nameref{sec:falling-damage}~due to landing on a hard surface or an object while unconscious. This usually will be equivalent of a fall from 6 feet. An unconscious man hitting an awkward surface (the corner of a table, a street curb, etc.) could suffer a far more severe injury than the actual blow that rendered him unconscious.

\paragraph{\label{par:Stability}Stability}---\quad
When a character makes a Stability Test (either as specified by a wound or effect, or because their CP has been reduced to the negatives by Stun or Pain, or other reasons at GM’s discretion) roll an AGI test. The RS will vary depending on the cause of the event that called for a Stability Test. Failure renders the character Prone.

\paragraph{\label{par:Prone}Prone}---\quad
Whenever a character is laying flat on the ground, they are considered to be Prone. While a character is prone, their maximum combat pools for all melee proficiencies and their Mobility are halved, and the effective Reach of any weapon they’re using (including unarmed “weapons” like fists, legs, etc) is reduced by 2 stages. While Prone, all Maneuvers except for Block are made at +1 TN.

Being knocked prone (after failing a Stability test, being knocked down, etc.) causes a character to immediately take Stun equal to half their maximum combat pool. Until the character is standing during a Refresh, their maximum combat pool remains halved, and any Refresh when prone only restores them to that half-maximum.

The use of the \nameref{man:Rapid Rise}~Maneuver allows a character to stand up immediately upon falling down, if they have the remaining CP. Check \autoref{ch:Maneuvers}: \nameref{ch:Maneuvers} for more information. 

Otherwise, a character must spend an entire action standing up, and cannot use any Maneuvers for that action except \nameref{man:Quick Defense}

\subsection{Falling Damage}\label{sec:falling-damage}
When a character falls, they roll on the first table below to determine what body part they land with. They then take a wound as if struck with a blunt object to the area hit at Strength determined by the distance fallen, and the surface fallen onto. Neither Armor nor TOU reduce this damage. The character also takes additional damage to the adjacent body parts as determined by the falling table at half damage.

\begin{table}[!ht]
	\centering
	\caption{Falling Damage Hit Location}
	\label{tab:falling-damage-hit-location}
	\begin{tabu} spread 0pt{X[-1,c]XX[2]}
d10 Result& Area Hit &Adjacent Areas\\
1 &Head&Neck and Shoulder \\
2 &Neck&Shoulder and Head \\
3 &Upper Back&Lower Back and Head \\
4 &Chest&Upper Abdomen and Head \\
5 &Right Foot&Thigh and Lower Leg (Right) \\
6 &Left Foot&Thigh and Lower Leg (Left) \\
7 &R. Side&Upper Arm and Thigh \\
8 &L. Side&Upper Arm and Thigh\\
9 &R. Forearm&Hand and Upper arm\\
10&L. Forearm&Hand and Upper arm\\
	\end{tabu}
\end{table}

\begin{table}[!ht]
	\caption{Falling Damage}
	\label{tab:Falling Damage}
	\begin{tabu} spread 0pt{X[2]X}
Ground Type&Damage per Feet fallen\\\toprule
Hard (Cement, Stone, Packed Dirt, Cobblestones, Hardwood floor, a thoughtlessly placed pile of bricks)&1 per 2’ fallen\\
Solid Earth, carpeted floor, collapsible structure (falling through a cart, wagon, canopy, an armored man)&1 per 4’ fallen\\
Soft (Straw, trampled battlefield earth, mud, Sand, a Person, a pile of dead people)&1 per 8’ fallen\\
	\\
Rough Slope (Rocky Hill, Flight of Stairs, Shingled Rooftop)&1 per 10’ tumbled\\
Deadly Slope (Cliffside, Jagged Mountain Slope, Castle Ramparts)&1 per 6’ tumbled\\
Hard Angle (Corner of a Table, Edge of a bin, street curb, etc)&1 per 1’ fallen, may inflict Cutting or Piercing damage\\ %COULD USE A FOOTNOTE
Water (clumsily fallen)&1 per 30’ fallen\\
Water (controlled dive)&1 per 60’ fallen\\
	\end{tabu}
\end{table}

\subsubsection{Falling onto Slopes}
When falling down a slope, a character sustains damage at a different rate than when falling from the air 
into a surface, and each step of the fall should be calculated separately.

For example: An assassin is flung off a balcony after a failed attempt on the life of a King, and then falls 
fifty feet onto a hard rooftop, where he then rolls another fifty feet before falling off this second roof, 
down another twenty feet onto the cobblestone street, he would sustain three separate Falls. Let’s run 
through an example of what happens: 
\begin{itemize}
\item First Fall: 50’ onto a Hard surface. The assassin sustains 25 damage to his Right Arm, and 12 
Damage to his Hand and Upper Arm. This results in 3 Level 5 Wounds. 
\item Second: 50’ on Rough Slope (5 damage) The assassin rolls down the slope, and suffers 5 damage 
to his neck, and 2 damage to his shoulder and head. At this moment, the assassin dies. 
\item Third: 20’ on Hard (10 damage) The assassin, who is already dead, finally lands on his back, 
suffering 10 damage to his upper back, and 5 damage to his head and lower back. This results in 
three more level 5 wounds. 
\end{itemize}

Every bone in his right arm is shattered, his neck is broken, his spine is broken is two places, and his head 
is smashed upon. He is dead. BUT, what if, like a smart Assassin, he had aimed for the conspicuous pile 
of straw instead of the rooftop? 

Well, assuming the slope of the roof was fairly steep and actually covered about 30 feet of vertical space, 
that’d mean that the total distance fallen would’ve been 100 feet exactly. 

Now let’s say that the the Assassin leaps into the pile of straw from the same total height (about 100 feet) 
instead. He instead suffers 12 damage to one body part, and 6 damage to two others. If he’s lucky, he 
might only shatter one of his legs or arms, but he’s just as likely to break his neck, spine, or rupture all of 
his organs.

Falling onto certain materials or hazards might justify changing the damage type inflicted from Bludgeoning to Cutting or Piercing damage. Use your judgment, GM.

\subsection{Burning \& Fire Damage}
When characters come into contact with fire, bad things tend to happen. Fire damage is much simpler 
than damage from blunt, cutting, or piercing weapons. Its table is generic, and any part of the body may suffer a Burn Wound using the same chart, seen below. 

When a character suffers Fire Damage, the format looks like this: $\dfrac{\mathrm{X}}{\mathrm{TN5}}$

X is the number of dice rolled, and the TN is always 5. Each success on this roll inflicts a level of Burn 
Damage to the body part (or body parts) affected. 

Burn Damage is not reduced by TOU, but it is reduced minorly by Armor.
Cloth, Leather and other light armors do nothing to reduce burn damage. In fact, they might make it worse (Quilted Armor, if worn 
without Plate over it, increases the Burning Time of any Fire Damage by 1. Mail, Laminar, Brigandine, 
etc. reduces Burn Damage by 1. Plate reduces Burn Damage by 2).

Burn Damage is Cumulative. If a character suffers a Level 1 Burn, and then next Round suffers two
additional Burn Damage, this does not create a Level 2 Burn, but instead turns the Level 1 into a Level 3.

Burning damage often lasts for multiple Rounds, inflicting damage at the beginning of each, and clothing or armor may also catch fire, continuing the blaze for still longer! 

\begin{table}[!ht]
	\centering
	\caption{Burn Examples}
	\label{tab:Burn Area Examples}
	\begin{tabu} to \linewidth {X[-0.5] X X[-2,c] X}
\rowfont[c]{}Example Burn Source
&	Coverage 
&		Damage 
&			Burning Time\\\toprule

Boiling Oil
&	Usually head, shoulders, neck, back, possibly face or arms
&		$ \dfrac{\mathrm{4}}{\mathrm{TN5}} $
&			4 Round\\

Flaming Arrow
&	Point of Impact
&		$\dfrac{2}{\mathrm{TN5}} $
&			10 Rounds\\

Hair Caught Fire
&	Head(?)
&		$ \dfrac{1}{\mathrm{TN5}} $
&			1-3 Rounds (length of hair)\\

Burning Sleeve
&	Arm
&		$ \dfrac{2}{\mathrm{TN5}}$
&			5 Rounds\\

Bonfire
&	Full Body 
&		$\dfrac{3}{\mathrm{TN5}}$
&			Indefinitely, 3 Rounds after exiting Bonfire\\
	\end{tabu}
\end{table}

Extinguishing a fire on one’s person is either very easy (patting out, smothering with a blanket or cloth, 
dropping and rolling) or excessively difficult (boiling oil, burning pitch, chainmail made dangerously hot 
in a fire, acid). You may try to Extinguish a source of burn damage by spending a Full Action in the 
Bystander Phase doing so. This is an Agility check, with the RS ranging from 1 (burning sleeve) to 8 (covered in burning pitch). %REFERENCE TO BYSTANDER PHASE?

So, a character may suffer a pot of boiling oil being dumped on his head while climbing up a siege ladder. 
The GM estimates this as a 4/TN5 burn attack that hits the upper head, neck, shoulders, and back. The 
roll is made, 4 at TN 5.
2 successes are scored, however the character is wearing a Plate helmet and a Mail shirt, which reduces 
burn damage to the upper head by 2, the back and shoulders by 1. The neck suffers all damage. 
The result is a Level 1 Burn to both shoulders (two wounds), and a Level 2 burn to the neck. The next 
Round, he suffers additional damage. 

\begin{table}[!ht]
	\centering
	\caption{Burn Damage}
	\label{wound:Burn Damage}
	\begin{tabu} to \linewidth {X[-1,l] *{3}{X[-3,c]} X[2]X[4]}
\rowfont[c]{}Wound Level &Stun &Pain &Bleed &Special &Description\\\toprule
Level 1 &4 &4 &- &[-3 Infection Chance]&1st degree Burn. Painful, but not serious.\\
Level 2 &8 &6 &- &[-1 Infection Chance]&Minor 2nd degree burn with swelling, but no probable scarring.\\
Level 3 &12 &10 &- &[+1 Infection Chance] &Bad 2nd degree burn, skin lost, scarring probable.\\
Level 4 &20 &20 &- &[+3 Infection Chance] &3rd degree burn, serious scarring and injury. Consider disfigurement Banes.\\
Level 5 &Total &30 &- &[+5 Infection Chance] &4th degree burn or worse, flesh destroyed, bone exposed and charred, almost certainly lethal if located on vitals (HLT at RS 5)\\
	\end{tabu}
\end{table}


\subsection{Electrical Damage}
Being electrocuted is not something that happens often in a medieval setting, but it can happen. Electrical 
Damage is made against TOU, but it is not reduced by any sort of armor, and in fact is increased by 
metallic armor (mail and scale increases Electrical damage by 1, Laminar and Plate of all varieties 
increases it by 2).
A character does not suffer Electrical Wounds to any particular part of their body, the Wound applies to 
the character’s whole frame. However, in extreme cases (Damage 5 and Higher) a Burn Wound may be 
inflicted (generally 1 at 5, +1 per 2 points of Electricity Damage past 5) at the point of electrocution. 
Multiple Electrical Wounds combine, creating a new Wound with a Level equal to the highest previous 
Wound +1. (So if a character has a Level 3 Electrical Wound, and suffers an additional Level 1-3 
Electrical Wound, he instead upgrades the original Level 3 Wound to a Level 4, and suffers its effects 
instead.)
Each Action (second) that a character stays in contact with the source of Electricity, he suffers the damage 
again.
On top of any actual damage inflicted, Electricity always inflicts Shock equal to the amount of damage 
dealt.
Electrical Wounds cannot be Infected, and heal at twice the speed of other Wounds. 


\begin{table} %SHOULD probably be a wraptable
	\centering
	\caption{Electricity Damage by Source}
	\label{tab:Electricity Damage by Source}
	\begin{tabu} spread 0pt {Xr}
Source &Damage\\\cmidrule{1-2}
Mild Shock (power outlet) &3\\
Severe Shock (Car Battery)&5\\
Dangerous Shock (Transformer) &6\\
Lethal Shock (Struck by Lightning) &8\\
Cataclysmic Shock (Power Lines) &10\\
	\end{tabu}
\end{table}

\begin{table}[!ht]
	\centering
	\caption{Electrical Damage}
	\label{wound:Electrical Damage}
	\begin{tabu} spread 0pt {X[-1,l] *{3}{X[-3,c]} X[2]}
Wound Level &Stun &Pain &Bleed &Special\\\toprule %&Description\\
Level 1 &8  &0 &- &-\\
Level 2 &10 &2 &- &KO at 1\\
Level 3 &12 &4 &- &KO at 2\\
Level 4 &16 &20 &-&KO at 3\\
Level 5 &Total &Total &- &KO at 4, HET at 3 vs Death\\
	\end{tabu}
\end{table}

\subsection{Cold Damage}
Something something, Cold can target Limbs, TOU reduces, Armor doesn’t, Warm Clothing can slow 
effects, need rules for Freezing, yadda yadda, what killed the dinosaurs? Get to the Choppa. 

\begin{table}[!ht]
	\centering
	\caption{Cold Damage}
	\label{wound:Cold Damage}
	\begin{tabu} spread 0pt {X[-1,l] *{3}{X[-3,c]} X[2]}
\rowfont[c]{}Wound Level &Stun &Pain &Bleed &Special\\\toprule %&Description
Level 1 &2 &2 &- &-\\
Level 2 &4 &4 &- &-\\
Level 3 &6 &2 &- &Reduces all Pain in Zone by 2 \\
Level 4 &8 &0 &-
	&Reduces all Pain in Zone by 4,\newline
	Difficult Surgery vs Frostbite if Limb,\newline
	HET test at 3 vs Death if not Limb\\
Level 5 &10& 0& -
	&Reduces all Pain in Zone by 10,\newline
	Miracle Surgery vs Amputation if Limb,\newline
	HET test at 4 vs Death if not Limb\\
	\end{tabu}
\end{table}

\section{Infection}\label{sec:Infection}
The shock of a wound can kill a man, or he can bleed out before the injury can be staunched, but the most 
insidious and terrible way to die of a wound is by corruption, or infection. In the course of healing, a 
wound can become infested with bacteria and parasites which quickly cause the wound to become 
gangrenous or septic, which can eventually result in necrosis and death. 

Infection has a chance of occurring 24 hours after the infliction of the wound. Roll a HLT Test at RS
equal to the Infection Level of the Wound. The Infection Level equal to the Wound Level, plus any
miscellaneous modifiers. Both Wound descriptions and exterior circumstances (see \autoref{tab:Infection Modifiers}) can
increase or decrease Infection Chance.

Unarmed Wounds cannot cause Infection.

\begin{table} %SHOULD probably be a wraptable
	\centering
	\caption{Infection Modifiers}
	\label{tab:Infection Modifiers}
	\begin{tabu} spread 0pt {Xr}
Infection Modifiers &Modifier Amount\\
Wound was successfully treated with Surgery immediately after it was inflicted. &-1 per 2 BS\\
Wound is treated with an expensive poultice during Surgery &-2\\
Wound is the result of an Amputation &-2\\
Wounded party is allowed to rest immediately &-1\\
Wound was inflicted in a fairly clean environment free of filth (being cut with a well maintained dress sword during a parade, being injured by a practice weapon). &-2\\
Wound was inflicted by a bludgeoning weapon. &-1\\
Wound was inflicted by a piercing weapon. &+1\\
Wound was inflicted in an unclean environment. &+1\\
Wound was inflicted with a filthy weapon. &+2\\
Wounded party has strenuously exerted himself since being injured. &+1\\
Wounded party did not receive adequate care after the injury. &+2\\
Wound was intentionally infected after being sustained. &+3\\
	\end{tabu}
\end{table}
If the character succeeds, then the Wound does not become infected, and heals normally. However, if 
something occurs to infect the Wound again (the wound is exposed to filth, or reopened, or stripped of its 
coverings and left unprotected for some time) then an additional roll may be necessary. 

If the Infection Roll fails, then the Wound becomes Infected. Once a Wound becomes Infected, the 
character’s body begins to fight against the infection by all means possible, some of which are quite 
detrimental. Characters with Infections become Feverish within 24 hours, and remain Feverish until the 
Infection is defeated. (See Illness and Disease section for the effects of Fever. Short version: The 
character suffers a Halved Combat and Mobility pool, and gains Exhaustion at double the normal rate). 

Every 24 hours, the character must roll a Health Check at RS equal to the Infection Level of the Wound. 
This roll can be modified with Medical skill, as detailed on Table 7.2, and this will greatly increase a 
character’s chance of survival. If unsuccessful, the character’s Health is reduced temporarily by 1. This 
Health does not recover until after the Infection is defeated. If the character’s Health is reduced to 1 by 
Infection, then all current Infections become Septic. 

If a character succeeds a number of these Health Checks equal to the Infection Level of the wound, that 
Infection is defeated, and the character can begin recovering Health, assuming there are no other 
Infections to struggle against. 

In the case of multiple Infections developing, only one Health Check is rolled per 24 hour period, at the 
highest Infection Level of the infected wounds suffered. If the Check succeeds, all of the Infections are 
resisted. If the Check fails, the character only loses 1 point of Health, not 1 per Infection. 

Table 8.8 Infection Treatment
Modifier Amount
Character’s infections are treated by a Surgeon -2 
Character is kept in a comfortable, clean environment -2 
Character has enough to eat/drink -1 
Character partakes of no strenuous activity -1 

\subsection{Septic Infections}
Septic Infections (when the injury becomes gangrenous and decaying) are almost invariably lethal. After 
a wound becomes Septic, it rapidly spreads across the body, eventually reaching the heart or brain, and 
killing the victim. The character becomes delusional, weak, and likely unconscious, and dies within 
2d10+HLT hours. If the character can pass a WIL+HLT test at RS 5, the character can struggle on for 
another 24 hours, but will still die before the next day. 

There is no treatment for Sepsis, but if the Infection is in a limb, it can be Amputated immediately upon 
the onset of Sepsis to prevent it from spreading. This increases the Infection Level of the resulting 
Amputated wound by 1, whereas Amputating before Sepsis develops does not increase the Infection 
Level at all. Surgeons must be careful when deciding whether or not to amputate too early on what 
might’ve been a salvageable limb. 

\subsection{Amputation}
When a limb is Amputated (which itself requires a Surgery check at RS 3), the limb suffers the equivalent 
of a level 5 cutting wound on the location where the Infection is located, but with no Bleed. This 
automatically removes the Infection that occupied the amputated limb, and the wound can then be treated 
like a normal wound for terms of surgery checks to end blood loss and begin the healing process. 
Amputations also confer an automatic -2 to the Infection Chance of the wound, due to the care that goes 
into the procedure to make it as clean as possible, and may benefit from other factors as well, such as a 
clean environment with proper tools, to minimize the chance of the new wound becoming Infected as 
well.

If a wound becomes Septic before being amputated, as detailed in the Septic Infection section above, then 
the Amputation’s -2 to Infection Chance is reduced to -1, increasing the possibility of repeated Infection, 
as some of it may have already spread beyond the limb. 

\chapter{Wound Charts}\label{ch:Damage Tables}
\startcontents[chapters]
\printcontents[chapters]{}{1}{}
\clearpage
\begin{center}
	This is the part where you learn about what it’s like to die.
\end{center}

\section{Target Zones and Hit Locations}
When a character is hit by an attack, there are two key aspects of the hit that need to be noted: the Target 
Zone (which is the general area of the body that the attack hits) and the Hit Location (which is the precise point of the body that is hit).

Each Target Zone contains three to four Hit Locations, which are listed below. Swinging and Thrusting attacks have different Target Zones available to them.

If an attack hits, roll 1d10 to decide the Hit Location that the attack actually lands on, and then calculate 
damage based on the following formula: 
\[\text{Total Damage Done}~(\text{Maneuver Damage}+\text{Miscellaneous}) ~-~ \text{Total Defensive Value}~(\text{TOU}+\text{AV} + \text{Miscellaneous})\]

If the Total Damage exceeds the Total Defense, the result is a Wound of a level equal to the excess 
damage.

The amount of damage done is then applied on the damage charts located later in this chapter.

\subsection*{Which Thigh, though?}
When declaring an attack against an arm or a leg, or even when attacking the neck, be sure to declare 
which limb, or from which side the cut is made from, in the case of neck or torso cuts. Obviously if you 
cut the right arm, all of the hit results should be on that arm. The same applies when making a cut to the 
neck and hitting the shoulder. Specify which side you’re attacking.
However, when making, say, downwards or upwards swings, you may hit a shoulder or the thigh. But 
which one?

Here’s an easy way to decide if you don’t feel like flipping a coin: fighters tend to lead with the dominant 
side of their bodies when fighting, so if an opponent is right-handed, and you land a hit on his thigh with 
an upwards swing, you probably hit him on his right thigh.

Alternatively, flip a coin or roll for it. GMs, be sure to be consistent about this sort of thing!

\newpage
\subsection{Hit Location Tables} %NEED REF TO BACK ATTACKS AND MISSILES?
\begin{table}[!hb]
{	\captionsetup{textformat=empty, labelformat=blank}\vspace{-2em}
	\caption{Target Hit Locations}
	\label{tab:Hit Locations}}
	\parbox{0.5\linewidth}{
		\centering
		\caption{Swinging Attacks}
		\begin{tabu} spread 0pt{X[1.2]X}
	\rowfont[c]{}Target Zone&Hit Locations\\\toprule
	\multirow{4}{*}{Downwards Head Swing}
	&	1-3: \nameref{sec:shoulder}\\
	&	4-5: \nameref{sec:lower-head}\\
	&	6-7: \nameref{sec:face}\\
	&	8-10 Upper Head\\
		\\
	\multirow{4}{*}{Upwards Head Swing}
	&	1-3: \nameref{sec:chest}\\
	&	4-5: \nameref{sec:neck}\\
	&	6-8: \nameref{sec:lower-head}\\
	&	9-10: \nameref{sec:face}\\
		\\
	\multirow{4}{*}{Neck Swing}
	&	1-3: \nameref{sec:shoulder}\\
	&	4-6: \nameref{sec:neck}\\
	&	7-8: \nameref{sec:lower-head}\\
	&	9-10: \nameref{sec:face}\\
		\\
	\multirow{3}{*}{Torso Swing (Horizontal)}
	&	1-3: \nameref{sec:belly}\\
	&	4-7: \nameref{sec:side}\\
	&	8-10: \nameref{sec:chest}\\
		\\
	\multirow{3}{*}{Upper Arm Swing}
	&	1-3: \nameref{sec:elbow}\\
	&	4-7: \nameref{sec:upper-arm}\\
	&	8-10: \nameref{sec:shoulder}\\
		\\
	\multirow{3}{*}{Lower Arm Swing}
	&	1-3: \nameref{sec:hand}\\
	&	4-7: \nameref{sec:forearm}\\
	&	8-10: \nameref{sec:elbow}\\
		\\
	\multirow{3}{*}{Upper Leg Swing}
	&	1-3: \nameref{sec:knee}\\
	&	4-7: \nameref{sec:thigh}\\
	&	8-10: \nameref{sec:hip}\\
		\\
	\multirow{3}{*}{Lower Leg Swing}
	&	1-3: \nameref{sec:foot}\\
	&	4-7: \nameref{sec:shin}\\
	&	8-10: \nameref{sec:knee}\\
		\\
	\multirow{3}{*}{Upwards Groin Swing}
	&	1-3: \nameref{sec:thigh}\\
	&	4-7: \nameref{sec:groin}\\
	&	8-10: \nameref{sec:belly}\\
		\end{tabu}
	}
\hfill
	\parbox{0.5\linewidth}{
		\centering
		\caption{Thrusting Attacks}
		\begin{tabu} spread 0pt {X[1.2]X}
	Target Zone&Hit Locations\\\toprule
	\multirow{3}{*}{Face Thrust}
	&	1-3: \nameref{sec:neck}\\
	&	4-7: \nameref{sec:face}\\
	&	8-10: \nameref{sec:upper-head}\\
		\\
	\multirow{3}{*}{Neck Thrust}
	&	1-3: \nameref{sec:chest}\\
	&	4-7: \nameref{sec:neck}\\
	&	8-10: \nameref{sec:face}\\
		\\
	\multirow{4}{*}{Chest Thrust}
	&	1-3: \nameref{sec:belly}\\
	&	4-7: \nameref{sec:chest}\\
	&	8-9: \nameref{sec:shoulder}\\
	&	10: \nameref{sec:neck}\\
		\\
	\multirow{4}{*}{Belly Thrust}
	&1: \nameref{sec:groin}\\
	&2: \nameref{sec:hip}\\
	&3-7: \nameref{sec:belly}\\
	&8-10: \nameref{sec:chest}\\
		\\
	\multirow{3}{*}{Upper Arm Thrust}
	&	1-3: \nameref{sec:elbow}\\
	&	4-7: \nameref{sec:upper-arm}\\
	&	8-10: \nameref{sec:shoulder}\\
		\\
	\multirow{3}{*}{Lower Arm Thrust}
	&	1-3: \nameref{sec:hand}\\
	&	4-7: \nameref{sec:forearm}\\
	&	8-10: \nameref{sec:elbow}\\
		\\
	\multirow{3}{*}{Groin Thrust}
	&	1-3: \nameref{sec:thigh}\\
	&	4-7: \nameref{sec:groin}\\
	&	8-10: \nameref{sec:belly}\\
		\\
	\multirow{3}{*}{Upper Leg Thrust}
	&	1-3: \nameref{sec:knee}\\
	&	4-7: \nameref{sec:thigh}\\
	&	8-10: \nameref{sec:hip}\\
		\\
	\multirow{3}{*}{Lower Leg Thrust}
	&	1-3: \nameref{sec:foot}\\
	&	4-7: \nameref{sec:shin}\\
	&	8-10: \nameref{sec:knee}\\
		\end{tabu}
	}
\end{table}
\clearpage
\section{Wound Qualities}

\paragraph[Stability vs. ]{\label{par:Stability vs}Stability vs [X] RS}---\quad
You must make a \nameref{par:Stability}~(AGI+Misc) test at X RS or be rendered \nameref{par:Prone}.

\vspace{-8pt}\paragraph[Disarm vs. ]{\label{par:Disarm vs}Disarm vs [X] RS}---\quad
You must make a Disarm (AGI+Misc) test at X RS or drop anything held in the affected arm.

\vspace{-5pt}\paragraph[KO vs. ]{\label{par:KO vs}KO vs [X] RS }---\quad
You must make a \nameref{par:Knockout}~(END+Misc) test at X RS or be rendered unconscious for 1d10 minutes.

\vspace{-5pt}\paragraph[Surgery vs. ]{\label{par:Surgery vs}[X] Surgery vs [Y]}---\quad
You immediately gain Y Bane, which cannot be removed without a successful Surgery test at X difficulty before the wound heals. If the wound heals (or the surgery test fails) then Y Bane becomes permanent.

\vspace{-5pt}\paragraph{\label{par:Autodrop}Autodrop}---\quad 
You automatically drop anything held with this limb when this Wound is inflicted. Increase the RS of any \nameref{par:Stability}~test made to maintain your grip on a weapon with this limb by the Wound level. %WAIT WHAT THE FUCK DO YOU MEAN STABILITY TO MAINTAIN GRIP

\vspace{-5pt}\paragraph{\label{par:Auto-Prone}Auto-Prone}---\quad
You are automatically rendered Prone. You cannot Stand Up in the same action, or in the action after receiving this Wound. Increase all further Stability tests RS by the Wound level, until the Wound is fully healed.

\vspace{-5pt}\paragraph{\label{par:Auto-KO}Auto-KO}---\quad
Character is rendered unconscious for 1d10*6 minutes. Increase the RS of any further KO rolls by the Wound level, until the Wound is fully healed.

\vspace{-5pt}\paragraph[Infection Risk]{\label{par:Infection Risk}Infection Risk+[X]}---\quad
This Wound suffers an increased~\nameref{sec:Infection}~Risk equal to X.

\vspace{-5pt}\paragraph[Total]{\label{par:Stun: Total}Stun: Total}---\quad
Automatically reduces Combat Pool to 0, and immediately renders you Prone without a test.

\vspace{-5pt}\paragraph[Total]{\label{par:Pain: Total}Pain: Total}---\quad
Your effective Pain is considered infinite until this Wound has healed.

\vspace{-5pt}\paragraph[Limb Severed!]{\label{par:Limb Severed}Limb Severed}---\quad
This limb is cut off! You receive the \hyperref[bane:Severed Limb/Appendage]{Severed Limb}~Bane. 

\vspace{-5pt}\paragraph{\label{par:Dead}Dead}---\quad
You’re Dead. You get the \nameref{bane:Dead}~Bane.

\vspace{-5pt}\paragraph{\label{par:Internal Bleeding}Internal Bleeding}---\quad
Halting this bleeding requires a Surgery roll at +3 RS from normal.
\newpage
		\captionsetup{font=small} %Set captions for wound tables to smaller size
		\AtBeginEnvironment{tabu}{\scriptsize}{}{}
\section{Upper Head}\label{sec:upper-head}
\begin{table}[!hb] %Upper Head - Cutting
%	\captionsetup{textformat=empty, labelformat=blank}
%	\caption{Upper Head - Cutting}\vspace{-15pt}
%	\label{wound:Upper Head - Cutting}
	\begin{tabu} to 1\linewidth {X[1.1] *{3}{X[-2,c]} X[3]X[5]}
Wound~Level&\nameref{par:Stun} &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1 &7 &1 &- &Shallow cut on scalp, not serious.\\
Level 2& 2 &9 &3 &- &Deep cut on scalp, blood runs into face.\\
Level 3 &4 &13&5 &\nameref{par:KO vs}4 RS &Cracked skull, nearly invisible, but devastating.\\
Level 4& Total &Total &8
	&\nameref{par:Auto-KO},\newline
		Difficult \nameref{par:Surgery vs}\nameref{bane:Brain Damage},\newline
		\nameref{par:Internal Bleeding}.
	&Severe skull fracture, brain damage, immediate unconsciousness.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead} &Blade cleaves through top of skull down to the teeth. Instant death.\\
%	\end{tabu}
%\end{table}
%	\marginpar{
%		\begin{sideways}
%			\qquad \quad Cutting
%		\end{sideways}
%	}
%
%\begin{table}[hb] %Upper Head - Piercing
%	\captionsetup{textformat=empty, labelformat=blank}
%	\caption{Upper Head - Piercing}\vspace{-15pt}
%	\label{wound:Upper Head - Piercing}
%	\begin{tabu} to \linewidth {X *{3}{X[-2,c]} X[3]X[4]}
	\\ 
%Wound Level &Stun &Pain &Bleed &Special &Description\\\toprule
		\\
	\\ 
	&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1 &0 &4 &0 &- &Raking glance to the side of the head. \\
Level 2 &1 &6 &3 &- &Deeper rake across scalp, lots of blood.\\
Level 3 &2 &10 &5 &\nameref{par:KO vs}3 RS &Skull cracked, possible unconsciousness. \\
Level 4 &Total &Total &20
	&\nameref{par:Auto-KO},\newline
		Miracle~\nameref{par:Surgery vs}\nameref{bane:Brain Damage},\newline
		\nameref{par:Internal Bleeding}
	&Skull penetrated, instantly unconscious.\\
\rowcolor{red!30} Level 5 &- &- &- &\nameref{par:Dead} &Point penetrates straight through skull into brain, instant death.\\
%	\end{tabu}
%\end{table}
%	\marginpar{
%		\begin{sideways}
%			\quad \qquad Piercing
%		\end{sideways}
%	}
%\begin{table}[!hb] %Upper Head - Bludgeoning
%	\captionsetup{textformat=empty, labelformat=blank}
%	\caption{Upper Head - Bludgeoning}\vspace{-15pt}
%	\label{wound:Upper Head - Bludgeoning}
%	\begin{tabu} to \linewidth {X *{3}{X[-2,c]} X[3.3]X[4.3]}
	\\ 
%Wound Level &Stun &Pain &Bleed &Special &Description\\\toprule
		\\
	\\ 
	&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 1& 4& 1& - &Blow to the side of the head. Ow.\\
Level 2& 3& 6& 2& \nameref{par:KO vs}1 RS &Temple is glanced. May lose consciousness.\\
Level 3& 5& 8& 4& \nameref{par:KO vs}5 RS &Skull fracture, may lose consciousness.\\
Level 4& Total &Total &6 
	&\nameref{par:Auto-KO},\newline
	Difficult \nameref{par:Surgery vs}~\nameref{bane:Brain Damage},\newline
	\nameref{par:Internal Bleeding}.
&Skull shatters, immediately unconscious.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead} &Skull crushed. Brain and bone everywhere. Death. \\
%	\end{tabu}
%\end{table}
%	\marginpar{
%		\begin{sideways}
%			\qquad Bludgeoning
%		\end{sideways}
%	}
%\begin{table}[!hb] %Upper Head - Unarmed
%	\captionsetup{textformat=empty, labelformat=blank}
%	\centering
%	\caption{Upper Head - Unarmed}\vspace{-15pt}
%	\label{wound: Upper Head - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[1.5]X[5]}
	\\ 
%Wound Level &Stun &Pain &Bleed &Special &Description\\\toprule
		\\
	\\ 
	&\nameref{par:Stun}&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Blow to the forehead. Negligible effect.\\
Level 2& 0& 0& 0& - &Graze to the side of the head. Painful but nothing special.\\
Level 3& 1& 1& 0& \nameref{par:KO vs}1 RS &Hard blow to the dome of the head. Bruising, possible blackout.\\
Level 4& 3& 3& 0& \nameref{par:KO vs}3 RS &Straight blow to the temple. Probably unconscious. \\
Level 5& Total& 10 &3 
	&\nameref{par:Auto-KO}, \newline
	\nameref{par:Internal Bleeding}.
&Severe blow to the temple, fractured bones, possible brain damage.\\
	\end{tabu}
\end{table}
%	\marginpar{
%		\begin{sideways}
%			\qquad Unarmed
%		\end{sideways}
%	}
	\clearpage

\section{Face}  \label{sec:face}
\begin{table}[!hb] %Face - Cutting
%	\caption{Face - Cutting}\vspace{-3pt}
%	\label{wound:Face - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3.8]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 6& 1& - &Cut vertically across the chin or lips. May leave a scar.\\
Level 2& 2& 8& 3& - &Deeper cut, possible chipping of jaw, may contact chin, lip, cheek, or brow.\\
Level 3& 4& 16& 8&Miracle \nameref{par:Surgery vs}~\nameref{bane:One-Eyed} &Slash catches eye! Severe damage, may lose eye.\\
Level 4& 6& 20& 18
	&Miracle \nameref{par:Surgery vs}~One-Eyed OR,\newline
	Miracle \nameref{par:Surgery vs}~Facial Deformity,\newline
	\nameref{par:Infection Risk} +2
&Face bisected, extreme damage to jaw, teeth, and skull. May lose eye (1-5) or nose (6-10), serious medical complications either way.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Head cloven through to the teeth, instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Face - Piercing
%	\caption{Face - Piercing}\vspace{-3pt}
%	\label{wound:Face - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]} %UGLY FACE PIERCING TABLE
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 5& 1& - &Facial scratch.\\
Level 2& 2& 6& 2& - &Point glances off facial bones, bleeding occurs, but nothing serious.\\
Level 3& 3& 8& 3& - &Point hits, but fails to penetrate facial bone, very painful, but still no significant damage.\\
Level 4& 5& 15& 10
	&Difficult~\nameref{par:Surgery vs}\nameref{bane:Mute}~OR, \newline
	Miracle~\nameref{par:Surgery vs}\nameref{bane:One-Eyed}~OR, \newline
	+5 to wound’s Bleed.
&Point penetrates face, either through mouth (1-3) eye (4-6) or facial bones (7-10.) Severe injury, lots of bleeding, extremely painful.\\ %CHECK THIS WOUND OUT IF WE WANNA SPREAD THE d10 LOCATION
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Head impaled, instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Face - Bludgeoning
%	\caption{Face - Bludgeoning}\vspace{-3pt}
%	\label{wound:Face - Bludgeoning}
%	\begin{tabu} to \linewidth {X[-1] *{3}{X[-2,c]} X[1.5]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1 &2 &4 &0 &- &Blow to the jaw, head rolls with it well. \\
Level 2 &3 &5 &2 &- &Firmer blow, blood in mouth, possible lost tooth. Youch. \\
Level 3 &5 &8 &3 &\nameref{par:KO vs}3 RS &Serious damage to jaw, fracture maybe, head jerked hard, may lose consciousness and several teeth.\\
Level 4 &7 &12&7 &\nameref{par:KO vs}5 RS &\nameref{bane:Mute} as the Bane until healed Broken jaw, almost certain unconsciousness.\\
\rowcolor{red!30}Level 5 &- &- &- &\nameref{bane:Dead} &Jaw smashed, skull shattered, neck broken. Instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Face - Unarmed
%	\caption{Face - Unarmed}\vspace{-3pt}
%	\label{wound:Face - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[1.5]X[5]}
\\
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1 &1 &0 &0& - &Grazing blow.\\
Level 2 &2 &0 &0& - &Firmer blow, might leave a bruise. \\
Level 3 &3 &2 &0 &\nameref{par:KO vs}2 RS &Hard hit, swelling!\\
Level 4 &4 &10&3 &\nameref{par:KO vs}5 RS &Broken nose or orbital fracture of eye socket. Severely debilitating and painful, probable KO.\\
Level 5& Total& 15 &3 &\nameref{par:Auto-KO} &Perfect hit, nose probably broken, Several teeth probably lost, instant KO.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Lower Head}  \label{sec:lower-head}
\begin{table}[hb] %Lower Head - Cutting
%	\caption{Lower Head - Cutting}\vspace{-3pt}
%	\label{wound:Lower Head - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[4.3]X[6]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 6& 1& - &Cross cut on the face, probably on the cheek, mouth or chin.\\
Level 2& 2& 8& 4& - &Direct hit on side of face, classic German dueling scar.\\
Level 3& 3& 12& 8& \nameref{par:KO vs}3 RS &Direct hit cleaves through cheek, damaging several teeth.\\
Level 4& 4& 18& 12
	&\nameref{par:KO vs}4 RS,\newline
		Difficult~\nameref{par:Surgery vs}\nameref{bane:Facial Deformity}.
	&Blade cuts cheek from corner of mouth to behind the ear, lots of teeth lost, cheek muscle damage.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Blade either cleaves through entire side of head, decapitating above the jaw, or most of the way through. Death is mercifully instantaneous.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Lower Head - Piercing
%	\caption{Lower Head - Piercing}\vspace{-3pt}
%	\label{wound:Lower Head - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 0& 4& 2& - &Grazed jaw.\\
Level 2& 1& 6& 4& - &Stabs through cheek, chipped tooth!\\
Level 3& 2& 9& 8& \nameref{par:KO vs}3 RS &Several teeth broken, cheek ripped open, jaw fractured.\\
Level 4& 3& 14& 14
	&\nameref{par:KO vs}5 RS,\newline
	Difficult~\nameref{par:Surgery vs}\nameref{bane:Mute}
&Jaw shattered, many teeth destroyed, may lose tongue.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Point slams through brainstem. Instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Lower Head - Bludgeoning
%	\caption{Lower Head - Bludgeoning}\vspace{-3pt}
%	\label{wound:Lower Head - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 3& 0& - &Tapped jaw, possibly bruised.\\
Level 2& 4& 4& 1& \nameref{par:KO vs}3 RS &Solid strike to jaw, jarring!\\
Level 3& 6& 6& 3& \nameref{par:KO vs}4 RS &Fractured jaw, possibly several lost teeth! \\
Level 4& 8& 10&4
	&\nameref{par:KO vs}5 RS,\newline
		Routine~\nameref{par:Surgery vs}~\nameref{bane:Facial Deformity}.
	&Broken jaw.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Neck Broken. Instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Lower Head - Unarmed
%	\caption{Lower Head - Unarmed}\vspace{-3pt}
%	\label{wound:Lower Head - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 2& 0& 0& - &Grazing blow to the jaw.\\
Level 2& 3& 0& 0& \nameref{par:KO vs}1 RS &Hit on the jaw. Better hope it ain’t glass... \\
Level 3& 5& 2& 0& \nameref{par:KO vs}3 RS &Good hit, head jerked to the side, KO is a distinct possibility. \\
Level 4& 10& 5& 0& \nameref{par:KO vs}8 RS &Jaw battered, significant chance of unconsciousness. \\
Level 5& Total &15 &3 &\nameref{par:Auto-KO} &Jaw broken, instant unconsciousness. \\
	\end{tabu}
\end{table}
	\clearpage

\section{Neck} \label{sec:neck}
\begin{table}[!hb] %Neck - Cutting
%	\caption{Neck - Cutting}\vspace{-3pt}
%	\label{wound:Neck - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 4& 4& - &Cut on the neck, painful but not deep.\\
Level 2& 2& 8& 6& - &Serious cut to the side of neck, passes dangerously close to windpipe.\\
Level 3& 3& 12& 12& Difficult~\nameref{par:Surgery vs}~\nameref{bane:Mute} &Slashed throat, possible vocal chord damage.\\
Level 4& 5& 14& 20& Miracle~\nameref{par:Surgery vs}~\nameref{bane:Mute}&Throat and jugular cut, spray of blood, needs a surgeon immediately.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Decapitation. Instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Neck - Piercing
%	\caption{Neck - Piercing}\vspace{-3pt}
%	\label{wound:Neck - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 5& 6& - &Puncture wound in neck, not good!\\
Level 2& 2& 5& 10& - &Deeper puncture wound, neck’s major artery!\\
Level 3& 3& 8& 20& Difficult~\nameref{par:Surgery vs}~\nameref{bane:Mute} &Punctured carotid or jugular, extreme bleeding! Possible vocal chord damage.\\
Level 4& 5& 12& 25& Miracle~\nameref{par:Surgery vs}~\nameref{bane:Mute} &Severed carotid or jugular, cracked vertebrae, vocal chord damage!\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Throat impaled, neck possibly broken. Exsanguination is almost instant. Death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Neck - Bludgeoning
%	\caption{Neck - Bludgeoning}\vspace{-3pt}
%	\label{wound:Neck - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 2& 0& - &Painful blow to neck, bruised flesh. \\
Level 2& 4& 3& 0& - &Heavy blow, momentary choking, significantly stunned.\\
Level 3& 6& 6& 3& - &Severely bruised windpipe, vocal cords damaged. \\
Level 4 &8 &Total &7 
	&\nameref{par:KO vs}3 RS, \newline
		Hard~\nameref{par:Surgery vs}~\nameref{bane:Mute}.
	&Severe damage to windpipe, brachial nerve damage, chipped spinal cord.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Snapped neck, instantaneous death. \\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Neck - Unarmed
%	\caption{Neck - Unarmed}\vspace{-3pt}
%	\label{wound:Neck - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 1& 0& 0& - &Chin too tucked in for a good hit, glances off.\\
Level 2& 1& 0& 0& - &Glancing blow to side of neck. Minimal effect.\\
Level 3& 3& 2& 0& - &Solid hit to throat. Might be hard to talk for a while.\\
Level 4& 5& 5& 0& \nameref{par:KO vs}3 RS &Severe hit to the throat, breathing is difficult. Possible unconsciousness.\\
Level 5& Total &15 &3 
	&\nameref{par:Auto-KO}, \newline
		HLT vs 3 RS vs Death.
	&Windpipe collapsed, instant unconsciousness, possible death.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Shoulder} \label{sec:shoulder}
\begin{table}[!hb] %Shoulder - Cutting
%	\caption{Shoulder - Cutting}\vspace{-3pt}
%	\label{wound:Shoulder - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 0& 4& 1& - &Shallow cut to the shoulder.\\
Level 2& 0& 6& 3& - &Deep cut, significant pain and bleeding.\\
Level 3& 1& 12& 6& \nameref{par:Autodrop} &Cleaves into the meat of shoulder, halts at collarbone.\\
Level 4& 2& 14& 8
	&\nameref{par:Autodrop}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Limb}.
	&Broken collarbone and cracked ribs, shoulder severely injured\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Cloven from shoulder to hip. Near or near-instant death.\\
%	\end{tabu}
%\end{table}
%\begin{table}[hb] %Shoulder - Piercing
%	\caption{Shoulder - Piercing}\vspace{-3pt}
%	\label{wound:Shoulder - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 0& 3& 2& - &Shallow stab to the shoulder. Painful, bloody, not serious.\\
Level 2& 0& 5& 5& - &Deeper stab, point halts in muscle.\\
Level 3& 1& 9& 7& - &Deep into the muscle, chipped collarbone.\\
Level 4& 2& 12& 12& - &Collarbone cracked, nicked artery, lots of blood.\\
Level 5& 5& 15& 20& - &Penetration into the subclavicular area. Artery severed.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Shoulder - Bludgeoning
%	\caption{Shoulder - Bludgeoning}\vspace{-3pt}
%	\label{wound:Shoulder - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 3& 0& - &Blow to the shoulder. No real damage.\\
Level 2& 4& 5& 2& - &Hard blow to the shoulder, severe muscle bruising.\\
Level 3& 6& 6& 4
	&\nameref{par:Autodrop}, \newline
		Routine~\nameref{par:Surgery vs}Ruined limb. %THIS bane doesn't exist
	&Dislocated shoulder, arm useless.\\
Level 4& 8& 10& 5
	&\nameref{par:Autodrop}, \newline
		Difficult~\nameref{par:Surgery vs}Ruined Limb &Shoulder broken, may not recover properly.\\ %RUINED LIMB?
Level 5& 12& 15& 12
	&\nameref{par:Autodrop}, \newline
		Miracle~\nameref{par:Surgery vs}Ruined Limb, \newline
		\nameref{par:Internal Bleeding}.
	&Shoulder collapsed, internal bleeding, recovery unlikely, death possible.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Shoulder - Unarmed
%	\caption{Shoulder - Unarmed}\vspace{-3pt}
%	\label{wound:Should - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Light thump, almost friendly.\\
Level 2& 0& 0& 0& - &Slightly painful, though nothing lasting.\\
Level 3& 1& 3& 0& - &Shoulder deadened, arm feels funny for quite some time.\\
Level 4& 2& 8& 0
	&\nameref{par:Autodrop}, \newline
		Difficult~\nameref{par:Surgery vs}~\hyperref[bane:Crippled Limb/Appendage]{Crippled Limb}.
	&Shoulder joint dislocated, tendons torn. Arm may never be the same again.\\
Level 5& 8& 18& 0
	&\nameref{par:Autodrop}, \newline
		Difficult~\nameref{par:Surgery vs}Ruined Limb. %RUINED LIMB?
	&Shoulder joint smashed, arm may twist in unnatural ways.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Chest}  \label{sec:chest}
\begin{table}[!hb] %Chest - Cutting
%	\caption{Chest - Cutting}\vspace{-3pt}
%	\label{wound:Chest - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 0& 4& 0& - &Shallow cut across chest, ribs do their job, no serious effect.\\
Level 2& 1& 5& 2& - &Deeper cut, raking ribs.\\
Level 3& 2& 10& 5& - &Cracked ribs, laceration within ribcage, no organ damage but very painful.\\
Level 4& 3& 15& 20& - &Ribs broken or cloven, bleeding in chest cavity.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Blade cleaves through ribs and into center of chest cavity. Lungs, heart, major blood vessels cut, near-instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Chest - Piercing
%	\caption{Chest - Piercing}\vspace{-3pt}
%	\label{wound:Chest - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 0& 3& 2& - &Attack rakes across ribs.\\
Level 2& 1& 4& 4& - &Minor penetration, point halted by ribs.\\
Level 3& 3& 7& 6& - &Penetration into chest, no organ damage but significant bleeding.\\
Level 4& 4& 14& 25& Difficult~\nameref{par:Surgery vs}Punctured Lung &Deep penetration into chest cavity! Extreme bleeding and organ damage!\\ %PUNCTURED LUNG?
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Heart or spine pierced, death is instantaneous.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Chest - Bludgeoning
%	\caption{Chest - Bludgeoning}\vspace{-3pt}
%	\label{wound:Chest - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 1& 3& 0& - &Blow to chest, shock absorbed by ribs.\\
Level 2& 2& 6& 0& - &Bruised ribs.\\
Level 3& 3& 8& 0& - &Cracked ribs, exceptional pain. \\
Level 4& 5& 12& 8& \nameref{par:Internal Bleeding} &Ribs broken, fragments lacerate flesh, severe bleeding internally\\
Level 5& Total& Total& 15 &\nameref{par:Internal Bleeding} &Ribcage shattered, possible organ damage, definite internal bleeding.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Chest - Unarmed
%	\caption{Chest - Unarmed}\vspace{-3pt}
%	\label{wound:Chest - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Blow to the chest. Trivial damage.\\
Level 2& 0& 1& 0& - &Good blow to the chest. Might leave a bruise.\\
Level 3& 1& 2& 0& - &Serious hit to the chest. Possible bone damage.\\
Level 4& 2& 5& 0& - &Rib cracking, significant damage.\\
Level 5& 3& 15& 5&\nameref{par:Internal Bleeding}&Severe rib damage, might have a rib in the lung, internal bleeding probable.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Side}  \label{sec:side}
\begin{table}[!hb] %Side - Cutting
%	\caption{Side - Cutting}\vspace{-3pt}
%	\label{wound:Side - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 0& 6& 2& - &Cut across the side and belly, painful but shallow.\\
Level 2& 1& 7& 4& - &Deeper cut, lots of blood.\\
Level 3& 2& 12& 6& \nameref{par:Infection Risk} +3 &Blade penetrates into side, serious but survivable. Possible organ damage.\\
Level 4& 3& 20& 20
	&\nameref{par:KO vs}5 RS, \newline
		\nameref{par:Infection Risk} +5.
	&Disembowelment, extreme damage and survival is close to unthinkable.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Individual is cloven nearly or totally in half. Death is instantaneous.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Side - Piercing
%	\caption{Side - Piercing}\vspace{-3pt}
%	\label{wound:Side - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 0& 4& 4& - &Shallow penetration to the side, bit of blood, nothing serious.\\
Level 2& 1& 5& 6& - &Penetration at an oblique angle, might be through-and-through, significant bleeding but no lasting harm.\\
Level 3& 2& 9& 9& - &Side pierced, severe bleeding but no organ damage.\\
Level 4& 3& 12& 14& \nameref{par:Infection Risk}~+2& Deep penetration, organ damage, severe bleeding, serious risk of infection. \\
Level 5& 5& 16& 18& \nameref{par:Infection Risk}~+5& Extreme penetration, possible exit wound, organ and intestinal damage!\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Side - Bludgeoning
%	\caption{Side - Bludgeoning}\vspace{-3pt}
%	\label{wound:Side - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 4& 0& - &Smacked in the side. Bad bruise, but that’s it.\\
Level 2& 4& 5& 0& - &Hard blow, winded, may cause nausea.\\
Level 3& 6& 7& 1& \nameref{par:Internal Bleeding} &Severe blow to side or belly, nausea and pain, but no internal damage.\\
Level 4& 8& 12& 5& \nameref{par:Internal Bleeding} &Blood in mouth, organ damage. Mercifully little risk of infection.\\
Level 5& Total& Total& 25 
	&\nameref{par:Infection Risk} +5, \newline
		\nameref{par:Internal Bleeding}.
	&Internal organs ruptured, survival would be nothing short of a miracle.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Side - Unarmed
%	\caption{Side - Unarmed}\vspace{-3pt}
%	\label{wound:Side - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 1& 0& 0& - &Grazing blow, little effect.\\
Level 2& 1& 1& 0& - &Blow absorbed mostly by muscle, little effect.\\
Level 3& 8& 8& 0& - &Liver blow! That’s gonna hurt tomorrow.\\
Level 4& 15& 10& 0& - &Severe liver blow! This would be a fight-ender for most people.\\
Level 5& 20& 15& 3& \nameref{par:Internal Bleeding} &Liver damaged, unbearable pain, possible internal bleeding.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Belly}  \label{sec:belly}
\begin{table}[!hb] %Belly - Cutting
%	\caption{Belly - Cutting}\vspace{-3pt}
%	\label{wound:Belly - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 5& 1& - &Cut straight up across navel. Not deep, but damn close.\\
Level 2& 3& 8& 4& - &Deeper cut, significant flesh wound, some muscle damage. Not pretty.\\
Level 3& 4& 10& 10& - &Serious cut belly, muscle damage, lots of bleeding, near organ damage.\\
Level 4& 5& 15& 15& \nameref{par:Infection Risk} +4 &Cut belly, intestinal and organ damage. Get a doctor. Get one fast.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Blade cleaves into belly, through organs and sternum. Instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Belly - Piercing
%	\caption{Belly - Piercing}\vspace{-3pt}
%	\label{wound:Belly - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 4& 3& - &Shallow stab to side of midriff. No real damage.\\
Level 2& 3& 6& 6& - &Deeper stab, muscle damage and lots of bleeding, but mercifully no organs hit.\\
Level 3& 5& 12& 10& \nameref{par:Infection Risk} +4 &Stabbed in the gut, organ and intestine damage. This is serious.\\
Level 4& 7& 14& 20& \nameref{par:Infection Risk} +4 &Severe internal damage. If this happened in a hospital today, your odds would be poor.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Point slams straight through guts to sever central artery of trunk. Exsanguination occurs within seconds, unconsciousness is instant.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Belly - Bludgeoning
%	\caption{Belly - Bludgeoning}\vspace{-3pt}
%	\label{wound:Belly - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 3& 4& 0& - &A solid blow to the stomach, winding but not serious.\\
Level 2& 6& 5& 0& - &A harder, bruising blow to the midriff.\\
Level 3& 9& 6& 0& \nameref{par:KO vs}3 RS &Severe blow to the stomach, brachial response might result in unconsciousness.\\
Level 4& 11& 7& 3& \nameref{par:KO vs}4 RS &Tremendous winding blow to the stomach, may vomit (END at 2), may lose consciousness.\\
Level 5& 15& 8& 5& \nameref{par:KO vs}6 RS &Possible organ damage, will be sick, will almost certainly lose consciousness.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Belly - Unarmed
%	\caption{Belly - Unarmed}\vspace{-3pt}
%	\label{wound:Belly - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 1& 0& 0& - &Gut shot.\\
Level 2& 3& 3& 0& - &A harder, bruising blow to the midriff.\\
Level 3& 5& 5& 0& \nameref{par:KO vs}3 RS &Severe blow to the stomach, brachial response might result in unconsciousness.\\
Level 4& 7& 7& 0& \nameref{par:KO vs}4 RS &Tremendous winding blow to the stomach, may vomit (END at 2), may lose consciousness.\\
Level 5& 9& 12& 3& \nameref{par:KO vs}6 RS &Possible organ damage, will be sick, will almost certainly lose consciousness.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Hip}  \label{sec:hip}
\begin{table}[!hb] %Hip - Cutting
%	\caption{Hip - Cutting}\vspace{-3pt}
%	\label{wound:Hip - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\rowfont[c]{}Wound~Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 5& 1& - &Shallow cut on the hip. No significant damage.\\
Level 2& 2& 7& 2& - &Deeper cut, bleeding, painful.\\
Level 3& 3& 12& 5& \nameref{par:Stability vs}3 RS &Blade chips hip bone, excruciatingly painful.\\
Level 4& 4& 16& 10& \nameref{par:Stability vs}4 RS &Broken hip, blade nicks artery.\\
Level 5& 5& 20& 20
	&\nameref{par:Auto-Prone}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Blade cleaves fully into hip, cuts artery, severe, and very likely lethal injury.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Hip - Piercing
%	\caption{Hip - Piercing}\vspace{-3pt}
%	\label{wound:Hip - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
\\
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 4& 1& - &Shallow penetration over the hip.\\
Level 2& 2& 6& 3& \nameref{par:Stability vs}1 RS &Deeper penetration, glances off bone.\\
Level 3& 3& 10& 6& \nameref{par:Stability vs}3 RS &Bone chips, but weapon does not penetrate deeper.\\
Level 4& 4& 12& 8& \nameref{par:Stability vs}4 RS &Hip fractures beneath penetration, severe bleeding.\\
Level 5& 5& 14& 12
	&\nameref{par:Stability vs}5 RS, \newline
		Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Hip is smashed, may never recover, potentially lethal.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Hip - Bludgeoning
%	\caption{Hip - Bludgeoning}\vspace{-3pt}
%	\label{wound:Hip - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 3& 0& - &Glancing blow to the hip.\\
Level 2& 3& 5& 1& \nameref{par:Stability vs}2 RS &Solid blow to hip, jarring and painful.\\
Level 3& 5& 10& 3& \nameref{par:Stability vs}3 RS &Cracked hip, bruised bone.\\
Level 4& 6& 12& 5
	&\nameref{par:Auto-Prone}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Hip broken, leg may be crippled permanently.\\
Level 5& 8& 18& 7
	&\nameref{par:Auto-Prone}, \newline
		Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Hip is destroyed. Leg probably won’t recover.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Hip - Unarmed
%	\caption{Hip - Unarmed}\vspace{-3pt}
%	\label{wound:Hip - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Grazing blow, on a very inopportune location.\\
Level 2& 0& 0& 0& - &Minor damage to the hip.\\
Level 3& 1& 1& 0& - &Bruised hip. It stings a bit.\\
Level 4& 2& 3& 0& - &Severely bruised hip. Still not much damage.\\
Level 5& 3& 6& 0& \nameref{par:Stability vs}3 RS &Possible fractured hip, loss of balance probable.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Groin} \label{sec:groin}
\begin{table}[!hb] %Groin - Cutting
%	\caption{Groin - Cutting}\vspace{-3pt}
%	\label{wound:Groin - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 2& 14& 2& - &Jimmies rustled, not a good day.\\
Level 2& 3& 16& 4& Difficult~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility} &Severe cut, significant damage. May lose consciousness.\\
Level 3& 5& 20& 8
	&\nameref{par:KO vs}3 RS, \newline
		Miracle~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}.
	&Excessive damage, blade strikes pelvis, internal bleeding and terrible damage.\\
Level 4& Total& Total& 20
	&\nameref{par:KO vs}5 RS, \newline
		\hyperref[bane:Barren/Sterility]{Sterility}.
	&Blade cleaves deep, artery cut, excessive damage to reproductive organs within and without.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Blade cleaves pelvis in twain, coming to rest deep in torso. Mercifully near-instant death.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Groin - Piercing
%	\caption{Groin - Piercing}\vspace{-3pt}
%	\label{wound:Groin - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2.3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 2& 12& 4& - &Your lizard brain cries, vehemently, ‘Do not make a habit of getting hit like this.’\\
Level 2& 3& 15& 8& Difficult~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}&Stabbed in groin. Excruciatingly painful, chance of reproductive organ damage.\\
Level 3& 4& 18& 15& Miracle~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility} &Deep stab in groin, nicks artery. Possible loss of reproductive capability.\\
Level 4& Total& Total& 25 
	&\nameref{par:KO vs}5 RS, \newline
		\hyperref[bane:Barren/Sterility]{Sterility}.
	&Artery severed, reproductive organs damaged beyond repair, unconsciousness likely.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Point slams through groin into abdomen, severing major artery and cracking pelvis. Death from shock and exsanguination follow instantly.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Groin - Bludgeoning
%	\caption{Groin - Bludgeoning}\vspace{-3pt}
%	\label{wound:Groin - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 4& 10& 0& - &A modest thump between the legs. Painful and debilitating.\\
Level 2& 6& 15& 0& \nameref{par:KO vs}2 RS &A harder blow, enough to cause bruising and extreme nausea. May lose consciousness.\\
Level 3& 8& 20& 3
	&\nameref{par:KO vs}3 RS, \newline
		Routine~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}.
	&Serious damage, reproductive organs may be damaged, surgical assessment advised. May lose consciousness.\\
Level 4& Total& Total& 10 
	&\nameref{par:KO vs}5 RS, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}.
	&Severe damage to reproductive organs, but also to pelvis and innards. Possible organ damage. May lose consciousness.\\
Level 5& Total& Total& 15
	&\nameref{par:Auto-KO}, \newline
		Miracle~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}.
	&The horrors borne within this wound level do not bear consideration. Severe internal organ damage, instant knockout, may never recover.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Groin - Unarmed
%	\caption{Groin - Unarmed}\vspace{-3pt}
%	\label{wound:Groin - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 2& 3& 0& - &The schoolyard standby, eh? Pain vanishes after 5 Rounds.\\
Level 2& 4& 5& 0& - &Blow to the groin, cringeworthy. Pain vanishes after 5 Rounds.\\
Level 3& 6& 8& 0& \nameref{par:KO vs}2 RS &Severe blow, agonizing.\\
Level 4& 8& 12& 0& \nameref{par:KO vs}5 RS &Hard blow, instant nausea, significant chance of losing consciousness.\\
Level 5& Total& Total& 3 
	&\nameref{par:Auto-KO}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Barren/Sterility]{Sterility}.
	&Mother of God.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Thigh}  \label{sec:thigh}
\begin{table}[!hb] %Thigh - Cutting
%	\caption{Thigh - Cutting}\vspace{-3pt}
%	\label{wound:Thigh - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 4& 2& - &Shallow cut on thigh.\\
Level 2& 2& 5& 4& - &Deeper cut on thigh. Superficial damage.\\
Level 3& 3& 12& 10& \nameref{par:Stability vs}3 RS &Serious cut cleaves to bone, anguish, possible knockdown.\\
Level 4& 4& 15& 20
	&Auto-prone,\newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
&Deep cut to thigh, chipped femur, severed artery. Medic!\\
Level 5& Total& Total& 25 
	&\nameref{par:Auto-Prone},\newline
		\nameref{par:Infection Risk} +5,\newline
		\nameref{par:Limb Severed}
	&Blade cleaves through the femur and either severs leg entirely or gets stuck in hip/pelvis. This is close to unsurvivable.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Thigh - Piercing
%	\caption{Thigh - Piercing}\vspace{-3pt}
%	\label{wound:Thigh - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 3& 4& - &Shallow penetration of thigh.\\
Level 2& 2& 5& 6& - &Deeper penetration of thigh, lots of blood.\\
Level 3& 3& 10& 10& \nameref{par:Stability vs}1 RS &Bone is glanced, artery nicked.\\
Level 4& 4& 12& 15& \nameref{par:Stability vs}3 RS &Femur cracked, artery severed.\\
Level 5& 5& 14& 20
	&\nameref{par:Auto-Prone}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Leg completely impaled, artery severed, bone broken. Automatic knockdown\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Thigh - Bludgeoning
%	\caption{Thigh - Bludgeoning}\vspace{-3pt}
%	\label{wound:Thigh - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 0& 0& - &Hit on thigh. Impact totally absorbed by dense muscle.\\
Level 2& 4& 2& 0& - &More solid hit, painful bruising on thigh, still nothing serious.\\
Level 3& 6& 4& 1& \nameref{par:Stability vs}2 RS &Severe bruising, minor femur damage, possible knockdown.\\
Level 4& 8& 10& 3
	&\nameref{par:Auto-Prone}, \newline
		Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Femur fractured, automatic knockdown.\\
Level 5& 10& 20& 5
	&\nameref{par:Auto-Prone}, \newline
		Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Compound fracture of femur. Unspeakably horrific, does not bear thinking about.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Thigh - Unarmed
%	\caption{Thigh - Unarmed}\vspace{-3pt}
%	\label{wound:Thigh - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Hit on thigh. Impact totally absorbed by dense muscle.\\
Level 2& 0& 0& 0& - &Hit on thigh. Impact mostly absorbed by dense muscle.\\
Level 3& 0& 1& 0& - &More solid hit, painful bruising on thigh, still nothing serious.\\
Level 4& 1& 2& 0& \nameref{par:Stability vs}~2 RS &Severe bruising, minor femur damage, possible knockdown.\\
Level 5& 2& 10& 3
	&\nameref{par:Auto-Prone}, \newline
		Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg}.
	&Femur fractured, automatic knockdown.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Knee}  \label{sec:knee}
\begin{table}[!hb] %Knee - Cutting
%	\caption{Knee - Cutting}\vspace{-3pt}
%	\label{wound:Knee - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 6& 0& - &‘Motherf-’\\
Level 2& 1& 8& 1& \nameref{par:Stability vs}1 RS &Deep cut to knee, painful contact with bone. May fall.\\
Level 3& 3& 10& 2& \nameref{par:Stability vs}3 RS &Chipped bone, jarring and painful.\\
Level 4& 5& 15& 3& \nameref{par:Auto-Prone}, \newline
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Knee shattered. Automatic Knockdown. Leg may never recover.\\
Level 5& Total& Total& 20 &
	\nameref{par:Auto-Prone}, \newline
		\nameref{par:Infection Risk} +3, \newline
		\nameref{par:Limb Severed}
	&Leg severed at knee. Knockdown automatic. Death by exsanguination likely to follow without medical attention.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Knee - Piercing
%	\caption{Knee - Piercing}\vspace{-3pt}
%	\label{wound:Knee - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 5& 2& - &A glancing stab to the knee.\\
Level 2& 1& 7& 4& \nameref{par:Stability vs}2 RS &Knee injured.\\
Level 3& 2& 9& 7& \nameref{par:Stability vs}4 RS &Knee fractures, pain is intense.\\
Level 4& 3& 16& 12& \nameref{par:Auto-Prone}, \newline
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee}. &Kneecap shattered. Probably won't recover. Automatic knockdown.\\
Level 5& 8& 20& 15& \nameref{par:Auto-Prone}, \newline
Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Knee shatters, leg bends wrong way. Automatic knockdown. Odds of recovery are dubious at best.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Knee - Bludgeoning
%	\caption{Knee - Bludgeoning}\vspace{-3pt}
%	\label{wound:Knee - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1 &2 &3 &0 &\nameref{par:Stability vs}2 RS &Glancing blow to the knee. Disruptive, but no major damage. May trip.\\
Level 2 &4 &5& 0& \nameref{par:Stability vs}3 RS &Solid blow to the knee, bruising, may trip.\\
Level 3 &6 &7& 2& \nameref{par:Stability vs}5 RS &Damaged knee, bruised or chipped bone, may fall down.\\
Level 4 &8 &10& 3& \nameref{par:Auto-Prone}, \newline
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Broken knee, automatic knockdown, may never recover.\\
Level 5 &Total &18 &5 &\nameref{par:Auto-Prone}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Knee smashed, automatic knockdown. Would take an act of God to recover.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Knee - Unarmed
%	\caption{Knee - Unarmed}\vspace{-3pt}
%	\label{wound:Knee - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Didn’t even feel it.\\
Level 2& 1& 0& 0& - &Unpleasant stinging sensation.\\
Level 3& 2& 5& 0&\nameref{par:Stability vs}5 RS &Knee hurt, possible knockdown.\\
Level 4& 3& 10& 0& \nameref{par:Auto-Prone}, \newline 
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Knee fractured, possible tendon/muscle tear. Could be a crippling injury. Extreme pain.\\
Level 5& 8& 18& 0& \nameref{par:Auto-Prone}, \newline
Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Knee} &Knee snapped backwards. Onlookers may faint. Leg will probably never recover.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Shin}  \label{sec:shin}
\begin{table}[!hb] %Shin - Cutting
%	\caption{Shin - Cutting}\vspace{-3pt}
%	\label{wound:Shin - Cutting}
	\begin{tabu} to \linewidth {X[-1] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 5& 1& - &Grazed shin, painful but no lasting harm.\\
Level 2& 2& 7& 4& \nameref{par:Stability vs}1 RS &Deep cut, bone contact.\\
Level 3& 3& 9& 7& \nameref{par:Stability vs}3 RS &Chipped shinbone, some blood, may fall.\\
Level 4& 4& 14& 10& \nameref{par:Auto-Prone}, \newline
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Deep cut cleaves into bone, automatic knockdown. May never recover.\\
Level 5& 5& 20& 15& \nameref{par:Auto-Prone}, \newline
\nameref{par:Infection Risk} +3, \newline
\nameref{par:Limb Severed} &Leg severed halfway down shin. Automatic knockdown.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Shin - Piercing
%	\caption{Shin - Piercing}\vspace{-3pt}
%	\label{wound:Shin - Piercing}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 4& 2& - &A light fracture to the shin.\\
Level 2& 2& 6& 5& \nameref{par:Stability vs}2 RS &Deep puncture, chipped bone.\\
Level 3& 3& 8& 8& \nameref{par:Stability vs}3 RS &Tendon grazed, leg freezes up. Possible bone fracture.\\
Level 4& 4& 10& 10& \nameref{par:Auto-Prone}, \newline
	Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Shinbone breaks, tendon damage. Severe muscle damage. Automatic knockdown.\\
Level 5& 5& 20& 17& \nameref{par:Auto-Prone}, \newline
Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Compound fracture of shinbone, leg ruined, may never recover. Automatic knockdown.\\
%	\end{tabu}
%\end{table}

%\begin{table}[hb] %Shin - Bludgeoning
%	\caption{Shin - Bludgeoning}\vspace{-3pt}
%	\label{wound:Shin - Bludgeoning}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 4& 0& - &Glancing blow to shin. No real damage, but man that stings.\\
Level 2& 3& 6& 0& \nameref{par:Stability vs}1 RS &Harder blow, may fall.\\
Level 3& 5& 8& 3& \nameref{par:Stability vs}4 RS &Fractured shin, agonizing, and may fall.\\
Level 4& 7& 14& 5& \nameref{par:Auto-Prone}, \newline
	Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Shin breaks, excessive pain. Automatic knockdown.\\
Level 5& 9& 16& 10& \nameref{par:Auto-Prone}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Compound fracture of the shinbone, excessive damage. Leg may never recover. Automatic knockdown.\\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Shin - Unarmed
%	\caption{Shin - Unarmed}\vspace{-3pt}
%	\label{wound:Shin - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Glancing blow to shin. No real damage, but man that stings. \\
Level 2& 1& 0& 0& \nameref{par:Stability vs}1 RS &Glancing blow to shin. No real damage, but man that stings. Might fall.\\
Level 3& 2& 5& 0& \nameref{par:Stability vs}3 RS &Bruising, minor damage to shinbone\\
Level 4& 3& 8& 0& \nameref{par:Stability vs}4 RS &Fractured shin, agonizing, and may fall.\\
Level 5& 6& 14& 0& \nameref{par:Auto-Prone}, \newline
	Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Ever seen Silva vs Weidman?\\
	\end{tabu}
\end{table}
	\clearpage

\section{Foot} \label{sec:foot}
\begin{table}[!hb] %Foot - Cutting
%	\caption{Foot - Cutting}\vspace{-3pt}
%	\label{wound:Foot - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 4& 0& - &Light cut on foot, unpleasant.\\
Level 2& 2& 5& 1& - &Deep cut into the meat of foot. Painful but not serious.\\
Level 3& 3& 6& 1& \nameref{par:Stability vs}3 RS &Serious damage to foot, possible bone fractures.\\
Level 4& 4& 10& 3& \nameref{par:Auto-Prone} &Broken foot, tendon damage, possible toe loss. Automatic knockdown.\\
Level 5& 5& 12& 10& \nameref{par:Auto-Prone}, \newline
	\nameref{par:Infection Risk} +3, \newline
	\nameref{par:Limb Severed} &Foot destroyed or severed at the ankle. Automatic knockdown, severe bleeding.\\
	%\end{tabu}
%\end{table}

%\begin{table}[hb] %Foot - Piercing
%	\caption{Foot - Piercing}\vspace{-3pt}
%	\label{wound:Foot - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 2& 2& - &Glancing blow to foot, no major damage.\\
Level 2& 2& 5& 3& \nameref{par:Stability vs}1 RS &Foot stabbed.\\
Level 3& 3& 10&4& \nameref{par:Stability vs}3 RS &Foot impaled, fracture of small bones.\\
Level 4& 4& 12&5& \nameref{par:Stability vs}5 RS &Foot is nailed to the floor.\\
Level 5& 5& 18&7& \nameref{par:Auto-Prone}, \newline
	Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Foot impaled, ankle damaged, tendons damaged. Automatic knockdown, may not regain functionality.\\
	%\end{tabu}
%\end{table}

%\begin{table}[hb] %Foot - Bludgeoning
%	\caption{Foot - Bludgeoning}\vspace{-3pt}
%	\label{wound:Foot - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 2& 0& - &Foot thumped. Unfortunately, feet are kind of designed for that. Minimal damage.\\
Level 2& 3& 3& 0& \nameref{par:Stability vs}2 RS &Foot hit harder, actual pain, may trip.\\
Level 3& 5& 7& 2& \nameref{par:Stability vs}4 RS &Equivalent to an ankle sprain, painful and debilitating.\\
Level 4& 7& 10& 3& \nameref{par:Auto-Prone}, \newline
	Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Foot is broken, automatic knockdown. May not recover, stay off it.\\
Level 5& 9& 18& 5& \nameref{par:Auto-Prone}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Foot is smashed entirely, multiple compound fractures. May not recover, amputation worth considering.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Foot - Unarmed
%	\caption{Foot - Unarmed}\vspace{-3pt}
%	\label{wound:Foot - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Foot stings for a minute.\\
Level 2& 0& 1& 0& \nameref{par:Stability vs}1 RS &Slightly more painful. Like, hopping up and down painful.\\
Level 3& 1& 3& 0& \nameref{par:Stability vs}2 RS &Foot bruised, actually painful.\\
Level 4& 2& 7& 0& \nameref{par:Stability vs}3 RS &Equivalent to an ankle sprain, painful and debilitating.\\
Level 5& 3& 10& 0& \nameref{par:Auto-Prone}, \newline
	Routine~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Leg} &Foot is broken, automatic knockdown. May not recover, stay off it.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Upper Arm}  \label{sec:upper-arm}
\begin{table}[!hb] %Upper Arm - Cutting
%	\caption{Upper Arm - Cutting}\vspace{-3pt}
%	\label{wound:Upper Arm - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 4& 1& - &Cut on arm. Trivial.\\
Level 2& 2& 7& 2& - &Painful cut on upper arm. No serious damage, but some blood.\\
Level 3& 3& 10& 3& - &Blade cuts through muscle, chips humerus.\\
Level 4& 4& 14& 6& Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Blade breaks humerus, arm is severely damaged, may never recover use of arm.\\
Level 5& 5& 15& 20&
	\nameref{par:Limb Severed} \newline
	\nameref{par:Infection Risk} +3 &Arm severed just below the shoulder! Extreme blood loss, seek help!\\
	%\end{tabu}
%\end{table}

%\begin{table}[hb] %Upper Arm - Piercing
%	\caption{Upper Arm - Piercing}\vspace{-3pt}
%	\label{wound:Upper Arm - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 4& 2& - &Shallow puncture on upper arm. Trivial injury.\\
Level 2& 2& 5& 3& - &Deeper puncture, some muscle damage, bleeding.\\
Level 3& 3& 8& 5& - &Penetration deep into humerus, terrible bleeding and extreme pain.\\
Level 4& 4& 12& 8&Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Grievous bone damage, going to be difficult to set.\\
Level 5& 5& 15& 15&Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}&Bone shattered, clipped artery, may never recover, extreme bleeding.\\
	%\end{tabu}
%\end{table}

%\begin{table}[hb] %Upper Arm - Bludgeoning
%	\caption{Upper Arm - Bludgeoning}\vspace{-3pt}
%	\label{wound:Upper Arm - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 2& 0& - &Bruise on the upper arm, no significant harm.\\
Level 2& 3& 3& 0& - &Harder blow, severe bruising on tricep.\\
Level 3& 5& 5& 1& \nameref{par:Autodrop} &Intense trauma, fractured humerus.\\
Level 4& 7& 9& 2& \nameref{par:Autodrop}, \newline
	Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Broken humerus, extreme pain, needs treatment.\\
Level 5& 9& 12& 5& \nameref{par:Autodrop}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}, \newline
	\nameref{par:Infection Risk} +2 &Humerus shattered, fragments driven into surrounding flesh, limb may be unsalvageable.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Upper Arm - Unarmed
%	\caption{Upper Arm - Unarmed}\vspace{-3pt}
%	\label{wound: Upper arm - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1 &0 &0 &0 &- &Two for flinching.\\
Level 2 &1 &0 &0 &- &Dead arm.\\
Level 3 &2 &  &0 &- &Bruise, nothing serious.\\ %MISSING NUMBER IN ORIGINAL DOC
Level 4 &3 &3 &0 &- &Bruised bone in the upper arm. Damn that hurts!\\
Level 5 &4 &5 &0 &\nameref{par:Autodrop} &Severe bruise, minor fractures in upper arm bones.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Elbow}  \label{sec:elbow}
\begin{table}[!hb] %Elbow - Cutting
%	\caption{Elbow - Cutting}\vspace{-3pt}
%	\label{wound:Elbow - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 5& 1& - &Shallow cut to elbow.\\
Level 2& 2& 6& 3& \nameref{par:Disarm vs}2 RS &Deeper cut to elbow, nerves nicked, may drop item.\\
Level 3& 3& 8& 5& \nameref{par:Autodrop} &Nerves cut, bone chipped, item automatically dropped. Arm will not function until healed.\\
Level 4& 4& 12& 7& \nameref{par:Autodrop}, \newline
	Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Nerves severed, arm nearly cloven off, bone broken. Item automatically dropped. Arm may not recover functionality.\\
Level 5& 5& 20& 20& \nameref{par:Autodrop}, \newline
	\nameref{par:Infection Risk} +3, \newline
	\nameref{par:Limb Severed} &Arm severed at the elbow, massive blood loss, limb irrecoverable, infection is a serious risk.\\
		%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Elbow - Piercing
%	\caption{Elbow - Piercing}\vspace{-3pt}
%	\label{wound:Elbow - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 4& 3& Disarm vs RS 1 &Painful stab in elbow. May drop item.\\
Level 2& 2& 6& 6& Disarm vs RS 3 &Deeper penetration, nerve injury, may drop item.\\
Level 3& 3& 10& 10& \nameref{par:Autodrop}, \newline
	Difficult~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Spike penetrates clear through elbow, automatic weapon drop, arm may not recover.\\
Level 4& 4& 14& 14& \nameref{par:Autodrop}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Extreme injury to elbow joint, nerves destroyed, arm loses functionality and may not recover.\\
Level 5& 5& 18& 20& \nameref{par:Autodrop}, \newline
	Arm automatically \hyperref[bane:Crippled Limb/Appendage]{crippled}, \newline
	\nameref{par:Infection Risk} +4 &Arm rendered totally unusable, will probably need to be amputated, definitely will not recover functionality.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Elbow - Bludgeoning
%	\caption{Elbow - Bludgeoning}\vspace{-3pt}
%	\label{wound:Elbow - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 4& 0& \nameref{par:Disarm vs}1 RS &Straight blow to the elbow. Arm goes temporarily numb, but no lasting damage.\\
Level 2& 4& 6& 0& \nameref{par:Disarm vs}3 RS &More severe impact, bruising, impact to nerves, painful tingling and numbness in whole arm.\\
Level 3& 6& 10& 1& \nameref{par:Disarm vs}5 RS &Fractured elbow, severe pain, hopefully no lasting damage.\\
Level 4& 8& 14& 3& \nameref{par:Autodrop}, \newline
	Easy~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Elbow broken or severely dislocated. Needs treatment.\\
Level 5& 10& 18& 6& \nameref{par:Autodrop}, \newline
	Miracle~\nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Elbow smashed to pieces, arm bent the wrong way, exquisite pain. Unfathomable horror.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Elbow - Unarmed
%	\caption{Elbow - Unarmed}\vspace{-3pt}
%	\label{wound:Elbow - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &No effect.\\
Level 2& 0& 0& 0& - &Still no effect.\\
Level 3& 1& 0& 0& - &Funnybone effect. Minor bruising.\\
Level 4& 2& 2& 0& - &Painful bruising, but still nothing serious. Arm might be numb for a bit.\\
Level 5& 3& 5& 0& \nameref{par:Autodrop} &Bruised bone and injured tendons. Arm won’t work quite the same for a while.\\
	\end{tabu}
\end{table}
	\clearpage

\section{Forearm}  \label{sec:forearm}
\begin{table}[!hb] %Forearm - Cutting
%	\caption{Forearm - Cutting}\vspace{-3pt}
%	\label{wound:Forearm - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
		Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
		Level 1& 1& 3& 1& - &Shallow cut on the forearm. Will be fine without treatment.\\
		Level 2& 2& 6& 3& - &Deep cut, painful but not debilitating.\\
		Level 3& 3& 10& 6& \nameref{par:Disarm vs}4 RS &Serious cut, muscle and tendon damage, possible fracture. May drop held item.\\
		Level 4& 4& 15& 8& \nameref{par:Autodrop}, \newline
		Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Broken bones, serious muscle damage, automatically drop held item, arm may not recover.\\
		Level 5& 5& 20& 20& \nameref{par:Autodrop}, \newline
	\nameref{par:Infection Risk} +3, \newline
	\nameref{par:Limb Severed}
&Arm cloven off just beneath the elbow. Serious blood loss.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Forearm - Piercing
%	\caption{Forearm - Piercing}\vspace{-3pt}
%	\label{wound:Forearm - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 2& 2& - &Puncture wound on the forearm.\\
Level 2& 2& 5& 5& - &Deep puncture on the forearm, more blood, no vital damage.\\
Level 3& 3& 7& 10& Disarm vs 2 RS &A deep puncture to the forearm, chipping bone, nicked vein. May drop item. \\
Level 4& 4& 9& 14& Disarm vs 3 RS &Veins are torn, blood everywhere. May drop item.\\
Level 5& 5& 12& 18& \nameref{par:Autodrop}, \newline
Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}, \newline
\nameref{par:Infection Risk} +4 &Complete penetration of the arm, blood loss serious. Item automatically dropped.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Forearm - Bludgeoning
%	\caption{Forearm - Bludgeoning}\vspace{-3pt}
%	\label{wound:Forearm - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 2& 0& - &Sharp blow to the forearm. That stings!\\
Level 2& 4& 4& 1& \nameref{par:Disarm vs}3 RS &Hard strike to the wrist, severe bruise, may drop held items.\\
Level 3& 6& 10& 3
&\nameref{par:Disarm vs}4 RS, \newline
Routine \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm} &Broken bones, possible snapped wrist. May drop held items.\\
Level 4& 8& 14& 6
	&\nameref{par:Autodrop}, \newline
	Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}
	&Forearm is shattered, automatically drop held items.\\
Level 5& 10& 15& 10
	&\nameref{par:Autodrop}, \newline
		Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}
		&Forearm crushed, staggering damage, automatically drop item, limb may never recover.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Forearm - Unarmed
%	\caption{Forearm - Unarmed}\vspace{-3pt}
%	\label{wound:Forearm - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Ow.\\
Level 2& 1& 0& 0& - &Ow!\\
Level 3& 2& 1& 0& - &Seriously, stop.\\
Level 4& 3& 2& 0& - &Agh! You jerk!\\
Level 5& 5& 8& 3
	&\nameref{par:Disarm vs}4 RS, \newline
		Routine \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Arm}
	&FFFFFFFF \\
	\end{tabu}
\end{table}
\clearpage

\section{Hand}  \label{sec:hand}
\begin{table}[!hb] %Hand - Cutting
%	\caption{Hand - Cutting}\vspace{-3pt}
%	\label{wound:Hand - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
Level 1& 1& 4& 1& - &Painful cut on hand.\\
Level 2& 2& 6& 3& \nameref{par:Disarm vs}2 RS &Deep cut, bruised bones, may drop held item.\\
Level 3& 3& 9& 5& \nameref{par:Disarm vs}5 RS &Broken hand, may drop held item.\\
Level 4& 4& 12& 8& \nameref{par:Autodrop} &Severe damage to hand, 1-3 fingers severed, automatically drop held items.\\
Level 5& 5& 15& 15
	&\nameref{par:Autodrop}, \newline
		\nameref{par:Infection Risk} +5, \newline
		\nameref{par:Limb Severed}
	&Hand destroyed or severed. Jet of blood erupts from wound.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Hand - Piercing
%	\caption{Hand - Piercing}\vspace{-3pt}
%	\label{wound:Hand - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1& 1& 3& 1& - &Glancing hit on back of hand.\\
Level 2& 2& 4& 2& \nameref{par:Disarm vs}1 RS &Deeper penetration of hand. May drop held items.\\
Level 3& 3& 6& 4
	&\nameref{par:Disarm vs}3 RS, \newline
		Routine \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand}
	&Hand impaled, item automatically dropped, may not recover.\\
Level 4&4&10&4
	&\nameref{par:Autodrop}, \newline
		Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand} 
	&Hand seriously damaged, probably won't recover.\\
Level 5& 5& 15& 18
	&\nameref{par:Autodrop}, \newline
		Miracle \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand}
	&Point tears through hand to forearm, tearing open vein. Extreme blood loss.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Hand - Bludgeoning
%	\caption{Hand - Bludgeoning}\vspace{-3pt}
%	\label{wound:Hand - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 2& 3& 1& - &Painful strike on hand.\\
Level 2& 4& 5& 2& \nameref{par:Disarm vs}2 RS &A severe blow to hand. Fractured phalanges. May drop items.\\
Level 3& 6& 6& 2
	&\nameref{par:Disarm vs}4 RS, \newline
		Routine \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand}
	&Hand smashed, items probably dropped.\\
Level 4& 8& 8& 3
	&\nameref{par:Autodrop}, \newline
		Difficult \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand}
	&Hand bones shattered, damage severe, item automatically dropped, may not recover.\\
Level 5& 10& 12& 8
	&\nameref{par:Autodrop}, \newline
		Miracle \nameref{par:Surgery vs}\hyperref[bane:Crippled Limb/Appendage]{Crippled Hand}, \newline
		\nameref{par:Infection Risk} +3
	&Hand completely smashed. Will probably need to be amputated. Time to consider a hook.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Hand - Unarmed
%	\caption{Hand - Unarmed}\vspace{-3pt}
%	\label{wound:Hand - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1& 0& 0& 0& - &Stings for a second.\\
Level 2& 0& 0& 0& - &Trivial bruising.\\
Level 3& 2& 1& 0& - &Bruised hand.\\
Level 4& 5& 5& 0& \nameref{par:Autodrop} &Fractured bones in hand.\\
Level 5& 10& 10& 0& \nameref{par:Autodrop} &Broken fist.\\
	\end{tabu}
\end{table}
\clearpage

\section{Upper Back}  \label{sec:upper-back}
\begin{table}[!hb] %Upper Back - Cutting
%	\caption{Upper Back - Cutting}\vspace{-3pt}
%	\label{wound:Upper Back - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
		Wound Level &Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
		Level 1& 2& 2& 2& - &Light lacerations, noticeable discomfort.\\
		Level 2& 4& 4& 4& - &Long cuts, heavy bleeding.\\
		Level 3& 7& 9& 6& - &Deep gashes, wounds noticeable. Scars that will never heal.\\
		Level 4& 10& 15& 8& \nameref{par:Stability vs}3 RS &Spinal fracture, shoulder blades cracked, no organ damage, but extreme pain.\\
\rowcolor{red!30} 		Level 5& - &- &- &\nameref{par:Dead}& Shoulders cleaved, spine severed. Spine destroyed. While death is not instantaneous, it will come within 24 hours.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Upper Back - Piercing
%	\caption{Upper Back - Piercing}\vspace{-3pt}
%	\label{wound:Upper Back - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
		Level 1& 1& 3& 2& - &Light piercing, minor bleeding, noticeable pain.\\
		Level 2& 3& 4& 5& - &Minor penetration, glances off shoulder blade. \\
		Level 3& 5& 9& 8& - &Penetrates through flesh, halted deep in shoulder blade. \\
		Level 4& 10& 13& 20& Miracle \nameref{par:Surgery vs}Punctured Lung &Point hits home, piercing deep and causing significant bleeding. Possible organ damage.\\ %PUNCTURED LUNG?
\rowcolor{red!30} 		Level 5& - &- &- &\nameref{par:Dead}& Spine severed or heart pierced, it’s all bad. Spine destroyed. While death is not instantaneous,it will come within 24 hours.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Upper Back - Bludgeoning
%	\caption{Upper Back - Bludgeoning}\vspace{-3pt}
%	\label{wound:Upper Back - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 4& 2& 0& - &Light bruising, minor displacement.\\
Level 2& 7& 3& 1& - &Heavy bruising, feels like something chipped.\\
Level 3& 10& 4& 2& \nameref{par:Stability vs}1 RS &Severe bruising, noticeable bleeding.\\
Level 4& 13& 5& 3
	&\nameref{par:Auto-Prone}, \newline
		Paralyzed (Not in Beta)
	&Something definitely made a cracking sound. Spine damaged, may never walk again.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead}& Spine completely destroyed, severe rupturing of internal organs, instant death.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Upper Back - Unarmed
%	\caption{Upper Back - Unarmed}\vspace{-3pt}
%	\label{wound:Upper Back - Unarmed}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
		&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
		Level 1& 0& 0& 0& - &Negligible effect.\\
		Level 2& 1& 0& 0& - &Still nothing.\\
		Level 3& 3& 1& 0& - &Bruising. Starting to hurt now.\\
		Level 4& 5& 2& 0& - &Severe bruising, starting to hurt, possible damage to spine and ribs.\\
		Level 5& 10& 5& 0& - &Minor fracture to ribs, bruised spine.\\
	\end{tabu}
\end{table}
\clearpage

\section{Lower Back}
\begin{table}[!hb] %Lower Back - Cutting
%	\caption{Lower Back - Cutting}\vspace{-3pt}
%	\label{wound:Lower Back - Cutting}
	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[2]X[5]}
		Wound~Level&Stun &Pain &Bleed &Special{\hfill \large Cutting ~~~} &Description\\\toprule
		Level 1& 3& 5& 1& - &Light lacerations, painful.\\
		Level 2& 5& 8& 5& - &Long cuts, significant bleeding.\\
		Level 3& 8& 10& 10& \nameref{par:Stability vs}2 RS &Deep gashes, wounds noticeable. Terrible scarring.\\
		Level 4& 12& 15& 15
		&\nameref{par:Auto-Prone}, \newline
			Paralyzed (Not in Beta)
		&Blow cracks spine, and cleaves muscle and flesh. Possible paralysis.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead} &Spine severed, severe intestinal damage, central artery in trunk severed. Near-instant death.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Lower Back - Piercing
%	\caption{Lower Back - Piercing}\vspace{-3pt}
%	\label{wound:Lower Back - Piercing}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Piercing~~~~}&\\\toprule
Level 1 &5 &2 &3 &- &Shallow penetration halted by back muscles.\\
Level 2& 7& 4& 6& - &Deeper penetration into the lower back, severe bleeding but no organ damage.\\
Level 3& 12& 10& 10
	&\nameref{par:Stability vs}1 RS, \newline
		\nameref{par:Infection Risk} +2
	&Deep penetration, possible organ damage, severe bleeding.\\
Level 4& 14& 12& 20
	&\nameref{par:Auto-Prone}, \newline
		\nameref{par:Infection Risk} +3
	&Deep penetration, kidney, liver, or intestinal damage, possible spinal damage.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead} &Total impalement, main artery of the trunk severed, major organ damage, instant death.\\
	%\end{tabu}
%\end{table}

%\begin{table}[!hb] %Lower Back - Unarmed
%	\caption{Lower Back - Bludgeoning}\vspace{-3pt}
%	\label{wound:Lower Back - Bludgeoning}
	%\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\ 
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Bludgeoning}&\\\toprule
Level 1& 5& 4& 0& - &Light bruising, shocking pain.\\
Level 2& 8& 5& 0& - &Heavy bruising, stunning pain.\\
Level 3& 10& 6& 1 
	&\nameref{par:Stability vs}3 RS, \newline
		\nameref{par:Internal Bleeding}
	&Blow to the kidneys from behind. Agony.\\
Level 4& 15& 7& 3
	&\nameref{par:Auto-Prone}, \newline
		Paralyzed (Not in Beta), \newline
		\nameref{par:Internal Bleeding}.
	&Spine damaged, footing lost instantly.\\
\rowcolor{red!30} Level 5& - &- &- &\nameref{par:Dead} &Spine broken just above tailbone, intestinal rupture, unconsciousness instant, death follows shortly. \\
%	\end{tabu}
%\end{table}

%\begin{table}[!hb] %Lower Back - Unarmed
%	\caption{Lower Back - Unarmed}\vspace{-3pt}
%	\label{wound:Lower Back - Unarmed}
%	\begin{tabu} to \linewidth {X[c] *{3}{X[-2,c]} X[3]X[5]}
	\\
&Stun&Pain&Bleed&\multicolumn{1}{r}{\large Unarmed~~~~}&\\\toprule
Level 1 &3 &1 &0 &- &Blow to the kidney. Painful.\\
Level 2 &5 &2 &0 &- &Severe blow to the kidney.\\
Level 3 &6 &4 &0 & \nameref{par:KO vs}2&Savage kidney blow. Possible unconsciousness.\\
Level 4 &5 &2 &0 & \nameref{par:KO vs}3&Agonizing kidney shot, unconsciousness likely.\\
Level 5 &10 &5 &1 
	&\nameref{par:KO vs}4, \newline
		\nameref{par:Internal Bleeding}.
	&Kidney damage! Jesus Christ!\\
	\end{tabu}
\end{table}
\end{document}
