\documentclass[a4paper, twocolumn, openany]{book}
\usepackage[top=2cm, bottom=2cm, left=1.2cm, right=1.2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[colorlinks=true, hidelinks]{hyperref} % Use \phantomsection\label{labelName} to add a link to the line below

% use for aligning columns in tables with m{} or b{}; not sure if I want that or keep it as is
%\usepackage{array}

% make new command to easily indent a paragraph. Have to use \tabto{0} after the paragraph, to reset the indentation.
\newlength{\indentlen}
\setlength{\indentlen}{7mm}
\newcommand{\tabto}[1]{\setlength{\leftskip}{#1\indentlen}}
%\setlength{\paperwidth}{50cm}

% change the default depth the ToC lists from 3 to four, thus including subsubsections and paragraphs in it
\setcounter{tocdepth}{5}

\begin{document}
\pagestyle{empty}
{\normalsize % if changing this, some passages need manual tuning. Mostly pagebreaks are made manually (\newpage and \twocolumn[{ ... }] that is)
\twocolumn[{
	{\centering
	{\LARGE
	Call of the Void:\\
	Ballad of the Laser Whales\\
	} % close Large 
	Alpha 1.7\\[\baselineskip]
	} % close centering
	{\itshape
	The ship moved silently through the dust of Underspace, like a viper in the grass. The crew sat
	or stood in the pale blue light of the ship’s interior, silently awaiting orders from their captain.
	The navigator, suspended in a tank of greenish gel, twitched occasionally as visual data was fed
	into his brain through the cords plugged into his skull. His voice, tinny and haunting, whispered
	course corrections into his handler’s ears through a headset, who then implemented the
	instructions through a series of dials and levers.\\
	Captain Voorstadt, seated on a raised throne near the back of the bridge, also listened in
	through a small speaker near his ear, and occasionally made his own corrections. Navigators
	were good at short-range plotting, shaving valuable seconds off the journey by steering around
	eddies and avoiding rogue currents, but they tended to become overstimulated and
	myopic--they lacked long-term planning. It took a more experienced crew than Voorstadt had to
	police the navigator without his help.\\
	He had taken command of Unterweltraumboot-9481 two weeks ago, after the scuttling of his old
	command to prevent the Eschatology Code from falling into enemy hands. He had almost been
	crucified by the brass for destroying his ship without a fight, but cooler heads had recognized
	that he had been right to do so, to deny the enemy access to vital codes.\\
	“Ships spotted, sir, bearing 345” the watchman’s voice came in tinny through the speaking tube.
	Voorstadt gestured to his XO to answer for him. The man rattled off a course adjustment to get
	a closer look. The Captain glanced at Vlatka, his Zell, who was idly sketching the
	communications officer in charcoal. She looked back.\\
	“What do the whales say?” he asked. She put down the sketch pad, and took his winecup out of
	his hand. She drank, and then smacked her lips a few times. Then she nodded.
	“Sad. No friends around. No shanties to warm the void. The sailors here are cold and bitter.
	They’re at war. It’s not a civilian ship.”\\
	“Good girl. Go pour yourself another,” he waved her away. She climbed up out of his lap and
	poured herself another drink from the pitcher on the counter. Voorstadt looked to the XO, who
	had gone over to check the periscope himself. He drew his eyes away and looked at the
	Captain grimly.\\
	“It’s a convoy, sir. Looks like two oil tankers and four destroyers. They’re big, sir.”
	“Four escorts for a pair of tankers? They must be desperate. Alert Muller and Eberard, and
	verify those targets.”\\

	Voorstadt and Muller’s U-Boats watched the convoy as Eberard caught up. He had been out of
	formation checking a ping up North. Fortunately nothing went awry in the time it took him to get
	back into place. The wolfpack waited until dark, and then slipped into the column, closing in on
	the tankers.\\

	On his bridge, Voorstadt waited for the green light from the other two ships before sending the
	order.\\

	“Muller, Eberard. Two eels each. We’ll overwatch.”\\

	The torpedoes exploded out of Underspace like flish leaping from a stream. They streaked for a
	time on momentum before the rockets kicked in, and they became incandescent streams of
	flame in the night sky. The two tankers had no chance to avoid them. Three of the four hit. The
	first tanker erupted into flames, its cargo igniting instantly in a monstrous cookoff that obliterated
	the vessel in a titanic explosion. The second was struck near the stern, its engines ruined, and
	began to slow as it lost power.The destroyers scrambled, pinging and rushing towards the
	sources of the torpedoes. Voorstadt watched the surviving tanker, heedless of the XO’s
	warnings.\\
	“Two eels, 5 degree dispersal, bearing 48. Fire.” he made a commanding gesture as his officers
	looked at him in astonishment. There was an instant’s silence, then they threw themselves to
	the task. The torpedoes launched five seconds later.\\
	“Sir, destroyers incoming, they know we’re here sir!”\\
	“Crash dive. Go to 120 meters,” Voorstadt said, turning his head slightly to accept a grape from
	Vlatka, who had been peeling them for him for the last twenty minutes, “take her down.”
	} % close textit 
}] % close twocolumn

\tableofcontents

\chapter{Introduction to Concepts}%\\[\baselineskip]
\pagestyle{myheadings}
\markboth{Call of the Void: Ballad of the Laser Whales}{Call of the Void: Ballad of the Laser Whales}

	{\noindent\bfseries Strahlenkanonenweltraumwal (Laser-Whale)\\}
	{\bfseries\itshape Imperator Astrocephalus\\}
	The Laser-Whale. Spanning anywhere from 120 to 12,000 meters, these creatures drift through
	open space, moving from star to star, feeding off of ice-asteroids, void-plankton, and
	occasionally consuming stars.\\

	Inside their massive heads there is a liquid, called Star-Oil, which can be used to purify
	Moonstone into a form usable as fuel for Voidships (called Whaleblood) and for countless other
	purposes. This oil-reservoir allows the Laser-Whale to survive in space by converting
	moonstone into biological energy for itself, as well as giving it a potent defense against
	predators in the form of the Strahlenkanone—the eponymous laser of the Laser-Whale.\\

	Though majestic, beautiful, and gentle, the Laser-Whales have since time immemorial been
	hunted by man. In their great voidships, star-whalers have risen up to harvest these great
	leviathans, slaying them with moonstone-tipped harpoons, and siphoning off their delicious,
	essential oil. The bones, which are naturally buoyant in air, are harvested for various purposes,
	and their flesh provides cheap sustenance for the poor. The population of whales has fallen,
	and with it the population of Star-Krakens has risen, as they are natural enemies. Nevertheless,
	the industry continues, to fuel the burgeoning industries of the Whaleblood Revolution.\\

	But some reject the temptation of easy energy and quick power. Void Sailors, powering their
	vessels not with Whaleblood, but with star-sails that catch the astral currents and carry them
	between worlds. Though slower, the Void Sailors are harder to detect, and do not require fuel.
	Pirates, smugglers, guardians of the whales, enemies of the State, the Void Sailors seek to
	push back the oppressive march of progress and keep the oceans free for all.\\

\newpage
\noindent
\section{Factions}

	\noindent
	{\large\bfseries The States\\}

	{\large Sacred Chiron\\}
	A fascist state born out of a failed attempt by the UC to uplift an ultra-religious feudal society.
	Chiron is militaristic, aggressive, and puritan to the extreme. Their practice of eugenics in the
	pursuit of “The Godly Man” have made them very unpopular, but they are nevertheless on the
	ascent, challenging the older States with aggressive and innovative military doctrines. They
	have also, to their credit, made enormous leaps in securing the great trade routes against
	piracy, at least in their waters.\\

	{\large The Unified Culture\\}
	A nominally peaceful post-scarcity (except for Whalebood) society, the people of the UC live in
	great vessels called Thoughtships which possess a sort of intelligence because of their unique
	Whalebone construction. The Thoughtships require a constant supply of Whaleblood to sustain
	the hedonistic, decadent lifestyles of their crews, and as such are constantly involved in
	territorial skirmishes with the other States. To preemptively resolve these conflicts, a secret
	service called Unpleasant Circumstances (UC, or UCUC) is employed to manipulate foreign
	politics and try to bring about regimes favorable to the Unified Culture.\\

	{\large The Hive\\}
	A recent revolution in a previously Legalistic society resulted in The Hive (called The People’s
	Republic by its own people, at least publically) a collectivist police state aimed at creating a
	perfectly harmonious society. The Hive may seem callous to foreigners with its draconian
	policies, and a populace essentially living in brutal serfdom, but alone of the States, The Hive
	has never experienced meaningful upheaval or internal strife since the time of the revolution.
	With the sheer industrial might of an incalculable mass of people behind it, The Hive’s only real
	Movement outside of the Hive) has pulled several attempted coups in countries like Chiron, the
	CIS and The Technocracy. Thusfar they have all been ruthlessly put down, but the UC and
	elements of The Hive support their endeavors in the hopes of sabotaging their rival states.\\

	{\large Pilgrims\\}
	Religious apostates from Chiron, the Pilgrim movement worship the same god as their
	parent-nation, but believe that no one organization should rule in matters of faith. The Pilgrims
	wander the Void now, seeking new lands free of religious oppression, establishing missions and
	colonies where they can, and seeking God among the endless stars. Known to practice a form
	of martial arts utilizing pistols called The Way of Five.\\

	{\large The Commercial League\\}
	A free association of many small mercantile states, the League is a purely capitalist organization
	that believes in the dissolution of governments and the end of imperialism. The League believes
	that free trade and commerce are anathema to warfare and conflict, and its people seek to open
	trade relations wherever possible, in the hopes that one day all peoples will be too busy
	enjoying the benefits of modern industry and commerce to think about killing each other.\\

	{\large The Zells\\}
	The Zells are an enigmatic race of humanoids with long pointed ears and strange behaviors.
	They rarely speak, but seem to be able to communicate with their crewmates non-verbally. They
	are intelligent, and very similar to humans--though they come off as a bit lethargic--and like most
	of the same things humans do: Food, drink, dancing, and violence. Here the similarities end.
	Zells are naturally drawn on the opposite direction by gravity that humans are--up into the sky.
	Their ships fly upside-down as a result, and they treat the Overspace much as normal folks treat
	the Underspace. They are also all, to a man, Void-Gifted. Their ships, the Zellislava, are
	terrifying vessels of great size that seem to be partially alive, despite having no flesh or blood.
	They also fly upside down, but steer themselves, and will fight to the last even after the entire
	crew is killed. Sometimes, depopulated Zellislavas can be seen dipping in and out of
	Underspace, wailing miserably and lashing out at anything that comes too close.\\

\newpage
\section{Technology}

	{\large Guns\\}
	Firearms exist about as advanced as modern day firearms. However, because penetrating the
	hulls of Voidships can have terrible consequences, most on-board weapons are pistol-caliber
	SMGs or carbines with frangible rounds.
	Special weapons exist that use magic to produce unusually powerful effects, but these and their
	ammunitions are rare.\\

	{\large Armor\\}
	Because firearms tend to be low-velocity and have poor penetration, armor is very effective
	against them. Ceramic plate armor can potentially resist a dozen rounds before one penetrates,
	giving armored men a great advantage in combat.\\

	{\large Melee Weapons\\}
	Moonstone Blades can penetrate any material like butter, which makes such weapons highly
	effective against armored enemies. However, closing to range is difficult when heavily armored,
	meaning that melee fighters must be lightly armored—a death sentence if they approach an
	enemy through their line of fire. Melee fighters must outflank in order to be effective.\\%[\baselineskip]

\section{The Void}

	{\large Overspace\\}
	Overspace is the high misty realm above the rest of the Void, where the air is thin and one can
	see the stars forever. Voidships cannot normally venture this high—Void Zeppelins are needed
	to reach this sort of altitude. Zeppelins are fragile and ill-suited for war, much less for whaling,
	so they are mostly used for exploration, observation, or as carriers for Voidships and Fighters.
	Overspace begins about 50 kilometers above the edge of Underspace.\\
	The Technocracy is known for its excellent Airships, though the Chironites’ Sternwaffe has a
	long and bitter rivalry with them.\\

	{\large Space\\}
	Space is the habitable central belt of the void, where most Voidships sail, and where most of the
	habitable islands reside. For analogy, consider this the surface world, with Overspace being the
	sky, and Underspace being underwater. Travel is fastest in this region of space, and requires
	the least fuel, so most vessels traverse here. Regular Space is a belt of 50 miles between
	Overspace and Underspace.\\
	The Unified Culture’s sentient warships are widely known to be the most powerful ships in
	regular space.\\

	{\large Underspace\\}
	Below Space there is a realm filled with dark energy that the human eye cannot penetrate.
	Here, a technology called Sonar is used to send ripples through this energy. Those ripples
	which return can identify the location of other vessels, allowing Underspace Ships to avoid
	collision… Or to detect enemies. At high altitudes just below regular Space, periscopes can be
	used to spy for enemies manually, since Sonar cannot extend above the murk. Underspace
	extends (as far as anyone knows) infinitely below Regular Space. Ships that venture too low
	tend to be destroyed by eldritch energies, or consumed by monsters. Sonar must be used
	sparingly, lest it attract the attention of these predators.\\
	Chiron is known for its powerful Underspace Navy, which it uses to annihilate the shipping of its
	rivals with surprise attacks.\\

	{\large Islands/Worlds\\}
	Mostly situated in Regular Space, these horizontally-stretching expanses of land act as the
	primary population centers of mankind and (most of) the other races. Some of them are
	thousands of kilometers wide and have fully functioning ecosystems.\\

\section{Magic}
	{\large Void Magic\\}
	Those who dwell on the edge of habitable space, who submerge themselves too
	often in the Underspace, sometimes find themselves developing strange powers. Contacted by
	mysterious entities, they become living conduits for terrible power. The ability to move between
	spaces instantly, change shape, slow down the passage of time, and even control the minds of
	others are all Void Magics.\\

	{\large Qin-Tao Magic\\}
	Qin-Tao Magic: A magic based on harnessing the internal power of one’s own body, this ancient
	form of magic requires no contract with otherworldly beings, and thus is less mercurial and dark
	in implication than Void Magic. However, it is also much more physically draining, and can be
	used less casually because it draws power from the user.\\

\chapter{Character Creation}
{\large\bfseries Differences from Song of Swords\\}

{\bfseries Toughness\\}
Toughness as a Compound Attribute no longer exists. All human characters have Toughness 4
at base.\\

{\bfseries Social Class\\}
Social Class no longer exists. All men are equal, on the sea. Except those with guns. People
with guns tend to be a bit higher up than unarmed men.\\

{\bfseries Ship Contribution\\}
Characters pool their Ship Contribution points together at Chargen to see what sort of vessel
they share.\\

{\bfseries Luck\\}
Luck points are a multi-purpose narrative power that allow characters to survive dangerous
situations, do awesome stuff, and avoid hideous death, just as you’d expect characters in a
game called “Ballad of the Laser Whales” to.\\

{\bfseries Magic\\}
For some reason designing magic for a pulp Age Of Sail/Science Fiction/Dieselpunk game is
easier than designing magic for a medieval swordplay simulator. As such, we already have a
system for Void Magic. You can still take 0 points in Magic.\\

{\bfseries Proficiencies\\}
Minor changes to how Talents work. You now only get Talents at set points in your total
Proficiency levels, rather than for individual Proficiencies. New Talents incoming.\\

{\bfseries Skills\\}
Don’t pretend that you care about skills. But if you do, skill packets are back.\\

{\bfseries Boons \& Banes\\}
New ones! You can only have -20 Bane Points, at all. If you have 1 PCP in B\&B, you can only
have up to +5 points from Banes.\\

{\bfseries Races*\\}
Optionally, a Player may spend 3 PCP in order to play a non-human race. This requires GM
permission -- non-humans may not be appropriate for your game.\\

{\large\itshape Zells\\}
Attributes: Agi +1, Per +2\\

\tabto{1}
	\underline{\bfseries The Zellish Touch}\\
	So long as a Zell is using a missile weapon with a rotating part (Revolvers, most
	commonly) they always add their Perception to their dice for missile attacks. This is
	cumulative with the effects of Aim and other modifiers.\\

	\underline{\bfseries Echolocation}\\
	Zells gain a +4 bonus to detecting opponents in Concealment, and suffers only half
	penalties for poor visibility and other vision-based penalties to Missile Attacks.\\

	\underline{\bfseries Racial Glossophobia}\\
	Zells have very keen senses of hearing, and this actually proves a disadvantage when
	communicating verbally, because the echoes of both their own and other voices makes it very
	difficult to pick out small details like words and inflections. As such, Zells often seem awkward,
	are prone to pausing in conversation to ‘sift’ through the words they have just heard, and also
	tend to prefer talking either very softly, or very loudly, with little room in between. Zells suffer a
	-4 pool penalty to CHA related tests involving communicating with others verbally. This does not
	apply for Intimidation checks, or for communicating with other Zells.\\

	\underline{\bfseries Race of Sailors}\\
	Zells gain a +2 racial pool bonus to Sailing tests. This bonus increases to +4 if the boat being
	operated is a Zellislava. This bonus decreases to -2 if the boat being operated is a Zellislava of a
	Chaos God hostile to the Zell in question.\\

	\underline{\bfseries Resistance to Disease}\\
	Zells cannot contract some diseases, including any transmitted by insects or rats, and some
	vitamin deficiencies, like scurvy. They are also curiously immune to botulism, though it does
	upset their stomachs.\\
% zell entry continued after printing the twocolumn

\twocolumn[{
\centering
	\begin{tabular}{lll}
		Campaign Power & Example & PCP \\ \hline \\
		Brutal, Grim Low-Poer & No Man's Land & 22\\
		Gritty Low Poer & Saving Private Ryan & 25\\
		Average Power & Enemy at the Gates & 28\\
		Savage High Power & Wolfenstein & 32\\
		Glorious High Adventure & Legend of the Galactic Heroes & 36\\
		\\\hline
	\end{tabular}\\[\baselineskip]
	\begin{tabular}{llllllll}
		PCP & Luck & Attributes & Skills & Profs & Wealth/Ship Contribution & Magic & Boons and Banes\\ \hline \\
		1 & 1 & 88 & 6 & 0 & 0 gp / 0 & 5 & -15\\
		2 & 2 & 92 & 9 & 3 & 5 gp / 1 & 10 & -10\\
		3 & 3 & 96 & 12 & 6 & 15 gp / 2 & 15 & -5\\
		4 & 4 & 108 & 15 & 9 & 25 gp / 3 & 20 & 0\\
		5 & 5 & 124 & 18 & 12 & 40 gp / 4 & 25 & 5\\
		6 & 6 & 140 & 21 & 15 & 80 gp / 5 & 30 & 10\\
		7 & 7 & 160 & 24 & 18 & 150 gp / 6 & 35 & 15\\
		8 & 8 & 180 & 27 & 21 & 300 gp / 10 & 40 & 20\\
		9 & 9 & 200 & 30 & 24 & 800 gp / 15 & 45 & 25\\
		10 & 10 & 224 & 33 & 27 & 1500 gp / 20 & 50 & 30\\
		\\ \hline
	\end{tabular}\\[2\baselineskip]
}]

% Zell entry continued
	\underline{\bfseries Strange Tastes}\\
	Zells suffer a -2 penalty to PER to detect poison (or anything else) in food or drink, because they
	have a dulled sense of taste. However, they may survive drinking seawater, and cannot suffer
	nausea from smells or tastes, no matter how vile.\\

	\underline{\bfseries Void Touched}\\
	Zells start with 5 free Magic Points which are automatically invested in Void Magic, making
	them Novice Void-Users.\\

	\underline{\bfseries Incompatible}\\
	Zells start with -2 Purity, to a minimum of 0.\\

\tabto{0}\noindent

\section{Attributes}
When creating a character, all Attributes start at 1. You receive a number of Temporary Arc
Points indicated on the Character Creation Table X.X in Chapter X, and may spend them to
upgrade your Attributes as-per Advancing with Arc. The higher an Attribute is at the start, the
more Arc it costs. To simplify this, here is a table with the total cost of upgrading an Attribute
from 1 at Character Creation.\\

The Attributes are:\\

{\centering
\noindent
\begin{tabular}{ll}
	Strength (STR)	& Willpower (WIL)	\\
	Agility	(AGI)	& Wit (WIT)			\\
	Health (HLT)	& Intelligence (INT)\\
	Endurance (END)	& Perception (PER)	\\
\end{tabular}\\[\baselineskip] } % close centering
From this existing Attributes, you gain a number of Compound Attributes that are used for
specialized tasks in combat and other circumstances.\\

\noindent{\small 
\begin{tabular}{cc}
	Adroitness (ADR) & Mobility (MOB) \\
	(Agility+Wit)/2  & (Strength+Agility+Endurance)/2 \\ \\ % may be a bit too wide for the column, but whatevs
	Carry Weight (CAR) & Charisma (CHA)\\
	Strength+Endurance & (Willpower+Wit+Perception)/2\\
\end{tabular}\\[\baselineskip] } % close small

{\centering
	\begin{tabular}{ll}
		Attribute Level & Arc Cost \\ \hline \\
		1					& - \\
		2					& 4 \\
		3					& 8 \\
		4 (Human Average)	& 12 \\
		5					& 16 \\
		6					& 20 \\
		7					& 24 \\
		8					& 28 \\
		9					& 34 \\
		10 (Human Maximum)	& 40 \\
		11					& 46 \\
		12 					& 52 \\
		13 (Mortal Maximum) & 58 \\
		\\ \hline
	\end{tabular}\\[\baselineskip] } % close centering

\section{Skills}

{\large Skill Packets\\}
Skill Packets cost 3 points at Chargen. Taking Packets multiple times increases the level of all
Skills within the Packet by 1, same with Packets that have overlapping skills.\\

{\centering
\begin{tabular}{llll}
	Skill Level & Skill Pool & Skill Bonus & Arc Price \\ \hline \\
	1  & +1  & Trained 	& 1\\
	2  & +2  & 			& 1\\
	3  & +3  & 			& 1\\
	4  & +4  & Teacher 	& 1\\
	5  & +5  & 			& 2\\
	6  & +6  & 			& 2\\
	7  & +7  & 			& 2\\
	8  & +8  & 			& 3\\
	9  & +9  & 			& 3\\
	10 & +10 & Mastery 	& 3\\
	\\ \hline
\end{tabular}\\[2\baselineskip] } % close centering

\hspace*{1cm}Available Skill Packets:\\

{\centering
\begin{tabular}{ll}
{\bfseries Aristocrat}					& {\bfseries Criminal}				\\
Knowledge: History						& Subterfuge						\\
Knowledge: Politics 					& Intimidate						\\
Persuasion 								& Knowledge: Streets				\\
Subterfuge 								& Gather Information				\\
\\
{\bfseries Sailor}						& {\bfseries Marine}				\\
Athletics								& Athletics							\\
Awareness								& Cooking, Crafting, or Ride 		\\
Profession 								& Drill 							\\
Sailing 								& Tactics (Foot) 					\\
\\
{\bfseries Scout}						& {\bfseries Doctor}				\\
Athletics								& Surgery x2 						\\
Stealth									& Persuasion 						\\
Awareness								& Knowledge: Science				\\
Tactics (Foot)							&									\\
\\
{\bfseries Academic}					& {\bfseries Officer}				\\
Knowledge x2 							& Persuasion, Ride, 				\\
Knowledge X2 							& or Navigation						\\
(pick two different						& Drill								\\
 knowledge-types)						& Leadership 						\\
										& Strategy							\\
										& Tactics (Pick one) 				\\
\\
{\bfseries Ensign}						& {\bfseries Spy}					\\
Sensors, Gunnery, 						& Stealth							\\
or Engineering							& Subterfuge						\\
Sailing									& Awareness							\\
Knowledge: Voidlore						& Gather Information				\\
Drill									& 									\\
\\
{\bfseries Assassin} 					& {\bfseries Pilot} 				\\
Athletics 								& Piloting 							\\
Awareness								& Navigation 						\\
Stealth									& Knowledge: Voidlore 				\\
Subterfuge								& Engineering 						\\
\\
{\bfseries Engineer} 					& {\bfseries Laborer} 				\\
Engineering x3 							& Profession x3 					\\
Knowledge: Science 						& Cooking 							\\
\\
\end{tabular}\\ }

\newpage
{\large Skill List\\}

\noindent
{\bfseries Athletics (Mob)}: Running, Climbing, and other stuff you’d do in PE.\\
{\bfseries Awareness (Per)}: Noticing shit happening around you.\\
{\bfseries Cooking (Int)}: Cooking fucking food.\\
{\bfseries Crafting (Int)}: Making stuff.\\
{\bfseries Drill (Wip)}: Military decorum and discipline.\\
{\bfseries Engineering (Int)}: Maintaining, repairing and managing machines.\\
{\bfseries Gather Information }(Cha): What it says on the tin.\\
{\bfseries Gunnery (Per)}: Coordinating the fire of the gunnery systems on a vessel.\\
{\bfseries Intimidate (Cha)}: Intimidating people into backing off, giving up information, etc.\\
{\bfseries Knowledge (Int)}: Streets/\linebreak[0]History/\linebreak[0]Politics/\linebreak[0]Science/\linebreak[0]Nature/\linebreak[0]Arcane/\linebreak[0]Voidlore\\
{\bfseries Leadership (Cha)}: Leading people.\\
{\bfseries Navigation (Int)}: Navigating places.\\
{\bfseries Perform (Cha)}: Performing for people.\\
{\bfseries Persuasion (Cha)}: Convincing people to do what you want.\\
{\bfseries Piloting (Vessel) (Per)}: Piloting vessels.\\
{\bfseries Profession (Wip)}: Doing some sort of job.\\
{\bfseries Riding (Agi)}: Riding a horse or other animal.\\
{\bfseries Sailing (End)}: Doing the menial work of making a boat move. For steering the vessel, see Piloting.\\
{\bfseries Sensors (Int)}: Detecting things using sonar, radar, and other ship industries.\\
{\bfseries Stealth (Agi)}: Not being seen or detected.\\
{\bfseries Strategy (Int)}: Directing troops over the course of a campaign, assembling troops in good order for a battle.\\
{\bfseries Subterfuge (Cha)}: Lying.\\
{\bfseries Surgery (Int)}: Keeping people from dying.\\
{\bfseries Survival (Int)}: Keeping yourself from dying in the wild.\\
{\bfseries Swimming (Mob)}: Swimming. Rarer than you’d think since there are no oceans.\\
{\bfseries Tactics (Foot) (Wit)}: The tactics of fighting in an infantry battle.\\
{\bfseries Tactics (Naval) (Wit)}: The tactics of fighting in a naval battle.\\
{\bfseries Tactics (Aerial) (Wit)}: The tactics of fighting in an aerial battle.\\

\section{Proficiencies}
Schools are no longer a thing. Proficiencies must be taken individually. However, all melee
Proficiencies have access to all maneuvers in their trees at level 1, and Talents are gained not
through individual proficiencies, but as a whole. When you get 4 Proficiency levels in any
combination of Proficiencies, you earn your first Talent, at 8 you earn your second, etc.\\

\noindent{\centering
\begin{tabular}{p{1.7cm}p{1.7cm}p{.06cm}lp{1.2cm}}
Proficiency Level 	& Proficiency Point / Arc Point Cost 	&\  & Talents 	& Total Prof Levels \\ \cline{1-2}\cline{4-5} &&\\
1 					& 1  									&\  & 1st 		& 4					\\
2 					& 1  									&\  & 2nd 		& 8					\\
3 					& 1  									&\  & 3rd 		& 12				\\
4 					& 1  									&\  & 4th 		& 16				\\
5 					& 2  									&\  & 5th 		& 20				\\
6 					& 2  									&\  & 6th 		& 24				\\
7 					& 2  									&\  & 7th 		& 28				\\
8 					& 2  									&\  & 8th 		& 32				\\
9 					& 3  									&\  & 9th 		& 36				\\
10					& 3  									&\  & 10th		& 40				\\ 
11					& 3  									&\  \\ \cline{4-5}
12					& 3  									&\  \\
13					& 6  									&\  \\
14					& 6  									&\  \\
15					& 6  									&\  \\
16					& 6  									&\  \\
17					& 10 									&\  \\
18					& 10 									&\  \\
19					& 10 									&\  \\
20					& 10 									&\  \\
 &&\\ \cline{1-2}
\end{tabular}\\[\baselineskip] }

{\large Buying Talents\\}
Once you have at least one proficiency at level 4 or higher, you may opt to buy talents
individually. As long as you meet the requirements for said talent, you may purchase it for 3 Arc.
This represents focusing on a single aspect of your shooting style, rather than getting better
overall.\\

{\large New Proficiencies and Defaults\\}
\noindent
{\bfseries Pistol} Defaults to SMG at -1, Shotgun at -2, Rifle at -3, and Machinegun at -3.\\
{\bfseries Submachine Gun} Defaults to Pistol at -1, Shotgun at -1, Rifle at -2, and Machinegun at -2.\\
{\bfseries Machinegun} Defaults to Rifle at -2, SMG at -2, Shotgun at -3, and Pistol at -3.\\
{\bfseries Shotgun} Defaults to Pistol at -2, SMG at -1, and Rifle at -1, Machinegun at -3.\\
{\bfseries Rifle} Defaults to Pistol at -3, SMG at -2, Shotgun at -1, and Machinegun at -2.\\
{\bfseries Rocket} No Defaults.\\
{\bfseries Flamethrower} No Defaults.\\

{\bfseries Talents do not apply when defaulting.\\}

\newpage
\section{New Talents}
{\bfseries\hspace*{\parindent} Fire and Maneuver}\\
{\bfseries Req}:  Agi 4\\
Effect: When Moving into Cover, you gain the Cover Bonus to Defense immediately instead of
at the beginning of the next Phase.\\

	\tabto{1}
	{\bfseries Watch Your Corners}\\
	{\bfseries Req}: Agi 6, Fire and Maneuver\\
	Effect: Your Base Defense increases by an additional 1 when moving.\\

		\tabto{2}
		{\bfseries Point Man}\\
		{\bfseries Req}:  Agi 6, Watch Your Corners\\
		Effect: Your Base Defense increases by an additional 1 when moving.\\ \hrule\ \\

\tabto{0} % reset indentation
{\bfseries Akimbo}\\
{\bfseries Req}: Agi 5\\
{\bfseries Effect}: When dual-wielding, the penalty for shooting is reduced by 2.\\

	\tabto{1}
	{\bfseries Macedonian}\\
	{\bfseries Req}:  Agi 5, Akimbo\\
	{\bfseries Effect}: When dual-wielding non-Single-Action pistols (both pistols must have more than
	3 rounds left in them) you may perform a Suppression or Autofire Action at ROF 4 by
	firing both at once. This always expends all ammunition in both pistols regardless of
	ROF, but only a maximum of 4 rounds can hit targets.\\ \hrule\ \\

\tabto{0} % reset indentation
{\bfseries Rebel Yell}\\
{\bfseries Req}: Wip 6\\
{\bfseries Effect}: When declaring a Charge against an opponent, for the rest of the Phase any shooting
attacks that character makes against you suffer a -4 penalty. If you enter Melee, you gain a +2
bonus to Initiative in the first Round of Melee Combat.\\

	\tabto{1}
	{\bfseries Give Them the Bayonet!}\\
	{\bfseries Req}: Wip 6, Rebel Yell\\
	{\bfseries Effect}: When making a Charge as-per “Rebel Yell,” all characters suffer the -4 penalty to
	attacking you, not just the character against whom you have declared the Charge.\\ \hrule\ \\

\tabto{0} % reset indentation
{\bfseries Rapid Reload}\\
{\bfseries Req}: Agi 4\\
{\bfseries Effect}: If you do not move in this Phase, you may Reload without using up your Action for the
Phase.\\

	\tabto{1}
	{\bfseries Rapid Unjam}\\
	{\bfseries Effect}: If you do not move in this Phase, you may Unjam your weapon without using up
	your Action for the Phase.\\ \hrule\ \\

\tabto{0}
{\bfseries Light Infantry Doctrine}\\ 
{\bfseries Req}: Per 4\\
{\bfseries Effect}: Your Defense Bonus from all Cover increases by 1.\\

	\tabto{1}
	{\bfseries Skirmisher}\\
	{\bfseries Req}: Per 6, Light Infantry Doctrine\\
	{\bfseries Effect}: Your Defense Bonus from all Cover increases by 1.\\ \hrule\ \\

\tabto{0}
{\bfseries Controlled Bursts}\\
{\bfseries Req}: Agi 4, Per 4\\
{\bfseries Effect}: When making a Burst Fire attack, your first BS causes two additional bullets to hit
instead of one.\\

	\tabto{1}
	{\bfseries Sustained Fire}\\
	{\bfseries Req}:  Agi 6, Per 6, Controlled Bursts\\
	{\bfseries Effect}: When firing Full Auto, you only use half of your weapon’s ROF in ammunition. If
	you would hit with more bullets than that number, you expend that many rounds of
	ammunition instead.\\

		\tabto{2}
		Hail of Lead\\
		{\bfseries Req}: Agi 8, Per 8, Sustained Fire\\
		{\bfseries Effect}: When firing Full Auto, your first BS causes two additional bullets to hit
		instead of one.\\ \hrule\ \\

\tabto{0}
{\bfseries Double Tap}\\
{\bfseries Req}: Agi 4\\
{\bfseries Effect}: When making a Regular Fire action, you may choose to only roll Hit Location once, and
have both shots (assuming they hit) land on that location.\\

	\tabto{1}
	{\bfseries Triple Tap}\\
	{\bfseries Req}: Agi 6, Double Tap\\
	{\bfseries Effect}: When making a Regular Fire action, you may fire three times instead of two. If
	you use Double Tap, only the first two shots hit the same location automatically, the third shot is
	still random.\\

		\tabto{2}
		{\bfseries Mozambique Drill}\\
		{\bfseries Req}: Agi 7, Triple Tap\\
		{\bfseries Effect}: When making a Regular Fire action using Triple Tap, you may choose to
		have your first two attacks (assuming they hit at all) automatically hit the Chest (Vitals),
		and the third automatically hit the Head, without rolling for Hit Location. You cannot be
		further than 10m away from your target.\\

		{\bfseries Hand of the Devil}\\
		{\bfseries Req}: Agi 7, Triple Tap\\
		{\bfseries Effect}: When firing a weapon that is fed from a cylinder, tube or barrel (not from
		a magazine) you may reload one round into the weapon as part of a Regular Fire action.
		This may occur at any point in the action--for example, if you have a double barrel
		shotgun, you may fire twice, then reload one round and fire a third time using Triple Tap.\\ \hrule\ \\

\tabto{0}
{\bfseries Shotgun Surgery}\\
{\bfseries Req}: Agi 6\\
{\bfseries Effect}: When making an attack with a shotgun with Shot, you may chose to forego the
scatter-damage to adjacent Hit Locations, and instead add increase the weapon’s base damage
by half for the primary attack.\\

	\tabto{1}
	{\bfseries This is my Boomstick}\\
	{\bfseries Req}: Agi 6\\
	{\bfseries Effect}: When making an attack with a shotgun with Shot within 1 Range Increment, your
	scatter-damage may land on the Critical Hit location without being redirected to the
	Chest location as normal.\\ \hrule\ \\

\tabto{0}
{\bfseries That was Left Handed!}\\
{\bfseries Req}: Per 6\\
{\bfseries Effect}: The Aim action provides a +3 bonus to attacks per Phase spent aiming, instead of +2.\\

	\tabto{1}
	{\bfseries Yes, I am a Wizard}\\
	{\bfseries Req}: Per 8, That was Left Handed!\\
	{\bfseries Effect}: If you aim for 3 Phases at a target within the base Range of your weapon, that
	target’s Cover bonus to Defense is halved for this attack. This does not apply to targets that
	have Full Cover.\\ \hrule\ \\

\tabto{0}
{\bfseries I am Huge}\\
{\bfseries Req}: Str 6\\
{\bfseries Effect}:  If you do not move in an Action, you count as Bracing for the purposes of Brace.\\

	\tabto{1}
	{\bfseries I am In Control}\\
	{\bfseries Req}:  Str 8, I am Huge\\
	{\bfseries Effect}: You always count as Bracing for the purposes of Brace.\\

		\tabto{2}
		{\bfseries I will Crush Them}\\
		{\bfseries Req}:  Str 8, I am In Control\\
		{\bfseries Effect}: You gain +2 dice to all attacks when Suppressing.\\ \hrule\ \\

\tabto{0}
{\bfseries Nope!}\\
{\bfseries Req}: Hlt 6\\
{\bfseries Effect}: When using a Riot Shield, you gain +2 AV against all attacks.\\ \hrule\ \\

\newpage % manual pagebreakma
	\tabto{1}
	{\bfseries Gave Peace a Chance}\\
	{\bfseries Req}: Hlt 6, Nope!\\
	{\bfseries Effect}: When using a Riot Shield, so long as you move towards the enemy in this
	Phase, your shield provides you with an additional +1 to Defense.\\

		\tabto{2}
		{\bfseries I’m Fucking Invincible!}\\
		{\bfseries Req}: Hlt 8, Gave Peace a Chance\\
		{\bfseries Effect}: When using a Riot Shield, you are immune to Critical Hits, so long as you
		exclaim loudly that you are, in fact, invincible at the start of the Round. Critical
		Hits count as hitting the Chest instead.\\ \hrule\ \\

\tabto{0}
{\bfseries Chironite Style}\\
{\bfseries Req}: Agi 4\\
{\bfseries Effect}: When using a single pistol at Point Blank Range, you do not suffer the -2 penalty
normally incurred by the weapon’s length.\\

	\tabto{1}
	{\bfseries SMK Close Stance}\\
	{\bfseries Req}: Agi 4, Chironite Style\\
	{\bfseries Effect}: When using an SMG or Shotgun at Point Blank Range, you do not suffer the -3
	penalty normally incurred by the weapon’s length. You suffer a -1 penalty instead.\\

		\tabto{2}
		{\bfseries Hip-Fire}\\
		{\bfseries Req}: Agi 4, SMK Close Stance\\
		{\bfseries Effect}: When shooting a Rifle at Point Blank Range, you do not suffer the -6
		penalty normally incurred by the weapon’s length. You suffer a -2 penalty
		instead.\\ \hrule\ \\

\tabto{0}
{\bfseries Advance!}\\
{\bfseries Req}: Agi 6\\
{\bfseries Effect}: You suffer only a -2 penalty to shooting while moving.\\

	\tabto{1}
	{\bfseries Stormtrooper}\\
	{\bfseries Req}: Agi 8\\
	{\bfseries Effect}: You suffer no penalty for shooting while moving.\\

		\tabto{2}
		{\bfseries Blitzkrieg}\\
		{\bfseries Req}: Wip 8, Stormtrooper\\
		{\bfseries Effect}: When advancing towards an enemy through Suppression fire, the
		distance you can move before they get a +2 bonus to attack is doubled from 3m
		to 6m.\\ \hrule\ \\

\tabto{0}
{\bfseries Fancy Shooting}\\
{\bfseries Req}: Agi 6, Per 6\\
{\bfseries Effect}:  When making a Regular Fire action with a weapon chambered in Pistol Rounds, you
may try to ricochet an attack around cover at your target. If you do so, your attack does not
benefit from its AP bonus, but your opponent’s Cover bonus to defense is reduced by 2, and all
Hit Locations can be hit. This does not apply to Full Cover.\\

	\tabto{1}
	{\bfseries Pretty Good}\\
	{\bfseries Req}:  Agi 8, Per 8, Fancy Shooting\\
	{\bfseries Effect}:  When using Some Fancy Shooting, you may reduce your opponent’s Cover
	bonus to defense by 4 instead of 2.\\

		\tabto{2}
		{\bfseries Never let your prey escape}\\
		{\bfseries Req}:  Agi 10, Per 10, Pretty Good\\
		{\bfseries Effect}:  When using Pretty Good, you may ignore your opponent’s entire Cover
		Bonus instead of just 4.\\
		Additionally, you may now hit characters in Full Cover as though they were not in
		cover at all, but you suffer a -8 penalty to doing so.\\ \hrule\ \\

\tabto{0}
{\bfseries Ranch Hand}\\
{\bfseries Req}:  Agi 5\\
{\bfseries Effect}:  Reduce the penalty for fan-firing a Single-Action weapon by 2.\\

	\tabto{1}
	{\bfseries Cowboy}\\
	{\bfseries Req}:  Agi 6, Ranch Hand\\
	{\bfseries Effect}:  Reduce the penalty for fan-firing a Single-Action weapon by 2.\\ \hrule\ \\

\tabto{0}
{\bfseries I See You}\\
{\bfseries Req}:  Per 6\\
{\bfseries Effect}:  When shooting at a target in a low-light or other obscured environment, if that target has
fired a weapon or otherwise created a visible flash of light surrounding it in the Phase thus far,
halve all penalties to your attack against that character.\\ \hrule\ \\

{\bfseries Work The Bolt}\\
{\bfseries Req}: Agi 6\\
{\bfseries Effect}:  When Regular Firing a Single-Shot weapon, you may now fire up to two shots. However,
the second shot is at a -3 penalty, and cannot benefit from Aim. Weapons with Reliable only
suffer a -2 penalty, while weapons with Jam suffer -4.\\ \hrule\ \\

{\bfseries Off-Hand Training}\\
{\bfseries Req}: Agi 4\\
{\bfseries Effect}: The penalty for firing a weapon from your off-hand/non-dominant side is reduced from -4
MP to -2 MP.\\ \hrule\ \\

\newpage % manual pagebreak

\section{Luck}

You know what’s not very heroic? Swinging onto the deck of a Chironite warship on a grappling
hook and immediately being hosed down by six members of the SMK with submachineguns.
Luck is what distinguishes dead idiots from heroes. Every character has an amount of luck
per-session that they can use to change their fortune. Luck can be used in the following ways:\\

{\large Salvation\\}
If you’re about to die in some way which is avoidable, but which you have failed to avoid either
through a botched skill roll or some other factor, you may spend three Luck points to survive by
the skin of your teeth. This may also apply in events which are not necessarily life-threatening,
but extremely tense and with high consequences for failure (i.e., getting caught pickpocketing a
king)\\

{\large Lucky Son of a Bitch\\}
If you’re in a situation where you’re in a large group of potential targets and an enemy is
targeting you, you may, with the GM’s approval, spend 1 Luck to have the enemy instead target
someone else. It’s fine, you can always get more friends. This probably won’t work if the guy
targeting you very specifically wants to kill you.\\

{\large Wound Reduction\\}
If you suffer a Wound that is not instantly lethal, you may spend a Luck point to reduce it by 2
stages. For instantly lethal wounds, you may spend 2 Luck to reduce them by 1 stage, and no
more.\\

{\large Lucky Shot\\}
When you make an attack of any sort, after you roll you may spend a point of Luck to roll 4
additional dice and add any successes to the total. Limit once per attack.\\

{\large Lucky Break\\}
When you make a skill test of any sort, after you roll you may spend a point of Luck to roll 4
additional dice and add any successes to the total. Limit once per test.\\

{\large Comedic Timing\\}
If something really, really appropriate could happen right now, and your GM agrees, you can
spend 2-5 points of Luck to cause that thing to occur (GM decides the exact cost).\\

{\itshape Ex: Paladin Charon has been yelling at the scoundrel Vick not to play with the Kriegsmesser he
keeps on his desk for weeks. Charon has now revealed himself to be a villain, and is holding the
party at gunpoint. with that same pistol. Vick’s player notices this and asks the GM if his
character could have broken the firing pin on Charon’s gun while playing with it a week ago and
tried to hide that fact to avoid getting in trouble. The GM decides that this would be hilarious,
and assigns it a cost of 3 Luck. Vick pays the cost, and steps forward. Charon grins evilly, and
pulls the trigger...}\\

\section{New Boons \& Banes}

{\Large\bfseries Boons\\}

{\bfseries Rank (3/6/9)\\}
You hold a rank in some organization or government. This gives you a degree of authority,
respect or privileges in that role.\\
3: Police Sergeant\\
6: Army Lieutenant\\
9: Captain of the Queen’s Navy\\
\hrule\ \\

{\bfseries Crack Shot (6/12/18)\\}
You’re a born marksman. It might be a mind for trajectories, a knack for leading targets, or
maybe you just like shooting people. Regardless, you know how to shoot on a purely instinctual
level.\\
6: Add +1 to your MP.\\
12: Add +2 to your MP\\
18: Add +3 to your MP\\
\hrule\ \\

{\bfseries Genetic Marvel (10)\\}
Your body is exceptionally compatible with whale-nerve prosthetics. Your Purity is now
(Willpower+Health+Misc.)*2.\\
\hrule\ \\

{\bfseries Big Guy (15)\\}
You’re a pretty big guy. For you this means you’re harder to hurt and more likely to recover from
injuries. Your Toughness is 5, instead of 4, and you enjoy a +2 bonus on Infection and
Bloodloss rolls. This is not compatible with Scrawny or Skinny.\\
\hrule\ \\

{\bfseries Secret Agent (Good Guy) (3/6/9)\\}
You secretly work for a government to help the party. They don’t need to know about it (and
they probably shouldn’t) but you can call in favors from your government to help them achieve
their goals.\\
3: The government considers it of minor concern but can get the party off of minor crimes,
arrange indirect support, etc.\\
6: The government considers the party relevant to its plans, and will help them as much as it
can without openly showing its hand.\\
9: The government has staked the success of serious plans on the party’s success, and if needs
be will send a squadron of warships to bail them out.\\
\hrule\ \\

{\bfseries Monster Hunter (6)\\}
You’re not afraid of Star-Krakens, Laser Whales, Leviathans, or Zellislava. You’ve seen worse.
Where? Don’t ask. You only suffer half the normal Panic Penalty when facing a Monster.\\
\hrule\ \\

{\bfseries I Know A Guy (3)\\}
You know every scoundrel, hooligan and rat bastard this side of Black Reach. This Boon is sort
of like Contacts, but it applies specifically to scumbags and criminals, and also allows you to call
in favors, within GM discretion.\\

{\Large\bfseries Banes\\}

{\bfseries Duty (-2/-4/-6)\\}
You’ve got obligations, and these may conflict with your desires.\\
-2: You are generally autonomous but have some obligations (detective, privateer, ranger)\\
-4: You answer directly to an authority figure but still have a long leash (officer, ship captain)\\
-6: You are part of a rigid command chain and are more or less constantly on orders (soldier,
crewman, police officer)\\
\hrule\ \\

{\bfseries Demi-Zell (-10)\\}
You’re part Zell. If you’re a woman, you’re probably unusually tall. Your ears are tapered and
long. A faint smell like quinine lingers around you no matter how often you bathe. Zells are
almost universally distrusted by civilized peoples in the Void, mostly because they’re weird, and
have a reputation for violence, piracy, and generally antisocial behavior. No major State can
afford to be openly hostile to Zells, as they are too important to maritime politics, but that doesn’t
mean anyone has to like them. Being part Zell is the worst of both worlds. Humans distrust and
despise you, both because you resemble a Zell, and also because you represent a betrayal of
your race by your human parent. Zells pity and patronize you. You have neither the eternal life
of the Zell, nor the endless ambition and mortal vigor of the human. What you do have is human
foolishness, Zellish laziness, human rage, Zellish glossophobia, and worst of all--Zellish
tastebuds. It’s awful, you can’t taste sour at all.\\
You suffer -4 to any Charisma test with normal humans, and they are never likely to fully trust
you or recognize your achievements. You suffer no mechanical penalty with Zells, but they will
never fully trust you, and will treat you like a child no matter how high you rise in the world.\\
\hrule\ \\

{\bfseries Alcoholic (-10/-15)\\}
You drink. A lot. Maybe a bit too much. Okay definitely too much, but you can stop any time you
want. You just don’t want to. Why would you? The world is so gray all the time… And a bit of
drink helps! You also only drink with friends, okay? What friends? You’ve never met Jack? Oh
man, you’ll love Jackie-D, let me go find him...\\
-10: Any time combat occurs, you have a 30\% chance to be inebriated, a 30\% chance to be
drunk, and a 40\% chance of being sober.\\
When Inebriated, you suffer a -2 penalty to all attack rolls (shooting and in melee) and your
Mobility is reduced by 2. You also suffer a -2 penalty to all Skill rolls.\\
When Drunk, you suffer a -4 penalty to shooting rolls, a -2 penalty to melee rolls, your
Perception is reduced by 2, and your Mobility is reduced by 2. You also suffer a -4 penalty to all
Skill rolls.\\
-15: Unless someone physically stops you from drinking, you will be Inebriated at all times. You
will be Drunk 60\% of the time. The penalties are as above. If you are prevented from drinking for
more than 24 hours, you will enter Withdrawal, which has the same penalties as being
Inebriated, plus extreme irritability and Pain which increases by 2 every hour that you don’t
drink, to a maximum of [20-WIP].\\
\hrule\ \\

{\bfseries Fucking Nazi (-15)\\}
Wait a minute, you guys are Nazis! You’re a dyed in the wool fascist. Or a communist, or
whatever flavor of ultra-idealogue. You might be a member of the Chironite triple-S or a UC
“Cleanup Crew” member, or part of the deepest inner-circle of the Transhuman Death Cult of
the Technocracy. Whatever the case, you’re the sort of person who sets off screaming alarm
bells for anyone who isn’t what you are. You might be a nice guy in theory, but what you believe
is utterly unpalatable to normal people. You suffer -8 to any Charisma test (except Intimidate)
with anyone who isn’t friendly to your group. You may choose to have this quality be a secret
(combines well with the Secret Agent Bane, or even better, the Secret Agent Boon!) but if it gets
out, your party is likely to lynch you. Unless of course they’re also nazis. Wait, is anyone here
NOT a Nazi? Jesus Christ, what the fuck is the Abwehr doing!?\\
\hrule\ \\

{\bfseries Secret Agent (Bad Guy) (-10)\\}
You’re secretly work for a government whose interests do NOT align with the party. Your job is
to make sure that the party doesn’t screw things up for your bosses. You can do it subtly, or you
can do it overtly when the time comes, but if you fail you’re probably going to end up in a gulag
somewhere. Talk to your GM about what exactly you’re trying to prevent the party from doing. It
shouldn’t just be the main goal of the campaign--unless you’re really into being a teamkilling
fucktard--but something secondary that the party could conceivably be persuaded not to do or
prevented from doing without ruining their long-term goals.\\
\hrule\ \\

{\bfseries Prejudiced (-4)\\}
Man, you really hate Zells. Always making up bullshit like “the ocean” and “lobsters,” what a
bunch of lying bastards.\\
Pick a group of any sort, religious, ethnic, racial, political, etc. You just cannot resist spewing
bile at them whenever you get the chance. You must make a Willpower test at RS 3 to avoid
suffering a -8 Charisma penalty when interacting with a member of this group. This does not
apply to Intimidation attempts, or attempts to insult or taunt. If anything, it might help, at the
GM’s discretion.\\
\hrule\ \\

{\bfseries Scrawny (-20)\\}
You’re just not a very substantial person, you’re easy to injure and always seem to come down
with something. Your Toughness is 3 instead of 4, and you suffer a -2 penalty against Infection
and Bloodloss rolls. This is not compatible with Fat or Big Guy.

\chapter{Combat}

{\itshape“Don’t be such a pussy, it’s only death!”}\\
\hspace*{\fill} -Captain Flint Scarsborough DRAC\\

So you’re minding your own business trying to sell some blue jeans to some direly
fashion-deprived Hiver kids, when a bunch of thugs in stupid hats and uniforms show up to ruin
your parade. Turns out they don’t like free enterprise too much around here, something about
disrupting faith in party leadership.\\
You’re not too keen on spending the rest of your life on a truck to the Lubyanka Building, so you
shout something about bread lines, pull out your Liberator 2112, and start shooting.\\

Hey, you gave peace a chance, right?\\

Combat is an important part of Ballad of the Laser Whales. People die in gunfights all the time,
you know. If you don’t want to join that honored register of the dead, read these rules carefully.\\

{\bfseries Rounds\\}
A Round represents about six seconds of fighting. During the Round, characters will act once or
more depending on their Initiative. At the end of the Round, if anyone’s still alive, a new Round
starts and the fighting continues. Combat ends when the shooting stops. Simple enough, right?\\

{\bfseries 3-Phase Combat Round\\}
Each Round has three Phases: 1, 2, and 3. Every character can act in Phase 3, but only
characters with higher ADR (as shown in the table below) can act in Phases 2 and 1. This
represents the speed of the fighters involved. A fast man can get a lot more done in a short time
than a slow one. Characters who can’t act in a Phase simply do not take any Actions.\\

Characters declare Actions in reverse ADR order (from lowest to highest), but the Actions are
resolved in ADR Order (from highest to lowest.) If characters tie in ADR, the characters with the
higher PER is treated as having the higher ADR. If PERs are tied, it goes to WIL. If WILs tie, flip
a coin, or some other random determination.\\

{\centering
\noindent\begin{tabular}{llll}
	Initiative & ADR 1-3 & ADR 4-6 & ADR 7+\\ \hline \\
	Phase 1 & N & N & Y \\
	Phase 2 & N & Y & Y \\
	Phase 3 & Y & Y & Y \\
	\\ \hline
\end{tabular}\\ }\vspace*{\baselineskip} % close centering

{\bfseries Stealing the Initiative\\}
In the general Round, you can temporarily raise your ADR to act in lower Phases and to
improve your place in the Initiative Order, but doing so incurs a penalty to all Actions this Round
equal to the amount by which you raise your ADR. (This penalty also applies to your Combat
Pool if you end up in Melee this Round) You cannot raise your ADR to more than double its
normal amount this way. Opponents and other characters may raise their ADR in response to
any other character doing so, in the same way.\\
Beware overusing this option, as the penalties can quickly become so high as to make you
useless, even if you can act!\\

{\itshape
Krauser and Manaheim face each other in a tense staredown. Manaheim knows Krauser is
quicker than him, and he doesn’t want to die. He incurs a -5 penalty to raise his ADR from 5 to
10 for Initiative, allowing him to act in all 3 Phases, and almost certainly to act before Krauser,
who has ADR 7. Krauser decides he’d rather not suffer a penalty, so he does not raise his in
response. The two draw their pistols.\\
Manaheim now acts before Krauser in Phase 1, as he has higher ADR, but his -5 penalty
severely hinders his accuracy, and he misses both of his shots. Krauser, by contrast, takes aim,
and calmly shoots his opponent in the chest, killing him instantly.\\
Sometimes, it’s better to be cool than quick.\\
}

{\bfseries Actions}\\
You may perform one Action per phase. Sometimes Actions can span multiple phases if they
take longer than a single beat in combat.\\

{\bfseries Base Movement}: You can move a number of meters equal to your Mobility per Phase normally,
or more when Sprinting. If moving into Cover, you do not gain the Cover Bonus to your Defense
until the beginning of the next Phase.\\
Moving normally does not count as an Action, but any Action taken while moving suffers a
{\bfseries -4} Penalty. Conversely, moving raises your Base Defense by {\bfseries +1}.\\
Moving slowly (less than or equal to 1/2 MOB) does not count as an Action, but any Action
taken while moving slowly suffers a {\bfseries -2} penalty. Your Base Defense is not raised when moving
slowly.\\
Sprinting and Active Defense Actions are exempt from these effects, since they are forms of
Movement, and have their own penalties/bonuses.\\

{\bfseries Action Types}:
\underline{Aim}: For Missile Attacks only. You focus in on one target, and increase your odds of hitting it.
Your next Regular Fire or Burst Fire action against that target gains a +2 bonus. You may Aim
for multiple Phases in a row, each time gaining a cumulative +2 bonus to your next attack, to a
maximum of +6 for 3 Phases of aiming. If you successfully Aim three times at a target in
succession, you gain an additional bonus equal to your Perception to the attack that follows. If
you shift targets, lose sight of the target, or are otherwise disrupted, you lose the Aim Bonus.\\

\underline{Shoot}: You may perform a shooting action. There are a few varieties of these. In the event that
the target you were attempting to shoot at is incapacitated or becomes unavailable before you
get to resolve your Shoot, you may pick a new target against which to make the attack, but only
at half your base dice. Other modifiers apply as normal.\\

	\tabto{1}
	{\slshape Regular Fire} lets you fire up to two shots at the same or different targets. If your weapon
	is Pump or Bolt action, you can only fire one shot. Regular Fire has no inherent accuracy
	bonuses, but instead adds BS to damage. For Bonus Successes over 1, you may shift
	the location on the Missile Hit table of your shot by 1. No shot can be shifted by more
	than 1 however.\\
	Instead of firing two shots with Regular Fire, you may Aim as-per the Aim Action and fire
	one shot.\\

	{\slshape Burst Fire} lets you fire 3 rounds at your target. This confers a +4 bonus to shooting, and
	each BS causes an additional round to hit--up to 3. Bonus Success do not raise your
	damage on this maneuver, but for every Bonus Success over 2, you may shift the
	location on the Missile Hit table of one of your shots by 1. No shot can be shifted by
	more than 1 however.\\
	Burst Fire requires that a weapon have a ROF of 3 or better.\\

	{\slshape Full Auto Fire} lets you spray bullets at your enemy en-masse. You fire a number of
	rounds equal to your weapon’s ROF (up to the remaining rounds in your Magazine,
	which must be more than 3 in order for you to use this option) and gain a +8 bonus to
	shooting. Like Burst Fire, each BS causes an additional round to hit--up to the number of
	rounds fired. In the unlikely event that you have more Bonus Successes than bullets
	fired, you may shift the hit location of your shots as with Burst Fire. Full Auto Fire cannot
	benefit from the Aim Action. Additionally Full Auto cannot cause Critical Hits. Any results
	of a Critical Hit instead hit the Chest.\\
	Full Auto Fire requires that a weapon have a ROF of 4 or better.\\

	{\slshape Suppressive Fire} lets you fire a sustained blast of fire over a large area, potentially hitting
	everyone in that area, and denying them access to it without risk of being shot. You fire a
	number of rounds equal to your weapon’s ROF (up to the remaining rounds in your
	Magazine, which must be more than 3 in order for you to use this option.) Choose an
	area up to 2m wide per ROF of the weapon in a line perpendicular to your character, any
	distance away within range. Imagine a cone between your character and that line, and
	extending on behind. That cone is the Area of Suppression. Characters who move
	through this area suffer attacks as listed below. {\bfseries Treat these attacks as Full Auto Fire
		attacks for the purposes of Bonus Successes.}\\
	Suppressive Fire requires that a weapon have a ROF of 4 or better.\\
	-Any figure who moves into our out of the Area of Suppression laterally (moving parallel
	to the line) suffers an attack by you at half your normal Dice, +2 dice per 3m they move
	through the Area of Suppression.\\
	-Any figure who moves towards you through the Area of Suppression (moving
	perpendicular to the line) suffers an attack at full Dice +2 dice per 3m they move through
	the Area of Suppression.\\
	-Any figure who performs an Action, while in the Area of Suppression, suffers an attack
	by you at full Dice before they resolve their Action. Regular Fire only counts as one
	Action for these purposes, and only incurs one attack.\\
	-Any figure who ends the Phase in your Area of Suppression suffers an attack from you
	at +4 dice, or +6 if they made an attack during this Phase.\\
	-Making an attack while in an Area of Suppression is difficult--Characters suffer a -4 to
	any attacks made while under Suppressive Fire.\\
	Weapons with ROF 1 or Single Shot cannot Suppress. Single-Actions can be used to
	Suppress at ROF 3, but suffer the Fan-Fire penalty (-4) as normal.\\
	{\bfseries If/When the total number of HITS equals the firing weapon’s RoF, the Suppression
	ends immediately.}\\

	{\slshape Covering Fire} lets you wait for an enemy to show himself, and then shoot him if he does.
	Covering Fire suffers an additional -4 penalty if you’re Moving. If a character steps into
	your field of view or otherwise exposes himself (you can choose another trigger, like
	waiting for the guy to draw his gun before shooting him, etc) you may immediately take a
	Regular Fire, Burst Fire or Full Auto Fire Action. You cannot Suppress with Covering
	Fire. Covering Fire lasts until the next Action you take, but you can choose to simply
	sustain Covering Fire every Action until something happens. This can allow a character
	who cannot normally act in Phase 1 to act in Phase 1 if he Covered in Phase 3.\\

	{\slshape Blind Fire} lets you make any of the above attacks except for Covering Fire at extreme
	penalties, while also making yourself much harder to hit. When Blind Firing you must be
	in at least $1/2$ Cover. ALL Hit Locations immediately become “covered” except for your
	Weapon Arm, and you gain +2 Defense. You suffer a -8 Penalty to Hit with your attacks,
	cannot benefit from the Aim Action, but may otherwise make attacks normally.\\
	Suppressive Fire is made as normal, but at an additional -2 Penalty on your part. You
	cannot aim for a specific Hit Location while Blind Firing.\\

	{\slshape Multi-Barrel Shot} fires all (loaded) barrels of the weapon simultaneously. This confers no
	bonus to shooting, and each BS causes an additional round to hit--up to number of
	barrels fired. Bonus Successes do not raise your damage on this maneuver, but for
	every Bonus Success over 1, you may shift the location on the Missile Hit table of one of
	your shots by 1. No shot can be shifted by more than 1 however.\\

\tabto{0}
\underline{Throw}: You spend the Phase throwing a grenade or other object. You may throw the object up
to its Throwing Distance. Roll ADR at TN 7. The RS for landing the grenade within a meter of
the target-area is 4. If you fail, the throw will deviate in a random direction (1-3 too close, 4-6 too
far, 7-8 to the left, 9-10 to the right) by 2 meters for every success you fail by. If you critically fail,
roll deviation. The object will land 10 meters closer to you in a straight line from its deviation
point.\\
Throwing a weapon with the intent of actually hitting an enemy directly to injure them is resolved
differently, under Fling.\\

\underline{Active Defense}: You spend the Phase ducking and dodging enemy attacks. Roll an ADR Test
and add BS to your Defense RS. If you declare Active Defense, roll it before resolving any other
Actions--it applies against attacks that occur before your Initiative in the Phase, as well as those
after. This can be performed either standing still behind one piece of cover, or while running
along or between close (2-3 meters apart) pieces of cover.\\

\underline{Take Cover}: Similar to Active Defense, but only usable behind at least Half Cover, you simply
duck down and hide. You now have Full Cover, and cannot be hit by any attacks that cannot
penetrate that Cover or otherwise circumvent it (you are still vulnerable to flanking attacks,
though you count as Prone against them.) including Suppression. You cannot move faster than
2 meters per Phase while Taking Cover. If the cover can be penetrated by an attack, treat Full
Cover as 8/10ths cover, but with all locations except Critical Hit covered.\\

\underline{Reload}: You spend the Action reloading your weapon. For weapons with Box magazines (or
Cylinder/Internal Magazine weapons which you have a speedloaders or stripper clips for), this
means you instantly reload the weapon. For weapons with Cylinders or Internal Magazines, this
means you reload an number of rounds equal to your ADR. This also applies if you don’t have
an additional magazine for your Box-Fed weapon, and have to input rounds manually into the
magazine of the weapon either separate from the gun or through the slide-breech.\\

\underline{Sprint}: You move up to double your Mobility score in meters this Action in lieu of taking any
non-Mobility Actions. While sprinting, you gain a +2 bonus to Defense.\\

\underline{Melee Attack}: Move up to 10m into contact (2m away from) with a target, and Enter Melee
Combat with a target for 1 Round. If you attack from a flank or behind, this counts as a Surprise
Attack. If you cannot come within 2m of the target, you cannot initiate Melee Combat. You may
choose to Charge into combat, gaining +4 CP for the first round of combat, but you may only
declare Aggressive/Red as your Orientation. Reach is chosen by the character who holds
initiative at the end of Orientation, with the maximum being the Reach of the longer weapon
between the two (LL if there a missile weapon in play), and the minimum being the reach of the
shortest weapon (S if unarmed or armed with a firearm). If neither character has initiative or the
reach is somehow in question, default to the maximum reach.\\

\underline{Magic}: Spend the Phase using magic of some variety, such as a Void-Push or Shift.\\

{\bfseries Defense RS\\}
The RS to hit a stationary man-sized target not in cover is 4. Various forms of cover increase
this RS depending on their amount. The Action Active Defense increases this number further
through an ADR roll on the defender’s part. Additionally, environmental effects like lighting and
visibility can increase Defense, as detailed on the table below.\\
A character with a Camouflage Bonus adds it to RS against any attacks made from more than
50 meters distance away.\\
Offensive Void powers halve the Defense RS to hit their intended target.\\

Environmental effects that deal in visibility (lighting, smoke, etc.) do NOT raise Defense against
Suppressive Fire.\\

{\centering
\begin{tabular}{p{2.3cm}p{1.5cm}p{2cm}}
Environmental Effect 					& Defense Increase & Counter\\ \hline \\
Dimly Lit 								& +1 				& Flashlight, Night Vision Goggles\\ \\
Darkness 								& +2 				& Flashlight, Night Vision Goggles\\ \\
Smoke, Mist, Fog etc. 					& +2 				& None\\ \\
On a ship that is moving erratically 	& +1 				& Be a Zell\\ \\ \hline
\end{tabular}\\[\baselineskip] }\vspace*{\fill}

{\bfseries Camouflage \& Concealment\\}
Camouflage and Concealment play dual roles. Firstly, they can affect a character’s Defense RS,
as detailed above. Secondly, they can make it difficult for an attacker to actually target a
character in the first place.\\

When attempting to make an attack against enemies further than 50m away, a character must
make an Awareness roll at RS equal to 0+Camouflage+Concealment, rolled against all enemies
who currently benefit from Camouflage/Concealment. Roll only once and compare the roll to all
enemies’ RS. If the roll equals or exceeds the RS of a character, they can be targeted as
normal. If not, they cannot be targeted at all.\\
Once within 50m, only characters who succeed at their Stealth checks remain
undetected--otherwise they can be targeted normally.\\

{\centering
\begin{tabular}{p{6cm}p{1.2cm}}
	Concealment 				& Camo Bonus \\ \hline \\
	Poor Lighting 				& +1 \\
	Darkness 					& +2 \\
	Smoke, Mist, Fog, etc 		& +2 \\
	Thin Brush, Rocky Terrain 	& +1 \\ 
	Heavy Brush, Jungle, Cliffside & +2 \\
	Dense Urban/Interior Environment (Ruins, Street Battlefield, Cargo Bay) & +2 \\ \\ 
	Character is Crouching & +1 \\
	Character is Prone & +2 \\
	Half Cover* & +2 \\
	3/4ths Cover* & +3 \\
	9/10ths Cover* & +4 \\
	Every 100m distance after the first & +1 \\
	Character has fired a weapon without a Flash Suppressor in this or the previous Phase & -4 \\ \\
	Character is in extremely visible attire & -4 \\
	Character is moving very slowly (Half Mob) & -1 \\ \\
	Character is moving regular speed & -2 \\
	Character is running/sprinting & -4 \\

	\\ \hline
\end{tabular}\\ }
*Cover gained from shields don’t count for the purposes of camouflage\\

{\bfseries Making Attacks\\}
A character makes an attack with his Pool (Missile Pool for missile weapons) plus any other
modifiers he might have.\\
Each multiple of the weapon’s Range that the Target is away after the first increases the
Defense RS by 1.\\
Aiming for a specific Target Location costs 4 dice, but the attack automatically hits the chosen
Target Location. You cannot “aim” for Critical Hits, however, and this option does remove the
chance that those will occur.\\
A successful hit results if the attacker scores {\bfseries equal to or more} successes than the target’s
Defense RS. Hit Location is rolled on the Random Hit Table unless the attack targeted a specific
location as detailed above.\\
An attack using Void Magic halves the target’s Defense RS.\\

{\bfseries Critical Failure\\}
A Critical Failure is any Check or Opposed Roll which is failed or lost, and in which more 1s
were rolled than successes.\\

{\bfseries Dual-Wielding\\}
A character may Dual Wield pistols (using two at once) with the following Benefits:\\

	\tabto{1}\noindent
	{\bfseries +}Can perform a Shoot Action of any variety available to the weapon (Regular, Burst, etc) with
	each weapon\\
	{\bfseries +}Can reload both weapons simultaneously if stationary (no Move Action that phase)\\

	And the following Penalties:\\
	{\bfseries -4 to MP} with all attacks (-2 MP if the Character has the Ambidextrous Boon)
	Cannot use the Aim Action normally with either weapon.\\

\tabto{0}
{\bfseries Outflanking\\}
A character’s Defense is lowered by 2 against attacks that come from the 90 degrees to their
left and right. Cover that does not extend all the way around them is also ignored by such an
attack.
Attacks from the rear 90 degrees lower that Defense by 4, and similarly ignore cover that does
not extend behind the character.

{\bfseries One-Handed Shooting\\}
All weapons, barring Pistols, are assumed to be fired using both hands. If forced to fire a
non-pistol in one hand, you suffer a penalty to MP equal to the weapon’s WT multiplied by 2.
Additionally, you cannot benefit from Aim actions, and MP bonuses (such as from Burst or Auto
fire) are halved. Firing weapons from your off-hand suffers a -4 MP penalty, this is negated by
Ambidextrous, and lessened to -2 MP by Off-Hand Training.\\

{\bfseries Shooting Point Blank\\}
If adjacent to a target (2m or closer) and the target is not helpless or unaware, some weapons
take a severe penalty to shooting. Aiming is also impossible at this range.\\
These penalties apply to Initiative Contests made with the Melee Shoot maneuver in melee
combat. {Your chance to shoot them easily already passed!} {Don’t use a gun in a knife
fight!}\\

	\tabto{1}\noindent
	Pistols suffer -2 to attacks.\\
	Shotguns and SMGs suffer -3 to attacks.\\
	Rifles suffer -6 to attacks.\\
	Machineguns suffer -10 to attacks.\\

\tabto{0}
However, if an enemy is this close and helpless/unaware of the attacker, you suffer none of the
above penalties, their Defensive RS is 0, and all dice of the attack automatically succeed.\\

{
\centering
\begin{tabular}{ll}
Roll 	& Location \\ \hline \\
10 		& Critical Hit \\
9 		& Head \\
8 		& Chest (Vitals) \\
7 		& Chest (non-Vitals) \\
6		& Abdomen \\
5		& Abdomen \\
4		& R Arm \\
3		& L Arm \\
2		& R Leg \\
1		& L Leg \\
\\ \hline
\end{tabular}\\ } % close centering

\newpage % manual pagebreak
{\bfseries Critical Hits\\}
A Critical Hit result means something special happens. Here are some sample Critical Hits--if
you want something different and narratively appropriate to happen (say, disarming an
opponent of his sword, or shooting the feather off of his cap, etc) consult with your GM when it
comes up. For those of you who hate narrativist systems, Critical Hits allow you to choose the
location which you hit, and ignore 50\% of the AV on that location.\\
Full Auto Fire attacks cannot score Critical Hits. A Critical Hit result is instead a Chest
(Non-Vitals) result. Shotguns’ primary attacks can score Critical Hits, but the secondary “scatter”
attacks cannot, and count as hitting the Chest instead.\\

{\bfseries Weapon Jams and Malfunctions\\}
Guns jam on critical failure, or as a result of the Jams quality. When that happens, a roll on the
jam table is required. Check and resolve Jams before resolving the shot that caused it!\\

{\centering
\begin{tabular}{lp{7cm}}
Roll 			&	Result \\ \hline \\
1-3				&	Stoppage! Failure to Eject, Failure to Extract, or something
					similar. Requires a Reload action to clear, shot is resolved as
					normal. Ignore this result if the weapon uses a cylinder
					mechanism.\\ \\
4-8				&	Failure to Fire! Light primer strike, Failure to Feed, or something
					similar. Shot doesn’t resolve at all. Requires a Reload action to
					remediate, though not if the weapon uses a cylinder mechanism,
					in which case the next shot fires normally.\\ \\
9				&	Major Stoppage! Possible double-feed, case head separation, or
					possibly something more esoteric, like chambering a wasp.
					Requires 3 Reload Actions to clear, 2 if you possess the Rapid
					Unjam Talent. Shot is resolved as normal.\\ \\
10				&	Uh-oh. The first time you get this result while firing a specific
	weapon, reroll on the table. The second time you receive this
	result, something in the firearm is damaged/set out of time during
	operation. This affects the weapon’s accuracy, raising the MTN
	by 1. If this result is rolled a third time while firing the same
	weapon, the weapon suffers a catastrophic failure. The weapon
	is functionally destroyed, and the wielder suffers half the
	weapon’s normal damage to a random location, rolled on the
	Missile Hit Table (Critical Hits count as Head). This counter is
	reset when the weapon is cleaned and/or repaired.\\ \\\hline
\end{tabular}\\[\baselineskip] }

Various factors can add (or add to!) various levels of the Jams quality to the weapon, as
described in the following table. If a Weapon already has Jams, add +1 to it, or if the Level
suggested exceeds the current level +1, change it to the suggestion, and add +1 to that.\\

{\centering
\begin{tabular}{p{3cm}p{3cm}}
Level of Jams														& Appropriate Conditions \\ \hline \\
Lower Jams by 1, to a minimum of Jams [0] (Weapon already has Jams)	& Good Ammo (125\% Price), Just Cleaned, Sterile Range \\ \\
Jams [0]															& Cheap Ammo (80\% Price), Infrequent Cleaning, Dropped in the Sand \\ \\
Jams [2]															& Really Cheap Ammo (60\% Price), Can’t Remember the Last Cleaning, Left in the Sand \\ \\
Jams [4]															& Dangerously Cheap Ammo (40\% Price), What’s a Cleaning Rod?, Arizona Mud and a Wheelbarrow \\
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Cover\\}
When in Cover, you gain a Defense Bonus against enemy attacks that the Cover protects
against (if the cover isn’t actually between you and the shooter, it’s obviously useless) and some
of your Random Hit Locations are “covered,” meaning that attacks that land on those Zones are
rerolled on the hit location table, until they land on an uncovered location. These are numbers
on the Random Hit Table, numbers 1-9 (10 can never be obscured) and are chosen either by
the GM, or by you at the GM’s discretion. Usually how you take cover behind the object dictates
what parts are covered (shooting over a box that grants Half Cover would probably cover 1-5,
for example) but in some cases this might change, like crouching around the side, etc. At GM’s
discretion, once you have been in cover for a full Phase, you may choose which Parts are
Covered at the start of each Phase.\\
If your Head is ever a Part Covered, you cannot make a normal Shoot action, and can only
Blindfire.\\

{\centering
\begin{tabular}{lll}
Cover 		& Defense Bonus & Parts Covered \\ \hline \\
Half 		& +2 			& 5 \\
$3/4$ 		& +4 			& 7 \\
$8/10$ths 	& +6 			& 8 \\
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Piercing Cover\\}
Some Cover can be penetrated, reducing its effectiveness and making it less useful against
some weapons. Cover has a Durability, and if a weapon’s Base Damage+Armor Penetration is
higher than that Durability, then the weapon can penetrate the cover, and potentially hit a target
hiding behind it.\\
An attack with higher Base Damage+Armor Penetration than the Cover’s Durability halves the
Cover’s Defense Bonus. Additionally, when rolling for Hit Location, a result that lands on a Part
Covered is not re-rolled, but instead hits that part, but reduces its AP by the Cover’s Durability.\\

{\bfseries\slshape Cover Durability\\}

{\center
\begin{tabular}{p{5cm}l}
Cover Type & Durabulity \\ \hline \\
Sheet Rock/Plywood & 4 \\
2cm thick Wooden Wall, Cardoor & 7 \\ \\
Cinder Blocks, Store Shelves & 9 \\
Sheet Metal, Small Tree & 11 \\
Brick Wall, Shipping Crates & 15 \\
Ship Hull, Large Tree, 1 Meter of Wood & 20 \\ \\
Sandbags/1’ of packed earth, Concrete Bunker Wall & 40 \\
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Dealing Damage\\}
The damage done by an attack is compared to the Damage Threshold of the location hit. For
each point by which the damage done exceeds the Target’s DT, a level of Wound is inflicted.
{\bfseries An attack that hit with 0 BS can never inflict higher than a Level 1 Wound, regardless of
 damage done.} This applies only to attacks that were rolled--the additional hits for Burst or Full
Auto Fire count as having BS for the purposes of damage, even though they do not add those
BS to damage. Additionally, attacks that combine damage may inflict higher than a level 1
Wound so long as at least one of the attacks scored more than 0 BS.\\
If multiple attacks from the same source (both shots from a Regular Fire action, or rounds from
a Burst Fire action for example) hit the same Hit Location simultaneously, the damage from
these attacks partially combine, the first Attack gaining half of the damage of subsequent
attacks.\\

{\itshape Krauser shoots Danger Redcliff twice with his Ritterschwert. Both shots land on Redcliff’s Chest.
Rather than inflicting two Strength 8 attacks, these combine into one Strength 12 attack (8+4)
which is enough to penetrate Redcliff’s armor and injure him.\\}

\twocolumn[{
{\bfseries Bloodloss\\}
Bloodloss is tested at the end of every Round, after all 3 Phases have concluded.\\

{\bfseries Head\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		& 													& Close Call, clothing, helmet or hair clipped. \\ \\
Level 1 		& 3 	& 1 	& 1 		& 													& Graze to temple or cheek. \\ \\
Level 2 		& 5 	& 3 	& 2 		& KO RS 2 											& Glance to skull, possible unconsciousness. \\ \\
Level 3 		& 10 	& 10 	& 8 		& KO RS 4 											& Shot to mouth, teeth lost, jaw fractured, unconsciousness likely. \\ \\
Level 4 		& All 	& All 	& 15 		& Automatic KO, Difficult Surgery vs Brain Damage 	& Shot in the head, significant damage to brain. This is actually very survivable in some situations--sometimes with no major side-effects, even.\\ \\
Level 5 					& Dead 			& Dead & Dead & Dead & Brains splattered over the opposite wall. Dead.\\
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Chest (Vitals)\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		&													& Close Call. \\ \\
Level 1 		& 2 	& 0 	& 0 		&													& Flesh wound, nothing serious. \\ \\
Level 2 		& 3 	& 1 	& 5 		& 													& Penetration into the body, no organs hit but dangerous bleeding. \\ \\
Level 3 		& 5 	& 3 	& 10 		& 													& Artery Clipped! Severe bleeding, call a medic! \\ \\
Level 4 		& 10 	& 6 	& 20 		& KO RS 4											& Punctured Lung, blood filling chest cavity. Death likely. \\ \\
Level 5 		& Dead 	& Dead 	& Dead 		& Dead 												& Shot in the Heart. Dead. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Chest (non-Vitals)\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		&													& Close Call. \\ \\
Level 1 		& 2 	& 0 	& 0 		&													& Flesh wound, nothing serious. \\ \\
Level 2 		& 3 	& 4 	& 2 		& 													& Clips a rib, painful! \\ \\
Level 3 		& 5 	& 8 	& 4 		& 													& Cracked ribs, exceptionally painful! \\ \\
Level 4 		& 10 	& 12 	& 8 		& Internal Bleeding									& Breastbone hit, fragments lacerate internals, severe bleeding. \\ \\
Level 5 		& 20 	& 20 	& 15 		& Internal Bleeding									& Ribcage shattered, possible organ damage, have fun picking out bone shards. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

}]
\twocolumn[{

{\bfseries Abdomen\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		&													& Close Call. \\ \\
Level 1 		& 2 	& 0 	& 0 		&													& Flesh wound, nothing serious. \\ \\
Level 2 		& 3 	& 2 	& 2 		& 													& Cuts through some fat, but otherwise nothing serious. \\ \\
Level 3 		& 5 	& 4 	& 4 		& 													& Penetration through oblique muscles, painful, but not lethal. \\ \\
Level 4 		& 10 	& 6 	& 6 		& Infection Chance [+3]								& Gut shot. The pain is unbearable, but bleed-out is unlikely with medical attention. Infection, though... \\ \\
Level 5 		& 20 	& 10 	& 10 		& Infection Chance [+4]								& Severe organ damage, incomparable suffering. Death will come, but probably not quickly. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Arm\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		&													& Close Call. \\ \\
Level 1 		& 1 	& 0 	& 0 		&													& Graze of the upper arm. \\ \\
Level 2 		& 2 	& 0 	& 0 		& Disarm vs [1] RS									& Flesh wound, surprising, might drop weapon. \\ \\
Level 3 		& 4 	& 2 	& 3 		& Disarm vs [4] RS									& Shot in the upper arm. Muscle damage, but little bleeding. Will probably drop weapon. \\ \\
Level 4 		& 6 	& 4 	& 8 		& Disarm vs [6] RS									& Shot in the firearm, tendons, weapon dropping is very likely. \\ \\
Level 5 		& 10 	& 6 	& 10 		& Autodrop, [RS 2] Surgery vs [Crippled Limb (arm)]	& Artery clipped, arm broken, weapon dropped. Bloodloss is a serious threat. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Leg\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		& 													& Close Call. \\ \\
Level 1 		& 3 	& 0 	& 1 		& Stability vs [1] RS								& Grazed thigh. \\ \\
Level 2 		& 4 	& 1 	& 3 		& Stability vs [3] RS								& Flesh wound to the thigh or shin. \\ \\
Level 3 		& 6 	& 2 	& 6 		& Auto Prone										& Shot in the shin or knee, muscle spasms cause leg to fail. \\ \\
Level 4 		& 8 	& 4 	& 10 		& AutoProne, [RS 2] Surgery vs [Crippled Limb (leg)]& Shot in the shin or knee, bone broken, may not recover. \\ \\
Level 5 		& 12 	& 6 	& 15 		& AutoProne, [RS 4] Surgery vs [Crippled Limb (leg)]& Shot in the thigh, artery clipped, may never recover--bloodloss may kill. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

{\bfseries Leg\\}

{\centering
\begin{tabular}{llllp{3cm}p{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 0 		& 1 	& 0 	& 0 		& 													& Close Call. \\ \\
Level 1 		& 3 	& 0 	& 1 		& Stability vs [1] RS								& Grazed thigh. \\ \\
Level 2 		& 4 	& 1 	& 3 		& Stability vs [3] RS								& Flesh wound to the thigh or shin. \\ \\
Level 3 		& 6 	& 2 	& 6 		& Auto Prone										& Shot in the shin or knee, muscle spasms cause leg to fail. \\ \\
Level 4 		& 8 	& 4 	& 10 		& AutoProne, [RS 2] Surgery vs [Crippled Limb (leg)]& Shot in the shin or knee, bone broken, may not recover. \\ \\
Level 5 		& 12 	& 6 	& 15 		& AutoProne, [RS 4] Surgery vs [Crippled Limb (leg)]& Shot in the thigh, artery clipped, may never recover--bloodloss may kill. \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

}]
\twocolumn[{
{\bfseries Burn Damage\\}
Burn Damage is not reduced by TOU. Burn Damage is Cumulative If a character suffers a Level
1 Burn, and then next Round suffers two additional Burn Damage, this does not create a Level 2
Burn, but instead turns the Level 1 into a Level 3.\\

{\centering
\begin{tabular}{lllllp{7cm}}
Wound Level 	& Stun 	& Pain 	& Bleed 	& Special 											& Description \\ \hline \\
Level 1 		& 4 	& 4 	& - 		& [-3 Infection Chance]								& 1st degree Burn, painful but not serious \\ \\
Level 2 		& 8 	& 6 	& - 		& [-1 Infection Chance]								& Minor 2nd degree burn with swelling, but no probable scarring. \\ \\
Level 3 		& 12 	& 10 	& - 		& [+1 Infection Chance]								& Bad 2nd degree burn, skin lost, scarring probable. \\ \\
Level 4 		& 20 	& 20 	& - 		& [+3 Infection Chance]								& 3rd degree burn, serious scarring and injury. Consider disfigurement Banes. \\ \\
Level 5 		& Total & 30 	& - 		& [+5 Infection Chance]								& 4th degree burn or worse, flesh destroyed, bone exposed and charred, almost certainly lethal if located on vitals (HLT at RS 5) \\ 
\\ \hline
\end{tabular}\\[\baselineskip] }

}] % last damage table done

\chapter{Equipment}

{\bfseries Firearm Stats\\}

\tabto{1}
Weapon\\
The name of the Weapon described.\\

Type\\
What Proficiency is used while firing the weapon, weapons with two can be used with either
listed.\\

Range Increment\\
The weapon’s range, firing past a weapon’s range increment grants the target +1 Defense per
multiple of the range increment.\\

MTN\\
The TN used when firing the weapon.\\

Round\\
The cartridge the weapon fires, weapons with multiple can fire either.\\

Mag and Capacity\\
The amount of rounds the weapon holds. The first number is the capacity of weapon’s normal
magazine, second number indicates the capacity of an extended magazine if one exists for the
weapon in question. In the parentheses, the type of feeding mechanism is noted. (Box) denotes
that the weapon feeds from a detachable box or drum magazine. (Int) denotes internal
mechanisms, whether that be a tubular magazine, internal magazine, enbloc clip, or a simple
break-action. (Cyl) denotes a revolving cylinder, for revolvers. (Belt) fed weapons have no
specific capacity, it being tied to the belt being used at the moment. Internal Magazines and
most Cylinders cannot be changed on the fly, requiring the use of stripper clips, speedloaders,
or simply loose rounds. Box fed firearms can have an extra round in the chamber.\\

RoF/Action\\
The weapon’s rate of fire and Action, used for determining if the weapon is capable of Burst,
Full-Auto or Suppressive fire.\\

Special\\
Special Rules applied to the weapon.\\

WT (Weight)\\
The weight of the weapon added to Encumbrance. Firearms apply weight regardless of whether
they are slung or held in hand.\\

Cost\\
The weapon’s base cost. Almost certainly varies regionally.\\

\tabto{0}
% pdftex specific commands to change the document dims mid document
%\eject \pdfpagewidth=42cm \pdfpageheight=29.7cm
\twocolumn[{ 
	\section{Weapons}
	% put the equipment in a dedicated pdf in landscape mode; alternatively stretch the table over two adjacent pages, so the table fills an a3 landscape page
	%{\bfseries Pistols\\}
	\subsection{Pistols}

	\noindent{\centering
	\begin{tabular}{p{3cm}lllp{2cm}llp{2.7cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	Abitante “Verro” MR/34					& Pistol & 10m 	& 6 	& 10mm 								& 6 (Cyl) 		& 1		& Conceal [2], \newline Reliable 				& 0.5 	& 6 gp 	\\
	Canton PM210							& Pistol & 10m 	& 6 	& 9mm 								& 8 (Box) 		& 1		& Conceal [2], \newline Reliable 				& 0.5 	& 10 gp \\
	C08 Eroberer							& Pistol & 10m 	& 6 	& 7.5mm 							& 8 (Int) 		& 1		& Conceal [2], \newline Reliable 				& 0.5 	& 15 sp \\
	C18 Streithammer						& Pistol & 10m 	& 7 	& 9x19mm 							& 10/40 (Box) 	& 1		& Conceal [2] 							& 0.5 	& 2 gp 	\\
	C19 Ritterschwert						& Pistol & 10m 	& 6 	& 9x19mm 							& 8/32 (Box) 	& 1		& Conceal [2] 							& 0 	& 5 gp 	\\
	C19S Kriegsmesser						& Pistol & 10m 	& 6 	& 9x19mm 							& 8/32 (Box) 	& 10	& Conceal [2] 							& 0 	& 10 gp \\
	C20 Sturmpistole						& Pistol & 10m 	& 7 	& 9x19mm 							& 10/40 (Box) 	& 30	& Conceal [2] 							& 0.5 	& 5 gp 	\\
	Republican Arms Conquest SAA			& Pistol & 10m 	& 6 	& 11.5mm \newline Rahoo 			& 6 (Cyl) 		& 1		& Conceal [2], \newline Reliable, \newline Single-Action 	& 0.5 	& 5 gp 	\\
	Republican Arms Liberator 1991			& Pistol & 10m 	& 7 	& 11.5mm \newline Rahoo 			& 7 (Box) 		& 1		& Conceal [2], \newline Reliable 				& 0.5 	& 1 gp 	\\
	Republican Arms Defender 1991			& Pistol & 10m 	& 7 	& 11.5mm \newline Rahoo 			& 6 (Box) 		& 1		& Conceal [3], \newline Reliable 				& 0 	& 3 gp 	\\
	Republican Arms Emancipator 1991		& Pistol & 10m 	& 7 	& 11.5mm \newline Rahoo 			& 7/25 (Box) 	& 20	& Conceal [2] 							& 0.5 	& 6 gp 	\\
	Model 9 “Tianleng”						& Pistol & 10m 	& 7 	& 7.5mm 							& 8 (Box) 		& 1		& Conceal [2], Jams [2] 				& 0 	& 5 sp 	\\
	Type 17 “Yenhsi”						& Pistol & 10m 	& 8 	& 11.5mm \newline Rahoo 			& 10 (Int) 		& 3		& Conceal [2], Jams [2] 				& 1 	& 1 gp 	\\
	Suncrusher .357							& Pistol & 10m 	& 8 	& 9x33mm 							& 9 (Box) 		& 1		& Conceal [2] 							& 1 	& 7 gp 	\\
	Buccaneer .69							& Pistol & 10m 	& 7 	& 17.56mm 							& 1 (Int) 		& 1		& Conceal [2] 							& 1 	& 5 gp 	\\
	Blackbeard .69							& Pistol & 10m 	& 8 	& 17.56mm 							& 4 (Cyl) 		& 1		& Conceal [1], \newline Reliable, \newline Single-Action 	& 1.5 	& 10 gp \\
	Eugene-Longhill “Stonebreaker” ‘36 		& Pistol & 10m	& 7		& 9x33mm							& 6 (Cyl) 		& 1 	& Single-Action, Conceal [2] 			& 0.5	& 3 gp \\
	Norminster Pistol 						& Pistol & 10m	& 7		& 11.5mm \newline Rahoo				& 6 (Cyl) 		& 1 	& Conceal [2], \newline Reliable 				& 0.5	& 1 gp \\
	Hector Arms Mustang .357 				& Pistol & 10m	& 7/6	& 9x33mm/\linebreak[0]9x29 (-1 TN)	& 6 (Cyl) 		& 1 	& Conceal [2], \newline Reliable 				& 1		& 5 gp \\
	Hector Arms Police Snub .357 			& Pistol & 10m	& 8/7	& 9x33mm/\linebreak[0]9x29 (-1 TN)	& 5 (Cyl) 		& 1 	& Conceal [3], \newline Reliable 				& 0.5	& 4 gp \\
	Hector Arms Mule .38 					& Pistol & 10m	& 6		& 9x29mm				& 6 (Cyl) 	& 1 			& Conceal [2], \newline Reliable 				& 0.5	& 1 gp \\
	Kaufman-Isaev King Eagle 				& Pistol & 10m	& 8		& 7.62x33mm				& 8 (Box) 	& 1 			& Conceal [1] 							& 1		& 23 gp \\
	UC Naval Pistol 						& Pistol & 10m	& 6		& 7.5mm					& 12 (Box) 	& 1 			& Conceal [2], \newline Reliable 				& 0.5	& 2 gp \\
	UC Agent Pistol 						& Pistol & 10m	& 6		& 7.5mm					& 5 (Box) 	& 1 			& Conceal [3], \newline Silenced 				& 0		& 10 gp \\
	\end{tabular}\\[\baselineskip] } 
	\large\flushright \underline{Continued on next page$\rightarrow$\textgreater\textgreater}
}]
\twocolumn[{
	\noindent{\centering
	\begin{tabular}{p{3cm}lllp{2cm}llp{2cm}ll}
	UC “Insurgent” Palm Pistol 				& Pistol & 2m	& 7		& 11.5mm \newline Rahoo			& 1 (Int) 		& 1 	& Conceal [4], Jams[1], Single Shot 	& 0		& 1 sp \\
	Kuropatkin Silenced Revolver 			& Pistol & 10m	& 7		& 7.5mm					& 6 (Cyl) 		& 1 	& Conceal [3], \newline Single Action, Silenced 	& 0		& 10 gp \\
	Republican Arms “Wrath of God” 1776 	& Pistol & 10m	& 8		& 17x76mm				& 4 (Cyl) 		& 1 	& Conceal [1], \newline Single Action 			& 1.5	& 45 gp \\
	Jenner Micropistol 						& Pistol & 10m	& 6		& 6x16mm				& 8 (Box) 		& 1 	& Conceal [4] 							& 0		& 3 sp \\
	Herstal “HAP” Automatic Pistol 			& Pistol & 10m	& 7		& 9x19mm				& 13 (Box) 		& 1 	& Conceal [2], Reliable 				& 0.5	& 6 gp \\
	Dixon Phoenix 							& Pistol & 10m	& 7		& 10mm					& 10 (Box) 		& 1 	& Conceal [2] 							& 0.5	& 5 gp \\
	Bittenfield “Sonderbarkeit” Großmacht 	& Pistol & 10m	& 6		& 9mm					& 12 (Box) 		& 6 	& Conceal [3], Reliable 				& 0.5	& 75 gp \\
	Herstal Model 28 Pocket Hammerless 		& Pistol & 10m	& 6		& 9x17mm				& 7 (Box) 		& 1 	& Conceal [3], Reliable 				& 0		& 15 sp \\
	Rahoo Headhunter 						& Pistol & 20m	& 8		& 7.62x63mm				& 1 (Int) 		& 1 	& Conceal [2], Singleshot 				& 0.5	& 8 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip]

	\begin{tabular}{p{2.5cm}lllp{2cm}llp{2cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	Caster & Pistol & 10m & N/A & Caster Shells & 1(Int) & 1 & Single Shot & 0.5 & 2 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	%{\bfseries SMGs\\}
	\subsection{SMGs}

	{\centering
	\begin{tabular}{p{2.5cm}lllp{2cm}llp{2cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	C19S Kriegsmesser w/Stock 	 & SMG & 50m & 6 & 9x19mm 			& 8/32 (Box) 	& 10 & 							& 1 	& 10 gp \\
	C20 Sturmpistole w/Stock 	 & SMG & 50m & 7 & 9x19mm 			& 10/40 (Box) 	& 30 & 							& 1.5 	& 5 gp 	\\
	Nachtmacher C45 			 & SMG & 50m & 7 & 9x19mm 			& 32/50 (Box) 	& 15 & Reliable 				& 2 	& 10 gp \\
	Weißenschwert C41 			 & SMG & 50m & 7 & 9x19mm 			& 35/75 (Box) 	& 35 & 							& 3 	& 8 gp 	\\
	Soltar Carbine .30 			 & SMG & 50m & 7 & 7.62x33mm 		& 30 (Box) 		& 20 & Jams [2] 				& 3 	& 8 gp 	\\
	Type 71 					 & SMG & 50m & 7 & 8x22mm 			& 30/60 (Box) 	& 25 & Jams [2] 				& 2 	& 3 gp 	\\
	Rahoo M28 “Viper” 			 & SMG & 50m & 7 & 11.5mm \newline Rahoo 	& 20 (Box) 		& 20 & Jams [1], \newline Silenced 		& 2 	& 8 gp 	\\
	Rahoo “Glue Gun” M2 		 & SMG & 50m & 7 & 11.5mm \newline Rahoo 	& 30 (Box) 		& 15 & Conceal [1], Jams [4] 	& 2 	& 3 gp 	\\
	Rawn Gun 					 & SMG & 50m & 7 & 9x19mm 			& 32 (Box) 		& 20 & Jams [1] 				& 2.5 	& 2 gp 	\\
	UC Naval SMG 				 & SMG & 50m & 6 & 7.5mm 			& 20 (Box) 		& 30 & Conceal [2] 				& 1 	& 5 gp 	\\
	Sark SMG 					 & SMG & 50m & 8 & 9x33mm 			& 20 (Box) 		& 10 & Jams [2] 				& 2 	& 6 gp 	\\
	Silver SMG 					 & SMG & 50m & 7 & 9x19mm 			& 34 (Box) 		& 20 & 							& 1.5 	& 5 gp 	\\
	Blish 19 					 & SMG & 50m & 7 & 11.5mm \newline Rahoo 	& 20/100 (Box) 	& 20 & 							& 4 	& 15 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }
}]
\twocolumn[{
	%{\bfseries Shotguns\\}
	\subsection{Shotguns}

	{\centering
	\begin{tabular}{p{2.5cm}lllp{2cm}llp{2.1cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	Wellington Pump 							& Shot & 20m & 6 & 12-G & 5 (Int) & 1 & Single Shot 				& 4 	& 3 gp \\
	Wellington Pump Sawed-Off 					& Shot & 10m & 5 & 12-G & 5 (Int) & 1 & Conceal [1], Single Shot 	& 2.5 	& 3 gp \\
	Wellington Over-Under 						& Shot & 20m & 6 & 12-G & 2 (Int) & 1 & [2] Barrels 				& 3.5 	& 2 gp \\
	Wellington Sawed-Off 						& Shot & 10m & 5 & 12-G & 2 (Int) & 1 & Conceal [1], [2] Barrels 	& 2 	& 2 gp \\
	Kuropatkin Revolving 						& Shot & 20m & 6 & 12-G & 5 (Cyl) & 1 &  							& 4 	& 6 gp \\
	Shotgun 									& Shot & 10m & 5 & 12-G & 5 (Cyl) & 1 & Conceal [1] 				& 2.5 	& 6 gp \\
	Kuropatkin Cut-Down Lancaster E1M1 			& Shot & 20m & 6 & 12-G & 4 (Int) & 1 & [4] Barrels 				& 4 	& 8 gp \\
	Tompkins ‘09 Lever Action 					& Shot & 20m & 6 & 12-G & 5 (Int) & 1 & Single-Action 				& 4 	& 5 gp \\
	Burgh “Constable”Folding Shotgun 			& Shot & 20m & 6 & 12-G & 6 (Int) & 1 & Conceal [1], Single Shot 	& 4 	& 8 gp \\
	Herstal “Hunchback” Self-Feeding Shotgun 	& Shot & 20m & 6 & 12-G & 4 (Int) & 4 & Reliable 					& 4 	& 10 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	%{\bfseries Machineguns\\}
	\subsection{Machineguns}

	{\centering
	\begin{tabular}{p{2.5cm}lllp{2cm}llp{2cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	Kurschat-Lickert MG71 	& MG & 350m & 10 & 13mm 		& Belt 		& 30 	& Brace [3] 			& 15 	& 75 gp \\
	Kurschat-Lickert MG66 	& MG & 350m & 10 & 7.92x57mm 	& Belt 		& 40 	& Brace [3], Reliable 	& 12 	& 50 gp \\
	Rahoo BZK-50 			& MG & 350m & 10 & 12.7x99 		& Belt 		& 20 	& Brace [3], Reliable 	& 25 	& 20 gp \\
	Rahoo BZK-30 			& MG & 350m & 10 & 7.62x63mm 	& Belt 		& 30 	& Brace [3], Reliable 	& 20 	& 18 gp \\
	Rahoo BZK-303 			& MG & 350m & 10 & .303 Albish 	& Belt 		& 30 	& Brace [3], Reliable 	& 20 	& 20 gp \\
	Type 702 				& MG & 350m & 10 & 7.92x57mm 	& 30 (Box) 	& 15 	& Brace [3] 			& 8 	& 15 gp \\
	Janovich 1488 			& MG & 350m & 10 & 7.62x33mm 	& Belt 		& 120 	& Brace [2], Ludicrous 	& 30 	& 70 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }
}]
\twocolumn[{
	%{\bfseries Rifles\\}
	\subsection{Rifles}

	{\centering
	\begin{tabular}{p{2.6cm}lllp{2cm}llp{2cm}ll}
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\ \hline \\
	Abitante VB/15 											& Rifle 	& 150m 	& 8 & 7.92x57mm 	& 5 (int) 		& 1 	& Longshot, Single Shot, Reliable 						& 4 	& 2 gp 	\\
	Bacci 79 LMG 											& Rifle/MG 	& 200m 	& 9 & 7.92x57mm 	& 20/80 (Box) 	& 20 	& Brace [1], Longshot 									& 8 	& 15 gp \\
	Schwartzvald C5 										& Rifle 	& 150m 	& 8 & 7.92x57mm 	& 8 (Int) 		& 1 	& Longshot, Single Shot 								& 4 	& 5 gp 	\\
	Schwartzritter C6 										& Rifle 	& 150m 	& 8 & 7.92x57mm 	& 10 (Box) 		& 3 	& Longshot, Reliable 									& 4 	& 8 gp 	\\
	Volksverteidiger C10 									& Rifle 	& 100m 	& 7 & 7.92x33mm 	& 30 (Box) 		& 20 	&  														& 3.5 	& 15 gp \\
	Flammenschwert C7 										& Rifle 	& 150m 	& 8 & 7.92x57mm 	& 20 (Box) 		& 25 	& Longshot 												& 6 	& 20 gp \\
	Hexenjäger EK21 										& Rifle 	& 100m 	& 7 & 7.62x33mm 	& 15 (Box) 		& 1 	& Night Sights 											& 2 	& 12 gp \\
	Type 70 Qi-Ar-Mi 										& Rifle 	& 150m 	& 8 & 7.92x57mm 	& 8 (Int) 		& 1 	& Jams [2], Longshot, Single Shot 						& 4 	& 2 gp 	\\
	Westridge-Ansem IE Mk II 								& Rifle 	& 150m 	& 8 & .303 Albish 	& 10 (Box) 		& 1 	& Longshot, Single Shot 								& 4 	& 5 gp 	\\
	Republican Arms Model 9 Condor 							& Rifle 	& 200m 	& 8 & 17.56mm 		& 6 (Box) 		& 1 	& Longshot, Reliable, Single Shot 						& 6 	& 18 gp \\
	Republican Arms “King Chuck” 							& Rifle 	& 200m 	& 8 & 12.7x99mm 	& 5 (Box) 		& 1 	& Longshot, Single Shot 								& 10 	& 20 gp \\
	Nelson Deluge 											& Rifle/MG 	& 60m 	& 7 & 6x16mm 		& 120 (Box) 	& 40 	& Jams [1] 												& 6 	& 8 gp 	\\
	Republican Arms 1776 Special 							& Rifle 	& 200m 	& 8 & 17x76mm 		& 1 (Int) 		& 1 	& Longshot, Reliable, Single Shot 						& 8 	& 55 gp \\
	Republican Arms “Volgin Contract” Lever Action Model 95 & Rifle 	& 150m 	& 7 & 7.62x54mm 	& 5 (Int) 		& 1 	& Single-Action 										& 4 	& 6 gp 	\\
	UC “Tipper” Field Rifle 								& Rifle 	& 100m 	& 8 & 7.62x54mm 	& 25 (Box) 		& 15 	& Longshot 												& 4 	& 16 gp \\
	Rahoo R-32 												& Rifle 	& 150m 	& 8 & 7.62x63mm 	& 5 (Int) 		& 1 	& Longshot, Single Shot 								& 4 	& 4 gp 	\\
	Rahoo SLR-58 											& Rifle 	& 150m 	& 8 & 7.62x63mm 	& 8 (Int) 		& 1 	& Longshot												& 4 	& 10 gp \\
	Rahoo SLC-60 											& Rifle 	& 100m 	& 7 & 7.62x33mm 	& 15/30 (Box) 	& 1 	& 														& 2 	& 8 gp 	\\
	Rahoo SLC-60B “Bushwhacker” 							& Rifle 	& 50m 	& 7 & 7.62x33mm 	& 15/30 (Box) 	& 1 	& Silenced, Single Shot (if used with Cold Load Rounds)	& 2 	& 12 gp \\
	Rahoo ASGR 												& Rifle 	& 50m 	& 7 & 9x33mm 		& 30 (Box) 		& 20 	& Jams [2]												& 3.5 	& 6 gp 	\\
	Herstal Monitor Rifle 									& Rifle/MG 	& 200m 	& 8 & 7.62x63mm 	& 20 (Box) 		& 15 	& Brace [1]												& 8 	& 15 gp \\
	Godfray Saboteur 										& Rifle 	& 50m 	& 7 & 11.5mm Rahoo 	& 7 (Box) 		& 1 	& Silenced, Single Shot									& 2 	& 10 gp \\
	Albish Royal Arms Squad Support Rifle 					& Rifle/MG 	& 200m 	& 9 & .303 Albish 	& 30/100 (Box) 	& 20 	& Brace [1], Longshot									& 8 	& 18 gp \\
	\end{tabular}\\ }
}]
\twocolumn[{

	{\centering
	\begin{tabular}{p{2.5cm}lllp{2cm}llp{2cm}ll}
	Eugene-Longhill Lever-Action Rifle 						& Rifle 	& 100m 	& 7 & 9x33mm 		& 12 (Int) 		& 1 	& Reliable, Single Action								& 4 	& 8 gp 	\\
	Eugene-Longhill “Bitterroot” Lever-Action Carbine 		& Rifle 	& 50m 	& 7 & 9x33mm 		& 6 (Int) 		& 1 	& Conceal [1], \newline Single-Action							& 2 	& 6 gp 	\\
	Eugene-Longhill “Good Mother” Falling-block Rifle 		& Rifle 	& 200m 	& 8 & 13mm 			& 1 (Int) 		& 1 	& Longshot, Single-Shot									& 6 	& 14 gp \\
	McCollum Arms 4-Bore Stopping Rifle 					& Rifle 	& 100m 	& 8 & 26.7x100mm 	& 1 (Int) 		& 1 	& Reliable, Single Shot									& 6 	& 8 gp 	\\
	PTSD Anti-Tank Rifle 									& Rifle 	& 200m 	& 9 & 14.5x114mm 	& 1 (Int) 		& 1 	& Brace [1], Longshot, Single-Shot						& 12 	& 20 gp \\
	\\ \hline \\
	Weapon & Type & Range & MTN & Round & Cap. & RoF & Spec. & WT & Cost\\ 
	&&&&& (Mag Type) \\
	\end{tabular}\\[2\baselineskip] }

	%{\bfseries Ammunition\\}
	\subsection{Ammunition}

	{\centering
	\begin{tabular}{llllll}
	Round & Damage & AP & Special & Cost/round & Type  \\ \hline \\
	6x16mm 						& 3 	& 2 	& 					& 1/10th cp 		& Pistol Cartridge 	\\
	7.5mm 						& 5 	& 4 	& 					& 1/5th cp 			& Pistol Cartridge 	\\
	8x22mm 						& 4 	& 1 	& 					& 1/10th cp 		& Pistol Cartridge 	\\
	9x17mm 						& 6 	& 1 	& 					& 1/5th cp 			& Pistol Cartridge 	\\
	9x19mm 						& 6 	& 2 	& 					& 1 cp 				& Pistol Cartridge 	\\
	9x29mm 						& 7 	& 1 	& 					& 1/2th cp 			& Pistol Cartridge 	\\
	10mm 						& 6 	& 3 	& 					& 1 cp 				& Pistol Cartridge 	\\
	11.5mm Rahoo 				& 7 	& 2 	& 					& 1 cp 				& Pistol Cartridge 	\\
	9x33mm 						& 7 	& 5 	& 					& 2 cp 				& Pistol Cartridge 	\\
	17.56mm 					& 9 	& 6 	& BP				& 8 cp 				& Pistol Cartridge 	\\
	12 Gauge Buck 				& 12\dag& Shot 	& 					& 4 cp 				& Shotgun Shell 	\\
	12 Gauge Slug 				& 12 	& 2 	& 					& 6 cp 				& Shotgun Shell 	\\
	7.62x33mm 					& 7 	& 8 	& 					& 3 cp 				& Rifle Cartridge 	\\
	7.92x33mm 					& 7 	& 12 	& 					& 2 cp 				& Rifle Cartridge 	\\
	.303 Albish 				& 8 	& 14 	& 					& 2 cp 				& Rifle Cartridge 	\\
	7.62x54mm 					& 8 	& 16 	& 					& 1 cp 				& Rifle Cartridge 	\\
	7.62x63mm 					& 8 	& 18 	& 					& 3 cp 				& Rifle Cartridge 	\\
	7.92x57mm 					& 9 	& 15 	& 					& 4 cp 				& Rifle Cartridge 	\\
	12.7x99 (.50) 				& 15 	& 15 	& Over-Sized		& 3 cp 				& Rifle Cartridge 	\\
	13mm 						& 15 	& 20 	& Over-Sized		& 6 cp 				& Rifle Cartridge 	\\
	26.7x100mm 					& 25 	& 3 	& Over-Sized, BP	& 1 sp 				& Rifle Cartridge 	\\
	17x76mm 					& 20 	& 25 	& Over-Sized		& 8 sp 				& Rifle Cartridge 	\\
	14.5x114mm 					& 20 	& 35 	& 					& 10 sp 			& Rifle Cartridge	\\
	Caster Shell 				& NA 	& NA 	& 					& See Caster Shells & Sorcery Cartridge	\\
	Normal Magazine/Clip 		& - 	& - 	& 					& 3 cp 				& 					\\
	Expanded/Drum Magazine 		& - 	& - 	& 					& 1 sp 				& 					\\
	Cylinder Speedloader 		& - 	& - 	& 					& 1 cp 				& 					\\
	Stripper Clip 				& - 	& - 	& 					& 1/2 cp 			& 					\\
	\\ \hline \\
	\end{tabular}\\ }
	\dag Any Hit with more than 0 BS with a Shot Round will automatically hit the two adjacent
	Hit Locations (see Critical Hits, above, Rolls of 10 or 1 only count 9 or 2 as adjacent)) on
	the Random Hit Table as well. 0 BS hits only hit Target Location.\\
}]
\twocolumn[{
	%{\bfseries Firearm Special Qualities\\}
	\subsection{Firearm Special Qualities}

	{\centering
	\noindent
	\begin{tabular}{lp{14cm}}
	Special Quiality 	& Effect \\ \hline \\
	Brace [X] 			& While this weapon is braced against a surface or mounted on a bipod, etc. (see Aim for more details) the weapon’s TN is reduced by X.\\ \\
	BP 					& This round uses black-powder propellants, rather than modern nitro based propellant. BP Cartridges cannot use Hot Load Rounds.\\ \\
	{[}X{]} Barrels 	& This weapon can make a Multi-Barrel Shot using a number of barrels up to [X].\\ \\
	Jams [X] 			& When shooting, a weapon with this quality does not need a critical failure to Jam. It instead Jams if, during a roll involving the firearm, more 1s are rolled than successes. Additionally, it converts [X] failed dice to 1s. GMs can add this condition to firearms based on environment, maintenance, ammo quality, etc.\\ \\
	Longshot 			& Attacks benefitting from Aim enjoy a -1 bonus to TN.\\ \\
	Ludicrous 			& The Area of Suppression of this weapon is doubled. Also, when using Burst or Auto fire, this weapon inflicts twice as many Hits as normal.\\ \\
	Over-Sized 			& Any Chest (Non-Vitals) hit with this round automatically becomes a Chest (Vitals) hit.\\ \\
	Reliable 			& Increase the number of 1s needed for a Malfunction to occur by 2.\\ \\
	Single-Action 		& Can be Fan-Fired, giving a ROF of 3 at the cost of -4 dice on the attack. Can only shoot once with Regular Fire instead of twice, unless Fan-Fired. You cannot benefit from Aim while Fan-Firing.\\ \\
	Single Shot 		& Can only shoot once with Regular Fire instead of twice.\\ \\
	Silenced 			& Requires an Awareness Test at RS 3 to notice firing within 50m, RS 1 within 10m. Impossible to hear further away.\\ \\
	Shot 				& Damage is Halved against any sort of armor. -1 Damage per Range Increment past the first.\\ \\
	Conceal [X] 		& Can be hidden on your person with a proper holster, depending on your clothing. A weapon with Conceal [1] can be hidden under a coat, [2] can be hidden under a shirt, [3] can be hidden in your pants pocket, and [4] can fit in your sleeve. Rifles with Conceal [1] are treated as SMGs/Shotguns for the purposes of shooting at Point Blank Range.\\ \\
	Night Sights 		& Ignore the first +1 Defense a target receives from dim lighting.\\
	\\ \hline
	\end{tabular}\\[\baselineskip] }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	%{\bfseries Weapon Enhancements\\}
	\subsection{Weapon Enhancements}

	\noindent{\centering
	\begin{tabular}{lp{10cm}p{1.3cm}p{1.9cm}}
	Enhancement & Effect & Weight & Cost \\ \hline \\
	Marksman's Scope & Against targets further than 10m away, Aiming with this weapon gains an additional +1 bonus per Action. & 1 & 1 gp \\ \\
	Detachable Silencer & Weapon gains Silenced quality, and suffers -2 damage. Has a chance of breaking with each automatic fire action (Burst, Full Auto, Suppression, etc), roll 1d10. On a 1, it breaks. As a rule of thumb, incompatible with weapons using a cylinder feeding mechanism. & .5 & 10 sp \\ \\
	Custom Parts & Ignore first Jam result from Critical Failure or the Jam quality. & 0 & 1 gp \\ \\
	Stealth Adaption & This weapon turns invisible along with the bearer when using a Cloaking Device. & 0 & 10 gp \\ \\
	Detachable Pistol Stock & Changes a pistol’s used proficiency to SMG, multiples the range by 3, removes Conceal. These changes are lost once the stock is removed. Takes an action to put on or remove. Not all pistols can accept a stock factory, but most can be modified to fit one (1 gp or a RS 2 Craft skill check). Often doubles as a holster. Attached stocks can hold a loaded pistol of the same model as a backup. & 1 & 1 gp \\ \\
	Tritium Sights & Adds the Night Sights rule to your firearm. & 0 & 5 gp \\ \\
	Foregrip & Reduce the penalty to firing this weapon during or after Movement by -1. & .5 & 5 sp \\ \\
	Hot Load Rounds & +2 AP, +10\% range, Decrease the RS of the test to hear the weapon firing by 2. & 0 & +100\% Ammo Cost \\ \\
	Cold Load Rounds & -2 AP, -10\% Range, Increase the RS of the test to hear a silenced weapon firing by 2. & 0 & +100\% Ammo Cost \\ \\
	Rifle Cutdown & This modification cannot be reversed, and can generally only be used on rifles with the Single-Shot Rule. Remove Longshot or Brace rules from the weapon, and add Conceal [1]. Reduce the weapon’s range to 10\% of its original range. & -1/2 Weapon WT & 1 sp or RS 2 Craft \\
	\\ \hline \\
	\end{tabular}\\ }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	\subsection{Granades}

	{\bfseries Explosive Grenades\\}
	Grenades have fuses measured in Phases. When thrown as detailed in the Throw action, they
	will sit dormant until the aforementioned number of Phases passes and they explode.
	Remember: The defender has no way of knowing how long the fuse on a grenade is!
	Grenades always explode at the very beginning of the Phase. Every character within the blast
	radius suffers damage from the explosion. Roll Hit Location a number of times equal to the
	second number in the Damage section, and apply the first number in Damage to each section
	rolled. In the event that multiple Hits land in the same location, treat these like normal
	attacks--the second Hit adds half of its damage to the first Hit’s total. Only ever resolve one Hit
	per Location.\\
	Critical Hits count as an automatic miss for the purposes of explosions.
	If a character is in Cover, any Hit Results that land on a Covered body part are not
	re-rolled--they automatically miss.\\
	If a character goes Prone or uses Active Defense, reduce the second number by one.
	For every multiple of the Blast Radius away that a character is, halve the first number in
	Damage and reduce the second number by one. Once the second number is 0, the explosion is
	no longer effective.\\
	If the Fuse entry has multiple numbers, it means the Phase length of the Fuse can be chosen
	before throwing the grenade.\\


	{\centering
	\noindent
	\begin{tabular}{llllll}
	Grenade & Damage & Blast Radius & Throwing Range & Fuse & Cost \\ \hline \\
	Type 42 Stielhandgranate 			& 20/4 & 5m		& 40m & 2		& 1 sp \\			
	Type 19 Grenade  					& 15/4 & 5m		& 30m & 2/\linebreak[0]4/\linebreak[0]6/\linebreak[0]8	& 2 sp \\		
	Type 42x3 “Bundle” Stielhandgranate & 60/4 & 5m		& 10m & 2		& 3 sp \\						
	M27 Grenade 						& 20/4 & 5m		& 30m & 2/4		& 2 sp \\
	Bongland 77 						& 20/4 & 10m	& 20m & 1		& 4 sp \\	
	\\ \hline
	\end{tabular}\\[\baselineskip] }

	{\bfseries Specialty Grenades\\}
	Specialty Grenades are similar to explosive ones, except that they have a different function than
	to explode.\\

	{\centering
	\noindent
	\begin{tabular}{p{1.6cm}p{12cm}p{1.5cm}ll}
	Grenade & Effect & Throwing Range & Fuse & Cost \\ \hline \\
	Smoke Grenade & Creates 10m wide cloud of smoke. +2 Defense for any characters attacked through cloud. Does not affect Suppressive Fire. Cloud persists for 9 phases. & 40m & 1 & 6 cp \\ \\
	Flashbang & Blinds and Deafens any character within 10m for 3 Phases after detonation. Causes 20 Stun to any character within 10m. & 40m & 1 & 1 sp \\ \\
	Flare & Illuminates an area of 50m, bringing lighting levels up 1 (Darkness to Poor Lighting, Poor to Well Lit, Well to Bright, Bright to Blinding). An Area of 10m from the flare is brought up two stages (Darkness to Well, Well to Blinding, etc.) Flare grenades fired from a launcher go off one phase after being fired, and have an attached parachute to allow them to stay in the air. Flare persists for 10 phases. & 40m & 1 & 1 sp\\ \\
	Strangle Gas & Creates 10m wide cloud of tear gas. For every phase spent in the cloud without proper protection, a character suffers +3 pain, which stacks for every consecutive phase spent in the cloud. Once outside of the cloud, -1 pain per phase until all pain from the tear gas is gone.Cloud persists for 9 Phases. & 40m & 1 & 2 sp\\ \\
	Poison Gas & Creates 10m wide cloud of poisonous gas. Any character without a respirator within this cloud at the end of a Phase must make an HLT test at RS 4 or suffer 10 Pain and lose 1 HLT. At 0 HLT, character dies. Cloud persists for 8 phases. & 40m & 2 & 6 sp\\ \\
	Babel Grenade & Creates 10m wide swirling vortex of haunted ghost magic. +2 Defense for any characters attacked through vortex. Any character who moves through the vortex or ends a Phase inside of it must make an WIP test at RS 3 or suffer 10 Pain and lose 1 WIP. At 0 WIP, character dies. Vortex persists for 9 phases. 

	Overlapping Babel Fields increase WIP test RS by
	1 each. Only make 1 test.

	WIP lost in this manner returns after 24 hours.
	While moving through a Babel Field, a character’s
	MOB is reduced by half. & 40m & 1/3 & 8 sp\\
	\\ \hline
	\end{tabular}\\[\baselineskip] }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	%{\bfseries Grenade Launchers\\}
	\subsection{Grenade Launchers}
	Grenade Launchers that miss deviate as-per grenades, but triple the distance.\\

	\noindent{\centering
	\begin{tabular}{p{3cm}llllllp{2cm}ll}
	Weapon & Type & Range & MTN & Round & Capacity & RoF & Spec. & WT & Cost \\ \hline \\
	C602 Fackelwerfer 		& Pistol 	& 50mm & 6 	& 27mm Specialty 	& 1 (Int) & 1 & Singleshot, Conceal [2] & 0.5 	& 1 gp\\ \\
	C602G Granatpistole 	& Pistol 	& 50mm & 8 	& 30mm Grenade 		& 1 (Int) & 1 & Singleshot, Conceal [2] & 1 	& 2 gp\\ \\
	C802 Granatwerfer 		& Rifle 	& 100m & 7 	& 30mm Grenade 		& 4 (Int) & 1 & Single-Shot, Reliable 	& 4 	& 8 gp\\ \\
	Black Shore “Boop” 71 	& Rifle 	& 100m & 7 	& 40mm Grenade 		& 1 (Int) & 1 & Single Shot, Reliable 	& 3 	& 3 gp\\ \\
	Republican Arms Rhodes 	& Rifle 	& 100m & 8 	& 40mm Grenade 		& 6 (Cyl) & 1 &  						& 6 	& 12 gp\\ \\
	Tekidanto II 			& Rifle 	& 300m & 10 & Type 19 Grenade 	& 1 (Int) & 1 & Single Shot, Brace [2] 	& 3 	& 2 gp\\
	\\ \hline
	\end{tabular}\\[\baselineskip]

	\begin{tabular}{lllll}
	Grenade & Damage & Blast Radius & Fuse & Cost \\ \hline \\
	27mm Specialty 	& \dag 	& \dag 	& Impact 	& 1 sp 		\\
	30mm HE 		& 15/4 	& 5m 	& Impact 	& 2 sp 		\\
	30mm Specialty 	& \dag 	& \dag 	& Impact 	& 1sp \dag 	\\
	40mm HE 		& 20/4 	& 5m 	& Impact 	& 3 sp 		\\
	40mm Specialty 	& \dag 	& \dag 	& Impact 	& 2 sp \dag \\
	Type 19 		& 15/4 	& 5m 	& 2/4/6/8 	& 2 sp 		\\
	\\ \hline
	\end{tabular}\\ }
	\dag The effects of this grenade are the same as a Specialty Grenade. The cost is the listed amount
	plus the cost of a regular Specialty grenade of that type. The Fuse type is always Impact,
	however.\\
	A direct hit with a launched grenade deals 7 Bludgeoning Damage to the hit location.\\
	\vspace{1cm}

	%{\bfseries Rocket Launchers\\}
	\subsection{Rocket Launchers}

	\noindent{\centering
	\begin{tabular}{p{3.9cm}llllllp{1.8cm}ll}
	Weapon & Type & Range & MTN & Round & Capacity & RoF & Spec. & WT & Cost \\ \hline \\
	Panzerklatschen 				& Rocket & 75m & 8 & 88mm Rocket 	& 1 (Int) & 1 & Single Shot 				& 8 & 1 gp \\
	Panzerknüppel 					& Rocket & 30m & 7 & 149mm Charge 	& 1 (Int) & 1 & Single Shot, Single Use 	& 4 & 1 sp \\
	Canebreak “Endodontist” M77 	& Rocket & 75m & 7 & 57mm Rocket 	& 1 (Int) & 1 & Single Shot 				& 6 & 1 gp \\
	Type 4 Launcher 				& Rocket & 9sm & 8 & 70mm Rocket 	& 1 (Int) & 1 & Single Shot 				& 8 & 10 sp\\
	\\ \hline
	\end{tabular}\\[\baselineskip]

	\begin{tabular}{lllll}
	Rocket & Damage & Blast Radius & Special & Cost \\ \hline \\
	88mm AP Rocket 	& 100/1 & 2m & Armor Piercing 20, Oversized & 24 sp \\
	149mm Charge 	& 150/1 & 1m & Armor Piercing 25, Oversized & 24 sp \\
	57mm HE Rocket 	& 25/4 	& 5m &  							& 18 sp \\
	57mm AP Rocket 	& 80/1 	& 2m & Armor Piercing 20, Oversized & 18 sp \\
	70mm AP Rocket 	& 90/1 	& 2m & Armor Piercing 20, Oversized & 20 sp \\
	\\ \hline
	\end{tabular}\\[\baselineskip] }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	%{\bfseries Flamethrowers\\}
	\subsection{Flamethrowers}
	Flamethrowers are fired using the Suppressive Fire action. The arc of the flamethrower’s effect
	is determined in the same way as normal (by Rate of Fire) but when using a flamethrower, a
	character gains a +8 to hit any target within the arc, and instead of hitting characters with a
	bullet, the attack lights them on fire. You cannot benefit from Aim with a flamethrower.
	Flamethrowers use the Flamethrower Proficiency.\\
	The damage is delivered in the following format: X/Y/TN5\\
	X is the number of random hit locations which are affected by flame.\\
	Y is the number of dice rolled for each Hit Location every Phase to inflict burn damage. These
	dice are always rolled at TN 5.\\
	Burning damage continues indefinitely until smothered.\\

	{\centering
	\begin{tabular}{lllllll}
	Flamethrower & Damage & RoF & Range & Ammunition & WT & Cost \\ \hline \\
	Type 9 Flammenwerfer 		& 4/5/TN5 & 20 & 50m & 12 Actions & 16 & 20 gp \\
	BZI Flamechucker 			& 6/5/TN5 & 30 & 30m & 12 Actions & 18 & 20 gp \\
	Bongland GDI Flamethrower 	& 3/5/TN5 & 15 & 70m & 10 Actions & 14 & 20 gp \\
	\\ \hline
	\end{tabular}\\[\baselineskip] }
}]

{\bfseries Caster Shells\\}

Casters are a very specialized category of firearm, usually incorporating whalebone latticework.
The weapons are designed to project arcane energies from specialized “Caster Shells,” which
essentially contain powerful magic. Caster Shells are rare and powerful--they can only be
produced by powerful mystics, and many of them drain the life-force of the creators, limiting their
production to only a handful over a lifetime.\\

Caster Shells are ranked in power from 13 to 1--with 13 being the strongest, and 1 being the
weakest. They are also categorized by their effect.\\
When players find Caster Shells, if the type is not specified (perhaps for plot purposes) you
should roll on the following table for the kind found. Those found on the open market should be
similarly rolled randomly.\\

{\bfseries Caster Shell Cost}: [Powerx5] sp\\

{\centering
\begin{tabular}{lll}
Roll 2d10 & Power & Type \\
--- & 1 	& Fire 			\\
2 	& 1 	& Fire 			\\
3 	& 1 	& Fire 			\\
4 	& 2 	& Fire 			\\
5 	& 2 	& Fire 			\\
6 	& 2 	& Fire 			\\
7 	& 3 	& Electricity 	\\
8 	& 3 	& Electricity 	\\
9 	& 3 	& Electricity 	\\
10 	& 4 	& Electricity 	\\
11 	& 4 	& Electricity 	\\
12 	& 5 	& Explosion 	\\
13 	& 6 	& Explosion 	\\
14 	& 7 	& Explosion 	\\
15 	& 8 	& Explosion 	\\
16 	& 9 	& Explosion 	\\
17 	& 10 	& Void 			\\
18 	& 11 	& Void 			\\
19 	& 12 	& Crush 		\\
20 	& 13 	& Warp 			\\
\\\hline
\end{tabular}\\[\baselineskip] }

{\bfseries Power Ranking\\}
1-13\\
(The power ranking of a Caster Shell determines the number of dice rolled for its effect.)\\

{\bfseries Effect Type\\}

\tabto{1}
Fire (Ignores Cover, inflicts X Fire Damage distributed randomly)\\
Electricity (Ignores metal/conductive Cover, inflicts X Electricity damage, distributed randomly)\\
Explosion (Explosion with Power X)\\
Void: Ignores AV, Inflicts X+BS Damage, Critical Hit disintegrates instantly\\
Crush: Inflicts X+BS+AV Damage, leaves no body\\
Warp: Shoot TN -1, Inflicts X Damage, ignoring AV. Triples the regular amount of pain.\\

\tabto{0}
{\bfseries Random Caster Effects\\}
Caster Shells aren’t perfect. Part of the reason they have only been sporadically popular is their
tendency for unpredictable side effects. When a Caster Shell is acquired, roll on the following
table for Random Effects. Caster shells purchased from very reliable dealers or directly from
their creators may be exempt from this.\\ 

\noindent{\centering
\begin{tabular}{lp{7cm}}
1 	& Life drain. Shooter suffers 1d10 Pain. If the Pain from this effect ever exceeds WIP, character goes comatose for 2d10 days. If Pain from this effect exceeds WIPx2, character must make a HLT test at RS 4 or die.\\
2 	& Minor Life Drain. Shooter suffers 1d10 Stun.\\
3 	& Electric Shock. Shooter suffers 1d10/2 Stun.\\
4 	& No Effect.\\
5	& No Effect. \\
6	& No Effect. \\
7	& No Effect. \\
8	& Overcharge. Shooter suffers 1d10/2 Stun, but Shell has +1 Power. \\
9	& Supercharge. Shooter suffers 1d10 Stun, but Shell has +2 Power. \\
10	& Superior Shell. Shell has +3 Power. \\
\end{tabular}\\[\baselineskip] }

\twocolumn[{
	\setlength{\parindent}{10.00002pt}
	\section{Armors}

	%{\bfseries Snapcoat\\}
	\subsection{Snapcoat}
	Made of a non-newtonian fabric that stiffens against impact, snapcoat is a popular form of armor
	for those less inclined to cover themselves in ballistic steel plates. Though effective enough to
	slow most pistol cartridges into the realm of survivability, it should not be trusted to repel
	anything of particularly high caliber. Totally useless against rifle cartridges. Thicker “heavy”
	vests are available, though these are somewhat cumbersome.\\

	\noindent{\centering
	\begin{tabular}{llllll}
	Armor & AV & Weight & Coverage & Spec. & Cost \\ \hline \\
	Snapcoat Coat					& 4 	& 1 & Chest, Abdomen, Arms 			& Half AV vs Melee, Layer [1] 	& 15 sp \\
	Snapcoat Vest					& 4 	& 0 & Chest, Abdomen 				& Half AV vs Melee, Layer [1] 	& 10 sp \\
	Snapcoat Jumpsuit/Trenchcoat	& 4 	& 1 & Chest, Abdomen, Arms, Legs 	& Half AV vs Melee, Layer [1] 	& 2 gp 	\\
	Heavy Snapcoat Vest				& 6 	& 2 & Chest, Abdomen 				& Half AV vs Melee, Layer [1] 	& 1 gp 	\\
	Snapcoat Sleeves				& 4 	& 0 & Arms 							& Half AV vs Melee, Layer [1] 	& 6 sp 	\\
	Snapcoat Leggings				& 4 	& 0 & Legs 							& Half AV vs Melee, Layer [1] 	& 6 sp 	\\
	Snapcoat Hat					& 4 	& 0 & Head 							& Half AV vs Melee, Layer [1] 	& 4 sp 	\\
	Additional Layering!			& +1 	& 1 & Any one location 				& Layer [1] 					& 10 sp \\
	\\ \hline
	\end{tabular} \\[\baselineskip] }

	%{\bfseries Sneaking Suit\\}
	\subsection{Sneaking Suit}
	Made of a combination of experimental plastics, silk and ceramics, and designed with
	cutting-edge polymers in the soles and other contact surfaces, the Sneaking Suit was originally
	designed by Desotia’s AMDRA group for use by RIS agents in the field. The sneaking suit
	combines impressive ballistic protection with noise reduction and silhouette-breaking
	camouflage patterns on light-sensitive surfaces--optimizing the pattern automatically to account
	for light exposure and humidity. The “Skull Helmet” includes a bulletproof skullcap, a facemask
	made of air-conditioned plastic, anti-fog auto-darkening eyepieces, and the same camouflage
	system as the rest of the suit.\\
	The whole suit is compatible with the proprietary Timber Wolf Active Stealth System (sold
	separately) which can effectively turn the wearer invisible for a short period of time.\\

	\noindent{\centering
	\begin{tabular}{lllp{3cm}ll}
	Armor & AV & Weight & Coverage & Spec. & Cost \\ \hline \\
	Sneaking Suit 	& 4 & 3 & Chest, Abdomen, Arms, Legs 	& Stealth Enhancements [2], Stealth Suite compatible & 5 gp\\
	Skull Helmet 	& 8 & 1 & Head 							& Stealth Enhancements [1], Stealth Suite compatible & 5 gp\\
	\\ \hline
	\end{tabular} \\[\baselineskip] }

	%{\bfseries Battle Plate\\}
	\subsection{Battle Plate}
	Made of carefully designed plates of Whaleblood-Steel, plastics, and snapcoat fibers, Battle
	Plate was developed for boarding actions and close-quarters combat in the intense environment
	of Aethership interiors. Combining excellent protection with adequate coverage, the armor
	nevertheless is no match for armor-piercing rifle cartridges. Thankfully, those are rarely seen in
	the aether due to the risk of penetrating the hulls of ships, potentially flooding the compartments
	with Underspace Dust.\\

	\noindent{\centering
	\begin{tabular}{llllll}
	Armor & AV & Weight & Coverage & Spec. & Cost \\ \hline \\
	Battle Cuirass 			& 10 & 3 & Chest, Abdomen 	&  				& 10 gp \\
	Heavy Battle Cuirass 	& 10 & 6 & Chest, Abdomen 	& Bulletproof 	& 30 gp \\
	Battle Leggings 		& 10 & 6 & Legs 			&  				& 3 gp 	\\
	Battle Arms 			& 10 & 4 & Arms 			&  				& 3 gp 	\\
	Battle Helmet 			& 10 & 1 & Head 			& -2 Perception & 3 gp 	\\
	\\ \hline
	\end{tabular} \\[\baselineskip] }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	%{\bfseries Steel Plate\\}
	\subsection{Steel Plate}
	A favourite of the cheap, the desperate, and the luddite, this crude Steel plating is an easy
	method of armouring oneself against harm against person and dignity. Mostly against person. A
	far cry from the stylish armor of old, but if it worked for Ned, it’ll work for you.\\

	\noindent{\centering
	\begin{tabular}{llllll}
	Armor & AV & Weight & Coverage & Spec. & Cost \\ \hline \\
	Steel Cuirass 	& 6 & 3 & Chest, Abdomen &  				& 1 gp 	\\
	Steel Leggings 	& 6 & 6 & Legs 			 &  				& 10 sp \\
	Steel Arms 		& 6 & 4 & Arms 			 &  				& 10 sp \\
	Steel Helmet 	& 6 & 1 & Head 			 & -2 Perception 	& 10 sp \\
	\\ \hline
	\end{tabular} \\[\baselineskip] }

	%{\bfseries Clayplate\\}
	\subsection{Clayplate}
	Using multiple laminated layers of ceramics, Clayplate, or Flowerpot as it is colloquially known
	dissipates the energy from firearms incredibly well, but falls short against sustained attack, or
	against large caliber weapons and melee attacks.\\

	\noindent{\centering
	\begin{tabular}{llllll}
	Armor & AV & Weight & Coverage & Spec. & Cost \\ \hline \\
	Clayplate Cuirass 		& 10 	& 3 & Chest, Abdomen 	& Ablative [3], Bulletproof 				& 10 gp \\
	Light Clayplate Cuirass & 8 	& 1 & Chest, Abdomen 	& Ablative [3], Bulletproof 				& 8 gp 	\\
	Heavy Clayplate Cuirass & 15 	& 6 & Chest, Abdomen 	& Ablative [3], Bulletproof 				& 15 gp \\
	Clayplate Leggings 		& 10 	& 6 & Legs 				& Ablative [3], Bulletproof 				& 3 gp 	\\
	Clayplate Arms 			& 10 	& 4 & Arms 				& Ablative [3], Bulletproof 				& 3 gp 	\\
	Clayplate Helmet 		& 10 	& 1 & Head 				& Ablative [3], Bulletproof, -2 Perception 	& 3 gp 	\\
	\\ \hline
	\end{tabular} \\[\baselineskip] }

	%{\bfseries Shields\\}
	\subsection{Shields}
	When using Shields, the character counts as having a certain amount of Cover, but unlike
	normal Cover, when an attack lands on an area protected by the Shield, it resolves, but with the
	AV of the shield added to the area. Attacks do NOT combine damage when hitting an area
	protected by a Shield, instead resolve each attack individually. This may cause the Shield to
	deteriorate more quickly because of the properties of the Ablative special rule.
	Shooting a weapon while carrying a shield incurs a -2 penalty, shooting a two-handed weapon
	while holding one is next to impossible.\\

	\noindent{\centering
	\begin{tabular}{llllllll}
	Shield Name & Bash TN & Block TN & AV & Coverage & Special & Wt. & Cost \\ \hline \\
	Ballistic Shield 	& 7(+1b) & 7 & 15 & Half 			& Ablative [1] 				& 8 	& 1 gp 	\\
	Bunker Shield 		& 8(+1b) & 7 & 20 & 3/4ths Cover 	& Bulletproof, Ablative [1] & 15 	& 		\\
	Doorway Shield 		& 8(+1b) & 7 & 15 & 3/4ths Cover 	& Ablative [1] 				& 12 	& 		\\
	\\ \hline
	\end{tabular} \\[\baselineskip] }

	\noindent{\centering
	\begin{tabular}{lp{13cm}}
	Special Quality & Effect \\ \hline \\
	Ablaitive & Any attack that inflicts more than 50\% of this armor’s AV reduces its AV by X after the attack has resolved. This cannot be repaired, you’ll have to buy a new one. \\ \\
	Bulletproof & Weapons do not benefit from AP against this armor. \\ \\
	Layer [x] & If worn on the same Hit Location as another armor component with equal to or higher AV, this armor increases the AV on that area by X. Multiple instances of Layer do not combine. \\ \\
	Stealth Enhancements [X] & While wearing this armor, you gain a bonus equal to X to any Stealth skill tests you make. Additionally, you count as having Camouflage X in dark, indoor or urban environments. \\
	\\ \hline
	\end{tabular} \\[\baselineskip] }
}]

%{\bfseries Backpacks\\}
\section{Backpacks}

\noindent{\centering
\begin{tabular}{p{3cm}p{3cm}ll}
Pack & Gear & Wt. & Cost \\ \hline \\
Marine Assault Pack 					& Grappling Hook, \newline Buoyancy Device, \newline Respirator 			& 4 & 1 gp 	\\ \\
Engineer’s Pack 						& Engineering Tools, \newline Buoyancy Device, \newline Respirator 		& 4 & 1 gp 	\\ \\
Stealth Suite [Requires Sneaking Suit] 	& Cloaking Device, \newline Respirator 							& 4 & 25 gp \\ \\
Luftpack 								& Jetpack, \newline Glider Module, \newline Buoyancy Device, \newline Respirator 	& 8 & 10 gp \\
\\ \hline
\end{tabular} \\[\baselineskip] }

{\bfseries Grappling Hook\\}
You can launch a grappling hook up to 50 meters, where it will attach itself to most surfaces
(including metal via magnets, wood, etc) and pull yourself up at a rate of 10m per phase.\\

{\bfseries Buoyancy Device\\}
You may activate this device instantly. An inflating balloon causes you to become “buoyant”
compared to the Dust of Underspace. You will float several meters above the surface
indefinitely.\\

{\bfseries Respirator\\}
This device contains a thirty-minute supply of oxygen that can be fed to the mouth and nose
through a facemask. It is usually integrated into a helmet. The oxygen flow can be toggled on
and off, but the respirator also has a simple filtration system so that regular air (or air with minor
contamination like dust, smoke, etc) can be filtered and breathed normally.\\

{\bfseries Engineering Tools\\}
This device allows for easy access to welding, steel-lathing, and other engineering tools without
much fiddling or rummaging. Essential for work in the field.\\

{\bfseries Cloaking Device\\}
This device works in conjunction with a Stealth Suit to turn the wearer effectively invisible for ten
minutes. The power cell for this device costs 10 gp, and weighs ten pounds. While invisible, a
character has Camouflage 10 in all environments, and gains a +20 bonus to stealth.
HOWEVER: Objects not treated with a Stealth Adaptation remain visible, and can be seen
normally. A floating gun is obviously a dead giveaway of the bearer’s presence, and negates the
bonuses. Equipment small enough to be stored in the backpack or in pouches on the suit are
exempt from this penalty.\\
Any sort of sharp impact (firing a weapon, tripping and falling, being hit by a weapon or bumped
into strongly) will cause the wearer to become visible for a full Round.\\

%{\bfseries Melee Weapons\\}
\section{Melee Weapons}

{\bfseries Conventional\\}
Conventional Melee Weapons have the advantage of passing through all forms of energy field.
Some rely on raw striking power to do their damage, whereas others can be used to target the
joints and cracks in armor. {\bfseries (All armor components in Ballad of the Laser Whales have
Joints for the purposes of Joint Thrust.)}\\

{\bfseries Moonblades\\}
Moonblades are made from the mysterious Moonstone, a sort of crystalline mineral which floats
unnaturally on the surface of Underspace, and can be collected with nets or poles. Moonblades
are weapons fashioned of this material and reinforced with metal for weight and handiness.
When charged with an energy current, glow with an unearthly light and become capable of
cutting through any material except for other moonstones. This is obviously extremely useful,
and moonblades are highly prized and very expensive. Most Moonblades are swords, intended
to be used in close quarters. Moonblade harpoons (Starpoons) are used to penetrate the thick
hide of Laser Whales, though, and are sometimes used as weapons by desperate (or deranged)
warriors.\\
Weapons with the Moonblade Tag ignore all AV of all varieties. They still have to overcome
target Toughness.\\

{\bfseries Heat Bayonets\\}
A sort of “poor man’s Moonblade,” a Heat Bayonet is a tungsten spike through which a current
is run that heats it up to the point where it can melt through most metals on contact. This makes
it especially useful against Snapcoat, as well as Battle Armor, which has many plastic
components which can melt like butter. These are usually attached to the ends of firearms,
though they can also be put at the end of poles. Often called Heat-Pikes, these polearms are
surprisingly useful in some situations, and many ships keep a stock of them for emergencies.\\

\twocolumn[{

	\noindent{\centering
	\begin{tabular}{p{2cm}p{1.2cm}llllp{1.3cm}p{3.3cm}ll}
	Weapon Name & Type & Hands & Reach & Swing & Thrust & Defense (Guard) & Special & Weight & Cost \\ \hline \\
	Moonblade Cutlass 		& Sword & 1H & S & 7(+1c) & 8(+0p) & 7(3) 	& Moonblade 							& 0.5 	& 120 gp\\ \\
	Moonblade Harpoon* 		& Spear & 1H & L & 8(+0c) & 7(+1p) & 8 		& Moonblade , \newline Fluid Thrusts 			&  		& 100 gp\\ \\
	Moonblade Harpoon 		& Spear & 2H & L & 6(+1c) & 6(+2p) & 6 		& Moonblade , \newline Fluid Thrusts, Hand-Off 	& 4 	& 100 gp\\ \\
	Moonblade Templar Sword & Sword & 1H & M & 7(+0c) & 7(+0p) & 7(0) 	& Moonblade , Draw 2, Cavalry Sword 	& 0.5 	& 150 gp\\
	\\ \hline \\
	Heat Bayonet (SMG) 			& Spear 		& 2H & S 	& 9(-3c) & 8(+6p) & 8 		& Heavy Weapon, \newline Thin Blade 																& 0 	& 1 gp \\ \\
	Heat Bayonet (Rifle) 		& Spear 		& 2H & L 	& 9(-3c) & 8(+6p) & 8 		& Heavy Weapon, \newline Thin Blade 																& 0 	& 1 gp \\ \\
	Heat Pike 					& Spear 		& 2H & VL 	& 9(-3c) & 8(+6p) & 8 		& Heavy Weapon, \newline Thin Blade 																& 0 	& 1 gp \\
	\\ \hline \\
	Knife (Large) 				& Dagger		& 1H & 1H 	& 6(-1c) & 6(+0p) & 9(0) 	& Draw 1, Light Blade 																	& 0 	& 7 cp \\ \\
	Degan 						& Dagger, Sword & 1H & 1H 	& 6(-1c) & 6(+1p) & 7(0) 	& Fluid Thrusts, \newline Light Blade 															& 0 	& 1 sp \\ \\
	Cutlass 					& Sword	 		& 1H & 1H 	& 7(+1c) & 8(+0p) & 7(3) 	&  																						& 0.5 	& 1 gp \\ \\
	Streitaxt 					& Blunt	 		& 2H & 2H 	& 7(+4c) & 8(+0p) & 8 		& AP Swing 2, Crushing 2, Heavy Weapon, Hook, Shock 2, Swinging Slot, Thrusting Slot 	& 2 	& 7 sp \\ \\
	Claymore (Basket Hilted) 	& Sword			& 1H & 1H 	& 7(+1c) & 7(+0p) & 7(4) 	& Tight Grip 2 																			& 0.5 	& 1 gp \\ \\
	Templar Sword 				& Sword 		& 1H & M 	& 7(+0c) & 7(+0p) & 7(0) 	& Draw 2, \newline Cavalry Sword 																& 0.5 	& 15 sp \\
	\end{tabular}\\[2\baselineskip] }
}]

\section{Prosthetics}

Laser Whales, as has been recently discovered, are dual-base life forms that incorporate both
silicate and carbon structures in their biology. Their nervous systems are enormously complex,
and can connect highly disparate materials--including crystalline and fleshy parts. For centuries
the nervous system of the Laser Whale has been treasured as a piece of natural art--massive,
branching, crystalline, and glowing with an unearthly light.\\
Now, in recent years, a new use for the beautiful material has been discovered. Whalenerve, if
treated carefully and arranged properly, can be bonded to exposed human nerves to create an
artificial extension to the nervous system that can then, using the material’s qualities, be
connected to artificial mechanical limbs, allowing them to be directed as naturally as one might
control one’s organic arm.\\
This was the birth of the field of automated prosthetics, which have now become mainstream,
and are widely available across the void. Given the amount of war and violence, there has been
no shortage of clients on which to test these new technologies. However, the technology is not
without its risks. Whalebone nerve replacements emit low-level radiation when charged, which
can be harmful over long periods of time, and can ultimately result in a host of negative
side-effects.\\
A character’s resistance to this radiation is called Purity. Each prosthesis lowers Purity,
increasing the risk of side-effects with subsequent enhancements.\\
When installing a prosthesis, the person obtaining a prosthetic rolls a Purity Test. Successes
must meet an RS equal to the Purity Loss of the prosthesis. Failure could mean Cybernetic
Rejection.\\

{\bfseries Purity/Compatibility\\}
Willpower+Health+Misc. (Hale and Hearty, etc.)
Represents the purity of the nervous system, and how total the control of a character over his
own body is. Reductions in Purity can result in nerve damage, phantom pain, psychosis, etc.\\

{\bfseries Purity Loss\\}
Every time you install a prosthesis, regardless of successful installation, you lose Purity based
on extremeness of prosthesis.\\

When installed, the limb it is replacing is Amputated (if it’s still there at all), inflicting a Severed
Limb bane there (or One-Eyed for an Eye, or Facial Deformity (8) for a full Facial
Reconstruction).\\
After installation, a prosthetic negates the effects of the Bane associated with it, though the
bane is still there.\\

Be sure to track total Purity Reduction, as Purity Reduction also directly affects Void Capacity.\\

{\bfseries Damaging Prosthetic Limbs\\}
Prosthetic Limbs have a TOU based on their Quality level, but otherwise suffer injury the same
way that regular limbs do. However, Prosthetic limbs, when damaged, suffer Stun and Specials,
but they do not suffer Pain or Bleed damage, since they have no nerve endings and there is no
blood to be spilled. Low Performance limbs have a TOU of 6, High Performance limbs have
TOU 8, and Extreme Performance limbs have TOU 10. Surgery tests versus Crippled Limbs
become Engineering tests. Repairing a wounded/damaged prosthetic is an Engineering test at
RS equal to the Wound level suffered.\\

\subsection{Prosthetic Limbs}

\subsubsection{Arm Prosthetics}
A replacement of the arm. Arm prosthetics are probably the most common in the Void, as a
great many people have suffered maiming at some point, and replacements allow them to live
and work normally.\\
Prosthetic arms do not normally increase the strength of the wearer, as they must be light
enough not to burden the wearer, and must also be mounted to flesh, meaning that excessive
force might rip them out of their sockets.\\
Prosthetic arms, lacking sensory nerves in flesh, suffer -4 dice to any test involving fine tactile
sensation, such as trying to pick a lock, feel for something in the dark, Quick-Drawing a weapon,
or anything along those lines. This penalty can be negated by purchasing the Tactile Sense
augmentation.\\

Strength augmentations are only available for Full limb replacement, which includes part of the
torso, thus reinforcing the arm’s connection to the body, making strength enhancements
possible.\\
When getting a second prosthetic Arm, the Purity Reduction for the new limb is Halved.\\

\noindent{\centering
\begin{tabular}{p{4cm}lp{1.7cm}}
Prosthetic 	& Monetary 	& Purity\\ 
			& Cost 		& Reduction \\ \hline \\
Prosthetic Hand/Forearm 		& 8 gp 	& -1 	\\ \\
Prosthetic Full Arm 			& 15 gp & -2 	\\ \\
Prosthetic Full Arm \newline w/ Torso Rig & 25 gp & -3 	\\ \\
Low Performance \newline (Civilian) 		& x1 	& - 	\\ \\
High Performance \newline (Military) 	& x3 	& - 	\\ \\
Extreme Performance \newline (Corporate) 	& x5 	& - 	\\
\\ \hline
\end{tabular}\\[\baselineskip ] }
Low Performance Prosthetics have one Augmentation Slots, and 6 TOU.\\
High Performance Prosthetics have two Augmentation Slots, and 8 TOU.\\
Extreme Performance Prosthetics have three Augmentation Slots, and 10 TOU.\\

\paragraph{Arm Augmentations} (Cost of Augment/Augmentation Slots Used)\\

\tabto{1}
	{\slshape Armored (5gp/1)\\}
	Your prosthesis has +4 AV. This augmentation can be taken multiple times.\\

	Strength (Full Arm w/ Torso Rig only) (12 gp/1)\\
	Your Strength for any action taken involving this limb (punch, climb, etc.) is 8, instead of your
	regular Strength. If the action would involve two arms, and you only have one augmented arm,
	your Strength counts as 6--or your base strength, whichever is higher--instead.\\

	{\slshape Combat Hand (5 gp/1)\\}
	Your hand is literally made of steel, and can move in ways that a human hand can’t. You count
	as having knuckledusters/gauntlet for the purposes of punching damage, and your STR counts
	as 2 higher for the purposes of Grappling Maneuvers.\\

	{\slshape Electrified Fist (8 gp/2)\\}
	You hand can be electrified, allowing you to channel electricity into your punches. Once the fist
	is activated, hitting an opponent with it causes +4 Stun, regardless of wounds inflicted. Keeping
	the fist in constant contact with an enemy, such as trapping them in a grapple, inflicts this stun
	once per Action. The fist can only be electrified for 5 Rounds, requiring a 10 Round recharge
	period.\\

	{\slshape Claw Hand (5 gp/1)\\}
	Your fingers are tipped with sharpened steel, able to rip and tear with ease. Your hook punches
	can now inflict cutting damage, and your straight punches can now inflict piercing damage.
	These forms of punches cannot benefit from the Superior version of the maneuver, and
	obviously don’t get damage bonuses from knuckledusters.\\

	{\slshape Free Rotation (2gp/1)\\}
	Your arm’s joints have no limit on their rotation, allowing unusual feats of flexibility. Add +3 dice
	to any test where this effect could come into play.\\

	{\slshape Tools (10 gp/1)\\}
	You have a set of integrated tools in your arm for a specific purpose. Pick one non-Knowledge
	Skill. You gain a +2 bonus to performing that Skill so long as this limb is still usable and can be
	applied to the task. You can take this augmentation multiple times, for different skills each time.\\

	{\slshape Grappling Hook (Full Arm w/ Torso Rig only) (6gp/1)\\}
	The arm has a built-in grappling hook. Functions as a Grappling Hook; You can launch a
	grappling hook up to 50 meters, where it will attach itself to most surfaces (including metal via
	magnets, wood, etc) and pull you up at a rate of 10m per phase.\\

	{\slshape Integrated Weapon (10+X gp/Variable) [X=Cost of Weapon to be Integrated]\\}
	You conceal a weapon inside your limb that can be activated and used at will.\\
	For a Hand/Forearm sized limb, this can be a firearm of Pistol, SMG or Shotgun size, or a melee
	weapon no longer than M-Length.\\
	For a Full Arm or larger limb, this can be a firearm of Pistol, SMG, Shotgun or Rifle size, or a
	Melee weapon no longer than L-Length.\\
	Pistols and S-Length melee weapons require 1 Augmentation Slot.\\
	SMGs, Shotguns and M-Length melee weapons require 2 Augmentation Slots.\\
	Rifles and L-Length melee weapons require 3 Augmentation Slots.\\
	The weapon can be readied normally, or Quickdrawn, but a Quickdraw with an Integrated
	Weapon automatically succeeds. Noticing that a limb has this augmentation by sight alone is an
	RS 4 Awareness check, RS 2 if you are allowed to handle the limb. Integrated Weapons have
	WT 0.\\

	Tactile Sense (2 gp/1)\\
	You can feel again. Remove the -4 dice penalty towards any action requiring/involving tactile
	sensations.\\

	Gyro-Arm (Full Arm only) (10 gp/1)\\
	Your arm is gyroscopically stabilized, making shooting and similar dextrous tasks easier. You
	have +2 dice to any missile rolls, and to any relevant skill tests.\\

	Folding Shield (Full Arm only) (20 gp/2)\\
	Your arm can deploy a small folding shield from itself, about the size of a buckler. It takes 1
	phase to do so completely. The folding shield doesn’t function quite the same as normal shield,
	providing no bonus to defense RS. Also, instead of covering a number of hit locations equal to
	an equivalent cover, it covers two hit locations, one being the Arm it is installed on, then another
	of the player’s choosing. The covered location can be changed phase by phase, but not while
	actions are taking place (you can’t change the shield to cover your head as someone is rolling
	to shoot you in the head). The shield has the following stats:\\

	% look for other solutions than making the table text fucking tiny
	\noindent{\centering\tiny
	\begin{tabular}{p{.8cm}lllp{1cm}p{1.3cm}ll}

	Shield Name & \multicolumn{2}{c}{TN} & AV & Coverage & Special & Wt. & Cost \\
	& Bash & Block \\ \hline \\
	Folding Shield & 7(+1b) & 7 & 15 & Two Locations & Bulletproof, Ablative[1] & 0 & 0 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\slshape Repetitive Task Function (Full Arm only) (4gp/1)\\}
	Your arm has an unusual function, in that it can be programmed to perform a specific, simple,
	one-handed task. This can be something like bringing a drink up to your mouth every half
	minute and holding it there for 5 seconds, writing out a specific section of a form, or signing your
	name. It takes about a minute to program the prosthesis, consisting of setting the clockwork
	mechanism up, then performing the task a single time. Once programmed, the prosthesis will
	perform the task exactly how it was performed during programming. The prosthesis will retain
	this programming until it is overridden by new programming. The method of activating the
	function is determined by the buyer, common methods being a pull string, a small button, or a
	turn key. A pair of full arm prosthetics can be programmed as one, allowing two-handed tasks.\\

\tabto{0}\newpage
\subsubsection{Leg Prosthetics}
A replacement of the leg. Prosthetic legs do not normally increase the strength or speed of the
wearer, as they must be light enough not to burden the wearer, and must also be mounted to
flesh, meaning that excessive force might rip them out of their sockets.\\
When getting a second prosthetic Leg, the Purity Cost for the new limb is Halved.\\

\noindent{\centering
\begin{tabular}{p{4cm}lp{1.7cm}}
Prosthetic 	& Monetary 	& Purity\\ 
			& Cost 		& Reduction \\ \hline \\
Prosthetic Foot/Shin 							& 8 gp 	& -1 	\\ \\
Prosthetic Full Leg 							& 15 gp & -2 	\\ \\
Prosthetic Full Leg \newline w/ Pelvic Reinforcement 	& 25 gp & -3 	\\ \\
Low Performance \newline (Civilian) 						& x1 	& - 	\\ \\
High Performance \newline (Military) 					& x3 	& - 	\\ \\
Extreme Performance \newline (Corporate) 				& x5 	& - 	\\
\\ \hline
\end{tabular}\\[\baselineskip] }
Low Performance Prosthetics have one Augmentation Slots, and 6 TOU.\\
High Performance Prosthetics have two Augmentation Slots, and 8 TOU.\\
Extreme Performance Prosthetics have three Augmentation Slots, and 10 TOU.\\

\paragraph{Leg Augmentations} (Cost of Augment/Augmentation Slots Used)\\

\tabto{1}
	{\slshape Armored (5gp/1)\\}
	Your prosthesis has +4 AV. This augmentation can be taken multiple times.\\

	{\slshape Combat Leg (5gp/1)\\}
	You’ve got a leg made of steel, and joints that rotate outside the human range of motion. You
	count as having sabatons for the purpose of kicking damage, and in addition, when using the
	Whip Kick rule, you may invest up to 10 CP to increase damage, rather than the usual 2.\\

	{\slshape Dagger Foot (5gp/1)\\}
	A dagger has been installed in the sole of your prosthetic foot, allowing it to be used in combat.\\
	Swinging Kicks can now deal piercing damage.\\

	{\slshape Mobility Enhancement (Two Leg Prosthetics only) (5gp/1)\\}
	These prosthetics have been optimized for movement of all kinds, from sprinting, to running, to
	jumping! This augmentation adds +2 to your MOB, and can be purchased multiple times.\\

	{\slshape Claw Stabilizers (Two Full Legs w/ Pelvic Reinforcement only) (8gp/1)\\}
	Stabilizing claws have been installed in your legs, letting you dig into the ground. Whenever
	appropriate, you may spend a move action to dig your claws into the ground. Doing so gives
	you +3 dice to stability tests (or any other appropriate test, GM discretion), and allows you to
	Brace a weapon.\\

	{\slshape Free Rotation (2gp/1)\\}
	This prosthesis’s joints have no limit on their rotation, allowing unusual feats of flexibility. Add +3
	dice to any test where this effect could come into play.\\

	{\slshape Rocket Legs (2 Full Legs w/ Pelvic Reinforcement only) (20gp/2)\\}
	A small rocket booster has been installed into both of your legs, allowing you to jump massive
	distances. Sadly, it only has enough fuel for one jump. Once per fuel tank (fuel tanks can be
	bought for 5gp), you may automatically pass one Jumping Skill test, within GM discretion.\\
	Jumping the gap between two close together voidships may be reasonable, jumping up into
	Overspace to engage an Airship is not. The soft limit of distance is 33m horizontal, and about
	17m straight vertical.\\

	{\slshape Leg Compartment (2gp/1)\\}
	A small hidden compartment has been installed somewhere on your prosthesis. This could be a
	holster of some kind, or just a small box for storing miscellaneous items. Noticing that a limb has
	this augmentation by sight alone is an RS 5 Awareness check, RS 2 if you are allowed to
	handle the limb.\\

\tabto{0}\newpage
\subsection{Other Prosthetics}

\subsubsection{Head Prosthetics}
A prosthesis on or of the skull. Generally the least common of replacements, but covering the
widest range of injuries. One man may, after taking a blast of shrapnel to the face, may require
a Prosthetic Jaw, Prosthetic Eye, or even a full Prosthetic Facial Reconstruction.\\
By default, any prosthesis on your face counts as a Facial Deformity (2).\\
Prosthetic Eyes, by default, cannot see color, and suffer a -2 dice penalty to any PER check.
When getting a second Prosthetic Eye, the Purity Cost for the new eye is Halved.\\

\noindent{\centering
\begin{tabular}{p{4cm}lp{1.7cm}}
Prosthetic 	& Monetary 	& Purity\\ 
			& Cost 		& Reduction \\ \hline \\
Prosthetic Jaw 						& 2 gp 	& -1 	\\ \\
Prosthetic Eye 						& 10 gp & -2 	\\ \\
Prosthetic Facial \newline Reconstruction 	& 30 gp & -4 	\\ \\
Low Performance \newline (Civilian) 			& x1 	& - 	\\ \\
High Performance \newline (Military) 		& x3 	& - 	\\ \\
Extreme Performance \newline (Corporate) 		& x5 	& - 	\\
\\ \hline
\end{tabular}\\[\baselineskip] }
Low Performance Prosthetics have one Augmentation Slots.\\
High Performance Prosthetics have two Augmentation Slots.\\
Extreme Performance Prosthetics have three Augmentation Slots.\\

\paragraph{Head Augmentations} (Cost of Augment/Augmentation Slots Used)\\

\tabto{1}
	{\slshape Combat Jaw (Prosthetic Jaw) (2gp/1)\\}
	Your teeth and jaw have been reinforced, and potentially even sharpened! Your Jaw now
	increases the AV of the Lower Head by 6. Additionally, Bite Maneuvers now deal [STR+BS]
	rather than [Half STR+Half BS].\\

	{\slshape Technicolor Vision (Prosthetic Eye) (5gp/1)\\}
	You can see in color once more! It’s not quite up to par with normal human vision, but it’s
	enough. Remove the dice penalty related to prosthetic eyes not being able to color.\\

	{\slshape Starlight Vision (Prosthetic Eye) (8gp/1)\\}
	Your eye can now take in extra light, allowing you to see better in the dark. When activated,
	remove any penalty related to darkness affecting your vision (shooting in dark ship, looking for
	something in the dark, etc.). If suddenly exposed to light while this augmentation is active, suffer
	4 stun immediately.\\

	{\slshape Facial Molding (Prosthetic Facial Reconstruction only) (20gp/2)\\}
	An unusual enhancement, this augmentation allows a facial prosthetic to reshape itself,
	potentially changing your face completely. After starting the process, this augmentation will
	slowly change your face to whatever form (still human face) you wish, over the course of 1
	week.\\

	{\slshape Voidsight (Prosthetic Eye) (20gp/2)\\}
	Your eye can now project Moonstone radiation, which can reflect back into special sensors,
	allowing you to see things that normal light could not convey to you. Additionally, Moonstone
	radiation penetrates most materials, allowing you to briefly see through walls or other obstacles.
	Voidsight is powered by the moonstone battery used to project the radiation, but there is
	another cost--something the Zells call “The Wheel’s Price.” Events tend to conspire against
	those who utilize the Voidsight for too long, as though they are being punished for violating
	some natural law.\\

	Activating Voidsight requires the expenditure of 1 Luck point. This keeps it active for about thirty
	seconds (5 Rounds in Combat), during which time you can see things normally invisible to the
	naked eye (ghosts, magical spirits, the trails left by Void Users, etc) and can also see living
	creatures through walls and other obstacles.\\
	If shooting at an opponent or other target hiding behind cover that your weapon can penetrate,
	you may completely negate their Cover Defense bonus when shooting at them, since you can
	see them clearly and their cover provides no meaningful deterrence to the weapon.\\

\tabto{0}\newpage
\paragraph{Torso Augmentations} (Cost of Augment/Purity Loss)\\
Augmentations of the Torso work slightly differently from other prosthetics. They have no
augmentation slots, just being a flat cost and purity loss for a specific effect.\\

\tabto{1}
	{\slshape Major (Organ Replacement) (15gp/-3 Purity)\\}
	Has no specific positive effects. Organ Replacement is done to prevent death from organ failure
	on account of illness, surgery or injury.\\
	When a character would be about to perish due to Septic Infection (see Wounds) and assuming
	that they are in a medical facility capable of performing such a surgery, an Organ Replacement
	Surgery may be used to immediately negate the Sepsis. Health is returned to normal, but
	another Infection roll must be made (counts as sterile and in good conditions, cutting type
	wound level 3) to see if the organ is rejected. No Cybernetic Rejection roll is need for this
	augmentation.\\

	{\slshape Nerve Replacement (20gp/-4 Purity per Level)\\}
	The replacement of the central nervous system (where possible) with Whalebone fibers. This
	increases reaction time, coordination, reflexes and instills in the subject an extreme clarity of
	thought in stressful situations. However, it has its drawbacks. Difficulty sleeping is not
	uncommon, as the nerves react powerfully to stimuli, even in dreams. A normal man might sit up
	suddenly in bed after a nightmare-- a man with whalebone nerves might jump up so hard he hits
	the ceiling.\\
	Each level of Nerve Replacement increases ADR by 1. However, in the event of any surprise,
	alarm or other “trigger” for immediate reaction, a character must make a Willpower Test at RS
	2+1 per level of this Augmentation to avoid reacting in the most immediate, primal and stupid
	way possible.\\

	{\slshape Hemostatic Clamp (10gp/-2 Purity)\\}
	A small augmentation situated near the heart, this unit helps minimize blood-loss in combat
	conditions. Reduce the RS of Blood Loss tests by 1.\\

\tabto{0}
\subsection{Prothesis Installation \& Cybernetic Rejection}
When installing a prosthesis, the Surgeon responsible for its installation must make an RS 4
Surgery test. This surgery takes about 4 hours of continuous work under carefully controlled
conditions. Meanwhile, the person undergoing surgery must make a WIL+Grit test, at RS equal
to the Purity Reduction of the Prosthetic, in order to stay conscious while the surgery is taking
place, so connecting nerves can be properly tested. If the person fails to stay conscious, the
surgery cannot be finished until they wake up and stay awake (repeat the test). Multiple failures,
or an unreachable RS, could mean that the surgery is simply too invasive for the person to
safely endure.\\

After a prosthesis is installed, a Purity test must be made by the person receiving the
prosthesis. Failure risks Cybernetic Rejection. The Purity test is RS 3. Success means the
prosthesis has been successfully installed, and no ill effects other than purity loss is suffered.
On failure, various ill effects are suffered, based on the degree of failure, see the table below.
Prosthetics can be removed, restoring the old bane they negated, and removing any negative
effects of rejection. However, Purity Reductions are not reversed, and are permanent. You can
never be clean again. Removing a prosthesis is an RS 4 Surgery Test.\\

\noindent{\centering
\begin{tabular}{p{3cm}p{5cm}}
Degrees of Failure & Effect \\ \hline \\
Failed by 1 & Minor chronic pain is suffered. Gain Lasting Pain (Minor). \\ \\
Failed by 2 & Major chronic pain is suffered. Gain Lasting Pain (Major). \\ \\
Failed by 3 & Major chronic pain is suffered. In addition, lesions and open wounds develop around the socket. Gain Lasting Pain (Major). The location the prosthesis was installed on has a constant Infection Risk of 1. This risk is reduced as normal by things like having enough to eat, and being in a clean environment. \\ \\
Critically Failed (or 0 Purity) & Total Cybernetic Rejection. An immediate full-body septic infection is suffered, persisting until the prosthesis is removed, or until the subject is placed on a regimen of immunosuppressants drugs. These drugs cost 4gp per month if paying out of pocket, and reduce a character’s HLT to 0 for the purposes of infection and sickness. \\
\\ \hline
\end{tabular}\\[\baselineskip] }
Post installation/removal recovery time is equal to that of a Level 4 wound.\\

\chapter{The Ship}
So you’re a bunch of Void Pirates. You’d better have a Voidship! Each PC brings a certain
number of Ship Points to the Party. Pool these together and use them to create a ship for the
Party. This should be a collaborative process, though the GM could just do it himself if you can’t
agree, or if he has something special in mind.\\

\noindent{\centering
\begin{tabular}{p{3cm}p{5cm}}
PCP  & Ship Contribution \\ \hline
1 	& 0 	\\
2 	& 1 	\\
3 	& 2 	\\
4 	& 3 	\\
5 	& 4 	\\
6 	& 5 	\\
7 	& 6 	\\
8 	& 10 	\\
9	& 15	\\
10	& 20	\\
\\ \hline
\end{tabular}\\[2\baselineskip] }

Other Sources of Ship Points\\

\noindent{\centering
\begin{tabular}{p{5cm}p{3cm}}
Source & Points \\ \hline \\
50-Year Mortgage & (7,500 gp debt per Point) \\ \\
State Navy (All characters gain Duty Bane for the country they serve. Captain gains Rank 6.) & 5 per character \\ \\
Salvaged (50\% of ship systems are damaged at start) & +25\% Points\\
\\ \hline
\end{tabular}\\[2\baselineskip] }

Ship Type\\

\noindent{\centering
\begin{tabular}{ll}
Type & Point Cost \\ \hline \\
Voidship & 0 \\
Zeppelin & 5 \\
Underspace Boat & 10 \\
\\ \hline
\end{tabular}\\[2\baselineskip] }

{\bfseries Ship Rooms\\}
Corridor\\
Crew Quarters\\
Storage\\
Magazine\\
Fuel Storage\\
Bridge\\
Armored Bulkhead\\
Mess Hall\\
Lounge/Entertainment Area\\
Armory\\
Officer Quarters\\
Engine Supercharger\\
Science Lab\\
Containment Cells\\
Medbay\\
Damage Control Center\\
Gun Pod\\

{\bfseries Shields\\}
Ship shields protect the\\

\subsection{Ship Combat}
Ship Combat is divided into 1-minute Rounds.\\
The Captains of each ship write down the actions of their vessels simultaneously in secret, and
reveal them all. The actions are then resolved in the following order:\\

\noindent
1: Orders\\
2: 1st Half Movement\\
3: Attack Actions\\
A ship can make one Attack Action a Round per Fire Control Seat or Targeting Computer it has.\\

\tabto{1}
{\slshape Barrage}: Choose a target and a weapon type on your ship. All weapons of that type that
are capable of doing so fire at the target. Fire Control Officer rolls Gunnery+Gunner
Quality+Number of Weapons Firing+Misc against enemy Base Defense+Angle
Bonus/Penalty+Range Modifier+Misc. Success results in 1 shot hitting, +1 per BS.\\

\noindent{\itshape
Broadside\\
Launch Torpedoes\\
Launch Fighters\\
Main Battery, Fire!\\
}

\tabto{0}
\noindent{\centering
\begin{tabular}{ll}
Angle & Bonus/Penalty to Defense \\ \hline \\
Approaching & 0 \\
Abeam & +1 \\
Fleeing & -1 \\
\\ \hline
\end{tabular}\\[2\baselineskip] }

\noindent
4: Damage Resolution\\

\tabto{1}
{\itshape
Going back to the prior example, the enemy destroyer has just been hit with 9 Light Gun
shots. Hit Location is rolled for each weapon based on Firing Angle (Abeam.) 7 of the
attacks hit the Left Flank, and 2 hit the Bow. 7d6 damage is rolled against the Left Flank
for a total of 20, 2d6 are rolled against the Bow for a total of 6.\\
The Left Flank has sustained 20 damage, which is twice-over 20\% of the total Hull value
of that Flank (50), meaning two Critical Hits have been inflicted.
The Bow has sustained 6 damage.\\
The Critical Hits are rolled. One Fire Control Seat is destroyed, and the Damage Control
Room is hit, casualties are suffered and the system is down. This is good! It will make it
hard for the enemy to recover from future attacks.\\
}

\tabto{0}
\noindent
5: 2nd Half Movement\\
6: Damage Control\\

{\bfseries Torpedoes\\}
When fired, Torpedoes will advance towards their targets in one of three ways depending on
their guidance system. As soon as a Torpedo makes contact with a target vessel, resolve
Anti-Torpedo measures (if any) and then resolve the torpedo attack with the remaining
torpedoes, rolling for random hit location and then damage like any other attack.\\

Detecting Torpedoes requires an RS 5 Sensors test for regular torpedoes, and an RS 8 Sensors
test for stealth torpedoes.\\

Torpedoes can be shot with attacks. A wave of torpedoes has Defense 5 against weapons
without the Anti Aircraft quality, and Defense 2 against weapons with that quality. Any impact
will destroy a torpedo. A Fighter Squadron of any sort will completely destroy any Torpedo
Flight.\\

{\bfseries Dumbfire}: Dumbfire Torpedoes will simply proceed in a straight line towards their destination
until they either reach the target, encounter another potential target, or are destroyed. Dumbfire
torpedoes that pass within 1km of a potential target will change targets to that ship instead. A
Gunnery Test and a Sensors Test, both at RS 4, will determine where the torpedoes would have
to be launched to hit a target based on its current trajectory. Unless the target changes speed or
turns, the torpedoes will hit automatically.\\

{\bfseries Guided}: Guided Torpedoes will adjust their trajectory at the start of each Movement phase, and
then proceed in a straight line towards the target. Guided Torpedoes will avoid or move around
obstacles or other vessels in pursuit of their target.\\

{\bfseries Assault}: Assault Torpedoes are a bit different, in that they’re really man-piloted vessels full of
marines intent on doing grievous harm to the enemy. Assault Torpedoes can be controlled
manually each Movement Phase, and can be instructed to go wherever. Once they hit a target,
they can conduct a {\bfseries Boarding Action} with Manpower 1 per Assault Torpedo in the volley.\\

{\bfseries Fighters\\}
Much like Torpedoes, Fighters are launched during the Shooting phase, but move during the
Movement phases. Unlike Torpedoes, Fighters can be ordered to abandon their old target and
adopt a new one, or to hold their position, or return to their host-ship.\\

When Fighters make contact with other Fighters, an Air Superiority test is made between the
two squadrons. If one squadron is more numerous by 50\% or more, it gains a +1 bonus to its Air
Superiority quality per 50\% advantage it has.\\
The victor of the contest continues as normal. The loser is scattered and returns to their carrier,
unable to continue their mission.\\

Fighters can be shot at by ship weapons. A squad of fighters has Defense 5 against weapons
without the Anti Aircraft quality, and Defense 2 against weapons with that quality. Any hit will
destroy a fighter.\\

When Fighters contact an enemy Vessel, first resolve the Antiaircraft fire from the ship. Every
weapon with the Anti Aircraft quality shoots its ROF at the Squadron. Any hit will destroy a
fighter. Every surviving fighter may then inflict damage equal to their Bomber qualities.
Afterwards, the fighter squadron will have to retreat back to the mothership to reload on bombs.\\

\twocolumn[{
	\setlength{\parindent}{10.00002pt}

	{\bfseries Ship Class\\}

	\noindent{\centering
	\begin{tabular}{p{2.5cm}lp{6.5cm}llll}
	Class & Lengths/ & Normal & Hull & Base & Point & Monetary \\
		  & Squares & Armament & & Defense & Cost & Cost  \\\hline \\
	Tiny (Cutter) 			& 10m/12 	& 1 gun 											& 10 	& 3 	& 5 	& 25,000 gp 	\\ \\
	Small (Sloop) 			& 25m/30 	& 4 guns 											& 50 	& 2 	& 10 	& 50,000 gp 	\\ \\
	Medium (Frigate) 		& 50m/96 	& 2 batteries, 4 guns 								& 100 	& 1 	& 20 	& 100,000 gp 	\\ \\
	Large (Destroyer) 		& 100m/144 	& 4 batteries, 8 guns, 2 heavy guns 				& 200 	& 0 	& 30 	& 150,000 gp 	\\ \\
	V. Large (Battleship) 	& 150m/270 	& 6 batteries, 12 guns, 4 heavy guns, 1 spinal gun 	& 400 	& -1 	& 50 	& 250,000 gp 	\\ \\
	Huge (Dreadnought) 		& 250m/360 	& 8 batteries, 16 guns, 6 heavy guns, 1 spinal gun 	& 800 	& -2 	& 200 	& 1,000,000 gp 	\\ \\
	Leviathanic (Leviathan) & 300m/480 	& 10 batteries, 20 guns, 8 heavy guns, 1 spinal gun & 1600 	& -3 	& 240 	& 1,200,000 gp 	\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\bfseries Propulsion\\}

	\noindent{\centering
	\begin{tabular}{llll}
	Type & Speed & Squares & Point Cost \\ \hline \\
	Whaleblood Engine 	& 50 knots & 1/2/4/6/8/10/15 	& 2\\
	Solar Sails 		& 30 knots & 0/1/2/3/4/5/6 		& 2\\
	Manual Screws\dag 	& 10 knots & 0/1/2/3/4/5/6 		& 1\\
	\\ \hline
	\end{tabular}\\[\baselineskip] }
	\dag Can be taken in addition to other options as an emergency propulsion system\\

	Ships have a “base” number of guns which are free, and take up no space. Additional guns take
	up space as shown below.\\

	{\bfseries Light Guns} (All Light Guns take up one square)\\

	\noindent{\centering
	\begin{tabular}{lllllp{5cm}}
	Gun Type & Range & Attack & Hull & Arc & Special \\ 
			&		&	TN 		& Damage \\ \hline \\
	Light Gun 			& 1000m & 7 & 1d10 		& 180 degrees & Blast Radius 10m 												\\
	Light Rocket 		& 5000m & 9 & 3d10 		& 180 degrees & Blast Radius 25m 												\\
	Flamethr ower 		& 100m 	& 4 & NA 		& 180 degrees & Fire damage 1d10/2 for 5 rounds 								\\
	Harpoon Launcher 	& 500m 	& 7 & 1d10/2 	& 180 degrees & Boarding Action 												\\
	Flak Gun 			& 5000m & 6 & 1 		& 360 degrees & ROF 4, Anti-Aircraft, Blast Radius 10m, Half Damage vs Ships 	\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\bfseries Heavy Guns} (All Heavy Guns take up 2 Squares)\\

	\noindent{\centering
	\begin{tabular}{lllllp{5cm}}
	Gun Type & Range & Attack & Hull & Arc & Special \\ 
			&		&	TN 		& Damage \\ \hline \\
	Heavy Gun 				& 3000m 	& 7 & 3d10 	& 180 degrees & Blast Radius 25m 												\\
	Heavy Rocket Launcher 	& 10,000m 	& 9 & 10d10 & 180 degrees & Blast Radius 50m 												\\
	Anti-Aircraft Array 	& 5,000m 	& 6 & 1 	& 360 degrees & ROF 16, Anti-Aircraft, Blast Radius 10m, Half Damage vs Ships 	\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }
}]
\twocolumn[{
	\setlength{\parindent}{10.00002pt}
	
	{\bfseries Batteries} (Batteries take up 10 Squares, these must all be on the edge of the ship.)\\

	\noindent{\centering
	\begin{tabular}{lllllp{5cm}}
	Gun Type & Range & Attack & Hull & Arc & Special \\ 
			&		&	TN 		& Damage \\ \hline \\
	Broadside Array 	& 1000m 		& 6 			& 1d10 			& 90 degrees & ROF 20 								\\
	Torpedo Array 		& See Torpedoes & See Torpedoes & See Torpedoes & 90 degrees & 6 Tubes, Launches Torpedoes 			\\
	Shield Generator 	&  				&  				&  				& 90 degrees & Not a Weapon, see Shield Array entry \\
	Aircraft Catapults 	&  				&  				&  				& 90 degrees & Carries Aircraft (10/20/30) 			\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\bfseries Spinal Cannon\\}
	Spinal Cannons take up a single line of squares that extends the length of the vessel. The
	number of squares this actually takes up will depend on the design of the vessel, but should be
	counted and listed as X for stat purposes as shown below.\\

		\noindent{\centering
	\begin{tabular}{lllllp{4.4cm}}
	Gun Type & Range & Attack & Hull & Arc & Special \\ 
			&		&	TN 		& Damage \\ \hline \\
	Moonstone Cannon 		& 15,000m+[X*100m] 	& 7	& 3d10*10+X 	& Dead Ahead &  							\\
	Annihilator Lance 		& 500m+[X*10m] 		& 6	& 10d10*10+X 	& Dead Ahead &  							\\
	Venator Cannon 			& 12,000m+[X*100m] 	& 7	& 1d10*10+X 	& Dead Ahead & ROF 5 						\\
	Rapture Artillery Array & 20,000m+[X*100m] 	& 7	& 1d10 			& 90 degrees & ROF 30, Blast Radius 10 		\\
	Airstrip 				&  					& 	&  				& Dead Ahead & Carries Aircraft (30/60/90) 	\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\bfseries Torpedoes\\}

	\noindent{\centering
	\begin{tabular}{llllll}
	Torpedo & Speed & Maximum & Damage & Special & Cost \\
	Type & 			& Range \\ \hline \\
	Dumbfire Anti-Ship Torpedo 	& 4 kpm 	& 10km 	& 1d10x100 	& 			&		\\
	Guided Anti-Ship Torpedo 	& 2 kpm 	& 5km 	& 1d10x100 	& 			&		\\
	Dumbfire Stealth Torpedo 	& 1 kpm 	& 5 km 	& 1d10x50 	& 			&		\\
	Guided Stealth Torpedo 		& .5 kpm 	& 2 km 	& 1d10x50 	& 			&		\\
	Assault Torpedo 			& 4 kpm 	& 50 km & NA 		& 			&		\\
	Stealth  Assault Torpedo 	& 2 kpm 	& 25 km & NA 		& 			&		\\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }

	{\bfseries Fighters\\}\hrule

	\noindent{\centering
	\begin{tabular}{p{5cm}llp{4.4cm}p{2.4cm}l}
	Fighter & Speed & Max & Armament & Special & Cost \\
	Type 	& 		& Range \\ \hline \\
	Sternfleischer-109 							& 12 kpm 	& 250 km 	& \scriptsize 2 13mm Kurschat-Lickert MG71 Machineguns, \newline 1 20mm cannon, 1x250 kg Bomb 	& Air Superiority 6, Bomber [1] 	& 1200 gp 	\\
	Schneider-Tannenhauser-702 					& 10 kpm 	& 250 km 	& \scriptsize 4 13mm Kurschat-Lickert MG71 Machineguns, \newline 2 20mm cannons 					& Air Superiority 7 				& 1100 gp 	\\
	Sternfleischer-87 							& 8 kpm 	& 200 km 	& \scriptsize 5 13mm Kurschat-Lickert MG71 Machineguns, \newline 4x400kg bombs 					& Air Superiority 3, Bomber [4] 	& 3000 gp 	\\
	UC “Can’t we all just get along?” Mk VII 	& 12 kpm 	& 100 km 	& \scriptsize 6 20mm cannons 															& Air Superiority 5 				& 1000 gp 	\\
	Merlin Firebrand Mk XIII 					& 10 kpm 	& 150 km 	& \scriptsize 8 12x99mm machineguns, \newline 1 37mm cannon 										& Air Superiority 6 				& 1100 gp 	\\
	D10 Stallion 								& 10 kpm 	& 500km 	& \scriptsize 8 12x99mm machineguns 													& Air Superiority 6 				& 1200 gp 	\\
	CIS “Eadwine”Bomber 						& 10 kpm 	& 500 km 	& \scriptsize 4 .303 machineguns, \newline 10x1000lb bombs 										& Air Superiority 1, Bomber [10] 	& 10,000 gp \\
	\\ \hline
	\end{tabular}\\[2\baselineskip] }
}]

\chapter{Void Magic}
{\itshape“Can you hear it, human? Like a mother’s lullaby. Like the song of a whale. The winds of the
netherworld whisper through darkness. The secrets, the secrets!”}

Void Magic, also called Black Magic, is a form of spellcasting granted to certain individuals by
the “Gods Below,” the beings of Underspace, or the Netherworld.\\
Nobody is certain what causes an individual to be chosen. Some search their entire lives for the
power without ever finding it, others have it dropped into their laps without even knowing what it
is. The dark powers are unpredictable, and capricious--however, they do not revoke their gifts
once given.\\

Of the many strange abilities granted to Void-Users, the most universal is that they can now
understand the songs of whales. The eerie wailing that permeates the Void is now clear to
them. The whales sing very slowly--a sentence in a day, perhaps. They are lonely, so very
lonely, and call out to their fellows from across the darkness.\\

At Character Creation, a PC with points in Magic may become a Void-User by spending the
points awarded as detailed in the chart below.\\

If a character has prosthetics of any kind, apply any Purity Reductions directly to Void Capacity.\\

\noindent{\centering
\begin{tabular}{lp{2cm}ll}
Magic & User & Void Powers & Void \\
Points & Level & /levels 1/2/3) & Capacity \\ \hline \\
5 	& Novice 				& 1/0/0 & 3+Wip 	\\
15 	& Initiate 				& 3/1/0 & 6+Wip 	\\
25 	& Adept 				& 6/3/1 & 9+Wip 	\\
35 	& Master of \newline the Temple 	& 8/6/3 & 12+Wip 	\\
\\ \hline
\end{tabular}\\ }
You may forfeit a level two or level three power gained from your user level to gain two level one
powers. You may forfeit a level 3 power to gain two level 2 powers.\\

{\bfseries Leveling up Void Capacity costs 8 Arc per VC point.\\}

{\bfseries Learning new Void Powers\\}
(Powers with multiple levels must be bought in order)\\

\noindent{\centering
\begin{tabular}{ll}
Power Level & Arc Cost \\ \hline \\
Level 1 & 8 \\
Level 2 & 12 \\
Level 3 & 16 \\
\\ \hline
\end{tabular}\\[2\baselineskip] }

% maybe put this stuff on a twocolumn page
{\bfseries Powers\\}
Powers can be used with Void. You regain Void equal to (and up to) Void Capacity at the
beginning of every Round. Offensive Void powers halve the Defense RS to hit their intended
target.\\

{\bfseries\slshape Using Void Powers as an Action\\}

{\bfseries\large Translocate\\}
You can dissolve yourself into ether, and reappear elsewhere at will! This is essentially
teleportation. Many find this experience disturbing, as it takes a second between dissolving for
them to reappear, and all that anyone can remember of that second is a lurching sense of
weightlessness and a great bloodshot eye staring out of a field of eternal darkness.\\
But hey, it’s fast, right?\\

	\tabto{1}
	{\bfseries Level 1}: Flash [5]\\
	As a Movement, you may instantly move up to 30m to any location that you can see
	directly, ignoring all of the terrain in between. You can bring equipment with you up to
	your own bodyweight. You may spend another 10 Void to bring an additional person with
	you, or up to an additional person’s weight worth of inanimate materials.\\
	For the purposes of shooting, all penalties double when Translocating. Enemies suffer a
	-4 penalty when shooting at you, and you suffer a -8 penalty when shooting. This penalty
	applies to Offensive Void powers as well. You may make melee attacks after Flashing
	adjacent to someone. This “counts” as a Charge, but you suffer a -10 penalty to CP from
	Flashing. In Melee Combat, you may activate this Power as a Flee Maneuver, at -1 TN.\\

		\tabto{2}
		{\bfseries Level 2}: Shift [8]\\
		As a Movement, you may instantly move up to 60m to any location, regardless of
		whether or not you can actually see it directly. You have to know that the location
		is there, though. If for some reason the location you had anticipated is no longer
		there, the Shift fails, and you waste your Action. You may spend another 10 Void
		to bring an additional person with you, or up to an additional person’s weight
		worth of inanimate materials. e penalties to attacks \& actions are the same as
		with Flash, with the exception of the penalty to melee attacks, which is is reduced
		to -6. In Melee Combat, you may activate this Power as a Flee Maneuver, at -1
		TN.\\

			\tabto{3}
			{\bfseries Level 3}: Translocate [15]\\
			As a Magic Action, you may instantly move up to 120m to any location,
			regardless of whether or not you can actually see it directly. You may also
			instantly move any distance to a Location where you have a Tefillin.\\
			A Tefillin is a piece of yourself (usually a vial of blood, some hair, or a bit
			of fingernails) suspended in a container of Underspace Ether.\\
			Constructing such a reliquary costs 1 gp. However, Tefillin can be broken
			easily, so it is best to only install them in safe locations or among friends.
			You may spend another 10 Void to bring an additional person with you, or
			up to an additional person’s weight worth of inanimate materials.\\
			The penalties to attacks \& actions are the same as with Shift, with the
			exception of the penalty to melee attacks, which is is reduced to -4. In
			Melee Combat, you may activate this Power as a Flee Maneuver, at -1
			TN.\\

\tabto{0}
{\bfseries\large Spiderwalk [5 Void]\\}
You may walk on walls or the ceiling with no more difficulty than if you were walking on the floor.
Curiously, your clothing and other equipment do not seem to obey the laws of gravity as you do
this--they “hang” in whichever direction the surface you are clinging to is relative to you. If you
are clinging to multiple surfaces opposed to each other, they seem to float as though
weightless.\\
You may activate this Power at any time by paying the cost.\\

	\tabto{1}
	{\bfseries Level 1}: 1 Round\\

		\tabto{2}
		{\bfseries Level 2}: 1 Minute (10 Rounds)\\

			\tabto{3}
			{\bfseries Level 3}: 1 Hour (600 Rounds)\\

\tabto{0}
{\bfseries\large Influence Technology [5 Void]\\}
You can cause technological devices (those involving electricity, motors, etc) to obey your will to
a certain extent. At low levels you can simply cause them to fail, but at higher levels you can
control them as extensions of your body. Make a Willpower Test at RS 1 for simple devices, 3
for complex devices, and 5 for extremely complex devices (GM’s discretion, but as a rule, a car
battery or a gun is simple, an electric motor or a diesel engine is complex, and a computer is
extremely complex.)\\
Purely mechanical devices can also be manipulated (guns, cranks, a hand-powered rail car) at
+1 RS.\\

	\tabto{1}
	{\bfseries Level 1}: Short-Circuit\\
	The device fails at whatever its intended task was, or activates once at your command.\\

		\tabto{2}
		{\bfseries Level 2}: Destroy\\
		The device not only fails, you may choose to break it as well. Reduce RS by 1.\\

			\tabto{3}
			{\bfseries Level 3}: Control\\
			You may cause the device to perform one action at your bidding within its
			capabilities (cause a crane to turn left instead of right, cause a weapon to
			fire in a specific way on your timing and command, etc) Reduce RS by 2.\\
			If the device would require electricity or manual input to function, you may
			force it to function as though something were supplying it with the
			necessary force, so long as you concentrate and take no other action.\\

\tabto{0}
{\bfseries\large Field of Darkness [10 Void]\\}
You create a field of darkness around you that reduces visibility to Darkness in its area. This
darkness cannot be penetrated by conventional lighting, only by Moonstone Lamps. It
overwhelms all natural and artificial lighting sources, except the aforementioned. Field of
Darkness lasts for the entire round it was activated. At the start of the next round, before
Phases begin, Field of Darkness can activated again for the same cost.\\

	\tabto{1}
	{\bfseries Level 1}: The field is up to 10 meters in all directions from you.\\

		\tabto{2}
		{\bfseries Level 2} : The field is up to 10 meters in all directions from you, and your visibility
		remains as though it were normal lighting, regardless of what it was before.\\

\tabto{0}
{\bfseries\large Winds of the Netherworld\\}
You can cause tendrils of dark energy to lash out at your foes, dissolving their flesh into
subspace dust. If used against the corpses of already dead creatures no larger than a horse or
cow, the bodies and all other evidence disintegrate automatically.\\

	\tabto{1}
	{\bfseries Level 1}: Vortex [3+X]\\
	As a Shoot Action you can make an attack at TN 6 with X that ignores Movement
	penalties to attack. This is only one attack with X. If it hits it ignores AV and inflicts
	WIP+BS damage. Effective Range is no further than 30m. In melee combat, this may be
	performed as a Melee Shoot maneuver, with the X CP of Melee Shoot not exceeding the
	X of Void Capacity invested in Vortex.\\

		\tabto{2}
		{\bfseries Level 2}: Infest [5+X]\\
		As a Shoot Action you can make an attack at TN 6 with X that ignores Movement
		penalties to attack. This is only one attack with X. If it hits it ignores AV and
		inflicts WIP+BS damage. Next Phase, the target suffers the same amount of
		Damage to a different randomly-chosen Hit Location. Effective Range is no
		further than 30m.\\

			\tabto{3}
			{\bfseries Level 3}: Dust to Dust [8+X+Y]\\
			As a Shoot Action you can make an attack at TN 6 with X that ignores
			Movement penalties to attack. This is Y attacks with X. These attacks
			ignore AV and inflicts WIP+BS damage. A Critical Hit instantly
			disintegrates any target. Effective Range is no further than 30m.\\

\tabto{0}
{\bfseries\large Void-Push\\}

	\tabto{1}
	{\bfseries Level 1}:  Minor [X]\\
	As a Shoot Action you can make an attack at TN 7 with X that ignores Movement and
	Cover penalties. If successful, the enemy is pushed back BS meters directly away from you, and
	must make a Stability Test at RS equal to your BS or be knocked prone. Effective Range is no
	further than 10m. In melee combat, this may be performed as a Melee Shoot maneuver, with
	the X CP of Melee Shoot not exceeding the X of Void Capacity invested in Voidpush.\\

		\tabto{2}
		{\bfseries Level 2}: Major [X+4]\\
		As a Shoot Action you can make an attack with X that ignores Movement and
		Cover penalties. If successful, the enemy is pushed back 2xBS meters directly away
		from you, and must make a Stability Test at RS equal to your BS or be knocked prone.
		Effective Range is no further than 10m. In melee combat, this may be performed as a
		Melee Shoot maneuver, with the X CP of Melee Shoot not exceeding the X of Void
		Capacity invested in Voidpush.\\

			\tabto{3}
			{\bfseries Level 3}:  Superior [X+6]\\
			As a Shoot Action you can make an attack with X that ignores Movement
			and Cover penalties. If successful, the enemy suffers a Bludgeoning Attack to the
			Chest: Non-Vitals at Strength equal to BS, is pushed back 2xBS meters directly
			away from you, and must make a Stability Test at RS equal to your BS or be
			knocked prone. Effective Range is no further than 10m. In melee combat, this
			may be performed as a Melee Shoot maneuver, with the X CP of Melee Shoot
			not exceeding the X of Void Capacity invested in Voidpush.\\

\tabto{0}
{\bfseries\large Freeze Time\\}

			\tabto{3}
			{\bfseries Level 3}: The World is Mine [15]\\
			Activate at the very start of a Round. For this Round, there is now a 0th Phase which you
			may act in before any other Phases begin. You also gain +4 to your Base Defense for
			the 0th and 1st Phase of this Round. If multiple characters use this power
			simultaneously, they may all act in the “0th Phase.”\\
			Outside of Combat, this ability allows you to freeze time for just over one full second.
			Objects you put into motion will move, but only sluggishly until time returns to its regular
			pace.\\

\tabto{0}
{\bfseries\large Dustform\\}

	\tabto{2}
	{\bfseries Level 2}:  Immaterial [10]\\
	Activate at the start of a Round. You gain +4 to your Base Defense for the entire Round.
	Moonstone weapons can still attack you normally, and you gain no bonus to your Base Defense
	against them. Attacks you make this Round only inflict half of their normal damage, be they
	missile, melee or void-magic, as you are not entirely “there,” and neither are your bullets.

		\tabto{3}
		{\bfseries Level 3}:  Unreal [15]\\
		Activate at the start of a Round. You gain a +8 to your Base Defense for
		the entire Round., but you cannot make any attacks or interactions with other
		characters or objects. You can move through solid objects with ease, so long as
		they are not Electrified or Moonstone-Laced. Moonstone weapons can still attack
		you normally, and you gain no bonus to your Base Defense against them.\\

\tabto{0}
{\bfseries\large Bend Mind\\}

		\tabto{2}
		{\bfseries Level 2}: Kneel! [X+10]\\
		You may activate this ability as an action in Combat. When activating this ability, you
		may issue a command, generally no longer than two or three words, directed at a living being.
		That being must make a Willpower test at RS equal to 2+(X/2) or do its best to obey the
		command. If the command is dangerous or potentially suicidal, the target gains a bonus of +3 to
		its Willpower test. If the command is overtly suicidal (“Die”is a valid command, causing the
		target’s heart to stop) the target doubles its Willpower for the purposes of the test. Kneel! is only
		binding for about ten seconds (3 Rounds in combat) after which the effects wear off. Multiple
		targets may be selected for Kneel! At a cost of +2 Void Capacity per target.\\

\tabto{0}
Hex\\
You can bring misfortune on those unlucky enough to earn your ire. Their children will be born
invalid, their wives will turn barren, and moths will eat their favorite coat. Well, perhaps not that
extreme. For a short while, anything that can go wrong for your target, will.\\

	\tabto{1}
	{\bfseries Level 1}:  Jinx [2+X] You may activate this ability as an action in Combat, targeting one
	character. Choose 1 location/number on the Missile Hit Location table. This location is treated
	as a Critical Hit. This is in effect for Phases equal to [X]. Additionally, Full Auto Attacks can now
	inflict Critical Hits, as can Explosives and Shotgun Scatter. You still cannot aim for Critical Hits,
	and aiming for a Hexed location is treated as if aiming for the original location. This applies to
	any attack targeting the character for the phases. Additionally, incidental bad luck will befall
	them during this time.\\

		\tabto{2}
		{\bfseries Level 2}: Curse [3+X] As above, but you may extend the Hex to two characters,
		or choose 2 locations on the Missile Hit Location for one character.\\

			\tabto{3}
			{\bfseries Level 3}:  Damn [4+X] As above, but you may extend the Hex to three
			characters, or choose 3 locations on the Missile Hit Location for one character.\\

\tabto{0}
{\bfseries\large Shadow Guard [5]\\}
The shadows embrace you, protecting you from those that would do you harm. A shadowy mist
forms around your body, and attacks are significantly impeded.\\

	\tabto{1}
	{\bfseries Level 1}: You may activate this as a free action in combat. You gain 3 AV across all hit
	locations. This lasts for 3 Phases, beginning the phase you activate this ability.\\

		\tabto{2}
		{\bfseries Level 2} : As above, but you gain 6 AV.\\

\tabto{0}
{\bfseries\large Shadow Walker [15]\\}
Your mastery of the supernatural has enabled you to become one with the shadows of the Void.
This is not an experience for the faint of heart, as it brings you face to face with everything else
that lurks there.\\

			\tabto{3}
			{\bfseries Level 3}: You may activate this as an action in combat. After activating this power, when
			in Pitch Darkness, you become invisible and intangible. You cannot make any attacks or
			interactions with other characters or objects, but you cannot be hit or injured by attacks. When
			in shadow, you can may teleport instantaneously to any other shadow, within a kilometer, that
			would be large enough to contain you. This effect lasts for 1 minute (10 Rounds in Combat), or
			until you end the effect.\\

\tabto{0}
{\bfseries\large Blur [X]\\}

	\tabto{1}
	{\bfseries Level 1}: You may activate this as a free action in combat. While this is active, any ranged attack
	made against you suffers an MP penalty equal to the Void invested, up to a maximum of 5. If
	this penalty reduces the attack to 0, the attack is still made, but is automatically failed. This
	effect lasts for 3 Phases, beginning the Phase it is activated.\\

		\tabto{2}
		{\bfseries Level 2}: As above, but the maximum void investment is increased to 10.\\

\tabto{0}

\chapter{The Way of Five}\vspace*{-1.7cm}
{\bfseries\large
\noindent
Up is Down.\\
Drink the Tar.\\

\noindent
Forward is Back.\\
Drink the Tar.\\

\noindent
Love is Hate.\\
Drink the Tar.\\

\noindent
War is Peace.\\
Drink the Tar.\\

\noindent
Good is Evil.\\
Drink the Tar.\\ } % close bfseries

Babel Tar has always played an important role in Chironite society. A powerful hallucinogenic and narcotic, it
is frequently used (illegally) as a recreational drug, and less illicitly as an anaesthetic. However, a less
publicized use was discovered first by the monks of St. Bertulf. Very small quantities, taken regularly,
produce a state of impenetrable calm, objectivity, and, at times, precognition. Others had found ways to
achieve this state without use of the drug, but by “brute-forcing” the condition, enough practitioners were
assembled to make serious strides in understanding the state through sheer force of numbers.\\

The monastic order used this altered state to develop a martial arts system. Initially, the system was used as
an exercise, to maintain a religious experience despite violent and hectic surroundings, imposing a sense of
order and piety on a chaotic world. Eventually, however, the system was tested in battle and found to be
superbly effective. The monks were called upon to defend the faith against infidels and outsiders, and
eventually became a militant order.\\

Now, centuries later, The Way of Five has been adapted for modern use, and elements of the style have
found their way into military shooting stances, drills, and reloading routines--but the true art, the art of
precognition and unnaturally ordered and efficient movements, has remained the field of mystics and
warrior-monks. There are three primary groups who hold to this tradition--the Paladins, who serve the State,
the Clerics, who serve the Church, and the Pilgrims, who serve nobody. There are other practitioners and
imitators throughout the Void, but most take some inspiration from the original three.\\

{\bfseries Templars\\}
A blanket term for a related international network of Military Societies who experimented with Babel Tar as
far back as 500 years ago. The Templars are firm believers that violence is a desirable activity, a necessary
part of the human experience, and, in the case of religious orders, a sacrament to God. However, they also
believe that the only honorable form of violence is personal violence. Violence up close, where men can look
each other in the eyes as they kill. As a result, the Templars have steadfastly refused to take up modern
arms in war, and have at times violently protested the use of artillery, bombs, and other impersonal weapons
in battle. They tolerated muskets because the bayonet was still essential, but since the emergence of the
rifle, they have been a constant thorn in the side of every nation's war machine.\\
Modern Templars have agreed to cooperate with their national governments in return for rules of war being
implemented which prohibit gas attacks, firebombings and other acts of indiscriminate slaughter, but they
still aren't happy about the whole “artillery” thing, or the existence of planes and tanks. Even so, they are
excellent warriors, and many have found their ways into the military as NCOs or Officers of stormtrooper
units.\\
The Templars use melee weapons primarily—though the kind varies. The Order of the Stave from Albion
use quarterstaffs, as absurd as that sounds, while the Order of the Rose from the same country prefer
sabers. The Ximbri Brotherhood famously wield two-handed messers, and the Lucani Extraodinarii prefer to
use spears. A small group in Chiron called The House of the Trench are infamous for eschewing fancy
weapons in favor of shovels. Whatever the weapon, a Templar's abilities render him a terrible and lethal
opponent to any foe.\\

{\bfseries Pilgrims\\}
The Pilgrim Order was originally a splinter-faction of Templars who took their ideology one step further, and
rejected warfare entirely as inhuman and contrary to the values of their “perfect violence.” Pilgrims despise
the idea of the State, and reject all human authority. They seek violence on their own terms, and they are
not so concerned with the personal nature of killing as they are in the personal nature of conflict. A Pilgrim
will only engage an enemy whom he has a personal desire to fight, but he will do so with whatever means
he sees fit—a gun, a knife, a bomb, whatever. The Pilgrims are best known for fighting with a pair of pistols
in a very distinctive style that is both hypnotic and terrifying. Pilgrims are often hired as mercenaries, though
the tenants of their faith require them to be persuaded that the cause is just before they will accept payment
for intervening. It is common knowledge that Pilgrims will usually protect those in need, so many of them try
not to advertise their status too openly—those who do, however, do so in the form of dress, wearing puritan
outfits with no buttons and no adornment, and most particularly wide-brimmed buckled hats.\\

The use of the Way of Five revolves around Focus. Focus is the state of supreme calm that allows one to
make use of the superb abilities of the martial art. A character’s supply of Focus replenishes at the start of
every Round.\\

{\bfseries Way of Five at Character Creation\\}

\noindent{\centering
\begin{tabular}{lll}
Magic & User & Focus \\
Points & Level & Limit \\ \hline \\
5 & Novice & WIP \\
\\ \hline
\end{tabular}\\[\baselineskip] }
Purchasing additional Focus Limit at Character Creation costs 5 Magic Points for 4.\\
Purchasing additional Focus Limit with Arc during gameplay costs 8 Arc Points for 1.\\
Being part of a Tradition at Character Creation costs 10 Magic Points in addition to the initial 5.\\

{\bfseries\large The Five Axioms\\}
Unless specified otherwise, each of the listed effects lasts for the entire Round. A character may activate
any number of Axiom abilities in any combination, so long as they have the Focus required. Axioms must be
activated at the start of the Round before any Phases are resolved.\\
Activating both Abilities in any given Axiom also activates the [Axiom] Ability.\\
Templars and Pilgrims only need to activate the [Templar] or [Pilgrim] ability to activate the [Axiom]. If it is a
Passive ability, then they are always considered to have the Axiom activated.\\

A character's Focus is reduced by Pain.\\

{\bfseries Way the First Up is Down\\}
(The Way of Avoidance)\\
\begin{itemize}
\item Calculated Evasion [3]: You gain a +4 bonus to Defense if you are not in Cover.
\item Mobility [3]: You gain a +10 bonus to Athletics this Round.
\item {[}Pilgrim{]} [Passive]: You may move up to your MOB in meters while performing an Active Defense.
\item {[}Templar{]} [Passive]: You gain a +2 bonus to Defense when performing an Active Defense.
\item Axiom—Up is Down: If you remain stationary, you may perform an Active Defense without expending an Action, and attack normally.\\[\baselineskip]
\end{itemize}

{\bfseries Way the Second Forward is Back\\}
(The Way of Defense)\\
\begin{itemize}
\item Fight Geometry [3]: You may count Cover as one level of coverage higher than it actually
is, up to Full Cover.
\item Calculated Suppression [3]: You gain +2 Defense against any enemy against whom you
have made a Shoot attack this Round.
\item {[}Pilgrim{]} [4]: This round, every character who declares an attack targeting you in a Phase
increases your Base Defense by 1 for that Phase, against all attacks.
\item {[}Templar{]} [Passive] Precognition: So long as you have a melee weapon drawn, you gain
a +2 bonus to Defense.
\item Axiom—Forward is Back: You gain a +2 Defense in all situations.\\[\baselineskip]
\end{itemize}

{\bfseries Way the Third Love is Hate\\}
(The Way of Dueling)\\
\begin{itemize}
\item Bullseye [4]: You gain a +8 bonus to any Regular Fire attack if you only fire one shot. You may
change the Hit Location of this shot to Chest: Non-Vitals after rolling for hit location.
\item Quickdraw [4]: You gain +4 AGI for the purposes of Initiative.
\item {[}Templar{]} Carpe Diem [5]: Your gain +4 AGI for the purposes of Initiative, and you may Charge an
additional 20 meters.
\item Axiom—Love is Hate: You gain a +4 bonus to Defense this Round. For each enemy after the first
who makes an attack of any sort against you, your Defense reduces by 3 for the remainder of the
Round. This penalty can exceed 4.\\[\baselineskip]
\end{itemize}

\newpage % manual pagebreak
{\bfseries Way the Fourth War is Peace\\}
(The Way of War)\\
\begin{itemize}
\item Ride the Storm [4]: You gain +2 Defense against attacks incurred by Suppressive Fire.
\item Turkey Shoot [4]: You may count enemy Cover as being 1 stage less than it actually is
for the purposes of Shooting.
\item {[}Templar{]} [8]: If you have a melee weapon drawn, the first attack incurred against you by
Suppressive Fire each Phase automatically fails. You gain a +4 bonus to Initiative
Contests in any Melee you engage in this Round.
\item Axiom—War is Peace: You do not suffer Pain or Stun penalties this Round, except to
Focus.\\[\baselineskip]
\end{itemize}

{\bfseries Way the Fifth Good is Evil\\}
(The Way of Desperation)\\
\begin{itemize}
\item Not today! [1]: You gain +8 to Defense. Next Round, your Focus is reduced by 3.
\item No no no no NO [1]: Reduce the damage of any Explosion you are exposed to by half.
Next Round, your Focus is reduced by 3.
\item {[}Templar{]} [We are ALL going to die!] [15-X]: This ability's cost is reduced by 1 per
Wound you currently have. You may ignore all Pain and Stun penalties this Round,
except to Focus. You may charge an additional 20 meters, you gain a +4 bonus to
Defense this Round, you gain a +4 bonus to Agility this Round for the purposes of
Initiative, the first attack incurred against you by Suppressive Fire each Phase
automatically fails. You gain a +4 bonus to Initiative Contests in any Melee you engage
in this Round. At the end of each Phase, you suffer 1 Bloodloss per Wound you currently
have.
\item Axiom—Good is Evil: Any Wound of Level 5 you suffer this Round cannot be prevented
with Luck. Any Wound of Level 4 or lower that you suffer this Round is reduced to Level
0.\\[\baselineskip]
\end{itemize}

\chapter{Monsters}

Monsters fight differently from regular characters, and must be attacked differently as well.\\

To begin with, each Body Part on a Monster is treated as a single target in combat--do not
bother rolling on the Random Hit Location table against Monsters.\\

Secondly, Monsters always attack in Phase 3 of the Combat, and no other Phases.\\

Third, Monsters have a Base Defense of 0, as opposed to the Base Defense of 4 for humans.
They’re so huge that it’s almost impossible to miss them.\\

Fourth, Monsters sustain damage differently from humanoid targets. A Monster’s Toughness
depends on its size, and its Armor is determined by the toughness of its hide. However, unlike
human opponents, damage that penetrates Armor inflicts damage differently.\\

If an attack inflicts damage greater than the Monster’s Armor, take any excess damage and
compare it to the Monster’s Toughness. For each multiple of the Monster’s Toughness that the
damage reaches, it inflicts 1 level of Wound.\\

Dietrich, after carefully aiming, shoots a Star Kraken in the head with his Republican 1776, and
scores 10 successes. Since Monsters have a Base Defense of 0 and this one isn’t taking cover
(behind what!?) or moving, he hits with 10 BS. He inflicts 35 total damage, and halves the
Kraken’s Armor. The Kraken’s armor is 20. The Kraken’s Toughness is 20. So, after surpassing
the Kraken’s 10 AV, he manages to inflict 25 damage to its Toughness, inflicting a Level 1
Wound. The Star Kraken suffers a very mild injury. Dietrich immediately regrets his choice in
careers.\\

{\bfseries Monster Attacks\\}
Monster Body Parts can have Attack functions. Each Body Part can make an Attack in Phase 3,
up to the limit determined by the Monster’s Head. There are two types of Monster Attacks:\\

{\bfseries General\\}
General Monster Attacks affect an area detailed in the Monster’s description. Every character
within that area is attacked with the listed number of dice, and suffers damage if hit.\\

{\bfseries Aimed\\}
Aimed Monster Attacks target a specific object or character. These Attacks are resolved more or
less normally, with the Monster’s attack bonus being compared to the character’s Base Defense
plus other factors, just like a shooting attack.\\

{\bfseries Kraken\\}

{\bfseries Head\\}
{\bfseries (Core)\\}
Armor 20, Tou 20\\
Movement (Water Jet) 100m\\
Attacks Allowed: 4\\
Ram: Anti Ship Attack, inflicts 10 Structure Points of Damage\\
Beak Bite: Aimed Attack at +20 against Entangled Character. Kills Instantly\\

{\bfseries Tentacles (8)\\}
Armor 10, TOU 10\\
Slap: General Attack, 20m Area, +8, 20 Damage\\
Grab: Aimed Attack at +10. Entangles Character (-10 to all Actions, can be freed by killing
Tentacle)\\
Constrict: Aimed Attack at +20 vs Entangled Character, 20 Damage\\
Ship Grab: Anti Ship Attack, Tentacle is now attached to Ship, and inflicts 1 Structure Point of
Damage per Turn attached. Ship must make Engine Test at RS equal to Attached Tentacles to
break free of grab.\\

\vfill{\centering\Large\bfseries
All rights reserved (c) 2015 Opaque Industries \

}
\end{document}